### 2011-03-29 pdis/presentation1/presentation.tex
Added presentation 1 of pdis.

\documentclass{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\usetheme{Luebeck}
\useinnertheme{rounded}
\usecolortheme{whale}
\graphicspath{{pics/}}

\usepackage{tikz}
\usetikzlibrary{arrows}
\tikzstyle{block}=[draw opacity=0.7,line width=1.4cm]

\usepackage{pgfpages}
\pgfpagesuselayout{2 on 1}[a4paper,border shrink=5mm]

\setbeameroption{show notes}
\setbeameroption{notes on second screen}
\setbeamertemplate{note page}[plain]
\setbeamertemplate{footline}[frame number]
\beamertemplatenavigationsymbolsempty

\title{Novelty Detection for Visual Place Classification}
\author{André Susano Pinto\inst{1}}

\institute[FEUP] {
 \inst{1}Faculdade de Engenharia da Universidade do Porto
}

\begin{document}
\begin{frame}
 \titlepage
\end{frame}

\begin{frame}{Outline}
 \tableofcontents
\end{frame}

\section{Introduction}

\subsection{Project}
\begin{frame}{Master Thesis}

\begin{block}{}
\textbf{Title}: Novelty Detection for Visual Place Classification

\textbf{Student}: André Susano Pinto (student at FEUP currently Erasmus in KTH).
\end{block}

\begin{columns}[t]
  \column{.5\textwidth}
  \begin{block}{}
    \textbf{Supervisor}: Luis Paulo Reis at FEUP
  \end{block}
  \column{0.5\textwidth}
  \begin{block}{}
    \textbf{Supervisor}: Andrzej Pronobis at CVAP (Computer Vision and Active Perception) in KTH
  \end{block}
\end{columns}

\note{
\begin{itemize}
\item Hi my name is André and at the moment I am studying in KTH in Stockholm as an exchange student (free mover).
\item My master thesis topic will be on "Novelty detection for Visual Place Classification".
\item And it will also be performed in KTH on the Computer Vision and Active Perception lab. My supervisor there will be Andrzej Pronobis.
\item And my supervisor in FEUP will be Luis Paulo Reis.
\end{itemize}
}

\end{frame}

\subsection{Motivation}
\begin{frame}{Motivation}
\begin{itemize}
\item A.I. and robotics focus in cognitive robots able to interact with humans and man-made environments.
\item The ability to classify the environment leads to reliable navigation but also in higher-level semantic space interpretation.
\item Novelty detection allows to distinguish between what is known and what is unknown.
\end{itemize}
%-And will allow the robot to detect surprise.

\note{
\begin{itemize}
\item There has been several efforts in the area of Robotics and Artifical Inteligence in creating robots that are able to interact with humans.
\item Those robots often have to navigate on indoor rooms such as kitchens, corridors and the ability for the robot to navigate in those is crucial.
\item A correct identification of the type of room also brings easier human interaction as a human would be able to say: "Bring me a tea cup from kitchen". And the robot should be smart enough to identify the kitchen.
\item Also crucial is the ability of the robot to operate on unknown environments as often the robot cannot be trained with the environment it will be deployed to.
\item In that context understanding the properties that represent a room category and be able deal with suprise and unknown becomes crucial too. Hence the importance of be able to detect novelty.
\end{itemize}
}
\end{frame}

\section{Problem Description}

\subsection{Robot Knowledge Layers}
\begin{frame}

\begin{columns}[t]
  \begin{column}{.5\textwidth}
    \includegraphics[width=.95\textwidth]{RobotLayers}
  \end{column}
%  \cite{spatial_knowledge}

%\begin{itemize}
%\item Sensorial
%\item Place
%\item Categorical
%item Conceptual Layer
%\end{itemize}

  \begin{column}{.5\textwidth}
	\includegraphics[width=.95\textwidth]{screenshot-0003}
  \end{column}

\end{columns}
\note{
I will now describe in more detail the robot architecture in which the problem is inserted on.

\begin{itemize}
\item As the robot navigates through the environment peforming SLAM (Simulataneous Localization and Mapping) its sensor laser captures data from the environ. That data is composed of scans, imagery from cameras and depth information.
\item From that data the robot extracts given properties for its current position. We can see on the image the robot classification of the rooms as square, elongated.
\item At the same time the robot creates a graph of its positions and estimates a distribution of those properties for each node.
\item Then at an higer level the robot performs classification of each room using the extracted properties and the knowledge it has from each class. In the picture we can see the robot classification for each type of room. For example it was able to detect the corridor.
\end{itemize}
}
\end{frame}

\subsection{Property based mapping}
\begin{frame}{Property based mapping}
A property mapping allows the system to explain how rooms are categorized.

\begin{block}{Example}
\begin{itemize}
 \item A corridor has elongated shape and is usually empty.
 \item A kitchen usually contains cereal boxes.
\end{itemize}
\end{block}

\note{
\begin{itemize}
\item It has been sugested that certain groups of features appear and disappear together between diferent room classses.
\item Those can be seen as set of features generated by some property of the room.
\item For example an elongated shape room or a room containing a cereal box will often triggerthe same subset of features.
\item Performing such a mapping is important for the problem in question as it greatly increase robot-human iteraction by allowing a robot to explain the properties that lead to decisions.
\end{itemize}
}
\end{frame}

\subsection{Place Classification with Novelty detection}
\begin{frame}
\includegraphics[width=0.95\textwidth]{NoveltyDetection}

\note{
With the give description on the plataform I can now explain where classification is performed.
\begin{itemize}
\item As said before the sensor layer will extract several feature descriptors from the input data.
\item Those features are then classified or well mapped to properties.
\item Then at a later stage those properties together with the place graph is used to classify the room.
\end{itemize}

So on the diagram those 2 classification parts are represented in green.
}

\end{frame}

\section{Solution Perspectives}

\begin{frame}{Solution - Place Classification}
\begin{block}{}
Given an input determine to which classe of a set C it is believed to be generated from.
\end{block}

\begin{itemize}
\item Usually performed by multiclass classifiers. Eg.: SVM (One-Against-All, One-againts-One).
\item High dimensional data leads to poor results as it fails to define a correct boundary over data.
\item Lacks ability to detect novel cases.
\end{itemize}

% \includegraphics<1>[width=0.5\textwidth]{place_classification}
\note{
\begin{itemize}
\item The problem of classification is one in which an algorithm has to assign an input into one of given categories.
\item It is often solved using several 2-class classifiers (for example Support Vector Machines).
\item Though it is know that for data with high dimension its becomes incredibly hard to define boundaries between the diferent classes.
\item Also those methods often lack ability to understand novel cases.
\end{itemize}
}
\end{frame}

\subsection{Novelty Detection}
% Used in methods where its hard to gather positive/negative sample of a class (eg.: Breast cancer, digit recognition)
\begin{frame}{Solution - Novelty Detection}
\begin{block}{}
Given an input determine if it was generated by the same class as the one seen before.
\end{block}

\begin{itemize}
\item Also known as Single Class Classifier.
\item Used when its hard or impossible to gather representative negative samples of a class (eg.: used on breast cancer detection).
\item Some of these methods are Kernel PCA and one-class SVMs.
\end{itemize}

\note{
\begin{itemize}
\item The solution passes then by implementing novelty detection methods, also known as single class classifiers.
\item Such a system task becomes then: given an input determine if it was generated by the same class as it was trained with.
\item And during my thesis diferent methods of performing novelty detection will be implemented and analised.
\item Such as Kernel PCA and one class-SVMs.
\end{itemize}
}

\end{frame}

\section{Technical Aspects}

\subsection{Databases and Test data}
\begin{frame}{Databases and Test data}

\begin{columns}[t]
\column{.5\textwidth}

\begin{block}{COsy Location Database}
\includegraphics[width=0.95\textwidth]{cold}
\end{block}

\column{.5\textwidth}

\begin{block}{ImageCLEF}
\includegraphics[width=1\textwidth]{imageclef-logo}
\end{block}

\end{columns}

\note{
\begin{itemize}
\item For training data databases such as COSY will be used. COsy is a database that contains imagery recorded from several labs.
\item It is also planned that the developed system will be partipate in the ImageCLEF competition.
\end{itemize}
}
\end{frame}

\section{Work Plan}
\begin{frame}{Work Plan}
\begin{itemize}
\item Literature Review (3 weeks)
\item Study of the property-base semantic mapping system (2 weeks)
\item Implementing novelty detection algorithms (3 weeks)
\item Implementing global and local features (3 weeks)
\item Testing and evaluating methods (5 weeks)
\item Thesis writting (4 weeks)
\end{itemize}
\end{frame}

\section{}
\begin{frame}{Thank You!}
 \titlepage

\end{frame}

\end{document}

### 2011-03-29 pdis/presentation2/cold-database-sample.tex
Added pdis presentation 2.

\begin{figure}
\begin{center}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_PA}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_LO}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_2PO}

\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_RL}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_PA}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_CR}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA}

\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR(1)}
\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR(1)}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO(1)}
\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL(1)}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB(1)}
\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA(1)}
\end{center}
\caption{Sample data from the COLD database. Images were acquired from three diferent indoor laboratory enviroments located in diferent European cities: Saarbrücken, Freiburg and Ljubjlana.}
\end{figure}

### 2011-03-29 pdis/presentation2/presentation.tex
Added pdis presentation 2.

\documentclass[compress]{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\useinnertheme{rounded}
\usecolortheme{whale}
\useoutertheme[subsection=false]{miniframes}
\graphicspath{{figures/}}

\setbeamertemplate{footline}[frame number]
\beamertemplatenavigationsymbolsempty

\title{Novelty Detection for Visual Place Classification}
\author{André Susano Pinto\inst{1}}
\date{February 1, 2011}

\institute[FEUP] {
 \inst{1}Faculdade de Engenharia da Universidade do Porto
}

\begin{document}

\AtBeginSection[]{\begin{frame}\frametitle{Outline}\tableofcontents[currentsection]\end{frame}}

\begin{frame}
 \titlepage

    Supervisors:
    \begin{itemize}
    \item Luis Paulo Reis
    \item Andrzej Pronobis
    \end{itemize}
\end{frame}

\begin{frame}{Outline}
 \tableofcontents
\end{frame}

\section{Problem Definition and Motivation}
\subsection{Place Classification}
\begin{frame}{Place Classification}
\begin{figure}
\includegraphics[width=0.75\textwidth]{place_classification}
\end{figure}
Mapping of robot location to concepts:
\begin{itemize}
\item Improves human interaction.
\item High-level and robust localization information.
\item Knowledge about room categories and their properties.
\item Context-aware decisions\cite{dey2000towards}.
\end{itemize}
\end{frame}

\subsection{Novelty Detection}
\begin{frame}{Novelty Detection}
\begin{figure}
\includegraphics[width=0.75\textwidth]{novelty_detection}
\end{figure}

\begin{quotation}
Novelty detection is the identification of new or unknown data or signal that a machine learning system is not aware of during training\cite{markou2003novelty}.
\end{quotation}
\begin{itemize}
\item Operate in unknown environs.
\item Model of what is known.
\item Self-extend knowledge.
\end{itemize}
\end{frame}

\subsection{Goals}
\begin{frame}{Goals}

\begin{itemize}
\item Evaluation of a property-based semantic mapping system for mobile robots on a real world visual database\cite{pronobis2011exploiting}.
\vfill
\item Applying state-of-the-art novelty detection machine learning algorithms to the problem of visual place categorization.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background and State-of-Art}

\subsection{Classification}

\begin{frame}{Classification}
\framesubtitle{Probability Estimation}

\begin{columns}[t]
    \column{.5\textwidth}
    \begin{figure}
        \includegraphics[width=.9\textwidth]{classification.pdf}
    \end{figure}
    \begin{quote}This is harder than the classification problem\cite{cortes1995support}\end{quote}.
    \column{.5\textwidth}
    Several approaches\cite{bishop2006pattern}:
    \begin{itemize}
        \item Nearest Neighbour
        \item Neural Networks
        \item Gaussian Mixture Models
        \item several others and extensions\dots 
    \end{itemize}
\end{columns}
\end{frame}

\begin{frame}{Classification}
\framesubtitle{Support Vector Machines}
    \begin{columns}[t]
    \column{.5\textwidth}
    \begin{figure}
        \includegraphics[width=.75\textwidth]{Svm_max_sep_hyperplane_with_margin}
    \end{figure}
        SVM uses a max-margin hyperplane to separate two classes.
    \column{.5\textwidth}
    \begin{block}{Kernel-Trick}
        Can be extended to non-linear classification:\\
        $K(x, y) = \phi(x)\cdot\phi(y)$
    \end{block}
    \begin{block}{Multi-class}
        Can be used in multi-class scenarios:
        \begin{itemize}
            \item One-Versus-One
            \item One-Versus-All
        \end{itemize}
    \end{block}
    \end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Features}

\begin{frame}{Features}
\framesubtitle{Reasoning}
    \begin{itemize}
        \item Feature is a piece of information revealing useful information.
        \item Should be invariant and repeatable.
        \item Remove noise and useless information.
        \item Reduce the classification difficulty.
    \end{itemize}
\end{frame}

\begin{frame}{Features}
\framesubtitle{Visual Features - Local and Global}
    \begin{columns}[t]

    \column{.5\textwidth}
    \begin{figure}
        \includegraphics[width=\textwidth]{sift/sift}
        \caption{SIFT and other local features have been proven useful in object detection.}
    \end{figure}
    \column{.5\textwidth}
    \begin{figure}
    \includegraphics[width=\textwidth]{crfh_model}
    \caption{Composed Receptive Field Histograms can be used to extract appearance.}
    \end{figure}
    \end{columns}
\end{frame}

\begin{frame}{Features}
\framesubtitle{Others\dots}
    \begin{figure}
    \includegraphics[width=0.60\textwidth]{laser_and_depth_input}
    \caption{Lasers can be used to extract useful features.}
    \end{figure}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Novelty Detection}
\begin{frame}{Novelty Detection}
\framesubtitle{Probability}
    \begin{figure}
        \includegraphics[width=0.75\textwidth]{novelty.pdf}
    \end{figure}
    Novelty detection can be modeled by calculating the probability of a given input being generated and triggering on that value\cite{bishop1994novelty}.
\end{frame}

\begin{frame}{Novelty Detection}
\framesubtitle{Kernel PCA}
    \begin{columns}[t]
    \column{.5\textwidth}
        \begin{itemize}
        \item Principal Component Analysis
        \item Extended for non-linear (kernel-trick)
        \item Radial Basis Function Kernel
        \item Use reconstruction error as novelty measure
        \item Good results on several applications
        \end{itemize}
    \column{.5\textwidth}
    \begin{figure}
        \includegraphics[width=0.9\textwidth]{ringlinesquare_kpca}
    \end{figure}
        Kernel PCA used on a strange shaped class\cite{Hoffmann2007863}.
    \end{columns}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Graphical Models}

\begin{frame}{Graphical Models}
    \begin{itemize}
        \item Represent a system of random variables
        \item Allow to model conditional independence
        \item Act as a filter of probability distributions
        \item Can calculate joint probabilities (inference).
    \end{itemize}
    \begin{columns}[t]
    \column{0.3\textwidth}
        \begin{figure}
            \includegraphics[width=0.5\textwidth]{graphical-models/BayesNet.pdf}
            \caption{Bayes Networks}
        \end{figure}
    \column{0.3\textwidth}
        \begin{figure}
            \includegraphics[width=0.5\textwidth]{graphical-models/MarkovRandomField}
            \caption{Markov Random Field}
        \end{figure}
    \column{0.3\textwidth}
        \begin{figure}
            \includegraphics[width=0.60\textwidth]{graphical-models/FactorGraph.pdf}
            \caption{Factor Graph}
        \end{figure}
    \end{columns}
\end{frame}

\subsection{Related Work}
\begin{frame}{Related Work}
\begin{itemize}
    \item Object search studied for 20 years\cite{shubina2010}.
    \item Relation between object and room category\cite{galindo2005}.
    \item Bayes reasoning on top of deterministic presence of objects\cite{vasudevan2008}.
    \item Global and Local features for places indoor\cite{quattoni2009recognizing}.
\end{itemize}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Approach}
\subsection{Layers}
\begin{frame}{Layers}
    \begin{figure}
        \includegraphics[width=0.65\textwidth]{system-representation/layers.pdf}
    \end{figure}
\end{frame}

\begin{frame}{Layers}
\framesubtitle{Conceptual Graph}
    \begin{figure}
        \includegraphics[width=\textwidth]{system-representation/graph.pdf}
    \end{figure}
\end{frame}

\begin{frame}{Layers}
\framesubtitle{Novelty Detection}
    \begin{figure}
        \includegraphics[width=\textwidth]{system-representation/novelty-detection.pdf}
    \end{figure}
\end{frame}

\subsection{Databases}
\begin{frame}{Databases}
\input{cold-database-sample}
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Work Plan}
\begin{frame}{Work Plan}
\begin{figure}
\includegraphics[width=\textwidth]{gantt.pdf}
\end{figure}

%    \begin{itemize}
%        \item Literature Review (3 weeks)
%        \item Study of the property-base semantic mapping system (2 weeks)
%        \item Implementing novelty detection algorithms (6 weeks)
%        \item Testing and evaluating methods (5 weeks)
%        \item Thesis writing (4 weeks)
%    \end{itemize}
\end{frame}

\appendix

\begin{frame}[allowframebreaks]{Bibliography}
\bibliographystyle{alpha}
\bibliography{myrefs}
\end{frame}

%\begin{frame}{Bibliography}
%Please consult the technical report.
%\end{frame}

\begin{frame}
    \titlepage
    Supervisors:
    \begin{itemize}
    \item Luis Paulo Reis
    \item Andrzej Pronobis
    \end{itemize}
\end{frame}

\end{document}

### 2011-03-30 pdis/techreport/abstract.tex
Added tech-report for pdis.

\chapter*{Abstract}
This work documents a thesis proposal for \emph{Novelty Detection on Visual Place Classification}.
Motivation and description of the topic is presented showing the usefulness and need of search for solutions for the problem.

Background and State-of-Art knowledge is given as it is a requisite to understand the whole system and interactions between the approach layers.
Low-level features are briefly documented as understanding the nature of them helps to address the \emph{novelty detection} problem, which requires case-specific solutions.

The used approach for place classification is described and directions on extending it to perform \emph{novelty detection} are given.
The databases and evaluation methods are also presented as data always plays an important factor on any machine-learning problem and its performance.

At last an high-level plan for the thesis duration is given.

### 2011-03-30 pdis/techreport/approach.tex
Added tech-report for pdis.

\chapter{Approach}
\label{chap:approach}
In this chapter an overview of the approach used to address the problem of place classification in mobile robotics and concerns regarding implementing novelty detection on it are presented.

\section{Overview}
Graphical models have shown to be useful as they allow to unify the whole set of beliefs that influence a given variable, which is expected to lead to better results than exploiting a single source of information (\autoref{sec:related-work}).

Our approach is based on a \emph{chain graph} representation that combines general purpose knowledge with spatial knowledge and uses probabilistic perception to model variables.

Using a map of the environment or other knowledge (for example given by a user) the graph is built with the variables that the system wants to model (rooms, objects, sensed perceptions).
Connectivity between those variables is then created using default knowledge: such as probabilities of room connectivity or object existence on a given room category.
The low-level sensors are passed through classifiers and their results are plugged into the model.

Using that \emph{graph} the robot can then perform inferences on the variables and estimate the most likely scenario. 
It can be used for example to query about specific variables: that a given room is a kitchen.
Or used to perform high-level reasoning operations such as strategy-planning.

\section{Layers}
The system can be decomposed in four layers:
\emph{Conceptual Layer}, \emph{Categorical Layer}, \emph{Place Layer} and \emph{Sensory Layer}.
Each of them will be presented to the reader in the next sections starting from the high-level \emph{Conceptual Layer} and diving depth into the low-level ones.
\autoref{fig:system-structure} shows the 4 layers and draws lines related to their interaction. A look at the figure with a simultaneous reading of the layers is encouraged.

Special interest for novelty detection and place classification is on the \emph{Conceptual} and \emph{Categorical Layer} as they are the ones responsible for probabilistic modeling and classification.

\subsection{Conceptual Layer}
% TODO graphics:
% World <--> | wall | <--> sensors <--> Graphical model
While the robot moves through the environ it builds a graphical model (\autoref{sec:graphical-models}).
That graph models and relates the unknown variables that the robot tries to extract from reality.
The connections between those random variables can be seen as a filter of possible probability distributions for them \citep{bishop2006pattern}. It is therefore important the way the graph is connected and which variables it models since that is the world model the robot is using.

The description of the modeled variables and factors is given on the next subsections.
\subsubsection*{Modelled Variables}
\begin{description}
\item[Room Category] represents the category the room belongs to. This is the variable used as output to the the place classification.
\item[Room Shape] represents the type of room shape, such as elongated, square. The output of the laser scan is used to condition this variable.
\item[Room Appearance] represents what the room \emph{looks like}, this is associated with the global visual feature classifier.
\item[Object Presence] represents a given object. This variable is conditioned by the object detector which uses local visual features.
\end{description}

\subsubsection*{Modelled Factors}
\begin{description}
\item[Room-Room Connectivity] models the probability that two given rooms are connected. For example class rooms are very likely to be connected to a corridor.
\item[Room-Object Connectivity] models the object presence in a given room.
\item[Observed Room Shape] models the sensed room shape by the classification layer.
\item[Observed Room Appearance] models the sensed room appearance by the classification layer.
\end{description}

\subsection{Categorical Layer}
This layer is responsible from extracting properties from the sensors.
It has classifiers trained to detect several properties from the visual and laser data captured from the robot sensors.

So far the following properties are extracted:
\begin{description}
\item[Object presence] using \emph{local visual features} (\autoref{sec:local-features}).
\item[Place appearance] using \emph{global visual features} (\autoref{sec:global-features}).
\item[Room shape and size] using \emph{geometric features} from 2D laser data to classify properties as: square/elongated and room size.
\item[Landmark detection] uses 2D laser data to detect doors.
\end{description}

\subsection{Place Layer}
The place layer is responsible for the low-level movement and mapping of the environment.
The robot spreads several virtual places markers around the whole environment and maintains connectivity information between them.

By using the door landmarks detected by the categorical layer, it becomes then possible to group the place markers to create the concept of room.

Also this low-level places together with information associating the robot view point with extracted properties allows the system to perform a clever integration over the room properties, avoiding sampling bias, for example coming due to the robot spend too much time in a part of the room (\autoref{sec:accumulation}).

\subsection{Sensory Layer}
This layer captures data from the robots sensors, those include images from cameras, 2D laser scans and odometry from encoders in the movement motors.

\begin{figure}[h]
\includegraphics[width=\textwidth]{system-representation/structure}
\caption{In the figure its possible to visualize the interactions of the 4 layers that composed the proposed approach. Novelty detection will have special impact on the \emph{Conceptual} and \emph{Categorical Layers},}
\label{fig:system-structure}
\end{figure}

\section{Novelty Detection}
\label{sec:approach-novelty}
In the previous sections the used approach for the problem of visual place classification was presented.
That system has been used with success by \cite{pronobis2011exploiting} and has demonstrated to be efficient for planning and high-level reasoning under dynamic ambients.

Nonetheless the system has no knowledge about novelty and when presented with a classification task it has to classify the room into one of the room classes it knows.
This poses problems and performance decrease when the robot is deploying into new environments where the room categories or other world variables are novel to the system.

The inability of the robot to handle novel cases (a room which belongs to a category the system was not trained with) is translated at the graphical model not being prepared to handle a unknown variable value.
The detection of novel values (\autoref{sec:novelty-detection}) can be done by triggering on the unconditional probability density for the given input.
Since the presented approach uses a \emph{generative model} we expect such an approach to be feasible.

At a lower-level the system also needs to be extended to handle novelty at the classifiers used to extract properties. In that case novel inputs from the sensors translate as the inability for the classifier to operate correctly and so less confidence must be given to its results. That 
Those classifiers after training do not provide a \emph{generative models} rendering a probabilistic estimation approach infeasible.
For that case other approaches such as \gls{K-PCA} (\autoref{sec:kernel-pca}) will be used to address the problem.

### 2011-03-30 pdis/techreport/background.tex
Added tech-report for pdis.

\chapter{Background}
\label{chap:background}

In this chapter the background needed to understand the problem of visual place classification and novelty detection will be presented.
It should be enough to introduce several concepts to a reader unfamiliar with computer vision and machine learning areas.
And for readers seeking deeper detail all the concepts have been referenced to papers presenting them.

\section{Visual Place Classification}
\label{sec:place-classification}

\begin{figure}[here]
\begin{center}
\includegraphics[width=0.9\textwidth]{place_classification}
%\caption{Place classification deals with the problem of mapping a given location to a set of known concepts}
\end{center}
\end{figure}

Place Classification deals with the problem of mapping a given location to a set of known concepts.
In this thesis work its assumed the existence of labelled data (sample inputs where the correct label is known) that the algorithm can use during a training step, this is known as supervised-learning.
After the learning step the algorithm is feed with new data and is expected to guess the correct label.

It may feel natural then to pose the classification problem as a probability estimation problem of all the classes and pick the most likely. And several algorithms exists to perform that kind of analysis on data.
Though the problem of probability density estimation is harder than the problem of classification as it requires a full estimation over input space instead of a single decision between classes.
And one should not try to solve an harder problem.

So the usage of discriminative solutions, such as Support Vector Machines, is expected to lead to easier solutions.

There exists several methods that can be used for classification problem such as usage of \emph{Neural Networks}, \emph{Nearest-neighbours}, \emph{Decision Trees}, \emph{Bayesian Networks} and other statistical tools for regression and statistical analysis.
Though since only \glspl{SVM} are used on this work only a short overview over them is provided. For extended knowledge in the methods the reader is advised to look on other materials~\citep{bishop2006pattern}.

\subsection{Support Vector Machines}
\Glspl{SVM} where introduced by \cite{cortes1995support} and can be seen as discriminative linear based classifier.
They are based on a strong mathematical foundation and have powerful generalization capabilities.
In their original form \gls{SVM} separates two classes of points in an hyper-space with a \emph{maximal margin hyperplane}~(\autoref{fig:svm-sample}).

Later they were extended to deal with noisy data by using \emph{soft margins}.
And to handle non-linear spaces as seen on \autoref{sec:kernel-trick}.
Although they only allow for two class-classification several methods have been proposed to build multi-class classifiers based on binary classifiers as seen on \autoref{sec:multiclass-classifiers}.

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.5\textwidth]{Svm_max_sep_hyperplane_with_margin}
\end{center}
\caption{{SVM} separating two class of points by a \emph{maximal margin hyperplane}. The hyperplane can be described by the collection of support vectors and associated weights, marked in the image as sample points with large borders.}
\label{fig:svm-sample}
\end{figure}

They have been used in several classification and recognition problems and are in fact a standard across machine learning techniques.
Their efficiency, exact training results and generalization made their suitable for many tasks. Such as text categorization, digit-recognition, spam-classification.
They have also been extensively used in visual place classification~\citep{pronobis2005msc}.

\subsection{The Kernel Trick}
\label{sec:kernel-trick}
\gls{SVM} in their basic form are only able to handle linear spaces.
Nonetheless the classes are most of the time not linearly separable in the input space.
Although there might exists a transformation $\phi$ from the original space into a space $H$ where the input becomes linearly separable.

The Kernel trick allows to extend the \gls{SVM} definition to work on such space $H$ without ever performing an implicit transformation between spaces.
Being enough for that to have a Kernel function defining an inner-product inside such space: $K(x_i, x_j) = \phi(x_i)\cdot\phi(x_j)$.
In that sense a Kernel function can be seen as a function that calculates some similarity measure between two inputs.

Several kernel functions have been proposed being the most commonly used:

\begin{description}
\item[Polynomial Kernel] - $K(x, y) = (x \cdot y + p)^d$
\item[Radial Basis Function] - $K(x, y) = e^{-\gamma\|x - y \|^2}$
\item[Histogram intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where $\chi^2(x,y) = \sum_{i=1}^{N}\frac{(x_i-y_i)}{x_i+y_i}$ introduced by \cite{barla2003histogram} allows to compute histogram similarity.
\item[Matching Kernels] - mimic matching similarity and are used when each sample is represented as a set of features~\citep{boughorbel2005intermediate}.
\end{description}

\subsection{Extending to Multi-class}
\label{sec:multiclass-classifiers}
\Glspl{SVM} were designed as two-class classifier. And so they need to be adapted for using in multi-class problems.
The most common usage is to design multi-class classifiers recurring at the usage of multiple two-class classifiers and methods for combining them.
The most typical approaches for a multi-class classification on $c$ classes using \glspl{SVM} are:

\begin{description}
\item[One Against All] - in this method $c$ distinct classifiers are trained to distinguish any class of the remaining ones.
The output of all those classifiers (distance to the separating hyperplane) is then used to categorize the output.
The most common approach is to pick the class with the largest hyperplane distance.
Other variations exists as is the case of using the minimal distance to the average classification distance of each class~\citep{pronobis2007confidence}.

\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to distinguish between each pair of classes. The final decision is based on the output of all those classifiers being common to use a majority vote strategy.
\end{description}

\subsection{Features}\label{sec:features}
A feature is a piece of information which is expected to reveal information for solving a specific task.
In that sense features are task-dependant and they will yield different performance based on the type of task they are applied to.

A wanted property on features is its repeatability under similar conditions for the problem in hand. This is: they should be stable and invariant across unwanted types of transformations and noise.
For example a visual feature for object detection should be present even if the target object was translated, scaled, rotated, the light-conditions have changed or even if the object is partially occluded.

Extracting features with those properties allows to greatly reduce the size of input by removing unwanted noise and useless information from the captured data.
Turning the classification problem easier, more reliable and more efficient.

Often several and different types of features need to be extracted. It has been reported by \cite{pronobis2010ijrr} that using multiple features provides a great benefit in the context of place classification.
And \cite{quattoni2009recognizing} has showed that different types have different impact in indoor scene recognition based on the type of scene matching.
Namely was seen that some room-categories are more likely to be recognized by the presence of some objects and others by it generic appearance.

In the context of robotics, sensors such as cameras, laser scans are used to sense the surrounding environment, and features can be extracted from all those. Nonetheless as described in \autoref{sec:visual_motivation} the visual sensor is incredibly rich and most of our features will come from it.

Visual features can be seen as belonging to two categories: \emph{local features} and \emph{global features}.
\emph{Local features} describe fine grain properties of a part of image. For example the existence of specific corner or an edge. \Gls{SIFT} (\autoref{sec:sift}) is an example of such feature and has been proven useful for matching points between images and subsequence extension to object detection.

\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or {gist}~(\autoref{sec:gist}) try to describe the whole image. Either by statistical analysis of features over all the image or by a structured distribution of textures findable in the image.

\subsubsection{{SIFT} - Scalar Invariant Feature Transform}
\label{sec:sift}\label{sec:local-features}
The detection of interesting points has been studied for several years and is the base of several computer vision problems solution. It allows to perform point matching which can be used in several areas from image stitching, 3D reconstruction, video tracking, object detection, etc\dots

The most used method was presented by \cite{lowe1999object}. And its based on building a feature vector for each image. Each of those features is based on \emph{interesting points} detected by detecting maxima and minima of a difference of Gaussian functions applied in a scale-space.
The scale space is used to provide scale invariant detection. Gaussian functions are used as they are the only way to model a linear scale-space.
Each interesting point is then described by a contained that is rotation invariant.

By seeing an object as a set of features points, index and matching is then performed by a high-dimensional search on a database of know objects. After matching objects can be verified for geometric coherence between features.
\Gls{SVM} classifiers can also be trained to detect objects based on this type of local features by using \emph{Matching kernels} (\autoref{sec:kernel-trick}).

\begin{figure}[h]
    \includegraphics[width=\textwidth]{sift/sift.pdf}
    \caption{{SIFT} and other local features have been proven useful in object detection.}
\end{figure}

\subsubsection{Gist of a Scene}
\label{sec:gist}
\cite{oliva2006building} argue that fast scene recognition does not need to be built on top of object recognition but can be analyzed by scene-centered mechanisms.
They defend that position by pointing out behaviours on human vision:
when provided with a glance of a shot a person can identify the meaning of that given shot or "gist of a scene" without remembering specific details.

\begin{figure}[h]
\center
\includegraphics[width=0.60\textwidth]{gist.jpg}
\caption{An illustration of the gist of an image. Top row: original image I; bottom row:
noise image J for which gist(I) = gist(J). We see that the gist captures the dominant
textural features of the overall image, and their coarse spatial layout~\citep{murphy2006object}.}
\end{figure}

\subsubsection{{CRFH} - Composed Receptive Field Histograms}
\label{sec:crfh}\label{sec:global-features}
Composed Receptive Field Histograms are a multidimensional statistical representation of the occurrence of several image descriptors applied to an image.
They can be seen as an high-dimension histogram where each cell records how many pixels of the image have the cell response for the applied descriptors.
Such high-dimensional histogram is expected to be able to better global information contained in the image by capturing several properties of the image as well the relations between them.

\begin{figure}[h]
\begin{center}
\includegraphics[width=1\textwidth]{crfh_model.jpg}
\end{center}
\caption{A two dimensional histogram of the image built out from two image descriptors: $Lx$ and $Ly$. First-order Gaussian derivatives of image luminance in horizontal and vertical direction applied at a scale 4.}
\end{figure}

Several type of image descriptors can be applied.
For example Gaussian derivatives (such as $L_x$, $L_y$, $L_{xx}$, $L_{yy}$) are partial derivatives of the image luminance after applying a Gaussian filter of a given scale ($\sigma$) on the image.
Gradient magnitude descriptors (such as $|\nabla L|$) can also be used as a rotation invariant descriptor.

One of the problems of high-dimensional histograms is the memory and computationally complexity needed to handle them.
\cite{linde2004object} suggested using a sparse form to represent those by storing only the non-zero-cells in an array structure.

Multidimensional histograms have proven to be useful in the context of object recognition~\citep{schiele1996object}. And have also been previously used in the context of visual place classification~\cite{pronobis2010ijrr}.
\emph{Histogram intersection kernels} (\autoref{sec:kernel-trick}) can be used to train classifiers based on this feature.

\subsection{Other features}
Robots are often equipped with other types of sensors besides cameras and those can be used to extract information from the surrounding environment and yield useful information for the problem of place classification.

An example of such a sensor is a laser scan. Many robots are equipped with them and use it for low-level navigation.
From them its possible to extract a few set of geometric features, such as elongatedness ratios, or room size, which definitely yield important information for place classification.

Currently, with the deploy of the \gls{Kinect} device there has been an increasing usage of depth information available.
We expect that very soon usable features from that type of sensors will be developed.

\begin{figure}[h!]
\center
\includegraphics[width=0.5\textwidth]{laser_and_depth_input.pdf}
\caption{Using current laser sensors its possible to extract both 2D and 3D depth information from the environment.}
\end{figure}

\subsection{Spatio-temporal Accumulation}
\label{sec:accumulation}
In the context of robotics, data is not only available as a single-instant (eg.: single frame) but its acquired across time and space as the robot moves through the environment.
This can be used to further improve place classification by accumulating the several clues extracted by the robot for a given position and time.

By using odometry data (such as position and heading) its possible to create a view-point dependent histogram of the robot observations, which can be integrated using discriminative techniques and confidence level estimations. This approach greatly increases the reliability of classification methods as turns the classification more robust against noise or unbalanced data (such as a robot staring at a wall). \cite{pronobis2010ijrr} uses this technique.

\newpage
\section{Novelty Detection}
\label{sec:novelty-detection}
Novelty detection, also called outlier or anomaly detection, is the classification problem related with identification of new or unknown data that the system was not aware of during the training.

It has several applications such as fault detection, intrusion detection, detection of masses in mammograms, hand written digit recognition and many others~\citep{markou2003novelty}.
And has seen an increase of its application in the area of robotics and control.

It is posed as a one class classification where there is only access to the positive samples.
The objective then is to classify new inputs as being generated by the class it was trained or if they belong to some new unknown one.

Either by impossibility of generating samples of "unknown", or by extreme difficulty to acquire negative-examples, two-class classifiers cannot be applied to this problem.
Several other methods were developed to address this problem and they often can be categorized in the following approaches:
based on probability density estimation;
based on using the feature space to separate the given class from others, such as one-class \gls{SVM};
and based on reconstruction-error from a lower-dimension.

\subsection{Density estimation}
One of the simplest approach for novelty detection is pose the problem as a probability density estimation problem and use the expected probability of a given input to trigger the case as novel.

Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
In order to reduce the input size, those approaches often employ dimension-reduction techniques such as \gls{PCA}.

This approach is used in \cite{bishop1994novelty} where \emph{novelty detection} is added to a \emph{neural network} system.
On it the density estimation on the input is calculated using a \emph{Parzen window} over training-data and that value is used as a novelty measure. This way they detect novel data that the their network has not seen before and so the expected output of their system is not reliable.

\subsection{One-class {SVM}}
One-class \gls{SVM} approaches try to distinguish the class by separating the training set from all other points in the input space. They try to archive that by enclosing the training set by some structure (example: hyper-sphere)~\citep{scholkopf2000support}.
This leads to a decision boundary and follows the same principal as Vapnik solving classification with \gls{SVM}: do not solve something harder.

\subsection{Reconstruction Error}
Error reconstruction methods for novelty detection use the assumption that the class to be defined lies on a manifold embedded on the input space of higher dimensions.
By using dimensional reduction techniques, they try to defined that manifold and calculate how far a new input is from it.

\label{sec:kernel-pca}
One of the most common methods used for that is \gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick (\autoref{sec:kernel-trick}) to extend \gls{PCA} and perform a nonlinear dimensionality reduction on the input. This technique has been successfully used in novelty detection by \cite{Hoffmann2007863}.

\begin{figure}
\center
\includegraphics[width=0.50\textwidth]{ringlinesquare_kpca}
\caption{{Kernel PCA} used on a ring-square-circle distribution~\citep{Hoffmann2007863}.}
\end{figure}

\section{Graphical Models}
\label{sec:graphical-models}
Graphical models usage can be tracked back to earlies 1920 but they only become popular in mid-eighties when researchers started to use \emph{Bayesian Networks} to model expert systems~\citep{borgelt2002graphical}.

They serve as a better tool to model \emph{random variables} (nodes on the graph) and their probabilities as they model the conditional dependence between variables (edges on the graph).
Important to note that here \emph{random variables} does not denotes a truly random variable but one that is unknown by the system and is conditioned by other variables/evidences.

This type of graphs provide a generative model where the probability of any given scenario can be determined. As opposed to non-generative models where a given probability can only be calculated if certain data is given.
This means when a graphical model is learned, it can be used to generate new samples from the learned distribution.

They have been successfully used in several machine-learning task such as: information extraction, speech recognition, computer vision. They are also useful due to their ability to deal with semantic (high-level) features~\citep{boutell2006factor} and ability to represent properties of the reality they try to model.

Two main types of graphical models have been used: Bayesian Networks which model directed edges between variables and Markov Random Fields where variables are connected by a potential but no special direction is given to edges.

An important property of these graphs is the \emph{Markov-blanket} of a node.
For a given variable $a$, a \emph{Markov-blanket} is a set of variables in the surroundings of $a$ that when given the value of $a$ becomes independent of the rest of the graph~\citep{pearl1988probabilistic}.
On non-directional graphs it is directly determined by the nodes connected to $a$.
This allows the usage of graph algorithms such as \emph{min-cut} to quickly determine most likely scenarios.
In the case of directed edges a node blanket is also influenced by the direction of the edges and a more complex schemes need to be used.

As \cite{lauritzen2002chain} points this two types of models can be represented as a \emph{chain graph} where both directed and undirected edges can co-exist.
Though this generalization is hard to implement due to mis-understandings on the concepts the graph-models use.

Another useful interpretation of graphical-models are \emph{factor graphs}. Those are able to handle both \emph{Bayesian Networks} and \emph{Markov Random Fields}. Under this interpretation a graph is seen as a bipartite graph that connects variables with factors that influence them~\citep{bishop2006pattern}.
This gives a very useful framework to develop belief propagation on them by seeing a message-passing mechanism between nodes. Belief propagation is used to calculate marginal-probabilities.

\begin{figure}[ht]
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=0.8\textwidth]{graphical-models/BayesNet.pdf}
        \caption{Bayes Networks}
    \end{minipage}
    \hspace{0.5cm}
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=0.8\textwidth]{graphical-models/MarkovRandomField.pdf}
        \caption{Markov Random Field}
    \end{minipage}
    \hspace{0.5cm}
    \begin{minipage}[b]{0.3\linewidth}
        \centering
        \includegraphics[width=\textwidth]{graphical-models/FactorGraph.pdf}
        \caption{Factor Graph}
    \end{minipage}
\end{figure}

### 2011-03-30 pdis/techreport/glossary.tex
Added tech-report for pdis.

\newacronym{SVM}{SVM}{Support Vector Machine}
\newacronym{CRFH}{CRFH}{Composed Receptive Field Histogram}
\newacronym{SIFT}{SIFT}{Scalar Invariant Feature Transform}
\newacronym{PCA}{PCA}{Principal Component Analysis}
\newacronym{K-PCA}{K-PCA}{Kernel Principal Component Analysis}
\newacronym{COLD}{COLD}{COsy Location Database}
\newacronym{KTH}{KTH}{Kungliga Tekniska Högskolan}

\newglossaryentry{ImageCLEF}
{
  name={ImageCLEF},
  description={is the cross-language image retrieval track which is run as part of the Cross Language Evaluation Forum}
}

\newglossaryentry{Kinect}
{
  name={Kinect},
  description={is a motion sensor introduced by Microsoft to create a controller-free gaming and entertainment experience. Since then it has been used by several researchers and hobbyist in the area of robotics}
}

### 2011-03-30 pdis/techreport/introduction.tex
Added tech-report for pdis.

\chapter{Introduction}
\label{chap:introduction}

%\section*{}
%This chapter gives a generic overview of the problem, its motivation and goals. It also describes how the rest of the document is organized.

\section{Motivation}
There has been several efforts in the area of Artificial Intelligence and Robotics in creating robots that are able to interact with humans and their environments.
One of the existing problems is a reliable high-level localization method that can be deployed into new and unknown environments.

That task is specially difficult due to the constant change in those environments, either introduced by human interaction or by other external factors such as light-conditions.
Also the generalization requisite on such task requires highly generic and stable features to be extracted.

This thesis will focus on man-made indoor environments such as houses, offices, labs.
Where it would be desirable to map robot position to an high-level description such as kitchen, corridor, printer-area, office.
Such a classification can then be used to:
\footnote{Besides the motivation scenario, visual place classification has uses on other areas like augmented reality, content-base image retrieval and context awareness~\citep{dey2000towards}.}
\begin{itemize*}
\item Improve human interaction by mapping the robot localization to human concepts.
\item Improve robot localization methods with a high-level and robust localization information.
\item Extend knowledge about room categories and their properties.
\item Provide the robot with the ability to perform context-aware decisions.
\end{itemize*}

The robots should be able to operate in unknown environments as often they cannot be trained on the same environment they will operate on.
And under those circumstances the ability to distinguish between the known and unknown becomes a key point for reliability since it allows a robot to not trust the results it gets on new types of rooms.

It is therefore important to develop and access the quality of methods to identify novel cases.
Being the detection of novelty a key point for several tasks such as:
\begin{itemize*}
\item Operation in unknown environments.
\item Modelling what is know.
\item Ability to self-extend knowledge.
\end{itemize*}

\subsection{Visual features}
\label{sec:visual_motivation}
% Why Visual Place Classification!!
A robot often has several sensors that capture characteristics of its surrounding environment.
From those, vision is the most interesting and rich one and nowadays it is very easy to incorporate.
Making it a primary source of information for place classification.

Although its also the richness of the vision sensors that make it noisy and harder do interpret as the appearance of places varies over time due to illumination, human activity and view change.
It becomes then important to extract stable features from the visual input.
Visual features will be the main features explored during the thesis although other methods will also be used.

\section{Related Work}
\label{sec:related-work}
\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor scenes when compared to outdoor scenes results.
Since the properties that characterize rooms changes conforming its category. Namely corridors are well described by global properties and bookstores are well described by the presence of specific objects (books).
It became obvious then to use information provided by several sources. Their work uses both global and local features for scene recognition and does not address any specific information available in the context of mobile robotics.

This relation between room category and object has also been studied in the object search field.
Object search mainly focus on geometric properties but \cite{galindo2005multi} defines a bidirectional relation between object and room category, where object defines a room category and a room category provides information on where objects may be found.

Probabilistic representations are used in several localised functions in robots operating in the real-world~\citep{gross2009toomas,maierprobabilistic}. And some employ, up to some extent, a probabilistic representation across some subsystems~\citep{kraft2008exploration}.
\citet{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.

Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.

While there has been active research on visual properties and place classification, novelty detection applied to this problem has not seen much work on it~\citep{caputo2009overview}.
As \cite{markou2003novelty} reviews, novelty detection is an incredibly complex problem and requires specific techniques and methods to each problem.

\cite{bishop1994novelty} has showed that unconditional probability density can be use to provide a novelty measure. Though that probability is in most cases impractical to measure and even in those cases its necessary to find a correct threshold for \emph{novelty detection}. In higher-dimensions this method loses precision due to a spread out of the probability density function as most probability will be spread out on the tails of the function~\citep{markou2003novelty-part2}.

For that reason several other approaches have been developed for novelty detection.
One of those is the work of \cite{Hoffmann2007863} which applied non-linear statistical analysis to detect novel cases.
It has shown good results when applied to several problems such as digit recognition and cancer detection.

This type of approach although suffers from high computational and memory needs and often techniques need to be adapted to allow a online behaviour~\citep{sofman2010anytime}.

\section{Contribution/Goals}
\label{sec:goals}
During the thesis a thorough evaluation of a recently proposed property-based semantic mapping system for mobile robots~\citep{pronobis2011exploiting} on a real world visual database will be performed.

That system will be extended by usage of state-of-art novelty detection machine learning algorithms to the problem of visual place categorization.
As a final evaluation step, the developed method will be submitted to the Robot Vision Task on \gls{ImageCLEF}.

\section{Outline}
The rest of this technical report is organized as follows:

\begin{description}
\item[Chapter \ref{chap:background}] introduces the background for handling the presented problem.
It introduces the generic classification problem and techniques used to address it. Later developing on the specific visual features normally used as input for the place classification.

It also presents the novelty detection problem as well the techniques used to address it in the current context.

\item[Chapter \ref{chap:approach}] presents our approach to the problem. It introduces the platform over which the visual place classification is performed and points in the direction of extending such a platform for novelty detection.

\item[Chapter \ref{chap:testing}] introduces the testing and evaluation methodologies.
It also presents the databases used and the \gls{ImageCLEF} competition to which the developed work will be submitted.

\item[Chapter \ref{chap:workplan}] lists and elaborates on the planned tasks to be completed during the master thesis work and gives an estimated schedule for the working time. 
\end{description}

### 2011-03-30 pdis/techreport/mieic-distrib/abbrevs.tex
Added tech-report for pdis.

\chapter*{Abreviaturas e Símbolos}
\chaptermark{ABREVIATURAS E SÍMBOLOS}

\begin{flushleft}
\begin{tabular}{l p{0.8\linewidth}}
ADT      & Abstract Data Type\\
ANDF     & Architecture-Neutral Distribution Format\\
API      & Application Programming Interface\\
CAD      & Computer-Aided Design\\
CASE     & Computer-Aided Software Engineering\\
CORBA    & Common Object Request Broker Architecture\\
UNCOL    & UNiversal COmpiler-oriented Language\\
Loren    & Lorem ipsum dolor sit amet, consectetuer adipiscing
elit. Sed vehicula lorem commodo dui\\
WWW      & \emph{World Wide Web}
\end{tabular}
\end{flushleft}

### 2011-03-30 pdis/techreport/mieic-distrib/abstract.tex
Added tech-report for pdis.

\chapter*{Resumo}

O Resumo fornece ao leitor um sumário do conteúdo da dissertação.
Deverá ser breve mas conter detalhe suficiente e, uma vez que é a porta
de entrada para a dissertação, deverá dar ao leitor uma boa impressão
inicial.

Este texto inicial da dissertação é escrito no fim e resume numa
página, sem referências externas, o tema e o contexto do trabalho, a
motivação e os objectivos, as metodologias e técnicas empregues, os
principais resultados alcançados e as conclusões.

Este documento ilustra o formato a usar em dissertações na \Feup.
São dados exemplos de margens, cabeçalhos, títulos, paginação, estilos
de índices, etc. 
São ainda dados exemplos de formatação de citações, figuras e tabelas,
equações, referências cruzadas, lista de referências e índices.
%Este documento não pretende exemplificar conteúdos a usar. 
É usado texto descartável, \emph{Loren Ipsum}, para preencher a
dissertação por forma a ilustrar os formatos.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Sed vehicula
lorem commodo dui. Fusce mollis feugiat elit. Cum sociis natoque
penatibus et magnis dis parturient montes, nascetur ridiculus
mus. Donec eu quam. Aenean consectetuer odio quis nisi. Fusce molestie
metus sed neque. Praesent nulla. Donec quis urna. Pellentesque
hendrerit vulputate nunc. Donec id eros et leo ullamcorper
placerat. Curabitur aliquam tellus et diam.

Fusce sed ipsum vel velit imperdiet dictum. Sed nisi purus, dapibus
ut, iaculis ac, placerat id, purus. Integer aliquet elementum
libero. Phasellus facilisis leo eget elit. Nullam nisi magna, ornare
at, aliquet et, porta id, odio. Sed volutpat tellus consectetuer
ligula. Phasellus turpis augue, malesuada et, placerat fringilla,
ornare nec, eros. Class aptent taciti sociosqu ad litora torquent per
conubia nostra, per inceptos himenaeos. Vivamus ornare quam nec sem
mattis vulputate. Nullam porta, diam nec porta mollis, orci leo
condimentum sapien, quis venenatis mi dolor a metus. Nullam
mollis. Aenean metus massa, pellentesque sit amet, sagittis eget,
tincidunt in, arcu. Vestibulum porta laoreet tortor. Nullam mollis
elit nec justo. In nulla ligula, pellentesque sit amet, consequat sed,
faucibus id, velit. Fusce purus. Quisque sagittis urna at quam. Ut eu
lacus. Maecenas tortor nibh, ultricies nec, vestibulum varius, egestas
id, sapien.

\chapter*{Abstract}

Here goes the abstract written in English.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Sed vehicula
lorem commodo dui. Fusce mollis feugiat elit. Cum sociis natoque
penatibus et magnis dis parturient montes, nascetur ridiculus
mus. Donec eu quam. Aenean consectetuer odio quis nisi. Fusce molestie
metus sed neque. Praesent nulla. Donec quis urna. Pellentesque
hendrerit vulputate nunc. Donec id eros et leo ullamcorper
placerat. Curabitur aliquam tellus et diam.

Ut tortor. Morbi eget elit. Maecenas nec risus. Sed ultricies. Sed
scelerisque libero faucibus sem. Nullam molestie leo quis
tellus. Donec ipsum. Nulla lobortis purus pharetra turpis. Nulla
laoreet, arcu nec hendrerit vulputate, tortor elit eleifend turpis, et
aliquam leo metus in dolor. Praesent sed nulla. Mauris ac augue. Cras
ac orci. Etiam sed urna eget nulla sodales venenatis. Donec faucibus
ante eget dui. Nam magna. Suspendisse sollicitudin est et mi.

Fusce sed ipsum vel velit imperdiet dictum. Sed nisi purus, dapibus
ut, iaculis ac, placerat id, purus. Integer aliquet elementum
libero. Phasellus facilisis leo eget elit. Nullam nisi magna, ornare
at, aliquet et, porta id, odio. Sed volutpat tellus consectetuer
ligula. Phasellus turpis augue, malesuada et, placerat fringilla,
ornare nec, eros. Class aptent taciti sociosqu ad litora torquent per
conubia nostra, per inceptos himenaeos. Vivamus ornare quam nec sem
mattis vulputate. Nullam porta, diam nec porta mollis, orci leo
condimentum sapien, quis venenatis mi dolor a metus. Nullam
mollis. Aenean metus massa, pellentesque sit amet, sagittis eget,
tincidunt in, arcu. Vestibulum porta laoreet tortor. Nullam mollis
elit nec justo. In nulla ligula, pellentesque sit amet, consequat sed,
faucibus id, velit. Fusce purus. Quisque sagittis urna at quam. Ut eu
lacus. Maecenas tortor nibh, ultricies nec, vestibulum varius, egestas
id, sapien.

Phasellus ullamcorper justo id risus. Nunc in leo. Mauris auctor
lectus vitae est lacinia egestas. Nulla faucibus erat sit amet lectus
varius semper. Praesent ultrices vehicula orci. Nam at metus. Aenean
eget lorem nec purus feugiat molestie. Phasellus fringilla nulla ac
risus. Aliquam elementum aliquam velit. Aenean nunc odio, lobortis id,
dictum et, rutrum ac, ipsum. Aenean tellus magna, lacinia eget,
bibendum ut, interdum sit amet, ipsum. Class aptent taciti sociosqu ad
litora torquent per conubia nostra, per inceptos himenaeos. Mauris
felis lacus, dapibus sit amet, pretium feugiat, aliquet non,
purus.

### 2011-03-30 pdis/techreport/mieic-distrib/chapter1.tex
Added tech-report for pdis.

\chapter{Introdução} \label{chap:intro}

\section*{}

O primeiro capítulo da dissertação deve servir para apresentar o
enquadramento e a moti\-va\-ção do trabalho e para identificar e
definir os problemas que a dissertação aborda.
Deve resumir as metodologias utilizadas no trabalho e termina
apresentando um breve resumo de cada um dos capítulos
posteriores.

Este documento ilustra o formato a usar em dissertações na \Feup, não
servindo de exemplo sobre os conteúdos a usar.
São dados exemplos de margens, cabeçalhos, títulos, paginação, estilos
de índices, etc. 
São ainda dados exemplos de formatação de citações, figuras e tabelas,
equações, referências cruzadas, lista de referências e índices.

Uma recolha de normas existentes sobre este assunto pode ser
encontrada em~\citet{kn:Mat93}.

\begin{quote}
  ``Like the Abstract, the Introduction should be written to engage the
  interest of the reader. It should also give the reader an idea of
  how the dissertation is structured, and in doing so, define the
  thread of the contents.''~\citep[chap.\ Introduction]{kn:Tha01} 
\end{quote}

Neste primeiro capítulo ilustra-se a utilização de citações e de
referências biblio\-grá\-fi\-cas.
Para além de dar um exemplo de utilização de uma citação, a citação
anterior, introduz uma referência que pode ser consultada, entre
muitas outras referências bibliográficas
interessantes~\citep{kn:Tha01,kn:PP05}.

\section{Contexto/Enquadramento} \label{sec:context}

Esta secção descreve a área em que o trabalho se insere, podendo
referir um eventual projecto de que faz parte e apresentar uma breve
descrição da empresa onde o trabalho decorreu.

Lorem ipsum~\citep{kn:Lip08} dolor sit amet, consectetuer adipiscing
elit. 
Sed eget nunc. Phasellus interdum, risus viverra mollis laoreet, felis
justo iaculis ante, eget ornare purus augue non urna. Nam in magna. In a
est. Phasellus a tellus vitae enim vehicula imperdiet. Etiam sit amet
elit. In hac habitasse platea dictumst. Quisque eget turpis vel felis
elementum tempus. Curabitur sit amet tortor id libero dapibus
pretium. Integer mattis eros eu lorem. Duis erat tellus, porttitor
sed, blandit eget, fringilla et, lacus. Phasellus tristique nibh nec
orci. Mauris sed leo. Suspendisse fringilla tempor dolor. Donec sapien
enim, congue in, porta et, sollicitudin in, quam. Curabitur semper,
mauris ut vestibulum eleifend, diam ipsum tincidunt quam, et
vestibulum velit mauris ut risus.

Sed eget libero. Nulla facilisi. Proin eget tortor. Morbi
gravida. Donec arcu risus, blandit a, rutrum at, ornare ut,
nisl. Etiam consectetuer tortor eu odio. Etiam blandit molestie
ligula. Nulla facilisi. Nam a augue non justo laoreet hendrerit. Nam
aliquam, purus eu ultricies dictum, urna purus posuere neque, vel
tempus tellus enim a arcu.

\section{Projecto} \label{sec:proj}

Na continuação da secção anterior, e apenas no caso de ser um Projecto
e não uma Dissertação, esta secção apresenta resumidamente o projecto.

Nulla nec eros et pede vehicula aliquam. Aenean sodales pede vel
ante. Fusce sollicitudin sodales lacus. Maecenas justo mauris,
adipiscing vitae, ornare quis, convallis nec, eros. Etiam laoreet
venenatis ipsum. In tellus odio, eleifend ac, ultrices vel, lobortis
sed, nibh. Fusce nunc augue, dictum non, pulvinar sed, consectetuer
eu, ipsum. Vivamus nec pede. Pellentesque pulvinar fringilla dolor. In
sit amet pede. Proin orci justo, semper vel, vulputate quis, convallis
ac, nulla. Nulla at justo. Mauris feugiat dolor. Etiam posuere
fermentum eros. Morbi nisl ipsum, tempus id, ornare quis, mattis id,
dolor. Aenean molestie metus suscipit dolor. Aliquam id lectus sed
nisl lobortis rhoncus. Curabitur vitae diam sed sem aliquet
tempus. Sed scelerisque nisi nec sem.

\section{Motivação e Objectivos} \label{sec:goals}

Apresenta a motivação e enumera os objectivos do trabalho terminando
com um resumo das metodologias para a prossecução dos objectivos.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Morbi sit
amet nibh. Fusce faucibus, enim vel ultrices ornare, est mauris
ultricies velit, vitae consequat sem erat vel nunc. Nam libero eros,
mattis eget, sagittis nec, imperdiet at, sapien. Aliquam lacus. Aenean
adipiscing nibh in orci. Aliquam vestibulum, elit at fringilla
dignissim, metus diam lobortis urna, a laoreet nunc odio ac ipsum. Sed
at urna. Integer vehicula fringilla augue. Nulla lacus eros, rhoncus
sit amet, posuere ut, vehicula ac, nibh. Ut eleifend, eros eu placerat
vehicula, justo turpis blandit dolor, eu tincidunt felis risus at
ante. Aenean suscipit nisl eget eros. Ut laoreet libero eget
enim. Cras tempus pellentesque felis. Vestibulum vitae erat ac nibh
posuere eleifend.

Integer nec quam. Sed fermentum. Nunc vitae leo. Etiam sit amet
quam. Nunc vestibulum massa in mauris. Duis eget nulla. Fusce
ultricies arcu eu nibh volutpat feugiat. Maecenas urna pede, commodo
quis, porta eu, bibendum elementum, pede. Sed eros massa, molestie
eget, mattis non, rutrum ac, magna. Duis dui. Maecenas eget tortor ut
dolor semper mattis. Maecenas auctor, tellus et ultricies tempor, elit
est placerat lacus, in posuere mauris lorem et arcu.

\section{Estrutura da Dissertação} \label{sec:struct}

Para além da introdução, esta dissertação contém mais x capítulos.
No capítulo~\ref{chap:sota}, é descrito o estado da arte e são
apresentados trabalhos relacionados. 
No capítulo~\ref{chap:chap3}, ipsum dolor sit amet, consectetuer
adipiscing elit.
No capítulo~\ref{chap:chap4} praesent sit amet sem. 
No capítulo~\ref{chap:concl}  posuere, ante non tristique
consectetuer, dui elit scelerisque augue, eu vehicula nibh nisi ac
est.

### 2011-03-30 pdis/techreport/mieic-distrib/chapter2.tex
Added tech-report for pdis.

\chapter{Revisão Bibliográfica} \label{chap:sota}

\section*{}

Neste capítulo é descrito o estado da arte e são
apresentados trabalhos relacionados para mostrar o que existe no
mesmo domínio e quais os problemas em aberto.
Deve deixar claro que existe uma oportunidade de desenvolvimento que
cobre alguma falha concreta .

O capítulo deve também efectuar uma revisão tecnológica às principais
ferramentas utilizáveis no âmbito do projecto, justificando futuras
escolhas.

\section{Introdução}

Neste capítulo é ilustrada a utilização de macros \LaTeX\ para definir
entradas no índice remissivo e são feitas diversas referências
bibliográficas, usando-se texto de um artigo apresentado na Conferência 
XATA2006~\citep{kn:MVL06-xata}.

Nos últimos tempos têm surgido diversas soluções, apresentadas por
empresas do sector Automação de Sistemas para a disponibilização de
sistemas \scadadms{} na \textit{Web}.

Aliquam sollicitudin facilisis sapien. Mauris tincidunt tristique
diam. Mauris sollicitudin pede at tellus varius volutpat. Integer vel
leo. Nunc massa diam, egestas eu, venenatis at, porttitor ac,
sapien. Sed magna elit, vulputate in, lacinia sed, lobortis ac,
urna. Proin cursus massa id risus. Vestibulum libero. Curabitur
venenatis augue. Mauris eu libero eget lectus tempus tempor. In
tincidunt, justo in varius adipiscing, ipsum enim gravida massa, eget
ornare ante lacus id est. Praesent vitae est ut elit convallis
convallis. Aenean tincidunt, purus id consectetur volutpat, sem leo
pulvinar libero, nec semper sem purus ultricies nibh \cite{kn:Fra94-thesis}.

Fusce risus mi, tristique eu, consectetuer id, auctor sed, elit. Donec
laoreet. Duis consectetuer interdum libero. Etiam eu orci. In eu
arcu. Fusce luctus diam eget lectus. Duis interdum lacus sed
ligula. Proin vestibulum felis eget lacus. Vivamus vestibulum, tellus
ut congue viverra, mauris lacus tempor turpis, eu congue nisi magna at
dolor. Ut molestie vehicula libero. Praesent in neque sed risus tempus
ornare. Donec hendrerit, erat eu semper aliquam, pede nulla dapibus
risus, ut pretium orci pede et neque.
Etiam eget tortor a metus convallis viverra. Quisque eget nisi sed
orci facilisis interdum. Aliquam non felis.

\section{Secção Exemplo}\label{sec:dialecto}

\emph{Scalable Vector Graphics}\index{SVG}\index{XML!SVG} é uma
linguagem em formato XML que descreve gráficos de duas dimensões. 
Este formato padronizado pela W3C (\emph{World Wide Web Consortium})
é livre de patentes ou direitos de autor e está totalmente
documentado, à semelhança de outros W3C
standards~\citep{kn:svgdoc}.

Sendo uma linguagem XML, o \svg{} herda uma série de vantagens: a
possibilidade de transformar \svg{} usando técnicas como
XSLT\index{XML!XSLT}, de embeber \svg{} em qualquer documento
XML\index{XML} usando \textit{namespaces} ou até de  
estilizar \svg{} recorrendo a CSS\index{CSS} (\emph{Cascade Style Sheets}). 
De uma forma geral, pode dizer-se que \svg{}s interagem bem com as
actuais tecnologias ligadas ao XML e à Web, tal como referido
em~\citep{kn:svgibm,kn:svgw3c}.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Donec a
eros. Phasellus non nulla non massa venenatis convallis. In
porta. Mauris quis magna. Proin mauris eros, aliquet id, eleifend
vitae, semper quis, erat. Aliquam id lectus non odio dignissim
blandit. Vestibulum porttitor arcu ut ligula. Nunc quis
erat. Curabitur ipsum tortor, ornare vitae, dapibus pretium, hendrerit
sed, urna. Vestibulum ante ipsum primis in faucibus orci luctus et
ultrices posuere cubilia Curae; Phasellus bibendum, nulla eget varius
aliquam, tortor nulla sollicitudin quam, vel vestibulum nisl magna at
sem. Aliquam velit sapien, ultrices viverra, tempus quis, ultrices at,
dui. Aliquam sit amet justo.

Quisque tristique, metus eu iaculis
sagittis, urna leo bibendum diam, a ultricies sem diam a augue. Mauris
consectetuer, libero vel euismod tincidunt, nisi metus viverra ante,
quis pretium sapien odio nec risus. Nunc semper auctor
nulla\footnote{Exemplo de nota de rodapé.}.

\subsection{Sub-secção Exemplo} \label{batik}

Batik é um conjunto de bibliotecas baseadas em \textit{Java} que
permitem o uso de imagens \svg{} (visualização, geração ou
manipulação) em aplicações ou \textit{applets}~\citep{kn:batik}.  
O projecto Batik\index{Batik} destina-se a fornecer ao programador
alguns módulos que permitem desenvolver soluções especificas usando
\svg~\citep{kn:svgdoc}.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Nunc eu
nulla. Pellentesque vitae nibh ultrices quam iaculis
convallis. Aliquam purus eros, varius eget, volutpat sodales,
imperdiet nec, lacus. Curabitur in elit sed sem rutrum posuere. Class
aptent taciti sociosqu ad litora torquent per conubia nostra, per
inceptos himenaeos. Duis sem. Praesent ultricies odio vel
sapien. Integer faucibus malesuada libero. Cras semper, dolor id
ullamcorper varius, magna risus volutpat felis, id pellentesque nulla
ante at erat. Integer sodales.

Quisque sit amet odio. In at risus sit amet turpis interdum
posuere. Maecenas iaculis vehicula sem. Ut leo arcu, malesuada vel,
imperdiet id, dignissim a, purus. Duis eleifend, lectus non venenatis
dignissim, risus libero imperdiet mi, nec gravida massa libero sed
mauris. Nullam lobortis libero non sapien. Integer convallis iaculis
erat. Morbi dictum. Ut ultrices pellentesque velit. Cras ac
ante. Etiam in neque tincidunt lacus gravida vehicula. Proin et nisi.

Vivamus non nunc nec risus tempor varius. Quisque bibendum mi at
dolor. Aliquam consectetuer condimentum risus. Aliquam luctus pulvinar
sem. Duis aliquam, urna et vulputate tristique, dui elit aliquet nibh,
vel dignissim magna turpis id sapien. Duis commodo sem id
quam. Phasellus dolor. Class aptent taciti sociosqu ad litora torquent
per conubia nostra, per inceptos himenaeos.

\subsection{Sub-secção Exemplo}

Loren ipsum dolor sit amet, consectetuer adipiscing elit. 
Praesent sit amet sem. Maecenas eleifend facilisis leo. Vestibulum et
mi. Aliquam posuere, ante non tristique consectetuer, dui elit
scelerisque augue, eu vehicula nibh nisi ac est. Suspendisse elementum
sodales felis. Nullam laoreet fermentum urna.

Duis eget diam. In est justo, tristique in, lacinia vel, feugiat eget,
quam. Pellentesque habitant morbi tristique senectus et netus et
malesuada fames ac turpis egestas. Fusce feugiat, elit ac placerat
fermentum, augue nisl ultricies eros, id fringilla enim sapien eu
felis. Vestibulum ante ipsum primis in faucibus orci luctus et
ultrices posuere cubilia Curae; Sed dolor mi, porttitor quis,
condimentum sed, luctus in.

\section{Resumo ou Conclusões}

No final do capítulo deverá ser apresentado um resumo com as 
principais conclusões que se podem tirar.

Vivamus non nunc nec risus tempor varius. Quisque bibendum mi at
dolor. Aliquam consectetuer condimentum risus. Aliquam luctus pulvinar
sem. Duis aliquam, urna et vulputate tristique, dui elit aliquet nibh,
vel dignissim magna turpis id sapien. Duis commodo sem id
quam. Phasellus dolor. Class aptent taciti sociosqu ad litora torquent
per conubia nostra, per inceptos himenaeos.

### 2011-03-30 pdis/techreport/mieic-distrib/chapter3.tex
Added tech-report for pdis.

\chapter{Visualização de Sinópticos SVG}\label{chap:chap3}

\section*{}

Este capítulo deve começar por fazer uma apresentação detalhada do
problema a resolver\footnote{Na introdução a apresentação do
  problema foi breve.} podendo mesmo, caso se justifique,
constituir-se um capítulo com essa finalidade.

Deve depois dedicar-se à apresentação da solução sem detalhes de
implementação. 
Dependendo do trabalho, pode ser uma descrição mais teórica, mais
``arquitectural'', etc.

\section{Secção Exemplo}

Neste capítulo apresentam-se exemplos de formatação de figuras e
tabelas, equações e referências cruzadas.

Apresenta-se de seguida um exemplo de equação, completamente fora do contexto:
\begin{eqnarray}
CIF_1: \hspace*{5mm}F_0^j(a) &=& \frac{1}{2\pi \iota} \oint_{\gamma} \frac{F_0^j(z)}{z - a} dz\\
CIF_2: \hspace*{5mm}F_1^j(a) &=& \frac{1}{2\pi \iota} \oint_{\gamma} \frac{F_0^j(x)}{x - a} dx \label{eq:cif}
\end{eqnarray}

Na Equação~\ref{eq:cif} lorem ipsum dolor sit amet, consectetuer
adipiscing elit. Suspendisse tincidunt viverra elit. Donec tempus
vulputate mauris. Donec arcu. Vestibulum condimentum porta
justo. Curabitur ornare tincidunt lacus. Curabitur ac massa vel ante
tincidunt placerat. Cras vehicula semper elit. Curabitur gravida, est
a elementum suscipit, est eros ullamcorper quam, sed cursus velit
velit tempor neque. Duis tempor condimentum ante.

Phasellus imperdiet, orci vel pretium sollicitudin, magna nunc
ullamcorper augue, non venenatis dui nunc quis massa. Pellentesque
dolor elit, dapibus venenatis, viverra ultricies, accumsan cursus,
orci. Aliquam erat volutpat. Mauris ornare tristique leo. Maecenas
eros. Curabitur velit nunc, tincidunt vitae, dictum posuere, pulvinar
nec, diam. In suscipit mauris a nunc. Pellentesque gravida. Morbi quam
lacus, pretium eget, tincidunt vulputate, interdum sed,
turpis. Curabitur quis est. Sed lectus lorem, congue vel, dignissim
laoreet, blandit a, nisi. Aenean nunc ligula, tincidunt eu, hendrerit
vel, suscipit non, erat. Aliquam gravida. Integer non pede. In laoreet
augue id leo. Mauris placerat.

A arquitectura do visualizador assenta sobre os seguintes conceitos
base~\citep{kn:ZPMD97}:

\begin{itemize}
\item \textbf{Componentes} --- Suspendisse auctor mattis augue \emph{push};
\item \textbf{Praesent} --- Sit amet sem maecenas eleifend facilisis leo;
\item \textbf{Pellentesque} --- Habitant morbi tristique senectus et netus.
\end{itemize}

\subsection{Exemplo de Figura}

É apresentado na Figura~\ref{fig:arch} %da página~\pageref{fig:arch}
um exemplo de figura flutuante.

\begin{figure}[t]
  \begin{center}
    \leavevmode
    \includegraphics[width=0.86\textwidth]{puzzle}
    \caption{Arquitectura da Solução Proposta}
    \label{fig:arch}
  \end{center}
\end{figure}

Loren ipsum dolor sit amet, consectetuer adipiscing elit. 
Praesent sit amet sem. Maecenas eleifend facilisis leo. Vestibulum et
mi. Aliquam posuere, ante non tristique consectetuer, dui elit
scelerisque augue, eu vehicula nibh nisi ac est. Suspendisse elementum
sodales felis. Nullam laoreet fermentum urna.

Duis eget diam. In est justo, tristique in, lacinia vel, feugiat eget,
quam. Pellentesque habitant morbi tristique senectus et netus et
malesuada fames ac turpis egestas. Fusce feugiat, elit ac placerat
fermentum, augue nisl ultricies eros, id fringilla enim sapien eu
felis. Vestibulum ante ipsum primis in faucibus orci luctus et
ultrices posuere cubilia Curae; Sed dolor mi, porttitor quis,
condimentum sed luctus.

\subsection{Exemplo de Tabela}

É apresentado na Tabela~\ref{tab:exemplo1} um exemplo de tabela
flutuante e na Tabela~\ref{tab:exemplo2} um exemplo de tabela
flutuante, um pouco mais complicada.

\begin{table}[t]
  \centering
  \caption{Uma Tabela Simples}
\begin{tabular}{| l | p{45mm} |}
	\hline
\textbf{Acrónimo} & \textbf{Significado}\\
	\hline
	\hline
        ADT   & \emph{Abstract Data Type}\\\hline
        ANDF  & \emph{Architecture-Neutral Distribution Format}\\\hline
        API   & \emph{Application Programming Interface}\\
	\hline
\end{tabular}
  \label{tab:exemplo1}
\end{table}

Integer quis pede. Fusce nibh. Fusce nec erat vel mi condimentum
convallis. Sed at tortor non mauris pretium aliquet. In in lacus in
dolor molestie dapibus. Suspendisse potenti. Pellentesque sagittis
porta erat. Mauris sodales sapien id augue. Nam eu dolor. Donec sit
amet turpis non orci rhoncus commodo. Etiam condimentum commodo
libero.

Mauris pede. Curabitur faucibus dictum nibh. Proin tincidunt diam
vitae mauris. Sed hendrerit dolor vel ipsum. Nullam dapibus. Vivamus
tellus diam, egestas sit amet, vulputate non, vulputate id, eros. Nunc
sit amet nibh eget nibh imperdiet ornare. Cras vehicula mattis
ipsum. Sed diam arcu, semper at, gravida vitae, fermentum et,
nulla. Aenean massa orci, tristique nec, rutrum id, fringilla eget,
erat. Curabitur nulla ipsum, aliquam sed, rutrum vitae, semper quis,
ante. Fusce at nunc in dolor condimentum tempor. Duis sit amet massa.

Curabitur convallis nulla quis risus. Nulla mollis porttitor
purus. Fusce ultricies odio at ligula pellentesque suscipit. Nulla
velit libero, blandit a, aliquet quis, hendrerit id, arcu. Phasellus
porttitor porttitor purus. Suspendisse velit tortor, fringilla sit
amet, commodo a, ultrices et, mi. Donec eu metus in erat ornare
adipiscing. Praesent varius mi ac nunc. Vestibulum leo lacus,
elementum in, vestibulum sit amet, hendrerit at, justo. Sed sit amet
neque. Donec libero risus, commodo sit amet, dignissim ut, tincidunt
a, eros. Ut non lacus quis tortor mattis ullamcorper. Vivamus
consequat augue vel erat. Sed tincidunt. Sed leo eros, ornare a,
pulvinar non, mattis quis, nibh. Aliquam faucibus mi ac nisi.

Pellentesque habitant morbi tristique senectus et netus et malesuada
fames ac turpis egestas. Duis aliquet, libero sit amet ornare viverra,
augue erat interdum dolor, vitae tincidunt lorem erat a lacus. Sed
lectus nisi, auctor in, hendrerit a, molestie vel, lectus. Cum sociis
natoque penatibus et magnis dis parturient montes, nascetur ridiculus
mus. Duis lacinia tempor dui. Vivamus rhoncus, tellus a viverra
dignissim, pede dui adipiscing odio, non faucibus metus mi gravida
eros. Nullam a tellus ut velit elementum tempus. Aenean rutrum
convallis tellus. Vestibulum nulla ante, dapibus ut, lobortis ut,
varius sed, nisl. Fusce lobortis. Sed ac lorem. Nulla tincidunt nulla
eget leo. Maecenas ac lectus eu neque ultrices pharetra. Curabitur a
risus nec arcu placerat tempor. Suspendisse magna nisl, viverra a,
adipiscing eget, ornare ultricies, ligula. Maecenas eu ligula vitae
eros convallis dignissim.

\begin{table}[t]
  \centering
  \caption{Uma Tabela Mais Complicada}
\begin{tabular}{|c|r@{.}lr@{.}lr@{.}l||r|}
	\hline
\multicolumn{8}{|c|}
	{\rule[-3mm]{0mm}{8mm}Iteração $k$ de $f(x_n)$} \\
\textbf{\em k}
	& \multicolumn{2}{c}{$x_1^k$}
	& \multicolumn{2}{c}{$x_2^k$}
	& \multicolumn{2}{c||}{$x_3^k$}
	& comentários \\ \hline \hline
0   & -0&3                 & 0&6                 &  0&7   & - \\
1   &  0&47102965 & 0&04883157 & -0&53345964  & $\delta<\epsilon$ \\
2   &  0&49988691 & 0&00228830 & -0&52246185  & $\delta < \varepsilon$ \\
3   &  0&49999976 & 0&00005380 & -0&523656   &   $N$ \\
4   &  0&5                 & 0&00000307 & -0&52359743  & \\
\vdots	& \multicolumn{2}{c}{\vdots}
	& \multicolumn{2}{c}{$\ddots$}
	& \multicolumn{2}{c||}{\vdots}  & \\
7   &  0&5   & 0&0    & \textbf{-0}&\textbf{52359878}
		 & $\delta<10^{-8}$ \\ \hline
\end{tabular}
  \label{tab:exemplo2}
\end{table}

Loren ipsum dolor sit amet, consectetuer adipiscing elit. 
Praesent sit amet sem. Maecenas eleifend facilisis leo. Vestibulum et
mi. Aliquam posuere, ante non tristique consectetuer, dui elit
scelerisque augue, eu vehicula nibh nisi ac est. Suspendisse elementum
sodales felis. Nullam laoreet fermentum urna.

Duis eget diam. In est justo, tristique in, lacinia vel, feugiat eget,
quam. Pellentesque habitant morbi tristique senectus et netus et
malesuada fames ac turpis egestas. Fusce feugiat, elit ac placerat
fermentum, augue nisl ultricies eros, id fringilla enim sapien eu
felis. Vestibulum ante ipsum primis in faucibus orci luctus et
ultrices posuere cubilia Curae; Sed dolor mi, porttitor quis,
condimentum sed luctus.

\section{Secção Exemplo}

Loren ipsum dolor sit amet, consectetuer adipiscing elit. 
Praesent sit amet sem. Maecenas eleifend facilisis leo. Vestibulum et
mi. Aliquam posuere, ante non tristique consectetuer, dui elit
scelerisque augue, eu vehicula nibh nisi ac est. Suspendisse elementum
sodales felis. Nullam laoreet fermentum urna.

Duis eget diam. In est justo, tristique in, lacinia vel, feugiat eget,
quam. Pellentesque habitant morbi tristique senectus et netus et
malesuada fames ac turpis egestas. Fusce feugiat, elit ac placerat
fermentum, augue nisl ultricies eros, id fringilla enim sapien eu
felis. Vestibulum ante ipsum primis in faucibus orci luctus et
ultrices posuere cubilia Curae; Sed dolor mi, porttitor quis,
condimentum sed luctus.

\section{Resumo e Conclusões}

Resumir e apresentar as conclusões que se podem tirar no fim deste
capítulo.

### 2011-03-30 pdis/techreport/mieic-distrib/chapter4.tex
Added tech-report for pdis.

\chapter{Implementação}\label{chap:chap4}

\section*{}

Este capítulo pode ser dedicado à apresentação de detalhes de nível
mais baixo relacionados com o enquadramento e implementação das
soluções preconizadas no capítulo anterior.
Note-se no entanto que detalhes desnecessários à compreensão do
trabalho devem ser remetidos para anexos.

Dependendo do volume, a avaliação do trabalho pode ser incluída neste
capítulo ou pode constituir um capítulo separado.

\section{Secção Exemplo}

Neste capítulo mostra-se apenas o formato da dissertação.

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Integer
hendrerit commodo ante. Pellentesque nibh libero, aliquam at, faucibus
id, commodo a, velit. Duis eleifend sem eget leo. Morbi in
est. Suspendisse magna sem, varius nec, hendrerit non, tincidunt quis,
quam. Aenean congue. Vivamus vel est sit amet sem iaculis
posuere. Cras mollis, enim vel gravida aliquam, libero nunc
ullamcorper dui, ullamcorper sodales lectus nulla sed urna. Morbi
aliquet porta risus. Proin vestibulum ligula a purus. Maecenas a
nulla. Maecenas mattis est vitae neque auctor tempus. Etiam nulla dui,
mattis vitae, porttitor sed, aliquet ut, enim. Cras nisl magna,
aliquet et, laoreet at, gravida ac, neque. Sed id est. Nulla dapibus
dolor quis ipsum rhoncus cursus.

Etiam nisi est, dignissim sodales, fermentum id, pulvinar ac,
eros. Duis id orci. Nam pretium nisl ac augue. Ut adipiscing magna
eget est. Curabitur varius. Nulla facilisi. Pellentesque sit amet
neque ac dui accumsan blandit. Donec mauris felis, egestas sit amet,
convallis ac, dignissim quis, dolor. Maecenas cursus tortor vel
leo. Quisque tristique. Nunc augue odio, tincidunt in, dapibus sed,
ultricies sit amet, lorem. In hac habitasse platea dictumst. Praesent
iaculis, lacus hendrerit tempor sodales, libero tellus aliquet orci,
ut rhoncus massa lectus quis erat. Pellentesque quis dolor nec tortor
rhoncus convallis. Aliquam erat volutpat. Fusce placerat, magna eu
imperdiet lobortis, augue massa blandit turpis, a consectetuer quam
arcu sit amet risus.

\section{Mais uma Secção}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Quisque
purus sapien, interdum ut, vestibulum a, accumsan ullamcorper,
erat. Mauris a magna ut leo porta imperdiet. Donec dui odio, porta in,
pretium non, semper quis, orci. Quisque erat diam, pharetra vel,
laoreet ac, hendrerit vel, enim. Donec tristique luctus risus. Fusce
dolor est, eleifend id, elementum sit amet, varius vitae, neque. Morbi
at augue. Ut sem ligula, auctor vitae, facilisis id, pharetra non,
lectus. Nulla lacus augue, aliquam eget, sollicitudin sed, hendrerit
eu, leo. Suspendisse ac tortor. Mauris at odio. Etiam vehicula. Nam
lacinia purus at nibh. Aliquam fringilla lorem ac justo. Ut nec
enim.

Quisque ullamcorper. Aliquam vel magna. Sed pulvinar dictum
ligula. Sed ultrices dolor ut turpis. Vivamus sagittis orci malesuada
arcu venenatis auctor. Proin vehicula pharetra urna. Aliquam egestas
nunc quis nisl. Donec ullamcorper. Nulla purus. Ut suscipit lacus
vitae dui. Mauris semper. Ut eget sem. Integer orci. Nam vitae dui
eget nisi placerat convallis.

Sed id lorem. Proin gravida bibendum lacus. Sed molestie, urna quis
euismod laoreet, diam dolor dictum diam, vitae consectetuer leo ipsum
id ante. Integer eu lectus non mauris pharetra viverra. In feugiat
libero ut massa. Morbi cursus, lorem sollicitudin blandit semper,
felis magna pellentesque lacus, ut rhoncus leo neque at tellus. Sed
mattis, diam eget eleifend tincidunt, ligula eros tincidunt diam,
vitae auctor turpis est vel nunc. In eu magna. Donec dolor metus,
egestas sit amet, ultrices in, faucibus sed, lectus. Etiam est enim,
vehicula pharetra, porta non, viverra vel, nunc. Ut non sem. Etiam nec
neque.

\section{Resumo ou Conclusões}

Proin vehicula pharetra urna. Aliquam egestas
nunc quis nisl. Donec ullamcorper. Nulla purus. Ut suscipit lacus
vitae dui. Mauris semper. Ut eget sem. Integer orci. Nam vitae dui
eget nisi placerat convallis.

### 2011-03-30 pdis/techreport/mieic-distrib/chapter5.tex
Added tech-report for pdis.

\chapter{Conclusões e Trabalho Futuro} \label{chap:concl}

\section*{}

Deve ser apresentado um resumo do trabalho realizado e apreciada a
satisfação dos objectivos do trabalho, uma lista de contribuições
principais do trabalho e as direcções para trabalho futuro.

A escrita deste capítulo deve ser orientada para a total compreeensãoo
do trabalho, tendo em atenção que, depois de ler o Resumo e a
Introdução, a maioria dos leitores passará à leitura deste capítulo de
conclusões e recomendações para trabalho futuro.

\section{Satisfação dos Objectivos}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Etiam non
felis sed odio rutrum ultrices. Donec tempor dolor. Vivamus justo
neque, tempus id, ullamcorper in, pharetra non, tellus. Praesent eu
orci eu dolor congue gravida. Sed eu est. Donec pulvinar, lectus et
eleifend volutpat, diam sapien sollicitudin arcu, a sagittis libero
neque et dolor. Nam ligula. Cras tincidunt lectus quis nunc. Cras
tincidunt congue turpis. Nulla pede velit, sagittis a, faucibus vitae,
porttitor nec, ante. Nulla ut arcu. Cras eu augue at ipsum feugiat
hendrerit. Proin sed justo eu sapien eleifend elementum. Pellentesque
habitant morbi tristique senectus et netus et malesuada fames ac
turpis egestas. Vivamus quam lacus, pharetra vel, aliquam vel,
volutpat sed, nisl.

Nullam erat est, vehicula id, tempor non, scelerisque at,
tellus. Pellentesque tincidunt, ante vehicula bibendum adipiscing,
lorem augue tempor felis, in dictum massa justo sed metus. Suspendisse
placerat, mi eget molestie sodales, tortor ante interdum dui, ac
sagittis est pede et lacus. Duis sapien. Nam ornare turpis et
magna. Etiam adipiscing adipiscing ipsum. Fusce sodales nisl a
arcu. Cras massa leo, vehicula facilisis, commodo a, molestie
faucibus, metus. Suspendisse potenti. Duis sagittis. Donec porta. Sed
urna. Maecenas eros. Vivamus erat ligula, pharetra sit amet, bibendum
et, fermentum sed, dolor. Nullam eleifend condimentum nibh. Integer
leo nibh, consequat eget, mollis et, sagittis ac, felis. Duis viverra
pede in pede. Phasellus molestie placerat leo. Praesent at tellus a
augue congue molestie. Proin sed justo eu sapien eleifend
elementum. Pellentesque habitant morbi tristique senectus et netus et
malesuada fames ac turpis egestas.

\section{Trabalho Futuro}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Aliquam
tempor tristique risus. Suspendisse potenti. Fusce id eros. In eu
enim. Praesent commodo leo. Nullam augue. Pellentesque tellus. Integer
pulvinar purus a dui convallis consectetuer. In adipiscing, orci vitae
lacinia semper, sapien elit posuere sem, ac euismod ipsum elit tempus
urna. Aliquam erat volutpat. Nullam suscipit augue sed
felis. Phasellus faucibus accumsan est.

Aliquam felis justo, facilisis sit amet, bibendum ut, tempus ac,
dolor. Sed malesuada. Nunc non massa. In erat. Nulla
facilisi. Phasellus blandit, est in accumsan cursus, libero augue
elementum leo, vitae auctor mauris nisl ac tortor. Cras porttitor
ornare elit. Fusce at lorem. Sed lectus tortor, vestibulum id, varius
a, condimentum nec, lectus. Maecenas in nisi et magna pretium
aliquam. Pellentesque justo elit, feugiat nec, tincidunt a, dignissim
vel, ipsum. Sed nunc. Vestibulum ante ipsum primis in faucibus orci
luctus et ultrices posuere cubilia Curae; Aliquam tempus rhoncus
leo. Donec neque quam, cursus sit amet, ultricies varius, semper non,
pede. Donec porttitor. Sed aliquet feugiat elit.

\vspace*{12mm}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Phasellus
tellus pede, auctor ut, tincidunt a, consectetuer in, felis. Mauris
quis dolor et neque accumsan pellentesque. Donec dui magna,
scelerisque mattis, sagittis nec, porta quis, nulla. Vivamus quis
nisl. Etiam vitae nisl in diam vehicula viverra. Sed sollicitudin
scelerisque est. Nunc dapibus. Sed urna. Nulla gravida. Praesent
faucibus, risus ac lobortis dignissim, est tortor laoreet mauris,
dictum pellentesque nunc orci tincidunt tellus. Nullam pulvinar, leo
sed vestibulum euismod, ante ligula elementum pede, sit amet dapibus
lacus tortor ac nisl. Morbi libero. Integer sed dolor ac lectus
commodo iaculis. Donec ut odio.

### 2011-03-30 pdis/techreport/testing.tex
Added tech-report for pdis.

\chapter{Datasets and Evaluation}
\label{chap:testing}
This chapter describes the methods and databases used to gather data and evaluate the developed work. It also presents the \gls{ImageCLEF} competition to which the work will be submitted to.

\section{Common Sense Databases}
There are several databases that can be used to represent common sense for human reasonning.
From those databases several probabilistic relations can be derived in order to increase the quality of the model the robot uses for reasoning at high-level.
Examples of those databases are:

{OpenMind Common Sense}~\citep{singh2010open} which contains several sentence-templates filled in by humans by creating meaningful sentences (\autoref{fig:openmind-sample}). With those its possible to extract knowledge of finding certain objects in room categories.

\begin{figure}[!h]
\centering
\fbox{
\begin{minipage}{13 cm}
A kitchen floor is dirty when a kitchen floor is sticky.\\
A room where you generally find a muffin is the kitchen.\\
You generally find a blender in a kitchen.\\
An object that you might find in an office is a answering machine.\\
An object that you might find in an office is a briefcase.
\end{minipage}
}
\caption{Examples of sentences available on OpenMind Indoor Common Sense}
\label{fig:openmind-sample}
\end{figure}

\begin{figure}[!h]
\begin{center}
\includegraphics[width=0.7\textwidth]{labelme_kitchen_sample}
\end{center}
\caption{Sample annotated image from the LabelMe online database.}
\label{fig:labelme_sample}
\end{figure}

LabelMe~\citep{russell2008labelme} which is an online database for image annotation (\autoref{fig:labelme_sample}.
At the moment it counts with hundreds of thousands labelled objects.
There is both images and sequences of images that have been annotated.
The annotation does not only goes for objects but also for scene categories,
turning this database useful for extraction of data related with probabilities of finding objects in room categories. As well their appearance.

\section{Evaluation Databases}\label{sec:databases}
In order to train and evaluate the system real databases are required. In that sense we point to a previously used database: \gls{COLD}~\citep{pronobis09ijrr-cold}.

It contains image sequences captured using a regular and omni-directional cameras together with laser range scans and odometry data.
The data were recorded using three different mobile robot platforms and the same camera setup, under various weather and illumination conditions (during cloudy weather, sunny weather and at night) over several days.
In each of the three labs, the acquisition was performed in several rooms of different functionality.

\input{cold-database-sample}

Consequently, the \gls{COLD} database is an ideal testbed for assessing the robustness of localization and place recognition/categorization algorithms with respect to both categorical and dynamic changes (introduced by illumination variations and human activity).

The \gls{COLD} database is also in the process of being extend with additional data from Stockholm labs in \gls{KTH}.

\section{{ImageCLEF} - Robot Vision Task}
\subsection{History}
\gls{ImageCLEF} is the cross-language image retrieval track which is run as part of the Cross Language Evaluation Forum (CLEF). \gls{ImageCLEF} has already seen participation from both academic and commercial research groups worldwide from communities~\citep{imageclef}.
Together with each task run, datasets of training and validation data are made available providing the community with useful tools for benchmarking their methods.

It exists since 2003 and the tasks proposed by it have been evolving over-time and in 2010 it counted four main tasks: Medical Retrieval, Photo Annotation, Robot Vision and Wikipedia Retrieval.
Of which the Robot Vision is of interest to this thesis.

\subsection{Robot Vision Task in 2011}
The robot vision task~\citep{pronobis2010imageclef} which originally has been part of the \gls{ImageCLEF} since 2009 will no longer be part of it.
Nonetheless according to the organizers the competition will keep its structure and will remain active as a sole competition not part of the forum.

A transcription of the challenge description follows:

\begin{quotation}``The third edition of the challenge (2011) will focus on the problem of visual place classification, with a special focus on generalization.
Participants will be asked to classify rooms and functional areas on the basis of image sequences, captured by a stereo camera mounted on a mobile robot within an office environment.
The test sequence will be acquired within the same building but at a different floor than the training sequence.
It will contain rooms of the same categorical type ("corridor", "office", "bathroom") and it will also contain room categories not seen in the training sequence ("meeting room", "library").

The system built by participants should be able to answer the question "where are you?" when presented with a test sequence imaging a room category seen during training, and it should be able to answer "I do not know this category" when presented with a new room category.'' -- Robot Vision Task
\end{quotation}

### 2011-03-30 pdis/techreport/workplan.tex
Added tech-report for pdis.

\chapter{Work Plan}
\label{chap:workplan}
In order to accomplish the goals described in \autoref{sec:goals} several work needs to be done.
This chapter presents the several task under which the work has been split into.
Those cover the basic literature review and study, developing, testing, evaluation and writing the final report and presentation of this thesis.

\section{Tasks Description}
\subsection{Literature Review}
Although during the preparation of this report several concepts and theories were already studied, during the thesis they will need to be deeper analysed and studied.
This task is expected to be spread out during the whole thesis preparation though with a bigger impact in the beginning.

\subsection{Study Property-base semantic mapping system}
Adding to the base literature review, the novel proposed system \cite{pronobis2011exploiting} requires additional study and understanding as one of the focus of this thesis is the evaluation of that system. This will be one of the first tasks to be completed.

\subsection{Implement novelty detection algorithms}
As described in \autoref{sec:approach-novelty} novelty detection can be seen as two separate methods.
Those two methods will be implemented, starting with the identification of novel room types directly on the \emph{graphical model}.
Later the graphical model will be expanded to be able to handle novelty signals from the classifiers used to extract sensed properties.

\subsection{Testing and Evaluation of developed methods}
A step of this thesis is an evaluation of the used approach and determination of its efficiency.
The developed methods will be run through the databases described in \autoref{sec:databases}.

The results will be compared and used for tunning parameters and redirect research efforts.
For final evaluation and comparison the developed method will be submitted to robot task of \gls{ImageCLEF}.

\subsection{Thesis Writing}
As final stage all the study and results will be written in the Master thesis and its presentation will be prepared.

\section{Tasks Schedule}
Due to the nature of the tasks at hand its impossible to predict exact dates or give precise and detailed description on each of them.

We point to the fact that novelty research on itself is not an easy topic as it requires specific knowledge to each case. And also the approaches detailed here are novel on itself and its impossible to predict the problems that may appear as we move forward on the research.

Its however possible to give an high-level level overview of the tasks and their order of execution.
As presented on the graph below:
\begin{figure}[!h]
\center
\includegraphics[width=\linewidth]{gantt.pdf}
\caption{Scheme of tasks to be completed over time.}
\end{figure}

As shown the literature review, thesis writing and a small part of the system study were the main focus until the moment. Since that was required knowledge to the writing this report.

Next focus will be on implementation of novelty detection on top off the graphical-model. Test and evaluation also start earlier as scripts and performance measures need to be established in order to correctly implement the developed methods. Some writing will be involved after that in order to report results on the developed method.

After that focus will be on integration of novelty detection at the level of properties classifiers. And respective changes on the graphical-model to handle that additional information provided by the classifiers.

At last efforts will be redirected to the conclusion of the thesis writing and preparation of its presentation.

### 2011-04-01 dataset/Sampler.py
Temp commit

@@ -19,7 +19,7 @@
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
 import random
-from Utils import Histogram
+from util import Histogram
 
 class Distribution:
   def Probability(self, sample):

### 2011-04-01 dataset/__init__.py

@@ -1,0 +1,6 @@
+from Sampler import *
+
+def ExtractLabel(samples):
+  for sample in samples:
+    yield sample['label'], sample
+

### 2011-04-01 ml/__init__.py

@@ -1,0 +1,5 @@
+from Classifier import *
+
+
+def DiscreteProbabilityEstimator(samples):
+  return None

### 2011-04-01 run/seminovelty.py

@@ -1,0 +1,72 @@
+#!/usr/bin/python3
+# -*- coding: utf-8 -*-
+#
+# Novelty Detection Scripts
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+import ml
+import dataset
+
+
+Room = dataset.DefineIndependentPropertySet({
+  'Appearance': dataset.DefineProperty('appearance', ['kitchen', 'office', 'corridor']),
+  'RoomShape': dataset.DefineProperty('room_shape', ['square', 'elongated']),
+})
+
+kitchen = Room({
+             'Appearance': {'kitchen': 0.70,    'office': 0.29, 'corridor':0.01},
+             'RoomShape' : {'square':  0.60, 'elongated': 0.4},
+         })
+
+corridor = Room({
+             'Appearance':{'kitchen': 0.01,    'office': 0.01, 'corridor':0.9},
+             'RoomShape' :{'square':  0.05, 'elongated': 0.95},
+         })
+office = Room({
+             'Appearance':{'kitchen': 0.30,   'office': 0.69, 'corridor':0.01},
+             'RoomShape' :{'square':  0.6, 'elongated': 0.4},
+         })
+
+room_categories = {
+  'kitchen': kitchen,
+  'corridor': corridor,
+  'office': office
+}
+
+
+world_distrib = dataset.ClassDistribution( dataset.DiscreteDistribution(room_categories.keys()), room_categories)
+
+
+
+known_labels = {}
+
+unlabelled_data = []
+labelled_data = []
+test_data = []
+
+unconditional_prob = ml.DiscreteProbabilityEstimator(unlabelled_data)
+conditional_prob = ml.DiscreteProbabilityEstimator(
[…]

### 2011-04-01 run/test_case.py

@@ -19,9 +19,9 @@
 # along with this program; if not, write to the Free Software Foundation,
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
-from Sampler import *
-from Test import *
-import Classifier
+import dataset.Sampler as Sampler
+import ml.Test
+import ml.Classifier
 import pylab as pl
 import random
 random.seed(0)

### 2011-04-01 dataset/Sampler.py
Advacing with structure of code that allows me to create scripts to run experiments and print results.

@@ -186,6 +186,7 @@ def DefineIndependentPropertySet(properties):
 
 # TODO(andressp): create something like this in sampler module.
 def SampleN(max_samples, data):
+  data = iter(data)
   for x in range(max_samples):
     yield next(data)
 

### 2011-04-01 dataset/__init__.py

@@ -2,5 +2,5 @@ from Sampler import *
 
 def ExtractLabel(samples):
   for sample in samples:
-    yield sample['label'], sample
+    yield sample[0], sample[1]
 

### 2011-04-01 ml/__init__.py

@@ -2,4 +2,6 @@ from Classifier import *
 
 
 def DiscreteProbabilityEstimator(samples):
-  return None
+  def Density(sample):
+    return 1.0
+  return Density

### 2011-04-01 run/seminovelty.py

@@ -23,9 +23,12 @@ import ml
 import dataset
 
 
+show_data_samples = 10
+
+
 Room = dataset.DefineIndependentPropertySet({
-  'Appearance': dataset.DefineProperty('appearance', ['kitchen', 'office', 'corridor']),
-  'RoomShape': dataset.DefineProperty('room_shape', ['square', 'elongated']),
+  'Appearance': dataset.DefineProperty('place_appearance_property', ['kitchen', 'office', 'corridor']),
+  'RoomShape': dataset.DefineProperty('place_shape_property', ['square', 'elongated']),
 })
 
 kitchen = Room({
@@ -42,31 +45,67 @@ office = Room({
              'RoomShape' :{'square':  0.6, 'elongated': 0.4},
          })
 
-room_categories = {
-  'kitchen': kitchen,
-  'corridor': corridor,
-  'office': office
-}
+
+def UnlabelledData(samples = 100):
+  room_categories = {
+    'kitchen': kitchen,
+    'corridor': corridor,
+    'office': office
+  }
+  world = dataset.ClassDistribution(dataset.DiscreteDistribution(room_categories.keys()), room_categories)
+  return list(dataset.SampleN(100, world.Generator()))
+
+def LabelledData(samples = 100):
+  room_categories = {
+    'kitchen': kitchen,
+    'corridor': corridor
+  }
+  world = dataset.ClassDistribution(dataset.DiscreteDistribution(room_categories.keys()), room_categories)
+  return list(dataset.SampleN(100, world.ClassGenerator()))
+
+def TestData():
+  """This data represents the 12 places present on rocs/data/sample/ConceptualGraph/sample2 logs"""
+  data = [
+   [('roomId', 0), ('placeId', 0), ('place_appearance_property', 'office'), ('place_shape_property', 'elongated')],
+   [('roomId', 0), ('placeId', 1), ('place_appearance_property', 'corridor'), ('place_shape_property', 'square')],
+   [('roomId', 0), ('placeId', 2), ('place_appearance_property', 'office'), ('place_shape_property', 'elongated')],
+   [('roomId', 1), ('placeId', 4), ('place_appearance_property', 'corridor'), ('place_shape_property', 'elongated')],
+   [('roomId', 2), ('placeId', 6), ('place_appearance_property', 'kitchen'), ('place_shape_property', 'square')],
[…]

### 2011-04-02 dataset/__init__.py
Added simpler distribution definition and sampling methods. Only discrete distributions supported.
Others can be implemented by creating objects with a method GenerateSample(self).

@@ -1,4 +1,25 @@
-from Sampler import *
+# -*- coding: utf-8 -*-
+#
+# Novelty Detection Scripts
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+from Sampler import SampleN
+from distribution import DiscreteDistribution, GenerateSample, Generator
 
 def ExtractLabel(samples):
   for sample in samples:

### 2011-04-02 run/seminovelty.py

@@ -21,14 +21,14 @@
 #
 import ml
 import dataset
-
+import dataset.Sampler as sampler
 
 show_data_samples = 10
 
 
-Room = dataset.DefineIndependentPropertySet({
-  'Appearance': dataset.DefineProperty('place_appearance_property', ['kitchen', 'office', 'corridor']),
-  'RoomShape': dataset.DefineProperty('place_shape_property', ['square', 'elongated']),
+Room = sampler.DefineIndependentPropertySet({
+  'Appearance': sampler.DefineProperty('place_appearance_property', ['kitchen', 'office', 'corridor']),
+  'RoomShape': sampler.DefineProperty('place_shape_property', ['square', 'elongated']),
 })
 
 kitchen = Room({
@@ -52,7 +52,7 @@ def UnlabelledData(samples = 1000):
     'corridor': corridor,
     'office': office
   }
-  world = dataset.ClassDistribution(dataset.DiscreteDistribution(room_categories.keys()), room_categories)
+  world = sampler.ClassDistribution(sampler.DiscreteDistribution(room_categories.keys()), room_categories)
   return list(dataset.SampleN(100, world.Generator()))
 
 def LabelledData(samples = 1000):
@@ -61,7 +61,7 @@ def LabelledData(samples = 1000):
 #    'kitchen': kitchen,
 #    'corridor': corridor
   }
-  world = dataset.ClassDistribution(dataset.DiscreteDistribution(room_categories.keys()), room_categories)
+  world = sampler.ClassDistribution(sampler.DiscreteDistribution(room_categories.keys()), room_categories)
   return list(dataset.SampleN(100, world.ClassGenerator()))
 
 def TestData():

### 2011-04-02 ml/__init__.py
Added missing license preamble

@@ -1,3 +1,23 @@
+# -*- coding: utf-8 -*-
+#
+# Novelty Detection Scripts
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
 from Classifier import *
 from util import Histogram
 

### 2011-04-02 dataset/__init__.py
Converted run/seminovelty to the new distribution classes.

@@ -21,7 +21,10 @@
 from Sampler import SampleN
 from distribution import DiscreteDistribution, GenerateSample, Generator
 
-def ExtractLabel(samples):
-  for sample in samples:
-    yield sample[0], sample[1]
+def ExtractLabel(sample):
+  assert sample[0][0] == 'label'
+  return sample[0][1], tuple(list(sample)[1:])
 
+def FilterLabel(sample):
+  label, sample = ExtractLabel(sample)
+  return sample

### 2011-04-02 ml/__init__.py

@@ -21,9 +21,9 @@
 from Classifier import *
 from util import Histogram
 
-def DiscreteProbabilityEstimator(samples, input_state_size = 0):
+def DiscreteProbabilityEstimator(samples, state_size = 0):
   counter = Histogram(samples)
-  total = input_state_size + sum(map(lambda x: x[1], counter.items()), 0)
+  total = state_size + sum(map(lambda x: x[1], counter.items()), 0)
   def Density(sample):
     if sample in counter:
       return (counter[sample]+1.0) / total 

### 2011-04-02 run/seminovelty.py

@@ -21,48 +21,54 @@
 #
 import ml
 import dataset
-import dataset.Sampler as sampler
 
 show_data_samples = 10
 
-
-Room = sampler.DefineIndependentPropertySet({
-  'Appearance': sampler.DefineProperty('place_appearance_property', ['kitchen', 'office', 'corridor']),
-  'RoomShape': sampler.DefineProperty('place_shape_property', ['square', 'elongated']),
-})
-
-kitchen = Room({
-             'Appearance': {'kitchen': 0.70,    'office': 0.29, 'corridor':0.01},
-             'RoomShape' : {'square':  0.60, 'elongated': 0.4},
-         })
-
-corridor = Room({
-             'Appearance':{'kitchen': 0.01,    'office': 0.01, 'corridor':0.9},
-             'RoomShape' :{'square':  0.05, 'elongated': 0.95},
-         })
-office = Room({
-             'Appearance':{'kitchen': 0.30,   'office': 0.69, 'corridor':0.01},
-             'RoomShape' :{'square':  0.6, 'elongated': 0.4},
-         })
+def Label(label):  return ('label', label)
+def Appearance(d): return ('place_appearance_property', dataset.DiscreteDistribution(d))
+def Shape(d):      return ('place_shape_property',      dataset.DiscreteDistribution(d))
 
 
-def UnlabelledData(samples = 1000):
-  room_categories = {
-    'kitchen': kitchen,
-    'corridor': corridor,
-    'office': office
-  }
-  world = sampler.ClassDistribution(sampler.DiscreteDistribution(room_categories.keys()), room_categories)
-  return list(dataset.SampleN(100, world.Generator()))
+kitchen = ( Label('kitchen'),
+            Appearance([(0.70, 'kitchen'),
+                        (0.29, 'office' ),
+                        (0.01, 'corridor')]),
+            Shape([(0.60, 'square'),
+                   (0.40, 'elongated')])
+          )
 
-def LabelledData(samples = 1000):
-  room_categories = {
-    'office': office,
-#    'kitchen': kitchen,
-#    'corridor': corridor
-  }
-  world = sampler.ClassDistribution(sampler.DiscreteDistribution(room_categories.keys()), room_categories)
-  return list(dataset.SampleN(100, world.ClassGenerator()))
+corridor = ( Label('corridor'),
+             Appearance([(0.05, 'kitchen'),
+                         (0.05, 'office'),
[…]

### 2011-04-04 run/seminovelty.py
Added state_size as defined variable (no hardcoded values)

@@ -27,6 +27,7 @@ show_data_samples = 10
 def Label(label):  return ('label', label)
 def Appearance(d): return ('place_appearance_property', dataset.DiscreteDistribution(d))
 def Shape(d):      return ('place_shape_property',      dataset.DiscreteDistribution(d))
+state_size = 6
 
 
 kitchen = ( Label('kitchen'),
@@ -108,8 +109,8 @@ if show_data_samples:
 
 
 
-unconditional_prob = ml.DiscreteProbabilityEstimator(unlabelled_data, state_size = 6)
-conditional_prob   = ml.DiscreteProbabilityEstimator(map(dataset.FilterLabel, labelled_data), state_size = 6)
+unconditional_prob = ml.DiscreteProbabilityEstimator(unlabelled_data, state_size = state_size)
+conditional_prob   = ml.DiscreteProbabilityEstimator(map(dataset.FilterLabel, labelled_data), state_size = state_size)
 
 def threshold(sample):
   return conditional_prob(sample) / unconditional_prob(sample)

### 2011-04-04 dataset/__init__.py
Several changes on functions

@@ -28,3 +28,11 @@ def ExtractLabel(sample):
 def FilterLabel(sample):
   label, sample = ExtractLabel(sample)
   return sample
+
+def UnlabelledSample(distribution, samples = 1000):
+  gen = Generator(distribution)
+  return list(map(FilterLabel, SampleN(samples, gen)))
+
+def LabelledSample(distribution, samples = 1000):
+  gen = Generator(distribution)
+  return list(SampleN(samples, gen))

### 2011-04-04 graph/ROC.py

@@ -5,7 +5,7 @@
 import numpy as N
 import pylab as P
 
-def ROC(data):
+def ROC(data, label = None):
   """Plots a ROC curve."""
   
   def acc(data, value):
@@ -20,11 +20,11 @@ def ROC(data):
   if fpr[-1] != 0: fpr /= fpr[-1]
   if tpr[-1] != 0: tpr /= tpr[-1]
   
-  P.figure()
-  P.plot(fpr, tpr)
+  P.plot(fpr, tpr, label = label)
   P.xlim([-0.05,1.05])
   P.ylim([-0.05,1.05])
   P.xlabel('False Positive Rate')
   P.ylabel('True Positive Rate')
   P.title('Receiver operating characteristic')
+  P.legend(loc = 'best')
 

### 2011-04-04 graph/__init__.py

@@ -1,2 +1,7 @@
+import pylab as pl
 from .HintonDiagrams import hinton
 from .ROC import ROC as roc
+
+
+def show():
+  pl.show()

### 2011-04-04 ml/__init__.py

@@ -23,6 +23,7 @@ from util import Histogram
 
 def DiscreteProbabilityEstimator(samples, state_size = 0):
   counter = Histogram(samples)
+  print counter
   total = state_size + sum(map(lambda x: x[1], counter.items()), 0)
   def Density(sample):
     if sample in counter:

### 2011-04-04 run/seminovelty.py
Semi-novelty sample code working.
Results are very strange since there is only 6 different possible outputs.

@@ -20,9 +20,10 @@
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
 import ml
+import graph
 import dataset
 
-show_data_samples = 10
+show_data_samples = 0
 
 def Label(label):  return ('label', label)
 def Appearance(d): return ('place_appearance_property', dataset.DiscreteDistribution(d))
@@ -54,22 +55,13 @@ office   = ( Label('office'),
                     (0.4, 'elongated')])
            )
 
-print list(dataset.SampleN(2, dataset.Generator(kitchen)))
+all_classes = [(0.2, corridor), (0.4, kitchen), (0.4, office)]
+known_labels = ['kitchen', 'corridor']
+known_classes = [all_classes[1], all_classes[0]]
 
-all_classes = [corridor, kitchen, office]
-known_classes = [ corridor, kitchen ]
+all_classes   = dataset.DiscreteDistribution(all_classes)
+known_classes = dataset.DiscreteDistribution(known_classes)
 
-
-
-def UnlabelledData(samples = 10000, classes = all_classes):
-  distribution = dataset.DiscreteDistribution([(1, c) for c in classes])
-  gen = dataset.Generator(distribution)
-  return list(map(dataset.FilterLabel, dataset.SampleN(100, gen)))
-
-def LabelledData(samples = 10000, classes = known_classes):
-  distribution = dataset.DiscreteDistribution([(1, c) for c in classes])
-  gen = dataset.Generator(distribution)
-  return list(dataset.SampleN(100, gen))
 
 def TestData():
   """This data represents the 12 places present on rocs/data/sample/ConceptualGraph/sample2 logs"""
@@ -95,8 +87,8 @@ def TestData():
 
 
 
-unlabelled_data = UnlabelledData()
-labelled_data   = LabelledData()
+unlabelled_data = dataset.UnlabelledSample(all_classes, samples = 10000)
+labelled_data   = dataset.LabelledSample(known_classes, samples = 1000)
 test_data       = TestData()
 
 if show_data_samples:
@@ -112,22 +104,32 @@ if show_data_samples:
 unconditional_prob = ml.DiscreteProbabilityEstimator(unlabelled_data, state_size = state_size)
 conditional_prob   = ml.DiscreteProbabilityEstimator(map(dataset.FilterLabel, labelled_data), state_size = state_size)
 
-def threshold(sample):
[…]

### 2011-04-04 semisupervised/notes.tex
Added notes on novelty detection when there is access to unlabelled data.
This might have a better name, maybe concept detection.
This is mainly brainstorming and random text at the moment.

\documentclass[a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[round]{natbib}
\include{macros}

\title{Notes on Semi-Supervised `Novelty Detection'}
\author{André Susano Pinto}

\begin{document}
\maketitle

\section{Introduction}
The case presented here assumes that a machine learning system besides having access to samples of the known class also has access to unlabelled data representative of the whole world.
This may happen in some machine learning problems where a set of samples from a given class has been grouped and identified and although its possible to gather new samples its infeasible to classify them (eg.: the human classifier got bored)\footnote{Other examples of situations are given on \autoref{sec:detection}.}.

Under that assumption its possible to model the functions $P(x|know)$ and $P(x)$ representing both a conditional and an unconditional density function for a given input.
Using those its possible to define a threshold function for novelty detection as seen on \autoref{sec:threshold}.

In \autoref{sec:threshold-analysis} the described threshold is analyzed in more detail, and in \autoref{sec:experiment} experiments to test the described methods are presented.

\subsection{(Novelty?) Detection}
\note{I believe calling \emph{novelty detection} to the threshold described here is incorrect.
I am not sure which would be the correct term.}

This new assumption of access to unlabelled data makes it dubious whether it can still be classified as a \emph{novelty detection} problem:
\begin{quotation}
`Novelty detection is the identification of new or unknown data or signal that a machine learning system is not aware of during training.' - \cite{markou2003novelty}.
\end{quotation}

But it is also different from a binary classification problem since instead of negative samples there is access to unlabelled data\footnote{Though several detection systems use binary classification techniques where negative samples are obtained by randomly sampling.}.

\section{Concept Detection}
\label{sec:detection}

\note{
This lacks literature background, and writing it on an airplane does not helps.
Gotta find some papers on the subject
}

In several machine learning problems the target is to detect whether a concept is present.
(eg.: face detection, intrusion detection).

In cases where negative samples are \emph{abundant} (face detection, music/speech detection), systems often use a random sample of input space to represent negative samples and then perform detection using a normal 2-class classification.

On systems where negative samples are very rare (intrusion-detection, cancer-detection), system need to use other approaches based on novelty detection\dots

Some cases lie somewhere in between: negative samples are easy to obtain, but they are not that common that we can assume a random sample does not contains positive samples in it. Hence we cannot treat a random sample as a negative.

In those cases we point to the usage of the ratio between conditional and unconditional probabilities as a detection function.
Examples of cases where this type of detection is may or may not appear are given below:

\subsubsection*{Unconstrainted labelled data}
On a system where a user is asked to tag samples with words representing concepts it is fairly easy to filter positive samples for a given concept.
Although getting negative samples is an impossible task, as the non-presence of a given tag on a sample does not necessarily makes that a negative sample for the concept.

Eg.: associating concepts with images, tagging objects on images, ...

\subsubsection*{Tag-prediction systems}
Music, image, text, news, auto-tagging\dots

\subsubsection*{Good approximation of density functions}
A good approximation of probability functions for a given input is necessary for the performance of the presented method.
So its expected to have uses where the input space is small or it can be modelled with some precision (eg.: independent features).

\section{Threshold}
\label{sec:threshold}

The threshold is based on the probability that a given concept is present on a sample
knowing both that the concept samples and the unlabelled samples come from the same distribution $P$.

In that case our optimal threshold would be:
$P(concept | x) > 1/2$, which means we assume the concept is present if we have more than chance probability of getting it right.

Using bayes rules we obtain:

\begin{equation}\frac{P(x|concept) P(concept)}{P(x)} > \frac{1}{2}\end{equation}
\begin{equation}\frac{P(x|concept)}{P(x)} > \frac{1}{2 P(concept)}\end{equation}
\begin{equation}\frac{P(x|concept)}{P(x)} > K_{threshold}\end{equation}

\note{Novelty detection techniques often assume a constant $P(x)$ and then threshold is performed only on the conditional probability density function.}

\section{Threshold Analysis}
\label{sec:threshold-analysis}

\begin{equation}P(x) = \sum_{c}^{} P(x|c) P(c)\end{equation}

\begin{equation}0 \le P(x|c) P(c) \le P(x)\end{equation}
\begin{equation}0 \le P(c) \le \frac{P(x)}{P(x|c)}\end{equation}

\section{Experiment}
\label{sec:experiment}

\begin{itemize}
\item Define a set of features (all extractable from visual or user-tags).
\item Define a dataset of inputs (eg.: indoor environs at LabelMe)
\item Filter those for \emph{kitchen}.
\item Use ML methods to model $P(features | kitchen)$ and $P(features)$.
\item Test results.
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{refs}

\end{document}

### 2011-04-04 graph/__init__.py
Added made pylab savefig available through graph module.

@@ -5,3 +5,6 @@ from .ROC import ROC as roc
 
 def show():
   pl.show()
+
+def savefig(path):
+  pl.savefig(path)

### 2011-04-04 run/seminovelty.py
Added output of input space sorted by threshold value.

@@ -89,7 +89,8 @@ def TestData():
 
 unlabelled_data = dataset.UnlabelledSample(all_classes, samples = 10000)
 labelled_data   = dataset.LabelledSample(known_classes, samples = 1000)
-test_data       = TestData()
+test_data       = dataset.LabelledSample(all_classes, samples = 1000)
+#test_data       = TestData()
 
 if show_data_samples:
   print('# Unlabelled Data is like:')
@@ -123,7 +124,6 @@ def plot_roc(threshold, samples, known_labels = known_labels, title=''):
   graph.roc(roc_curve, label = title)
   pass
 
-test_data = dataset.LabelledSample(all_classes, samples = 1000)
 plot_roc(threshold = density_threshold,
          samples   = test_data,
          title= 'ROC for P(x|c) threshold')
@@ -132,4 +132,18 @@ plot_roc(threshold = semi_threshold,
          samples   = test_data,
          title= 'ROC for P(x|c)/P(x) threshold')
 
+graph.savefig('example1.pdf')
 graph.show()
+
+"""Input space is small, so we can analyze it."""
+def analyse(threshold, samples):
+  for a in sorted(zip(map(threshold, samples), samples), reverse=True):
+    print(a)
+
+inputs = list(set(unlabelled_data))
+print('P(x|c) Threshold')
+analyse(density_threshold, inputs)
+
+print('P(x|c)/P(x) Threshold')
+analyse(semi_threshold, inputs)
+

### 2011-04-04 dataset/__init__.py
Added FeatureFilter function to help clean test inputs and remove data.

@@ -36,3 +36,19 @@ def UnlabelledSample(distribution, samples = 1000):
 def LabelledSample(distribution, samples = 1000):
   gen = Generator(distribution)
   return list(SampleN(samples, gen))
+
+def FeatureFilter(accept=set(), block=set()):
+  def filter(sample):
+    out = []
+    print sample
+    for feature in sample:
+      fname, fvalue = feature
+      if accept and fname in accept:
+        out.append(feature)
+        continue
+      elif block and fname in block:
+        continue
+      else:
+        out.append(feature)
+    return tuple(out)
+  return filter

### 2011-04-04 semisupervised/dora-data/dora-data.py
tmp commit: start definition of dora ontology for spatial understanding.

@@ -1,0 +1,50 @@
+
+
+# This file exports the ontology defined on:
+# http://teambox.pronobis.pro/projects/iros-paper/pages/data
+#
+import dataset
+
+labels = set('''
+             anteroom bathroom computerlab conferencehall doubleoffice
+             hallway kitchen meetingroom professorsoffice robotlab office
+             singleoffice
+             '''.split())
+
+objects = set('''
+              book cerealbox computer robot stapler toiletpaper
+              '''.split())
+
+shapes = set('''
+             elongated rectangular square
+             '''.split())
+
+appearances = set('''
+                  bathroom-like hallway-like office-like kitchen-like
+                  lab-like meetingroom-like anteroom-like
+                  '''.split())
+
+sizes = set('''
+            small medium large
+            '''.split())
+
+def Label(label):
+  assert label in labels
+  return ('label', label)
+
+def Object(obj, gamma):
+  assert 
+  return ('object_%s'%obj, dataset.DiscreteDistribution(d))
+
+def Shape(d):
+  assert
+  return ('room_size', dataset.DiscreteDistribution(d))
+
+def Appearance(d):
+  assert
+  return ('room_size', dataset.DiscreteDistribution(d))
+
+def Size(d):
+  assert Valid(d, size)
+  return ('room_size', dataset.DiscreteDistribution(d))
+

### 2011-04-05 semisupervised/dora-data/ontology.py
Added module to load ontologies from dora semantic maps config files, and create distributions for the defined room types.

@@ -1,0 +1,109 @@
+#!/usr/bin/python3
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+# This module loads the ontology and creates distributions defined by:
+# default_knowledge-semmap.xml and defaultprobs-semmap.txt
+#
+import re
+import math
+import dataset
+from collections import defaultdict
+
+potential = dict()
+roomcats  = set()
+features  = set()
+feature_space = defaultdict(set)
+
+def Load():
+  print("Loading default knowledge...")
+  for line in open("default_knowledge-semmap.xml"):
+    match = re.match('^\s+<item\s+room_category1="(?P<roomcat>\w+)"\s+' + 
+                     '(?P<proptype>\w+)="(?P<propval>\w+)"\s+' + 
+                     '(?:potential|probability)="(?P<potential>\S+)"\s*\/>\s*$', line)
+    if match:
+      roomcat, feature, descriptor, prob = match.groups()
+      potential[roomcat, feature, descriptor] = float(prob)
+      roomcats.add(roomcat)
+      features.add(feature)
+      feature_space[feature].add(descriptor)
+    elif line.strip() == '':
+      pass
+    else:
+      print "Ignored line: ", line,
+  
+  
+  print('Loading object probabilities...')
+  for line in open("defaultprobs-semmap.txt"):
+    match = re.match('^INROOM (?P<object>\w+) (?P<roomcat>\w+) (?P<gamma>\S+)\n$', line)
+    if match:
+      obj, roomcat, gamma = match.groups()
+      feature = 'object_%s' % obj
+      gamma = float(gamma)
+      """Unzip the poisson distribution for up to 5+ samples"""
+      features.add(feature)
[…]

### 2011-04-05 semisupervised/dora-data/ontology.py
small changes

@@ -58,14 +58,14 @@ def Load():
       """Unzip the poisson distribution for up to 5+ samples"""
       features.add(feature)
       t_sum = 0.0
-      for k in range(0, 5):
+      for k in range(0, 4):
         descriptor = str(k)
         prob = gamma**k * math.exp(-gamma) / math.factorial(k)
         potential[roomcat, feature, descriptor] = prob
         t_sum += prob
         feature_space[feature].add(descriptor)
-      potential[roomcat, feature, '5+'] = 1.0-t_sum
-      feature_space[feature].add('5+')
+      potential[roomcat, feature, '4+'] = 1.0-t_sum
+      feature_space[feature].add('4+')
     elif line.strip() == '':
       pass
     else:
@@ -90,8 +90,9 @@ def MakeDistributions():
         try:
           pot = potential[roomcat, feature, descriptor]
         except KeyError:
+          pot = 0.0
           print('Missing potential for: %s, using 0.0' % ((roomcat, feature, descriptor),))
-        f_dist.append( (0.0, descriptor) )
+        f_dist.append( (pot, descriptor) )
       
       f_dist = dataset.DiscreteDistribution(f_dist)
       out.append((feature, f_dist))

### 2011-04-05 dataset/distribution.py
Added sort on discrete distribution

@@ -53,7 +53,7 @@ class DiscreteDistribution:
     '''Creates a discrete distribution over a list of distribution.'''
     self.acc_dists  = []
     self.total_prob = 0.0
-    for prob, dist in iter(dist_list):
+    for prob, dist in sorted(iter(dist_list), reverse=True):
       assert prob >= 0.0
       self.acc_dists.append((self.total_prob, dist))
       self.total_prob += prob

### 2011-04-05 dataset/distribution.py
Added function to iterate over discrete distributions.

@@ -67,3 +67,8 @@ class DiscreteDistribution:
       prev = dist
     return GenerateSample(prev)
 
+  def IterDistributions(self):
+    probs = map(lambda x:x[0], self.acc_dists[1:]) + [self.total_prob]
+    dists = map(lambda x:x[1], self.acc_dists)
+    return zip(probs, dists)
+

### 2011-04-05 semisupervised/notes.tex
Started scripts to pretty print results directly form python to LaTeX.

@@
 
+\section{Experiment - Simple}
+
+\note{This first experience is very small and was only used to better explain the effects of the threshold functions.}
+
+For a better understanding of how the knowledge of the unconditional probability $P(x)$ can help to improve the detection threshold
+a very small toy experiment was run:
+
+The detection system is given access to labelled data ($known = \{kitchen, corridor\}$)
+filtered from the distribution described in \autoref{tab:simple-distribution} as well to unlabelled samples from it.
+From that input it learns both $P(x)$ and $P(x|class \in known)$.
+Later the whole input space is sorted by the diferent thresholds for analysis.
+
+Since the input space is very small (6 possible inputs), simple accumulation tables where used to learn both distributions.
+
+\begin{table}[h]
+\begin{center}
+\input{simple-data/explain}
+\end{center}
+\caption{\label{tab:simple-distribution}Distribution used on the simple experiment.}
+\end{table}
+
+\begin{table}[h]
+\begin{center}
+
+\begin{tabular}{ |c| |c|}
+\hline
+$P(x|class \in know)$ & $\frac{P(x|class \in know)}{P(x)}$ \\ \hline
+(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
+(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
+(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
+(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
+(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
+\end{tabular}
+
+\end{center}
+\caption{\label{tab:simple-threshold}Input space sorted by threshold.}
+\end{table}
 \section{Experiment}

### 2011-04-05 semisupervised/simple-data/explain.py

@@ -1,0 +1,56 @@
+import dataset
+
+def ExplainDistributions(features, distributions):
+  out = []
+
+  def feature_descriptor_label(descriptor):
+    return '\\begin{sideways}%s\\end{sideways}' % descriptor
+  
+  table_format = '|c'
+  for feature, feature_space in features:
+    table_format += '|' + ''.join(['c' for x in range(len(feature_space))])
+  table_format += '|'
+  
+  out.append('\\begin{tabular}{%s}' % table_format)
+  out.append('\\hline')
+  '''Print Header'''
+  out.append('')
+  for feature, feature_space in features:
+    out[-1] += ' & \\multicolumn{%d}{|c|}{%s}' % (len(feature_space),feature)
+  out[-1] += '\\\\ \\hline'
+
+  out.append('')
+  for feature, feature_space in features:
+    for descriptor in feature_space:
+      out[-1] += ' & %s' % feature_descriptor_label(descriptor)
+  out[-1] += '\\\\ \\hline'
+  
+  def explain_distribution(name, definition):
+
+    out.append(feature_descriptor_label(name))
+    for feature, feature_space in features:
+      for descriptor in feature_space:
+        out[-1] += ' & $%.2f\\%%$' % 0.23
+    out[-1] += '\\\\ \\hline'
+  map(lambda x: explain_distribution(x[0], x[1]), distributions)
+  
+
+  out.append('\\end{tabular}')
+  return  '\n'.join(out)
+
+'''
+\begin{tabular}{|c|ccc|cc|}
+\hline
+Label & \multicolumn{3}{c}{Size} & \multicolumn{2}{c}{Shape} \\
+      & \begin{sideways}small\end{sideways} & \begin{sideways}medium\end{sideways} & \begin{sideways}large\end{sideways} & square & round \\
+\hline \\
+kitchen & 0.4 & 0.3 & 0.4 & 0.2 & 0.8 \\
+\hline
+\end{tabular}
+
+
+\begin{description}
+\item[Label] Kitchen
+\item[Room Size] 
+\end{description}
+'''

### 2011-04-06 util/__init__.py
Added writefile function. This should be used to all scripts that write to files.
The idea is that all filenames pass by this module and later to implement an argument parsing on it
so it becomes possible to change all output locations.

@@ -29,4 +29,8 @@ def Histogram(data = None):
       h[sample] += 1
   return h
 
-
+def WriteFile(filename, output):
+  print 'Writting:', filename
+  file = open(filename, 'w')
+  file.write(output)
+  file.close()

### 2011-04-06 dataset/distribution.py
Added method to iterate the distributions on a DiscreteDistributions.

@@ -69,6 +69,10 @@ class DiscreteDistribution:
 
   def IterDistributions(self):
     probs = map(lambda x:x[0], self.acc_dists[1:]) + [self.total_prob]
+    print probs
+    for i in range(len(probs)-1, 0, -1):
+      probs[i] -= probs[i-1]
+    print probs
     dists = map(lambda x:x[1], self.acc_dists)
     return zip(probs, dists)
 

### 2011-04-06 semisupervised/notes.tex
Initial commit of scrpits used to generate tables in tex about the experiments run.

@@
 \end{center}
-\caption{\label{tab:simple-distribution}Distribution used on the simple experiment.}
+\caption{\label{tab:simple-distribution}Distribution used on the simple experiment. Each column cell shows $P(feature|class)$}
 \end{table}
 
-\begin{table}[h]
-\begin{center}
+\begin{table}
+\centering
+\subfloat[Input space sorted by threshold $P(x|class \in known)$]{\input{simple-data/density_threshold}}
+\qquad
+\subfloat[Input space sorted by threshold $P(x|class \in known)/P(x)$]{\input{simple-data/semi_threshold}}
+\caption{\label{tab:simple-data-sorted}Input space sorted by threshold functions.}
+\end{table}
 
-\begin{tabular}{ |c| |c|}
-\hline
-$P(x|class \in know)$ & $\frac{P(x|class \in know)}{P(x)}$ \\ \hline
-(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
-(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
-(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
-(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
-(kitchen, square) = 7.12 & (corridor, elongated) = 2 \\ \hline
-\end{tabular}
+\begin{comment}
+\begin{table}
+\centering
+\begin{minipage}[b]{0.4\textwidth}\centering
+  \input{simple-data/density_threshold}
+\end{minipage}
+\begin{minipage}[b]{0.4\textwidth}\centering
+  \input{simple-data/semi_threshold}
+\end{minipage}
 
-\end{center}
-\caption{\label{tab:simple-threshold}Input space sorted by threshold.}
+\caption{Input space sorted by threshold functions.}
 \end{table}
+\end{comment}
+
+As possible to see in \autoref{tab:simple-data-sorted} the access to unlabelled data allowed the system to better
+
 \section{Experiment}

### 2011-04-06 semisupervised/simple-data/explain.py

@@ -26,31 +26,29 @@ def ExplainDistributions(features, distributions):
   out[-1] += '\\\\ \\hline'
   
   def explain_distribution(name, definition):
-
+    feature_distrib = dict([(f, d) for (f,d) in definition if f != 'label'])
     out.append(feature_descriptor_label(name))
     for feature, feature_space in features:
+      assert feature in feature_distrib
+      potential = dict([(k,p) for (p, k) in feature_distrib[feature].IterDistributions()])
       for descriptor in feature_space:
-        out[-1] += ' & $%.2f\\%%$' % 0.23
+        out[-1] += ' & $%.1f\\%%$' % (potential[descriptor]*100.0)
     out[-1] += '\\\\ \\hline'
   map(lambda x: explain_distribution(x[0], x[1]), distributions)
   
 
   out.append('\\end{tabular}')
-  return  '\n'.join(out)
-
-'''
-\begin{tabular}{|c|ccc|cc|}
-\hline
-Label & \multicolumn{3}{c}{Size} & \multicolumn{2}{c}{Shape} \\
-      & \begin{sideways}small\end{sideways} & \begin{sideways}medium\end{sideways} & \begin{sideways}large\end{sideways} & square & round \\
-\hline \\
-kitchen & 0.4 & 0.3 & 0.4 & 0.2 & 0.8 \\
-\hline
-\end{tabular}
+  return '\n'.join(out)
 
 
-\begin{description}
-\item[Label] Kitchen
-\item[Room Size] 
-\end{description}
-'''
+
+def SortThresholds(title, inputs, func, columns = 1):
+  out = []
+  col = ''.join(['l' for c in range(columns)])
+  out.append('\\begin{tabular}{%sr}' % col)
+  #out.append('\\hline \\multicolumn{2}{c}{'+title+'}\\\\ \\hline')
+  out.append('\\multicolumn{%d}{c}{sample} & threshold \\\\ \\hline' % columns)
+  for threshold, sample in sorted(map(lambda x: (func(x[1]), x[0]), inputs), reverse=True):
+    out.append('%s & %.3f \\\\' % (sample, threshold))
+  out.append('\\end{tabular}')
+  return '\n'.join(out)

### 2011-04-06 semisupervised/simple-data/run.py

@@ -1,0 +1,107 @@
+#!/usr/bin/python
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+import ml
+import graph
+import dataset
+import util
+import explain 
+
+def Label(label):  return ('label', label)
+def Appearance(d): return ('Appearance', dataset.DiscreteDistribution(d))
+def Shape(d):      return ('Shape',      dataset.DiscreteDistribution(d))
+
+features = [ ('Appearance', ['kitchen', 'office', 'corridor']),
+             ('Shape', ['square', 'elongated'])]
+
+input_space_size = reduce(lambda x,y:x*y, map(len, [x[1] for x in features]))
+
+input_space = [[]]
+for feature, descriptors in features:
+  tmp = []
+  for descriptor in descriptors:
+    for input in input_space:
+      tmp.append(input + [(feature,descriptor)])
+  input_space = tmp
+
+input_space = map(tuple, input_space)
+
+kitchen = ( Label('kitchen'),
+            Appearance([(0.70, 'kitchen'),
+                        (0.29, 'office' ),
+                        (0.01, 'corridor')]),
+            Shape([(0.60, 'square'),
+                   (0.40, 'elongated')])
+          )
+
+corridor = ( Label('corridor'),
+             Appearance([(0.05, 'kitchen'),
+                         (0.05, 'office'),
+                         (0.90, 'corridor')]),
+             Shape([(0.05, 'square'),
+                    (0.95, 'elongated')])
+           )
+
+office   = ( Label('office'),
+             Appearance([(0.30, 'kitchen'),
+                         (0.69, 'office'),
+                         (0.01, 'corridor')]),
+             Shape([(0.6, 'square'),
[…]

### 2011-04-06 semisupervised/notes.tex
Advances on the simple experiment to demonstrate the importance of estimating P(x)

--- 2011-04-04 semisupervised/notes.tex
@@
 
+Proposed detection method:
+\begin{itemize}
+\item From the environment draw labelled and unlabelled samples.
+\item From those learn both $P(x|class \in known)$ and $P(x)$.
+\item Define a detection threshold based on $P(x|class \in known)/P(x)$.
+\end{itemize}
+
 \section{Threshold Analysis}
@@
 
-\section{Experiment}
+\section{Experiments}
 \label{sec:experiment}
-
 \begin{itemize}
@@
 
+\subsection{Simple}
+\note{This first experience is very small and was only used to better explain the effects of the threshold functions.}
+
+For a better understanding of how the knowledge of the unconditional probability $P(x)$ can help to improve the detection threshold
+a very small toy experiment was run:
+
+\begin{table}[h]
+\begin{center}
+\input{simple-data/explain}
+\end{center}
+\caption{\label{tab:simple-distribution}Distribution used on the simple experiment. Each column cell shows $P(feature|class)$}
+\end{table}
+
+Using a simple distribution, described in \autoref{tab:simple-distribution},
+labelled ($known = \{kitchen, corridor\}$)
+and unlabelled samples where drawn.
+$P(x|class \in known)$ and $P(x)$ where then approximated using simple accumulation tables
+(only 6 diferent inputs).
+The input space was then sorted according to the presented thresholds obtaining
+the results as seen on \autoref{tab:simple-data-sorted}.
+
+\begin{table}
+\centering
+\subfloat[Input space sorted by threshold $P(x|class \in known)$]{\input{simple-data/density_threshold}}
+\qquad
+\subfloat[Input space sorted by threshold $P(x|class \in known)/P(x)$]{\input{simple-data/semi_threshold}}
+\caption{\label{tab:simple-data-sorted}Input space sorted by threshold functions.}
+\end{table}
+
+The results obtained demonstrate that using the conditional probability function
+as a novelty threshold leads to poor performance: it classified (corridor,square) as
+the most novel case.
+This happens because the likelihood of such a sample is very low.
+Nonetheless when using the factor $P(x)$ the system is able to take in consideration
+that the low probability of that sample on the labelled data arrives from the low
+probability of finding such a sample and not necessarily from a novel case.
+
+This simple toy experience demonstrates the importance of using $P(x)$ to estimate a novelty
+threshold and it importance when dealing with features that have a tedency to be highly bias
+such as the semantic data used on our sample.
+
+\subsection{Dora with world model}
+
 \bibliographystyle{plainnat}

### 2011-04-06 semisupervised/simple-data/run.py

@@ -17,11 +17,14 @@
 # along with this program; if not, write to the Free Software Foundation,
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
+import random
 import ml
 import graph
 import dataset
 import util
 import explain 
+
+random.seed(0)
 
 def Label(label):  return ('label', label)
 def Appearance(d): return ('Appearance', dataset.DiscreteDistribution(d))
@@ -66,7 +69,7 @@ office   = ( Label('office'),
                     (0.4, 'elongated')])
            )
 
-all_classes = [(0.4, corridor), (0.4, kitchen), (0.4, office)]
+all_classes = [(1, corridor), (1, kitchen), (1, office)]
 known_labels = ['corridor', 'kitchen']
 known_classes = [all_classes[0], all_classes[1]]
 
@@ -85,7 +88,6 @@ def density_threshold(sample):
   return conditional_prob(sample)
 
 def semi_threshold(sample):
-  print 'Sample = ', sample
   return conditional_prob(sample) / unconditional_prob(sample)
 
 

### 2011-04-06 semisupervised/dora-data/ontology.py
Some scripts to run experiments on the world considered on Dora.

@@ -29,6 +29,7 @@ potential = dict()
 roomcats  = set()
 features  = set()
 feature_space = defaultdict(set)
+poisson_expansion = 3
 
 def Load():
   print("Loading default knowledge...")
@@ -55,17 +56,17 @@ def Load():
       obj, roomcat, gamma = match.groups()
       feature = 'object_%s' % obj
       gamma = float(gamma)
-      """Unzip the poisson distribution for up to 5+ samples"""
+      """Unzip the poisson distribution for up to the given number of samples"""
       features.add(feature)
       t_sum = 0.0
-      for k in range(0, 4):
+      for k in range(0, poisson_expansion):
         descriptor = str(k)
         prob = gamma**k * math.exp(-gamma) / math.factorial(k)
         potential[roomcat, feature, descriptor] = prob
         t_sum += prob
         feature_space[feature].add(descriptor)
-      potential[roomcat, feature, '4+'] = 1.0-t_sum
-      feature_space[feature].add('4+')
+      potential[roomcat, feature, '%d+' % poisson_expansion] = 1.0-t_sum
+      feature_space[feature].add('%d+' % poisson_expansion)
     elif line.strip() == '':
       pass
     else:

### 2011-04-06 semisupervised/dora-data/run.py

@@ -1,0 +1,37 @@
+#!/usr/bin/python
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+import ontology
+import util
+import explain
+import random
+random.seed(0)
+
+
+ontology.Load()
+
+rooms = ontology.MakeDistributions()
+features = ontology.feature_space
+for f in features:
+  features[f] = sorted(features[f])
+del features['room_category2']
+
+util.WriteFile('explain.tex',
+    explain.ExplainDistributions(sorted(features.items()),
+                                 sorted(rooms.items())))

### 2011-04-06 dataset/__init__.py
Added function to calculate exact probability of a sample on a distribution

@@ -19,7 +19,7 @@
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
 from Sampler import SampleN
-from distribution import DiscreteDistribution, GenerateSample, Generator
+from distribution import DiscreteDistribution, GenerateSample, Generator, SampleProbability
 
 def ExtractLabel(sample):
   assert sample[0][0] == 'label'

### 2011-04-06 dataset/distribution.py

@@ -69,10 +69,26 @@ class DiscreteDistribution:
 
   def IterDistributions(self):
     probs = map(lambda x:x[0], self.acc_dists[1:]) + [self.total_prob]
-    print probs
     for i in range(len(probs)-1, 0, -1):
       probs[i] -= probs[i-1]
-    print probs
     dists = map(lambda x:x[1], self.acc_dists)
     return zip(probs, dists)
 
+def SampleProbability(distribution, sample):
+  if hasattr(distribution, 'IterDistributions'):
+    p = 0.0
+    tl = 0.0
+    for likelihood, d in distribution.IterDistributions():
+      p += likelihood * SampleProbability(d, sample)
+      tl += likelihood
+    return p/tl
+  elif type(distribution) == type(()):
+    if type(distribution[0]) == type(()) and distribution[0][0] == 'label':
+      distribution = list(distribution[1:])
+    p = 1.0
+    assert len(distribution) == len(sample)
+    for a, b in zip(distribution, sample):
+      p *= SampleProbability(a, b)
+    return p
+  else:
+    return distribution == sample

### 2011-04-06 semisupervised/notes.tex
Some scripts to run experiments on the world considered on Dora.
Made comments to visually split chapters on the LaTeX souce.
Added ROC curve plotting of novelty thresholds on dora world with perfect distirbution information.

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Concept Detection}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Threshold}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Threshold Analysis}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Experiments}
@@
 \subsection{Simple}
+\label{sec:simple-experiment}
 \note{This first experience is very small and was only used to better explain the effects of the threshold functions.}
@@
 a very small toy experiment was run:
-
-\begin{table}[h]
-\begin{center}
-\input{simple-data/explain}
-\end{center}
-\caption{\label{tab:simple-distribution}Distribution used on the simple experiment. Each column cell shows $P(feature|class)$}
-\end{table}
 
@@
 the results as seen on \autoref{tab:simple-data-sorted}.
+
+\begin{table}
+\begin{center}
+\input{simple-data/explain}
+\end{center}
+\caption{\label{tab:simple-distribution}Distribution used on the simple experiment. Each column cell shows $P(feature|class)$}
+\end{table}
 
@@
 
+This experience is similar to the simple experiment run in \autoref{sec:simple-experiment}.
+But this time using a more realistic distribution (\autoref{tab:dora-world-distribution}) with
+11 diferent room categories and with a feature space containing semantic data for:
+room size, room shape, visual appearance and 6 diferent objects.
+Also, the features are statistically independent from each other given the room category.
+
+The system objective was to detect novel data from a given known set of 6 categories.
+In order to use drawn results two tests where run:
+\begin{enumerate}
+\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{tab:dora-data-perfect}).
+\item A simple model was used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
+\end{enumerate}
+
+\begin{figure}
+\centering
+\includegraphics[width=\textwidth]{dora-data/roc-perfect.pdf}
+\caption{\label{tab:dora-data-perfect}ROC curve for thresholds based on perfect information running on the dora distribution}
+\end{figure}
+
+\begin{figure}
+\centering
+\include{dora-data/roc-simple.tex}
+\caption{\label{tab:dora-data-simple}ROC curve for thresholds learned on samples from samples in the dora distribution}
+\end{figure}
+
+\begin{sidewaystable}[h]
+\begin{center}
+\scalebox{0.35}{
+\input{dora-data/explain}
+}
+\end{center}
+\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
+\end{sidewaystable}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\cleardoublepage
+\phantomsection
+\addcontentsline{toc}{section}{References}
+
 \bibliographystyle{plainnat}

### 2011-04-06 semisupervised/dora-data/run.py

@@ -21,11 +21,26 @@ import ontology
 import util
 import explain
 import random
+import dataset
+import graph
+
 random.seed(0)
+
+def plot_roc(threshold, samples, known_labels, title=''):
+  def ThresholdAndLabel(sample):
+    label, sample = dataset.ExtractLabel(sample)
+    return threshold(sample), label
+
+  roc_curve = []
+  #for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x[0], reverse=True):
+  for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x, reverse=True):
+    roc_curve.append( label in known_labels )
+  graph.roc(roc_curve, label = title)
+  pass
+
 
 
 ontology.Load()
-
 rooms = ontology.MakeDistributions()
 features = ontology.feature_space
 for f in features:
@@ -35,3 +50,32 @@ del features['room_category2']
 util.WriteFile('explain.tex',
     explain.ExplainDistributions(sorted(features.items()),
                                  sorted(rooms.items())))
+
+known_labels = set(['kitchen', 'robotlab', 'singleoffice'])
+world = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items()])
+known = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items() if cat in known_labels])
+
+def perfect_conditional_prob(sample):
+  return dataset.SampleProbability(known, sample)
+
+def perfect_unconditional_prob(sample):
+  return dataset.SampleProbability(world, sample)
+
+def perfect_density_threshold(sample): return perfect_conditional_prob(sample)
+def perfect_semi_threshold(sample):    return perfect_conditional_prob(sample)/perfect_unconditional_prob(sample)
+
+
+test_data = dataset.LabelledSample(world, 5000)
+plot_roc(threshold = perfect_density_threshold,
+         samples   = test_data,
+         known_labels = known_labels,
+         title= 'ROC for P(x|c) threshold')
+
+plot_roc(threshold = perfect_semi_threshold,
+         samples   = test_data,
+         known_labels = known_labels,
+         title= 'ROC for P(x|c)/P(x) threshold')
+
+graph.savefig('perfect-roc.tex')
+graph.show()
+

### 2011-04-07 semisupervised/dora-data/ontology.py
Small fixs.

@@ -85,7 +85,7 @@ def MakeDistributions():
   def MakeDistribution(roomcat):
     out = []
     out.append(('label', roomcat))
-    for feature in features:
+    for feature in feature_space:
       f_dist = []
       for descriptor in feature_space[feature]:
         try:

### 2011-04-07 semisupervised/dora-data/run.py

@@ -32,7 +32,6 @@ def plot_roc(threshold, samples, known_labels, title=''):
     return threshold(sample), label
 
   roc_curve = []
-  #for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x[0], reverse=True):
   for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x, reverse=True):
     roc_curve.append( label in known_labels )
   graph.roc(roc_curve, label = title)
@@ -41,11 +40,13 @@ def plot_roc(threshold, samples, known_labels, title=''):
 
 
 ontology.Load()
-rooms = ontology.MakeDistributions()
+del ontology.feature_space['room_category2']
+
 features = ontology.feature_space
 for f in features:
   features[f] = sorted(features[f])
-del features['room_category2']
+
+rooms = ontology.MakeDistributions()
 
 util.WriteFile('explain.tex',
     explain.ExplainDistributions(sorted(features.items()),
@@ -76,6 +77,6 @@ plot_roc(threshold = perfect_semi_threshold,
          known_labels = known_labels,
          title= 'ROC for P(x|c)/P(x) threshold')
 
-graph.savefig('perfect-roc.tex')
+graph.savefig('perfect-roc.pdf')
 graph.show()
 

### 2011-04-08 semisupervised/notes.tex
Ideas and a lot of changes on the report.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
 \section{Threshold Analysis}
@@
 
+\subsection{Bounding $P(c)$}
+\note{This was just a random though... if we can bound $P(c)$ in some way then we get
+      some information on how sure we are about a decision. Though its only possible
+      to bound it from above, and even that not sure if its computational feasible.}
+
 \begin{equation}P(x) = \sum_{c}^{} P(x|c) P(c)\end{equation}
-
 \begin{equation}0 \le P(x|c) P(c) \le P(x)\end{equation}
 \begin{equation}0 \le P(c) \le \frac{P(x)}{P(x|c)}\end{equation}
+
+\subsection{What does a ROC curve looks like for an optimal detector}
+\note{The ROC curve shape shows if a detector is perfect given an input space.
+      (Or at least I think so (though I haven't write any proof)).
+      I guess somewhere there should be literature reference to such a simple concept}
+
+Proof would be something like:
+
+An element of the input space is responsible for a vector change of $(TP, FP)$ in the ROC
+curve.
+A threshold functions orders the input space.
+And the best is to take the greedy approach of sorting the input by $TP/FP$.
+That is equivalent to a convex ROC curve.
+
+This leads to the important property:
+
+whatever threshold function you use,
+if you get a convex ROC curve, there is no way to perform better using that input space.
+
+At the same time we can proof an ordering on $TP/FP$ to be the same as $TP/(FP+TP)$.
+Which is coherent with the threshold of: $P(x|concept)/P(x)$.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Experiments}
+\clearpage
+\section{Evaluation and Experiments}
 \label{sec:experiment}
+
+This section describes the two experiences that were run to evaluate the methods.
+\begin{description}
+\item[\autoref{sec:simple-experiment}] presents a toy experience that aims at
+  showing the importance of using $P(x)$ to estimate a novelty threshold and
+  it importance when dealing with features that have a tendency to be highly
+  biased such as the semantic data used on our sample.
+
+\item[\autoref{sec:dora-experiment}] runs comparisons of the presented threshold
+  on a real-world scenario, where a robot has to identify room categories based
+  on obtained semantic data, such as object presence and room size.
+  The experiment is run both with a fictitious distribution and with a real world
+  dataset.
+\end{description}
+
+Other possible experiences are:
 \begin{itemize}
-\item Define a set of features (all extractable from visual or user-tags).
-\item Define a dataset of inputs (eg.: indoor environs at LabelMe)
-\item Filter those for \emph{kitchen}.
-\item Use ML methods to model $P(features | kitchen)$ and $P(features)$.
-\item Test results.
+\item run this type of classification on datasets such as \emph{LabelMe} databases,
+  either by using user-supplied tags as semantic data or by extracting them with some
+  image analysis.
 \end{itemize}
 
-\subsection{Simple}
-\label{sec:simple-experiment}
-\note{This first experience is very small and was only used to better explain the effects of the threshold functions.}
-
+\subsection{Simple}\label{sec:simple-experiment}
 For a better understanding of how the knowledge of the unconditional probability $P(x)$ can help to improve the detection threshold
@@
 $P(x|class \in known)$ and $P(x)$ where then approximated using simple accumulation tables
-(only 6 diferent inputs).
+(only 6 different inputs).
 The input space was then sorted according to the presented thresholds obtaining
@@
 
-\begin{table}
+\begin{table}[hb]
 \begin{center}
@@
 
-This simple toy experience demonstrates the importance of using $P(x)$ to estimate a novelty
-threshold and it importance when dealing with features that have a tedency to be highly bias
-such as the semantic data used on our sample.
-
 \subsection{Dora with world model}
+\label{sec:dora-experiment}
 
@@
 But this time using a more realistic distribution (\autoref{tab:dora-world-distribution}) with
-11 diferent room categories and with a feature space containing semantic data for:
-room size, room shape, visual appearance and 6 diferent objects.
+11 different room categories and with a feature space containing semantic data for:
+room size, room shape, visual appearance and 6 different objects.
 Also, the features are statistically independent from each other given the room category.
@@
 The system objective was to detect novel data from a given known set of 6 categories.
-In order to use drawn results two tests where run:
+Two tests where run:
 \begin{enumerate}

### 2011-04-08 dataset/Sampler.py
Some cleaning.

@@ -1,192 +1,0 @@
-# -*- coding: utf-8 -*-
-#
-# Novelty Detection Scripts
-#
-# Copyright 2011: André Susano Pinto
-#
-# This program is free software; you can redistribute it and/or
-# modify it under the terms of the GNU General Public License
-# as published by the Free Software Foundation; either version 2
-# of the License, or (at your option) any later version.
-#
-# This program is distributed in the hope that it will be useful,
-# but WITHOUT ANY WARRANTY; without even the implied warranty of
-# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-# GNU General Public License for more details.
-#
-# You should have received a copy of the GNU General Public License
-# along with this program; if not, write to the Free Software Foundation,
-# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
-#
-import random
-from util import Histogram
-
-class Distribution:
-  def Probability(self, sample):
-    """Returns probability of the given input be generated by this distribution."""
-    raise NotImplemented
-  
-  def Generator(self):
-    """Returns an infinity iterator that generates samples for this distribution."""
-    raise NotImplemented
-
-class UniqueDistribution(Distribution):
-  """Defines a distribution that always returns the same value."""
-  
-  def __init__(self, value):
-    self.value = value
-  
-  def Probability(self, sample):
-    if sample == self.value:
-      return 1.0
-    return 0.0
-  
-  def Generator(self):
-    """Returns an iterator that always yield the same value."""
-    def sampler(value):
-      while True:
-        yield value
-    return sampler(self.value)
-
-
-class HistogramDistribution(Distribution):
-  """Defines a discrete distribution according to a given histogram."""
-  
-  def __init__(self, histogram, normalize = True):
-    accumulate = 0.0
-    self.distributions = []
-    for (key, value) in histogram.items():
-      accumulate += value
-      self.distributions.append((accumulate, key))
-    
-    if normalize:
-      """Normalize the histogram."""
-      for i in range(len(self.distributions)):
-        (acc, value) = self.distributions[i]
-        self.distributions[i] = (acc/accumulate, value)
-      self.distributions[-1] = (1.0, self.distributions[-1][1])
-      """Make sure it sums up to one and there is no rounding errors."""
-
-  
-  def Probability(self, sample):
-    last = 0.0
-    prob = 0.0
-    for (acc, value) in self.distributions:
-      if value == sample:
-        prob += acc-last
[…]

### 2011-04-08 dataset/__init__.py

@@ -18,8 +18,13 @@
 # along with this program; if not, write to the Free Software Foundation,
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
-from Sampler import SampleN
 from distribution import DiscreteDistribution, GenerateSample, Generator, SampleProbability
+
+def SampleN(max_samples, data):
+  data = iter(data)
+  for x in range(max_samples):
+    yield next(data)
+
 
 def ExtractLabel(sample):
   assert sample[0][0] == 'label'

### 2011-04-08 semisupervised/dora-data/run.py
Added experience 2 with dora world model.
In it, a class dependent and and a class independent probability distribution is learned on "known" density.
And a simple not-class-aware is used to model the unknown.

This is interesting as its makes it possible to model using a graphical models.

TODO: distributions are very strickyly calculated atm and they return 0 for values never seen.
That should be smoothed out.

@@ -26,6 +26,70 @@ import graph
 
 random.seed(0)
 
+import collections
+import itertools
+
+def NormalizedHistogram(data):
+  count_total = 0
+  count_value = collections.defaultdict(int)
+  for sample in data:
+    count_value[sample] += 1
+    count_total += 1
+  
+  def prob(sample):
+    return count_value[sample] / float(count_total)
+  
+  return prob
+
+def ClassDependentEstimator(data, class_estimator, classifier = lambda x: (x[0], x[1])):
+  class_data = collections.defaultdict(list)
+  class_labels = []
+  for sample in data:
+    cname, cdata = classifier(sample)
+    class_data[cname].append(cdata)
+    class_labels.append(cname)
+  class_prob = NormalizedHistogram(class_labels)
+  class_estimator = dict([(cname, class_estimator(cdata)) for cname, cdata in class_data.items()])
+  
+  def prob(sample):
+    p = 0
+    for cname, cestimator in class_estimator.items():
+      p += class_prob(cname)*cestimator(sample)
+    return p
+  return prob
+
+
+def IndependentFeatureEstimator(data, feature_estimator, feature_iter = iter):
+  feature_data = collections.defaultdict(list)
+  for sample in data:
+    for feature_id, feature_descriptor in feature_iter(sample):
+      feature_data[feature_id].append(feature_descriptor)
+  feature_prob = dict([(fid, feature_estimator(fdata)) for fid,fdata in feature_data.items()])
+  
+  def prob(sample):
+    p = 1.0
+    for fid, fsample in feature_iter(sample):
+      p *= feature_prob[fid](fsample)
+    return p
+  
+  return prob
+
+
+'''
+a = IndependentFeatureEstimator([
+  ( ('label', 'kitchen'),  ('size', 'small') ),
+  ( ('label', 'corridor'), ('size', 'long') ),
+  ( ('label', 'kitchen'),  ('size', 'long') )], NormalizedHistogram)
+
+print a( (('label', 'corridor'), ))
+print a( (('size', 'small'), ))
+print a( (('label', 'corridor'), ('size', 'small'), ))
+
+import sys
+sys.exit(0)
+'''
+
+#############################################################################
 def plot_roc(threshold, samples, known_labels, title=''):
   def ThresholdAndLabel(sample):
     label, sample = dataset.ExtractLabel(sample)
@@ -56,6 +120,9 @@ known_labels = set(['kitchen', 'robotlab', 'singleoffice'])
[…]

### 2011-04-09 graph/__init__.py
Added newfig

@@ -8,3 +8,6 @@ def show():
 
 def savefig(path):
   pl.savefig(path)
+
+def newfig():
+  pl.figure()

### 2011-04-09 ml/__init__.py
Added code that can be used to perform simple probabilities estimations.
Code was moved form the thesis repository.

@@ -20,14 +20,49 @@
 #
 from Classifier import *
 from util import Histogram
+import collections
+import itertools
 
-def DiscreteProbabilityEstimator(samples, state_size = 0):
-  counter = Histogram(samples)
-  print counter
-  total = state_size + sum(map(lambda x: x[1], counter.items()), 0)
-  def Density(sample):
-    if sample in counter:
-      return (counter[sample]+1.0) / total 
-    else:
-      return 1.0 / total
-  return Density
+def NormalizedHistogram(data):
+  count_total = 0
+  count_value = collections.defaultdict(int)
+  for sample in data:
+    count_value[sample] += 1
+    count_total += 1
+  
+  def prob(sample):
+    return count_value[sample] / float(count_total)
+  
+  return prob
+
+def ClassDependentEstimator(data, class_estimator, classifier = lambda x: (x[0], x[1])):
+  class_data = collections.defaultdict(list)
+  class_labels = []
+  for sample in data:
+    cname, cdata = classifier(sample)
+    class_data[cname].append(cdata)
+    class_labels.append(cname)
+  class_prob = NormalizedHistogram(class_labels)
+  class_estimator = dict([(cname, class_estimator(cdata)) for cname, cdata in class_data.items()])
+  
+  def prob(sample):
+    p = 0
+    for cname, cestimator in class_estimator.items():
+      p += class_prob(cname)*cestimator(sample)
+    return p
+  return prob
+
+def IndependentFeatureEstimator(data, feature_estimator, feature_iter = iter):
+  feature_data = collections.defaultdict(list)
+  for sample in data:
+    for feature_id, feature_descriptor in feature_iter(sample):
+      feature_data[feature_id].append(feature_descriptor)
+  feature_prob = dict([(fid, feature_estimator(fdata)) for fid,fdata in feature_data.items()])
+  
+  def prob(sample):
+    p = 1.0
+    for fid, fsample in feature_iter(sample):
+      p *= feature_prob[fid](fsample)
+    return p
+  
+  return prob

### 2011-04-09 semisupervised/notes.tex
Added experience 2 with dora world model.
In it, a class dependent and and a class independent probability distribution is learned on "known" density.
And a simple not-class-aware is used to model the unknown.

This is interesting as its makes it possible to model using a graphical models.

TODO: distributions are very strickyly calculated atm and they return 0 for values never seen.
That should be smoothed out.
Updates

@@
 \item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{tab:dora-data-perfect}).
-\item A simple model was used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
+\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
 \end{enumerate}
+
+\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
+      to the probabilities functions to not return 0. This is because at the moment if on the known
+      samples we never see a given value we don't know it belongs to the possible feature space.
+      Has to be corrected.}
 
@@
 \centering
-\include{dora-data/roc-simple.tex}
+\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
 \caption{\label{tab:dora-data-simple}ROC curve for thresholds learned on samples from samples in the dora distribution}
@@
 \begin{center}
-\scalebox{0.35}{
-\input{dora-data/explain}
-}
+\scalebox{0.35}{\input{dora-data/explain}}
 \end{center}
@@
 \end{sidewaystable}
+
+\subsection{How Dora models the world?}
+\label{sec:dora-models}
+
+Currently dora uses a \emph{probabilistic graphical model} to model the world.
+All variables that the system see are added and connected in between with the
+\emph{probabilistic factors} that the system is aware of.
+Those factors are learned from the labelled data that the system is given when
+training.
+
+The proposal here is to also use such an approach to model the unconditional
+probability.
+Under that a novelty detection threshold can be easily implemented
+reusing the \emph{probabilistic graphical model} $G$ that dora uses as
+representation of the world:
+\begin{description}
+\item[$P(x|know)$] is obtained evaluating $P(G)$.
+\item[$P(x)$] can be obtained by creating $G'$ as an instance of $G$ but with the
+variable node we want to perform novelty detection removed and the factors associated
+with it replaced with factors learned from the unlabelled data.
+\end{description}
+
+We point to an additional simplification that if the original graphical model factors
+depending on the variable to perform novelty detection were previously split on a
+conditional and unconditional factors. Then $G'$ could be obtained by simply removing
+all the factors associated with the variable to perform novelty detection on.
 

### 2011-04-09 semisupervised/dora-data/run.py

@@ -23,71 +23,9 @@ import explain
 import random
 import dataset
 import graph
+import ml
 
 random.seed(0)
-
-import collections
-import itertools
-
-def NormalizedHistogram(data):
-  count_total = 0
-  count_value = collections.defaultdict(int)
-  for sample in data:
-    count_value[sample] += 1
-    count_total += 1
-  
-  def prob(sample):
-    return count_value[sample] / float(count_total)
-  
-  return prob
-
-def ClassDependentEstimator(data, class_estimator, classifier = lambda x: (x[0], x[1])):
-  class_data = collections.defaultdict(list)
-  class_labels = []
-  for sample in data:
-    cname, cdata = classifier(sample)
-    class_data[cname].append(cdata)
-    class_labels.append(cname)
-  class_prob = NormalizedHistogram(class_labels)
-  class_estimator = dict([(cname, class_estimator(cdata)) for cname, cdata in class_data.items()])
-  
-  def prob(sample):
-    p = 0
-    for cname, cestimator in class_estimator.items():
-      p += class_prob(cname)*cestimator(sample)
-    return p
-  return prob
-
-
-def IndependentFeatureEstimator(data, feature_estimator, feature_iter = iter):
-  feature_data = collections.defaultdict(list)
-  for sample in data:
-    for feature_id, feature_descriptor in feature_iter(sample):
-      feature_data[feature_id].append(feature_descriptor)
-  feature_prob = dict([(fid, feature_estimator(fdata)) for fid,fdata in feature_data.items()])
-  
-  def prob(sample):
-    p = 1.0
-    for fid, fsample in feature_iter(sample):
-      p *= feature_prob[fid](fsample)
-    return p
-  
-  return prob
-
-
-'''
-a = IndependentFeatureEstimator([
-  ( ('label', 'kitchen'),  ('size', 'small') ),
-  ( ('label', 'corridor'), ('size', 'long') ),
-  ( ('label', 'kitchen'),  ('size', 'long') )], NormalizedHistogram)
-
-print a( (('label', 'corridor'), ))
-print a( (('size', 'small'), ))
-print a( (('label', 'corridor'), ('size', 'small'), ))
-
-import sys
-sys.exit(0)
-'''
 
 #############################################################################
 def plot_roc(threshold, samples, known_labels, title=''):
@@ -121,7 +59,14 @@ world = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items()])
[…]

### 2011-04-11 semisupervised/notes.tex
Some  more fixes

@@
 \note{The ROC curve shape shows if a detector is perfect given an input space.
-      (Or at least I think so (though I haven't write any proof)).
       I guess somewhere there should be literature reference to such a simple concept}
@@
 This leads to the important property:
-
-whatever threshold function you use,
-if you get a convex ROC curve, there is no way to perform better using that input space.
+\emph{whatever threshold function you use,
+if you get a convex ROC curve, there is no way to perform better using that input space.}
 
@@
 \begin{enumerate}
-\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{tab:dora-data-perfect}).
-\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{tab:dora-data-simple}).
+\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{fig:dora-data-perfect}).
+\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{fig:dora-data-simple}).
 \end{enumerate}
@@
 
+As possible to see in \autoref{fig:dora-data-perfect} if we are able to use the unlabelled data to build a correct $P(x)$ then we obtain an optimal detector.
+
+\autoref{fig:dora-data-simple} shows how close our solution approaches the optimal by using creating models to for the conditional and unconditional probability from samples.
+
+Important to notice that our model of unconditional probability is very simple, it assumes semantic data to be independent from each other.
+
+\note{Describe better which approach we used to model $P(G)/P(G')$}
+\note{Perform experience with real data}
+
 \begin{figure}
@@
 \includegraphics[width=\textwidth]{dora-data/roc-perfect.pdf}
-\caption{\label{tab:dora-data-perfect}ROC curve for thresholds based on perfect information running on the dora distribution}
+\caption{ROC curve for thresholds based on perfect information running on the dora distribution}
+\label{fig:dora-data-perfect}
 \end{figure}
@@
 \includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
-\caption{\label{tab:dora-data-simple}ROC curve for thresholds learned on samples from samples in the dora distribution}
+\caption{ROC curve for thresholds learned on samples from samples in the dora distribution}
+\label{fig:dora-data-simple}
 \end{figure}

### 2011-04-11 semisupervised/dora-data/run.py

@@ -103,15 +103,15 @@ graph.newfig()
 p_roc('ROC for optimal threshold',
       lambda sample: perfect_conditional_prob(sample)/perfect_unconditional_prob(sample))
 
-p_roc('ROC for CD(x|c)/P(x) threshold',
-      lambda sample: (0.0001+estimated_class_conditional_prob(sample)) / perfect_unconditional_prob(sample))
+#p_roc('ROC for CD(x|c)/P(x) threshold',
+#      lambda sample: (0.0001+estimated_class_conditional_prob(sample)) / perfect_unconditional_prob(sample))
 
-p_roc('ROC for CD(x|c) threshold',
+p_roc('ROC for P(G) threshold',
       lambda sample: 0.0001+estimated_conditional_prob(sample))
 
-p_roc('ROC for estimated CD(x|c)/EP(x) threshold',
+p_roc('ROC for estimated P(G)/P(G\') threshold',
       lambda sample: (0.0001+estimated_class_conditional_prob(sample)) / (0.0001+estimated_unconditional_prob(sample)))
 
-p_roc('ROC for estimated CI(x|c)/EP(x) threshold',
-      lambda sample: (0.0001+estimated_conditional_prob(sample)) / (0.0001+estimated_unconditional_prob(sample)))
+#p_roc('ROC for estimated CI(x|c)/EP(x) threshold',
+#      lambda sample: (0.0001+estimated_conditional_prob(sample)) / (0.0001+estimated_unconditional_prob(sample)))
 graph.savefig('roc-simple.pdf')

### 2011-04-11 semisupervised/simple-data/run.py

@@ -81,8 +81,9 @@ unlabelled_data = dataset.UnlabelledSample(all_classes, samples = 10000)
 labelled_data   = dataset.LabelledSample(known_classes, samples = 1000)
 test_data       = dataset.LabelledSample(all_classes, samples = 1000)
 
-unconditional_prob = ml.DiscreteProbabilityEstimator(unlabelled_data, state_size = input_space_size)
-conditional_prob   = ml.DiscreteProbabilityEstimator(map(dataset.FilterLabel, labelled_data), state_size = input_space_size)
+# TODO: change this to use the perfect probability estimators.
+unconditional_prob = ml.NormalizedHistogram(unlabelled_data)
+conditional_prob   = ml.NormalizedHistogram(map(dataset.FilterLabel, labelled_data))
 
 def density_threshold(sample):
   return conditional_prob(sample)

### 2011-04-12 graph/ROC.py
updates on roc plotting

@@ -5,7 +5,7 @@
 import numpy as N
 import pylab as P
 
-def ROC(data, label = None):
+def ROC(data, style = 'ro-', label = None):
   """Plots a ROC curve."""
   
   def acc(data, value):
@@ -20,11 +20,10 @@ def ROC(data, label = None):
   if fpr[-1] != 0: fpr /= fpr[-1]
   if tpr[-1] != 0: tpr /= tpr[-1]
   
-  P.plot(fpr, tpr, label = label)
+  P.plot(fpr, tpr, style, markevery=len(fpr)/10, label = label)
   P.xlim([-0.05,1.05])
   P.ylim([-0.05,1.05])
   P.xlabel('False Positive Rate')
   P.ylabel('True Positive Rate')
-  P.title('Receiver operating characteristic')
   P.legend(loc = 'best')
 

### 2011-04-12 semisupervised/dora-data/run.py
Changes on plotting so they can be seen on black and white

@@ -28,7 +28,7 @@ import ml
 random.seed(0)
 
 #############################################################################
-def plot_roc(threshold, samples, known_labels, title=''):
+def plot_roc(threshold, samples, known_labels, style, title=''):
   def ThresholdAndLabel(sample):
     label, sample = dataset.ExtractLabel(sample)
     return threshold(sample), label
@@ -36,7 +36,7 @@ def plot_roc(threshold, samples, known_labels, title=''):
   roc_curve = []
   for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x, reverse=True):
     roc_curve.append( label in known_labels )
-  graph.roc(roc_curve, label = title)
+  graph.roc(roc_curve, style, label = title)
   pass
 
 
@@ -54,7 +54,8 @@ util.WriteFile('explain.tex',
     explain.ExplainDistributions(sorted(features.items()),
                                  sorted(rooms.items())))
 
-known_labels = set(['kitchen', 'robotlab', 'singleoffice'])
+all_labels = rooms.keys()
+known_labels = all_labels[1:]
 world = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items()])
 known = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items() if cat in known_labels])
 
@@ -84,9 +85,10 @@ estimated_class_conditional_prob = ml.ClassDependentEstimator(known_samples,
 ## Plotting
 ################################################
 test_data = dataset.LabelledSample(world, 10000)
-
+styles = [ 'ro-', 'g^-', 'b*-', 'kh-', 'kh-', 'y4-']
 p_roc = lambda title, func: plot_roc(threshold = func,
                                      samples = test_data,
+                                     style = styles.pop(), 
                                      known_labels = known_labels,
                                      title = title)
 
@@ -106,11 +108,12 @@ p_roc('ROC for optimal threshold',
 #p_roc('ROC for CD(x|c)/P(x) threshold',
 #      lambda sample: (0.0001+estimated_class_conditional_prob(sample)) / perfect_unconditional_prob(sample))
 
+p_roc('ROC for P(G)/P(G\') threshold',
+      lambda sample: (0.0001+estimated_class_conditional_prob(sample)) / (0.0001+estimated_unconditional_prob(sample)))
+
 p_roc('ROC for P(G) threshold',
[…]

### 2011-04-14 semisupervised/dora-data/dora-data.py
Changes on build process and graph printing.

@@ -1,50 +1,0 @@
-
-
-# This file exports the ontology defined on:
-# http://teambox.pronobis.pro/projects/iros-paper/pages/data
-#
-import dataset
-
-labels = set('''
-             anteroom bathroom computerlab conferencehall doubleoffice
-             hallway kitchen meetingroom professorsoffice robotlab office
-             singleoffice
-             '''.split())
-
-objects = set('''
-              book cerealbox computer robot stapler toiletpaper
-              '''.split())
-
-shapes = set('''
-             elongated rectangular square
-             '''.split())
-
-appearances = set('''
-                  bathroom-like hallway-like office-like kitchen-like
-                  lab-like meetingroom-like anteroom-like
-                  '''.split())
-
-sizes = set('''
-            small medium large
-            '''.split())
-
-def Label(label):
-  assert label in labels
-  return ('label', label)
-
-def Object(obj, gamma):
-  assert 
-  return ('object_%s'%obj, dataset.DiscreteDistribution(d))
-
-def Shape(d):
-  assert
-  return ('room_size', dataset.DiscreteDistribution(d))
-
-def Appearance(d):
-  assert
-  return ('room_size', dataset.DiscreteDistribution(d))
-
-def Size(d):
-  assert Valid(d, size)
-  return ('room_size', dataset.DiscreteDistribution(d))
-

### 2011-04-14 semisupervised/dora-data/explain_ontology.py

@@ -1,0 +1,8 @@
+import script
+import explain
+from ontology import features, rooms
+
+out = explain.ExplainDistributions(sorted(features.items()),
+                                   sorted(rooms.items()))
+
+script.WriteFile(script.option('OUTPUT'), out)

### 2011-04-14 semisupervised/dora-data/ontology.py

@@ -20,20 +20,25 @@
 # This module loads the ontology and creates distributions defined by:
 # default_knowledge-semmap.xml and defaultprobs-semmap.txt
 #
+import script
 import re
 import math
 import dataset
 from collections import defaultdict
 
+POISSON_EXPANSION = 3
+DEFAULT_KNOWLEDGE = 'default_knowledge-semmap.xml'
+DEFAULT_OBJ_PROBS = 'defaultprobs-semmap.txt'
+
+
 potential = dict()
 roomcats  = set()
 features  = set()
 feature_space = defaultdict(set)
-poisson_expansion = 3
 
 def Load():
   print("Loading default knowledge...")
-  for line in open("default_knowledge-semmap.xml"):
+  for line in open(DEFAULT_KNOWLEDGE):
     match = re.match('^\s+<item\s+room_category1="(?P<roomcat>\w+)"\s+' + 
                      '(?P<proptype>\w+)="(?P<propval>\w+)"\s+' + 
                      '(?:potential|probability)="(?P<potential>\S+)"\s*\/>\s*$', line)
@@ -47,26 +52,26 @@ def Load():
       pass
     else:
       print "Ignored line: ", line,
-  
+
   
   print('Loading object probabilities...')
-  for line in open("defaultprobs-semmap.txt"):
+  for line in open(DEFAULT_OBJ_PROBS):
     match = re.match('^INROOM (?P<object>\w+) (?P<roomcat>\w+) (?P<gamma>\S+)\n$', line)
     if match:
-      obj, roomcat, gamma = match.groups()
+      obj, roomcat, p_at_least_one = match.groups()
       feature = 'object_%s' % obj
-      gamma = float(gamma)
+      gamma = - math.log(1.0 - float(p_at_least_one))
       """Unzip the poisson distribution for up to the given number of samples"""
       features.add(feature)
       t_sum = 0.0
-      for k in range(0, poisson_expansion):
+      for k in range(0, POISSON_EXPANSION):
         descriptor = str(k)
         prob = gamma**k * math.exp(-gamma) / math.factorial(k)
         potential[roomcat, feature, descriptor] = prob
         t_sum += prob
         feature_space[feature].add(descriptor)
-      potential[roomcat, feature, '%d+' % poisson_expansion] = 1.0-t_sum
-      feature_space[feature].add('%d+' % poisson_expansion)
+      potential[roomcat, feature, '%d+' % POISSON_EXPANSION] = 1.0-t_sum
[…]

### 2011-04-14 semisupervised/dora-data/run.py

@@ -17,54 +17,40 @@
 # along with this program; if not, write to the Free Software Foundation,
 # Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
 #
+import script
 import ontology
-import util
 import explain
 import random
 import dataset
 import graph
 import ml
 
-random.seed(0)
+KNOWN_LABELS       = script.option('KNOWN_LABELS').split()
+LABELLED_SAMPLES   = int(script.option('LABELLED_SAMPLES'))
+UNLABELLED_SAMPLES = int(script.option('UNLABELLED_SAMPLES'))
+TEST_SAMPLES       = int(script.option('TEST_SAMPLES'))
+PLOT         = script.option('PLOT').split()
+OUTPUT       = script.option('OUTPUT')
+
+
+features = ontology.features
+rooms = ontology.rooms
 
 #############################################################################
-def plot_roc(threshold, samples, known_labels, style, title=''):
-  def ThresholdAndLabel(sample):
-    label, sample = dataset.ExtractLabel(sample)
-    return threshold(sample), label
-
-  roc_curve = []
-  for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x, reverse=True):
-    roc_curve.append( label in known_labels )
-  graph.roc(roc_curve, style, label = title)
-  pass
 
 
-
-ontology.Load()
-del ontology.feature_space['room_category2']
-
-features = ontology.feature_space
-for f in features:
-  features[f] = sorted(features[f])
-
-rooms = ontology.MakeDistributions()
-
-util.WriteFile('explain.tex',
-    explain.ExplainDistributions(sorted(features.items()),
-                                 sorted(rooms.items())))
 
 all_labels = rooms.keys()
-known_labels = all_labels[1:]
 world = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items()])
-known = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items() if cat in known_labels])
+known = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items() if cat in KNOWN_LABELS])
 
 ################################################
 ## Generate training data
 ################################################
-known_samples = map(dataset.ExtractLabel, dataset.LabelledSample(known, 1000))
[…]

### 2011-04-14 semisupervised/dora-data/run.py
Added SCons rule to plot all roc curves for missing single room categories.

@@ -61,9 +61,13 @@ def perfect_conditional_prob(sample):
 def perfect_unconditional_prob(sample):
   return dataset.SampleProbability(world, sample)
 
-estimated_unconditional_prob = ml.IndependentFeatureEstimator(world_samples, ml.NormalizedHistogram)
-estimated_class_conditional_prob = ml.ClassDependentEstimator(known_samples,
-  lambda x: ml.IndependentFeatureEstimator(x, ml.NormalizedHistogram))
+estimated_unconditional_prob = None
+estimated_class_conditional_prob = None
+
+if 'P(G)' in PLOT or 'P(G)/P(G\')' in PLOT:
+  estimated_unconditional_prob = ml.IndependentFeatureEstimator(world_samples, ml.NormalizedHistogram)
+  estimated_class_conditional_prob = ml.ClassDependentEstimator(known_samples,
+    lambda x: ml.IndependentFeatureEstimator(x, ml.NormalizedHistogram))
 
 
 ################################################
@@ -91,12 +95,12 @@ plots = {
                     'b*-'),
   'P(x|k)/P(x)':
       lambda: p_roc(lambda sample: perfect_conditional_prob(sample)/perfect_unconditional_prob(sample),
-                    'P(x|c)/P(x)',
-                    'kh-'),
+                    'P(x|k)/P(x)',
+                    'ko-'),
   'P(x|k)':
       lambda: p_roc(lambda sample: perfect_conditional_prob(sample),
-                    'P(x|c)',
-                    'y+-')
+                    'P(x|k)',
+                    'rv-')
 }
 
 graph.newfig()

### 2011-04-14 semisupervised/notes.tex
Added stuff on experiement with dora world

@@
 room size, room shape, visual appearance and 6 different objects.
-Also, the features are statistically independent from each other given the room category.
 
-The system objective was to detect novel data from a given known set of 6 categories.
-Two tests where run:
-\begin{enumerate}
-\item Using perfect knowledge of the probability functions both thresholds where compared for performance (\autoref{fig:dora-data-perfect}).
-\item The models described in \autoref{sec:dora-models} were used to learn the probability functions and the thresholds were compared (\autoref{fig:dora-data-simple}).
-\end{enumerate}
+As expected, having perfect information of the distributions $P(x|c \in known)$ and $P(x)$
+allows us to build an optimal detector as seen on \autoref{fig:dora-data-perfect}.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.75\textwidth]{dora-data/roc-perfect.pdf}
+\caption{ROC curve for thresholds based on perfect information running on the dora distribution.
+         The convex curve for $P(x|k)/P(x)$ shows its optimal.}
+\label{fig:dora-data-perfect}
+\end{figure}
+
+Nonetheless to build that optimal detector we need information on $P(x)$ which often is not available.
+We focus then on the study of the bias introduced by using $P(x|k)$ as a threshold.
+For that we compared all roc curves based on perfect information for each missing
+classes (\autoref{fig:roc-curves-missing-1}).
+
+\begin{figure}[h]
+ \centering
+ \subfloat[Anteroom]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-anteroom.pdf}}
+ \subfloat[Bathroom]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-bathroom.pdf}}
+ \subfloat[Computer Lab]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-computerlab.pdf}} \\
+ \subfloat[Conference Hall]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-conferencehall.pdf}}
+ \subfloat[Double Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-doubleoffice.pdf}}
+ \subfloat[Hallway]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-hallway.pdf}} \\
+ \subfloat[Kitchen]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-kitchen.pdf}}
+ \subfloat[Meeting Room]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-meetingroom.pdf}}
+ \subfloat[Professors Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-professorsoffice.pdf}} \\
+ \subfloat[Robot Lab]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-robotlab.pdf}}
+ \subfloat[Single Office]{\includegraphics[width=0.33\textwidth]{dora-data/roc-missing-singleoffice.pdf}}
+
+\caption{\label{fig:roc-curves-missing-1}ROC curves for all one single missing category in the known data.}
+\end{figure}
+
+\note{TODO: Look into the distribution and explain why the harder new room categories to detect were
+      Computer Lab, Conference Hall, Meeting Room and Robot Lab.}
 
 \note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
-      to the probabilities functions to not return 0. This is because at the moment if on the known
+      so the probabilities functions do not return 0. This is because at the moment if on the known
       samples we never see a given value we don't know it belongs to the possible feature space.
@@
 
-As possible to see in \autoref{fig:dora-data-perfect} if we are able to use the unlabelled data to build a correct $P(x)$ then we obtain an optimal detector.
+\subsubsection{Approximation Models}
 
-\autoref{fig:dora-data-simple} shows how close our solution approaches the optimal by using creating models to for the conditional and unconditional probability from samples.
+In this section, I should describe how we use graphical models to model both $P(x|k)$ and $P(x)$.
+And made them as $P(G)$ and $P(G')$.
 
-Important to notice that our model of unconditional probability is very simple, it assumes semantic data to be independent from each other.
+Comparison of the estimated models with respect to the perfect information functions as been done on
+\autoref{fig:dora-data-simple}.
+Important to notice that our model of unconditional probability is very simple, still is able to improve
+our novelty detection.
 
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=\textwidth]{dora-data/roc-perfect.pdf}
-\caption{ROC curve for thresholds based on perfect information running on the dora distribution}
-\label{fig:dora-data-perfect}
-\end{figure}
-
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 \begin{center}
-\scalebox{0.35}{\input{dora-data/explain}}
+\scalebox{0.50}{\input{dora-data/explain}}
 \end{center}

### 2011-04-15 graph/ROC.py
Change the way roc curves are plotted.
Now decisions are grouped by threshold so that no longer its made out of "vertical/horizontal" lines.

@@ -8,19 +8,18 @@ import pylab as P
 def ROC(data, style = 'ro-', label = None):
   """Plots a ROC curve."""
   
-  def acc(data, value):
-    sum = 0
+  def acc(data):
+    sum = 0.0
     for v in data:
-      if v == value:
-        sum += float(1)
+      sum += v
       yield sum
   
-  fpr = N.array(list(acc(data, False)))
-  tpr = N.array(list(acc(data, True)))
+  fpr = N.array(list(acc(map(lambda x:x[0], data))))
+  tpr = N.array(list(acc(map(lambda x:x[1], data))))
   if fpr[-1] != 0: fpr /= fpr[-1]
   if tpr[-1] != 0: tpr /= tpr[-1]
   
-  P.plot(fpr, tpr, style, markevery=len(fpr)/10, label = label)
+  P.plot(fpr, tpr, style, markevery=(len(fpr)/10, len(fpr)/10), label = label)
   P.xlim([-0.05,1.05])
   P.ylim([-0.05,1.05])
   P.xlabel('False Positive Rate')

### 2011-04-15 semisupervised/dora-data/ontology.py
Changed ontology to be able to control which features are used to generate classes.

@@ -27,8 +27,9 @@ import dataset
 from collections import defaultdict
 
 POISSON_EXPANSION = 3
-DEFAULT_KNOWLEDGE = 'default_knowledge-semmap.xml'
-DEFAULT_OBJ_PROBS = 'defaultprobs-semmap.txt'
+DEFAULT_KNOWLEDGE = script.option('ONTOLOGY_DEFAULT_KNOWLEDGE')
+DEFAULT_OBJ_PROBS = script.option('ONTOLOGY_DEFAULT_OBJ_PROBS')
+DROP_FEATURES     = script.option('ONTOLOGY_DROP_FEATURES').split()
 
 
 potential = dict()
@@ -109,9 +110,12 @@ def MakeDistributions():
   room = room_d
   return room_d
   
+
 '''Generate ontology distributions.'''
 Load()
-del feature_space['room_category2']
+for feature in DROP_FEATURES:
+  del feature_space[feature]
+
 features = feature_space
 for f in features:
   features[f] = sorted(features[f])

### 2011-04-15 semisupervised/dora-data/run.py

@@ -23,6 +23,7 @@ import explain
 import random
 import dataset
 import graph
+import collections
 import ml
 
 KNOWN_LABELS       = script.option('KNOWN_LABELS').split()
@@ -77,10 +78,14 @@ def p_roc(threshold, title, style, samples = test_samples, known_labels = KNOWN_
   def ThresholdAndLabel(sample):
     label, sample = dataset.ExtractLabel(sample)
     return threshold(sample), label
-
+  
+  ct = collections.defaultdict(lambda: [0, 0])
+  for threshold, label in map(ThresholdAndLabel, samples):
+    ct[threshold][label in KNOWN_LABELS] += 1
+  
   roc_curve = []
-  for threshold, label in sorted(map(ThresholdAndLabel, samples), key = lambda x: x, reverse=True):
-    roc_curve.append( label in KNOWN_LABELS)
+  for threshold, score in sorted(ct.items(), key = lambda x: x[0], reverse=True):
+    roc_curve.append(tuple(score))
   graph.roc(roc_curve, style, label = title)
 
 

### 2011-04-15 semisupervised/notes.tex
Added notes on why some classes are harder for novelty detection.

@@
 
-\note{TODO: Look into the distribution and explain why the harder new room categories to detect were
-      Computer Lab, Conference Hall, Meeting Room and Robot Lab.}
-
 \note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
@@
       Has to be corrected.}
+
+\begin{figure}
+\missingfigure[figwidth=\textwidth]{}
+
+\caption{\label{fig:confusion-matrix-dora}Confusion matrix for the 11 classes that make up the dora distribution.}
+\end{figure}
+
+The room categories where $P(x|k)$ threshold gave worst results were:
+Computer Lab, Conference Hall, Meeting Room and Robot Lab.
+
+That was lead by the fact conditional density threshold gives very poor results on detecting classes
+that are easily confused with other classes using the given feature space (see confusion matrix in
+\autoref{fig:confusion-matrix-dora}).
+As the lack of knowledge of $P(x)$ limits the ability of the detector to correctly sort the feature
+space to in a way that a bigger value is given to the inputs where the probability of being correct is more accurate.
+
+Depending on the feature space used, this might be a very important property.
+As it will greatly impact a novelty detector. For the case of semantic data used in dora:
+very small feature spaces and lack of specific features to detect some room categories.
+
+For that we implemented a very simple approach to unconditional density probability and compared the results
+\dots
 

### 2011-04-16 graph/__init__.py
Script mode... no longer need X disp+lay.

@@ -1,3 +1,6 @@
+import matplotlib
+matplotlib.use('Agg')
+
 import pylab as pl
 from .HintonDiagrams import hinton
 from .ROC import ROC as roc

### 2011-04-26 graph/ROC.py
Added point (0,0) on ROC curve

@@ -14,8 +14,8 @@ def ROC(data, style = 'ro-', label = None):
       sum += v
       yield sum
   
-  fpr = N.array(list(acc(map(lambda x:x[0], data))))
-  tpr = N.array(list(acc(map(lambda x:x[1], data))))
+  fpr = N.array([0.0] + list(acc(map(lambda x:x[0], data))))
+  tpr = N.array([0.0] + list(acc(map(lambda x:x[1], data))))
   if fpr[-1] != 0: fpr /= fpr[-1]
   if tpr[-1] != 0: tpr /= tpr[-1]
   

### 2011-04-26 semisupervised/dora-data/classification.py
Classification confusion matrix.

@@ -1,0 +1,62 @@
+#!/usr/bin/python
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+import script
+import ontology
+import dataset
+import graph
+import collections
+import ml
+import numpy as np
+
+TEST_SAMPLES = int(script.option('TEST_SAMPLES'))
+OUTPUT       = script.option('OUTPUT')
+
+features = ontology.features
+rooms = ontology.rooms
+
+world = dataset.DiscreteDistribution([(1, dist) for cat,dist in rooms.items()])
+test_samples = dataset.LabelledSample(world, TEST_SAMPLES)
+#############################################################################
+
+def map_classifier(sample):
+  best = max([(dataset.SampleProbability(dist, sample), label) for label, dist in rooms.items()])
+  return best[1]
+
+
+confusion = collections.defaultdict(int)
+for label, sample in map(dataset.ExtractLabel, test_samples):
+  guess = map_classifier(sample)
+  confusion[label, guess] += 1
+
+
+'''Plot confusion matrix.'''
+labels = set()
+for ((label,guess), count) in confusion.items():
+  labels.add(label)
+  labels.add(guess)
+labels = list(sorted(labels))
+def correct(a, b):
+  return 1 if a == b else -1
+mat = [[confusion[labels[i],labels[j]]*correct(i, j) for j in range(len(labels))] for i in range(len(labels))]
+graph.hinton(np.array(mat), title="Confusion matrix", vlabels=labels, hlabels=labels)
+
+graph.savefig(OUTPUT)
+
+

### 2011-04-27 semisupervised/notes.tex
Some stuff

@@
 
+\clearpage
 \subsection{Dora with world model}
@@
 Nonetheless to build that optimal detector we need information on $P(x)$ which often is not available.
-We focus then on the study of the bias introduced by using $P(x|k)$ as a threshold.
+We focus then on the study of the bias introduced by only using $P(x|k)$ as a threshold.
+
 For that we compared all roc curves based on perfect information for each missing
@@
 
-\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
-      so the probabilities functions do not return 0. This is because at the moment if on the known
-      samples we never see a given value we don't know it belongs to the possible feature space.
-      Has to be corrected.}
+The room categories where $P(x|k)$ threshold gave worst results were:
+Computer Lab, Conference Hall, Meeting Room and Robot Lab.
 
+These are 
 \begin{figure}
 \missingfigure[figwidth=\textwidth]{}
-
 \caption{\label{fig:confusion-matrix-dora}Confusion matrix for the 11 classes that make up the dora distribution.}
 \end{figure}
-
-The room categories where $P(x|k)$ threshold gave worst results were:
-Computer Lab, Conference Hall, Meeting Room and Robot Lab.
 
@@
 
-\subsubsection{Approximation Models}
+\begin{sidewaystable}[h]
+\begin{center}
+\scalebox{0.50}{\input{dora-data/explain}}
+\end{center}
+\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
+\end{sidewaystable}
 
+\subsubsection{Modeling the World with Graphical Models}
+The previous experiences assume perfect knowledge on the probability distributions functions.
+This sections presents how Dora models those distributions and how to perform
+novelty detection with them.
+
+Dora uses \emph{probabilistic graphical models} to represent all extracted knowledge from the world.
+Those allow to easily model certain probabilistic properties such as conditional independence.
+And easy the process of several computational problems on the distributions.
+Several tools exists for this type of graphical manipulation such as: \emph{Bayesian networks}, \emph{Markov Random Fields}
+or \emph{Factor Graphs}.
+
+Dora underlying graphical methods are based on \emph{Factor Graphs}\citep{kschischang2001factor}, which are bipartites graph composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
+
+\begin{figure}
+Graphical model of Dora
+\end{figure}
+
+We will use $P(G)$.
+
+We will try to 
 In this section, I should describe how we use graphical models to model both $P(x|k)$ and $P(x)$.
@@
 
+\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
+      so the probabilities functions do not return 0. This is because at the moment if on the known
+      samples we never see a given value we don't know it belongs to the possible feature space.
+      Has to be corrected.}
+
 \begin{figure}[h]
@@
 
-\begin{sidewaystable}[h]
-\begin{center}
-\scalebox{0.50}{\input{dora-data/explain}}
-\end{center}
-\caption{\label{tab:dora-world-distribution}World model used in Dora. Each column cell shows $P(feature|class)$}
-\end{sidewaystable}
-
+\clearpage
 \subsection{How Dora models the world?}

### 2011-04-28 semisupervised/notes.tex
some changes

@@
 
-As expected, having perfect information of the distributions $P(x|c \in known)$ and $P(x)$
+Having perfect information of the distributions $P(x|c \in known)$ and $P(x)$
 allows us to build an optimal detector as seen on \autoref{fig:dora-data-perfect}.
@@
 
-The room categories where $P(x|k)$ threshold gave worst results were:
-Computer Lab, Conference Hall, Meeting Room and Robot Lab.
-
-These are 
 \begin{figure}
-\missingfigure[figwidth=\textwidth]{}
-\caption{\label{fig:confusion-matrix-dora}Confusion matrix for the 11 classes that make up the dora distribution.}
+\centering
+\includegraphics[width=0.75\textwidth]{dora-data/classification.pdf}
+\caption{\label{fig:confusion-matrix-dora}
+         Confusion matrix for the 11 classes that make up the dora distribution.}
 \end{figure}
@@
 
+\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
+      so the probabilities functions do not return 0. This is because at the moment if on the known
+      samples we never see a given value we don't know it belongs to the possible feature space.
+      Has to be corrected.}
+
 \clearpage

### 2011-04-28 semisupervised/notes.tex
more notes

@@
 
-\subsubsection{Modeling the World with Graphical Models}
+\clearpage
+\subsection{Modeling the World with Graphical Models}
 The previous experiences assume perfect knowledge on the probability distributions functions.
-This sections presents how Dora models those distributions and how to perform
-novelty detection with them.
+This sections presents how Dora models those distributions and how to perform novelty detection with them.
 
 Dora uses \emph{probabilistic graphical models} to represent all extracted knowledge from the world.
-Those allow to easily model certain probabilistic properties such as conditional independence.
-And easy the process of several computational problems on the distributions.
+Those allow to easily model certain probabilistic properties such as conditional independence
+and easy the process of several computational problems on the distributions.
 Several tools exists for this type of graphical manipulation such as: \emph{Bayesian networks}, \emph{Markov Random Fields}
@@
 
-Dora underlying graphical methods are based on \emph{Factor Graphs}\citep{kschischang2001factor}, which are bipartites graph composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
+Dora underlying graphical methods are based on \emph{Factor Graphs}\citep{kschischang2001factor}: bipartites graphs composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
 
-\begin{figure}
-Graphical model of Dora
+\begin{figure}[h]
+\begin{tikzpicture}
+\node[circle,fill=blue!20] (room1) {Room};
+\node[circle,fill=blue!20] (Appearance) {Apperance};
+\end{tikzpicture}
+\caption{Example of a factor graph used for Dora.}
 \end{figure}
 
-We will use $P(G)$.
+At any moment dora has a graph $G$ representing a set of random variables the system is aware of and all the probabilistic connections between them.
+The observed values are set by the sensed properties and inferences can be done on the remaining ones.
 
-We will try to 
-In this section, I should describe how we use graphical models to model both $P(x|k)$ and $P(x)$.
-And made them as $P(G)$ and $P(G')$.
+Assuming the graph edges capture the properties of the distribution $P(x|known)$.
+$G$ becomes indeed our replacement for that distribution and so $P(x|known) = P(x|G)$.
+
+In order to generate a simple and interesting properties we decided to model $P(x)$ also with a graphical model.
+
+Evaluating $P(x|G)/P(x|G')$...
+
+Method:
+\begin{enumerate}
+\item Train a graph model for unlabelled data without the labelled variables.
+\item Add the labelled variables and train the factors with the labelled data.
+\item Use $P(x|G)$ as novelty threshold.
+\end{enumerate}
 

### 2011-04-29 semisupervised/notes.tex
More changes...

@@
 
+Any probability distributuion can be represented by a \emph{factor graph} and we will denote the probability of a given set of variables outputs $x$ on a graph $G$ as $P_G(x)$.
+
+\tikzstyle{var} = [circle,
+                   thick,
+                   draw, fill=red!20,
+                   text width=4.5em,
+                   text badly centered,
+                   node distance=3cm,
+                   inner sep=0pt]
+
+\tikzstyle{factor} = [rectangle,
+                      thick,
+                      draw, fill=blue!20,
+                      text width=1em,
+                      text centered,
+                      minimum height=3em,
+                      minimum width=3em]
+
+\tikzstyle{line} = [draw]
+
 \begin{figure}[h]
-\begin{tikzpicture}
-\node[circle,fill=blue!20] (room1) {Room};
-\node[circle,fill=blue!20] (Appearance) {Apperance};
+\centering
+\begin{tikzpicture}[node distance = 2cm, auto]
+    \matrix[row sep=0.5cm,column sep=0.5cm] {
+      & \node [var] (room) {Room Category}; \\
+      \node [factor] (fAppearance) {}; &
+      \node [factor] (fShape) {}; &
+      \node [factor] (fSize) {}; \\
+
+      \node [var] (appearance) {Appearance}; &
+      \node [var] (shape) {Shape}; &
+      \node [var] (size) {Size}; \\
+    };
+  \path [line] (room) -- (fAppearance);
+  \path [line] (room) -- (fShape);
+  \path [line] (room) -- (fSize);
+
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
 \end{tikzpicture}
@@
 The observed values are set by the sensed properties and inferences can be done on the remaining ones.
+Assuming the graph $G$ captures the properties of the distribution $P(x|known)$. $G$ becomes indeed our replacement for that distribution and so $P(x|known) = P_G(x)$.
 
-Assuming the graph edges capture the properties of the distribution $P(x|known)$.
-$G$ becomes indeed our replacement for that distribution and so $P(x|known) = P(x|G)$.
+As seen before using the conditional probability as a novelty threshold doesn't gives optimal results. Therefore we try to approximate the unconditional probability to improve our threshold.
+For that we used a very simplistic model assuming independence between all variables that depend on the variable that represents the one we want to detect novelty.
+We build a graph $G'$ as an approximation for $P(x)$ with $P_{G'}(x)$.
 
-In order to generate a simple and interesting properties we decided to model $P(x)$ also with a graphical model.
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[node distance = 2cm, auto]
+    \matrix[row sep=0.5cm,column sep=0.5cm] {
+      \node [factor] (fAppearance) {}; &
+      \node [factor] (fShape) {}; &
+      \node [factor] (fSize) {}; \\
 
-Evaluating $P(x|G)/P(x|G')$...
+      \node [var] (appearance) {Appearance}; &
+      \node [var] (shape) {Shape}; &
+      \node [var] (size) {Size}; \\
+    };
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
+\end{tikzpicture}
+\caption{Factor graph to model unconditional probability. It assumes independency between\dots}
+\end{figure}
 
-Method:
-\begin{enumerate}
-\item Train a graph model for unlabelled data without the labelled variables.
-\item Add the labelled variables and train the factors with the labelled data.
-\item Use $P(x|G)$ as novelty threshold.
-\end{enumerate}
+Using the described $G$ and $G'$ we evaluated our novelty detectors in the distribution used before and compared with an optimal detector.
+As possible to see in \autoref{fig:dora-data-simple} our approaximation to $P(x)$ althouth very basic was able to improve a novelty detection based only on conditional probability.
 
-Comparison of the estimated models with respect to the perfect information functions as been done on
-\autoref{fig:dora-data-simple}.
-Important to notice that our model of unconditional probability is very simple, still is able to improve
-our novelty detection.
-
-\note{Describe better which approach we used to model $P(G)/P(G')$}
-\note{Perform experience with real data}
+In order to generate a simple model we decided to model the unconditional density probability function also with a graphical model $G'$ and so we approximate $P(x)$ as $P_{G'}(x)$.
 
@@
 
-\note{The graphics are a bit "cheated", since theres an added epsilon on the returned value
-      so the probabilities functions do not return 0. This is because at the moment if on the known
-      samples we never see a given value we don't know it belongs to the possible feature space.
-      Has to be corrected.}
+Evaluating $P(x|G)/P(x|G')$...
+
+Method:
+\begin{enumerate}
+\item Train a graph model for unlabelled data without the labelled variables.
+\item Add the labelled variables and train the factors with the labelled data.
+\item Use $P_G(x)$ as novelty threshold.
+\end{enumerate}
+
+\note{Perform experience with real data}
 

### 2011-05-03 semisupervised/notes.tex
more changes

@@
 or \emph{Factor Graphs}.
+Those tools are especially useful since they are generative models and allow to compute the probability of any given set of variables $x$.
 
-Dora underlying graphical methods are based on \emph{Factor Graphs}\citep{kschischang2001factor}: bipartites graphs composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
-
-Any probability distributuion can be represented by a \emph{factor graph} and we will denote the probability of a given set of variables outputs $x$ on a graph $G$ as $P_G(x)$.
+Dora underlying graphical methods are based on \emph{Chain Graphs} that can be translated into \emph{Factor Graphs}\citep{kschischang2001factor}: bipartite graphs composed of variable and factor nodes representing the random variables and several factor functions between the variables that are used to decomposed the distribution.
+Any probability distribution can be represented by a \emph{factor graph} and we will denote the probability of a given set of variables outputs $x$ on a graph $G$ as $P_G(x)$.
 
@@
 \end{tikzpicture}
-\caption{Example of a factor graph used for Dora.}
+\caption{Example of a factor graph $G$ used for Dora for modelling the labelled samples.}
 \end{figure}
@@
 \end{tikzpicture}
-\caption{Factor graph to model unconditional probability. It assumes independency between\dots}
+\caption{A sample factor graph to model the unconditional probability. It assumes independent properties. Factors are trained giving access to the whole set of unlabelled data.}
 \end{figure}
@@
 Using the described $G$ and $G'$ we evaluated our novelty detectors in the distribution used before and compared with an optimal detector.
-As possible to see in \autoref{fig:dora-data-simple} our approaximation to $P(x)$ althouth very basic was able to improve a novelty detection based only on conditional probability.
+As possible to see in \autoref{fig:dora-data-simple} our approximation $P_{G'}(x)$ to $P(x)$ although very basic was able to improve a novelty detection.
 
-In order to generate a simple model we decided to model the unconditional density probability function also with a graphical model $G'$ and so we approximate $P(x)$ as $P_{G'}(x)$.
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.75\textwidth]{dora-data/roc-simple.pdf}
+\caption{ROC curve for thresholds learned on samples from samples in the dora distribution.
+As possible to see even a simple approach to model $P(x)$ was able to considerably improve our novelty detector.}
+\label{fig:dora-data-simple}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[width=\textwidth]{dora-data/roc-simple.pdf}
-\caption{ROC curve for thresholds learned on samples from samples in the dora distribution}
-\label{fig:dora-data-simple}
-\end{figure}
-
-Evaluating $P(x|G)/P(x|G')$...
-
-Method:
-\begin{enumerate}
-\item Train a graph model for unlabelled data without the labelled variables.
-\item Add the labelled variables and train the factors with the labelled data.
-\item Use $P_G(x)$ as novelty threshold.
-\end{enumerate}
-
 \note{Perform experience with real data}
-
-\clearpage
-\subsection{How Dora models the world?}
-\label{sec:dora-models}
-
-Currently dora uses a \emph{probabilistic graphical model} to model the world.
-All variables that the system see are added and connected in between with the
-\emph{probabilistic factors} that the system is aware of.
-Those factors are learned from the labelled data that the system is given when
-training.
-
-The proposal here is to also use such an approach to model the unconditional
-probability.
-Under that a novelty detection threshold can be easily implemented
-reusing the \emph{probabilistic graphical model} $G$ that dora uses as
-representation of the world:
-\begin{description}
-\item[$P(x|know)$] is obtained evaluating $P(G)$.
-\item[$P(x)$] can be obtained by creating $G'$ as an instance of $G$ but with the
-variable node we want to perform novelty detection removed and the factors associated
-with it replaced with factors learned from the unlabelled data.
-\end{description}
-
-We point to an additional simplification that if the original graphical model factors
-depending on the variable to perform novelty detection were previously split on a
-conditional and unconditional factors. Then $G'$ could be obtained by simply removing
-all the factors associated with the variable to perform novelty detection on.
 

### 2011-05-03 semantic-nd/article.tex
Some ideas to make sort of a paper.

\documentclass[runningheads,a4paper]{llncs}

\usepackage{amssymb}
\setcounter{tocdepth}{3}
\usepackage{graphicx}

\usepackage{url}
\newcommand{\keywords}[1]{\par\addvspace\baselineskip
\noindent\keywordname\enspace\ignorespaces#1}

% Additional packages
\usepackage[utf8]{inputenc}

\begin{document}

\mainmatter  % start of an individual contribution

% first the title is needed
\title{Semantic Novelty Detection with Graphical Models
on Dora Architecture}

% a short form should be given in case it is too long for the running head
% \titlerunning{Lecture Notes in Computer Science: Authors' Instructions}

% the name(s) of the author(s) follow(s) next
%
% NB: Chinese authors should write their first names(s) in front of
% their surnames. This ensures that the names appear correctly in
% the running heads and the author index.
%
\author{André Susano Pinto}
%
%\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
% (feature abused for this document to repeat the title also on left hand pages)

% the affiliations are given next; don't give your e-mail address
% unless you accept that it will be published
\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
\url{andresusanopinto@gmail.com}\\
\url{http://url}}

\toctitle{Lecture Notes in Computer Science}
\tocauthor{Authors' Instructions}
\maketitle

\begin{abstract}
This papers presents an approach on implementing novelty detection at semantic level
for indoor room classification in Dora. \emph{Graphical models} are used to model
probabilistic knowledge and novelty detection is implemented on top of them.
The novelty threshold is then optimized using an unconditional probability density
function that is built from an unlabelled database.

\keywords{novelty detection, semantic data, graphical models, room classification,
indoor environments, robotics.}
\end{abstract}

\section{Introduction}

\begin{thebibliography}{4}

\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
Subsequences. J. Mol. Biol. 147, 195--197 (1981)

\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
Composing a Complex Biological Workflow through Web Services. In: Nagel,
W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
pp. 1148--1158. Springer, Heidelberg (2006)

\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
Infrastructure. Morgan Kaufmann, San Francisco (1999)

\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
Information Services for Distributed Resource Sharing. In: 10th IEEE
International Symposium on High Performance Distributed Computing, pp.
181--184. IEEE Press, New York (2001)

\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
Grid: an Open Grid Services Architecture for Distributed Systems
Integration. Technical report, Global Grid Forum (2002)

\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\end{document}

### 2011-05-04 semantic-nd/article.tex
More ideas organization

@@
 
+\subsection{Graphical Models}
+
+\section{Novelty Detection Threshold}
+Novelty detection deals with the detection of data comming from a new class,
+that the system is not aware during training.
+
+It is often implemented around a.
+
+When the data is all 
+In such cases the detection threshold can be seen as a function responsible to order
+the input space in such a way that .
+
+When dealing with exact
+
+\subsection{Conditional Density Probability}
+
+\subsection{Optimal Novelty Detector using Unconditional Probability}
+
 \begin{thebibliography}{4}

### 2011-05-04 semantic-nd/article.tex
Article ideas and organization.

--- 2011-05-03 semantic-nd/article.tex
@@
 \usepackage[utf8]{inputenc}
+\usepackage{tikz}
+
+% TiKZ styles
+\tikzstyle{var} = [circle,
+                   thick,
+                   draw, fill=red!20,
+                   text width=4.5em,
+                   text badly centered,
+                   node distance=3cm,
+                   inner sep=0pt]
+
+\tikzstyle{factor} = [rectangle,
+                      thick,
+                      draw, fill=blue!20,
+                      text width=1em,
+                      text centered,
+                      minimum height=3em,
+                      minimum width=3em]
+
+\tikzstyle{line} = [draw]
 
@@
 The novelty threshold is then optimized using an unconditional probability density
-function that is built from an unlabelled database.
+model that is trained from unlabelled data.
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
+\begin{itemize}
+\item Explain context of indoor room categorization and importance of novelty detection.
+\end{itemize}
 
+There has been several efforts in the area of Artificial Intelligence and Robotics in creating robots
+that are able to interact with humans and their environments.
+One of the existing problems is a reliable high-level localization method that can be deployed into new
+and unknown environments.
+In this article we focus on the mapping, using \emph{semantic data}, of indoor environments such as houses and offices to room categories
+such as kitchen, corridor, office.
+And how to perform novelty detection on them: \emph{detect a new room category that was not present in the labelled data}.
+
+Dora\cite{dora} (CogX: Dora) was used as a base system where this novelty detection would be implemented.
+As Dora moves through the environment its \emph{conceptual layer} builds a structural and probabilistic representation of the space
+instantiated as a \emph{graphical model}.
+That model connects the sensed properties together with the variables used to model the world.
+And allows to perform queries on the probabilities of aspects of the environment.
+For example: where is the robot most likely to find a cereal box\cite{exploiting}.
+
+\subsection{Semantic Data}
+\begin{itemize}
+\item Explain what we mean by semantic data and how its captured by Dora
+\end{itemize}
+
+Dora has a set of machine learning tools that are used to sense the environment around it.
+The output of those describe a certain high-level property or information that we refer as \emph{semantic data}.
+The following types of probabilistic information are extracted by Dora and used on this paper:
+\begin{description}
+\item[Object detection] is performed by running several pre-trained object detectors on the visual input. The system becomes
+aware of the presence of objects such as book, cereal box, computer, robot, stapler, toilet paper.
+
+\item[Room Size and Shape] are extracted by using the 2D laser scans. The system categorizes each room in terms of its size (large, medium, small) and shape (rectangular and elongated).
+
+\item[Room Appearance] is categorized using CRFH and a pre-trained set of 7 different models.
+
+\item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
+\end{description}
+
+\subsection{Graphical Models}
+\begin{itemize}
+\item Explain Dora use of Graphical Models
+\end{itemize}
+
+Dora uses \emph{chain graphs} as modelling tool. Those can be converted to \emph{factor graphs}\cite{factor}.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[node distance = 2cm, auto]
+    \matrix[row sep=0.5cm,column sep=0.5cm] {
+      & \node [var] (room) {Room Category}; \\
+      \node [factor] (fAppearance) {}; &
+      \node [factor] (fShape) {}; &
+      \node [factor] (fSize) {}; \\
+
+      \node [var] (appearance) {Appearance}; &
+      \node [var] (shape) {Shape}; &
+      \node [var] (size) {Size}; \\
+    };
+  \path [line] (room) -- (fAppearance);
+  \path [line] (room) -- (fShape);
+  \path [line] (room) -- (fSize);
+
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
+\end{tikzpicture}
+\caption{Example of a factor graph used for Dora for representing the environment.}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Novelty Detection}
+\begin{itemize}
+\item Explain 
+\item Approach with a threshold function
+\item Need of a normalizing factor for a threshold with dynamic functions
+\item Usage of conditional density
+\item Assuming of every output is equally likely
+\item Using unlabelled data
+\end{itemize}
+
+Novelty detection can be performed by a threshold function sorting the input space in an optimal way to
+decrease error in detection.
+
+Its harder than a classification problem as the system only has access to positive samples.
+Under that a common approach is to use conditional probability $P(x|known)$ as a threshold.
+
+\subsection{Stable threshold}
+Although in our case we are dealing with a dynamic graph $G$ that keeps changing through time as new variables
+are added. Which increases the dimension of our space and spreads the distribution over the variables.
+
+This turns it impossible to directly train a threshold function on $P_G(x)$, as we cannot assume $P(x)$ to be constant over any set of $x$.
+So we need to describe $P(x)$ to develop a stable threshold on $P_G(x)/P(x)$.
+
+A common assumption is that every possible input is equally likely then the threshold
+should be $P_G(x)/(\prod |x_i|)$.
+
+\subsection{Unlabelled data}
+Nonetheless in the case of robotics its often the case we can access large amounts of unlabelled data.
+This leads to an interesting point that instead of assuming a constant $P(x)$ we can actually try to build
+model for it.
+
+Under that we decided to test assuming that all probabilities are independent. That can be modelled as a graph $G'$.
+And corresponding function $P_G(x)/P_{G'}(x)$.
+
+This way our unconditional function perceive which features are slightly more biased to a certain value and will
+compensate on the probability calculation.
+
+This step is important to achieve a correct order of the input space.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+\begin{itemize}
+\item Show results and how much it improves by using a simple method for unconditional probability
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions and Future Work}
+\begin{itemize}
+\item Show a correct estimation of unconditional probability to be important in estabilishing a novelty detection threshold
+\item Show unlabelled data can be used for it.
+\item Although focus on Room Categorization we expect the results and techniques to be usable on other areas.
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{thebibliography}{4}

### 2011-05-04 semantic-nd/article.tex
some changes

@@
 \keywords{novelty detection, semantic data, graphical models, room classification,
-indoor environments, robotics.}
+indoor environments, robotics, multi-modal.}
 \end{abstract}
@@
 \item Explain context of indoor room categorization and importance of novelty detection.
+\item Introduce Dora and explain how it would benefit or use the semantic novel detection.
 \end{itemize}
@@
 And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is the robot most likely to find a cereal box\cite{exploiting}.
+For example: where is most likely to find a cereal box\cite{exploiting}.
+
+The ability to recognize novel categories on the modelled variables would allow to increase
+reliability and allow the creation of self-extending behaviours.
+
+\section{Related Work}
+Although this paper deals with novelty detection on the robotics area, it does so using very
+standard concepts and techniques such as semantic data and graphical models.
+Those are often found on areas related with information retrieval.
+
+Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{multi-model-image-retrieval}.
+
+\section{Dora Architecture Overview}
+\begin{itemize}
+\item Introduce Dora architecture aspects important for the paper
+\item Explain what we mean by semantic data and how its captured by Dora
+\item Introduce graphical models and how we use them (including notation on the paper)
+\end{itemize}
+
+Dora is composed by several layers, though for the sake of this paper only the conceptual layer is important.
 
 \subsection{Semantic Data}
-\begin{itemize}
-\item Explain what we mean by semantic data and how its captured by Dora
-\end{itemize}
-
 Dora has a set of machine learning tools that are used to sense the environment around it.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The edges and factors between the several variables can be seen as a kind of filter to
+the properties of the system.
+
+Graphical models provide a generative model for a distributions of variables.
+Those are mainly interesting since they allows to calculate the probability of any set of variables $x$
+
+As they can be seen as a method to represent distributions of variables.
+From now on we shall use the notation $P_G(x)$ to denote the probability that the distribution described by a graphical model $G$ generates the set $x$ of variables.
+
 \section{Novelty Detection}
@@
 \item Using unlabelled data
+\item What happens if the conditional and unconditional graphs differ only by the addition of a node?
+\item How to implement that in the case a graph is a tree
+\item Novelty detection on any variable of the graphical model!
 \end{itemize}

### 2011-05-05 semantic-nd/article.tex
advances on explaining dora and graphs

@@
 \usepackage[utf8]{inputenc}
+\usepackage[round]{natbib}
 \usepackage{tikz}
+\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
+\hypersetup{%
+   plainpages=false, 
+   pdfpagelayout=SinglePage,
+   bookmarksopen=false,
+   bookmarksnumbered=true,
+   breaklinks=true,
+   linktocpage,
+   colorlinks=true,
+   linkcolor=blue,
+   urlcolor=blue,
+   citecolor=blue,
+   anchorcolor=green
+}
 
@@
 % first the title is needed
-\title{Semantic Novelty Detection with Graphical Models
-on Dora Architecture}
+\title{Semantic Novelty Detection with Graphical Models}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related Work}
+\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor
+scenes when compared to outdoor scenes results.
+Since the properties that characterize rooms changes conforming its category. Namely corridors are
+well described by global properties and bookstores are well described by the presence of specific objects (books).
+Their work shows a multi-modal approach is expected to yield better performance by mixing several sources of information.
+
+\cite{galindo2005multi} defines a bidirectional relation between object and room category, where object defines a room category and a room category provides information on where objects may be found.
+
+Probabilistic representations are used in several localised functions in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
+And some employ, up to some extent, a probabilistic representation across some subsystems~\cite{kraft2008exploration}.
+\cite{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
+And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.
+
+Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.
+
 Although this paper deals with novelty detection on the robotics area, it does so using very
@@
 Those are often found on areas related with information retrieval.
+Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{zhang2005probabilistic}.
 
-Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{multi-model-image-retrieval}.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Dora Architecture Overview}
@@
 
-Dora is composed by several layers, though for the sake of this paper only the conceptual layer is important.
+Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
 
-\subsection{Semantic Data}
-Dora has a set of machine learning tools that are used to sense the environment around it.
-The output of those describe a certain high-level property or information that we refer as \emph{semantic data}.
-The following types of probabilistic information are extracted by Dora and used on this paper:
+From its architecture only the \emph{conceptual layer} is of interest to this article.
+Its role is to aggregate the following semantic information coming from other layers:
+
 \begin{description}
-\item[Object detection] is performed by running several pre-trained object detectors on the visual input. The system becomes
-aware of the presence of objects such as book, cereal box, computer, robot, stapler, toilet paper.
-
-\item[Room Size and Shape] are extracted by using the 2D laser scans. The system categorizes each room in terms of its size (large, medium, small) and shape (rectangular and elongated).
-
-\item[Room Appearance] is categorized using CRFH and a pre-trained set of 7 different models.
-
  \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
+ \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room.
+The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
+ \item[Object detection] is performed using the visual input in order to detect .
+The system keeps track of the number of object and types seen per each room.
+Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
+ \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
 \end{description}
 
-\subsection{Graphical Models}
-\begin{itemize}
-\item Explain Dora use of Graphical Models
-\end{itemize}
+With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
+in order to model all known variables and their relations.
+As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
+a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.
 
-Dora uses \emph{chain graphs} as modelling tool. Those can be converted to \emph{factor graphs}\cite{factor}.
+The use of graphical models to describe distributions of variables has useful properties.
+The edges between the variables can be seen as a kind of filter to the properties of the system.
+Properties those that we expect to capture by the conceptual knowledge.
+At the same time they are a generative models and therefore allow to calculate the probability
+on any given subset of variables on the graph allowing the system to work even when some information is missing.
+
+\begin{figure}[h]
+\centering
+\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge in order to
+         create a structured probabilistic representation of the world.}
+\end{figure}
+
+\subsection{Factor Graphs}
+Although the conceptual layer works with \emph{chain graph}, those can be converted into \emph{factor graphs}.
+We opted to use those as describing the ideas on this paper as they tend to be easier to understand.
+
+A \emph{factor graph} is a bipartite graph connecting nodes representing random variables and factors.
+Each factor represents a function dependent only on the variables to where its connected.
+As so given factor graph $G$ can be seen as description of a distribution function over a set of variables obtained by
+the product of all the factors. This manipulation allows to use algorithms such as the Sum-Product to efficiently calculate
+probabilities on any given subset $x$ of those variables by exploiting conditional independency between variables.
+We will use the notation $P_G(x)$ to represent that.
 
@@
 
-The edges and factors between the several variables can be seen as a kind of filter to
-the properties of the system.
-
-Graphical models provide a generative model for a distributions of variables.
-Those are mainly interesting since they allows to calculate the probability of any set of variables $x$
-
-As they can be seen as a method to represent distributions of variables.
-From now on we shall use the notation $P_G(x)$ to denote the probability that the distribution described by a graphical model $G$ generates the set $x$ of variables.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Novelty Detection}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{thebibliography}{4}
-
-\bibitem{jour} Smith, T.F., Waterman, M.S.: Identification of Common Molecular
-Subsequences. J. Mol. Biol. 147, 195--197 (1981)
-
-\bibitem{lncschap} May, P., Ehrlich, H.C., Steinke, T.: ZIB Structure Prediction Pipeline:
-Composing a Complex Biological Workflow through Web Services. In: Nagel,
-W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006. LNCS, vol. 4128,
-pp. 1148--1158. Springer, Heidelberg (2006)
-
-\bibitem{book} Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing
-Infrastructure. Morgan Kaufmann, San Francisco (1999)
-
-\bibitem{proceeding1} Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid
-Information Services for Distributed Resource Sharing. In: 10th IEEE
-International Symposium on High Performance Distributed Computing, pp.
-181--184. IEEE Press, New York (2001)
-
-\bibitem{proceeding2} Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The Physiology of the
-Grid: an Open Grid Services Architecture for Distributed Systems
-Integration. Technical report, Global Grid Forum (2002)
-
-\bibitem{url} National Center for Biotechnology Information, \url{http://www.ncbi.nlm.nih.gov}
-
-\end{thebibliography}
+\bibliographystyle{plainnat}
+\bibliography{refs}
 

### 2011-05-05 semantic-nd/article.tex
Several changes in both dora description and novelty detection.

@@
 \centering
+
+\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}
+\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}
 \caption{The conceptual layer structures the sensed environment together with the conceptual knowledge in order to
@@
 \begin{itemize}
-\item Explain 
-\item Approach with a threshold function
+\item Define novelty detection 
+\item Explain approach with a threshold function
 \item Need of a normalizing factor for a threshold with dynamic functions
@@
 
-Novelty detection can be performed by a threshold function sorting the input space in an optimal way to
-decrease error in detection.
+% Introduce novelty detection
+Novelty detection is an harder problem than that of classification due to its single-class nature.
+Essentially a system is given training data and when given new data has to decide if it belongs
+to the same concept it was trained with.
 
-Its harder than a classification problem as the system only has access to positive samples.
-Under that a common approach is to use conditional probability $P(x|known)$ as a threshold.
+% Explain threshold approach to novelty detection
+Due to the desired of robustness and generalization, a novelty detection system would always possess
+some threshold that describes on how strict the system should be to new data.
+This threshold is often seen as a distance measure to a normal input and several novelty detection
+techniques are based on it.
 
-\subsection{Stable threshold}
-Although in our case we are dealing with a dynamic graph $G$ that keeps changing through time as new variables
-are added. Which increases the dimension of our space and spreads the distribution over the variables.
+When dealing with a statistic point of view, noisy data or unstable features, a decision on a given input has
+an associated error rate. And in this case a novelty detection system is interested in defining an order
+on all the possible inputs equivalent to the order imposed by the error rate: $P(novel|x)$.
 
-This turns it impossible to directly train a threshold function on $P_G(x)$, as we cannot assume $P(x)$ to be constant over any set of $x$.
-So we need to describe $P(x)$ to develop a stable threshold on $P_G(x)/P(x)$.
+Such an ordering can be defined by a R-value function on the input space. And then a threshold can be trained
+on that function in order to accept an expected amount of cases at the cost of failing in others.
 
-A common assumption is that every possible input is equally likely then the threshold
-should be $P_G(x)/(\prod |x_i|)$.
+Based on bayes rule:
 
-\subsection{Unlabelled data}
-Nonetheless in the case of robotics its often the case we can access large amounts of unlabelled data.
-This leads to an interesting point that instead of assuming a constant $P(x)$ we can actually try to build
-model for it.
+\begin{equation}
+P(\overline{novel}|x) = \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
+\end{equation}
 
-Under that we decided to test assuming that all probabilities are independent. That can be modelled as a graph $G'$.
-And corresponding function $P_G(x)/P_{G'}(x)$.
+% Explain approach of using conditional probability
+\subsection{Conditional Probability: $P(x|\overline{novel})$}
+One of the hardest problems is that on most cases novelty detection system only have access to 
+the known data.
+Under that the best approach is to define a threshold based on $P(x|\overline{novel})/P(x)$
+considering that $P(x)$ is constant through all the input space.
 
+In the presented case, $P(x|\overline{novel})$ is approximated by the graphical model $G$ produced
+by the conceptual layer. As so we use $P_G(x)$ as a substitute.
+
+\label{sec:uniform-unconditional}
+In several cases assuming the unconditional probability $P(x)$ is constant leads to discarding the factor,
+nonetheless here the distributions are dynamically changing as the robot learns more on the environment.
+So we need to keep the normalizing argument $P(x)$ and re-calculate it for each new subset of $x$.
+Still assuming that the unconditional distributions generates all possible outcome with the same probability
+we can model it $P(x)=\prod{|x_i|}$, where $|x_i|$ denotes the state space size of variable $x_i$.
+In graphical model terms, this would be equivalent to build a factor graph $G'$ with the same variables but
+without any factors.
+
+Using a graphical model built to model the known data and assuming that in the unconditional case all possible outcomes
+are equally likely a novelty threshold would be given by: $P_G(x)/P_{G'}(x)$, as illustrated on \autoref{fig:unsupervised-threshold}.
+
+\begin{figure}[h]
+\caption{\label{fig:unsupervised-threshold}By producing graphical models $G$ and $G'$ to both conditional and unconditional probability
+         distributions on a set of variables, novelty can be detect on the threshold $P_G(x)/P_{G'}(x)$.
+         In this case the unconditional probability is modelled as a fully uniform distribution.}
+\end{figure}
+
+\subsection{Using Unlabelled Data}
+Nonetheless in several cases its often the case that there is access to large amounts of unlabelled data.
+Under that it becomes possible to obtain a better approach to the unconditional probability distribution that
+the uniform one assumed in \autoref{sec:uniform-unconditional}.
+
+For interesting and practical reasons it was assumed that the unconditional distribution could be modelled with
+all the variables independent from each other.
+Which is the same as assuming $P(x)=\prod{P_{x_i}(x_i)}$.
+And so from the unlabelled data we only need to estimate $P_{x_i}$.
+Once again \autoref{fig:semisupervised-threshold} illustrates this in terms of graphs $G$ and $G'$.
+
+\begin{figure}[h]
+\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to create a better approximation
+         for the unconditional probability. Which should lead to a better novelty threshold.}
+\end{figure}
+
+The described ratio between the two graphs can be understood as compensating for an existing bias
+on the unconditional distribution.
+Which is an important step to achieve a correct ordering of the input space for novelty thresholding.
 This way our unconditional function perceive which features are slightly more biased to a certain value and will
@@
 
-This step is important to achieve a correct order of the input space.
+\subsection{Generalizing to Novelty Detection on a variable}
+Until now, the presented methods describe how to detect data that belongs to a new class of input by creating
+generative models for both conditional and unconditional probabilities distributions.
+
+And although it serves as a method for detecting new classes in the input, it gives us no control in what
+variables novelty is being measured.
+
+Lets assume that the only difference between the conditional and unconditional graph is a single node and its connecting factors
+as seen on figure X.
 

### 2011-05-05 semantic-nd/article.tex
small changes

@@
 variables novelty is being measured.
+This is the novelty can come from any variable present on the graphs and we have no way to measure it.
 
 Lets assume that the only difference between the conditional and unconditional graph is a single node and its connecting factors
-as seen on figure X.
+as seen on \autoref{fig:measure-room-novelty}.
+
+\begin{figure}
+\caption{\label{fig:measure-room-novelty}In this case the only difference between $G$ and $G'$ is the room category variable,
+         that only exists in the later.}
+\end{figure}
+
+From now on we refer to the measure $P_G(x)/P_{G'}(x)$, when $G$ and $G'$ as a novelty measure on the variable $R_{category}$.
+As it attempts to measure on how much adding the variable $R_{category}$ helps to explain the given distribution of values.
+
+Expanding the calculation of $P_G(x)/P_{G'}(x)$ with the sum-product technique. Leads to:
 

### 2011-05-05 semantic-nd/article.tex
Added several tikz graphs explaining distributions.
Though for lack of time from now on other drawing methods will be used
and the tikz images will be developed later.

@@
 \usepackage[round]{natbib}
+\usepackage{subfig}
 \usepackage{tikz}
@@
                    draw, fill=red!20,
-                   text width=4.5em,
-                   text badly centered,
-                   node distance=3cm,
+                   font=\scriptsize,
+                   text width=3.5em,
+                   text centered,
+                   node distance=1em,
                    inner sep=0pt]
@@
                       draw, fill=blue!20,
-                      text width=1em,
-                      text centered,
-                      minimum height=3em,
-                      minimum width=3em]
+                      minimum height=2em,
+                      minimum width=2em]
 
@@
 \begin{figure}[h]
-\caption{\label{fig:unsupervised-threshold}By producing graphical models $G$ and $G'$ to both conditional and unconditional probability
-         distributions on a set of variables, novelty can be detect on the threshold $P_G(x)/P_{G'}(x)$.
-         In this case the unconditional probability is modelled as a fully uniform distribution.}
+\centering
+\subfloat[Factor graph $G$ used to model conditional probability]{
+\begin{tikzpicture}
+  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
+    \& \node [var] (room) {Room Category}; \& \\
+
+    \node [factor] (fAppearance) {}; \&
+    \node [factor] (fShape) {}; \&
+    \node [factor] (fSize) {}; \\
+
+    \node [var] (appearance) {Appearance}; \&
+    \node [var] (shape) {Shape}; \&
+    \node [var] (size) {Size}; \\
+  };
+  \path [line] (room) -- (fAppearance);
+  \path [line] (room) -- (fShape);
+  \path [line] (room) -- (fSize);
+
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
+\end{tikzpicture}}
+\hfill
+\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
+\begin{tikzpicture}
+  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
+    \& \node [var] (room) {Room Category}; \& \\
+
+    \node [var] (appearance) {Appearance}; \&
+    \node [var] (shape) {Shape}; \&
+    \node [var] (size) {Size}; \\
+  };
+\end{tikzpicture}}
+
+\caption{\label{fig:unsupervised-threshold}
+         Assuming a uniform unconditional distribution of any subset of the variables $x$,
+         a novelty threshold can be defined on the function $P_G(x)/P_{G'}(x)$.}
 \end{figure}
@@
 \begin{figure}[h]
+\centering
+\subfloat[Factor graph $G$ used to model conditional probability]{
+\begin{tikzpicture}
+  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
+    \& \node [var] (room) {Room Category}; \& \\
+
+    \node [factor] (fAppearance) {}; \&
+    \node [factor] (fShape) {}; \&
+    \node [factor] (fSize) {}; \\
+
+    \node [var] (appearance) {Appearance}; \&
+    \node [var] (shape) {Shape}; \&
+    \node [var] (size) {Size}; \\
+  };
+  \path [line] (room) -- (fAppearance);
+  \path [line] (room) -- (fShape);
+  \path [line] (room) -- (fSize);
+
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
+\end{tikzpicture}}
+\hfill
+\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
+\begin{tikzpicture}
+  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
+    \node [var] (appearance) {Appearance}; \&
+    \node [var] (shape) {Shape}; \&
+    \node [var] (size) {Size}; \\
+
+    \node [factor] (fAppearance) {}; \&
+    \node [factor] (fShape) {}; \&
+    \node [factor] (fSize) {}; \\
+  };
+  \path [line] (fAppearance) -- (appearance);
+  \path [line] (fShape) -- (shape);
+  \path [line] (fSize) -- (size);
+\end{tikzpicture}}
+
 \caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to create a better approximation

### 2011-05-05 semantic-nd/article.tex
Added drawings and refactor novelty part.

@@
 \subsection{Conditional Probability: $P(x|\overline{novel})$}
-One of the hardest problems is that on most cases novelty detection system only have access to 
+\label{sec:uniform-unconditional}
+
+One of the hardest problems is that on most cases a novelty detection system only have access to 
 the known data.
@@
 
-In the presented case, $P(x|\overline{novel})$ is approximated by the graphical model $G$ produced
-by the conceptual layer. As so we use $P_G(x)$ as a substitute.
-
-\label{sec:uniform-unconditional}
-In several cases assuming the unconditional probability $P(x)$ is constant leads to discarding the factor,
-nonetheless here the distributions are dynamically changing as the robot learns more on the environment.
-So we need to keep the normalizing argument $P(x)$ and re-calculate it for each new subset of $x$.
-Still assuming that the unconditional distributions generates all possible outcome with the same probability
-we can model it $P(x)=\prod{|x_i|}$, where $|x_i|$ denotes the state space size of variable $x_i$.
-In graphical model terms, this would be equivalent to build a factor graph $G'$ with the same variables but
-without any factors.
-
-Using a graphical model built to model the known data and assuming that in the unconditional case all possible outcomes
-are equally likely a novelty threshold would be given by: $P_G(x)/P_{G'}(x)$, as illustrated on \autoref{fig:unsupervised-threshold}.
+In the presented case the conditional probability is approximated by the graphical model $G$
+produced by the conceptual layer.
+The \autoref{fig:conditional-prob-graph} represents a possible graph $G$ built from the conceptual
+layer to represent the conditional probability on the sensed variables $x$.
+$P_G(x)$ is then used as an approximation for $P(x|\overline{novel})$.
 
@@
 \centering
-\subfloat[Factor graph $G$ used to model conditional probability]{
-\begin{tikzpicture}
-  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
-    \& \node [var] (room) {Room Category}; \& \\
+\includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
+\caption{\label{fig:conditional-prob-graph}The conceptual layer models the conditional
+         probability distribution of the sensed variables by using hidden variables.
+         Those are created from the conceptual knowledge and represent the structural
+         and connectivity information the system is aware of.
+         In this case only hidden variables $R_i$ were used to model the room categories
+         where each property was sensed.}
+\end{figure}
 
-    \node [factor] (fAppearance) {}; \&
-    \node [factor] (fShape) {}; \&
-    \node [factor] (fSize) {}; \\
+Its important to notice that in several cases assuming the unconditional probability $P(x)$ is constant leads to discarding the factor.
+Nonetheless, here the distributions are dynamically changing as the system learns more on the environment.
+So we need to keep the normalizing argument $P(x)$ and re-calculate it for each new subset of $x$.
 
-    \node [var] (appearance) {Appearance}; \&
-    \node [var] (shape) {Shape}; \&
-    \node [var] (size) {Size}; \\
-  };
-  \path [line] (room) -- (fAppearance);
-  \path [line] (room) -- (fShape);
-  \path [line] (room) -- (fSize);
+Assuming that the unconditional distributions generates all possible outcome with the same probability
+we can model it with $\prod{1/|x_i|}$, where $|x_i|$ denotes the state space size of variable $x_i$.
+In graphical model terms, this would be equivalent to build a factor graph $U$ with the same variables but
+without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
 
-  \path [line] (fAppearance) -- (appearance);
-  \path [line] (fShape) -- (shape);
-  \path [line] (fSize) -- (size);
-\end{tikzpicture}}
-\hfill
-\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
-\begin{tikzpicture}
-  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
-    \& \node [var] (room) {Room Category}; \& \\
+\begin{figure}
+\centering
+\includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
+\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a uniform
+         distribution over the set $x$.}
+\end{figure}
 
-    \node [var] (appearance) {Appearance}; \&
-    \node [var] (shape) {Shape}; \&
-    \node [var] (size) {Size}; \\
-  };
-\end{tikzpicture}}
-
-\caption{\label{fig:unsupervised-threshold}
-         Assuming a uniform unconditional distribution of any subset of the variables $x$,
-         a novelty threshold can be defined on the function $P_G(x)/P_{G'}(x)$.}
-\end{figure}
+Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_{U}(x)$.
 
@@
 Nonetheless in several cases its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability distribution that
+Under that it becomes possible to obtain a better approach to the unconditional probability distribution than
 the uniform one assumed in \autoref{sec:uniform-unconditional}.
 
-For interesting and practical reasons it was assumed that the unconditional distribution could be modelled with
-all the variables independent from each other.
-Which is the same as assuming $P(x)=\prod{P_{x_i}(x_i)}$.
+For interesting and practical reasons it was assumed that the unconditional distribution
+could be modelled with all the variables independent from each other.
+Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
 And so from the unlabelled data we only need to estimate $P_{x_i}$.
-Once again \autoref{fig:semisupervised-threshold} illustrates this in terms of graphs $G$ and $G'$.
+\autoref{fig:semisupervised-threshold} illustrates this in terms of graphs.
 
@@
 \centering
-\subfloat[Factor graph $G$ used to model conditional probability]{
-\begin{tikzpicture}
-  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
-    \& \node [var] (room) {Room Category}; \& \\
-
-    \node [factor] (fAppearance) {}; \&
-    \node [factor] (fShape) {}; \&
-    \node [factor] (fSize) {}; \\
-
-    \node [var] (appearance) {Appearance}; \&
-    \node [var] (shape) {Shape}; \&
-    \node [var] (size) {Size}; \\
-  };
-  \path [line] (room) -- (fAppearance);
-  \path [line] (room) -- (fShape);
-  \path [line] (room) -- (fSize);
-
-  \path [line] (fAppearance) -- (appearance);
-  \path [line] (fShape) -- (shape);
-  \path [line] (fSize) -- (size);
-\end{tikzpicture}}
-\hfill
-\subfloat[Factor graph $G'$ used to model a unconditional probability distribution]{
-\begin{tikzpicture}
-  \matrix[row sep=1em,column sep=1.5em, ampersand replacement=\&] {
-    \node [var] (appearance) {Appearance}; \&
-    \node [var] (shape) {Shape}; \&
-    \node [var] (size) {Size}; \\
-
-    \node [factor] (fAppearance) {}; \&
-    \node [factor] (fShape) {}; \&
-    \node [factor] (fSize) {}; \\
-  };
-  \path [line] (fAppearance) -- (appearance);
-  \path [line] (fShape) -- (shape);
-  \path [line] (fSize) -- (size);
-\end{tikzpicture}}
-
-\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to create a better approximation
-         for the unconditional probability. Which should lead to a better novelty threshold.}
+\includegraphics[width=0.70\textwidth]{figures/independent-prob-graph.pdf}
+\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to estimate
+         the unconditional probability for each variable and this way try to compensate
+         for a possible existent bias on some of the variables.}
 \end{figure}
 
-The described ratio between the two graphs can be understood as compensating for an existing bias
-on the unconditional distribution.
+Once again the novelty threshold would be given by $P_G(x)/P_U(x)$. This ratio between the two graphs
+can be understood as compensating for an existing bias on the unconditional distribution.
 Which is an important step to achieve a correct ordering of the input space for novelty thresholding.
-This way our unconditional function perceive which features are slightly more biased to a certain value and will
-compensate on the probability calculation.
+This way our unconditional function perceives which features are slightly more biased to a certain value
+and will compensate on the novelty calculation.
 

### 2011-05-06 semantic-nd/article.tex
More stuff on novelty detection.

@@
 
-Such an ordering can be defined by a R-value function on the input space. And then a threshold can be trained
-on that function in order to accept an expected amount of cases at the cost of failing in others.
-
-Based on bayes rule:
+Using bayes rule on the reversed order $P(\overline{novel}|x)$ and assuming a constant $P(\overline{novel})$
+we obtain a ratio between a conditional and unconditional probabilites on the input $x$.
+Such a ratio is a suitable function for implementing an optimal novelty detector system with
+thresholding.
 
 \begin{equation}
-P(\overline{novel}|x) = \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
+\label{eq:novelty-threshold}
+P(\overline{novel}|x) = \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)} = k\frac{P(x|\overline{novel})}{P(x)}
 \end{equation}
@@
 
-Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_{U}(x)$.
+Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all
+possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_{U}(x)$.
+
+Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of variables $x$ into the same
+measure units, such that a static threshold can be implemented.
+For example in very large sets of variables $x$, a large dimension space, the conditional probability would spread over
+the tails of the functions and would always yield very small values.
+As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to understand the observed result.
 
 \subsection{Using Unlabelled Data}
-Nonetheless in several cases its often the case that there is access to large amounts of unlabelled data.
+Nonetheless its often the case that there is access to large amounts of unlabelled data.
 Under that it becomes possible to obtain a better approach to the unconditional probability distribution than
@@
 And so from the unlabelled data we only need to estimate $P_{x_i}$.
-\autoref{fig:semisupervised-threshold} illustrates this in terms of graphs.
+\autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
 
@@
 
-Once again the novelty threshold would be given by $P_G(x)/P_U(x)$. This ratio between the two graphs
-can be understood as compensating for an existing bias on the unconditional distribution.
+Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
+This time the addition of the variables unconditional factors can be understood as an attempt to compensate
+for an existing bias on the unconditional distribution.
 Which is an important step to achieve a correct ordering of the input space for novelty thresholding.
@@
 And although it serves as a method for detecting new classes in the input, it gives us no control in what
-variables novelty is being measured.
-This is the novelty can come from any variable present on the graphs and we have no way to measure it.
+is novel in the input.
+Recall that the graph $G$ used by Dora gives us a lot of semantic and conceptual information that we are
+interested in analizying and improving with time. Being able to measure if the novel case derives from
+a novel room category or from any other variables would be highly desirable on the system.
 
@@
 
-\begin{figure}
-\caption{\label{fig:measure-room-novelty}In this case the only difference between $G$ and $G'$ is the room category variable,
-         that only exists in the later.}
+\begin{figure}[h]
+\subfloat[Conditional probability distribution $G$]{\includegraphics[width=0.45\textwidth]{figures/simple-conditional-prob-graph.pdf}}
+\hfill
+\subfloat[Unconditional probability distribution $G'$]{\includegraphics[width=0.45\textwidth]{figures/simple-unconditional-prob-graph.pdf}}
+
+\caption{\label{fig:measure-room-novelty}In this case the only difference between $G$ and $G'$ is the room category variable
+         and the associated factors that only exists in the former.}
 \end{figure}
 
-From now on we refer to the measure $P_G(x)/P_{G'}(x)$, when $G$ and $G'$ as a novelty measure on the variable $R_{category}$.
-As it attempts to measure on how much adding the variable $R_{category}$ helps to explain the given distribution of values.
+The notation $N_v$ is now used to refer to the measure $P_G(x)/P_{G'}(x)$, when $G$ and $G'$ differ only in the variable node $v$ and
+its associated factor nodes.
+And use it as a novelty measure of the variable $v$, as it attempts to measure on how much adding the variable $v$ helps to explain the
+given distribution of values.
 
-Expanding the calculation of $P_G(x)/P_{G'}(x)$ with the sum-product technique. Leads to:
+By assuming the graph is a tree and blabla, and expanding the calculation of $P_G(x)/P_{G'}(x)$ with the sum-product technique, leads to:
+
+\begin{equation}
+N_v(x) = \frac{\sum_{s \in States(v)}x / normalizing_{factor1}}{\prod{...} / normalizing_{factor2}}
+\end{equation}
+
+blablablaaa....
+I haven't done the math here yet.... but it should give us a method for measuring on how much each variable is contributing to the novelty
+threshold.
+Also due to the nature of the graphical models it should be possible to calculate in $O(N)$ time for all the $N$ variables.
+After the whole variables can be sorted by the contributions and presented to the user explaining where the novelty comes from.
+
+Quotting Andrzej: """ Being able to automatically explain why a class is novel and what
+makes it different from any of the other classes -- This is super important for communication with the user and active learning of new room categories."""
 
@@
 \end{itemize}
+
+In order to verify the improvements obtained by using unlabelled data, a fictitious simple dataset was generated.
+The fictitious distribution assumes that an independent and variable size set of features $x$ is generated by a
+given room category category. In whole there was 11 different room categories and 7 different measured properties types.
+
+The known data was extracted by generating samples from 6 of the 11 classes. Together with the generated known data, the
+room category that generated is given as label. Using that we built the factors for the generation of graphs representing
+the conditional distribution $G$.
+Using unlabelled data we trained also a graph $I$ representing an unconditional distribution where all features are independent.
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.60\textwidth]{figures/roc-simple.pdf}
+
+This graph is from a previous experience. Where there was even no need for the normalizing factor.
+
+\caption{\label{fig:fictitious-roc}ROC curve comparing novelty detection performance between an optimal detector,
+         $P_G(x)/P_U(x)$ a graph ratio considering an uniform unconditional distribution and
+         $P_G(x)/P_I(x)$ a graph ratio considering an independent unconditional distribution}
+\end{figure}
+
+Since the distribution is fictitious we were able to create an optimal novelty detector, as we can exactly measure $P(novel|x)$.
+That served as a comparison on how optimal a detector could be.
+A ROC curve was plotted for the three methods on \autoref{fig:fictitious-roc}.
+As possible to see a correct estimation on the unconditional probability is important for obtaining a correct ordering of the
+input space.
 

### 2011-05-08 semantic-nd/article.tex
Several changes before cleaning the article to remove the novelty introduced by a part of the graph.

@@
 model that is trained from unlabelled data.
+The concepts are then generalized to measure novelty introduced by each modelled
+variable allowing the system to explain why a class is novel.
 
-\keywords{novelty detection, semantic data, graphical models, room classification,
-indoor environments, robotics, multi-modal.}
+\keywords{novelty detection, semantic data, probabilistic graphical models,
+room classification, indoor environments, robotics, multi-modal classification.}
 \end{abstract}
@@
 At the same time they are a generative models and therefore allow to calculate the probability
-on any given subset of variables on the graph allowing the system to work even when some information is missing.
+on any given subset of variables on the graph allowing the system to work even when some
+information is missing.
 
@@
 \includegraphics[width=0.49\textwidth]{figures/chain-graph.png}
-\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge in order to
-         create a structured probabilistic representation of the world.}
+\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
+         in order to create a structured probabilistic representation of the world.}
 \end{figure}
@@
 Although the conceptual layer works with \emph{chain graph}, those can be converted into \emph{factor graphs}.
-We opted to use those as describing the ideas on this paper as they tend to be easier to understand.
+Those are used through this paper as they provide an easier manipulation in terms of factorization.
 
@@
 Each factor represents a function dependent only on the variables to where its connected.
-As so given factor graph $G$ can be seen as description of a distribution function over a set of variables obtained by
-the product of all the factors. This manipulation allows to use algorithms such as the Sum-Product to efficiently calculate
-probabilities on any given subset $x$ of those variables by exploiting conditional independency between variables.
-We will use the notation $P_G(x)$ to represent that.
+As so given factor graph $G$ can be seen as description of a distribution function over a set
+of variables obtained by the product of all the factors. In order to represent a probability
+a factor $Z$ is defined
+
+\begin{equation}
+P_G(X_V) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
+\end{equation}
+
+Describing the distribution function in terms of graphs allows to use algorithms such as the
+Sum-Product to efficiently calculate marginals on any given subset $x$ of variables by
+exploiting conditional independency between variables.
+
+\begin{equation}
+M_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
+\end{equation}
 
@@
 % Explain threshold approach to novelty detection
-Due to the desired of robustness and generalization, a novelty detection system would always possess
+Due to the desire of robustness and generalization, a novelty detection system would always possess
 some threshold that describes on how strict the system should be to new data.
@@
 
-When dealing with a statistic point of view, noisy data or unstable features, a decision on a given input has
-an associated error rate. And in this case a novelty detection system is interested in defining an order
-on all the possible inputs equivalent to the order imposed by the error rate: $P(novel|x)$.
+When dealing with a statistic point of view, noisy data or unstable features, a decision on a
+given input has an associated error rate. A novelty detection system is interested in defining an
+order on all the possible inputs equivalent to the order imposed by the error rate: $P(novel|x)$.
 
 Using bayes rule on the reversed order $P(\overline{novel}|x)$ and assuming a constant $P(\overline{novel})$
-we obtain a ratio between a conditional and unconditional probabilites on the input $x$.
+we obtain a ratio between a conditional and unconditional probabilities of the input $x$.
 Such a ratio is a suitable function for implementing an optimal novelty detector system with
@@
 \label{eq:novelty-threshold}
-P(\overline{novel}|x) = \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)} = k\frac{P(x|\overline{novel})}{P(x)}
+          P(\overline{novel}|x)
+  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
+  \propto \frac{P(x|\overline{novel})}{P(x)}
 \end{equation}
@@
 
-One of the hardest problems is that on most cases a novelty detection system only have access to 
+One of the hardest problems is that on most cases a novelty detection system only has access to 
 the known data.
@@
 layer to represent the conditional probability on the sensed variables $x$.
-$P_G(x)$ is then used as an approximation for $P(x|\overline{novel})$.
+$M_G(x)$ is used as an approximation for $P(x|\overline{novel})$.
 
@@
 
-Its important to notice that in several cases assuming the unconditional probability $P(x)$ is constant leads to discarding the factor.
+Its important to notice that in several cases assuming the unconditional probability $P(x)$
+is constant leads to discarding the factor.
 Nonetheless, here the distributions are dynamically changing as the system learns more on the environment.
-So we need to keep the normalizing argument $P(x)$ and re-calculate it for each new subset of $x$.
+So the normalizing argument $P(x)$ has to be evaluated for each new subset of $x$.
 
@@
 \includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
-\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a uniform
-         distribution over the set $x$.}
+\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a
+         uniform distribution over any set of its variables.}
 \end{figure}
@@
 Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all
-possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_{U}(x)$.
+possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_U(x)$.
 
-Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of variables $x$ into the same
-measure units, such that a static threshold can be implemented.
-For example in very large sets of variables $x$, a large dimension space, the conditional probability would spread over
-the tails of the functions and would always yield very small values.
-As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to understand the observed result.
+Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of
+variables $x$ into the same measure units (error rate), such that a static threshold can be implemented.
+For example in very large sets of variables $x$, a large dimension space, the conditional
+probability would spread over the tails of the functions and $P_G(x)$ would always yield very
+small values.
+As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
+understand the observed result.
 
@@
 Nonetheless its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability distribution than
-the uniform one assumed in \autoref{sec:uniform-unconditional}.
+Under that it becomes possible to obtain a better approach to the unconditional probability
+distribution than the uniform one assumed in \autoref{sec:uniform-unconditional}.
 
@@
 Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-And so from the unlabelled data we only need to estimate $P_{x_i}$.
+And so from the unlabelled data we only need to estimate the probability of each feature $x_i$.
 \autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
@@
 Which is an important step to achieve a correct ordering of the input space for novelty thresholding.
-This way our unconditional function perceives which features are slightly more biased to a certain value
-and will compensate on the novelty calculation.
 
-\subsection{Generalizing to Novelty Detection on a variable}
-Until now, the presented methods describe how to detect data that belongs to a new class of input by creating
-generative models for both conditional and unconditional probabilities distributions.
+\subsection{Generalizing to Novelty Detection on a Variable}
+Until now, the presented methods describe how to detect data that belongs to a new class of input
+by creating generative models for both conditional and unconditional probabilities distributions.
 
-And although it serves as a method for detecting new classes in the input, it gives us no control in what
-is novel in the input.
-Recall that the graph $G$ used by Dora gives us a lot of semantic and conceptual information that we are
-interested in analizying and improving with time. Being able to measure if the novel case derives from
-a novel room category or from any other variables would be highly desirable on the system.
-
-Lets assume that the only difference between the conditional and unconditional graph is a single node and its connecting factors
-as seen on \autoref{fig:measure-room-novelty}.
+And although it serves as a method for detecting new classes in the input, it gives no information
+in why a class is novel.
+The graphical model used by Dora gives a lot of semantic and conceptual information that would be
+interesting in exploiting to be able to communicate novelty with a user and to perform
+active learning of new classes.
+Being able to measure which part of the graph is responsible for triggering novelty would be
+a desirable property on the system.
+In that sense focus as been made on detecting the impact of any given node.
 
@@
 \subfloat[Conditional probability distribution $G$]{\includegraphics[width=0.45\textwidth]{figures/simple-conditional-prob-graph.pdf}}
-\hfill
+\qquad
 \subfloat[Unconditional probability distribution $G'$]{\includegraphics[width=0.45\textwidth]{figures/simple-unconditional-prob-graph.pdf}}
@@
 \end{figure}
+
+Let the conditional and unconditional probabilities distributions be modeled by graphs
+$G$ and $G'$.
+And the only difference between those be the non-existence of factors connected to a given
+variable $A$ as seen on \autoref{fig:measure-room-novelty}.
+Let $N_A$ denote the set of neighbour variables of node $A$.
+
+The ratio: $P_G(x)/P_{G'}(x)$ is then given by:
+
+\begin{equation}
+\frac{\sum \prod ... }{\sum \prod ...}  \frac{Z_{G'}}{Z_G}
+\end{equation}
 

### 2011-05-08 semantic-nd/article.tex
Cleaning from novelty detection on given variable of a node.

@@
 % first the title is needed
-\title{Semantic Novelty Detection with Graphical Models}
+\title{Using Graphical Models for Novelty Detection on Semantic Room Classification}
 
@@
 \begin{abstract}
-This papers presents an approach on implementing novelty detection at semantic level
-for indoor room classification in Dora. \emph{Graphical models} are used to model
-probabilistic knowledge and novelty detection is implemented on top of them.
+This papers presents an approach on implementing novelty detection for indoor room classification
+using semantic data.
+\emph{Graphical models} are used to model probabilistic knowledge and a novelty threshold is
+defined in terms of conditional and unconditional probabilities.
 The novelty threshold is then optimized using an unconditional probability density
-model that is trained from unlabelled data.
-The concepts are then generalized to measure novelty introduced by each modelled
-variable allowing the system to explain why a class is novel.
+model trained from unlabelled data.
 
@@
 
-There has been several efforts in the area of Artificial Intelligence and Robotics in creating robots
-that are able to interact with humans and their environments.
-One of the existing problems is a reliable high-level localization method that can be deployed into new
-and unknown environments.
-In this article we focus on the mapping, using \emph{semantic data}, of indoor environments such as houses and offices to room categories
-such as kitchen, corridor, office.
-And how to perform novelty detection on them: \emph{detect a new room category that was not present in the labelled data}.
+There has been several efforts in the area of Artificial Intelligence and Robotics in creating
+robots that are able to interact with humans and their environments.
+One of the existing problems is a reliable high-level localization method that can be deployed
+into new and unknown environments.
+This article focus on the mapping, using \emph{semantic data}, of indoor environments such as
+houses and offices to room categories such as kitchen, corridor, office.
+And how to perform novelty detection on them: \emph{detect a new room category that was not
+present in the labelled data}.
 
-Dora\cite{dora} (CogX: Dora) was used as a base system where this novelty detection would be implemented.
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and probabilistic representation of the space
-instantiated as a \emph{graphical model}.
+Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the presented novelty
+detection system.
+As Dora moves through the environment its \emph{conceptual layer} builds a structural and
+probabilistic representation of the space instantiated as a \emph{graphical model}.
 That model connects the sensed properties together with the variables used to model the world.
@@
 The edges between the variables can be seen as a kind of filter to the properties of the system.
-Properties those that we expect to capture by the conceptual knowledge.
+Properties those that are expected to be captured by the conceptual knowledge.
 At the same time they are a generative models and therefore allow to calculate the probability
@@
 \subsection{Factor Graphs}
-Although the conceptual layer works with \emph{chain graph}, those can be converted into \emph{factor graphs}.
+Although the conceptual layer works with \emph{chain graphs}, those can be converted into \emph{factor graphs}.
 Those are used through this paper as they provide an easier manipulation in terms of factorization.
 
-A \emph{factor graph} is a bipartite graph connecting nodes representing random variables and factors.
+A \emph{factor graph} $G$ is a bipartite graph connecting a set of nodes $V_G$ representing
+random variables and factors $\phi_G$.
 Each factor represents a function dependent only on the variables to where its connected.
@@
 of variables obtained by the product of all the factors. In order to represent a probability
-a factor $Z$ is defined
+a normalization factor $Z$ needs to be introduced:
 
@@
 \begin{equation}
-M_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
+P_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
 \end{equation}
@@
 \item Using unlabelled data
-\item What happens if the conditional and unconditional graphs differ only by the addition of a node?
-\item How to implement that in the case a graph is a tree
-\item Novelty detection on any variable of the graphical model!
 \end{itemize}
@@
 Assuming that the unconditional distributions generates all possible outcome with the same probability
-we can model it with $\prod{1/|x_i|}$, where $|x_i|$ denotes the state space size of variable $x_i$.
-In graphical model terms, this would be equivalent to build a factor graph $U$ with the same variables but
+we can model it with $\prod{1/\# x_i}$, where $\# x_i$ denotes the cardinality of the state space
+of variable $x_i$.
+In graphical model terms this is represented to a factor graph $U$ with the variables but
 without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
@@
 variables $x$ into the same measure units (error rate), such that a static threshold can be implemented.
-For example in very large sets of variables $x$, a large dimension space, the conditional
-probability would spread over the tails of the functions and $P_G(x)$ would always yield very
-small values.
+For example the conditional probability would yield very small values on large sets of variables
+$x$ than in small sets due to the spreading over the dimensions of the input space.
 As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
@@
 Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
-This time the addition of the variables unconditional factors can be understood as an attempt to compensate
-for an existing bias on the unconditional distribution.
-Which is an important step to achieve a correct ordering of the input space for novelty thresholding.
-
-\subsection{Generalizing to Novelty Detection on a Variable}
-Until now, the presented methods describe how to detect data that belongs to a new class of input
-by creating generative models for both conditional and unconditional probabilities distributions.
-
-And although it serves as a method for detecting new classes in the input, it gives no information
-in why a class is novel.
-The graphical model used by Dora gives a lot of semantic and conceptual information that would be
-interesting in exploiting to be able to communicate novelty with a user and to perform
-active learning of new classes.
-Being able to measure which part of the graph is responsible for triggering novelty would be
-a desirable property on the system.
-In that sense focus as been made on detecting the impact of any given node.
-
-\begin{figure}[h]
-\subfloat[Conditional probability distribution $G$]{\includegraphics[width=0.45\textwidth]{figures/simple-conditional-prob-graph.pdf}}
-\qquad
-\subfloat[Unconditional probability distribution $G'$]{\includegraphics[width=0.45\textwidth]{figures/simple-unconditional-prob-graph.pdf}}
-
-\caption{\label{fig:measure-room-novelty}In this case the only difference between $G$ and $G'$ is the room category variable
-         and the associated factors that only exists in the former.}
-\end{figure}
-
-Let the conditional and unconditional probabilities distributions be modeled by graphs
-$G$ and $G'$.
-And the only difference between those be the non-existence of factors connected to a given
-variable $A$ as seen on \autoref{fig:measure-room-novelty}.
-Let $N_A$ denote the set of neighbour variables of node $A$.
-
-The ratio: $P_G(x)/P_{G'}(x)$ is then given by:
-
-\begin{equation}
-\frac{\sum \prod ... }{\sum \prod ...}  \frac{Z_{G'}}{Z_G}
-\end{equation}
-
-The notation $N_v$ is now used to refer to the measure $P_G(x)/P_{G'}(x)$, when $G$ and $G'$ differ only in the variable node $v$ and
-its associated factor nodes.
-And use it as a novelty measure of the variable $v$, as it attempts to measure on how much adding the variable $v$ helps to explain the
-given distribution of values.
-
-By assuming the graph is a tree and blabla, and expanding the calculation of $P_G(x)/P_{G'}(x)$ with the sum-product technique, leads to:
-
-\begin{equation}
-N_v(x) = \frac{\sum_{s \in States(v)}x / normalizing_{factor1}}{\prod{...} / normalizing_{factor2}}
-\end{equation}
-
-blablablaaa....
-I haven't done the math here yet.... but it should give us a method for measuring on how much each variable is contributing to the novelty
-threshold.
-Also due to the nature of the graphical models it should be possible to calculate in $O(N)$ time for all the $N$ variables.
-After the whole variables can be sorted by the contributions and presented to the user explaining where the novelty comes from.
-
-Quotting Andrzej: """ Being able to automatically explain why a class is novel and what
-makes it different from any of the other classes -- This is super important for communication with the user and active learning of new room categories."""
+This time the addition of the variables unconditional factors can be understood as an
+attempt to compensate for an existing bias on the unconditional distribution.
+Which is an important step to achieve a correct ordering of the input space for novelty
+thresholding.
 
@@
 
+On this paper it was presented how to define a stable novelty threshold function on top of
+\emph{probabilistic graphical models} instantiated dynamically from sensed semantic data and
+conceptual knowledge.
+
+It was also shown that unlabelled data can be used to approximate unconditional probability and
+improve a novelty threshold.
+
+Quotting Andrzej: """ Being able to automatically explain why a class is novel and what
+makes it different from any of the other classes -- This is super important for communication with the user and active learning of new room categories."""
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-05-09 semantic-nd/article.tex
Added stuff specially on results section.

@@
 layer to represent the conditional probability on the sensed variables $x$.
-$M_G(x)$ is used as an approximation for $P(x|\overline{novel})$.
+$P_G(x)$ is used as an approximation for $P(x|\overline{novel})$.
 
@@
 Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-And so from the unlabelled data we only need to estimate the probability of each feature $x_i$.
+From the unlabelled data only the probability of each feature $x_i$ needs to be estimated.
 \autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
@@
 
-In order to verify the improvements obtained by using unlabelled data, a fictitious simple dataset was generated.
-The fictitious distribution assumes that an independent and variable size set of features $x$ is generated by a
-given room category category. In whole there was 11 different room categories and 7 different measured properties types.
+In order to verify the improvements obtained by using unlabelled data, a fictitious simple
+dataset was generated.
+The fictitious distribution assumes that an independent and variable size set of features
+$x$ is generated by a given room category.
+In whole there was 11 different room categories and 7 different measured properties types.
 
-The known data was extracted by generating samples from 6 of the 11 classes. Together with the generated known data, the
-room category that generated is given as label. Using that we built the factors for the generation of graphs representing
-the conditional distribution $G$.
-Using unlabelled data we trained also a graph $I$ representing an unconditional distribution where all features are independent.
+The known data was extracted by generating samples from 6 of the 11 classes.
+Together with the generated known data, the room category that generated is given as label.
+That known data is used to train factors to represent a conditional distribution.
+Using unlabelled data, factors are trained to approximate an unconditional distribution
+where all features are independent.
+\autoref{fig:simple-experiment} shows the graph structure used for approximate the conditional and
+unconditional distributions trained from the labelled and unlabelled data.
+
+\begin{figure}[h]
+\centering
+
+\subfloat[Graph structure used to model conditional probability graph.
+          The graph factors are trained using the labelled data.]{... conditional probability}
+\qquad
+\subfloat[Graph structure used to model unconditional probability graph.
+          The graph factors are trained using the unlabelled data.]{... unconditional probability}
+
+\caption{\label{fig:simple-experiment}The graph structures $G$ and $I$ used to model the
+         conditional and unconditional probability.
+         Note that graph $G$ correctly models the structure of the fictitious distribution, though
+         during training it has not seen all possible room categories represented by variable $R_1$.}
+
+\end{figure}
 
@@
 
-Since the distribution is fictitious we were able to create an optimal novelty detector, as we can exactly measure $P(novel|x)$.
-That served as a comparison on how optimal a detector could be.
+Since the distribution is fictitious its possible to create an optimal novelty detector, as
+its the error rate $P(novel|x)$ can be exactly measured.
+That serves as a comparison on how optimal a detector can be using the same input space.
 A ROC curve was plotted for the three methods on \autoref{fig:fictitious-roc}.
-As possible to see a correct estimation on the unconditional probability is important for obtaining a correct ordering of the
-input space.
+As possible to see the unconditional probability plays an important role in obtaining a correct
+ordering of the input space.
+
+The importance of a correct estimate of unconditional probability is expected to decrease as the
+input space increases and allows the several existing classes to become easily distinguished.
+
+For comparison ROC curves where plot for the developed threshold based on the number of input features.
+
+\begin{figure}
+\centering
+
+\subfloat[3 sensed variables]{5 sensed variables}
+\qquad
+\subfloat[5 sensed variables]{5 sensed variables}
+
+\subfloat[10 sensed variables]{10 sensed variables}
+\qquad
+\subfloat[50 sensed variables]{50 sensed variables}
+
+\caption{\label{fig:fictitious-roc-breakdown}ROC curves plotted showing performance of the
+         presented novelty detection method on graphs generated for different amount of
+         sensed variables.
+         As possible to see as the system gains more semantic information it becomes easier
+         to detect novelty and the effect of picking an approximation for the unconditional
+         probability decreases.}
+\end{figure}
 
@@
 \begin{itemize}
-\item Show a correct estimation of unconditional probability to be important in estabilishing a novelty detection threshold
+\item Show a correct estimation of unconditional probability to be important in establishing a novelty detection threshold
 \item Show unlabelled data can be used for it.
@@
 conceptual knowledge.
+The presented technique is based on the ratio between a conditional and unconditional probability.
 
-It was also shown that unlabelled data can be used to approximate unconditional probability and
+It was shown that unlabelled data can be used to approximate unconditional probability and
 improve a novelty threshold.
+And that a correct estimation of unconditional probability plays an important role specially on
+small input spaces.
 
-Quotting Andrzej: """ Being able to automatically explain why a class is novel and what
-makes it different from any of the other classes -- This is super important for communication with the user and active learning of new room categories."""
+Future will work will focus on how to explain why a class is novel and what makes it different
+other classes.
+That will lead to generation of useful information that can be used on communication with the user
+and perform active learning of new room categories.
 

### 2011-05-09 semantic-nd/article.tex
some cleanup on what to write on each section and added dummy images.

@@
 \section{Introduction}
-\begin{itemize}
-\item Explain context of indoor room categorization and importance of novelty detection.
-\item Introduce Dora and explain how it would benefit or use the semantic novel detection.
-\end{itemize}
-
 There has been several efforts in the area of Artificial Intelligence and Robotics in creating
@@
 \section{Dora Architecture Overview}
-\begin{itemize}
-\item Introduce Dora architecture aspects important for the paper
-\item Explain what we mean by semantic data and how its captured by Dora
-\item Introduce graphical models and how we use them (including notation on the paper)
-\end{itemize}
-
 Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
@@
 \section{Novelty Detection}
-\begin{itemize}
-\item Define novelty detection 
-\item Explain approach with a threshold function
-\item Need of a normalizing factor for a threshold with dynamic functions
-\item Usage of conditional density
-\item Assuming of every output is equally likely
-\item Using unlabelled data
-\end{itemize}
-
 % Introduce novelty detection
@@
 \section{Results}
-\begin{itemize}
-\item Show results and how much it improves by using a simple method for unconditional probability
-\end{itemize}
-
 In order to verify the improvements obtained by using unlabelled data, a fictitious simple
@@
 \centering
-\includegraphics[width=0.60\textwidth]{figures/roc-simple.pdf}
+\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}
 
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
 
-\subfloat[3 sensed variables]{5 sensed variables}
+\subfloat[3 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
 \qquad
-\subfloat[5 sensed variables]{5 sensed variables}
+\subfloat[5 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
 
-\subfloat[10 sensed variables]{10 sensed variables}
+\subfloat[10 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
 \qquad
-\subfloat[50 sensed variables]{50 sensed variables}
+\subfloat[50 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
 
@@
 \section{Conclusions and Future Work}
-\begin{itemize}
-\item Show a correct estimation of unconditional probability to be important in establishing a novelty detection threshold
-\item Show unlabelled data can be used for it.
-\item Although focus on Room Categorization we expect the results and techniques to be usable on other areas.
-\end{itemize}
 

### 2011-05-10 semantic-nd/article.tex
changed fictitious for synthetic

@@
 \section{Results}
-In order to verify the improvements obtained by using unlabelled data, a fictitious simple
-dataset was generated.
-The fictitious distribution assumes that an independent and variable size set of features
+In order to verify the improvements obtained by using unlabelled data, a synthetic dataset
+was generated.
+The synthetic distribution assumes that an independent and variable size set of features
 $x$ is generated by a given room category.
@@
          conditional and unconditional probability.
-         Note that graph $G$ correctly models the structure of the fictitious distribution, though
+         Note that graph $G$ correctly models the structure of the synthetic distribution, though
          during training it has not seen all possible room categories represented by variable $R_1$.}
@@
 
-\caption{\label{fig:fictitious-roc}ROC curve comparing novelty detection performance between an optimal detector,
+\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance between an optimal detector,
          $P_G(x)/P_U(x)$ a graph ratio considering an uniform unconditional distribution and
@@
 
-Since the distribution is fictitious its possible to create an optimal novelty detector, as
+Since the distribution is synthetic its possible to create an optimal novelty detector, as
 its the error rate $P(novel|x)$ can be exactly measured.
 That serves as a comparison on how optimal a detector can be using the same input space.
-A ROC curve was plotted for the three methods on \autoref{fig:fictitious-roc}.
+A ROC curve was plotted for the three methods on \autoref{fig:synthentic-roc}.
 As possible to see the unconditional probability plays an important role in obtaining a correct
@@
 
-\caption{\label{fig:fictitious-roc-breakdown}ROC curves plotted showing performance of the
+\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
          presented novelty detection method on graphs generated for different amount of

### 2011-05-10 semantic-nd/abstract.tex
Added rule for abstract blind submission

--- 2011-05-10 semantic-nd/article.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[round]{natbib}
-\usepackage{subfig}
-\usepackage{tikz}
-\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
-\hypersetup{%
-   plainpages=false, 
-   pdfpagelayout=SinglePage,
-   bookmarksopen=false,
-   bookmarksnumbered=true,
-   breaklinks=true,
-   linktocpage,
-   colorlinks=true,
-   linkcolor=blue,
-   urlcolor=blue,
-   citecolor=blue,
-   anchorcolor=green
-}
-
-% TiKZ styles
-\tikzstyle{var} = [circle,
-                   thick,
-                   draw, fill=red!20,
-                   font=\scriptsize,
-                   text width=3.5em,
-                   text centered,
-                   node distance=1em,
-                   inner sep=0pt]
-
-\tikzstyle{factor} = [rectangle,
-                      thick,
-                      draw, fill=blue!20,
-                      minimum height=2em,
-                      minimum width=2em]
-
-\tikzstyle{line} = [draw]
 
@@
 %
-\author{André Susano Pinto}
+\author{Ommited for Review}
 %
@@
 % unless you accept that it will be published
-\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
-\url{andresusanopinto@gmail.com}\\
-\url{http://url}}
+\institute{Institution ommited for Review}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-There has been several efforts in the area of Artificial Intelligence and Robotics in creating
-robots that are able to interact with humans and their environments.
-One of the existing problems is a reliable high-level localization method that can be deployed
-into new and unknown environments.
-This article focus on the mapping, using \emph{semantic data}, of indoor environments such as
-houses and offices to room categories such as kitchen, corridor, office.
-And how to perform novelty detection on them: \emph{detect a new room category that was not
-present in the labelled data}.
-
-Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the presented novelty
-detection system.
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
-That model connects the sensed properties together with the variables used to model the world.
-And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box\cite{exploiting}.
-
-The ability to recognize novel categories on the modelled variables would allow to increase
-reliability and allow the creation of self-extending behaviours.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor
-scenes when compared to outdoor scenes results.
-Since the properties that characterize rooms changes conforming its category. Namely corridors are
-well described by global properties and bookstores are well described by the presence of specific objects (books).
-Their work shows a multi-modal approach is expected to yield better performance by mixing several sources of information.
-
-\cite{galindo2005multi} defines a bidirectional relation between object and room category, where object defines a room category and a room category provides information on where objects may be found.
-
-Probabilistic representations are used in several localised functions in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
-And some employ, up to some extent, a probabilistic representation across some subsystems~\cite{kraft2008exploration}.
-\cite{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
-And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.
-
-Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.
-
-Although this paper deals with novelty detection on the robotics area, it does so using very
-standard concepts and techniques such as semantic data and graphical models.
-Those are often found on areas related with information retrieval.
-Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{zhang2005probabilistic}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dora Architecture Overview}
-Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
-
-From its architecture only the \emph{conceptual layer} is of interest to this article.
-Its role is to aggregate the following semantic information coming from other layers:
-
-\begin{description}
- \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
- \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room.
-The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
- \item[Object detection] is performed using the visual input in order to detect .
-The system keeps track of the number of object and types seen per each room.
-Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
- \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
-\end{description}
-
-With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
-in order to model all known variables and their relations.
-As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
-a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.
-
-The use of graphical models to describe distributions of variables has useful properties.
-The edges between the variables can be seen as a kind of filter to the properties of the system.
-Properties those that are expected to be captured by the conceptual knowledge.
-At the same time they are a generative models and therefore allow to calculate the probability
-on any given subset of variables on the graph allowing the system to work even when some
-information is missing.
-
-\begin{figure}[h]
-\centering
-
-\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}
-\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}
-\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
-         in order to create a structured probabilistic representation of the world.}
-\end{figure}
-
-\subsection{Factor Graphs}
-Although the conceptual layer works with \emph{chain graphs}, those can be converted into \emph{factor graphs}.
-Those are used through this paper as they provide an easier manipulation in terms of factorization.
-
-A \emph{factor graph} $G$ is a bipartite graph connecting a set of nodes $V_G$ representing
-random variables and factors $\phi_G$.
-Each factor represents a function dependent only on the variables to where its connected.
-As so given factor graph $G$ can be seen as description of a distribution function over a set
-of variables obtained by the product of all the factors. In order to represent a probability
-a normalization factor $Z$ needs to be introduced:
-
-\begin{equation}
-P_G(X_V) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
-\end{equation}
-
-Describing the distribution function in terms of graphs allows to use algorithms such as the
-Sum-Product to efficiently calculate marginals on any given subset $x$ of variables by
-exploiting conditional independency between variables.
-
-\begin{equation}
-P_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
-\end{equation}
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[node distance = 2cm, auto]
-    \matrix[row sep=0.5cm,column sep=0.5cm] {
-      & \node [var] (room) {Room Category}; \\
-      \node [factor] (fAppearance) {}; &
-      \node [factor] (fShape) {}; &
-      \node [factor] (fSize) {}; \\
-
-      \node [var] (appearance) {Appearance}; &
-      \node [var] (shape) {Shape}; &
-      \node [var] (size) {Size}; \\
-    };
-  \path [line] (room) -- (fAppearance);
-  \path [line] (room) -- (fShape);
-  \path [line] (room) -- (fSize);
-
-  \path [line] (fAppearance) -- (appearance);
-  \path [line] (fShape) -- (shape);
-  \path [line] (fSize) -- (size);
-\end{tikzpicture}
-\caption{Example of a factor graph used for Dora for representing the environment.}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Novelty Detection}
-% Introduce novelty detection
-Novelty detection is an harder problem than that of classification due to its single-class nature.
-Essentially a system is given training data and when given new data has to decide if it belongs
-to the same concept it was trained with.
-
-% Explain threshold approach to novelty detection
-Due to the desire of robustness and generalization, a novelty detection system would always possess
-some threshold that describes on how strict the system should be to new data.
-This threshold is often seen as a distance measure to a normal input and several novelty detection
-techniques are based on it.
-
-When dealing with a statistic point of view, noisy data or unstable features, a decision on a
-given input has an associated error rate. A novelty detection system is interested in defining an
-order on all the possible inputs equivalent to the order imposed by the error rate: $P(novel|x)$.
-
-Using bayes rule on the reversed order $P(\overline{novel}|x)$ and assuming a constant $P(\overline{novel})$
-we obtain a ratio between a conditional and unconditional probabilities of the input $x$.
-Such a ratio is a suitable function for implementing an optimal novelty detector system with
-thresholding.
-
-\begin{equation}
-\label{eq:novelty-threshold}
-          P(\overline{novel}|x)
-  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
-  \propto \frac{P(x|\overline{novel})}{P(x)}
-\end{equation}
-
-% Explain approach of using conditional probability
-\subsection{Conditional Probability: $P(x|\overline{novel})$}
-\label{sec:uniform-unconditional}
-
-One of the hardest problems is that on most cases a novelty detection system only has access to 
-the known data.
-Under that the best approach is to define a threshold based on $P(x|\overline{novel})/P(x)$
-considering that $P(x)$ is constant through all the input space.
-
-In the presented case the conditional probability is approximated by the graphical model $G$
-produced by the conceptual layer.
-The \autoref{fig:conditional-prob-graph} represents a possible graph $G$ built from the conceptual
-layer to represent the conditional probability on the sensed variables $x$.
-$P_G(x)$ is used as an approximation for $P(x|\overline{novel})$.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
-\caption{\label{fig:conditional-prob-graph}The conceptual layer models the conditional
-         probability distribution of the sensed variables by using hidden variables.
-         Those are created from the conceptual knowledge and represent the structural
-         and connectivity information the system is aware of.
-         In this case only hidden variables $R_i$ were used to model the room categories
-         where each property was sensed.}
-\end{figure}
-
-Its important to notice that in several cases assuming the unconditional probability $P(x)$
-is constant leads to discarding the factor.
-Nonetheless, here the distributions are dynamically changing as the system learns more on the environment.
-So the normalizing argument $P(x)$ has to be evaluated for each new subset of $x$.
-
-Assuming that the unconditional distributions generates all possible outcome with the same probability
-we can model it with $\prod{1/\# x_i}$, where $\# x_i$ denotes the cardinality of the state space
-of variable $x_i$.
-In graphical model terms this is represented to a factor graph $U$ with the variables but
-without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
-\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a
-         uniform distribution over any set of its variables.}
-\end{figure}
-
-Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all
-possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_U(x)$.
-
-Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of
-variables $x$ into the same measure units (error rate), such that a static threshold can be implemented.
-For example the conditional probability would yield very small values on large sets of variables
-$x$ than in small sets due to the spreading over the dimensions of the input space.
-As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
-understand the observed result.
-
-\subsection{Using Unlabelled Data}
-Nonetheless its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability
-distribution than the uniform one assumed in \autoref{sec:uniform-unconditional}.
-
-For interesting and practical reasons it was assumed that the unconditional distribution
-could be modelled with all the variables independent from each other.
-Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-From the unlabelled data only the probability of each feature $x_i$ needs to be estimated.
-\autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/independent-prob-graph.pdf}
-\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to estimate
-         the unconditional probability for each variable and this way try to compensate
-         for a possible existent bias on some of the variables.}
-\end{figure}
-
-Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
-This time the addition of the variables unconditional factors can be understood as an
-attempt to compensate for an existing bias on the unconditional distribution.
-Which is an important step to achieve a correct ordering of the input space for novelty
-thresholding.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-In order to verify the improvements obtained by using unlabelled data, a synthetic dataset
-was generated.
-The synthetic distribution assumes that an independent and variable size set of features
-$x$ is generated by a given room category.
-In whole there was 11 different room categories and 7 different measured properties types.
-
-The known data was extracted by generating samples from 6 of the 11 classes.
-Together with the generated known data, the room category that generated is given as label.
-That known data is used to train factors to represent a conditional distribution.
-Using unlabelled data, factors are trained to approximate an unconditional distribution
-where all features are independent.
-\autoref{fig:simple-experiment} shows the graph structure used for approximate the conditional and
-unconditional distributions trained from the labelled and unlabelled data.
-
-\begin{figure}[h]
-\centering
-
-\subfloat[Graph structure used to model conditional probability graph.
-          The graph factors are trained using the labelled data.]{... conditional probability}
-\qquad
-\subfloat[Graph structure used to model unconditional probability graph.
-          The graph factors are trained using the unlabelled data.]{... unconditional probability}
-
-\caption{\label{fig:simple-experiment}The graph structures $G$ and $I$ used to model the
-         conditional and unconditional probability.
-         Note that graph $G$ correctly models the structure of the synthetic distribution, though
-         during training it has not seen all possible room categories represented by variable $R_1$.}
-
-\end{figure}
-
-\begin{figure}[h!]
-\centering
-\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}
-
-This graph is from a previous experience. Where there was even no need for the normalizing factor.
-
-\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance between an optimal detector,
-         $P_G(x)/P_U(x)$ a graph ratio considering an uniform unconditional distribution and
-         $P_G(x)/P_I(x)$ a graph ratio considering an independent unconditional distribution}
-\end{figure}
-
-Since the distribution is synthetic its possible to create an optimal novelty detector, as
-its the error rate $P(novel|x)$ can be exactly measured.
-That serves as a comparison on how optimal a detector can be using the same input space.
-A ROC curve was plotted for the three methods on \autoref{fig:synthentic-roc}.
-As possible to see the unconditional probability plays an important role in obtaining a correct
-ordering of the input space.
-
-The importance of a correct estimate of unconditional probability is expected to decrease as the
-input space increases and allows the several existing classes to become easily distinguished.
-
-For comparison ROC curves where plot for the developed threshold based on the number of input features.
-
-\begin{figure}[h]
-\centering
-
-\subfloat[3 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
-\qquad
-\subfloat[5 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
-
-\subfloat[10 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
-\qquad
-\subfloat[50 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
-
-\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
-         presented novelty detection method on graphs generated for different amount of
-         sensed variables.
-         As possible to see as the system gains more semantic information it becomes easier
-         to detect novelty and the effect of picking an approximation for the unconditional
-         probability decreases.}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and Future Work}
-
-On this paper it was presented how to define a stable novelty threshold function on top of
-\emph{probabilistic graphical models} instantiated dynamically from sensed semantic data and
-conceptual knowledge.
-The presented technique is based on the ratio between a conditional and unconditional probability.
-
-It was shown that unlabelled data can be used to approximate unconditional probability and
-improve a novelty threshold.
-And that a correct estimation of unconditional probability plays an important role specially on
-small input spaces.
-
-Future will work will focus on how to explain why a class is novel and what makes it different
-other classes.
-That will lead to generation of useful information that can be used on communication with the user
-and perform active learning of new room categories.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{plainnat}
-\bibliography{refs}
-
 \end{document}

### 2011-05-11 semisupervised/dora-data/run.py
Changed Sconscript to generated pdfs using inkscape command line option --export-pdf=

@@ -19,7 +19,6 @@
 #
 import script
 import ontology
-import explain
 import random
 import dataset
 import graph

### 2011-05-11 semantic-nd/results/synthetic.py
Started organizing the script to run the synthetic experiments for the article.

@@ -1,0 +1,99 @@
+#!/usr/bin/python
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+# ####################################################################
+#
+# This file contains code that generates the synthetic experiment described on the paper.
+#
+# The following experiments want to be show:
+#  - The presented probability ratio is suitable for novelty detection,
+#  - Approximating the unconditional probability improves detection performance.
+#  - Measure effect of approximating unconditional probability as the available information increases.
+#
+
+import graph
+import ontology
+
+def generate_data(samples, classes, with_label):
+  while True:
+    raise NotImplemented
+
+def exact_conditional_prob(sample):
+  raise NotImplemented
+
+def exact_unconditional_prob(sample):
+  raise NotImplemented
+
+known_data = """"""
+unlabelled_data = """"""
+
+
+"""Train factors for conditional probability."""
+for label, sample in known_data:
+  for fid, descriptor in iter_features(sample):
+    ct[label][fid][descriptor] += 1
+
+for feature in features:
+  for value in feature.value_space:
+
+def conditional_prob(sample):
+  for fid, descriptor in iter_features(sample):
+    
+
+"""With knowledge on the feature space define a uniform unconditional probability function."""
+def uniform_unconditional_prob(sample):
+  p = 1.0
+  for fid, descriptor in iter_features(sample):
+    p *= 1.0/feature_space_size(fid)
+  return p
+
+"""Use unlabelled data for model an independent distribution on the unconditional data."""
+for sample in unlabelled_data:
+  for fid, descriptor in iter_features(sample):
+    ct[fit][descriptor] += 1
+
+def independent_unconditional_prob(sample):
+  raise NotImplemented
+
+
[…]

### 2011-05-12 semantic-nd/article.tex
Code to run experiments for paper added.

@@
 \centering
-\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}
-
-This graph is from a previous experience. Where there was even no need for the normalizing factor.
+\includegraphics[width=0.60\textwidth]{results/synthetic-all.pdf}
 
@@
 
-\subfloat[3 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
+\subfloat[3 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-3features.pdf}}
 \qquad
-\subfloat[5 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
+\subfloat[5 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-5features.pdf}}
 
-\subfloat[10 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
+\subfloat[10 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-10features.pdf}}
 \qquad
-\subfloat[50 sensed variables]{\includegraphics[width=0.40\textwidth]{figures/roc-simple.pdf}}
+\subfloat[50 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-50features.pdf}}
 

### 2011-05-12 semantic-nd/results/synthetic.py

@@ -26,50 +26,124 @@
 #  - Approximating the unconditional probability improves detection performance.
 #  - Measure effect of approximating unconditional probability as the available information increases.
 #
+import bisect
+import collections
+import random
+random.seed(0)
 
-import graph
 import ontology
+import graph
 
-def generate_data(samples, classes, with_label):
-  while True:
-    raise NotImplemented
+FEATURES    = ontology.features
+ALL_CLASSES = ontology.rooms.keys()
+POTENTIAL   = ontology.potential
+PROB        = dict()
 
-def exact_conditional_prob(sample):
-  raise NotImplemented
+KNOWN_CLASSES = """singleoffice kitchen robotlab""".split()
+for label in KNOWN_CLASSES: assert label in ALL_CLASSES
+
+"""Make potentials normalized. And calculate probabilities for the synthetic distribution."""
+for fid,fspace in FEATURES.items():
+  for cat in ALL_CLASSES:
+    S = sum([POTENTIAL[cat, fid, fvalue] for fvalue in fspace])
+    for fvalue in fspace:
+      POTENTIAL[cat, fid, fvalue] /= S
+    PROB[cat, fid] = []
+    acc_sum = 0.0
+    for fvalue in fspace:
+      PROB[cat,fid].append( (acc_sum, fvalue))
+      acc_sum += POTENTIAL[cat, fid, fvalue]
+
+def generate_samples(samples, labelled, classes, nfeatures):
+  def generate_sample():
+    label = random.choice(classes)
+    nf    = random.choice(nfeatures)
+    features = []
+    for i in xrange(nf):
+      feature = random.choice(FEATURES.keys())
+      findex  = max(0, bisect.bisect_left( PROB[label, feature], (random.random(), ''))-1)
+      fvalue  = PROB[label, feature][findex][1]
+      features.append((feature, fvalue))
+    if labelled:
+      return label, features
+    else:
+      return features
+      
+  for i in xrange(samples):
+    yield generate_sample()
+
+def exact_conditional_prob(sample, known=KNOWN_CLASSES):
+  p = 0.0
+  for label in known:
+    p_given_label = 1.0
+    for fid, fvalue in sample:
+      p_given_label *= POTENTIAL[label, fid, fvalue]
+    p += p_given_label * 1.0 / len(KNOWN_CLASSES)
+  return p
 
 def exact_unconditional_prob(sample):
-  raise NotImplemented
+  return exact_conditional_prob(sample, ALL_CLASSES)
 
-known_data = """"""
[…]

### 2011-05-15 semantic-nd/results/ontology.py
Changed build to run the script to generate the graphs for the paper.

@@ -1,0 +1,124 @@
+#!/usr/bin/python3
+# -*- coding: utf-8 -*-
+#
+# Copyright 2011: André Susano Pinto
+#
+# This program is free software; you can redistribute it and/or
+# modify it under the terms of the GNU General Public License
+# as published by the Free Software Foundation; either version 2
+# of the License, or (at your option) any later version.
+#
+# This program is distributed in the hope that it will be useful,
+# but WITHOUT ANY WARRANTY; without even the implied warranty of
+# MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+# GNU General Public License for more details.
+#
+# You should have received a copy of the GNU General Public License
+# along with this program; if not, write to the Free Software Foundation,
+# Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+#
+# This module loads the ontology and creates distributions defined by:
+# default_knowledge-semmap.xml and defaultprobs-semmap.txt
+#
+import script
+import re
+import math
+import dataset
+from collections import defaultdict
+
+POISSON_EXPANSION = 3
+DEFAULT_KNOWLEDGE = script.option('ONTOLOGY_DEFAULT_KNOWLEDGE')
+DEFAULT_OBJ_PROBS = script.option('ONTOLOGY_DEFAULT_OBJ_PROBS')
+DROP_FEATURES     = script.option('ONTOLOGY_DROP_FEATURES').split()
+
+
+potential = dict()
+roomcats  = set()
+features  = set()
+feature_space = defaultdict(set)
+
+def Load():
+  print("Loading default knowledge...")
+  for line in open(DEFAULT_KNOWLEDGE):
+    match = re.match('^\s+<item\s+room_category1="(?P<roomcat>\w+)"\s+' + 
+                     '(?P<proptype>\w+)="(?P<propval>\w+)"\s+' + 
+                     '(?:potential|probability)="(?P<potential>\S+)"\s*\/>\s*$', line)
+    if match:
+      roomcat, feature, descriptor, prob = match.groups()
+      potential[roomcat, feature, descriptor] = float(prob)
+      roomcats.add(roomcat)
+      features.add(feature)
+      feature_space[feature].add(descriptor)
+    elif line.strip() == '':
+      pass
+    else:
+      print "Ignored line: ", line,
+
+  
+  print('Loading object probabilities...')
+  for line in open(DEFAULT_OBJ_PROBS):
[…]

### 2011-05-16 semantic-nd/article.tex
Several changes on the results section.
That section should now be very close to final status.
Apart from the subsection titles.

@@
 % first the title is needed
-\title{Using Graphical Models for Novelty Detection on Semantic Room Classification}
+\title{Novelty Detection Using Graphical Models for Semantic Room Classification}
 
@@
 \begin{abstract}
-This papers presents an approach on implementing novelty detection for indoor room classification
-using semantic data.
-\emph{Graphical models} are used to model probabilistic knowledge and a novelty threshold is
-defined in terms of conditional and unconditional probabilities.
-The novelty threshold is then optimized using an unconditional probability density
-model trained from unlabelled data.
+This paper presents an approach to the problem of novelty
+detection in the context of semantic room categorization.
+The ability to assign semantic labels to areas in the environment is crucial for
+autonomous agents aiming to perform complex human-like tasks and human
+interaction.
+However, in order to be robust and naturally learn the semantics from
+the human user, the agent must be able to identify gaps in its own knowledge.
+To this end, we propose a method based on graphical models to identify novel
+input which does not match any of the previously learnt semantic descriptions.
+The method employs a novelty threshold defined in terms of conditional
+and unconditional probabilities. The novelty threshold is then optimized using
+an unconditional probability density model trained from unlabelled data.
 
@@
 
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[node distance = 2cm, auto]
-    \matrix[row sep=0.5cm,column sep=0.5cm] {
-      & \node [var] (room) {Room Category}; \\
-      \node [factor] (fAppearance) {}; &
-      \node [factor] (fShape) {}; &
-      \node [factor] (fSize) {}; \\
-
-      \node [var] (appearance) {Appearance}; &
-      \node [var] (shape) {Shape}; &
-      \node [var] (size) {Size}; \\
-    };
-  \path [line] (room) -- (fAppearance);
-  \path [line] (room) -- (fShape);
-  \path [line] (room) -- (fSize);
-
-  \path [line] (fAppearance) -- (appearance);
-  \path [line] (fShape) -- (shape);
-  \path [line] (fSize) -- (size);
-\end{tikzpicture}
-\caption{Example of a factor graph used for Dora for representing the environment.}
-\end{figure}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 % Introduce novelty detection
-Novelty detection is an harder problem than that of classification due to its single-class nature.
-Essentially a system is given training data and when given new data has to decide if it belongs
-to the same concept it was trained with.
+Novelty detection deals with detecting that a new input was generated by a class
+other than the ones the system knows about.
+It is harder than classification as only positive samples of a class are available
+rendering normal classification methods invalid.
 
 % Explain threshold approach to novelty detection
-Due to the desire of robustness and generalization, a novelty detection system would always possess
-some threshold that describes on how strict the system should be to new data.
-This threshold is often seen as a distance measure to a normal input and several novelty detection
-techniques are based on it.
+Due to the desire of robustness and generalization, a novelty detection system would
+always possess some threshold that describes on how strict the system should be to
+new data.
+This threshold describes a trade of between generalization and certainty.
+And is often seen as a distance measure to the known class.
 
@@
 Using bayes rule on the reversed order $P(\overline{novel}|x)$ and assuming a constant $P(\overline{novel})$
-we obtain a ratio between a conditional and unconditional probabilities of the input $x$.
+a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
 Such a ratio is a suitable function for implementing an optimal novelty detector system with
@@
 % Explain approach of using conditional probability
-\subsection{Conditional Probability: $P(x|\overline{novel})$}
-\label{sec:uniform-unconditional}
-
-One of the hardest problems is that on most cases a novelty detection system only has access to 
-the known data.
-Under that the best approach is to define a threshold based on $P(x|\overline{novel})/P(x)$
-considering that $P(x)$ is constant through all the input space.
+\subsection{Conditional Probability}
+\label{sec:conditional-prob}
 
@@
 produced by the conceptual layer.
-The \autoref{fig:conditional-prob-graph} represents a possible graph $G$ built from the conceptual
+\autoref{fig:conditional-prob-graph} represents a possible graph $G$ built from the conceptual
 layer to represent the conditional probability on the sensed variables $x$.
@@
 
-Its important to notice that in several cases assuming the unconditional probability $P(x)$
-is constant leads to discarding the factor.
-Nonetheless, here the distributions are dynamically changing as the system learns more on the environment.
+\subsection{Unconditional Probability}
+\label{sec:unconditional-prob}
+One of the hardest problems is that on most cases a novelty detection system only
+has access to the known data.
+Under that the best approach is to define a threshold assuming that $P(x)$ is
+constant through all the input space.
+
+Its important to notice that in several cases assuming it to be constant leads to
+discarding the factor.
+Nonetheless, here the distributions are dynamically changing as the system learns
+more on the environment.
 So the normalizing argument $P(x)$ has to be evaluated for each new subset of $x$.
 
-Assuming that the unconditional distributions generates all possible outcome with the same probability
-we can model it with $\prod{1/\# x_i}$, where $\# x_i$ denotes the cardinality of the state space
-of variable $x_i$.
-In graphical model terms this is represented to a factor graph $U$ with the variables but
-without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
+Assuming that the unconditional distributions generates all possible outcome with
+the same probability we can model it with $\prod{1/\# x_i}$,
+where $\# x_i$ denotes the cardinality of the state space of variable $x_i$.
+In graphical model terms this is represented to a factor graph $U$ with the
+variables but without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
 
@@
 
-Having now a graphical model built to model the known data distribution and assuming that in the unconditional case all
-possible outcomes are equally likely a novelty threshold would be given by: $P_G(x)/P_U(x)$.
-
+Having a graphical model $G$ built to model the known data distribution and a model
+$U$ for the unconditional probability a novelty threshold would be given by:
+$P_G(x)/P_U(x)$.
 Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of
@@
 
-\subsection{Using Unlabelled Data}
+\subsection{Semi Supervised: Using Unlabelled Data}
 Nonetheless its often the case that there is access to large amounts of unlabelled data.
 Under that it becomes possible to obtain a better approach to the unconditional probability
-distribution than the uniform one assumed in \autoref{sec:uniform-unconditional}.
+distribution than the uniform one.
 
@@
 \section{Results}
-In order to verify the improvements obtained by using unlabelled data, a synthetic dataset
-was generated.
+In order to verify the performance of the proposed thresholds a synthetic dataset
+was generated. To keep it simple only information regarding direct properties of
+a room were modelled and no structured knowledge such as room connectivity was taken
+in account.
 The synthetic distribution assumes that an independent and variable size set of features
 $x$ is generated by a given room category.
-In whole there was 11 different room categories and 7 different measured properties types.
+In whole there was 11 different room categories and 7 different measured feature
+types. Each feature type can be present more than once in (Eg.: room shape is
+extracted from 2D laser scans in more than one position in the room).
 
-The known data was extracted by generating samples from 6 of the 11 classes.
-Together with the generated known data, the room category that generated is given as label.
-That known data is used to train factors to represent a conditional distribution.
-Using unlabelled data, factors are trained to approximate an unconditional distribution
-where all features are independent.
-\autoref{fig:simple-experiment} shows the graph structure used for approximate the conditional and
-unconditional distributions trained from the labelled and unlabelled data.
+The room categories were chosen to mimic as close as possible real features and
+categories existent on reality. And they model different room categories with
+different levels of detail. For example: 1 person office, 2 person office, hallway,
+robot lab.
+
+From the distribution, 100 labelled samples for 5 of the 11 room categories were
+drawn to represent the known concepts and 1000 unlabelled samples were drawn from
+all the room categories for learning the unconditional probability distribution.
+Using those samples, factors were learnt for the graphs used to model the
+conditional distribution and the independent unconditional distribution.
+\autoref{fig:simple-experiment} shows the graph structure used for approximate the
+trained conditional and unconditional distributions.
+Its important to notice that graph $G$ used to model the known classes perfectly
+mimics the real distribution, with the exception that the system is not aware of
+all the room categories.
 
@@
 
-\subfloat[Graph structure used to model conditional probability graph.
-          The graph factors are trained using the labelled data.]{... conditional probability}
+\subfloat[Graph structure $G$ used to model conditional probability graph.]{... conditional probability}
 \qquad
-\subfloat[Graph structure used to model unconditional probability graph.
-          The graph factors are trained using the unlabelled data.]{... unconditional probability}
+\subfloat[Graph structure $U$ used to model unconditional probability graph.]{... unconditional probability}
+\qquad
+\subfloat[Graph structure $I$ used to model unconditional probability graph.]{... unconditional probability}
 
-\caption{\label{fig:simple-experiment}The graph structures $G$ and $I$ used to model the
-         conditional and unconditional probability.
-         Note that graph $G$ correctly models the structure of the synthetic distribution, though
-         during training it has not seen all possible room categories represented by variable $R_1$.}
-
+\caption{\label{fig:simple-experiment}The graph structures used to model the
+         conditional and unconditional probability for implementing the novelty
+         thresholds $P_G(x)/P_U(x)$ and $P_G(x)/P_I(x)$.}
 \end{figure}
 
-\begin{figure}[h!]
+% Describe how to obtain the 3 thresholds functions seen on the graphs.
+Using the learned models $G$, $U$ and $I$ two thresholds were trained:
+$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
+and $P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
+Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
+and so a perfect threshold function could also be created to test how far the
+presented thresholds are from optimal.
+
+\subsection{Suitable functions for implementing a Static Threshold}
+%%% Results 1
+% Show the threshold ratio is an optimal detector (if perfect information was available)
+% Show that the thresholds are suitable functions for implementing a static threshold.
+As first tryout, the performance of the novelty thresholds was plotted for a set
+of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
+Those samples contain different number of sensed properties for each room, miming
+the dynamic properties expected to see when implemented on a robot.
+
+\begin{figure}[h]
 \centering
@@
 
-\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance between an optimal detector,
-         $P_G(x)/P_U(x)$ a graph ratio considering an uniform unconditional distribution and
-         $P_G(x)/P_I(x)$ a graph ratio considering an independent unconditional distribution}
+\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
+         under samples with variable size of sensed properties.}
 \end{figure}
 
-Since the distribution is synthetic its possible to create an optimal novelty detector, as
-its the error rate $P(novel|x)$ can be exactly measured.
-That serves as a comparison on how optimal a detector can be using the same input space.
-A ROC curve was plotted for the three methods on \autoref{fig:synthentic-roc}.
-As possible to see the unconditional probability plays an important role in obtaining a correct
-ordering of the input space.
+The convex shape of the optimal threshold shows that the ratio between conditional
+and unconditional probability is indeed an optimal detector and is suitable for
+implementing a static threshold when the samples are taken from dynamic
+distributions (eg.: some samples where there is only access to room size versus
+samples where there is a lot of information on the room properties).
 
-The importance of a correct estimate of unconditional probability is expected to decrease as the
-input space increases and allows the several existing classes to become easily distinguished.
+% Discuss importance on approximating unconditional probability.
+Its also possible to see how important it is to estimating a correct unconditional
+probability in order to obtain a correct novelty ordering on the inputs.
+The assumption of an uniform unconditional probability has led to very poor results.
+That is probably explained by semantic properties as the ones modelled to be highly
+biased towards some values and understanding that bias plays an important step
+to detect whether a given sensed value is a valuable clue on the room category
+or not.
 
-For comparison ROC curves where plot for the developed threshold based on the number of input features.
+%%% Results 2
+% Measure performance of the thresholds as more information becomes available.
+\subsection{Performance impact as more information becomes available}
+In order to measure the performance impact as more semantic information becomes
+available ROC curves were plotted for samples grouped by the number of sensed
+semantic properties.
 
@@
          presented novelty detection method on graphs generated for different amount of
-         sensed variables.
+         sensed variables.}
+\end{figure}
+
 As possible to see as the system gains more semantic information it becomes easier
-         to detect novelty and the effect of picking an approximation for the unconditional
-         probability decreases.}
-\end{figure}
+to detect novelty as the input space size increases and allows the several existing
+classes to become more easily distinguished.
+
+The performance of the independent threshold decreases as the number of sensed
+variables increases. This is easily explained by the fact that the graph $I$ is not
+able to model the existent dependence between the variables and that becomes obvious
+as the number of variables increases (Eg.: graph $I$ perfectly models $P(x)$ in the
+case where only 1 variable is present).
+
+The uniform threshold shows a very poor performance specially on small size samples
+where its performs almost no better than random.
+It performance increases as the size of sensed variables increases but nonetheless
+its always very small when compared to how optimal a threshold could be.
+This shows that at least under semantic distributions as the one modelled with the
+synthetic data assuming an uniform unconditional distribution is a very bad choice.
 

### 2011-05-16 semantic-nd/results/synthetic.py

@@ -39,7 +39,7 @@ ALL_CLASSES = ontology.rooms.keys()
 POTENTIAL   = ontology.potential
 PROB        = dict()
 
-KNOWN_CLASSES = """singleoffice kitchen robotlab""".split()
+KNOWN_CLASSES = """bathroom hallway kitchen robotlab singleoffice""".split()
 for label in KNOWN_CLASSES: assert label in ALL_CLASSES
 
 """Make potentials normalized. And calculate probabilities for the synthetic distribution."""
@@ -171,14 +171,14 @@ def roc_performance(detector, data, known_labels):
     
 
 
-for output, nfeatures in [('synthetic-all.pdf', [3, 5, 7, 10, 15, 20, 25, 35, 50]),
+for output, nfeatures in [('synthetic-all.pdf', [5, 10, 15, 20, 35, 50]),
                           ('synthetic-3features.pdf',  [3]),
                           ('synthetic-5features.pdf',  [5]),
                           ('synthetic-10features.pdf', [10]),
                           ('synthetic-50features.pdf', [50])]:
   print "Generating ", output
   test_samples = list(generate_samples(classes = ALL_CLASSES,
-                                  samples = 1000,
+                                  samples = 5000,
                                   labelled = True,
                                   nfeatures = nfeatures))
   R1 = roc_performance(exact_novelty_detector,       test_samples, KNOWN_CLASSES)

### 2011-05-16 semantic-nd/article.tex
Changes in Results sections.
Key points have been summarized and some directions are given on future work.

@@
 \section{Conclusions and Future Work}
+On this paper it was presented how to define a stable novelty threshold function on
+top of \emph{probabilistic graphical models} instantiated dynamically from sensed
+semantic data.
+The presented technique is based on the ratio between a conditional and
+unconditional probability and when perfect information exists it performs an optimal
+novelty detection threshold.
 
-On this paper it was presented how to define a stable novelty threshold function on top of
-\emph{probabilistic graphical models} instantiated dynamically from sensed semantic data and
-conceptual knowledge.
-The presented technique is based on the ratio between a conditional and unconditional probability.
+It was also shown that a correct estimation of unconditional probability plays an
+important role specially on small input spaces. And that semi-supervised techniques
+implemented with the access to unlabelled data can be used to significantly improve
+novelty detection performance.
 
-It was shown that unlabelled data can be used to approximate unconditional probability and
-improve a novelty threshold.
-And that a correct estimation of unconditional probability plays an important role specially on
-small input spaces.
+On the used synthetic distribution, an assumption on an uniform
+distribution has led to very poor results. This same behaviour is expected to held
+in real world distributions also based on semantic data. And for that reason
+and due to easy access to unlabelled data special attention will be given on using
+semi-supervised techniques for novelty detection.
 
-Future will work will focus on how to explain why a class is novel and what makes it different
-other classes.
-That will lead to generation of useful information that can be used on communication with the user
-and perform active learning of new room categories.
+After this initial study on how to detect new semantic classes based on
+\emph{graphical models} future work will focus on how to use the structured
+information available from the conceptual layer and be able to detect which variable
+of the graph is novel and what makes it different from other previously learned
+classes. That will lead to generation of useful information that can be used on
+communication with the user and perform active learning of new room categories.
 

### 2011-05-16 semantic-nd/article.tex
Added images and some texts fixed to make it more coeherent.
Eg.: call inputs as sensed features all the time.

@@
 
-A \emph{factor graph} $G$ is a bipartite graph connecting a set of nodes $V_G$ representing
+A \emph{factor graph} is a bipartite graph connecting a set of nodes representing
 random variables and factors $\phi_G$.
 Each factor represents a function dependent only on the variables to where its connected.
-As so given factor graph $G$ can be seen as description of a distribution function over a set
+As so given factor graph can be seen as description of a distribution function over a set
 of variables obtained by the product of all the factors. In order to represent a probability
-a normalization factor $Z$ needs to be introduced:
+a normalization factor needs to be introduced:
 
 \begin{equation}
-P_G(X_V) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
+P_G(x) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
 \end{equation}
 
-Describing the distribution function in terms of graphs allows to use algorithms such as the
-Sum-Product to efficiently calculate marginals on any given subset $x$ of variables by
-exploiting conditional independency between variables.
+Describing the distribution function in terms of graphs allows to use algorithms
+such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
+marginals on any given subset $x$ of variables by exploiting conditional
+independence between variables.
 
@@
 Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about.
+other than the ones the system knows about \citep{markou2003novelty}.
 It is harder than classification as only positive samples of a class are available
@@
 
-\subfloat[Graph structure $G$ used to model conditional probability graph.]{... conditional probability}
+\subfloat[Graph structure $G$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-cond-graph.pdf}}
 \qquad
-\subfloat[Graph structure $U$ used to model unconditional probability graph.]{... unconditional probability}
+\subfloat[Graph structure $U$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-uniform-graph.pdf}}
 \qquad
-\subfloat[Graph structure $I$ used to model unconditional probability graph.]{... unconditional probability}
+\subfloat[Graph structure $I$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-independent-graph.pdf}}
 
@@
 
-\subfloat[3 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-3features.pdf}}
+\subfloat[3 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-3features.pdf}}
 \qquad
-\subfloat[5 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-5features.pdf}}
+\subfloat[5 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-5features.pdf}}
 
-\subfloat[10 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-10features.pdf}}
+\subfloat[10 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-10features.pdf}}
 \qquad
-\subfloat[50 sensed variables]{\includegraphics[width=0.40\textwidth]{results/synthetic-50features.pdf}}
+\subfloat[50 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-50features.pdf}}
 
@@
          presented novelty detection method on graphs generated for different amount of
-         sensed variables.}
+         sensed features.}
 \end{figure}
@@
 The performance of the independent threshold decreases as the number of sensed
-variables increases. This is easily explained by the fact that the graph $I$ is not
-able to model the existent dependence between the variables and that becomes obvious
-as the number of variables increases (Eg.: graph $I$ perfectly models $P(x)$ in the
-case where only 1 variable is present).
+features increases. This is easily explained by the fact that the graph $I$ is not
+able to model the existent dependence between the features and that becomes obvious
+as the number of features increases (Eg.: graph $I$ perfectly models $P(x)$ in the
+case where only 1 feature is sensed).
 
@@
 where its performs almost no better than random.
-It performance increases as the size of sensed variables increases but nonetheless
+It performance increases as the size of sensed features increases but nonetheless
 its always very small when compared to how optimal a threshold could be.

### 2011-05-16 semantic-nd/article.tex
Changes in introduction.

@@
 \section{Introduction}
-There has been several efforts in the area of Artificial Intelligence and Robotics in creating
-robots that are able to interact with humans and their environments.
-One of the existing problems is a reliable high-level localization method that can be deployed
-into new and unknown environments.
-This article focus on the mapping, using \emph{semantic data}, of indoor environments such as
-houses and offices to room categories such as kitchen, corridor, office.
-And how to perform novelty detection on them: \emph{detect a new room category that was not
-present in the labelled data}.
+There has been several efforts in the area of Artificial Intelligence and Robotics
+in creating robots that are able to interact with humans and their environments.
+One of the existing problems is a reliable high-level localization method that can
+be deployed into new and unknown environments or where its impossible to access a
+previously labelled map of the area. Nonetheless the need of reliability and ability
+to deal with new cases also arises due to dynamic changes on the environment through
+time and with human action.
+This article focus on the mapping, using \emph{semantic data}, of indoor
+environments such as houses and offices to room categories such as kitchen,
+corridor, office.
 
-Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the presented novelty
-detection system.
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
-That model connects the sensed properties together with the variables used to model the world.
-And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box\cite{exploiting}.
+Due to unknown and dynamic nature of the operating environment of the agent the
+mapping methods should be resilient to new room categories it has never seen
+before. This is: perform novelty detection on the concepts the agent is aware of.
 
-The ability to recognize novel categories on the modelled variables would allow to increase
-reliability and allow the creation of self-extending behaviours.
+Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the
+presented novelty detection system as it already possesses a developed architecture
+oriented on dealing with probabilistic and semantic information and performing
+reasoning over it.
+Implementing novelty detection capabilities on an agent such as Dora, would allow to
+to increase reliability and allow the creation of self-extending behaviours.
+
+Its with a final interest on integrated the developed methods on Dora that this
+initial work focus on the novelty detection using probabilistic graphical models
+that model the conceptual knowledge and information handled by Dora.
 
@@
 Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
+
+As Dora moves through the environment its \emph{conceptual layer} builds a structural and
+probabilistic representation of the space instantiated as a \emph{graphical model}.
+That model connects the sensed properties together with the variables used to model the world.
+And allows to perform queries on the probabilities of aspects of the environment.
+For example: where is most likely to find a cereal box\cite{exploiting}.
 

### 2011-05-16 semantic-nd/article.tex
Changes on Related Work section

@@
 
-Its with a final interest on integrated the developed methods on Dora that this
+Its with a final interest on integrating the developed methods on Dora that this
 initial work focus on the novelty detection using probabilistic graphical models
@@
 \section{Related Work}
-\cite{quattoni2009recognizing} showed that most scene recognition models work poorly in indoor
-scenes when compared to outdoor scenes results.
-Since the properties that characterize rooms changes conforming its category. Namely corridors are
-well described by global properties and bookstores are well described by the presence of specific objects (books).
-Their work shows a multi-modal approach is expected to yield better performance by mixing several sources of information.
+\cite{quattoni2009recognizing} showed that most scene recognition models work poorly
+in indoor scenes when compared to outdoor scenes results since the properties that
+characterize rooms changes conforming its category. Namely corridors are well
+described by global properties and bookstores are well described by the presence of
+specific objects (books). Their work shows a multi-modal approach is expected to
+yield better performance by mixing several sources of information.
 
-\cite{galindo2005multi} defines a bidirectional relation between object and room category, where object defines a room category and a room category provides information on where objects may be found.
+\cite{galindo2005multi} defines a bidirectional relation between object and room
+category, where object defines a room category and a room category provides
+information on where objects may be found.
 
-Probabilistic representations are used in several localised functions in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
-And some employ, up to some extent, a probabilistic representation across some subsystems~\cite{kraft2008exploration}.
-\cite{vasudevan2008bayesian} performed room categorization through Bayesian reasoning about the presence of objects but did not included observations models (perception is considered deterministic).
-And \cite{boutell2006factor} have studied outdoor scene classification using \emph{factor graphs} and modelling spatial relations between objects in the scene to extract better knowledge from semantic (high-level) features.
+Probabilistic representations are used in several localised functions in robots
+operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
+And some employ, up to some extent, a probabilistic representation across some
+subsystems~\cite{kraft2008exploration}.
+And \cite{boutell2006factor} have studied outdoor scene classification using
+\emph{factor graphs} and modelling spatial relations between objects in the scene
+to extract better knowledge from semantic (high-level) features.
 
-Its expected that using a unified probabilistic model from the whole system, such as \cite{pronobis2011exploiting}, more information can be reused to correctly predict a given random variable.
+% Its expected that using a unified probabilistic model from the whole system,
+% such as \cite{pronobis2011exploiting}, more information can be reused to correctly
+% predict a given random variable.
 
-Although this paper deals with novelty detection on the robotics area, it does so using very
-standard concepts and techniques such as semantic data and graphical models.
+Although this paper deals with novelty detection on the robotics area, it does so
+using standard concepts and techniques such as semantic data and graphical models.
 Those are often found on areas related with information retrieval.
-Interesting examples are works in automatic image annotation using an hidden concept layer between visual features and text information\cite{zhang2005probabilistic}.
+An interesting example is the work in automatic image annotation using an hidden
+concept layer between visual features and text information by \cite{zhang2005probabilistic}.
 
@@
 \subsection{Factor Graphs}
-Although the conceptual layer works with \emph{chain graphs}, those can be converted into \emph{factor graphs}.
-Those are used through this paper as they provide an easier manipulation in terms of factorization.
+Although the conceptual layer works with \emph{chain graphs}, those can be converted
+into \emph{factor graphs}. Those are used through this paper as they provide an
+easier manipulation in terms of factorization.
 
 A \emph{factor graph} is a bipartite graph connecting a set of nodes representing
-random variables and factors $\phi_G$.
+random variables and factors.
 Each factor represents a function dependent only on the variables to where its connected.
@@
 \begin{equation}
-P_G(x) = \frac{1}{Z}\prod_{}{\phi(...)},\qquad Z = \sum\prod{\phi(...)}
+P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
+Z = \sum_{x \in X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
 \end{equation}
@@
 such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
-marginals on any given subset $x$ of variables by exploiting conditional
+marginals on any given subset $x_m$ of variables by exploiting conditional
 independence between variables.
-
-\begin{equation}
-P_G(x) = \frac{1}{Z}\sum_{V \\ x}{\prod{\phi(...)}}
-\end{equation}
 

### 2011-05-16 semantic-nd/article.tex
Big changes on the related work area.
Now it shows related work on the area of using semantic data and graphical
models to model room classification and shows some insights on work performed
on novelty detection area.

@@
 
-Due to unknown and dynamic nature of the operating environment of the agent the
-mapping methods should be resilient to new room categories it has never seen
+Due to the unknown and dynamic nature of the operating environment the
+mapping methods should be resilient to new room categories the agent has never seen
 before. This is: perform novelty detection on the concepts the agent is aware of.
@@
 \section{Related Work}
-\cite{quattoni2009recognizing} showed that most scene recognition models work poorly
-in indoor scenes when compared to outdoor scenes results since the properties that
+The usage of multi-modal approaches, such as combining semantic data extracted from
+several sources or classifiers is expected to have better performance on scene
+recognition than single approaches according to \cite{quattoni2009recognizing}
+that showed that most scene recognition models work poorly in indoor scenes when
+compared to outdoor scenes results since the properties that
 characterize rooms changes conforming its category. Namely corridors are well
 described by global properties and bookstores are well described by the presence of
-specific objects (books). Their work shows a multi-modal approach is expected to
-yield better performance by mixing several sources of information.
+specific objects (books).
+\cite{galindo2005multi} also exploits this by defining a bidirectional relation
+between object and room category, where object defines a room category and a room
+category provides information on where objects may be found.
 
-\cite{galindo2005multi} defines a bidirectional relation between object and room
-category, where object defines a room category and a room category provides
-information on where objects may be found.
-
-Probabilistic representations are used in several localised functions in robots
-operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
+Probabilistic representations have also been used in several localised functions
+in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
 And some employ, up to some extent, a probabilistic representation across some
@@
 
-% Its expected that using a unified probabilistic model from the whole system,
-% such as \cite{pronobis2011exploiting}, more information can be reused to correctly
-% predict a given random variable.
-
-Although this paper deals with novelty detection on the robotics area, it does so
+Also although this paper deals with robotics area, it does so
 using standard concepts and techniques such as semantic data and graphical models.
@@
 concept layer between visual features and text information by \cite{zhang2005probabilistic}.
+
+Novelty detection has been studied for some years and there is several approaches
+based on statistical analysis as reported by \cite{markou2003novelty}.
+Graphical models have been used to learn distributions of variables, both in
+supervised and unsupervised ways and by using thresholds on those distributions
+based solely on the conditional probability as seen on \cite{bishop1994novelty}
+a novelty system can be trivially implemented.
+But as far as we known there is no reference on how to perform novelty detection
+by using graphs that are dynamically generated.
 
@@
 new data.
-This threshold describes a trade of between generalization and certainty.
+This threshold describes a trade between error and rejection.
 And is often seen as a distance measure to the known class.

### 2011-05-16 semantic-nd/article.tex
small changes (including some subtitles)

@@
 such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
-marginals on any given subset $x_m$ of variables by exploiting conditional
+marginals on any given subset of variables by exploiting conditional
 independence between variables.
@@
 
-\subsection{Suitable functions for implementing a Static Threshold}
+\subsection{Probability ratio comparison}
 %%% Results 1
@@
 % Measure performance of the thresholds as more information becomes available.
-\subsection{Performance impact as more information becomes available}
+\subsection{Performance changes with amount of available information}
 In order to measure the performance impact as more semantic information becomes
 available ROC curves were plotted for samples grouped by the number of sensed
-semantic properties.
+semantic features.
 
@@
 This shows that at least under semantic distributions as the one modelled with the
-synthetic data assuming an uniform unconditional distribution is a very bad choice.
+synthetic data assuming an uniform unconditional distribution leads to poor results.
 

### 2011-05-17 scripts/graph/ROC.py
changed roc marker size

@@ -19,7 +19,7 @@ def ROC(data, style = 'ro-', label = None):
   if fpr[-1] != 0: fpr /= fpr[-1]
   if tpr[-1] != 0: tpr /= tpr[-1]
   
-  P.plot(fpr, tpr, style, markevery=(len(fpr)/10, len(fpr)/10), label = label)
+  P.plot(fpr, tpr, style, markersize=10.0, markevery=(len(fpr)/10, len(fpr)/10), label = label)
   P.xlim([-0.05,1.05])
   P.ylim([-0.05,1.05])
   P.xlabel('False Positive Rate')

### 2011-05-17 semantic-nd/article.tex
Some of andrzej comments fixed until the novelty detection section.

@@
 \section{Related Work}
-The usage of multi-modal approaches, such as combining semantic data extracted from
-several sources or classifiers is expected to have better performance on scene
-recognition than single approaches according to \cite{quattoni2009recognizing}
-that showed that most scene recognition models work poorly in indoor scenes when
-compared to outdoor scenes results since the properties that
-characterize rooms changes conforming its category. Namely corridors are well
+Multi-modal approaches, such as combining semantic data extracted from
+several sources or classifiers are expected to have better performance on scene
+recognition than single-cue approaches. \cite{quattoni2009recognizing}
+showed that most scene recognition models work poorly in indoor scenes when
+compared to outdoor scenes since the properties that
+characterize rooms changes depending on the category. For instance corridors are well
 described by global properties and bookstores are well described by the presence of
@@
 
-Probabilistic representations have also been used in several localised functions
+Probabilistic representations have been frequently used for spatial modelling
 in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
-And some employ, up to some extent, a probabilistic representation across some
+FIX/REMOVE: And some employ, up to certain extent, a probabilistic representation across some
 subsystems~\cite{kraft2008exploration}.
-And \cite{boutell2006factor} have studied outdoor scene classification using
+\cite{boutell2006factor} have studied outdoor scene classification using
 \emph{factor graphs} and modelling spatial relations between objects in the scene
@@
 
-Also although this paper deals with robotics area, it does so
-using standard concepts and techniques such as semantic data and graphical models.
-Those are often found on areas related with information retrieval.
+Although our approach is presented in the context of mobile robotics it relies on
+standard concepts and techniques such as semantic data and graphical models.
+Those are often used in the area of information retrieval.
 An interesting example is the work in automatic image annotation using an hidden
@@
 
-Novelty detection has been studied for some years and there is several approaches
+Novelty detection has been studied for many years and there are several approaches
 based on statistical analysis as reported by \cite{markou2003novelty}.
@@
 a novelty system can be trivially implemented.
-But as far as we known there is no reference on how to perform novelty detection
-by using graphs that are dynamically generated.
+However to the knowledge of the authors there is no reference on how to perform
+novelty detection using graphs that are dynamically generated.
 
@@
 Although the conceptual layer works with \emph{chain graphs}, those can be converted
-into \emph{factor graphs}. Those are used through this paper as they provide an
-easier manipulation in terms of factorization.
-
-A \emph{factor graph} is a bipartite graph connecting a set of nodes representing
-random variables and factors.
-Each factor represents a function dependent only on the variables to where its connected.
-As so given factor graph can be seen as description of a distribution function over a set
-of variables obtained by the product of all the factors. In order to represent a probability
-a normalization factor needs to be introduced:
-
-\begin{equation}
-P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
-Z = \sum_{x \in X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
-\end{equation}
+into \emph{factor graphs}. Factor graphs are used throughout this paper as they provide an
+easier manipulation due to factorization.
+Moreover, there exists efficient implementation of inference engines operating on factor graph
+representations \cite{Mooij_libDAI_10}.
 
@@
 independence between variables.
+
+A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
+representing random variables and factors.
+Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
+to which the factor is connected.
+Thus, a factor graph can be seen as a description of probability density function obtained
+by a product of all the factors. In order to represent the probability,
+a normalization factor needs to be introduced, resulting in the following equation
+defining the distribution:
+
+\begin{equation}
+P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
+Z = \sum_{X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
+\end{equation}
 

### 2011-05-17 semantic-nd/article.tex
Fixed citting style to LNCS.

@@
 \usepackage[utf8]{inputenc}
-\usepackage[round]{natbib}
 \usepackage{subfig}
-\usepackage{tikz}
 \usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
@@
 }
-
-% TiKZ styles
-\tikzstyle{var} = [circle,
-                   thick,
-                   draw, fill=red!20,
-                   font=\scriptsize,
-                   text width=3.5em,
-                   text centered,
-                   node distance=1em,
-                   inner sep=0pt]
-
-\tikzstyle{factor} = [rectangle,
-                      thick,
-                      draw, fill=blue!20,
-                      minimum height=2em,
-                      minimum width=2em]
-
-\tikzstyle{line} = [draw]
 
@@
 several sources or classifiers are expected to have better performance on scene
-recognition than single-cue approaches. \cite{quattoni2009recognizing}
+recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
 showed that most scene recognition models work poorly in indoor scenes when
@@
 specific objects (books).
-\cite{galindo2005multi} also exploits this by defining a bidirectional relation
+Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
 between object and room category, where object defines a room category and a room
@@
 Probabilistic representations have been frequently used for spatial modelling
-in robots operating in the real-world~\cite{gross2009toomas,maierprobabilistic}.
-FIX/REMOVE: And some employ, up to certain extent, a probabilistic representation across some
-subsystems~\cite{kraft2008exploration}.
-\cite{boutell2006factor} have studied outdoor scene classification using
+in robots operating in the real-world\cite{gross2009toomas,maierprobabilistic}.
+Boutell et al.\cite{boutell2006factor} have studied outdoor scene classification using
 \emph{factor graphs} and modelling spatial relations between objects in the scene
@@
 Those are often used in the area of information retrieval.
-An interesting example is the work in automatic image annotation using an hidden
-concept layer between visual features and text information by \cite{zhang2005probabilistic}.
+An interesting example is the usage of an hidden concept layer between visual features
+and text information to provide automatic image annotation\cite{zhang2005probabilistic}.
 
 Novelty detection has been studied for many years and there are several approaches
-based on statistical analysis as reported by \cite{markou2003novelty}.
+based on statistical analysis\cite{markou2003novelty}.
 Graphical models have been used to learn distributions of variables, both in
 supervised and unsupervised ways and by using thresholds on those distributions
-based solely on the conditional probability as seen on \cite{bishop1994novelty}
+based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
 a novelty system can be trivially implemented.
@@
 Although the conceptual layer works with \emph{chain graphs}, those can be converted
-into \emph{factor graphs}. Factor graphs are used throughout this paper as they provide an
+into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
 easier manipulation due to factorization.
 Moreover, there exists efficient implementation of inference engines operating on factor graph
-representations \cite{Mooij_libDAI_10}.
+representations\cite{Mooij_libDAI_10}.
 
 Describing the distribution function in terms of graphs allows to use algorithms
-such as the Sum-Product~\citep{kschischang2001factor} to efficiently calculate
+such as the Sum-Product\cite{kschischang2001factor} to efficiently calculate
 marginals on any given subset of variables by exploiting conditional
@@
 Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about \citep{markou2003novelty}.
+other than the ones the system knows about\cite{markou2003novelty}.
 It is harder than classification as only positive samples of a class are available
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{plainnat}
+\bibliographystyle{unsrt}
 \bibliography{refs}

### 2011-05-17 semantic-nd/article.tex
More small changes to everything before novelty detection section.

@@
 \section{Related Work}
+% TODO(pronobis): general comment about many .... approches to place categorization.
+
 Multi-modal approaches, such as combining semantic data extracted from
@@
 a novelty system can be trivially implemented.
+
 However to the knowledge of the authors there is no reference on how to perform
 novelty detection using graphs that are dynamically generated.
+% TODO(andresp): anything more that can be said on why this paper is useful?
 
@@
 \centering
+\subfloat[Conceptual layer]{\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}}
+\subfloat[Chain graph]{\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}}
 
-\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}
-\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}
 \caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
@@
 representations\cite{Mooij_libDAI_10}.
-
-Describing the distribution function in terms of graphs allows to use algorithms
-such as the Sum-Product\cite{kschischang2001factor} to efficiently calculate
-marginals on any given subset of variables by exploiting conditional
+Describing the distribution function in terms of graphs allows to use those engines to
+efficently calculate marginals on any given subset of variables by exploiting conditional
 independence between variables.
@@
 by a product of all the factors. In order to represent the probability,
-a normalization factor needs to be introduced, resulting in the following equation
-defining the distribution:
+a normalization factor needs to be introduced, resulting in the following equation:
 

### 2011-05-17 semantic-nd/article.tex
Changed the initial part of novelty detection to show that the threshold is optimal.

@@
 % Explain threshold approach to novelty detection
-Due to the desire of robustness and generalization, a novelty detection system would
+Due to the desire of robustness, a novelty detection system would
 always possess some threshold that describes on how strict the system should be to
-new data.
-This threshold describes a trade between error and rejection.
-And is often seen as a distance measure to the known class.
+new samples. This threshold describes a trade between error and rejection: as the
+more the system tries to reduce rejection the more prone to produce
+errors it becomes.
+This threshold if often seen as a max allowed distance between a sample and the known
+class.
 
-When dealing with a statistic point of view, noisy data or unstable features, a decision on a
-given input has an associated error rate. A novelty detection system is interested in defining an
-order on all the possible inputs equivalent to the order imposed by the error rate: $P(novel|x)$.
+When dealing with a statistic point of view, noisy data or unstable features, a decision to
+classify a sample as novel as an associated error probability of $P(\overline{novel}|x)$.
+And the error rate of a system that classifies the set of inputs $N$ as novel is given by
+the equation:
 
-Using bayes rule on the reversed order $P(\overline{novel}|x)$ and assuming a constant $P(\overline{novel})$
+\begin{equation}
+P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+\end{equation}
+
+Therefore any novelty system that decides to acknowledge a sample
+$a$ as novel should also detect novelty on any sample $b$ satisfying
+$P(\overline{novel}|b) < P(\overline{novel}|a)$.
+As that would always reduce the error probability.
+This way a novelty detection system is interested in defining an
+order on all the possible inputs equivalent to the order defined by
+the error rate: $P(\overline{novel}|x)$.
+
+Using Bayes rule and assuming a constant $P(\overline{novel})$
 a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
-Such a ratio is a suitable function for implementing an optimal novelty detector system with
+Such a ratio is a suitable function for implementing a novelty detector system with optimal
 thresholding.

### 2011-05-17 semantic-nd/article.tex
Pronobis introduction in scene categorization and using visual and laser scans to classification

@@
 % TODO(pronobis): general comment about many .... approches to place categorization.
+The problem of room categorization based on visual information was first addressed
+in the computer vision community. In this case, the research focused mainly on the
+problem of classifying single images captured in indoor or outdoor environments
+(scene classification). At the same time, robotics researchers initially employed
+the 2D laser range sensor being much more robust to variations occurring in the
+environment and much easier to handle computationally in real time.
 

### 2011-05-17 semantic-nd/article.tex
Changes in Conclusions and Future Work

@@
 \section{Conclusions and Future Work}
-On this paper it was presented how to define a stable novelty threshold function on
+In this paper we presented how to define a stable novelty threshold function on
 top of \emph{probabilistic graphical models} instantiated dynamically from sensed
@@
 It was also shown that a correct estimation of unconditional probability plays an
-important role specially on small input spaces. And that semi-supervised techniques
+important role specially on small input spaces. Moreover, semi-supervised techniques
 implemented with the access to unlabelled data can be used to significantly improve
@@
 
-On the used synthetic distribution, an assumption on an uniform
-distribution has led to very poor results. This same behaviour is expected to held
-in real world distributions also based on semantic data. And for that reason
-and due to easy access to unlabelled data special attention will be given on using
+Given the synthetic distribution, an assumption on an uniform
+distribution has led to very poor results. The same behaviour is expected to
+in real world distributions based on semantic data. For that reason,
+and due to easy access to unlabelled data, special attention will be given to using
 semi-supervised techniques for novelty detection.
@@
 After this initial study on how to detect new semantic classes based on
-\emph{graphical models} future work will focus on how to use the structured
-information available from the conceptual layer and be able to detect which variable
+graphical models, future work will focus on how to use the structured
+information available from the conceptual layer to be able to detect which variable
 of the graph is novel and what makes it different from other previously learned
-classes. That will lead to generation of useful information that can be used on
-communication with the user and perform active learning of new room categories.
+classes. That will lead to generation of useful information that can be used for
+communication with the user and performing active learning of new room categories.
 

### 2011-05-17 semantic-nd/article.tex
removed missing refs.

@@
 
-Dora\cite{dora} (CogX: Dora) was used as a base system where to implement the
+Dora (CogX: Dora) was used as a base system where to implement the
 presented novelty detection system as it already possesses a developed architecture
@@
 And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box\cite{exploiting}.
+For example: where is most likely to find a cereal box.
 

### 2011-05-17 semantic-nd/article.tex
Fixed Results section according to Andrzej comments

@@
 \section{Results}
-In order to verify the performance of the proposed thresholds a synthetic dataset
-was generated. To keep it simple only information regarding direct properties of
+In order to verify the performance of the proposed threshold functions a synthetic dataset
+was generated. To keep it simple only information regarding direct features of
 a room were modelled and no structured knowledge such as room connectivity was taken
@@
 In whole there was 11 different room categories and 7 different measured feature
-types. Each feature type can be present more than once in (Eg.: room shape is
+types. Each feature type can be present more than once in (e.g.: room shape is
 extracted from 2D laser scans in more than one position in the room).
 
-The room categories were chosen to mimic as close as possible real features and
-categories existent on reality. And they model different room categories with
+The room categories were chosen to mimic as close as possible the real features and
+categories existing in reality. And they model different room categories with
 different levels of detail. For example: 1 person office, 2 person office, hallway,
@@
 trained conditional and unconditional distributions.
-Its important to notice that graph $G$ used to model the known classes perfectly
-mimics the real distribution, with the exception that the system is not aware of
-all the room categories.
+Its important to notice that graph $G$ used to model the known classes when given
+enough labelled data is able to exactly learn the conditional distribution as it
+uses the same structure as the created synthetic distribution.
 
@@
 % Describe how to obtain the 3 thresholds functions seen on the graphs.
-Using the learned models $G$, $U$ and $I$ two thresholds were trained:
+Using the learned models $G$, $U$ and $I$, two thresholds were trained:
 $P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
@@
 Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
-and so a perfect threshold function could also be created to test how far the
+and a perfect threshold function could also be created to test how far the
 presented thresholds are from optimal.
 
-\subsection{Probability ratio comparison}
+\subsection{Probability Ratio Comparison}
 %%% Results 1
@@
 % Show that the thresholds are suitable functions for implementing a static threshold.
-As first tryout, the performance of the novelty thresholds was plotted for a set
+As first tryout, the performance of the novelty threshold selection was plotted for a set
 of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
-Those samples contain different number of sensed properties for each room, miming
+Those samples contain different number of sensed properties for each room, mimicking
 the dynamic properties expected to see when implemented on a robot.
@@
 
-The convex shape of the optimal threshold shows that the ratio between conditional
+The convex shape for the optimal threshold shows that the ratio between conditional
 and unconditional probability is indeed an optimal detector and is suitable for
-implementing a static threshold when the samples are taken from dynamic
-distributions (eg.: some samples where there is only access to room size versus
-samples where there is a lot of information on the room properties).
+implementing a threshold when the samples are taken from dynamic
+distributions (e.g.: some samples where there is only access to room size versus
+samples where there is a lot of information about the room properties).
 
 % Discuss importance on approximating unconditional probability.
-Its also possible to see how important it is to estimating a correct unconditional
-probability in order to obtain a correct novelty ordering on the inputs.
-The assumption of an uniform unconditional probability has led to very poor results.
-That is probably explained by semantic properties as the ones modelled to be highly
-biased towards some values and understanding that bias plays an important step
-to detect whether a given sensed value is a valuable clue on the room category
-or not.
+Its also possible to see how important it is to estimate a correct unconditional
+probability in order to obtain a correct novelty measure on the inputs.
+The assumption of a uniform unconditional probability has led to very poor results.
+That is probably explained by the semantic properties being highly
+biased towards some values. And shows that bias plays an important step
+in detecting whether a given sensed value is a valuable cue about the room category.
 
@@
 % Measure performance of the thresholds as more information becomes available.
-\subsection{Performance changes with amount of available information}
+\subsection{Performance Changes With Amount of Available Information}
 In order to measure the performance impact as more semantic information becomes
@@
 
-As possible to see as the system gains more semantic information it becomes easier
-to detect novelty as the input space size increases and allows the several existing
+We can see that as the system gains more semantic information it becomes easier
+to detect novelty. The input space size increases and allows the several existing
 classes to become more easily distinguished.
@@
 features increases. This is easily explained by the fact that the graph $I$ is not
-able to model the existent dependence between the features and that becomes obvious
-as the number of features increases (Eg.: graph $I$ perfectly models $P(x)$ in the
+able to model the existent dependence between the features. This becomes obvious
+as the number of features increases (e.g.: graph $I$ perfectly models $P(x)$ in the
 case where only 1 feature is sensed).
 
-The uniform threshold shows a very poor performance specially on small size samples
+The uniform threshold shows a poor performance specially on small size samples
 where its performs almost no better than random.
 It performance increases as the size of sensed features increases but nonetheless
-its always very small when compared to how optimal a threshold could be.
-This shows that at least under semantic distributions as the one modelled with the
-synthetic data assuming an uniform unconditional distribution leads to poor results.
+its very small when compared to how optimal a threshold could be.
 

### 2011-05-17 semantic-nd/article.tex
Removed authors and institutions for blind review.
Changes on explaning the order of the input space for novelty thresholding.

@@
 % the running heads and the author index.
-%
-\author{André Susano Pinto}
+\author{Removed for Blind Review}
+%\author{André Susano Pinto, Andrzej Pronobis, Luis Paulo Reis}
 %
@@
 % unless you accept that it will be published
-\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
-\url{andresusanopinto@gmail.com}\\
-\url{http://url}}
+\institute{Removed for Blind Review}
+%\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
+%\url{andresusanopinto@gmail.com}\\
+%\url{http://url}}
 
-\toctitle{Lecture Notes in Computer Science}
-\tocauthor{Authors' Instructions}
 \maketitle
@@
 
+Adding capabilities to an agent to detect a normal sample allows to increase its
+reliability. As a novelty signal can then be interpreted by the system to proceed
+with more caution as its on the presence of a situation where its knowledge
+does not correctly describes reality.
+
 % Explain threshold approach to novelty detection
-Due to the desire of robustness, a novelty detection system would
-always possess some threshold that describes on how strict the system should be to
-new samples. This threshold describes a trade between error and rejection: as the
-more the system tries to reduce rejection the more prone to produce
-errors it becomes.
-This threshold if often seen as a max allowed distance between a sample and the known
-class.
+Due to the desire of robustness, a novelty detection system is often given the option
+to reject normal samples. This is implemented as a threshold that describes a trade
+between error and rejection: the more the system tries to reduce rejection, the more
+prone to produce errors it becomes.
 
@@
 This way a novelty detection system is interested in defining an
-order on all the possible inputs equivalent to the order defined by
+order relation on all the possible inputs equivalent to the order defined by
 the error rate: $P(\overline{novel}|x)$.

### 2011-05-17 semantic-nd/article.tex
Fixed comments on unconditional probabilities sections.

@@
 
-In the presented case the conditional probability is approximated by the graphical model $G$
+In the presented case, the conditional probability is approximated by the graphical model $G$
 produced by the conceptual layer.
@@
 \subsection{Semi Supervised: Using Unlabelled Data}
-Nonetheless its often the case that there is access to large amounts of unlabelled data.
+Nonetheless, its often the case that there is access to large amounts of unlabelled data.
 Under that it becomes possible to obtain a better approach to the unconditional probability
@@
 
-For interesting and practical reasons it was assumed that the unconditional distribution
+For practical reasons it was assumed that the unconditional distribution
 could be modelled with all the variables independent from each other.
 Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-From the unlabelled data only the probability of each feature $x_i$ needs to be estimated.
+From the unlabelled data, only the probability of each feature $x_i$ needs to be estimated.
 \autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
@@
 Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
-This time the addition of the variables unconditional factors can be understood as an
+This time the addition of the unconditional factors can be understood as an
 attempt to compensate for an existing bias on the unconditional distribution.
-Which is an important step to achieve a correct ordering of the input space for novelty
-thresholding.
+This is an important step to achieve a correct order-relation of the inputs sample
+for implementing a novelty threshold.
 

### 2011-05-17 semantic-nd/article.tex
small stuff: some refs and some figure improving.

@@
 problem of classifying single images captured in indoor or outdoor environments
-(scene classification). At the same time, robotics researchers initially employed
+(scene classification)\cite{oliva2006building,torralba2003contextual}.
+At the same time, robotics researchers initially employed
 the 2D laser range sensor being much more robust to variations occurring in the
-environment and much easier to handle computationally in real time.
+environment and much easier to handle computationally in real time\cite{mozos2005supervised}.
 
@@
 Probabilistic representations have been frequently used for spatial modelling
-in robots operating in the real-world\cite{gross2009toomas,maierprobabilistic}.
+in robots operating in the real-world\cite{gross2009toomas,maier2010probabilistic}.
 Boutell et al.\cite{boutell2006factor} have studied outdoor scene classification using
@@
 produced by the conceptual layer.
-\autoref{fig:conditional-prob-graph} represents a possible graph $G$ built from the conceptual
+\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
 layer to represent the conditional probability on the sensed variables $x$.
@@
 \label{sec:unconditional-prob}
-One of the hardest problems is that on most cases a novelty detection system only
-has access to the known data.
-Under that the best approach is to define a threshold assuming that $P(x)$ is
-constant through all the input space.
+With only access to labelled data a common approach is to define a threshold assuming
+that $P(x)$ is constant through all the samples.
 

### 2011-05-17 semantic-nd/article.tex
Fixed reference.
Only missing subsection: 4.1

@@
 An interesting example is the usage of an hidden concept layer between visual features
-and text information to provide automatic image annotation\cite{zhang2005probabilistic}.
+and text information to provide automatic image annotation\cite{zhang2006probabilistic}.
 

### 2011-05-17 semantic-nd/article.tex
Rewritting of subsection 4.1: conditional probability.
It now explains why we use the factor graph produced by the conceptual layer.

@@
 
-In the presented case, the conditional probability is approximated by the graphical model $G$
-produced by the conceptual layer.
-\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
-layer to represent the conditional probability on the sensed variables $x$.
-$P_G(x)$ is used as an approximation for $P(x|\overline{novel})$.
+The conditional probability models the distribution of variables given that the sample
+is not novel for the agent.
+This is: it should model the distribution of the sensed features considering that the
+knowledge the agent possesses holds true. Under that its natural to model it with a graphical
+model that combines the information acquired with the relations that the describe the
+relations learned by the agent between the variables.
+
+In Dora case, this is equivalent to use the graphical model used by it to describe its
+current believes on the variables modelled by the system.
+That graph is built, by the conceptual layer, by instancing the information extracted from
+other layers together with the conceptual knowledge such as: objects are properties of rooms,
+rooms are connected between each other.
 
@@
 \includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
-\caption{\label{fig:conditional-prob-graph}The conceptual layer models the conditional
-         probability distribution of the sensed variables by using hidden variables.
-         Those are created from the conceptual knowledge and represent the structural
-         and connectivity information the system is aware of.
-         In this case only hidden variables $R_i$ were used to model the room categories
-         where each property was sensed.}
+\caption{\label{fig:conditional-prob-graph}A sample of a factor graph modelling a distribution
+         of a set $x$ of sensed variables.}
 \end{figure}
+
+\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
+layer to represent the conditional probability on the sensed variables $x$.
+A set of hidden variables is added to represent the conceptual knowledge the system is aware of.
+In the presented graph variables $R_i$ were added to model the room categories that influence
+the directly sensed features on each physical room, as well connectivity factors between each
+room.
+
+The factors connecting the variables are trained by the system by searching databases of
+common knowledge to build potentials describing how likely it is that a specific set of
+values for a set of variables is more likely to occur over than any other set of values.
+For instance: it is very likely to find a cereal box in a kitchen; and it is unlikely to find
+bathroom connected to another bathroom.
+
+We propose using such a graph $G$ built by the conceptual layer for modelling $P_G(x)$
+as an approximation for $P(x|\overline{novel})$.
 

### 2011-05-17 semantic-nd/article.tex
Introduction.

--- 2011-05-17 semantic-nd/article.tex
@@
 \section{Introduction}
-There has been several efforts in the area of Artificial Intelligence and Robotics
+
+There has been several efforts in the areas of artificial intelligence and robotics
 in creating robots that are able to interact with humans and their environments.
-One of the existing problems is a reliable high-level localization method that can
-be deployed into new and unknown environments or where its impossible to access a
-previously labelled map of the area. Nonetheless the need of reliability and ability
-to deal with new cases also arises due to dynamic changes on the environment through
-time and with human action.
-This article focus on the mapping, using \emph{semantic data}, of indoor
-environments such as houses and offices to room categories such as kitchen,
-corridor, office.
+One of the important aspects is to endow those robots with a deeper understanding
+of human environments, not just for the purpose of navigation and obstacle avoidance, 
+but also in terms of human semantics and functionality. An important problem
+in creating reliable representations of space for robots that are to be deployed 
+in new and unknown realistic environments is to be able to automatically 
+identify gaps in robot's knowledge and act in order to fill those gaps.
 
-Due to the unknown and dynamic nature of the operating environment the
-mapping methods should be resilient to new room categories the agent has never seen
-before. This is: perform novelty detection on the concepts the agent is aware of.
+This article addresses the problem of \emph{novelty detection} within the context
+of semantic mapping i.e. generating maps containing \emph{semantic information} 
+about indoor environments, such as homes or offices. In that context, the ability 
+to detect that the observations result from a semantic concept unknown to the robot, 
+and cannot be explained by one of its models, is crucial for generating fully autonomous
+and reliable behavior. In order to be robust, the robot must identify 
+novel concepts and instead of making a costly error, refrain from the decision
+and initiate learning. In particular, we address the problem of novelty detection for
+room categorization i.e. detecting whether the area in the environment identified as a
+separate room can be assigned one of the semantic labels that the robot knows (e.g. as \emph{a kitchen} 
+or \emph{an office}) or belongs to a yet unknown semantic category.
 
-Dora (CogX: Dora) was used as a base system where to implement the
-presented novelty detection system as it already possesses a developed architecture
-oriented on dealing with probabilistic and semantic information and performing
-reasoning over it.
-Implementing novelty detection capabilities on an agent such as Dora, would allow to
-to increase reliability and allow the creation of self-extending behaviours.
+The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
+which already possesses a developed architecture based on proababilistic graphical models 
+oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
+One of the major problems in using the probabilistic models of the environments for novelty
+detection is selecting the optimal threshold above which the test sample is considered
+novel. This problem becomes more difficult when, as in case of dora, the representation
+grows as the robot explores the environment. Methods are required to find
+optimal parameters given the current structure of the model. To this end, this work 
+studies methods for novelty threshold selection using probabilistic graphical models.
 
-Its with a final interest on integrating the developed methods on Dora that this
-initial work focus on the novelty detection using probabilistic graphical models
-that model the conceptual knowledge and information handled by Dora.
+The rest of this paper is structured as follows. First, we briefly review the related work on
+room categorization and novelty detection (Section~\ref{}). Then, we give an outline of the 
+architecture of Dora and the structure of the conceptual map representing the semantic information
+(Section~\ref{}). Next, we discuss the methods for novelty detection (Section~\ref{}) 
+and present results of our prelimianary experiments (Section~\ref{}). This paper concludes with a 
+summary in Section~\ref{}.
 

### 2011-05-17 semantic-nd/article.tex
Added sections references in description of document structure.

@@
 The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
-which already possesses a developed architecture based on proababilistic graphical models 
+which already possesses a developed architecture based on probabilistic graphical models 
 oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
@@
 detection is selecting the optimal threshold above which the test sample is considered
-novel. This problem becomes more difficult when, as in case of dora, the representation
+novel. This problem becomes more difficult when, as in case of Dora, the representation
 grows as the robot explores the environment. Methods are required to find
@@
 The rest of this paper is structured as follows. First, we briefly review the related work on
-room categorization and novelty detection (Section~\ref{}). Then, we give an outline of the 
+room categorization and novelty detection (Section~\ref{sec:related}). Then, we give an outline of the 
 architecture of Dora and the structure of the conceptual map representing the semantic information
-(Section~\ref{}). Next, we discuss the methods for novelty detection (Section~\ref{}) 
-and present results of our prelimianary experiments (Section~\ref{}). This paper concludes with a 
-summary in Section~\ref{}.
+(Section~\ref{sec:overview}). Next, we discuss the methods for novelty detection (Section~\ref{sec:novelty-detection}) 
+and present results of our preliminary experiments (Section~\ref{sec:results}). This paper concludes with a 
+summary in Section \ref{sec:conclusion}.
 
@@
 \section{Related Work}
+\label{sec:related}
 % TODO(pronobis): general comment about many .... approches to place categorization.
@@
 \section{Dora Architecture Overview}
+\label{sec:overview}
 Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
@@
 \section{Novelty Detection}
+\label{sec:novelty-detection}
 % Introduce novelty detection
@@
 \section{Results}
+\label{sec:results}
+
 In order to verify the performance of the proposed threshold functions a synthetic dataset
@@
 \section{Conclusions and Future Work}
+\label{sec:conclusion}
 In this paper we presented how to define a stable novelty threshold function on

### 2011-05-17 semantic-nd/article.tex
Added changes of Pronobis.

@@
 \label{sec:overview}
-Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
-
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
-That model connects the sensed properties together with the variables used to model the world.
-And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box.
-
-From its architecture only the \emph{conceptual layer} is of interest to this article.
-Its role is to aggregate the following semantic information coming from other layers:
-
-\begin{description}
- \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
- \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room.
-The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
- \item[Object detection] is performed using the visual input in order to detect .
-The system keeps track of the number of object and types seen per each room.
-Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
- \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
-\end{description}
-
-With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
-in order to model all known variables and their relations.
-As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
-a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.
-
-The use of graphical models to describe distributions of variables has useful properties.
-The edges between the variables can be seen as a kind of filter to the properties of the system.
-Properties those that are expected to be captured by the conceptual knowledge.
-At the same time they are a generative models and therefore allow to calculate the probability
-on any given subset of variables on the graph allowing the system to work even when some
-information is missing.
 
@@
          in order to create a structured probabilistic representation of the world.}
+\label{fig:system}
 \end{figure}
+
+The Dora system~\cite{ijcai} consists of several co-operating sub-systems, all of which actively use or maintain the spatial knowledge representation (see Figure~\ref{fig:system}).
+Only the \emph{conceptual layer} of the representation is of interest to this article.
+Its role is to aggregate the following semantic information coming from other sub-systems:
+\begin{description}
+ \item[Doorway detection] is used to segment the continuous space into rooms and map connectivity between them.
+ \item[Room size and shape] are classified by using 2D laser scans data from laser range finder mounted on the robot and are used as properties of a given room.
+The system utilizes pre-trained set of classifiers to extract rooms sizes (either large, medium or small) and shapes (rectangular or elongated).
+ \item[Object detection] is performed in the images acquired by the robot through its camera.
+The system keeps track of the number of objects of each type in each room.
+Objects are detected by running a pre-trained set of detectors for the following object types: book, cereal box, computer, robot, stapler, toilet paper.
+ \item[Room appearance] is categorized from the visual input by using global visual features and a pre-trained set of 7 different models.
+\end{description}
+
+As Dora moves through the environment its \emph{conceptual layer} builds a structural and
+probabilistic representation of the space instantiated as a \emph{graphical model}.
+It includes taxonomy of human-compatible spatial concepts which are linked to the sensed 
+instances of these concepts drawn from lower layers. It is the conceptual layer which 
+contains the information that kitchens commonly contain cereal boxes and have certain 
+general appearance and allows the robot to infer that the cornflakes box in front of the 
+robot makes it more likely that the current room is a kitchen. The conceptual layer is 
+described in terms of a probabilistic ontology defining spatial concepts and linking 
+those concepts to instances of spatial entities (see the example of the ontology in Figure~\ref{fig:system}).
+
+\subsection{Conceptual Map}
+
+Based on this design, a \emph{chain graph} model is proposed as a 
+representation for performing inferences on the knowledge represented in the conceptual 
+layer. Chain graphs are probabilistic graphical models that combine the properties of 
+both Bayesian Networks and Random Markov Fields. This results in an efficient approach to 
+probabilistic modeling and reasoning about conceptual knowledge.
+
+An exemplary chain graph corresponding to the conceptual map ontology is presented in Fig.~\ref{fig:system}. 
+Each discrete place identified in the environment is represented by a set of random variables, 
+one for each class of relation linked to that place. These are each connected to a random variable
+over the categories of rooms, representing the ``is-a'' relation between rooms and their categories. 
+Moreover, the room category variables are connected by undirected links to one another according 
+to the topological map. The remaining variables represent: shape and appearance properties of space 
+as observed from each place, and the presence of objects. 
+These are connected to observations of features extracted directly from 
+the sensory input. Finally, the 
+distributions $p_{s}(\cdot|\cdot)$, $p_a(\cdot|\cdot)$, $p_{o_i}(\cdot|\cdot)$ 
+represent the common sense knowledge about shape, appearance, and object co-occurrence, respectively. 
+They allow for inference about other properties and room categories e.g. that the room is likely to be a kitchen,
+because you are likely to have observed cornflakes in it.
+
+The use of graphical models to describe distributions of variables has useful properties.
+First, the permit inference about uncertain conceptual knowledge. At the same time, they are 
+generative models and therefore allow to calculate the probability
+on any given subset of variables of the graph, allowing the system to work even when some
+information is missing.
 

### 2011-05-17 semantic-nd/article.tex
Fix last comments by Andrzej and split pics relative to the Dora overview.
Article almost ready for submission.
Just missing better quality pics.

@@
 problem of classifying single images captured in indoor or outdoor environments
-(scene classification)\cite{oliva2006building,torralba2003contextual}.
+(scene classification)~\cite{oliva2006building,torralba2003contextual}.
 At the same time, robotics researchers initially employed
 the 2D laser range sensor being much more robust to variations occurring in the
-environment and much easier to handle computationally in real time\cite{mozos2005supervised}.
+environment and much easier to handle computationally in real time~\cite{mozos2005supervised}.
 
@@
 several sources or classifiers are expected to have better performance on scene
-recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
+recognition than single-cue approaches. Quattoni and Torralba~\cite{quattoni2009recognizing}
 showed that most scene recognition models work poorly in indoor scenes when
@@
 specific objects (books).
-Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
+Galindo et al.~\cite{galindo2005multi} also exploits this by defining a bidirectional relation
 between object and room category, where object defines a room category and a room
@@
 Probabilistic representations have been frequently used for spatial modelling
-in robots operating in the real-world\cite{gross2009toomas,maier2010probabilistic}.
-Boutell et al.\cite{boutell2006factor} have studied outdoor scene classification using
+in robots operating in the real-world~\cite{gross2009toomas,maier2010probabilistic}.
+Boutell et al.~\cite{boutell2006factor} have studied outdoor scene classification using
 \emph{factor graphs} and modelling spatial relations between objects in the scene
@@
 An interesting example is the usage of an hidden concept layer between visual features
-and text information to provide automatic image annotation\cite{zhang2006probabilistic}.
+and text information to provide automatic image annotation~\cite{zhang2006probabilistic}.
 
 Novelty detection has been studied for many years and there are several approaches
-based on statistical analysis\cite{markou2003novelty}.
+based on statistical analysis~\cite{markou2003novelty}.
 Graphical models have been used to learn distributions of variables, both in
 supervised and unsupervised ways and by using thresholds on those distributions
-based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
+based solely on the conditional probability as seen on Bishop~\cite{bishop1994novelty}
 a novelty system can be trivially implemented.
@@
 \centering
-\subfloat[Conceptual layer]{\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}}
-\subfloat[Chain graph]{\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}}
-
-\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
-         in order to create a structured probabilistic representation of the world.}
-\label{fig:system}
+\includegraphics[width=0.9\textwidth]{figures/conceptual-layer.jpg}
+\caption{\label{fig:conceptual-layer}Interaction of the sub-systems
+         in Dora with special focus on the conceptual layer.}
 \end{figure}
 
-The Dora system~\cite{ijcai} consists of several co-operating sub-systems, all of which actively use or maintain the spatial knowledge representation (see Figure~\ref{fig:system}).
+The Dora system~\cite{ijcai} consists of several co-operating sub-systems,
+all of which actively use or maintain the spatial knowledge representation
+(see \autoref{fig:conceptual-layer}).
 Only the \emph{conceptual layer} of the representation is of interest to this article.
@@
 described in terms of a probabilistic ontology defining spatial concepts and linking 
-those concepts to instances of spatial entities (see the example of the ontology in Figure~\ref{fig:system}).
+those concepts to instances of spatial entities (see the example of the ontology in
+\autoref{fig:conceptual-layer}).
 
@@
 
-An exemplary chain graph corresponding to the conceptual map ontology is presented in Fig.~\ref{fig:system}. 
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.90\textwidth]{figures/chain-graph.png}
+\caption{\label{fig:chain-graph}Example chain-graph produced by the \emph{conceptual layer}.}
+\end{figure}
+
+%The conceptual layer structures the sensed environment together with the conceptual knowledge
+%in order to create a structured probabilistic representation of the world.
+An exemplary chain graph corresponding to the conceptual map ontology is presented
+in \autoref{fig:chain-graph}. 
 Each discrete place identified in the environment is represented by a set of random variables, 
@@
 Although the conceptual layer works with \emph{chain graphs}, those can be converted
-into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
+into \emph{factor graphs}~\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
 easier manipulation due to factorization.
 Moreover, there exists efficient implementation of inference engines operating on factor graph
-representations\cite{Mooij_libDAI_10}.
+representations~\cite{Mooij_libDAI_10}.
 Describing the distribution function in terms of graphs allows to use those engines to
@@
 Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about\cite{markou2003novelty}.
+other than the ones the system knows about~\cite{markou2003novelty}.
 It is harder than classification as only positive samples of a class are available
@@
 
-Adding capabilities to an agent to detect a normal sample allows to increase its
-reliability. As a novelty signal can then be interpreted by the system to proceed
-with more caution as its on the presence of a situation where its knowledge
-does not correctly describes reality.
+Adding capabilities to an agent to detect novel samples allows to increase its
+reliability. Novelty signal can then be interpreted by the system to proceed
+with more caution as its knowledge does not correctly describe reality.
 
@@
 Due to the desire of robustness, a novelty detection system is often given the option
-to reject normal samples. This is implemented as a threshold that describes a trade
+to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
 between error and rejection: the more the system tries to reduce rejection, the more
@@
 When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel as an associated error probability of $P(\overline{novel}|x)$.
+classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
 And the error rate of a system that classifies the set of inputs $N$ as novel is given by

### 2011-05-24 thesis/acknows.tex
Getting the thesis structure and build system in shape.

\chapter*{Acknowledgements}

Aliquam id dui. Nulla facilisi. Nullam ligula nunc, viverra a, iaculis
at, faucibus quis, sapien. Cum sociis natoque penatibus et magnis dis
parturient montes, nascetur ridiculus mus. Curabitur magna ligula,
ornare luctus, aliquam non, aliquet at, tortor. Donec iaculis nulla
sed eros. Sed felis. Nam lobortis libero. Pellentesque
odio. Suspendisse potenti. Morbi imperdiet rhoncus magna. Morbi
vestibulum interdum turpis. Pellentesque varius. Morbi nulla urna,
euismod in, molestie ac, placerat in, orci.

Ut convallis. Suspendisse luctus pharetra sem. Sed sit amet mi in diam
luctus suscipit. Nulla facilisi. Integer commodo, turpis et semper
auctor, nisl ligula vestibulum erat, sed tempor lacus nibh at
turpis. Quisque vestibulum pulvinar justo. Class aptent taciti
sociosqu ad litora torquent per conubia nostra, per inceptos
himenaeos. Nam sed tellus vel tortor hendrerit pulvinar. Phasellus
eleifend, augue at mattis tincidunt, lorem lorem sodales arcu, id
volutpat risus est id neque. Phasellus egestas ante. Nam porttitor
justo sit amet urna. Suspendisse ligula nunc, mollis ac, elementum
non, venenatis ut, mauris. Mauris augue risus, tempus scelerisque,
rutrum quis, hendrerit at, nunc. Nulla posuere porta orci. Nulla dui.

Fusce gravida placerat sem. Aenean ipsum diam, pharetra vitae, ornare
et, semper sit amet, nibh. Nam id tellus. Etiam ultrices. Praesent
gravida. Aliquam nec sapien. Morbi sagittis vulputate dolor. Donec
sapien lorem, laoreet egestas, pellentesque euismod, porta at,
sapien. Integer vitae lacus id dui convallis blandit. Mauris non
sem. Integer in velit eget lorem scelerisque vehicula. Etiam tincidunt
turpis ac nunc. Pellentesque a justo. Mauris faucibus quam id
eros. Cras pharetra. Fusce rutrum vulputate lorem. Cras pretium magna
in nisl. Integer ornare dui non pede.

\vspace{10mm}
\flushleft{The Name of the Author}

### 2011-05-24 thesis/conclusion.tex
Getting the thesis structure and build system in shape.

\chapter{Conclusions and Future Work}
\section*{}

In this thesis

### 2011-05-24 thesis/introduction.tex
Getting the thesis structure and build system in shape.

--- 2011-03-30 pdis/techreport/mieic-distrib/chapter1.tex
@@
-\chapter{Introdução} \label{chap:intro}
+\chapter{Introduction} \label{chap:intro}
 
@@
 
-O primeiro capítulo da dissertação deve servir para apresentar o
+AO primeiro capítulo da dissertação deve servir para apresentar o
 enquadramento e a moti\-va\-ção do trabalho e para identificar e
@@
 
-Este documento ilustra o formato a usar em dissertações na \Feup, não
+Este documento ilustra o formato a usar em dissertações na \gls{FEUP}, não
 servindo de exemplo sobre os conteúdos a usar.
@@
 Uma recolha de normas existentes sobre este assunto pode ser
-encontrada em~\citet{kn:Mat93}.
+encontrada em~\cite{kn:Mat93}.
 
@@
   how the dissertation is structured, and in doing so, define the
-  thread of the contents.''~\citep[chap.\ Introduction]{kn:Tha01} 
+  thread of the contents.''~\cite{kn:Tha01} 
 \end{quote}
 
-Neste primeiro capítulo ilustra-se a utilização de citações e de
-referências biblio\-grá\-fi\-cas.
-Para além de dar um exemplo de utilização de uma citação, a citação
-anterior, introduz uma referência que pode ser consultada, entre
-muitas outras referências bibliográficas
-interessantes~\citep{kn:Tha01,kn:PP05}.
+Blabla \gls{Kinect}
 
-\section{Contexto/Enquadramento} \label{sec:context}
+\section{Context} \label{sec:context}
 
@@
 
-Lorem ipsum~\citep{kn:Lip08} dolor sit amet, consectetuer adipiscing
-elit. 
-Sed eget nunc. Phasellus interdum, risus viverra mollis laoreet, felis
-justo iaculis ante, eget ornare purus augue non urna. Nam in magna. In a
-est. Phasellus a tellus vitae enim vehicula imperdiet. Etiam sit amet
-elit. In hac habitasse platea dictumst. Quisque eget turpis vel felis
-elementum tempus. Curabitur sit amet tortor id libero dapibus
-pretium. Integer mattis eros eu lorem. Duis erat tellus, porttitor
-sed, blandit eget, fringilla et, lacus. Phasellus tristique nibh nec
-orci. Mauris sed leo. Suspendisse fringilla tempor dolor. Donec sapien
-enim, congue in, porta et, sollicitudin in, quam. Curabitur semper,
-mauris ut vestibulum eleifend, diam ipsum tincidunt quam, et
-vestibulum velit mauris ut risus.
-
-Sed eget libero. Nulla facilisi. Proin eget tortor. Morbi
-gravida. Donec arcu risus, blandit a, rutrum at, ornare ut,
-nisl. Etiam consectetuer tortor eu odio. Etiam blandit molestie
-ligula. Nulla facilisi. Nam a augue non justo laoreet hendrerit. Nam
-aliquam, purus eu ultricies dictum, urna purus posuere neque, vel
-tempus tellus enim a arcu.
-
-\section{Projecto} \label{sec:proj}
-
-Na continuação da secção anterior, e apenas no caso de ser um Projecto
-e não uma Dissertação, esta secção apresenta resumidamente o projecto.
-
-Nulla nec eros et pede vehicula aliquam. Aenean sodales pede vel
-ante. Fusce sollicitudin sodales lacus. Maecenas justo mauris,
-adipiscing vitae, ornare quis, convallis nec, eros. Etiam laoreet
-venenatis ipsum. In tellus odio, eleifend ac, ultrices vel, lobortis
-sed, nibh. Fusce nunc augue, dictum non, pulvinar sed, consectetuer
-eu, ipsum. Vivamus nec pede. Pellentesque pulvinar fringilla dolor. In
-sit amet pede. Proin orci justo, semper vel, vulputate quis, convallis
-ac, nulla. Nulla at justo. Mauris feugiat dolor. Etiam posuere
-fermentum eros. Morbi nisl ipsum, tempus id, ornare quis, mattis id,
-dolor. Aenean molestie metus suscipit dolor. Aliquam id lectus sed
-nisl lobortis rhoncus. Curabitur vitae diam sed sem aliquet
-tempus. Sed scelerisque nisi nec sem.
-
-\section{Motivação e Objectivos} \label{sec:goals}
+\section{Motivation and Goals} \label{sec:goals}
 
@@
 
-Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Morbi sit
-amet nibh. Fusce faucibus, enim vel ultrices ornare, est mauris
-ultricies velit, vitae consequat sem erat vel nunc. Nam libero eros,
-mattis eget, sagittis nec, imperdiet at, sapien. Aliquam lacus. Aenean
-adipiscing nibh in orci. Aliquam vestibulum, elit at fringilla
-dignissim, metus diam lobortis urna, a laoreet nunc odio ac ipsum. Sed
-at urna. Integer vehicula fringilla augue. Nulla lacus eros, rhoncus
-sit amet, posuere ut, vehicula ac, nibh. Ut eleifend, eros eu placerat
-vehicula, justo turpis blandit dolor, eu tincidunt felis risus at
-ante. Aenean suscipit nisl eget eros. Ut laoreet libero eget
-enim. Cras tempus pellentesque felis. Vestibulum vitae erat ac nibh
-posuere eleifend.
+\section{Thesis Outline} \label{sec:outline}
 
-Integer nec quam. Sed fermentum. Nunc vitae leo. Etiam sit amet
-quam. Nunc vestibulum massa in mauris. Duis eget nulla. Fusce
-ultricies arcu eu nibh volutpat feugiat. Maecenas urna pede, commodo
-quis, porta eu, bibendum elementum, pede. Sed eros massa, molestie
-eget, mattis non, rutrum ac, magna. Duis dui. Maecenas eget tortor ut
-dolor semper mattis. Maecenas auctor, tellus et ultricies tempor, elit
-est placerat lacus, in posuere mauris lorem et arcu.
-
-\section{Estrutura da Dissertação} \label{sec:struct}
-
-Para além da introdução, esta dissertação contém mais x capítulos.
-No capítulo~\ref{chap:sota}, é descrito o estado da arte e são
-apresentados trabalhos relacionados. 
-No capítulo~\ref{chap:chap3}, ipsum dolor sit amet, consectetuer
-adipiscing elit.
-No capítulo~\ref{chap:chap4} praesent sit amet sem. 
-No capítulo~\ref{chap:concl}  posuere, ante non tristique
-consectetuer, dui elit scelerisque augue, eu vehicula nibh nisi ac
-est.
+The rest of this thesis is structured as follows\dots

### 2011-05-24 thesis/macros.tex
Getting the thesis structure and build system in shape.

--- 2011-05-17 semantic-nd/article.tex
@@
-\documentclass[runningheads,a4paper]{llncs}
-
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
-
-\usepackage{url}
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
-
 % Additional packages
@@
 }
-
-\begin{document}
-
-\mainmatter  % start of an individual contribution
-
-% first the title is needed
-\title{Novelty Detection Using Graphical Models for Semantic Room Classification}
-
-% a short form should be given in case it is too long for the running head
-% \titlerunning{Lecture Notes in Computer Science: Authors' Instructions}
-
-% the name(s) of the author(s) follow(s) next
-%
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
-\author{Removed for Blind Review}
-%\author{André Susano Pinto, Andrzej Pronobis, Luis Paulo Reis}
-%
-%\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-\institute{Removed for Blind Review}
-%\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
-%\url{andresusanopinto@gmail.com}\\
-%\url{http://url}}
-
-\maketitle
-
-\begin{abstract}
-This paper presents an approach to the problem of novelty
-detection in the context of semantic room categorization.
-The ability to assign semantic labels to areas in the environment is crucial for
-autonomous agents aiming to perform complex human-like tasks and human
-interaction.
-However, in order to be robust and naturally learn the semantics from
-the human user, the agent must be able to identify gaps in its own knowledge.
-To this end, we propose a method based on graphical models to identify novel
-input which does not match any of the previously learnt semantic descriptions.
-The method employs a novelty threshold defined in terms of conditional
-and unconditional probabilities. The novelty threshold is then optimized using
-an unconditional probability density model trained from unlabelled data.
-
-\keywords{novelty detection, semantic data, probabilistic graphical models,
-room classification, indoor environments, robotics, multi-modal classification.}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-There has been several efforts in the areas of artificial intelligence and robotics
-in creating robots that are able to interact with humans and their environments.
-One of the important aspects is to endow those robots with a deeper understanding
-of human environments, not just for the purpose of navigation and obstacle avoidance, 
-but also in terms of human semantics and functionality. An important problem
-in creating reliable representations of space for robots that are to be deployed 
-in new and unknown realistic environments is to be able to automatically 
-identify gaps in robot's knowledge and act in order to fill those gaps.
-
-This article addresses the problem of \emph{novelty detection} within the context
-of semantic mapping i.e. generating maps containing \emph{semantic information} 
-about indoor environments, such as homes or offices. In that context, the ability 
-to detect that the observations result from a semantic concept unknown to the robot, 
-and cannot be explained by one of its models, is crucial for generating fully autonomous
-and reliable behavior. In order to be robust, the robot must identify 
-novel concepts and instead of making a costly error, refrain from the decision
-and initiate learning. In particular, we address the problem of novelty detection for
-room categorization i.e. detecting whether the area in the environment identified as a
-separate room can be assigned one of the semantic labels that the robot knows (e.g. as \emph{a kitchen} 
-or \emph{an office}) or belongs to a yet unknown semantic category.
-
-The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
-which already possesses a developed architecture based on probabilistic graphical models 
-oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
-One of the major problems in using the probabilistic models of the environments for novelty
-detection is selecting the optimal threshold above which the test sample is considered
-novel. This problem becomes more difficult when, as in case of Dora, the representation
-grows as the robot explores the environment. Methods are required to find
-optimal parameters given the current structure of the model. To this end, this work 
-studies methods for novelty threshold selection using probabilistic graphical models.
-
-The rest of this paper is structured as follows. First, we briefly review the related work on
-room categorization and novelty detection (Section~\ref{sec:related}). Then, we give an outline of the 
-architecture of Dora and the structure of the conceptual map representing the semantic information
-(Section~\ref{sec:overview}). Next, we discuss the methods for novelty detection (Section~\ref{sec:novelty-detection}) 
-and present results of our preliminary experiments (Section~\ref{sec:results}). This paper concludes with a 
-summary in Section \ref{sec:conclusion}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-% TODO(pronobis): general comment about many .... approches to place categorization.
-The problem of room categorization based on visual information was first addressed
-in the computer vision community. In this case, the research focused mainly on the
-problem of classifying single images captured in indoor or outdoor environments
-(scene classification)\cite{oliva2006building,torralba2003contextual}.
-At the same time, robotics researchers initially employed
-the 2D laser range sensor being much more robust to variations occurring in the
-environment and much easier to handle computationally in real time\cite{mozos2005supervised}.
-
-Multi-modal approaches, such as combining semantic data extracted from
-several sources or classifiers are expected to have better performance on scene
-recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
-showed that most scene recognition models work poorly in indoor scenes when
-compared to outdoor scenes since the properties that
-characterize rooms changes depending on the category. For instance corridors are well
-described by global properties and bookstores are well described by the presence of
-specific objects (books).
-Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
-between object and room category, where object defines a room category and a room
-category provides information on where objects may be found.
-
-Probabilistic representations have been frequently used for spatial modelling
-in robots operating in the real-world\cite{gross2009toomas,maier2010probabilistic}.
-Boutell et al.\cite{boutell2006factor} have studied outdoor scene classification using
-\emph{factor graphs} and modelling spatial relations between objects in the scene
-to extract better knowledge from semantic (high-level) features.
-
-Although our approach is presented in the context of mobile robotics it relies on
-standard concepts and techniques such as semantic data and graphical models.
-Those are often used in the area of information retrieval.
-An interesting example is the usage of an hidden concept layer between visual features
-and text information to provide automatic image annotation\cite{zhang2006probabilistic}.
-
-Novelty detection has been studied for many years and there are several approaches
-based on statistical analysis\cite{markou2003novelty}.
-Graphical models have been used to learn distributions of variables, both in
-supervised and unsupervised ways and by using thresholds on those distributions
-based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
-a novelty system can be trivially implemented.
-
-However to the knowledge of the authors there is no reference on how to perform
-novelty detection using graphs that are dynamically generated.
-% TODO(andresp): anything more that can be said on why this paper is useful?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dora Architecture Overview}
-\label{sec:overview}
-Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
-
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
-That model connects the sensed properties together with the variables used to model the world.
-And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box.
-
-From its architecture only the \emph{conceptual layer} is of interest to this article.
-Its role is to aggregate the following semantic information coming from other layers:
-
-\begin{description}
- \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
- \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room.
-The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
- \item[Object detection] is performed using the visual input in order to detect .
-The system keeps track of the number of object and types seen per each room.
-Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
- \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
-\end{description}
-
-With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
-in order to model all known variables and their relations.
-As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
-a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.
-
-The use of graphical models to describe distributions of variables has useful properties.
-The edges between the variables can be seen as a kind of filter to the properties of the system.
-Properties those that are expected to be captured by the conceptual knowledge.
-At the same time they are a generative models and therefore allow to calculate the probability
-on any given subset of variables on the graph allowing the system to work even when some
-information is missing.
-
-\begin{figure}[h]
-\centering
-\subfloat[Conceptual layer]{\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}}
-\subfloat[Chain graph]{\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}}
-
-\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
-         in order to create a structured probabilistic representation of the world.}
-\end{figure}
-
-\subsection{Factor Graphs}
-Although the conceptual layer works with \emph{chain graphs}, those can be converted
-into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
-easier manipulation due to factorization.
-Moreover, there exists efficient implementation of inference engines operating on factor graph
-representations\cite{Mooij_libDAI_10}.
-Describing the distribution function in terms of graphs allows to use those engines to
-efficently calculate marginals on any given subset of variables by exploiting conditional
-independence between variables.
-
-A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
-representing random variables and factors.
-Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
-to which the factor is connected.
-Thus, a factor graph can be seen as a description of probability density function obtained
-by a product of all the factors. In order to represent the probability,
-a normalization factor needs to be introduced, resulting in the following equation:
-
-\begin{equation}
-P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
-Z = \sum_{X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Novelty Detection}
-\label{sec:novelty-detection}
-% Introduce novelty detection
-Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about\cite{markou2003novelty}.
-It is harder than classification as only positive samples of a class are available
-rendering normal classification methods invalid.
-
-Adding capabilities to an agent to detect a normal sample allows to increase its
-reliability. As a novelty signal can then be interpreted by the system to proceed
-with more caution as its on the presence of a situation where its knowledge
-does not correctly describes reality.
-
-% Explain threshold approach to novelty detection
-Due to the desire of robustness, a novelty detection system is often given the option
-to reject normal samples. This is implemented as a threshold that describes a trade
-between error and rejection: the more the system tries to reduce rejection, the more
-prone to produce errors it becomes.
-
-When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel as an associated error probability of $P(\overline{novel}|x)$.
-And the error rate of a system that classifies the set of inputs $N$ as novel is given by
-the equation:
-
-\begin{equation}
-P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
-\end{equation}
-
-Therefore any novelty system that decides to acknowledge a sample
-$a$ as novel should also detect novelty on any sample $b$ satisfying
-$P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that would always reduce the error probability.
-This way a novelty detection system is interested in defining an
-order relation on all the possible inputs equivalent to the order defined by
-the error rate: $P(\overline{novel}|x)$.
-
-Using Bayes rule and assuming a constant $P(\overline{novel})$
-a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
-Such a ratio is a suitable function for implementing a novelty detector system with optimal
-thresholding.
-
-\begin{equation}
-\label{eq:novelty-threshold}
-          P(\overline{novel}|x)
-  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
-  \propto \frac{P(x|\overline{novel})}{P(x)}
-\end{equation}
-
-% Explain approach of using conditional probability
-\subsection{Conditional Probability}
-\label{sec:conditional-prob}
-
-The conditional probability models the distribution of variables given that the sample
-is not novel for the agent.
-This is: it should model the distribution of the sensed features considering that the
-knowledge the agent possesses holds true. Under that its natural to model it with a graphical
-model that combines the information acquired with the relations that the describe the
-relations learned by the agent between the variables.
-
-In Dora case, this is equivalent to use the graphical model used by it to describe its
-current believes on the variables modelled by the system.
-That graph is built, by the conceptual layer, by instancing the information extracted from
-other layers together with the conceptual knowledge such as: objects are properties of rooms,
-rooms are connected between each other.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
-\caption{\label{fig:conditional-prob-graph}A sample of a factor graph modelling a distribution
-         of a set $x$ of sensed variables.}
-\end{figure}
-
-\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
-layer to represent the conditional probability on the sensed variables $x$.
-A set of hidden variables is added to represent the conceptual knowledge the system is aware of.
-In the presented graph variables $R_i$ were added to model the room categories that influence
-the directly sensed features on each physical room, as well connectivity factors between each
-room.
-
-The factors connecting the variables are trained by the system by searching databases of
-common knowledge to build potentials describing how likely it is that a specific set of
-values for a set of variables is more likely to occur over than any other set of values.
-For instance: it is very likely to find a cereal box in a kitchen; and it is unlikely to find
-bathroom connected to another bathroom.
-
-We propose using such a graph $G$ built by the conceptual layer for modelling $P_G(x)$
-as an approximation for $P(x|\overline{novel})$.
-
-\subsection{Unconditional Probability}
-\label{sec:unconditional-prob}
-With only access to labelled data a common approach is to define a threshold assuming
-that $P(x)$ is constant through all the samples.
-
-Its important to notice that in several cases assuming it to be constant leads to
-discarding the factor.
-Nonetheless, here the distributions are dynamically changing as the system learns
-more on the environment.
-So the normalizing argument $P(x)$ has to be evaluated for each new subset of $x$.
-
-Assuming that the unconditional distributions generates all possible outcome with
-the same probability we can model it with $\prod{1/\# x_i}$,
-where $\# x_i$ denotes the cardinality of the state space of variable $x_i$.
-In graphical model terms this is represented to a factor graph $U$ with the
-variables but without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
-\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a
-         uniform distribution over any set of its variables.}
-\end{figure}
-
-Having a graphical model $G$ built to model the known data distribution and a model
-$U$ for the unconditional probability a novelty threshold would be given by:
-$P_G(x)/P_U(x)$.
-Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of
-variables $x$ into the same measure units (error rate), such that a static threshold can be implemented.
-For example the conditional probability would yield very small values on large sets of variables
-$x$ than in small sets due to the spreading over the dimensions of the input space.
-As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
-understand the observed result.
-
-\subsection{Semi Supervised: Using Unlabelled Data}
-Nonetheless, its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability
-distribution than the uniform one.
-
-For practical reasons it was assumed that the unconditional distribution
-could be modelled with all the variables independent from each other.
-Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-From the unlabelled data, only the probability of each feature $x_i$ needs to be estimated.
-\autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/independent-prob-graph.pdf}
-\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to estimate
-         the unconditional probability for each variable and this way try to compensate
-         for a possible existent bias on some of the variables.}
-\end{figure}
-
-Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
-This time the addition of the unconditional factors can be understood as an
-attempt to compensate for an existing bias on the unconditional distribution.
-This is an important step to achieve a correct order-relation of the inputs sample
-for implementing a novelty threshold.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-
-In order to verify the performance of the proposed threshold functions a synthetic dataset
-was generated. To keep it simple only information regarding direct features of
-a room were modelled and no structured knowledge such as room connectivity was taken
-in account.
-The synthetic distribution assumes that an independent and variable size set of features
-$x$ is generated by a given room category.
-In whole there was 11 different room categories and 7 different measured feature
-types. Each feature type can be present more than once in (e.g.: room shape is
-extracted from 2D laser scans in more than one position in the room).
-
-The room categories were chosen to mimic as close as possible the real features and
-categories existing in reality. And they model different room categories with
-different levels of detail. For example: 1 person office, 2 person office, hallway,
-robot lab.
-
-From the distribution, 100 labelled samples for 5 of the 11 room categories were
-drawn to represent the known concepts and 1000 unlabelled samples were drawn from
-all the room categories for learning the unconditional probability distribution.
-Using those samples, factors were learnt for the graphs used to model the
-conditional distribution and the independent unconditional distribution.
-\autoref{fig:simple-experiment} shows the graph structure used for approximate the
-trained conditional and unconditional distributions.
-Its important to notice that graph $G$ used to model the known classes when given
-enough labelled data is able to exactly learn the conditional distribution as it
-uses the same structure as the created synthetic distribution.
-
-\begin{figure}[h]
-\centering
-
-\subfloat[Graph structure $G$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-cond-graph.pdf}}
-\qquad
-\subfloat[Graph structure $U$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-uniform-graph.pdf}}
-\qquad
-\subfloat[Graph structure $I$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-independent-graph.pdf}}
-
-\caption{\label{fig:simple-experiment}The graph structures used to model the
-         conditional and unconditional probability for implementing the novelty
-         thresholds $P_G(x)/P_U(x)$ and $P_G(x)/P_I(x)$.}
-\end{figure}
-
-% Describe how to obtain the 3 thresholds functions seen on the graphs.
-Using the learned models $G$, $U$ and $I$, two thresholds were trained:
-$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
-and $P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
-Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
-and a perfect threshold function could also be created to test how far the
-presented thresholds are from optimal.
-
-\subsection{Probability Ratio Comparison}
-%%% Results 1
-% Show the threshold ratio is an optimal detector (if perfect information was available)
-% Show that the thresholds are suitable functions for implementing a static threshold.
-As first tryout, the performance of the novelty threshold selection was plotted for a set
-of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
-Those samples contain different number of sensed properties for each room, mimicking
-the dynamic properties expected to see when implemented on a robot.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.60\textwidth]{results/synthetic-all.pdf}
-
-\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
-         under samples with variable size of sensed properties.}
-\end{figure}
-
-The convex shape for the optimal threshold shows that the ratio between conditional
-and unconditional probability is indeed an optimal detector and is suitable for
-implementing a threshold when the samples are taken from dynamic
-distributions (e.g.: some samples where there is only access to room size versus
-samples where there is a lot of information about the room properties).
-
-% Discuss importance on approximating unconditional probability.
-Its also possible to see how important it is to estimate a correct unconditional
-probability in order to obtain a correct novelty measure on the inputs.
-The assumption of a uniform unconditional probability has led to very poor results.
-That is probably explained by the semantic properties being highly
-biased towards some values. And shows that bias plays an important step
-in detecting whether a given sensed value is a valuable cue about the room category.
-
-%%% Results 2
-% Measure performance of the thresholds as more information becomes available.
-\subsection{Performance Changes With Amount of Available Information}
-In order to measure the performance impact as more semantic information becomes
-available ROC curves were plotted for samples grouped by the number of sensed
-semantic features.
-
-\begin{figure}[h]
-\centering
-
-\subfloat[3 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-3features.pdf}}
-\qquad
-\subfloat[5 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-5features.pdf}}
-
-\subfloat[10 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-10features.pdf}}
-\qquad
-\subfloat[50 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-50features.pdf}}
-
-\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
-         presented novelty detection method on graphs generated for different amount of
-         sensed features.}
-\end{figure}
-
-We can see that as the system gains more semantic information it becomes easier
-to detect novelty. The input space size increases and allows the several existing
-classes to become more easily distinguished.
-
-The performance of the independent threshold decreases as the number of sensed
-features increases. This is easily explained by the fact that the graph $I$ is not
-able to model the existent dependence between the features. This becomes obvious
-as the number of features increases (e.g.: graph $I$ perfectly models $P(x)$ in the
-case where only 1 feature is sensed).
-
-The uniform threshold shows a poor performance specially on small size samples
-where its performs almost no better than random.
-It performance increases as the size of sensed features increases but nonetheless
-its very small when compared to how optimal a threshold could be.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and Future Work}
-\label{sec:conclusion}
-In this paper we presented how to define a stable novelty threshold function on
-top of \emph{probabilistic graphical models} instantiated dynamically from sensed
-semantic data.
-The presented technique is based on the ratio between a conditional and
-unconditional probability and when perfect information exists it performs an optimal
-novelty detection threshold.
-
-It was also shown that a correct estimation of unconditional probability plays an
-important role specially on small input spaces. Moreover, semi-supervised techniques
-implemented with the access to unlabelled data can be used to significantly improve
-novelty detection performance.
-
-Given the synthetic distribution, an assumption on an uniform
-distribution has led to very poor results. The same behaviour is expected to
-in real world distributions based on semantic data. For that reason,
-and due to easy access to unlabelled data, special attention will be given to using
-semi-supervised techniques for novelty detection.
-
-After this initial study on how to detect new semantic classes based on
-graphical models, future work will focus on how to use the structured
-information available from the conceptual layer to be able to detect which variable
-of the graph is novel and what makes it different from other previously learned
-classes. That will lead to generation of useful information that can be used for
-communication with the user and performing active learning of new room categories.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{unsrt}
-\bibliography{refs}
-
-\end{document}

### 2011-05-25 thesis/abstract.tex
More structuring...

\chapter*{Abstract}

This should probably be written as last thing.

\chapter*{Resumo}
Let $E$ denote the set of all possible texts in English and $P$ the set of
all possible texts in Portuguese. A transformation $T_{E->P}(x)$ can be defined
in $E \times P$, such that it maps $x \in E$ to a translated version $y \in P$ such
that they hold the same semantic meaning to a reader.

Assume this text to be $T_{E->P}(abstract)$.
\footnote{What I mean is that this will be a translation of the english text present
on abstract and don't expect it to be filled in before the final version of this
document!}

### 2011-05-25 thesis/contents.tex
More structuring...

\chapter{Introduction}
\section*{}
In \cite{kn:Mat93} some thesis rules are defined in portuguese.

\section{Context}
\section{Motivation and Goals}
\section{Thesis Outline}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Background}
\section*{}

\section{Classification}
\subsection{Recognition and Categorization}
\subsection{Support Vector Machines}

\section{Feature Extraction}
\subsubsection{Local Features}
\subsubsection{Global Features}

\section{Semantics}
\section{Semantic Tagging Mapping}
\section{Label Propagation}

% On novelty detection and uninformed guess.
\section{Principle of Maximum Entropy}

\section{Probabilistic Graphical Models}
\subsection{Factor Graphs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Semantic Mapping}
\section*{}
\section{Related Work}
\section{Conceptual Map}
\section{Spatial-Temporal Accumulation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Novelty Detection}
\section*{}
\section{Related Work}
\section{Novelty Detection as a Threshold}
\section{Using unlabelled Data}
\section{Detecting New Classes} 
\section{Detecting New Structure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Conclusions and Future Work}
\section*{}

\section{Conclusions}
\section{Future Work}

### 2011-05-26 thesis/contents.tex
Still refining the thesis structure and some likely content.

@@
+% What should the title be?
+% Novelty Detection for Visual Indoor Categorization
+% Novelty Detection on Semantic Representations.
+
 \chapter{Introduction}
-\section*{}
-In \cite{kn:Mat93} some thesis rules are defined in portuguese.
+% Introduce:
+%  Mobile Robotics, A.I.
+%  Interaction with Humans
+%  Need for concepts/semantic information
+%  Dynamic human environment
+%  Non-feasibility of hard-coded concepts
+%  Requirements to detect novelty and handle it.
+For a long time humanity has fantasized that one day robots will walk among us.
+They will move and be able to interact with us. Understand our concepts and
+be able to reason.
 
-\section{Context}
-\section{Motivation and Goals}
+They need to possess ability to adapt to situations as its infeasible to rely
+on extensive man-work to tag objects, map space and code all the aspects that
+make up our human-reality.
+
+% == Introduce the need of semantic representations ==
+%
+% There is a lot of low-level methods (or specific knowledge to robots)
+% But the creation of semantic representations allow to bridge that low-level
+% sensing with high-level concepts that facilitate several high-level Tasks.
+
+% Importance of mapping and localization in a robot as well as reasoning on
+% properties of each space.
+% Humans give labels to space that characterized the properties and activities
+% expected to be performed on them.
+% Such information is of great interest to be structured and organized in such
+% a way reasoning and planning can be implemented.
+%
+% It also holds an important role in long-term planning and stability as it
+% hides the space-time local details and allows to focus on high-level thinking.
+
+% Deal with uncertainty -> need for probabilistic models.
+% Deal with semantic    -> need of structured models such as graphical models.
+
+Example of what kind of an high-level thinking on space and its semantics can
+provide: Get me a beer/milk concept.
+
+% == Reasoning about need for Novelty Detection ==
+%
+% Dynamic environment (inability to know the environment or even to be possible
+% to map all semantic concepts to interact with humans).
+%
+% Knowledge-Awareness (knowing what we know... And detecting novel cases)
+% Identify gaps in the semantic knowledge.
+% Automatic detection and learning of novel concepts.
+%
+% Increase robustness
+% Self-Extending
+%
+
+\section{Problem and Goals}
+% Extend a probabilistic graphical modelling framework with capabilities to
+% detect novelty.
+% Both in terms of novel classes as of novel structure.
+%
+Given a probabilistic structure representing the sensed conceptual knowledge
+obtained by the agent develop methods to be able to detect novelty present
+either on new semantic concepts (new classes) or even on structure that was
+previously unknown.
+
+% Yeah thats what you call dream :P
+In an extreme ideal case the agent should be able to go from zero-knowledge to
+understanding presence of certain objects (as generators of a set of sensed
+sensed properties grouped locally), understand areas within rooms (sink-area on
+on a kitchen), rooms as separated by doors, and understand environments such
+as office, home, warehouse, spaceship.
+
+For that the probabilistic semantic representation presented by
+\cite{andrzej2011phd} is used.
+
+% If everything goes fine... 2 of the "Future Directions" proposed are 
+% Novelty Detection and Learning of Novel Concepts:
+%  Identify Gaps in Spatial and Semantic Knowledge -> Addressed
+%  Performing learning of new concepts -> Outside Scope / No Time
+%
+% Using Properties for Space Segmentation
+%  If we move on detecting novel structures on the graph we are addressing this
+%  issue.
+
 \section{Thesis Outline}
+The rest of this thesis is structured as follows:
+
+\autoref{chap:background} introduces background concepts for understanding the
+presented thesis. In particular \autoref{sec:graphical-models} introduces
+\emph{probabilistic graphical models} that lay the base modelling tool for
+the semantic structured representation over which novelty detection
+methods are developed.
+
+\autoref{chap:semantic-mapping} describes the system proposed by
+\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
+that is responsible for the creation of the \emph{probabilistic semantic
+representation} that this thesis aims at improving by developing methods with
+the capability to identify knowledge gaps.
+
+\autoref{chap:novelty-intro} introduces novelty detection under a
+statistical view point and how to interpret it as a threshold function.
+A simple approach on how to perform novelty on very simple graphs
+is given together with some results on the impact of using semi-supervised
+novelty detection to improve the system performance.
+
+\autoref{chap:novelty} presents the developed techniques developed to identify
+novelty on the semantic representation.
+
+\autoref{chap:conclusions} draws conclusions on the developed work and presents
+interesting directions for future works.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Background}
-\section*{}
+\chapter{Background}\label{chap:background}
+% This chapter will perform a breath-first explanation on the contents related
+% with the thesis. The extension of the contents explained here can be seen as
+% monotone function over the number of missing pages to attain the requirements.
+%
+% Some topics are indeed important as its the case of factor graphs.
+% Others like:
+%  - classification (SVM, multi-class, kernel-trick),
+%  - features (SIFT, GIST),
+%  - Basic Bayes theory and probabilistic review
+% are included for fun of the reader in case it inquires himself how to deal
+% with the low-levels features we propose using.
+%
+% ---
+% PS: This is the chapter that will be used to fill with unnecessary crap
+% in case they really annoy me with the number of pages. Every other chapter
+% will strive to be really required and be a masterpiece.
+%
 
@@
 \subsection{Support Vector Machines}
+\subsection{Multi-Class Classification}
+%\subsection{Kernel-Trick}
 
-\section{Feature Extraction}
-\subsubsection{Local Features}
-\subsubsection{Global Features}
+\section{Visual Features}
+\subsection{Local Features}
+\subsection{Global Features}
 
-\section{Semantics}
-\section{Semantic Tagging Mapping}
-\section{Label Propagation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Important sections of this chapter begin here.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Probability Theory}
 
-% On novelty detection and uninformed guess.
 \section{Principle of Maximum Entropy}
+The principle of maximum entropy states that given a set of distributions that
+are coherent with the acquired knowledge, the one which maximizes entropy should
+be picked.
+
+In the case of non-available information this is the uniform distributions.
+In cases where we only know the mean and standard deviation a normal
+distribution shall be picked, etc\dots
 
 \section{Probabilistic Graphical Models}
+\label{sec:graphical-models}
 \subsection{Factor Graphs}
+\subsection{Inference Engines}
+Exact Inference and Approximate Inference methods
+
+% In case we expand in some finding structure in graphs maybe the following will
+% be useful
+\section{Label Propagation}
+\section{Min-Cuts}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Semantic Mapping}
-\section*{}
-\section{Related Work}
+\chapter{Semantic Mapping}\label{chap:semantic-mapping}
+
+% Advantages of Multi-modal approaches
+%
+
+\section{Dora Architecture Overview}
+\subsection{System organization}
+\subsubsection{Sensory Layer}
+\subsubsection{Categorical Layer}
+\subsubsection{Place Layer}
+\subsubsection{Conceptual Layer}
+
+\section{Features}
+\section{Conceptual Knowledge}
 \section{Conceptual Map}
-\section{Spatial-Temporal Accumulation}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Novelty Detection}
-\section*{}
-\section{Related Work}
+\chapter{Novelty Detection}\label{chap:novelty-intro}
 \section{Novelty Detection as a Threshold}
-\section{Using unlabelled Data}
+\subsection{Conditional Probability}
+\subsection{Unconditional Probability}
+
+\section{Semi-Supervised Novelty Detection: Using Unlabelled Data}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
 \section{Detecting New Classes} 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Conclusions and Future Work}
-\section*{}
-
-\section{Conclusions}
+\chapter{Conclusions and Future Work}\label{chap:conclusions}
 \section{Future Work}

### 2011-05-27 thesis/abstract.tex
More thesis structure

@@
 
+% Portuguese Version of the above.
 \chapter*{Resumo}
-Let $E$ denote the set of all possible texts in English and $P$ the set of
-all possible texts in Portuguese. A transformation $T_{E->P}(x)$ can be defined
-in $E \times P$, such that it maps $x \in E$ to a translated version $y \in P$ such
-that they hold the same semantic meaning to a reader.
-
-Assume this text to be $T_{E->P}(abstract)$.
-\footnote{What I mean is that this will be a translation of the english text present
-on abstract and don't expect it to be filled in before the final version of this
-document!}
+Assume this text to be $T_{E \rightarrow P}(abstract)$.
+\footnote{What I mean is that this will be a translation of the English text
+present on abstract. Also don't expect it to be filled in before the final
+version of this document!}

### 2011-05-27 thesis/contents.tex
More thesis structure
Added content from the tech report. Its not at all the final content but it helps structuring.

@@
 \emph{probabilistic graphical models} that lay the base modelling tool for
-the semantic structured representation over which novelty detection
-methods are developed.
+the used structured semantic representation.
 
@@
 \subsection{Support Vector Machines}
+\Glspl{SVM} where introduced by \cite{cortes1995support} and can be seen as
+discriminative linear based classifier. They are based on a strong mathematical
+foundation and have powerful generalization capabilities. In their original form
+\gls{SVM} separates two classes of points in an hyper-space with a
+\emph{maximal margin hyperplane}~(\autoref{fig:svm-sample}).
+
+Later they were extended to deal with noisy data by using \emph{soft margins}.
+And to handle non-linear spaces as seen on \autoref{sec:kernel-trick}.
+Although they only allow for two class-classification several methods have been
+proposed to build multi-class classifiers based on binary classifiers as seen on
+\autoref{sec:multiclass-classifiers}.
+
+\begin{figure}[h]
+\begin{center}
+% TODO: get a decent picture to illustrate SVMs
+\includegraphics[width=0.4\textwidth]{figures/Svm_max_sep_hyperplane_with_margin}
+\end{center}
+\caption{{SVM} separating two class of points by a
+         \emph{maximal margin hyperplane}. The hyperplane can be described by
+         the collection of support vectors and associated weights, marked in the
+         image as sample points with large borders.}
+\label{fig:svm-sample}
+\end{figure}
+
+They have been used in several classification and recognition problems and are
+in fact a standard across machine learning techniques. Their efficiency, exact
+training results and generalization made their suitable for many tasks. Such as
+text categorization, digit-recognition, spam-classification.
+They have also been extensively used in visual place classification.
+
 \subsection{Multi-Class Classification}
-%\subsection{Kernel-Trick}
+\label{sec:multiclass-classifiers}
+\Glspl{SVM} were designed as two-class classifier. And so they need to be
+adapted for using in multi-class problems. The most common usage is to design
+multi-class classifiers recurring at the usage of multiple two-class classifiers
+and methods for combining them. The most typical approaches for a multi-class
+classification on $c$ classes using \glspl{SVM} are:
 
-\section{Visual Features}
+\begin{description}
+\item[One Against All] - in this method $c$ distinct classifiers are trained to
+distinguish any class of the remaining ones. The output of all those classifiers
+(distance to the separating hyperplane) is then used to categorize the output.
+The most common approach is to pick the class with the largest hyperplane
+distance. Other variations exists as is the case of using the minimal distance
+to the average classification distance of each class~\citep{pronobis2007confidence}.
+
+\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to
+distinguish between each pair of classes. The final decision is based on the
+output of all those classifiers being common to use a majority vote strategy.
+\end{description}
+
+\subsection{Kernel-Trick}
+\label{sec:kernel-trick}
+\gls{SVM} in their basic form are only able to handle linear spaces.
+Nonetheless the classes are most of the time not linearly separable in the input
+space. Although there might exists a transformation $\phi$ from the original
+space into a space $H$ where the input becomes linearly separable.
+
+The Kernel trick allows to extend the \gls{SVM} definition to work on such space
+$H$ without ever performing an implicit transformation between spaces. Being
+enough for that to have a Kernel function defining an inner-product inside such
+space: $K(x_i, x_j) = \phi(x_i)\cdot\phi(x_j)$.
+In that sense a Kernel function can be seen as a function that calculates some
+similarity measure between two inputs.
+
+Several kernel functions have been proposed being the most commonly used:
+
+\begin{description}
+\item[Polynomial Kernel] - $K(x, y) = (x \cdot y + p)^d$
+\item[Radial Basis Function] - $K(x, y) = e^{-\gamma\|x - y \|^2}$
+\item[Histogram intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where
+$\chi^2(x,y) = \sum_{i=1}^{N}\frac{(x_i-y_i)}{x_i+y_i}$ introduced by
+\cite{barla2003histogram} allows to compute histogram similarity.
+\item[Matching Kernels] - mimic matching similarity and are used when each
+sample is represented as a set of features~\citep{boughorbel2005intermediate}.
+\end{description}
+
+\section{Features}
+A feature is a piece of information which is expected to reveal information for
+solving a specific task. In that sense features are task-dependant and they will
+yield different performance based on the type of task they are applied to.
+
+A wanted property on features is its repeatability under similar conditions for
+the problem in hand. This is: they should be stable and invariant across
+unwanted types of transformations and noise. For example a visual feature for
+object detection should be present even if the target object was translated,
+scaled, rotated, the light-conditions have changed or even if the object is
+partially occluded.
+
+Extracting features with those properties allows to greatly reduce the size of
+input by removing unwanted noise and useless information from the captured data.
+Turning the classification problem easier, more reliable and more efficient.
+
+Often several and different types of features need to be extracted. It has been
+reported by \cite{pronobis2010ijrr} that using multiple features provides a
+great benefit in the context of place classification.
+And \cite{quattoni2009recognizing} has showed that different types have
+different impact in indoor scene recognition based on the type of scene
+matching. Namely was seen that some room-categories are more likely to be
+recognized by the presence of some objects and others by it generic appearance.
+
+In the context of robotics, sensors such as cameras, laser scans are used to
+sense the surrounding environment, and features can be extracted from all those.
+Nonetheless as described in \autoref{sec:visual-motivation} the visual sensor is
+incredibly rich and most of our features will come from it.
+
+Visual features can be seen as belonging to two categories:
+\emph{local features} and \emph{global features}.
+\emph{Local features} describe fine grain properties of a part of image.
+For example the existence of specific corner or an edge. \Gls{SIFT}
+(\autoref{sec:sift}) is an example of such feature and has been proven useful
+for matching points between images and subsequence extension to object detection.
+
+\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
+{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
+statistical analysis of features over all the image or by a structured
+distribution of textures findable in the image.
+
 \subsection{Local Features}
+\label{sec:local-features}
+
+\subsubsection{SIFT}
+\label{sec:sift}
+The detection of interesting points has been studied for several years and is
+the base of several computer vision problems solution. It allows to perform
+point matching which can be used in several areas from image stitching,
+3D reconstruction, video tracking, object detection, etc\dots
+
+The most used method was presented by \cite{lowe1999object}. And its based on
+building a feature vector for each image. Each of those features is based on
+\emph{interesting points} detected by detecting maxima and minima of a
+difference of Gaussian functions applied in a scale-space.
+The scale space is used to provide scale invariant detection. Gaussian functions
+are used as they are the only way to model a linear scale-space.
+Each interesting point is then described by a container that is rotation
+invariant.
+
+By seeing an object as a set of features points, index and matching is then
+performed by a high-dimensional search on a database of know objects. After
+matching objects can be verified for geometric coherence between features.
+\Gls{SVM} classifiers can also be trained to detect objects based on this type
+of local features by using \emph{Matching kernels} (\autoref{sec:kernel-trick}).
+
+\begin{figure}[h]
+    \includegraphics[width=\textwidth]{figures/sift/sift.pdf}
+    \caption{{SIFT} and other local features have been proven useful in object
+             detection.}
+\end{figure}
+
 \subsection{Global Features}
+
+\subsubsection{Gist of a Scene}
+\label{sec:gist}
+\cite{oliva2006building} argue that fast scene recognition does not need to be
+built on top of object recognition but can be analyzed by scene-centered
+mechanisms.
+They defend that position by pointing out behaviours on human vision:
+when provided with a glance of a shot a person can identify the meaning of that
+given shot or "gist of a scene" without remembering specific details.
+
+\begin{figure}[h]
+\center
+\includegraphics[width=0.60\textwidth]{figures/gist.jpg}
+\caption{An illustration of the gist of an image. Top row: original image I;
+         bottom row: noise image J for which gist(I) = gist(J). We see that the
+         gist captures the dominant textural features of the overall image, and
+         their coarse spatial layout~\citep{murphy2006object}.}
+\end{figure}
+
+\subsubsection{{CRFH} - Composed Receptive Field Histograms}
+\label{sec:crfh}\label{sec:global-features}
+Composed Receptive Field Histograms are a multidimensional statistical
+representation of the occurrence of several image descriptors applied to an
+image. They can be seen as an high-dimension histogram where each cell records
+how many pixels of the image have the cell response for the applied descriptors.
+Such high-dimensional histogram is expected to be able to better global
+information contained in the image by capturing several properties of the image
+as well the relations between them.
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=1\textwidth]{figures/crfh_model.jpg}
+\end{center}
+\caption{A two dimensional histogram of the image built out from two image
+         descriptors: $Lx$ and $Ly$. First-order Gaussian derivatives of image
+         luminance in horizontal and vertical direction applied at a scale 4.}
+\end{figure}
+
+Several type of image descriptors can be applied.
+For example Gaussian derivatives (such as $L_x$, $L_y$, $L_{xx}$, $L_{yy}$) are
+partial derivatives of the image luminance after applying a Gaussian filter of a
+given scale ($\sigma$) on the image. Gradient magnitude descriptors
+(such as $|\nabla L|$) can also be used as a rotation invariant descriptor.
+
+One of the problems of high-dimensional histograms is the memory and
+computationally complexity needed to handle them.
+\cite{linde2004object} suggested using a sparse form to represent those by
+storing only the non-zero-cells in an array structure.
+
+Multidimensional histograms have proven to be useful in the context of object
+recognition~\citep{schiele1996object}. And have also been previously used in
+the context of visual place classification~\cite{pronobis2010ijrr}.
+\emph{Histogram intersection kernels} (\autoref{sec:kernel-trick}) can be used
+to train classifiers based on this feature.
+
+\subsection{Laser Features}
 
@@
 \section{Probability Theory}
+\subsection{Bayes Theorem}
+
+\section{Probabilistic Graphical Models}
+\label{sec:graphical-models}
+\subsection{Factor Graphs}
+\subsection{Inference Engines}
+Exact Inference and Approximate Inference methods
 
@@
 
-\section{Probabilistic Graphical Models}
-\label{sec:graphical-models}
-\subsection{Factor Graphs}
-\subsection{Inference Engines}
-Exact Inference and Approximate Inference methods
-
 % In case we expand in some finding structure in graphs maybe the following will
 % be useful
-\section{Label Propagation}
-\section{Min-Cuts}
+\section{Label Propagation, Min-Cuts}
 

### 2011-05-27 thesis/contents.tex
Added a lot of content from the techreport and some more structed and ideas.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
 \section{Probability Theory}
@@
 \label{sec:graphical-models}
+% THIS IS ALL JUNK TEXT COPIED
+Graphical models usage can be tracked back to earlies 1920 but they only become
+popular in mid-eighties when researchers started to use \emph{Bayesian Networks}
+to model expert systems~\citep{borgelt2002graphical}.
+
+They serve as a better tool to model \emph{random variables}
+(nodes on the graph) and their probabilities as they model the conditional
+dependence between variables (edges on the graph). Important to note that here
+\emph{random variables} does not denotes a truly random variable but one that is
+unknown by the system and is conditioned by other variables/evidences.
+
+This type of graphs provide a generative model where the probability of any
+given scenario can be determined. As opposed to non-generative models where a
+given probability can only be calculated if certain data is given.
+This means when a graphical model is learned, it can be used to generate new
+samples from the learned distribution.
+
+They have been successfully used in several machine-learning task such as:
+information extraction, speech recognition, computer vision.
+They are also useful due to their ability to deal with semantic (high-level)
+features~\citep{boutell2006factor} and ability to represent properties of the
+reality they try to model.
+
+Two main types of graphical models have been used: Bayesian Networks which model
+directed edges between variables and Markov Random Fields where variables are
+connected by a potential but no special direction is given to edges.
+
+An important property of these graphs is the \emph{Markov-blanket} of a node.
+For a given variable $a$, a \emph{Markov-blanket} is a set of variables in the
+surroundings of $a$ that when given the value of $a$ becomes independent of the
+rest of the graph~\citep{pearl1988probabilistic}.
+On non-directional graphs it is directly determined by the nodes connected to $a$.
+This allows the usage of graph algorithms such as \emph{min-cut} to quickly
+determine most likely scenarios. In the case of directed edges a node blanket is
+also influenced by the direction of the edges and a more complex schemes need to
+be used.
+
+As \cite{lauritzen2002chain} points this two types of models can be represented
+as a \emph{chain graph} where both directed and undirected edges can co-exist.
+Though this generalization is hard to implement due to mis-understandings on the
+concepts the graph-models use.
+
+Another useful interpretation of graphical-models are \emph{factor graphs}.
+Those are able to handle both \emph{Bayesian Networks} and
+\emph{Markov Random Fields}. Under this interpretation a graph is seen as a
+bipartite graph that connects variables with factors that influence
+them~\citep{bishop2006pattern}.
+This gives a very useful framework to develop belief propagation on them by
+seeing a message-passing mechanism between nodes. Belief propagation is used to
+calculate marginal-probabilities.
+
+\begin{figure}[ht]
+    \begin{minipage}[b]{0.3\linewidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/graphical-models/BayesNet.pdf}
+        \caption{Bayes Networks}
+    \end{minipage}
+    \hspace{0.5cm}
+    \begin{minipage}[b]{0.3\linewidth}
+        \centering
+        \includegraphics[width=0.8\textwidth]{figures/graphical-models/MarkovRandomField.pdf}
+        \caption{Markov Random Field}
+    \end{minipage}
+    \hspace{0.5cm}
+    \begin{minipage}[b]{0.3\linewidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/graphical-models/FactorGraph.pdf}
+        \caption{Factor Graph}
+    \end{minipage}
+\end{figure}
+
 \subsection{Factor Graphs}
+A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
+representing random variables and factors.
+Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
+to which the factor is connected.
+Thus, a factor graph can be seen as a description of probability density function obtained
+by a product of all the factors. In order to represent the probability,
+a normalization factor needs to be introduced, resulting in the following equation:
+
+\begin{equation}
+P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
+Z = \sum_{X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
+\end{equation}
+
+More often we are only interested in a marginal of the probability distribution.
+
+\begin{equation}
+M_G(x) = \frac{1}{Z}\sum_{X_G \backslash x}{\prod_{\phi \in F_G}{\phi(x_{\phi})}}
+\end{equation}
+
 \subsection{Inference Engines}
@@
 \chapter{Novelty Detection}\label{chap:novelty-intro}
+\begin{itemize}
+\item Introduction
+\item History
+\item Approaches to Novelty Detection
+\end{itemize}
+\section{Novelty Detection Review and Related Work}
+
 \section{Novelty Detection as a Threshold}
+When dealing with a statistic point of view, noisy data or unstable features, a decision to
+classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
+And the error rate of a system that classifies the set of inputs $N$ as novel is given by
+the equation:
+
+\begin{equation}
+P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+\end{equation}
+
+Therefore any novelty system that decides to acknowledge a sample
+$a$ as novel should also detect novelty on any sample $b$ satisfying
+$P(\overline{novel}|b) < P(\overline{novel}|a)$.
+As that would always reduce the error probability.
+This way a novelty detection system is interested in defining an
+order relation on all the possible inputs equivalent to the order defined by
+the error rate: $P(\overline{novel}|x)$.
+
+Using Bayes rule and assuming a constant $P(\overline{novel})$
+a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
+Such a ratio is a suitable function for implementing a novelty detector system with optimal
+thresholding.
+
+\begin{equation}
+\label{eq:novelty-threshold}
+          P(\overline{novel}|x)
+  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
+  \propto \frac{P(x|\overline{novel})}{P(x)}
+\end{equation}
+
 \subsection{Conditional Probability}
@@
 \chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
+On the previous chapter, novelty as been introduced as a statistical
+measure on the error-likelihood ratio between a conditional and unconditional
+probability of a given .
+
 \section{Detecting New Classes}
+Explain exactly and formally how to obtain a novelty measure on a single
+variable of a graph.\footnote{I am sure on the math of this.}
+
+\section{Copping with Multiple Novel Variables}
+On the previous section we show how to detect a novel class but that assumes
+the rest of the graph holds true.
+
+In this section we show how to handle novelty in several variables
+in simultaneous.\footnote{I am almost sure the math here will work out the
+way I think}
+
+\subsection{Handling novelty signals from the low-level classifiers}
+Explain how to handle novelty signals from low-level classifiers, now that we
+have a model with the ability to still reason when several variables can be
+novel.
+
 \section{Detecting New Structure}

### 2011-05-29 thesis/contents.tex
Lots of changes in the novelty-intro chapter. With information from the semantic-nd article.

@@
 
+\subsection{Categorical and Multinomial Distribution}
+
 \section{Probabilistic Graphical Models}
@@
 \section{Principle of Maximum Entropy}
+\label{sec:max-entropy}
 The principle of maximum entropy states that given a set of distributions that
@@
 \chapter{Novelty Detection}\label{chap:novelty-intro}
+
+This chapter presents a statis
+
+Laying down base concepts that are then used in \autoref{chap:novelty} to
+implement novelty detection on the semantic spatial representation.
+
+It start by presenting a brief overview on novelty detection.
+
+It introduces novelty as a
+
+Shows novelty has an associated ordering.
+Shows novelty has a threshold function.
+
+At last a practical example using semantic data and probabilistic graphical
+models is presented: it shows how to use the introduced ratio to obtain a
+novelty detection system and analyses the performance impact by obtaining more
+data and by using an approximation on the unconditional probability.
+
 \begin{itemize}
@@
 \end{itemize}
+
 \section{Novelty Detection Review and Related Work}
+% Introduce novelty detection
+Novelty detection deals with detecting that a new input was generated by a class
+other than the ones the system knows about~\cite{markou2003novelty}.
+It is harder than classification as only positive samples of a class are available
+rendering normal classification methods invalid.
+
+Adding capabilities to an agent to detect novel samples allows to increase its
+reliability. Novelty signal can then be interpreted by the system to proceed
+with more caution as its knowledge does not correctly describe reality.
+
+% Explain threshold approach to novelty detection
+Due to the desire of robustness, a novelty detection system is often given the option
+to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
+between error and rejection: the more the system tries to reduce rejection, the more
+prone to produce errors it becomes.
 
 \section{Novelty Detection as a Threshold}
-When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
-And the error rate of a system that classifies the set of inputs $N$ as novel is given by
-the equation:
+When dealing with a statistic point of view, noisy data or unstable features,
+a decision to classify a sample $n$ as novel has an associated error probability
+$P(\overline{novel}|n)$.
+Accordingly, on a system that classifies a set of inputs $N$ as novel, the
+probability of incorrectly classifying a sample as novel is given by the
+equation:
 
 \begin{equation}
-P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+\label{eq:false-positive}
+P(\textnormal{false positive}|\textnormal{positive}) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
 \end{equation}
@@
 $P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that would always reduce the error probability.
-This way a novelty detection system is interested in defining an
-order relation on all the possible inputs equivalent to the order defined by
-the error rate: $P(\overline{novel}|x)$.
-
-Using Bayes rule and assuming a constant $P(\overline{novel})$
-a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
-Such a ratio is a suitable function for implementing a novelty detector system with optimal
-thresholding.
+As that always reduce the error probability described in
+\autoref{eq:false-positive} as shown:
 
 \begin{equation}
-\label{eq:novelty-threshold}
+\textnormal{Proof: to be optimal any system accepting }a\textnormal{ should also accept }b
+\end{equation}
+
+For the sake of completeness its show here that it also reduces the probability
+of false negatives.
+
+\begin{equation}
+P(\textnormal{false negative}|\textnormal{negative}) = \frac{\sum_{n \notin N}{P(novel|n)P(n)}}{\sum_{n \notin N}{P(n)}}
+\end{equation}
+
+Based on this, it can be said that an optimal novelty detection system is
+interested in defining an order relation on all the possible inputs equivalent
+to the order defined by the error rate: $P(\overline{novel}|x)$.
+And any optimal detector can be described by the largest $P(\overline{novel}|x)$
+accepted by it. Which is seen as threshold.
+
+\section{Conditional and Unconditional Probability Ratio}
+
+On the previous section it was shown that an optimal novelty detector can be
+implemented with a threshold on top of the order-relation defined by
+$P(\overline{novel}|x)$ over $x$. Performing some manipulations with
+Bayes theorem and assuming a constant $P(\overline{novel})$ a more usable
+form can be attained:
+
+\begin{equation}
+\label{eq:novelty-ratio}
           P(\overline{novel}|x)
@@
 
+Since there is only interest in maintaining the same order-relation as
+$P(\overline{novel}|x)$ the absolute value is not needed and any constant
+factor can be dropped. Leaving a ratio between a \emph{conditional} and
+\emph{unconditional probabilities} suitable for implementing novelty detection
+via  thresholding.
+
 \subsection{Conditional Probability}
+The conditional probability $P(x|\overline{novel})$ describes the distribution
+of the non-novel samples. In case the labelled data available for the agent
+to learn a concept comes from this same distribution the correct approach
+is to use it as prior-information for modelling the conditional probability.
+
+Note that it is important for the labelled data to come from the same
+distribution as where the system will run, as otherwise it will contain some
+bias (for example introduced by an incorrect filtering)
+
+Therefore the correct approach is to try to model it
+
 \subsection{Unconditional Probability}
+The unconditional probability $P(x)$ plays an important role on obtaining a
+correct order relation for performing novelty detection.
+It serves as a normalizing component that allows the system to  whether
+a given sample conditional probability arises from it belonging to the
+known concept or from it likelihood of being sampled.
 
-\section{Semi-Supervised Novelty Detection: Using Unlabelled Data}
+On lack of any information about the unconditional probability and conforming to
+the principle of maximum entropy (\autoref{sec:max-entropy}) a uniform
+distribution must be chosen.
+Though, as discussed in \autoref{sec:unlabelled-data}, by using unlabelled data
+it becomes possible to obtain prior-information and achieve a better
+approximation.
+
+%\subsection{Novelty Detection on a Variable Set of Features}
+Note that very often, novelty detection is applied on a fixed set of features
+together with an assumption of a uniform unconditional probability.
+On those cases $P(x)$ becomes a constant and therefore a novelty threshold
+can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
+But in the case where the set of features $x$ is variable it cannot be
+discarded. There $P(x)$ plays a role in levering all the conditional
+probabilities on different sets of variables into the same measure units
+such that a threshold can be implemented.
+
+\section{A Practical Example}
+\label{sec:unlabelled-data}
+In order to verify the performance of the proposed threshold functions a synthetic dataset
+was generated. To keep it simple only information regarding direct features of
+a room were modelled and no structured knowledge such as room connectivity was taken
+in account.
+The synthetic distribution assumes that an independent and variable size set of features
+$x$ is generated by a given room category.
+In whole there was 11 different room categories and 7 different measured feature
+types. Each feature type can be present more than once in (e.g.: room shape is
+extracted from 2D laser scans in more than one position in the room).
+
+The room categories were chosen to mimic as close as possible the real features and
+categories existing in reality. And they model different room categories with
+different levels of detail. For example: 1 person office, 2 person office, hallway,
+robot lab.
+
+From the distribution, 100 labelled samples for 5 of the 11 room categories were
+drawn to represent the known concepts and 1000 unlabelled samples were drawn from
+all the room categories for learning the unconditional probability distribution.
+Using those samples, factors were learnt for the graphs used to model the
+conditional distribution and the independent unconditional distribution.
+\autoref{fig:simple-experiment} shows the graph structure used for approximate the
+trained conditional and unconditional distributions.
+Its important to notice that graph $G$ used to model the known classes when given
+enough labelled data is able to exactly learn the conditional distribution as it
+uses the same structure as the created synthetic distribution.
+
+\begin{figure}[h]
+\centering
+
+\subfloat[Graph structure $G$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-cond-graph.pdf}}
+\qquad
+\subfloat[Graph structure $U$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-uniform-graph.pdf}}
+\qquad
+\subfloat[Graph structure $I$.]{
+         \includegraphics[width=0.27\textwidth]{figures/simple-independent-graph.pdf}}
+
+\caption{\label{fig:simple-experiment}The graph structures used to model the
+         conditional and unconditional probability for implementing the novelty
+         thresholds $P_G(x)/P_U(x)$ and $P_G(x)/P_I(x)$.}
+\end{figure}
+
+% Describe how to obtain the 3 thresholds functions seen on the graphs.
+Using the learned models $G$, $U$ and $I$, two thresholds were trained:
+$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
+and $P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
+Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
+and a perfect threshold function could also be created to test how far the
+presented thresholds are from optimal.
+
+\subsection{Probability Ratio Comparison}
+%%% Results 1
+% Show the threshold ratio is an optimal detector (if perfect information was available)
+% Show that the thresholds are suitable functions for implementing a static threshold.
+As first tryout, the performance of the novelty threshold selection was plotted for a set
+of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
+Those samples contain different number of sensed properties for each room, mimicking
+the dynamic properties expected to see when implemented on a robot.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.60\textwidth]{results/synthetic-all.pdf}
+
+\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
+         under samples with variable size of sensed properties.}
+\end{figure}
+
+The convex shape for the optimal threshold shows that the ratio between conditional
+and unconditional probability is indeed an optimal detector and is suitable for
+implementing a threshold when the samples are taken from dynamic
+distributions (e.g.: some samples where there is only access to room size versus
+samples where there is a lot of information about the room properties).
+
+% Discuss importance on approximating unconditional probability.
+Its also possible to see how important it is to estimate a correct unconditional
+probability in order to obtain a correct novelty measure on the inputs.
+The assumption of a uniform unconditional probability has led to very poor results.
+That is probably explained by the semantic properties being highly
+biased towards some values. And shows that bias plays an important step
+in detecting whether a given sensed value is a valuable cue about the room category.
+
+%%% Results 2
+% Measure performance of the thresholds as more information becomes available.
+\subsection{Performance Changes With Amount of Available Information}
+In order to measure the performance impact as more semantic information becomes
+available ROC curves were plotted for samples grouped by the number of sensed
+semantic features.
+
+\begin{figure}[h]
+\centering
+
+\subfloat[3 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-3features.pdf}}
+\qquad
+\subfloat[5 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-5features.pdf}}
+
+\subfloat[10 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-10features.pdf}}
+\qquad
+\subfloat[50 sensed features]{\includegraphics[width=0.40\textwidth]{results/synthetic-50features.pdf}}
+
+\caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
+         presented novelty detection method on graphs generated for different amount of
+         sensed features.}
+\end{figure}
+
+We can see that as the system gains more semantic information it becomes easier
+to detect novelty. The input space size increases and allows the several existing
+classes to become more easily distinguished.
+
+The performance of the independent threshold decreases as the number of sensed
+features increases. This is easily explained by the fact that the graph $I$ is not
+able to model the existent dependence between the features. This becomes obvious
+as the number of features increases (e.g.: graph $I$ perfectly models $P(x)$ in the
+case where only 1 feature is sensed).
+
+The uniform threshold shows a poor performance specially on small size samples
+where its performs almost no better than random.
+It performance increases as the size of sensed features increases but nonetheless
+its very small when compared to how optimal a threshold could be.
 

### 2011-05-29 thesis/background.tex
Splitted the chapters that are now some how big into other files.

--- 2011-05-27 thesis/contents.tex
@@
-% What should the title be?
-% Novelty Detection for Visual Indoor Categorization
-% Novelty Detection on Semantic Representations.
-
-\chapter{Introduction}
-% Introduce:
-%  Mobile Robotics, A.I.
-%  Interaction with Humans
-%  Need for concepts/semantic information
-%  Dynamic human environment
-%  Non-feasibility of hard-coded concepts
-%  Requirements to detect novelty and handle it.
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us. Understand our concepts and
-be able to reason.
-
-They need to possess ability to adapt to situations as its infeasible to rely
-on extensive man-work to tag objects, map space and code all the aspects that
-make up our human-reality.
-
-% == Introduce the need of semantic representations ==
-%
-% There is a lot of low-level methods (or specific knowledge to robots)
-% But the creation of semantic representations allow to bridge that low-level
-% sensing with high-level concepts that facilitate several high-level Tasks.
-
-% Importance of mapping and localization in a robot as well as reasoning on
-% properties of each space.
-% Humans give labels to space that characterized the properties and activities
-% expected to be performed on them.
-% Such information is of great interest to be structured and organized in such
-% a way reasoning and planning can be implemented.
-%
-% It also holds an important role in long-term planning and stability as it
-% hides the space-time local details and allows to focus on high-level thinking.
-
-% Deal with uncertainty -> need for probabilistic models.
-% Deal with semantic    -> need of structured models such as graphical models.
-
-Example of what kind of an high-level thinking on space and its semantics can
-provide: Get me a beer/milk concept.
-
-% == Reasoning about need for Novelty Detection ==
-%
-% Dynamic environment (inability to know the environment or even to be possible
-% to map all semantic concepts to interact with humans).
-%
-% Knowledge-Awareness (knowing what we know... And detecting novel cases)
-% Identify gaps in the semantic knowledge.
-% Automatic detection and learning of novel concepts.
-%
-% Increase robustness
-% Self-Extending
-%
-
-\section{Problem and Goals}
-% Extend a probabilistic graphical modelling framework with capabilities to
-% detect novelty.
-% Both in terms of novel classes as of novel structure.
-%
-Given a probabilistic structure representing the sensed conceptual knowledge
-obtained by the agent develop methods to be able to detect novelty present
-either on new semantic concepts (new classes) or even on structure that was
-previously unknown.
-
-% Yeah thats what you call dream :P
-In an extreme ideal case the agent should be able to go from zero-knowledge to
-understanding presence of certain objects (as generators of a set of sensed
-sensed properties grouped locally), understand areas within rooms (sink-area on
-on a kitchen), rooms as separated by doors, and understand environments such
-as office, home, warehouse, spaceship.
-
-For that the probabilistic semantic representation presented by
-\cite{andrzej2011phd} is used.
-
-% If everything goes fine... 2 of the "Future Directions" proposed are 
-% Novelty Detection and Learning of Novel Concepts:
-%  Identify Gaps in Spatial and Semantic Knowledge -> Addressed
-%  Performing learning of new concepts -> Outside Scope / No Time
-%
-% Using Properties for Space Segmentation
-%  If we move on detecting novel structures on the graph we are addressing this
-%  issue.
-
-\section{Thesis Outline}
-The rest of this thesis is structured as follows:
-
-\autoref{chap:background} introduces background concepts for understanding the
-presented thesis. In particular \autoref{sec:graphical-models} introduces
-\emph{probabilistic graphical models} that lay the base modelling tool for
-the used structured semantic representation.
-
-\autoref{chap:semantic-mapping} describes the system proposed by
-\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
-that is responsible for the creation of the \emph{probabilistic semantic
-representation} that this thesis aims at improving by developing methods with
-the capability to identify knowledge gaps.
-
-\autoref{chap:novelty-intro} introduces novelty detection under a
-statistical view point and how to interpret it as a threshold function.
-A simple approach on how to perform novelty on very simple graphs
-is given together with some results on the impact of using semi-supervised
-novelty detection to improve the system performance.
-
-\autoref{chap:novelty} presents the developed techniques developed to identify
-novelty on the semantic representation.
-
-\autoref{chap:conclusions} draws conclusions on the developed work and presents
-interesting directions for future works.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \chapter{Background}\label{chap:background}
@@
 \section{Label Propagation, Min-Cuts}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Semantic Mapping}\label{chap:semantic-mapping}
-
-% Advantages of Multi-modal approaches
-%
-
-\section{Dora Architecture Overview}
-\subsection{System organization}
-\subsubsection{Sensory Layer}
-\subsubsection{Categorical Layer}
-\subsubsection{Place Layer}
-\subsubsection{Conceptual Layer}
-
-\section{Features}
-\section{Conceptual Knowledge}
-\section{Conceptual Map}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Novelty Detection}\label{chap:novelty-intro}
-\begin{itemize}
-\item Introduction
-\item History
-\item Approaches to Novelty Detection
-\end{itemize}
-\section{Novelty Detection Review and Related Work}
-
-\section{Novelty Detection as a Threshold}
-When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
-And the error rate of a system that classifies the set of inputs $N$ as novel is given by
-the equation:
-
-\begin{equation}
-P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
-\end{equation}
-
-Therefore any novelty system that decides to acknowledge a sample
-$a$ as novel should also detect novelty on any sample $b$ satisfying
-$P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that would always reduce the error probability.
-This way a novelty detection system is interested in defining an
-order relation on all the possible inputs equivalent to the order defined by
-the error rate: $P(\overline{novel}|x)$.
-
-Using Bayes rule and assuming a constant $P(\overline{novel})$
-a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
-Such a ratio is a suitable function for implementing a novelty detector system with optimal
-thresholding.
-
-\begin{equation}
-\label{eq:novelty-threshold}
-          P(\overline{novel}|x)
-  =       \frac{P(x|\overline{novel}) P(\overline{novel})}{P(x)}
-  \propto \frac{P(x|\overline{novel})}{P(x)}
-\end{equation}
-
-\subsection{Conditional Probability}
-\subsection{Unconditional Probability}
-
-\section{Semi-Supervised Novelty Detection: Using Unlabelled Data}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
-On the previous chapter, novelty as been introduced as a statistical
-measure on the error-likelihood ratio between a conditional and unconditional
-probability of a given .
-
-\section{Detecting New Classes}
-Explain exactly and formally how to obtain a novelty measure on a single
-variable of a graph.\footnote{I am sure on the math of this.}
-
-\section{Copping with Multiple Novel Variables}
-On the previous section we show how to detect a novel class but that assumes
-the rest of the graph holds true.
-
-In this section we show how to handle novelty in several variables
-in simultaneous.\footnote{I am almost sure the math here will work out the
-way I think}
-
-\subsection{Handling novelty signals from the low-level classifiers}
-Explain how to handle novelty signals from low-level classifiers, now that we
-have a model with the ability to still reason when several variables can be
-novel.
-
-\section{Detecting New Structure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Conclusions and Future Work}\label{chap:conclusions}
-\section{Future Work}

### 2011-05-29 thesis/contents.tex
Splitted the chapters that are now some how big into other files.

--- 2011-05-27 thesis/contents.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Background}\label{chap:background}
-% This chapter will perform a breath-first explanation on the contents related
-% with the thesis. The extension of the contents explained here can be seen as
-% monotone function over the number of missing pages to attain the requirements.
-%
-% Some topics are indeed important as its the case of factor graphs.
-% Others like:
-%  - classification (SVM, multi-class, kernel-trick),
-%  - features (SIFT, GIST),
-%  - Basic Bayes theory and probabilistic review
-% are included for fun of the reader in case it inquires himself how to deal
-% with the low-levels features we propose using.
-%
-% ---
-% PS: This is the chapter that will be used to fill with unnecessary crap
-% in case they really annoy me with the number of pages. Every other chapter
-% will strive to be really required and be a masterpiece.
-%
-
-\section{Classification}
-\subsection{Recognition and Categorization}
-\subsection{Support Vector Machines}
-\Glspl{SVM} where introduced by \cite{cortes1995support} and can be seen as
-discriminative linear based classifier. They are based on a strong mathematical
-foundation and have powerful generalization capabilities. In their original form
-\gls{SVM} separates two classes of points in an hyper-space with a
-\emph{maximal margin hyperplane}~(\autoref{fig:svm-sample}).
-
-Later they were extended to deal with noisy data by using \emph{soft margins}.
-And to handle non-linear spaces as seen on \autoref{sec:kernel-trick}.
-Although they only allow for two class-classification several methods have been
-proposed to build multi-class classifiers based on binary classifiers as seen on
-\autoref{sec:multiclass-classifiers}.
-
-\begin{figure}[h]
-\begin{center}
-% TODO: get a decent picture to illustrate SVMs
-\includegraphics[width=0.4\textwidth]{figures/Svm_max_sep_hyperplane_with_margin}
-\end{center}
-\caption{{SVM} separating two class of points by a
-         \emph{maximal margin hyperplane}. The hyperplane can be described by
-         the collection of support vectors and associated weights, marked in the
-         image as sample points with large borders.}
-\label{fig:svm-sample}
-\end{figure}
-
-They have been used in several classification and recognition problems and are
-in fact a standard across machine learning techniques. Their efficiency, exact
-training results and generalization made their suitable for many tasks. Such as
-text categorization, digit-recognition, spam-classification.
-They have also been extensively used in visual place classification.
-
-\subsection{Multi-Class Classification}
-\label{sec:multiclass-classifiers}
-\Glspl{SVM} were designed as two-class classifier. And so they need to be
-adapted for using in multi-class problems. The most common usage is to design
-multi-class classifiers recurring at the usage of multiple two-class classifiers
-and methods for combining them. The most typical approaches for a multi-class
-classification on $c$ classes using \glspl{SVM} are:
-
-\begin{description}
-\item[One Against All] - in this method $c$ distinct classifiers are trained to
-distinguish any class of the remaining ones. The output of all those classifiers
-(distance to the separating hyperplane) is then used to categorize the output.
-The most common approach is to pick the class with the largest hyperplane
-distance. Other variations exists as is the case of using the minimal distance
-to the average classification distance of each class~\citep{pronobis2007confidence}.
-
-\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to
-distinguish between each pair of classes. The final decision is based on the
-output of all those classifiers being common to use a majority vote strategy.
-\end{description}
-
-\subsection{Kernel-Trick}
-\label{sec:kernel-trick}
-\gls{SVM} in their basic form are only able to handle linear spaces.
-Nonetheless the classes are most of the time not linearly separable in the input
-space. Although there might exists a transformation $\phi$ from the original
-space into a space $H$ where the input becomes linearly separable.
-
-The Kernel trick allows to extend the \gls{SVM} definition to work on such space
-$H$ without ever performing an implicit transformation between spaces. Being
-enough for that to have a Kernel function defining an inner-product inside such
-space: $K(x_i, x_j) = \phi(x_i)\cdot\phi(x_j)$.
-In that sense a Kernel function can be seen as a function that calculates some
-similarity measure between two inputs.
-
-Several kernel functions have been proposed being the most commonly used:
-
-\begin{description}
-\item[Polynomial Kernel] - $K(x, y) = (x \cdot y + p)^d$
-\item[Radial Basis Function] - $K(x, y) = e^{-\gamma\|x - y \|^2}$
-\item[Histogram intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where
-$\chi^2(x,y) = \sum_{i=1}^{N}\frac{(x_i-y_i)}{x_i+y_i}$ introduced by
-\cite{barla2003histogram} allows to compute histogram similarity.
-\item[Matching Kernels] - mimic matching similarity and are used when each
-sample is represented as a set of features~\citep{boughorbel2005intermediate}.
-\end{description}
-
-\section{Features}
-A feature is a piece of information which is expected to reveal information for
-solving a specific task. In that sense features are task-dependant and they will
-yield different performance based on the type of task they are applied to.
-
-A wanted property on features is its repeatability under similar conditions for
-the problem in hand. This is: they should be stable and invariant across
-unwanted types of transformations and noise. For example a visual feature for
-object detection should be present even if the target object was translated,
-scaled, rotated, the light-conditions have changed or even if the object is
-partially occluded.
-
-Extracting features with those properties allows to greatly reduce the size of
-input by removing unwanted noise and useless information from the captured data.
-Turning the classification problem easier, more reliable and more efficient.
-
-Often several and different types of features need to be extracted. It has been
-reported by \cite{pronobis2010ijrr} that using multiple features provides a
-great benefit in the context of place classification.
-And \cite{quattoni2009recognizing} has showed that different types have
-different impact in indoor scene recognition based on the type of scene
-matching. Namely was seen that some room-categories are more likely to be
-recognized by the presence of some objects and others by it generic appearance.
-
-In the context of robotics, sensors such as cameras, laser scans are used to
-sense the surrounding environment, and features can be extracted from all those.
-Nonetheless as described in \autoref{sec:visual-motivation} the visual sensor is
-incredibly rich and most of our features will come from it.
-
-Visual features can be seen as belonging to two categories:
-\emph{local features} and \emph{global features}.
-\emph{Local features} describe fine grain properties of a part of image.
-For example the existence of specific corner or an edge. \Gls{SIFT}
-(\autoref{sec:sift}) is an example of such feature and has been proven useful
-for matching points between images and subsequence extension to object detection.
-
-\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
-{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
-statistical analysis of features over all the image or by a structured
-distribution of textures findable in the image.
-
-\subsection{Local Features}
-\label{sec:local-features}
-
-\subsubsection{SIFT}
-\label{sec:sift}
-The detection of interesting points has been studied for several years and is
-the base of several computer vision problems solution. It allows to perform
-point matching which can be used in several areas from image stitching,
-3D reconstruction, video tracking, object detection, etc\dots
-
-The most used method was presented by \cite{lowe1999object}. And its based on
-building a feature vector for each image. Each of those features is based on
-\emph{interesting points} detected by detecting maxima and minima of a
-difference of Gaussian functions applied in a scale-space.
-The scale space is used to provide scale invariant detection. Gaussian functions
-are used as they are the only way to model a linear scale-space.
-Each interesting point is then described by a container that is rotation
-invariant.
-
-By seeing an object as a set of features points, index and matching is then
-performed by a high-dimensional search on a database of know objects. After
-matching objects can be verified for geometric coherence between features.
-\Gls{SVM} classifiers can also be trained to detect objects based on this type
-of local features by using \emph{Matching kernels} (\autoref{sec:kernel-trick}).
-
-\begin{figure}[h]
-    \includegraphics[width=\textwidth]{figures/sift/sift.pdf}
-    \caption{{SIFT} and other local features have been proven useful in object
-             detection.}
-\end{figure}
-
-\subsection{Global Features}
-
-\subsubsection{Gist of a Scene}
-\label{sec:gist}
-\cite{oliva2006building} argue that fast scene recognition does not need to be
-built on top of object recognition but can be analyzed by scene-centered
-mechanisms.
-They defend that position by pointing out behaviours on human vision:
-when provided with a glance of a shot a person can identify the meaning of that
-given shot or "gist of a scene" without remembering specific details.
-
-\begin{figure}[h]
-\center
-\includegraphics[width=0.60\textwidth]{figures/gist.jpg}
-\caption{An illustration of the gist of an image. Top row: original image I;
-         bottom row: noise image J for which gist(I) = gist(J). We see that the
-         gist captures the dominant textural features of the overall image, and
-         their coarse spatial layout~\citep{murphy2006object}.}
-\end{figure}
-
-\subsubsection{{CRFH} - Composed Receptive Field Histograms}
-\label{sec:crfh}\label{sec:global-features}
-Composed Receptive Field Histograms are a multidimensional statistical
-representation of the occurrence of several image descriptors applied to an
-image. They can be seen as an high-dimension histogram where each cell records
-how many pixels of the image have the cell response for the applied descriptors.
-Such high-dimensional histogram is expected to be able to better global
-information contained in the image by capturing several properties of the image
-as well the relations between them.
-
-\begin{figure}[h]
-\begin{center}
-\includegraphics[width=1\textwidth]{figures/crfh_model.jpg}
-\end{center}
-\caption{A two dimensional histogram of the image built out from two image
-         descriptors: $Lx$ and $Ly$. First-order Gaussian derivatives of image
-         luminance in horizontal and vertical direction applied at a scale 4.}
-\end{figure}
-
-Several type of image descriptors can be applied.
-For example Gaussian derivatives (such as $L_x$, $L_y$, $L_{xx}$, $L_{yy}$) are
-partial derivatives of the image luminance after applying a Gaussian filter of a
-given scale ($\sigma$) on the image. Gradient magnitude descriptors
-(such as $|\nabla L|$) can also be used as a rotation invariant descriptor.
-
-One of the problems of high-dimensional histograms is the memory and
-computationally complexity needed to handle them.
-\cite{linde2004object} suggested using a sparse form to represent those by
-storing only the non-zero-cells in an array structure.
-
-Multidimensional histograms have proven to be useful in the context of object
-recognition~\citep{schiele1996object}. And have also been previously used in
-the context of visual place classification~\cite{pronobis2010ijrr}.
-\emph{Histogram intersection kernels} (\autoref{sec:kernel-trick}) can be used
-to train classifiers based on this feature.
-
-\subsection{Laser Features}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Important sections of this chapter begin here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Probability Theory}
-\subsection{Bayes Theorem}
-
-\section{Probabilistic Graphical Models}
-\label{sec:graphical-models}
-\subsection{Factor Graphs}
-\subsection{Inference Engines}
-Exact Inference and Approximate Inference methods
-
-\section{Principle of Maximum Entropy}
-The principle of maximum entropy states that given a set of distributions that
-are coherent with the acquired knowledge, the one which maximizes entropy should
-be picked.
-
-In the case of non-available information this is the uniform distributions.
-In cases where we only know the mean and standard deviation a normal
-distribution shall be picked, etc\dots
-
-% In case we expand in some finding structure in graphs maybe the following will
-% be useful
-\section{Label Propagation, Min-Cuts}
+\input{background}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Novelty Detection}\label{chap:novelty-intro}
-\section{Novelty Detection as a Threshold}
-\subsection{Conditional Probability}
-\subsection{Unconditional Probability}
-
-\section{Semi-Supervised Novelty Detection: Using Unlabelled Data}
+\input{novelty-intro}
 

### 2011-05-29 thesis/novelty-intro.tex
Splitted the chapters that are now some how big into other files.

--- 2011-05-17 semantic-nd/article.tex
@@
-\documentclass[runningheads,a4paper]{llncs}
+\chapter{Novelty Detection}\label{chap:novelty-intro}
 
-\usepackage{amssymb}
-\setcounter{tocdepth}{3}
-\usepackage{graphicx}
+This chapter presents a statis
 
-\usepackage{url}
-\newcommand{\keywords}[1]{\par\addvspace\baselineskip
-\noindent\keywordname\enspace\ignorespaces#1}
+Laying down base concepts that are then used in \autoref{chap:novelty} to
+implement novelty detection on the semantic spatial representation.
 
-% Additional packages
-\usepackage[utf8]{inputenc}
-\usepackage{subfig}
-\usepackage[pdftex,pdfpagelabels,bookmarks,hyperindex,hyperfigures]{hyperref}
-\hypersetup{%
-   plainpages=false, 
-   pdfpagelayout=SinglePage,
-   bookmarksopen=false,
-   bookmarksnumbered=true,
-   breaklinks=true,
-   linktocpage,
-   colorlinks=true,
-   linkcolor=blue,
-   urlcolor=blue,
-   citecolor=blue,
-   anchorcolor=green
-}
+It start by presenting a brief overview on novelty detection.
 
-\begin{document}
+It introduces novelty as a
 
-\mainmatter  % start of an individual contribution
+Shows novelty has an associated ordering.
+Shows novelty has a threshold function.
 
-% first the title is needed
-\title{Novelty Detection Using Graphical Models for Semantic Room Classification}
+At last a practical example using semantic data and probabilistic graphical
+models is presented: it shows how to use the introduced ratio to obtain a
+novelty detection system and analyses the performance impact by obtaining more
+data and by using an approximation on the unconditional probability.
 
-% a short form should be given in case it is too long for the running head
-% \titlerunning{Lecture Notes in Computer Science: Authors' Instructions}
+\begin{itemize}
+\item Introduction
+\item History
+\item Approaches to Novelty Detection
+\end{itemize}
 
-% the name(s) of the author(s) follow(s) next
-%
-% NB: Chinese authors should write their first names(s) in front of
-% their surnames. This ensures that the names appear correctly in
-% the running heads and the author index.
-\author{Removed for Blind Review}
-%\author{André Susano Pinto, Andrzej Pronobis, Luis Paulo Reis}
-%
-%\authorrunning{Lecture Notes in Computer Science: Authors' Instructions}
-% (feature abused for this document to repeat the title also on left hand pages)
-
-% the affiliations are given next; don't give your e-mail address
-% unless you accept that it will be published
-\institute{Removed for Blind Review}
-%\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
-%\url{andresusanopinto@gmail.com}\\
-%\url{http://url}}
-
-\maketitle
-
-\begin{abstract}
-This paper presents an approach to the problem of novelty
-detection in the context of semantic room categorization.
-The ability to assign semantic labels to areas in the environment is crucial for
-autonomous agents aiming to perform complex human-like tasks and human
-interaction.
-However, in order to be robust and naturally learn the semantics from
-the human user, the agent must be able to identify gaps in its own knowledge.
-To this end, we propose a method based on graphical models to identify novel
-input which does not match any of the previously learnt semantic descriptions.
-The method employs a novelty threshold defined in terms of conditional
-and unconditional probabilities. The novelty threshold is then optimized using
-an unconditional probability density model trained from unlabelled data.
-
-\keywords{novelty detection, semantic data, probabilistic graphical models,
-room classification, indoor environments, robotics, multi-modal classification.}
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-There has been several efforts in the areas of artificial intelligence and robotics
-in creating robots that are able to interact with humans and their environments.
-One of the important aspects is to endow those robots with a deeper understanding
-of human environments, not just for the purpose of navigation and obstacle avoidance, 
-but also in terms of human semantics and functionality. An important problem
-in creating reliable representations of space for robots that are to be deployed 
-in new and unknown realistic environments is to be able to automatically 
-identify gaps in robot's knowledge and act in order to fill those gaps.
-
-This article addresses the problem of \emph{novelty detection} within the context
-of semantic mapping i.e. generating maps containing \emph{semantic information} 
-about indoor environments, such as homes or offices. In that context, the ability 
-to detect that the observations result from a semantic concept unknown to the robot, 
-and cannot be explained by one of its models, is crucial for generating fully autonomous
-and reliable behavior. In order to be robust, the robot must identify 
-novel concepts and instead of making a costly error, refrain from the decision
-and initiate learning. In particular, we address the problem of novelty detection for
-room categorization i.e. detecting whether the area in the environment identified as a
-separate room can be assigned one of the semantic labels that the robot knows (e.g. as \emph{a kitchen} 
-or \emph{an office}) or belongs to a yet unknown semantic category.
-
-The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
-which already possesses a developed architecture based on probabilistic graphical models 
-oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
-One of the major problems in using the probabilistic models of the environments for novelty
-detection is selecting the optimal threshold above which the test sample is considered
-novel. This problem becomes more difficult when, as in case of Dora, the representation
-grows as the robot explores the environment. Methods are required to find
-optimal parameters given the current structure of the model. To this end, this work 
-studies methods for novelty threshold selection using probabilistic graphical models.
-
-The rest of this paper is structured as follows. First, we briefly review the related work on
-room categorization and novelty detection (Section~\ref{sec:related}). Then, we give an outline of the 
-architecture of Dora and the structure of the conceptual map representing the semantic information
-(Section~\ref{sec:overview}). Next, we discuss the methods for novelty detection (Section~\ref{sec:novelty-detection}) 
-and present results of our preliminary experiments (Section~\ref{sec:results}). This paper concludes with a 
-summary in Section \ref{sec:conclusion}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Related Work}
-\label{sec:related}
-% TODO(pronobis): general comment about many .... approches to place categorization.
-The problem of room categorization based on visual information was first addressed
-in the computer vision community. In this case, the research focused mainly on the
-problem of classifying single images captured in indoor or outdoor environments
-(scene classification)\cite{oliva2006building,torralba2003contextual}.
-At the same time, robotics researchers initially employed
-the 2D laser range sensor being much more robust to variations occurring in the
-environment and much easier to handle computationally in real time\cite{mozos2005supervised}.
-
-Multi-modal approaches, such as combining semantic data extracted from
-several sources or classifiers are expected to have better performance on scene
-recognition than single-cue approaches. Quattoni and Torralba\cite{quattoni2009recognizing}
-showed that most scene recognition models work poorly in indoor scenes when
-compared to outdoor scenes since the properties that
-characterize rooms changes depending on the category. For instance corridors are well
-described by global properties and bookstores are well described by the presence of
-specific objects (books).
-Galindo et al.\cite{galindo2005multi} also exploits this by defining a bidirectional relation
-between object and room category, where object defines a room category and a room
-category provides information on where objects may be found.
-
-Probabilistic representations have been frequently used for spatial modelling
-in robots operating in the real-world\cite{gross2009toomas,maier2010probabilistic}.
-Boutell et al.\cite{boutell2006factor} have studied outdoor scene classification using
-\emph{factor graphs} and modelling spatial relations between objects in the scene
-to extract better knowledge from semantic (high-level) features.
-
-Although our approach is presented in the context of mobile robotics it relies on
-standard concepts and techniques such as semantic data and graphical models.
-Those are often used in the area of information retrieval.
-An interesting example is the usage of an hidden concept layer between visual features
-and text information to provide automatic image annotation\cite{zhang2006probabilistic}.
-
-Novelty detection has been studied for many years and there are several approaches
-based on statistical analysis\cite{markou2003novelty}.
-Graphical models have been used to learn distributions of variables, both in
-supervised and unsupervised ways and by using thresholds on those distributions
-based solely on the conditional probability as seen on Bishop\cite{bishop1994novelty}
-a novelty system can be trivially implemented.
-
-However to the knowledge of the authors there is no reference on how to perform
-novelty detection using graphs that are dynamically generated.
-% TODO(andresp): anything more that can be said on why this paper is useful?
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Dora Architecture Overview}
-\label{sec:overview}
-Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora. Short paragraph describing dora.
-
-As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
-That model connects the sensed properties together with the variables used to model the world.
-And allows to perform queries on the probabilities of aspects of the environment.
-For example: where is most likely to find a cereal box.
-
-From its architecture only the \emph{conceptual layer} is of interest to this article.
-Its role is to aggregate the following semantic information coming from other layers:
-
-\begin{description}
- \item[Doorway detection] is used to segment the low-level space into rooms and map connectivity between them.
- \item[Room size and shape] are classified by using 2D laser scans data and are associated as properties of a given room.
-The system utilizes pre-trained set of classifiers to label size as (large, medium, small) and shape as (rectangular or elongated).
- \item[Object detection] is performed using the visual input in order to detect .
-The system keeps track of the number of object and types seen per each room.
-Objects are once again detected by running a pre-trained set of detectors for: book, cereal box, computer, robot, stapler, toilet paper.
- \item[Room appearance] is categorized from the visual input by using CRFH and a pre-trained set of 7 different models.
-\end{description}
-
-With the extracted information and conceptual knowledge the conceptual layer creates a structured probabilistic representation
-in order to model all known variables and their relations.
-As so, in order to represent the knowledge the layer builds a \emph{chain graph}:
-a probabilistic graphical model that merges both Bayesian Networks and Random Markov Fields.
-
-The use of graphical models to describe distributions of variables has useful properties.
-The edges between the variables can be seen as a kind of filter to the properties of the system.
-Properties those that are expected to be captured by the conceptual knowledge.
-At the same time they are a generative models and therefore allow to calculate the probability
-on any given subset of variables on the graph allowing the system to work even when some
-information is missing.
-
-\begin{figure}[h]
-\centering
-\subfloat[Conceptual layer]{\includegraphics[width=0.50\textwidth]{figures/conceptual-layer.jpg}}
-\subfloat[Chain graph]{\includegraphics[width=0.49\textwidth]{figures/chain-graph.png}}
-
-\caption{The conceptual layer structures the sensed environment together with the conceptual knowledge
-         in order to create a structured probabilistic representation of the world.}
-\end{figure}
-
-\subsection{Factor Graphs}
-Although the conceptual layer works with \emph{chain graphs}, those can be converted
-into \emph{factor graphs}\cite{kschischang2001factor}. Factor graphs are used throughout this paper as they provide an
-easier manipulation due to factorization.
-Moreover, there exists efficient implementation of inference engines operating on factor graph
-representations\cite{Mooij_libDAI_10}.
-Describing the distribution function in terms of graphs allows to use those engines to
-efficently calculate marginals on any given subset of variables by exploiting conditional
-independence between variables.
-
-A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
-representing random variables and factors.
-Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
-to which the factor is connected.
-Thus, a factor graph can be seen as a description of probability density function obtained
-by a product of all the factors. In order to represent the probability,
-a normalization factor needs to be introduced, resulting in the following equation:
-
-\begin{equation}
-P_G(x) = \frac{1}{Z}\prod_{\phi \in F_G}{\phi(x_{\phi})},\qquad
-Z = \sum_{X_G}\prod_{\phi \in F_G}{\phi(x_{\phi})}
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Novelty Detection}
-\label{sec:novelty-detection}
+\section{Novelty Detection Review and Related Work}
 % Introduce novelty detection
 Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about\cite{markou2003novelty}.
+other than the ones the system knows about~\cite{markou2003novelty}.
 It is harder than classification as only positive samples of a class are available
@@
 
-When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel as an associated error probability of $P(\overline{novel}|x)$.
-And the error rate of a system that classifies the set of inputs $N$ as novel is given by
-the equation:
+\section{Novelty Detection as a Threshold}
+When dealing with a statistic point of view, noisy data or unstable features,
+a decision to classify a sample $n$ as novel has an associated error probability
+$P(\overline{novel}|n)$.
+Accordingly, on a system that classifies a set of inputs $N$ as novel, the
+probability of incorrectly classifying a sample as novel is given by the
+equation:
 
 \begin{equation}
-P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+\label{eq:false-positive}
+P(\textnormal{false positive}|\textnormal{positive}) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
 \end{equation}
@@
 
-% Explain approach of using conditional probability
+Since there is only interest in maintaining the same order-relation as
+$P(\overline{novel}|x)$ the absolute value is not needed and any constant
+factor can be dropped. Leaving a ratio between a \emph{conditional} and
+\emph{unconditional probabilities} suitable for implementing novelty detection
+via  thresholding.
+
 \subsection{Conditional Probability}
-\label{sec:conditional-prob}
+The conditional probability $P(x|\overline{novel})$ describes the distribution
+of the non-novel samples. In case the labelled data available for the agent
+to learn a concept comes from this same distribution the correct approach
+is to use it as prior-information for modelling the conditional probability.
 
-The conditional probability models the distribution of variables given that the sample
-is not novel for the agent.
-This is: it should model the distribution of the sensed features considering that the
-knowledge the agent possesses holds true. Under that its natural to model it with a graphical
-model that combines the information acquired with the relations that the describe the
-relations learned by the agent between the variables.
+Note that it is important for the labelled data to come from the same
+distribution as where the system will run, as otherwise it will contain some
+bias (for example introduced by an incorrect filtering)
 
-In Dora case, this is equivalent to use the graphical model used by it to describe its
-current believes on the variables modelled by the system.
-That graph is built, by the conceptual layer, by instancing the information extracted from
-other layers together with the conceptual knowledge such as: objects are properties of rooms,
-rooms are connected between each other.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
-\caption{\label{fig:conditional-prob-graph}A sample of a factor graph modelling a distribution
-         of a set $x$ of sensed variables.}
-\end{figure}
-
-\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
-layer to represent the conditional probability on the sensed variables $x$.
-A set of hidden variables is added to represent the conceptual knowledge the system is aware of.
-In the presented graph variables $R_i$ were added to model the room categories that influence
-the directly sensed features on each physical room, as well connectivity factors between each
-room.
-
-The factors connecting the variables are trained by the system by searching databases of
-common knowledge to build potentials describing how likely it is that a specific set of
-values for a set of variables is more likely to occur over than any other set of values.
-For instance: it is very likely to find a cereal box in a kitchen; and it is unlikely to find
-bathroom connected to another bathroom.
-
-We propose using such a graph $G$ built by the conceptual layer for modelling $P_G(x)$
-as an approximation for $P(x|\overline{novel})$.
+Therefore the correct approach is to try to model it
 
 \subsection{Unconditional Probability}
-\label{sec:unconditional-prob}
-With only access to labelled data a common approach is to define a threshold assuming
-that $P(x)$ is constant through all the samples.
+The unconditional probability $P(x)$ plays an important role on obtaining a
+correct order relation for performing novelty detection.
+It serves as a normalizing component that allows the system to  whether
+a given sample conditional probability arises from it belonging to the
+known concept or from it likelihood of being sampled.
 
-Its important to notice that in several cases assuming it to be constant leads to
-discarding the factor.
-Nonetheless, here the distributions are dynamically changing as the system learns
-more on the environment.
-So the normalizing argument $P(x)$ has to be evaluated for each new subset of $x$.
+On lack of any information about the unconditional probability and conforming to
+the principle of maximum entropy (\autoref{sec:max-entropy}) a uniform
+distribution must be chosen.
+Though, as discussed in \autoref{sec:unlabelled-data}, by using unlabelled data
+it becomes possible to obtain prior-information and achieve a better
+approximation.
 
-Assuming that the unconditional distributions generates all possible outcome with
-the same probability we can model it with $\prod{1/\# x_i}$,
-where $\# x_i$ denotes the cardinality of the state space of variable $x_i$.
-In graphical model terms this is represented to a factor graph $U$ with the
-variables but without any factors as illustrated on \autoref{fig:uniform-prob-graph}.
+%\subsection{Novelty Detection on a Variable Set of Features}
+Note that very often, novelty detection is applied on a fixed set of features
+together with an assumption of a uniform unconditional probability.
+On those cases $P(x)$ becomes a constant and therefore a novelty threshold
+can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
+But in the case where the set of features $x$ is variable it cannot be
+discarded. There $P(x)$ plays a role in levering all the conditional
+probabilities on different sets of variables into the same measure units
+such that a threshold can be implemented.
 
-\begin{figure}
-\centering
-\includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
-\caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a
-         uniform distribution over any set of its variables.}
-\end{figure}
-
-Having a graphical model $G$ built to model the known data distribution and a model
-$U$ for the unconditional probability a novelty threshold would be given by:
-$P_G(x)/P_U(x)$.
-Here $P_{U}(x)$ can be seen as a normalizing factor to lever all the $P_G(x)$ on any set of
-variables $x$ into the same measure units (error rate), such that a static threshold can be implemented.
-For example the conditional probability would yield very small values on large sets of variables
-$x$ than in small sets due to the spreading over the dimensions of the input space.
-As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
-understand the observed result.
-
-\subsection{Semi Supervised: Using Unlabelled Data}
-Nonetheless, its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability
-distribution than the uniform one.
-
-For practical reasons it was assumed that the unconditional distribution
-could be modelled with all the variables independent from each other.
-Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-From the unlabelled data, only the probability of each feature $x_i$ needs to be estimated.
-\autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.70\textwidth]{figures/independent-prob-graph.pdf}
-\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to estimate
-         the unconditional probability for each variable and this way try to compensate
-         for a possible existent bias on some of the variables.}
-\end{figure}
-
-Once again the novelty threshold would be given by $P_G(x)/P_I(x)$.
-This time the addition of the unconditional factors can be understood as an
-attempt to compensate for an existing bias on the unconditional distribution.
-This is an important step to achieve a correct order-relation of the inputs sample
-for implementing a novelty threshold.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
-\label{sec:results}
-
+\section{A Practical Example}
+\label{sec:unlabelled-data}
 In order to verify the performance of the proposed threshold functions a synthetic dataset
@@
 its very small when compared to how optimal a threshold could be.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions and Future Work}
-\label{sec:conclusion}
-In this paper we presented how to define a stable novelty threshold function on
-top of \emph{probabilistic graphical models} instantiated dynamically from sensed
-semantic data.
-The presented technique is based on the ratio between a conditional and
-unconditional probability and when perfect information exists it performs an optimal
-novelty detection threshold.
-
-It was also shown that a correct estimation of unconditional probability plays an
-important role specially on small input spaces. Moreover, semi-supervised techniques
-implemented with the access to unlabelled data can be used to significantly improve
-novelty detection performance.
-
-Given the synthetic distribution, an assumption on an uniform
-distribution has led to very poor results. The same behaviour is expected to
-in real world distributions based on semantic data. For that reason,
-and due to easy access to unlabelled data, special attention will be given to using
-semi-supervised techniques for novelty detection.
-
-After this initial study on how to detect new semantic classes based on
-graphical models, future work will focus on how to use the structured
-information available from the conceptual layer to be able to detect which variable
-of the graph is novel and what makes it different from other previously learned
-classes. That will lead to generation of useful information that can be used for
-communication with the user and performing active learning of new room categories.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliographystyle{unsrt}
-\bibliography{refs}
-
-\end{document}

### 2011-05-29 thesis/novelty-intro.tex
Added comment on wrong proof of novelty as threshold.

@@
 
-Therefore any novelty system that decides to acknowledge a sample
-$a$ as novel should also detect novelty on any sample $b$ satisfying
+Theres a problem with this proof. I need to review it. The original explanation
+for me to model it as a threshold function required to have both true positive
+and false positive ratios. But for the paper I though I could do it with only
+the FPR, but its clear wrong.
+
+Therefore any novelty system that considers to acknowledge a sample
+$a$ as novel should before detect novelty on any sample $b$ satisfying
 $P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that always reduce the error probability described in
+As that would always yield a reduced the error probability described in
 \autoref{eq:false-positive} as shown:

### 2011-05-30 thesis/novelty-intro.tex
More novelty detection introduction and history text from the tech report.

@@
 
-This chapter presents a statis
+This chapter presents concepts that are used as base when adding novelty
+detection capabilities on the semantic representation in \autoref{chap:novelty}.
+It starts by providing the reader with a brief overview on novelty detection
+techniques and related works.
 
-Laying down base concepts that are then used in \autoref{chap:novelty} to
-implement novelty detection on the semantic spatial representation.
-
-It start by presenting a brief overview on novelty detection.
-
-It introduces novelty as a
-
-Shows novelty has an associated ordering.
-Shows novelty has a threshold function.
+After it shows that an optimal novelty detection system can be implemented by
+thresholding on a order-relation defined over the inputs.
+And that an equivalent order-relation can be imposed by a ratio between
+conditional and unconditional probabilities.
+Some discussion on how to interpret the meaning of those factors is also
+presented.
 
@@
 
-\begin{itemize}
-\item Introduction
-\item History
-\item Approaches to Novelty Detection
-\end{itemize}
-
 \section{Novelty Detection Review and Related Work}
 % Introduce novelty detection
+%
+%
+%
+
+Novelty detection, also called outlier or anomaly detection, is the
+classification problem related with identification of new or unknown data that
+the system was not aware of during the training.
+
+It has several applications such as fault detection, intrusion detection,
+detection of masses in mammograms, hand written digit recognition and many
+others~\citep{markou2003novelty}.
+And has seen an increase of its application in the area of robotics and control.
+
+It is posed as a one class classification where there is only access to the
+positive samples. The objective then is to classify new inputs as being
+generated by the class it was trained or if they belong to some new unknown one.
+
+Either by impossibility of generating samples of "unknown", or by extreme
+difficulty to acquire negative-examples, two-class classifiers cannot be
+applied to this problem.
+Several other methods were developed to address this problem and they often can
+be categorized in the following approaches:
+based on probability density estimation;
+based on using the feature space to separate the given class from others, such as one-class \gls{SVM};
+and based on reconstruction-error from a lower-dimension.
+
+\subsubsection*{Density estimation}
+One of the simplest approach for novelty detection is pose the problem as a
+probability density estimation problem and use the expected probability of a
+given input to trigger the case as novel.
+
+Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
+In order to reduce the input size, those approaches often employ
+dimension-reduction techniques such as \gls{PCA}.
+
+This approach is used in \cite{bishop1994novelty} where \emph{novelty detection}
+is added to a \emph{neural network} system. On it the density estimation on the
+input is calculated using a \emph{Parzen window} over training-data and that
+value is used as a novelty measure. This way they detect novel data that their
+network has not seen before and so the expected output of their system is not
+reliable.
+
+\subsubsection*{One-class {SVM}}
+One-class \gls{SVM} approaches try to distinguish the class by separating the
+training set from all other points in the input space. They try to archive that
+by enclosing the training set by some structure (example: hyper-sphere)~\citep{scholkopf2000support}.
+This leads to a decision boundary and follows the same principle as Vapnik
+solving classification with \gls{SVM}: do not solve something harder.
+
+\subsubsection*{Reconstruction Error}
+Error reconstruction methods for novelty detection use the assumption that the
+class to be defined lies on a manifold embedded on the input space of higher
+dimensions.
+By using dimensional reduction techniques, they try to defined that manifold
+and calculate how far a new input is from it.
+
+\label{sec:kernel-pca}
+One of the most common methods used for that is
+\gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick
+(\autoref{sec:kernel-trick}) to extend \gls{PCA} and perform a nonlinear
+dimensionality reduction on the input. This technique has been successfully
+used in novelty detection by \cite{Hoffmann2007863}.
+
 Novelty detection deals with detecting that a new input was generated by a class
@@
 correct order relation for performing novelty detection.
-It serves as a normalizing component that allows the system to  whether
-a given sample conditional probability arises from it belonging to the
-known concept or from it likelihood of being sampled.
+It serves as a normalizing component that allows the system to figure out
+whether a given sample conditional probability arises from it belonging to
+the known concept or from it likelihood of being sampled.
 
@@
 %\subsection{Novelty Detection on a Variable Set of Features}
-Note that very often, novelty detection is applied on a fixed set of features
+Note also that often novelty detection is applied on a fixed set of features
 together with an assumption of a uniform unconditional probability.

### 2011-05-30 thesis/contents.tex
Advances on novelty detection on semantic representations.

@@
 Given a probabilistic structure representing the sensed conceptual knowledge
-obtained by the agent develop methods to be able to detect novelty present
+obtained by the agent develop methods able to detect novelty present
 either on new semantic concepts (new classes) or even on structure that was
@@
 \chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
-On the previous chapter, novelty as been introduced as a statistical
-measure on the error-likelihood ratio between a conditional and unconditional
-probability of a given .
+The semantic mapping representation presented in \autoref{chap:semantic-mapping}
+permits the agent to understand and reason over the human semantic concepts of
+space. By using semantic data not only it eases the process of communicating 
+but also reduces complexity and allows easy implementation of high-level
+decisions.
+Due to the nature of semantic data, variables have to be considered
+uncertain. For that reason the whole semantic representation is instantiated
+as a probabilistic structure: a \emph{conceptual map}.
+
+Although the conceptual map permits uncertain reasoning it does not incorporates
+any detection mechanism for identifying or copping with knowledge gaps.
+This means the system needs to consider the knowledge absolute and is not
+able to detect by itself that some of it does not correctly describes reality.
+
+The main goal of this thesis is to address this issue by developing methods
+able to detect both novel classes and novel structures on the semantic
+representation. Allowing a more efficient behaviour when performing on unknown
+and new environments. It is an important milestone on developing of high-level
+knowledge and active learning.
 
 \section{Detecting New Classes}
+Initially the work will focus on detecting a novel class on single variable of
+the conceptual map.
+As seen before on \autoref{chap:novelty-intro}, novelty detection can be
+done by implementing a threshold on $P(x|\overline{novel})/P(x)$.
+In order to correctly represent novelty detection on single variable of the
+graph
+
+The conceptual layer build a \emph{probabilistic graphical model} $G$ that
+represents the distribution of the sensed features assuming that both the known
+structure and the known classes hold in the sample.
+For that reason the correct approach is to use $P_G(x)$ as an approximation for
+the $P(x|\overline{novel})$.
+
+\begin{figure}[h]
+\centering
+\subfloat[Factor graph modelling the probability distribution in the case both
+          the variables types and graph structure hold true.]{
+          Hello get a graph here.}
+\qquad\qquad
+\subfloat[Factor graph modelling the probability distribution in the case
+          where only variable $a$ does not conforms to any class the system
+          has learned before.]{
+          Hello get a graph here.}
+\end{figure}
+
+able to detect novel classes those capabilities 
+and with the background introduced in the previous chapter,
+the developed methods are presented on this chapter.
+
+classes types on room categories
+or any other variable.
+This means the uncertain reasoning is not allowed to refrain from classifying
+a room in one of the known categories.
+
+On the previous chapter, novelty detection as been introduced as threshold
+function on top of a ratio between a conditional and unconditional
+probability of a given set of sensed features $x$.
+
 Explain exactly and formally how to obtain a novelty measure on a single

### 2011-05-30 thesis/contents.tex
Advances on explaning how to detect a single class novelty

@@
 \section{Detecting New Classes}
-Initially the work will focus on detecting a novel class on single variable of
-the conceptual map.
-As seen before on \autoref{chap:novelty-intro}, novelty detection can be
-done by implementing a threshold on $P(x|\overline{novel})/P(x)$.
-In order to correctly represent novelty detection on single variable of the
-graph
+This section focus on detecting a novel class on single variable of the
+conceptual map. As seen before on \autoref{chap:novelty-intro},
+novelty detection can be done by implementing a threshold on
+$P(x|\overline{novel})/P(x)$.
+In order to correctly measure novelty on a single variable both
+conditional and unconditional probabilities need to be coherently defined:
 
@@
 For that reason the correct approach is to use $P_G(x)$ as an approximation for
-the $P(x|\overline{novel})$.
+$P(x|\overline{novel})$.
+
+Assume now that everything the conceptual layer knows holds true except the
+class of a single variable $a$, which the system possess no information on. 
+In that case the correct approach is to replace the factors associated with
+that variable with factors representing the prior-information the system has on
+the unconditional distribution of those factors.
+In case no information is available this means replace by a uniform factor\footnote{In case the
+unconditional factors are uniform, this is the same as omitting the factors
+from the graph structure.} accordingly to the principle of maximum entropy.
+Note that in case some information was known on the prior distribution, the
+correct approach is to replace the factors by the distributions that have that
+information in account (as is the case in \autoref{sec:unlabelled-data}).
+Let this new factor graph with unconditional factors $\phi_u$ on the variable
+$a$ be $U$. $P(x)$ is then approximated by $P_U(x)$.
 
@@
           Hello get a graph here.}
+
+\caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
+         a class of a single variable $a$, assuming all the other variables
+         classes and graph structure are correctly modelled by the agent
+         knowledge.}
 \end{figure}
 
-able to detect novel classes those capabilities 
-and with the background introduced in the previous chapter,
-the developed methods are presented on this chapter.
-
-classes types on room categories
-or any other variable.
-This means the uncertain reasoning is not allowed to refrain from classifying
-a room in one of the known categories.
-
-On the previous chapter, novelty detection as been introduced as threshold
-function on top of a ratio between a conditional and unconditional
-probability of a given set of sensed features $x$.
-
-Explain exactly and formally how to obtain a novelty measure on a single
-variable of a graph.\footnote{I am sure on the math of this.}
-
 \section{Copping with Multiple Novel Variables}
-On the previous section we show how to detect a novel class but that assumes
-the rest of the graph holds true.
+On the previous section it was shown how to detect a novel class on a variable
+assuming that the rest of the graph variables and structure hold true with the
+knowledge acquired by the agent.
 

### 2011-05-31 thesis/contents.tex
Changes regarding detection with multiple variables.

@@
 On the previous section it was shown how to detect a novel class on a variable
-assuming that the rest of the graph variables and structure hold true with the
-knowledge acquired by the agent.
+assuming that the rest of the graph variables and structure is correctly modeled
+by the agent knowledge.
+But the test to detect a given variable $a$ is novel requires both a structure
+$G$ representing the case where no knowledge gaps exist and a structure $U$
+representing the case where $a$ was identified as a novel class.
+
+It may be tempting to use a deterministic approach on the system such as:
+when the agent determines a variable to have a novel class, it marks it as novel
+and proceeds.
+
+Such an approach would no longer be optimal and not be dynamic
+executes on a new
+structure $U$ 
+of the semantic structure it does not allows the
+semantic structure to be used
 

### 2011-05-31 thesis/contents.tex
A lot of things on how to cope with multiple novel variables.

@@
 
-\section{Detecting New Classes}
+\section{Detecting Novelty on a Single Variable}
 This section focus on detecting a novel class on single variable of the
@@
 
-The conceptual layer build a \emph{probabilistic graphical model} $G$ that
-represents the distribution of the sensed features assuming that both the known
-structure and the known classes hold in the sample.
-For that reason the correct approach is to use $P_G(x)$ as an approximation for
+The conceptual layer builds a \emph{probabilistic graphical model} $G$ that
+represents the distribution of the sensed features $x$ assuming that both the
+known structure and the known classes of all variables hold in the sample.
+The correct approach is then to use $P_G(x)$ as an approximation for
 $P(x|\overline{novel})$.
@@
 class of a single variable $a$, which the system possess no information on. 
-In that case the correct approach is to replace the factors associated with
-that variable with factors representing the prior-information the system has on
-the unconditional distribution of those factors.
-In case no information is available this means replace by a uniform factor\footnote{In case the
+In that case, the correct approach is to replace the factors associated with
+$a$ with factors representing the prior-information the system has on
+the unconditional distribution of those.
+Let this new graph with unconditional factors on the variable
+$a$ be $U_a$. $P(x)$ is then approximated by $P_{U_a}(x)$.
+
+Note that in case no prior-information is available on the unconditional
+factors this means replacing them by a uniform factor\footnote{In case the
 unconditional factors are uniform, this is the same as omitting the factors
-from the graph structure.} accordingly to the principle of maximum entropy.
-Note that in case some information was known on the prior distribution, the
-correct approach is to replace the factors by the distributions that have that
-information in account (as is the case in \autoref{sec:unlabelled-data}).
-Let this new factor graph with unconditional factors $\phi_u$ on the variable
-$a$ be $U$. $P(x)$ is then approximated by $P_U(x)$.
+from the graph structure.}. In case some information is known
+(i.e. unlabelled data) the correct approach is to incorporate it into the
+unconditional factors using the principle of maximum entropy as seen on
+\autoref{sec:unlabelled-data}.
 
@@
 
+Having now defined $P_G(x)$ and $P_{U_a}$ as models models for
+$P(x|\overline{novel})$ and $P(x)$ (\autoref{fig:nd-single-variable}),
+$P_G(x)/P_{U_a}(x)$ is used to define an order-relation on the novelty of $a$.
+Now all that is missing is a threshold $t_A$ to trigger on.
+Such a threshold can be obtained by calculating which value of $t_A$
+triggers a desired amount of the labelled data.
+Also, since there is no reason to distinguish between different instances of
+a variable type, only one threshold by variable type needs to be trained.
+
+Concluding: novelty on a variable $a$ of type $A$ should be detected when:
+
+\begin{equation}
+\frac{P_G(x)}{P_{U a}(x)} < t_A
+\end{equation}
+
+where $t_A$ is a threshold for variables of type $A$, $P_G(x)$ denotes the
+probability of the sensed variables $x$ being generated by a distribution
+where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
+by a distribution where only $a$ is novel.
+
 \section{Copping with Multiple Novel Variables}
@@
 by the agent knowledge.
+In this section methods are presented that not only allow to extended this to
+detection on multiple variables but also allow the system to incorporate
+novelty information back in the graphical model in order to continue its
+probabilistic and uncertain reasoning.
+
+When trying to extend the method to deal with multiple novel variables it may
+be tempting to follow a deterministic approach such as:
+when the agent determines a variable $a$ to have a novel class, it marks it as
+novel and proceeds from that moment by using a graphical model where the factors
+associated with the $a$ replaced by unconditional factors.
+But such an approach would not be able to optimally detect novel variables
+as later decisions are dependent on former ones.
+It would also not be able to adapt itself and reconsider previously decisions
+when more data becomes available for the agent.
+
+To handle those issues a fully probabilistic approach should be taken
+instead of a deterministic.
+That can be done by extending all the variables on the factor graph representing
+the semantic map with an additional class \emph{novel} and extending all the
+factors with information on the unconditional distribution.
+
+Such a factor graph would be able to handle novel variables and incorporate
+prior-information on the unconditional distributions since it is able to
+represent any state of the $X$ variables present on the graph where any
+subset of them is novel. But in order to add the ability to control the
+novelty thresholds a normalization factor $\phi_N$ needs to be introduced.
+
+In order to give full control on the threshold on which any subset of
+variables is considered novel the factor $\phi_N$ needs to be dependent on all
+the $X$ variables and act as a weighting function that correctly scales the
+likelihood of a given instance of the $X$ variables in order to make a
+\emph{MAP} operation on the graph to yield the desired value.
+Such a factor would require tunning $2^N$ parameters, describing how to
+distribute the weights over all the possible $2^N$ subsets of the $X$ variables.
+
+Aiming at feasibility a simpler approach should be used.
+Requiring
+
+This is the factor $\phi_N(X) = \prod_{i \in X} t_i$.
+As so the factor $\phi_N$ can be decomposed into several $\phi_{N_i} = t_i$, one
+for each variable, that measure on the threshold of detecting novelty on
+the single variable $i$.
+
+More formally, there is an interest in the new graph
+Nonetheless after that a mechanism to allow to control the thresholds on
+detecting novelty also needs to simple doing that the ability to control the thresholds on
+variables is lost
+In order to measure novelty of a set of variable the full set of approach
+Aiming at obtaining a correct behaviour for detecting novelty on a certain
+subset of variables
+Optimally the system should re-evaluate novelty for all subsets of variables
+each time new information is added.
+A method to define a threshold on any subset of the $N$ variables also needs to
+be introduced.
+
+After that the system would be still be able to reason over variables on the
+graph and deduce extra new novel variables if it was the case.
+But such an approach would be dependent on the order of variables that the
+system considers for novelty detection. And it would also suffer from inability
+to reconsider the novelty on any given variable as more data becomes available.
+
+The used approach can be extended to multiple variables. Though in order to
+expand it to all the $n$ variables in the graph, $n!$ unconditional
+probabilities would need to be evaluated.
+
+Aiming at a fully probabilistic integration of novelty detection on any
+number of variables the following method was introduced:
+
 But the test to detect a given variable $a$ is novel requires both a structure
@@
 representing the case where $a$ was identified as a novel class.
-
-It may be tempting to use a deterministic approach on the system such as:
-when the agent determines a variable to have a novel class, it marks it as novel
-and proceeds.
 
@@
 semantic structure to be used
-
-In this section we show how to handle novelty in several variables
-in simultaneous.\footnote{I am almost sure the math here will work out the
-way I think}
 

### 2011-05-31 thesis/contents.tex
Lots of text explanining how to integrate multi-novelty on the graph and how to control thresholds on it.

@@
 \section{Features}
+
 \section{Conceptual Knowledge}
 \section{Conceptual Map}
+\label{sec:clues-from-low-level}
 
@@
 \begin{equation}
-\frac{P_G(x)}{P_{U a}(x)} < t_A
+\frac{P_G(x)}{P_{U_a}(x)} < t_A
 \end{equation}
@@
 
+\begin{figure}[h]
+\centering
+Yeah get a figure for here.
+\caption{\label{fig:multiple-big-threshold}Factor graph extended to be able to
+         represent novelty on any subset of variables. Note the added factor
+         $\phi_N$ that servers as a normalization factor such that probability
+         with any subset of $Y$ variables considered novel can be scaled with
+         any value such that any thresholding decision mechanism can be
+         implemented.}
+\end{figure}
+
 In order to give full control on the threshold on which any subset of
@@
 the $X$ variables and act as a weighting function that correctly scales the
-likelihood of a given instance of the $X$ variables in order to make a
-\emph{MAP} operation on the graph to yield the desired value.
-Such a factor would require tunning $2^N$ parameters, describing how to
-distribute the weights over all the possible $2^N$ subsets of the $X$ variables.
+likelihood of a given instance of the $X$ variables.
+If the criterium for defining the threshold is the likelihood of all variables
+$X$ being non-novel has the max likelihood if and only if
+\autoref{eq:multiple-threshold} holds for all the non-empty subsets $Y$ of $X$.
+The weight for $\phi_N$ when the set $Y$ is classified as novel
+can be set for $t_Y$ and $\phi_N$ for the empty subset of novel variables is set
+to $1$.
 
-Aiming at feasibility a simpler approach should be used.
-Requiring
+\begin{equation}
+\label{eq:multiple-threshold}
+\frac{P_G(x)}{P_{U_Y}(x)} > t_Y
+\end{equation}
 
-This is the factor $\phi_N(X) = \prod_{i \in X} t_i$.
-As so the factor $\phi_N$ can be decomposed into several $\phi_{N_i} = t_i$, one
-for each variable, that measure on the threshold of detecting novelty on
-the single variable $i$.
+Still without further assumptions a threshold $t_Y$ would need to be defined for
+all the $2^N-1$ non-empty subsets.
+By taking a simple approach of considering the novelty on a variable independent
+from the novelty on any other variable $t_Y$ can be factorized:
 
-More formally, there is an interest in the new graph
-Nonetheless after that a mechanism to allow to control the thresholds on
-detecting novelty also needs to simple doing that the ability to control the thresholds on
-variables is lost
-In order to measure novelty of a set of variable the full set of approach
-Aiming at obtaining a correct behaviour for detecting novelty on a certain
-subset of variables
-Optimally the system should re-evaluate novelty for all subsets of variables
-each time new information is added.
-A method to define a threshold on any subset of the $N$ variables also needs to
-be introduced.
+\begin{equation}
+t_Y = \prod_{i \in Y}{t_i}
+\end{equation}
 
-After that the system would be still be able to reason over variables on the
-graph and deduce extra new novel variables if it was the case.
-But such an approach would be dependent on the order of variables that the
-system considers for novelty detection. And it would also suffer from inability
-to reconsider the novelty on any given variable as more data becomes available.
+This factorization yields then the interesting property to allow to split the
+threshold factor $\phi_N$ into $N$ different $\phi_{N_i}$ associated with each
+variable of the graph. A graph able to handle novelty detection on several
+variables with the described factorization is illustrated on
+\autoref{fig:multi-factorized-threshold}.
 
-The used approach can be extended to multiple variables. Though in order to
-expand it to all the $n$ variables in the graph, $n!$ unconditional
-probabilities would need to be evaluated.
+\begin{figure}[h]
+\centering
+Yeah get a figure for here.
+\caption{\label{fig:multi-factorized-threshold}Factor graph extended to be able
+         to represent novelty on any subset of variables. Note the added factor
+         $\phi_N$ has been split into a $\phi_{N_i}$ still it is able to perform
+         detection on multiple variables according to the factorization
+         described in the text.}
+\end{figure}
 
-Aiming at a fully probabilistic integration of novelty detection on any
-number of variables the following method was introduced:
-
-But the test to detect a given variable $a$ is novel requires both a structure
-$G$ representing the case where no knowledge gaps exist and a structure $U$
-representing the case where $a$ was identified as a novel class.
-
-Such an approach would no longer be optimal and not be dynamic
-executes on a new
-structure $U$ 
-of the semantic structure it does not allows the
-semantic structure to be used
+To sum up, the method described above allows to integrate novelty detection
+into the graph, providing the agent with tools to describe its belief on the
+novelty of variables on the system and to continue working with such a
+representation. The mechanism is also fully probabilistic and at no time
+the agent needs to take an irrevocable decision on the novelty of a variable,
+allowing the agent beliefs to change over time as more data becomes available.
 
 \subsection{Handling novelty signals from the low-level classifiers}
-Explain how to handle novelty signals from low-level classifiers, now that we
-have a model with the ability to still reason when several variables can be
-novel.
+On \autoref{sec:clues-from-low-level} it was shown that information from the
+low-level classifiers is integrated in the conceptual map in a probabilistic
+way by introducing a factor $\phi_o$ on each observed variable that describes
+the probabilistic distribution of that variable given by the classifier.
+
+Since in order to detect novelty all the variables have been extended with a
+new state indicating novelty. That same state can be used to integrate novelty
+signals coming from the low-level classifiers.
+
+TODO: maybe describe how those novelty signals could be generated from the
+low-level classifiers\dots decision boundary distance, kernel pca, etc\dots
 

### 2011-05-31 thesis/contents.tex
Moved Chapter Novelty on Semantic Representations to its own tex file.

--- 2011-05-31 thesis/contents.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
-The semantic mapping representation presented in \autoref{chap:semantic-mapping}
-permits the agent to understand and reason over the human semantic concepts of
-space. By using semantic data not only it eases the process of communicating 
-but also reduces complexity and allows easy implementation of high-level
-decisions.
-Due to the nature of semantic data, variables have to be considered
-uncertain. For that reason the whole semantic representation is instantiated
-as a probabilistic structure: a \emph{conceptual map}.
-
-Although the conceptual map permits uncertain reasoning it does not incorporates
-any detection mechanism for identifying or copping with knowledge gaps.
-This means the system needs to consider the knowledge absolute and is not
-able to detect by itself that some of it does not correctly describes reality.
-
-The main goal of this thesis is to address this issue by developing methods
-able to detect both novel classes and novel structures on the semantic
-representation. Allowing a more efficient behaviour when performing on unknown
-and new environments. It is an important milestone on developing of high-level
-knowledge and active learning.
-
-\section{Detecting New Classes}
-This section focus on detecting a novel class on single variable of the
-conceptual map. As seen before on \autoref{chap:novelty-intro},
-novelty detection can be done by implementing a threshold on
-$P(x|\overline{novel})/P(x)$.
-In order to correctly measure novelty on a single variable both
-conditional and unconditional probabilities need to be coherently defined:
-
-The conceptual layer build a \emph{probabilistic graphical model} $G$ that
-represents the distribution of the sensed features assuming that both the known
-structure and the known classes hold in the sample.
-For that reason the correct approach is to use $P_G(x)$ as an approximation for
-$P(x|\overline{novel})$.
-
-Assume now that everything the conceptual layer knows holds true except the
-class of a single variable $a$, which the system possess no information on. 
-In that case the correct approach is to replace the factors associated with
-that variable with factors representing the prior-information the system has on
-the unconditional distribution of those factors.
-In case no information is available this means replace by a uniform factor\footnote{In case the
-unconditional factors are uniform, this is the same as omitting the factors
-from the graph structure.} accordingly to the principle of maximum entropy.
-Note that in case some information was known on the prior distribution, the
-correct approach is to replace the factors by the distributions that have that
-information in account (as is the case in \autoref{sec:unlabelled-data}).
-Let this new factor graph with unconditional factors $\phi_u$ on the variable
-$a$ be $U$. $P(x)$ is then approximated by $P_U(x)$.
-
-\begin{figure}[h]
-\centering
-\subfloat[Factor graph modelling the probability distribution in the case both
-          the variables types and graph structure hold true.]{
-          Hello get a graph here.}
-\qquad\qquad
-\subfloat[Factor graph modelling the probability distribution in the case
-          where only variable $a$ does not conforms to any class the system
-          has learned before.]{
-          Hello get a graph here.}
-
-\caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
-         a class of a single variable $a$, assuming all the other variables
-         classes and graph structure are correctly modelled by the agent
-         knowledge.}
-\end{figure}
-
-\section{Copping with Multiple Novel Variables}
-On the previous section it was shown how to detect a novel class on a variable
-assuming that the rest of the graph variables and structure is correctly modeled
-by the agent knowledge.
-But the test to detect a given variable $a$ is novel requires both a structure
-$G$ representing the case where no knowledge gaps exist and a structure $U$
-representing the case where $a$ was identified as a novel class.
-
-It may be tempting to use a deterministic approach on the system such as:
-when the agent determines a variable to have a novel class, it marks it as novel
-and proceeds.
-
-Such an approach would no longer be optimal and not be dynamic
-executes on a new
-structure $U$ 
-of the semantic structure it does not allows the
-semantic structure to be used
-
-In this section we show how to handle novelty in several variables
-in simultaneous.\footnote{I am almost sure the math here will work out the
-way I think}
-
-\subsection{Handling novelty signals from the low-level classifiers}
-Explain how to handle novelty signals from low-level classifiers, now that we
-have a model with the ability to still reason when several variables can be
-novel.
-
-\section{Detecting New Structure}
+\input{novelty}
 

### 2011-05-31 thesis/novelty.tex
Moved Chapter Novelty on Semantic Representations to its own tex file.

\chapter{Novelty Detection on Semantic Representations}\label{chap:novelty}
The semantic mapping representation presented in \autoref{chap:semantic-mapping}
permits the agent to understand and reason over the human semantic concepts of
space. By using semantic data not only it eases the process of communicating 
but also reduces complexity and allows easy implementation of high-level
decisions.
Due to the nature of semantic data, variables have to be considered
uncertain. For that reason the whole semantic representation is instantiated
as a probabilistic structure: a \emph{conceptual map}.

Although the conceptual map permits uncertain reasoning it does not incorporates
any detection mechanism for identifying or copping with knowledge gaps.
This means the system needs to consider the knowledge absolute and is not
able to detect by itself that some of it does not correctly describes reality.

The main goal of this thesis is to address this issue by developing methods
able to detect both novel classes and novel structures on the semantic
representation. Allowing a more efficient behaviour when performing on unknown
and new environments. It is an important milestone on developing of high-level
knowledge and active learning.

\section{Detecting Novelty on a Single Variable}
This section focus on detecting a novel class on single variable of the
conceptual map. As seen before on \autoref{chap:novelty-intro},
novelty detection can be done by implementing a threshold on
$P(x|\overline{novel})/P(x)$.
In order to correctly measure novelty on a single variable both
conditional and unconditional probabilities need to be coherently defined:

The conceptual layer builds a \emph{probabilistic graphical model} $G$ that
represents the distribution of the sensed features $x$ assuming that both the
known structure and the known classes of all variables hold in the sample.
The correct approach is then to use $P_G(x)$ as an approximation for
$P(x|\overline{novel})$.

Assume now that everything the conceptual layer knows holds true except the
class of a single variable $a$, which the system possess no information on. 
In that case, the correct approach is to replace the factors associated with
$a$ with factors representing the prior-information the system has on
the unconditional distribution of those.
Let this new graph with unconditional factors on the variable
$a$ be $U_a$. $P(x)$ is then approximated by $P_{U_a}(x)$.

Note that in case no prior-information is available on the unconditional
factors this means replacing them by a uniform factor\footnote{In case the
unconditional factors are uniform, this is the same as omitting the factors
from the graph structure.}. In case some information is known
(i.e. unlabelled data) the correct approach is to incorporate it into the
unconditional factors using the principle of maximum entropy as seen on
\autoref{sec:unlabelled-data}.

\begin{figure}[h]
\centering
\subfloat[Factor graph modelling the probability distribution in the case both
          the variables types and graph structure hold true.]{
          Hello get a graph here.}
\qquad\qquad
\subfloat[Factor graph modelling the probability distribution in the case
          where only variable $a$ does not conforms to any class the system
          has learned before.]{
          Hello get a graph here.}

\caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
         a class of a single variable $a$, assuming all the other variables
         classes and graph structure are correctly modelled by the agent
         knowledge.}
\end{figure}

Having now defined $P_G(x)$ and $P_{U_a}$ as models models for
$P(x|\overline{novel})$ and $P(x)$ (\autoref{fig:nd-single-variable}),
$P_G(x)/P_{U_a}(x)$ is used to define an order-relation on the novelty of $a$.
Now all that is missing is a threshold $t_A$ to trigger on.
Such a threshold can be obtained by calculating which value of $t_A$
triggers a desired amount of the labelled data.
Also, since there is no reason to distinguish between different instances of
a variable type, only one threshold by variable type needs to be trained.

Concluding: novelty on a variable $a$ of type $A$ should be detected when:

\begin{equation}
\frac{P_G(x)}{P_{U_a}(x)} < t_A
\end{equation}

where $t_A$ is a threshold for variables of type $A$, $P_G(x)$ denotes the
probability of the sensed variables $x$ being generated by a distribution
where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
by a distribution where only $a$ is novel.

\section{Copping with Multiple Novel Variables}
On the previous section it was shown how to detect a novel class on a variable
assuming that the rest of the graph variables and structure is correctly modeled
by the agent knowledge.
In this section methods are presented that not only allow to extended this to
detection on multiple variables but also allow the system to incorporate
novelty information back in the graphical model in order to continue its
probabilistic and uncertain reasoning.

When trying to extend the method to deal with multiple novel variables it may
be tempting to follow a deterministic approach such as:
when the agent determines a variable $a$ to have a novel class, it marks it as
novel and proceeds from that moment by using a graphical model where the factors
associated with the $a$ replaced by unconditional factors.
But such an approach would not be able to optimally detect novel variables
as later decisions are dependent on former ones.
It would also not be able to adapt itself and reconsider previously decisions
when more data becomes available for the agent.

To handle those issues a fully probabilistic approach should be taken
instead of a deterministic.
That can be done by extending all the variables on the factor graph representing
the semantic map with an additional class \emph{novel} and extending all the
factors with information on the unconditional distribution.

Such a factor graph would be able to handle novel variables and incorporate
prior-information on the unconditional distributions since it is able to
represent any state of the $X$ variables present on the graph where any
subset of them is novel. But in order to add the ability to control the
novelty thresholds a normalization factor $\phi_N$ needs to be introduced.

\begin{figure}[h]
\centering
Yeah get a figure for here.
\caption{\label{fig:multiple-big-threshold}Factor graph extended to be able to
         represent novelty on any subset of variables. Note the added factor
         $\phi_N$ that servers as a normalization factor such that probability
         with any subset of $Y$ variables considered novel can be scaled with
         any value such that any thresholding decision mechanism can be
         implemented.}
\end{figure}

In order to give full control on the threshold on which any subset of
variables is considered novel the factor $\phi_N$ needs to be dependent on all
the $X$ variables and act as a weighting function that correctly scales the
likelihood of a given instance of the $X$ variables.
If the criterium for defining the threshold is the likelihood of all variables
$X$ being non-novel has the max likelihood if and only if
\autoref{eq:multiple-threshold} holds for all the non-empty subsets $Y$ of $X$.
The weight for $\phi_N$ when the set $Y$ is classified as novel
can be set for $t_Y$ and $\phi_N$ for the empty subset of novel variables is set
to $1$.

\begin{equation}
\label{eq:multiple-threshold}
\frac{P_G(x)}{P_{U_Y}(x)} > t_Y
\end{equation}

Still without further assumptions a threshold $t_Y$ would need to be defined for
all the $2^N-1$ non-empty subsets.
By taking a simple approach of considering the novelty on a variable independent
from the novelty on any other variable $t_Y$ can be factorized:

\begin{equation}
t_Y = \prod_{i \in Y}{t_i}
\end{equation}

This factorization yields then the interesting property to allow to split the
threshold factor $\phi_N$ into $N$ different $\phi_{N_i}$ associated with each
variable of the graph. A graph able to handle novelty detection on several
variables with the described factorization is illustrated on
\autoref{fig:multi-factorized-threshold}.

\begin{figure}[h]
\centering
Yeah get a figure for here.
\caption{\label{fig:multi-factorized-threshold}Factor graph extended to be able
         to represent novelty on any subset of variables. Note the added factor
         $\phi_N$ has been split into a $\phi_{N_i}$ still it is able to perform
         detection on multiple variables according to the factorization
         described in the text.}
\end{figure}

To sum up, the method described above allows to integrate novelty detection
into the graph, providing the agent with tools to describe its belief on the
novelty of variables on the system and to continue working with such a
representation. The mechanism is also fully probabilistic and at no time
the agent needs to take an irrevocable decision on the novelty of a variable,
allowing the agent beliefs to change over time as more data becomes available.

\subsection{Handling novelty signals from the low-level classifiers}
On \autoref{sec:clues-from-low-level} it was shown that information from the
low-level classifiers is integrated in the conceptual map in a probabilistic
way by introducing a factor $\phi_o$ on each observed variable that describes
the probabilistic distribution of that variable given by the classifier.

Since in order to detect novelty all the variables have been extended with a
new state indicating novelty. That same state can be used to integrate novelty
signals coming from the low-level classifiers.

TODO: maybe describe how those novelty signals could be generated from the
low-level classifiers\dots decision boundary distance, kernel pca, etc\dots

\section{Detecting New Structure}

### 2011-06-01 thesis/contents.tex
Spplited Introduction and Semantic Mpaping chapter out of contents.tex

@@
 
-\chapter{Introduction}
-% Introduce:
-%  Mobile Robotics, A.I.
-%  Interaction with Humans
-%  Need for concepts/semantic information
-%  Dynamic human environment
-%  Non-feasibility of hard-coded concepts
-%  Requirements to detect novelty and handle it.
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us. Understand our concepts and
-be able to reason.
-
-They need to possess ability to adapt to situations as its infeasible to rely
-on extensive man-work to tag objects, map space and code all the aspects that
-make up our human-reality.
-
-% == Introduce the need of semantic representations ==
-%
-% There is a lot of low-level methods (or specific knowledge to robots)
-% But the creation of semantic representations allow to bridge that low-level
-% sensing with high-level concepts that facilitate several high-level Tasks.
-
-% Importance of mapping and localization in a robot as well as reasoning on
-% properties of each space.
-% Humans give labels to space that characterized the properties and activities
-% expected to be performed on them.
-% Such information is of great interest to be structured and organized in such
-% a way reasoning and planning can be implemented.
-%
-% It also holds an important role in long-term planning and stability as it
-% hides the space-time local details and allows to focus on high-level thinking.
-
-% Deal with uncertainty -> need for probabilistic models.
-% Deal with semantic    -> need of structured models such as graphical models.
-
-Example of what kind of an high-level thinking on space and its semantics can
-provide: Get me a beer/milk concept.
-
-% == Reasoning about need for Novelty Detection ==
-%
-% Dynamic environment (inability to know the environment or even to be possible
-% to map all semantic concepts to interact with humans).
-%
-% Knowledge-Awareness (knowing what we know... And detecting novel cases)
-% Identify gaps in the semantic knowledge.
-% Automatic detection and learning of novel concepts.
-%
-% Increase robustness
-% Self-Extending
-%
-
-\section{Problem and Goals}
-% Extend a probabilistic graphical modelling framework with capabilities to
-% detect novelty.
-% Both in terms of novel classes as of novel structure.
-%
-Given a probabilistic structure representing the sensed conceptual knowledge
-obtained by the agent develop methods able to detect novelty present
-either on new semantic concepts (new classes) or even on structure that was
-previously unknown.
-
-% Yeah thats what you call dream :P
-In an extreme ideal case the agent should be able to go from zero-knowledge to
-understanding presence of certain objects (as generators of a set of sensed
-sensed properties grouped locally), understand areas within rooms (sink-area on
-on a kitchen), rooms as separated by doors, and understand environments such
-as office, home, warehouse, spaceship.
-
-For that the probabilistic semantic representation presented by
-\cite{andrzej2011phd} is used.
-
-% If everything goes fine... 2 of the "Future Directions" proposed are 
-% Novelty Detection and Learning of Novel Concepts:
-%  Identify Gaps in Spatial and Semantic Knowledge -> Addressed
-%  Performing learning of new concepts -> Outside Scope / No Time
-%
-% Using Properties for Space Segmentation
-%  If we move on detecting novel structures on the graph we are addressing this
-%  issue.
-
-\section{Thesis Outline}
-The rest of this thesis is structured as follows:
-
-\autoref{chap:background} introduces background concepts for understanding the
-presented thesis. In particular \autoref{sec:graphical-models} introduces
-\emph{probabilistic graphical models} that lay the base modelling tool for
-the used structured semantic representation.
-
-\autoref{chap:semantic-mapping} describes the system proposed by
-\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
-that is responsible for the creation of the \emph{probabilistic semantic
-representation} that this thesis aims at improving by developing methods with
-the capability to identify knowledge gaps.
-
-\autoref{chap:novelty-intro} introduces novelty detection under a
-statistical view point and how to interpret it as a threshold function.
-A simple approach on how to perform novelty on very simple graphs
-is given together with some results on the impact of using semi-supervised
-novelty detection to improve the system performance.
-
-\autoref{chap:novelty} presents the developed techniques developed to identify
-novelty on the semantic representation.
-
-\autoref{chap:conclusions} draws conclusions on the developed work and presents
-interesting directions for future works.
+\input{introduction}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\chapter{Semantic Mapping}\label{chap:semantic-mapping}
-
-% Advantages of Multi-modal approaches
-%
-
-\section{Dora Architecture Overview}
-\subsection{System organization}
-\subsubsection{Sensory Layer}
-\subsubsection{Categorical Layer}
-\subsubsection{Place Layer}
-\subsubsection{Conceptual Layer}
-
-\section{Features}
-
-\section{Conceptual Knowledge}
-\section{Conceptual Map}
-\label{sec:clues-from-low-level}
+\input{semantic-mapping}
 

### 2011-06-01 thesis/introduction.tex
Spplited Introduction and Semantic Mpaping chapter out of contents.tex

@@
-\chapter{Introduction} \label{chap:intro}
+\chapter{Introduction}
+% Introduce:
+%  Mobile Robotics, A.I.
+%  Interaction with Humans
+%  Need for concepts/semantic information
+%  Dynamic human environment
+%  Non-feasibility of hard-coded concepts
+%  Requirements to detect novelty and handle it.
+For a long time humanity has fantasized that one day robots will walk among us.
+They will move and be able to interact with us. Understand our concepts and
+be able to reason.
 
-\section*{}
+They need to possess ability to adapt to situations as its infeasible to rely
+on extensive man-work to tag objects, map space and code all the aspects that
+make up our human-reality.
 
-AO primeiro capítulo da dissertação deve servir para apresentar o
-enquadramento e a moti\-va\-ção do trabalho e para identificar e
-definir os problemas que a dissertação aborda.
-Deve resumir as metodologias utilizadas no trabalho e termina
-apresentando um breve resumo de cada um dos capítulos
-posteriores.
+% == Introduce the need of semantic representations ==
+%
+% There is a lot of low-level methods (or specific knowledge to robots)
+% But the creation of semantic representations allow to bridge that low-level
+% sensing with high-level concepts that facilitate several high-level Tasks.
 
-Este documento ilustra o formato a usar em dissertações na \gls{FEUP}, não
-servindo de exemplo sobre os conteúdos a usar.
-São dados exemplos de margens, cabeçalhos, títulos, paginação, estilos
-de índices, etc. 
-São ainda dados exemplos de formatação de citações, figuras e tabelas,
-equações, referências cruzadas, lista de referências e índices.
+% Importance of mapping and localization in a robot as well as reasoning on
+% properties of each space.
+% Humans give labels to space that characterized the properties and activities
+% expected to be performed on them.
+% Such information is of great interest to be structured and organized in such
+% a way reasoning and planning can be implemented.
+%
+% It also holds an important role in long-term planning and stability as it
+% hides the space-time local details and allows to focus on high-level thinking.
 
-Uma recolha de normas existentes sobre este assunto pode ser
-encontrada em~\cite{kn:Mat93}.
+% Deal with uncertainty -> need for probabilistic models.
+% Deal with semantic    -> need of structured models such as graphical models.
 
-\begin{quote}
-  ``Like the Abstract, the Introduction should be written to engage the
-  interest of the reader. It should also give the reader an idea of
-  how the dissertation is structured, and in doing so, define the
-  thread of the contents.''~\cite{kn:Tha01} 
-\end{quote}
+Example of what kind of an high-level thinking on space and its semantics can
+provide: Get me a beer/milk concept.
 
-Blabla \gls{Kinect}
+% == Reasoning about need for Novelty Detection ==
+%
+% Dynamic environment (inability to know the environment or even to be possible
+% to map all semantic concepts to interact with humans).
+%
+% Knowledge-Awareness (knowing what we know... And detecting novel cases)
+% Identify gaps in the semantic knowledge.
+% Automatic detection and learning of novel concepts.
+%
+% Increase robustness
+% Self-Extending
+%
 
-\section{Context} \label{sec:context}
+\section{Problem and Goals}
+% Extend a probabilistic graphical modelling framework with capabilities to
+% detect novelty.
+% Both in terms of novel classes as of novel structure.
+%
+Given a probabilistic structure representing the sensed conceptual knowledge
+obtained by the agent develop methods able to detect novelty present
+either on new semantic concepts (new classes) or even on structure that was
+previously unknown.
 
-Esta secção descreve a área em que o trabalho se insere, podendo
-referir um eventual projecto de que faz parte e apresentar uma breve
-descrição da empresa onde o trabalho decorreu.
+% Yeah thats what you call dream :P
+In an extreme ideal case the agent should be able to go from zero-knowledge to
+understanding presence of certain objects (as generators of a set of sensed
+sensed properties grouped locally), understand areas within rooms (sink-area on
+on a kitchen), rooms as separated by doors, and understand environments such
+as office, home, warehouse, spaceship.
 
-\section{Motivation and Goals} \label{sec:goals}
+For that the probabilistic semantic representation presented by
+\cite{andrzej2011phd} is used.
 
-Apresenta a motivação e enumera os objectivos do trabalho terminando
-com um resumo das metodologias para a prossecução dos objectivos.
+% If everything goes fine... 2 of the "Future Directions" proposed are 
+% Novelty Detection and Learning of Novel Concepts:
+%  Identify Gaps in Spatial and Semantic Knowledge -> Addressed
+%  Performing learning of new concepts -> Outside Scope / No Time
+%
+% Using Properties for Space Segmentation
+%  If we move on detecting novel structures on the graph we are addressing this
+%  issue.
 
-\section{Thesis Outline} \label{sec:outline}
+\section{Thesis Outline}
+The rest of this thesis is structured as follows:
 
-The rest of this thesis is structured as follows\dots
+\autoref{chap:background} introduces background concepts for understanding the
+presented thesis. In particular \autoref{sec:graphical-models} introduces
+\emph{probabilistic graphical models} that lay the base modelling tool for
+the used structured semantic representation.
+
+\autoref{chap:semantic-mapping} describes the system proposed by
+\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
+that is responsible for the creation of the \emph{probabilistic semantic
+representation} that this thesis aims at improving by developing methods with
+the capability to identify knowledge gaps.
+
+\autoref{chap:novelty-intro} introduces novelty detection under a
+statistical view point and how to interpret it as a threshold function.
+A simple approach on how to perform novelty on very simple graphs
+is given together with some results on the impact of using semi-supervised
+novelty detection to improve the system performance.
+
+\autoref{chap:novelty} presents the developed techniques developed to identify
+novelty on the semantic representation.
+
+\autoref{chap:conclusions} draws conclusions on the developed work and presents
+interesting directions for future works.

### 2011-06-01 thesis/semantic-mapping.tex
Spplited Introduction and Semantic Mpaping chapter out of contents.tex

\chapter{Semantic Mapping}\label{chap:semantic-mapping}

% Advantages of Multi-modal approaches
%

\section{Dora Architecture Overview}
\subsection{System organization}
\subsubsection{Sensory Layer}
\subsubsection{Categorical Layer}
\subsubsection{Place Layer}
\subsubsection{Conceptual Layer}

\section{Features}

\section{Conceptual Knowledge}
\section{Conceptual Map}
\label{sec:clues-from-low-level}

### 2011-06-02 thesis/macros.tex
Still refining the thesis structure and some likely content.
Added TiKZ templates for factor graphs.
Used templates from Laura Dietz:
http://www.mpi-inf.mpg.de/~dietz/probabilistic-models-tikz.zip

@@
 }
+
+% There seems to exist some kind of standard on using CHapter with capital.
+\def\chapterautorefname{Chapter}
+
+% Based on: http://lists.cs.princeton.edu/pipermail/topic-models/2010-December/001081.html
+% http://www.mpi-inf.mpg.de/~dietz/probabilistic-models-tikz.zip
+% by Laura Dietz (dietz at mpi-inf.mpg.de)
+%
+\usepackage{color}
+\usepackage{array}
+\usepackage{verbatim}
+\usepackage{float}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{esint}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
+%% Because html converters don't know tabularnewline
+\providecommand{\tabularnewline}{\\}
+\floatstyle{ruled}
+\newfloat{algorithm}{tbp}{loa}[chapter]
+\floatname{algorithm}{Algorithm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
+\usepackage{float}
+\floatstyle{ruled}
+\newfloat{algorithm}{tbp}{loa}
+\floatname{algorithm}{Algorithm}
+\newfloat{genmodel}{h}{loa}
+\floatname{genmodel}{Generative Process}
+\usepackage[noend]{algorithmic}
+\newcommand{\forbody}[1]{ #1 \ENDFOR}
+\newcommand{\ifbody}[1]{ #1  \ENDIF}
+\newcommand{\whilebody}[1]{ #1  \ENDWHILE}
+\renewcommand{\algorithmicprint}{\textbf{draw}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+
+\usepackage{euscript}
+
+\DeclareSymbolFont{rsfscript}{OMS}{rsfs}{m}{n}
+\DeclareSymbolFontAlphabet{\mathrsfs}{rsfscript}
+
+% PDF formatting instructions for A4
+\pdfpagewidth=210mm % for pdflatex
+\pdfpageheight=296mm % for pdflatex
+
+%%%%%%%% begin tikz %%%%%%
+\usepackage{tikz,tkz-base}
+\usetikzlibrary{shapes,decorations,shadows}
+\usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{decorations.shapes}
+\usetikzlibrary{fadings}
+\usetikzlibrary{patterns}
+\usetikzlibrary{calc}
+\usetikzlibrary{decorations.text}
+\usetikzlibrary{decorations.footprints}
+\usetikzlibrary{decorations.fractals}
+\usetikzlibrary{shapes.gates.logic.IEC}
+\usetikzlibrary{shapes.gates.logic.US}
+\usetikzlibrary{fit,chains}
+\usetikzlibrary{positioning}
+\usepgflibrary{shapes}
+\usetikzlibrary{scopes}
+\usetikzlibrary{arrows}
+\usetikzlibrary{backgrounds}
+
+\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
+minimum size=20pt, font=\fontsize{10}{10}\selectfont},
+obs/.style={latent,fill=gray!25},
+const/.style={rectangle, inner sep=0pt},
+factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=0pt},
+>={triangle 45}}
+
+% shapename, fitlist, caption, pos
+\newcommand{\plate}[4]{
+\node (invis#1) [draw, transparent, inner sep=1pt,rectangle,fit=#2] {};
+\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=2pt,yshift=1pt] {\footnotesize{#3}};
+\node (#1) [draw,inner sep=1pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+}
+
+\newcommand{\shiftedplate}[5]{
+\node (invis#1) [draw, transparent, inner sep=0 pt,rectangle,fit=#2] {};
+\node (capt#1) [#5, xshift=2pt] {\footnotesize{#3}};
+\node (#1) [draw,inner sep=2pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+}
+
+%shapename, pos, caption, in1, in2, out, captpos
+\newcommand{\twofactor}[7]{
+\node (#1) [factor] at #2 {};
+\node (capt#1) [#7 of #1]{\footnotesize{#3}};
+\draw [-] (#4) -- (#1) ;
+\draw [-] (#5) -- (#1) ;
+\draw [->,thick] (#1) -- (#6);
+}
+
+%shapename, pos, caption, in, out, captpos
+\newcommand{\factor}[6]{
+\node (#1) [factor] at #2 {};
+\node (capt#1) [#6 of #1]{\footnotesize{#3}};
+\draw [-] (#4) -- (#1) ;
+\draw [->,thick] (#1) -- (#5);
+}
+
+% name, --, caption, pos
+\newcommand{\nofactor}[4]{
+\node (#1) [factor, #2]  {};
+\node (capt#1) [#4 of #1]{\footnotesize{#3}};
+}
+
+%shapename,  fitlist, caption
+\newcommand{\namedgate}[3]{
+\node (invisgate#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (gatecapt#1) [ above right=0 pt of invisgate#1.north west, xshift=-1pt ] {\footnotesize{#3}};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgate#1)(gatecapt#1)]{};
+
+}
+
+%shapename,  fitlist, caption
+\newcommand{\gate}[3]{
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=#2]{};
+}
+
+%shapename,  fitlist1, fitlist2, caption1, caption2
+\newcommand{\vertgate}[5]{
+\node (invisgateleft#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (gatecaptleft#1) [ above left=0 pt of invisgateleft#1.north east, xshift=1pt ]{\footnotesize{#3}};
+\node (invisgateright#1) [rectangle, draw, transparent,  fit=#4] {};
+\node (gatecaptright#1) [ above right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
+\draw [-, dashed] (#1.north) -- (#1.south);
+}
+
+\newcommand{\vertgateSpec}[5]{
+\node (invisgateleft#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (gatecaptleft#1) [ above left=0 pt of invisgateleft#1.north east, xshift=1pt ]{\footnotesize{#3}};
+\node (invisgateright#1) [rectangle, draw, transparent,  fit=#4] {};
+\node (gatecaptright#1) [ above right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
+\draw [-, dashed] (#1.70) -- (#1.290);
+}
+
+\newcommand{\horgate}[5]{
+\node (invisgateleft#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (gatecaptleft#1) [ above right=0 pt of invisgateleft#1.south west, xshift=1pt ]{\footnotesize{#3}};
+\node (invisgateright#1) [rectangle, draw, transparent,  fit=#4] {};
+\node (gatecaptright#1) [ below right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
+\draw [-, dashed] (#1.west) -- (#1.east);
+}
+
+\newcommand{\horogate}[5]{
+\node (invisgateleft#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (invisgateright#1) [rectangle, draw, transparent,  fit=#4] {};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(invisgateright#1)]{};
+\node (gatecaptleft#1) [ above right=0 pt of #1.west, xshift=0pt ]{\footnotesize{#3}};
+\node (gatecaptright#1) [ below right=0 pt of #1.west, xshift=0pt ] {\footnotesize{#5}};
+
+\draw [-, dashed] (#1.west) -- (#1.east);
+}
+
+\newcommand{\vertogate}[5]{
+\node (invisgateleft#1) [rectangle, draw, transparent,  fit=#2] {};
+\node (invisgateright#1) [rectangle, draw, transparent,  fit=#4] {};
+\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(invisgateright#1)]{};
+\node (gatecaptleft#1) [ below left=0 pt of #1.north, xshift=0pt ]{\footnotesize{#3}};
+\node (gatecaptright#1) [ below right=0 pt of #1.north, xshift=0pt ] {\footnotesize{#5}};
+
+\draw [-, dashed] (#1.north) -- (#1.south);
+}

### 2011-06-02 thesis/novelty.tex
Added TiKZ templates for factor graphs.
Used templates from Laura Dietz:
http://www.mpi-inf.mpg.de/~dietz/probabilistic-models-tikz.zip

@@
           the variables types and graph structure hold true.]{
-          Hello get a graph here.}
+\begin{tikzpicture}
+\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+\&
+\&
+\node (room3) [latent] {$R_3$}; \& \\
+\& \& \& \\
+\&
+\node (room1) [latent] {$R_1$}; \& \&
+\node (room2) [latent] {$R_2$}; \\
+\& \& \& \\
+\node (shape1f) [factor] {}; \&
+\node (appearance1f) [factor] {}; \&
+\node (object1f) [factor] {}; \&
+\node (factor2) [factor] {}; \\
+\node (shape1l) [obs] {$S_p$}; \&
+\node (appearance1l) [obs] {$A_p$}; \&
+\node (object1l) [obs] {$O_p$}; \&
+\node (prop2) [obs] {$X_p$}; \\
+};
+
+\draw [-] (room1) -- (shape1f) -- (shape1l);
+\draw [-] (room1) -- (appearance1f) -- (appearance1l);
+\draw [-] (room1) -- (object1f) -- (object1l);
+\draw [-] (room2) -- (factor2) -- (prop2);
+
+\draw [-] (room1) -- (room2) node (r12f) [midway,factor] {};
+\draw [-] (room1) -- (room3) node (r13f) [midway,factor] {};
+\draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+
+\plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+\plate{places2}{(factor2)(prop2)}{$\dots$}{};
+
+\end{tikzpicture} %
+          }
 \qquad\qquad
@@
           has learned before.]{
-          Hello get a graph here.}
+\begin{tikzpicture}
+\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+\&
+\&
+\node (room3) [latent] {$R_3$}; \& \\
+\& \& \& \\
+\&
+\node (room1) [latent] {$R^N_1$}; \& \&
+\node (room2) [latent] {$R_2$}; \\
+\& \& \& \\
+\nofactor {shape1f}{}{$U$} {right=0pt}; \&
+\nofactor {appearance1f}{} {$U$} {right=0pt}; \&
+\nofactor {object1f}{} {$U$} {right=0pt}; \&
+\node (factor2) [factor] {}; \\
+\node (shape1l) [obs] {$S_p$}; \&
+\node (appearance1l) [obs] {$A_p$}; \&
+\node (object1l) [obs] {$O_p$}; \&
+\node (prop2) [obs] {$X_p$}; \\
+};
+
+\draw [-] (room1) -- (shape1f) -- (shape1l);
+\draw [-] (room1) -- (appearance1f) -- (appearance1l);
+\draw [-] (room1) -- (object1f) -- (object1l);
+\draw [-] (room2) -- (factor2) -- (prop2);
+
+\draw [-] (room1) -- (room2) node (r12f) [midway,factor] {};
+\draw [-] (room1) -- (room3) node (r13f) [midway,factor] {};
+\draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+\node (captr12f) [above=0pt of r12f] {\footnotesize{$U$}};
+\node (captr13f) [above=0pt of r13f] {\footnotesize{$U$}};
+
+\plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+\plate{places2}{(factor2)(prop2)}{$\dots$}{};
+
+\end{tikzpicture} %
+}
 
 \caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
-         a class of a single variable $a$, assuming all the other variables
+         a class of a single variable $R_1$, assuming all the other variables
          classes and graph structure are correctly modelled by the agent

### 2011-06-02 semantic-nd/article.tex
Added .jpgs in .pdf format
Commit changes in article that were stashed (I had forgotton about them).

@@
 \centering
-\includegraphics[width=0.9\textwidth]{figures/conceptual-layer.jpg}
+\includegraphics[width=0.9\textwidth]{figures/processes.pdf}
 \caption{\label{fig:conceptual-layer}Interaction of the sub-systems
@@
 \centering
-\includegraphics[width=0.90\textwidth]{figures/chain-graph.png}
+\includegraphics[width=0.80\textwidth]{figures/chaingraph.pdf}
 \caption{\label{fig:chain-graph}Example chain-graph produced by the \emph{conceptual layer}.}
@@
 % Introduce novelty detection
-Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about~\cite{markou2003novelty}.
-It is harder than classification as only positive samples of a class are available
-rendering normal classification methods invalid.
+Novelty detection deals with detecting that a data sample belongs to a class
+originated by a distribution other than the ones the system knows about \citep{markou2003novelty}.
+It is harder than classification as only positive samples of the class are available
+rendering normal classification methods unusable.
 

### 2011-06-02 semantic-nd/article.tex
Fixed citing style.

@@
 Novelty detection deals with detecting that a data sample belongs to a class
-originated by a distribution other than the ones the system knows about \citep{markou2003novelty}.
+originated by a distribution other than the ones the system knows about~\cite{markou2003novelty}.
 It is harder than classification as only positive samples of the class are available

### 2011-06-02 thesis/novelty-intro.tex
Changes on review of novelty detection.

--- 2011-05-29 thesis/novelty-intro.tex
@@
 models is presented: it shows how to use the introduced ratio to obtain a
-novelty detection system and analyses the performance impact by obtaining more
-data and by using an approximation on the unconditional probability.
-
-\begin{itemize}
-\item Introduction
-\item History
-\item Approaches to Novelty Detection
-\end{itemize}
+novelty detection system and analyses the performance impact by increasing
+the amount of sensed data and by using an approximation on the unconditional
+probability.
 
 \section{Novelty Detection Review and Related Work}
-% Introduce novelty detection
-Novelty detection deals with detecting that a new input was generated by a class
-other than the ones the system knows about~\cite{markou2003novelty}.
-It is harder than classification as only positive samples of a class are available
-rendering normal classification methods invalid.
+Novelty detection, also called outlier or anomaly detection, is the
+classification problem related with identification of new or unknown data
+patterns that the system is not aware of~\cite{markou2003novelty}.
 
-Adding capabilities to an agent to detect novel samples allows to increase its
-reliability. Novelty signal can then be interpreted by the system to proceed
-with more caution as its knowledge does not correctly describe reality.
+The ability to identify novel cases is crucial in any autonomous system
+that is deployed to unknown or to uncontrolled environments, as it gives the
+system the ability to detect something is not conforming to it knowledge and
+therefore it should proceed with caution.
+It has several applications such as fault detection, intrusion detection,
+detection of masses in mammograms, hand written digit recognition and many
+others.
+% TODO: list usage of novelty detection together with references to papers
+% with them.
+
+It is on the nature of unknown environments and/or in the infeasibility of
+training the system on all possible object classes that lies the complexity of
+novelty detection. It is only possible to obtain samples representing positives
+examples of known cases. The lack of negative examples renders normal
+classification methods unusable.
+
+In order to be usable in real world applications, novelty detection methods
+have to overcome a series of obstacles:
+be able to generalize while still detecting novelty,
+be resistant to noisy features,
+ability to scale in feature dimension,
+deal with multiple classes and performing detection efficiently (i.e. many
+autonomous systems require real-time or close to real-time performance).
+
+\subsection{Review}
+Many approaches have been made based on statistical methods. Those often model
+the training data with statistic distributions and then perform tests whether a
+given sample comes from the same distribution.
+
+Parametric approaches: probabilistic /gmm approaches, hidden markov models,
+hypothesis testing
+
+Non-Parametric: knn Approachs, parzen density estimation, string matching,
+clustering.
+
+One of the simplest approach for novelty detection is pose the problem as a
+probability density estimation problem and use the expected probability of a
+given input to trigger the case as novel.
+
+Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
+In order to reduce the input size, those approaches often employ
+dimension-reduction techniques such as \gls{PCA}.
+
+This approach is used in \cite{bishop1994novelty} where \emph{novelty detection}
+is added to a \emph{neural network} system. On it the density estimation on the
+input is calculated using a \emph{Parzen window} over training-data and that
+value is used as a novelty measure. This way they detect novel data that their
+network has not seen before and the output of their system is then not expected
+to be reliable.
+
+A slightly different approach is done by one-class \gls{SVM} approaches that
+try to distinguish the class by separating the traing set from all the other
+points in the input space. They try to achieve that by enclosing the training
+set by some structure (i.e. an hyper-sphere)\cite{bennett2000support}.
+Those approaches have been made in line with Vapnik principle of not solving
+something hard. As although having access to a perfect probability distribution
+of the input would solve the problem, creating such a function is harder than
+simply creating a boundary between known data and novel data
+\cite{scholkopf2000support}.
+
+Error reconstruction methods have also been used for novelty detection.
+They use the assumption that the class to be defined lies on a manifold embedded
+on the input space of higher dimensions. By using dimensional reduction
+techniques, they try to defined that manifold and calculate how far a new sample
+is from it.
+One of the most common methods used for that is
+\gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick to extend
+\gls{PCA} and perform a nonlinear dimensionality reduction on the input.
+This technique has been successfully used in novelty detection by
+\cite{Hoffmann2007863}.
+
+\subsection{Related Work}
+More in line with novelty detection in scene categorization and semantic
+structures, the work of PLISS\dots
+
+\section{Novelty Detection as a Threshold}
 
@@
 
-\section{Novelty Detection as a Threshold}
 When dealing with a statistic point of view, noisy data or unstable features,

### 2011-06-02 thesis/novelty.tex
Changes on review of novelty detection.

@@
 That can be done by extending all the variables on the factor graph representing
-the semantic map with an additional class \emph{novel} and extending all the
+the semantic map with an additional state \emph{novel} and extending all the
 factors with information on the unconditional distribution.

### 2011-06-03 thesis/macros.tex
Changed TikZ styles to draw factor graphs.

@@
 
-\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+
+\tikzset{latent/.style={circle,fill=white,draw=red,thick,inner sep=1pt, 
 minimum size=20pt, font=\fontsize{10}{10}\selectfont},
@@
 const/.style={rectangle, inner sep=0pt},
-factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=0pt},
+factor/.style={rectangle, fill=red,minimum size=7pt, inner sep=0pt},
+-/.style={color=red, thick},
 >={triangle 45}}
@@
 \node (invis#1) [draw, transparent, inner sep=1pt,rectangle,fit=#2] {};
-\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=2pt,yshift=1pt] {\footnotesize{#3}};
-\node (#1) [draw,inner sep=1pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=0pt,yshift=-9pt] {\raisebox{0pt}[0pt]{\footnotesize{#3}}};
+\node (#1) [draw=black!50,thick,inner sep=3pt,rectangle,rounded corners,fit=(invis#1) (capt#1),#4] {};
 }

### 2011-06-05 thesis/novelty-intro.tex
Several changes in novelty-intro chapter.
Fixed explanation on need of threshold function.
Reorganized Practical Example section.

@@
 system the ability to detect something is not conforming to it knowledge and
-therefore it should proceed with caution.
+therefore should proceed with caution.
 It has several applications such as fault detection, intrusion detection,
@@
 training the system on all possible object classes that lies the complexity of
-novelty detection. It is only possible to obtain samples representing positives
-examples of known cases. The lack of negative examples renders normal
+novelty detection: it is only possible to obtain samples representing positives
+examples of known cases and the lack of negative examples renders normal
 classification methods unusable.
@@
 Many approaches have been made based on statistical methods. Those often model
-the training data with statistic distributions and then perform tests whether a
-given sample comes from the same distribution.
+the training data with statistic distributions and then perform hypothesis tests
+For example performing checking whether two distributions are the same,
+distance to mean expected value or quartiles, etc\dots
+In some sense all they define a distance measure to normality and employ a
+threshold on that.
 
-Parametric approaches: probabilistic /gmm approaches, hidden markov models,
-hypothesis testing
+A common approach is to use density estimation and use the expected probability
+of a sample to trigger the sample as novel. Examples are usage of
+Gaussian Mixture Models and Parzen-window estimators. In order to be effective
+those use data as close as possible to the input features.
+And dimension-reduction techniques such as \gls{PCA} are used to make density
+estimation more feasible. Note that as dimension increases an exponential number
+of data samples are required to approach the density with a given quality.
 
-Non-Parametric: knn Approachs, parzen density estimation, string matching,
-clustering.
-
-One of the simplest approach for novelty detection is pose the problem as a
-probability density estimation problem and use the expected probability of a
-given input to trigger the case as novel.
-
-Examples of such approaches are Gaussian Mixtures Models and Parzen-window estimators.
-In order to reduce the input size, those approaches often employ
-dimension-reduction techniques such as \gls{PCA}.
-
-This approach is used in \cite{bishop1994novelty} where \emph{novelty detection}
-is added to a \emph{neural network} system. On it the density estimation on the
-input is calculated using a \emph{Parzen window} over training-data and that
-value is used as a novelty measure. This way they detect novel data that their
-network has not seen before and the output of their system is then not expected
-to be reliable.
+\cite{bishop1994novelty} uses that approach by employing a Parzen-window to
+estimate the density of the training data on a given input feed into a
+neural-network. By triggering on the calculated density they detect samples
+that differ from the training data and consider their neural-network output
+to be unreliable as the samples are distinct from what the network was trained
+with.
 
@@
 points in the input space. They try to achieve that by enclosing the training
-set by some structure (i.e. an hyper-sphere)\cite{bennett2000support}.
+set by some structure (i.e. an hyper-sphere)~\cite{bennett2000support}.
 Those approaches have been made in line with Vapnik principle of not solving
@@
 One of the most common methods used for that is
-\gls{K-PCA}~\citep{scholkopf1997kernel}, which uses the kernel-trick to extend
+\gls{K-PCA}~\cite{scholkopf1997kernel}, which uses the kernel-trick to extend
 \gls{PCA} and perform a nonlinear dimensionality reduction on the input.
@@
 More in line with novelty detection in scene categorization and semantic
-structures, the work of PLISS\dots
+structures, the work of \gls{PLISS}~\cite{ranganathan2010pliss}.\dots
 
@@
 When dealing with a statistic point of view, noisy data or unstable features,
-a decision to classify a sample $n$ as novel has an associated error probability
-$P(\overline{novel}|n)$.
-Accordingly, on a system that classifies a set of inputs $N$ as novel, the
-probability of incorrectly classifying a sample as novel is given by the
-equation:
+a decision to classify a sample $x$ as novel reduces rejection rate by
+$P(novel|x)P(x)$ by introducing an associated error probability of $P(\overline{novel}|x)P(x)$.
+Generalizing to a system that classifies a set of inputs $N$ as novel, the
+true positive and false negative rates arise:
+
+\begin{equation}
+\label{eq:true-positive}
+P(\textnormal{true positive}) = \sum_{x \in N}{P(novel|x)P(x)}
+\end{equation}
 
@@
 \label{eq:false-positive}
-P(\textnormal{false positive}|\textnormal{positive}) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+P(\textnormal{false positive}) = \sum_{x \in N}{P(\overline{novel}|x)P(x)}
 \end{equation}
 
-Theres a problem with this proof. I need to review it. The original explanation
-for me to model it as a threshold function required to have both true positive
-and false positive ratios. But for the paper I though I could do it with only
-the FPR, but its clear wrong.
+The performance of a novelty detection system can be measured on these two
+rates: the true-positive rate (\autoref{eq:true-positive}) describes the
+interest in detecting as much as possible the novel cases;
+and the false-positive rate (\autoref{eq:false-positive}) the interest in
+having the smallest error possible.
+As a given sample $x$ is included in $N$ both the true-positive and
+false-positive rates increase or stay equal\footnote{
+Note that both are described as summations of products of probabilities
+(non-negative functions).} describing the trade-of any detection system
+has to consider: by increasing it detection rate it also has to increase
+the error rate.
 
-Therefore any novelty system that considers to acknowledge a sample
-$a$ as novel should before detect novelty on any sample $b$ satisfying
-$P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that would always yield a reduced the error probability described in
-\autoref{eq:false-positive} as shown:
+An optimal novelty detector can then be formulated as selecting a subset $N$
+that achieves a given true-positive rate with the minimal false-positive
+rate possible. This is equivalent to the \emph{continuous knapsack problem}
+which allows a greedy solution by sorting the items by value per weight.
+Therefore a novelty system before classifying a sample $a$ as novel should
+classify any sample $b$ satisfying \autoref{eq:knapsack} as that would achieve
+a lower false-positive rate given a fixed true-positive rate.
 
 \begin{equation}
-\textnormal{Proof: to be optimal any system accepting }a\textnormal{ should also accept }b
+\label{eq:knapsack}
+\frac{P(\overline{novel}|b)P(b)}{P(novel|b)P(b)} < \frac{P(\overline{novel}|a)P(a)}{P(novel|a)P(a)}
 \end{equation}
 
-For the sake of completeness its show here that it also reduces the probability
-of false negatives.
+This relation between $a$ and $b$ can then be simplified into:
 
 \begin{equation}
-P(\textnormal{false negative}|\textnormal{negative}) = \frac{\sum_{n \notin N}{P(novel|n)P(n)}}{\sum_{n \notin N}{P(n)}}
+P(\overline{novel}|b) < P(\overline{novel}|a)
 \end{equation}
@@
 Since there is only interest in maintaining the same order-relation as
-$P(\overline{novel}|x)$ the absolute value is not needed and any constant
-factor can be dropped. Leaving a ratio between a \emph{conditional} and
-\emph{unconditional probabilities} suitable for implementing novelty detection
+$P(\overline{novel}|x)$ any constant factor can be dropped.
+Leaving a ratio, between a \emph{conditional} and
+\emph{unconditional probability}, suitable for implementing novelty detection
 via thresholding.
@@
 The conditional probability $P(x|\overline{novel})$ describes the distribution
-of the non-novel samples. In case the labelled data available for the agent
-to learn a concept comes from this same distribution the correct approach
-is to use it as prior-information for modelling the conditional probability.
+of the non-novel samples on the original distribution. In case the labelled data
+available for the agent to learn a concept comes from the same distribution
+where the system will run the correct approach is to use it as prior-information
+for modelling the conditional probability.
 
-Note that it is important for the labelled data to come from the same
-distribution as where the system will run, as otherwise it will contain some
-bias (for example introduced by an incorrect filtering)
-
-Therefore the correct approach is to try to model it
+Note that it is important for the labelled data to come from that same
+distribution, otherwise it will contain some bias and lead to an incorrect
+ordering. The bias, for example, can be introduced by an incorrect filtering or
+data balancing.
 
@@
 whether a given sample conditional probability arises from it belonging to
-the known concept or from it likelihood of being sampled.
+the known concept or from the likelihood of being sampled.
 
@@
 But in the case where the set of features $x$ is variable it cannot be
-discarded. There $P(x)$ plays a role in levering all the conditional
+discarded. There $P(x)$ also plays a role in levering all the conditional
 probabilities on different sets of variables into the same measure units
@@
 \label{sec:unlabelled-data}
-In order to verify the performance of the proposed threshold functions a synthetic dataset
-was generated. To keep it simple only information regarding direct features of
-a room were modelled and no structured knowledge such as room connectivity was taken
-in account.
-The synthetic distribution assumes that an independent and variable size set of features
-$x$ is generated by a given room category.
-In whole there was 11 different room categories and 7 different measured feature
-types. Each feature type can be present more than once in (e.g.: room shape is
-extracted from 2D laser scans in more than one position in the room).
+In order to sum up the presented concepts on novelty detection with a threshold
+function and exemplify how to use graphical models in the context of
+multi-modality room classification a synthetic dataset was generated.
+The dataset was kept simple by only modelling directly sensed features from a
+room, skipping any structural knowledge or sensing model such as room
+connectivity and extra hidden variables.
 
-The room categories were chosen to mimic as close as possible the real features and
-categories existing in reality. And they model different room categories with
-different levels of detail. For example: 1 person office, 2 person office, hallway,
-robot lab.
+In this dataset a room $r$ is seen as an hidden-variable generator of a set of
+features $X$ that are directly sensed by the agent.
+All the sensed features $x$ are all independent given the room category.
+In whole there was 11 different room categories and 7 different feature types.
+Each feature can be sensed more than once (i.e. room shape is extracted from 2D
+laser scans in more than one position in the room), but all those sensed
+instances are considered independent given the room category.
 
-From the distribution, 100 labelled samples for 5 of the 11 room categories were
-drawn to represent the known concepts and 1000 unlabelled samples were drawn from
-all the room categories for learning the unconditional probability distribution.
-Using those samples, factors were learnt for the graphs used to model the
-conditional distribution and the independent unconditional distribution.
-\autoref{fig:simple-experiment} shows the graph structure used for approximate the
-trained conditional and unconditional distributions.
-Its important to notice that graph $G$ used to model the known classes when given
-enough labelled data is able to exactly learn the conditional distribution as it
-uses the same structure as the created synthetic distribution.
+The room categories were chosen to mimic as close as possible the real features
+and categories existing in reality (i.e. 1 person office, 2 person office,
+hallway, robot lab, etc\dots). A table describing the used synthetic
+distribution is given in \autoref{extra:synthetic-distribution}.
+
+The objective then was to draw a system that, although only trained with
+labelled data from 5 of the 11 room categories, was able to detect novel
+room categories.
+For that 100 labelled samples for the 5 known categories were drawn and 1000 
+unlabelled samples were drawn from all the room categories for learning the
+unconditional probability distribution and measure effect of using unlabelled
+data.
+
+\subsection{Conditional Probability}
+Using the labelled samples, 7 factors $\phi_X(r,x)$ were created, one for each
+feature type, to represent the potential of sensing features $x$ of type $X$ on
+a room of category $r$.
+
+\begin{equation}
+\phi_X(r,x) = number of samples + C
+\end{equation}
+
+With those the probability of sensing a set of features $x$ on a room knowing
+room category is known to the agent can be modelled with a factor graph as
+illustrated on \autoref{fig:simple-cond-graph}.
 
@@
 \centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room) [latent] {$r$}; \& \& \\
+    \& \& \& \& \\
+    \node (f1) [factor] {}; \&
+    \node (f2) [factor] {}; \&
+    \node (f3) [factor] {}; \&
+    \node (fi) [] {\dots}; \&
+    \node (fn) [factor] {}; \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (room) -- (f1) -- (x1);
+  \draw [-] (room) -- (f2) -- (x2);
+  \draw [-] (room) -- (f3) -- (x3);
+  \draw [-] (room) -- (fi);
+  \draw [-] (room) -- (fn) -- (xn);
 
-\subfloat[Graph structure $G$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-cond-graph.pdf}}
-\qquad
-\subfloat[Graph structure $U$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-uniform-graph.pdf}}
-\qquad
-\subfloat[Graph structure $I$.]{
-         \includegraphics[width=0.27\textwidth]{figures/simple-independent-graph.pdf}}
-
-\caption{\label{fig:simple-experiment}The graph structures used to model the
-         conditional and unconditional probability for implementing the novelty
-         thresholds $P_G(x)/P_U(x)$ and $P_G(x)/P_I(x)$.}
+  \node (captf1) [right=2pt of f1] {\footnotesize{$\phi_{X_1}$}};
+  \node (captf2) [right=2pt of f2] {\footnotesize{$\phi_{X_2}$}};
+  \node (captf3) [right=2pt of f3] {\footnotesize{$\phi_{X_3}$}};
+  \node (captfn) [right=2pt of fn] {\footnotesize{$\phi_{X_n}$}};
+\end{tikzpicture}
+\caption{\label{fig:simple-cond-graph}A factor graph modelling the conditional
+probability of sensing a set of features $x$ known that the room category $r$ is
+one of the known classes.}
 \end{figure}
 
-% Describe how to obtain the 3 thresholds functions seen on the graphs.
-Using the learned models $G$, $U$ and $I$, two thresholds were trained:
-$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
-and $P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
+\subsection{Unconditional Probability}
+With no knowledge on the unconditional probability the correct approach is to
+assume a uniform distribution.
+That is represented in factor graphs by a graph without any factors.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+\end{tikzpicture}
+\caption{\label{fig:simple-uniform-graph}A factor graph modelling a uniform
+         distribution over the sensed set of features $x$.}
+\end{figure}
+
+Very often there is extra knowledge that can be obtained about the distribution
+of the variable that helps to model the unconditional distribution.
+In this practical example the access to unlabelled data is explored as access to
+it is very common on applications in robotics.
+
+Note that the sensed variables are dependent between each other when the room
+category $r$ is not known. With that the correct approach would be to train a
+factor that is able to correlate all the sensed variables as seen on
+\autoref{fig:simple-all-dependent}.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (fac) [factor] {}; \& \& \\
+    \& \& \& \& \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (fac) -- (x1);
+  \draw [-] (fac) -- (x2);
+  \draw [-] (fac) -- (x3);
+  \draw [-] (fac) -- (xn);
+  \draw [-] (fac) -- (xi);
+\end{tikzpicture}
+\caption{\label{fig:simple-all-dependent}A general factor graph able
+         to model any unconditional distribution on the sensed variables
+         requires a factor connecting all of them.}
+\end{figure}
+
+Nonetheless such an approach suffers from
+the \emph{curse of dimensionality} as the number of sensed features and feature
+types increases.
+
+A relaxed solution can be obtained by assuming in that the sensed features are
+independent conditional 
+Although the sensed variables are considered independent given the room category
+$r$, when $r$ is not able to model  hidden
+variable $r$. The approach should then be to 
+In order to correctly approach the unconditional probability the unlabelled
+data should be used 
+Without the assumption of the room category being known it is no longer easy
+to model the distribution between all sensed variables.
+Under the assumption
+
+Note that although sensed variables are independent, they are only so when $r$
+is given. Requiring a factor to be connect between all variables to model their
+dependence.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \& \& \\
+    \node (f1) [factor] {}; \&
+    \node (f2) [factor] {}; \&
+    \node (f3) [factor] {}; \&
+    \node (fi) [] {\dots}; \&
+    \node (fn) [factor] {}; \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (f1) -- (x1);
+  \draw [-] (f2) -- (x2);
+  \draw [-] (f3) -- (x3);
+  \draw [-] (fi);
+  \draw [-] (fn) -- (xn);
+\end{tikzpicture}
+\caption{\label{fig:simple-independent-graph}A factor graph modelling an
+         independent distribution over the sensed set of features $x$.}
+\end{figure}
+
+\subsection{Threshold Functions}
+Three threshold functions were created using the knowledge on the synthetic
+distribution and the models learnt from the sample data:
+$G$ on \autoref{fig:simple-cond-graph},
+$U$ on \autoref{fig:simple-uniform-graph} and
+$I$ on \autoref{fig:simple-independent-graph}.
+
+\begin{description}
+\item[exact]
+$P(x|\overline{novel})/P(x)$
 Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
@@
 presented thresholds are from optimal.
+
+\item[uniform]
+$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
+
+\item[independent]
+$P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
+\end{description}
 
@@
 
-We can see that as the system gains more semantic information it becomes easier
-to detect novelty. The input space size increases and allows the several existing
-classes to become more easily distinguished.
+It is possible to see that as the system gains more semantic information it
+becomes easier to detect novelty. The input space size increases and allows the
+several existing classes to become more easily distinguished.
 

### 2011-06-05 thesis/novelty-intro.tex
Explained independent approach for the unconditional factor.

@@
 \section{Novelty Detection as a Threshold}
-
 % Explain threshold approach to novelty detection
@@
 Nonetheless such an approach suffers from
-the \emph{curse of dimensionality} as the number of sensed features and feature
-types increases.
+the \emph{curse of dimensionality}: as the number of sensed features and feature
+types increases exponential ammounts of data are needed to describe it.
+Having interest in avoiding the issues associated with an high-connected factor
+an assumption on independent features can be done and then the unconditional
+probability can be modelled with a fully disconnected variables, but with
+factors that account for existent bias on each single feature (as visible on
+graph in \autoref{fig:simple-independent-graph}).
 
-A relaxed solution can be obtained by assuming in that the sensed features are
-independent conditional 
-Although the sensed variables are considered independent given the room category
-$r$, when $r$ is not able to model  hidden
-variable $r$. The approach should then be to 
-In order to correctly approach the unconditional probability the unlabelled
-data should be used 
-Without the assumption of the room category being known it is no longer easy
-to model the distribution between all sensed variables.
-Under the assumption
-
-Note that although sensed variables are independent, they are only so when $r$
-is given. Requiring a factor to be connect between all variables to model their
-dependence.
+Note than in the case there is only one sensed feature, the distribution
+probability modelled by an assumption of independent features correctly models
+unconditional probability. And it deviates from it as more features are sensed.
+Though the individual factors associated with each variable try to unwrapped any
+possible existent bias on each dimension and therefore it is expected to be a
+better estimate than assuming a uniform distribution.
 
@@
 \item[exact] 
-$P(x|\overline{novel})/P(x)$
-Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
-and a perfect threshold function could also be created to test how far the
-presented thresholds are from optimal.
+- since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
+and the ordering function $P(x|\overline{novel})/P(x)$ could be created
+to test how far the other presented thresholds are from optimal.
 
 \item[uniform]
-$P_G(x)/P_U(x)$ assuming a uniform unconditional distribution
+- by assuming a uniform unconditional distribution the ordering function is
+  given by $P_G(x)/P_U(x)$.
 
 \item[independent]
-$P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
+- using the unlabelled data the unconditional distribution can be approximated.
+  In this case, it was approximated with an independent distribution of the
+  sensed features. The independent threshold was implemented with $P_G(x)/P_I(x)$.
 \end{description}

### 2011-06-06 thesis/novelty-intro.tex
Fixes on novelty introduction chapter.

@@
 
-It is on the nature of unknown environments and/or in the infeasibility of
+It is on the nature of unknown environments and in the infeasibility of
 training the system on all possible object classes that lies the complexity of
@@
 
-In order to be usable in real world applications, novelty detection methods
-have to overcome a series of obstacles:
+In order to become practical in real world applications, novelty detection
+methods have to overcome a series of obstacles:
 be able to generalize while still detecting novelty,
@@
 ability to scale in feature dimension,
-deal with multiple classes and performing detection efficiently (i.e. many
-autonomous systems require real-time or close to real-time performance).
+deal with multiple classes and performing detection efficiently as many
+autonomous systems require real-time or close to real-time performance.
 
@@
 Many approaches have been made based on statistical methods. Those often model
-the training data with statistic distributions and then perform hypothesis tests
-For example performing checking whether two distributions are the same,
-distance to mean expected value or quartiles, etc\dots
+the training data with statistic distributions and then perform hypothesis tests.
+For example: testing whether two distributions are the same or using distance to
+mean value or quartiles.
 In some sense all they define a distance measure to normality and employ a
@@
 Gaussian Mixture Models and Parzen-window estimators. In order to be effective
-those use data as close as possible to the input features.
-And dimension-reduction techniques such as \gls{PCA} are used to make density
-estimation more feasible. Note that as dimension increases an exponential number
-of data samples are required to approach the density with a given quality.
+those use data as close as possible to the input features and use
+dimension-reduction techniques such as \gls{PCA} to make density
+estimation feasible. Note that as dimension increases an exponential number
+of data samples would be required to approach the density with the same quality.
 
@@
 A slightly different approach is done by one-class \gls{SVM} approaches that
-try to distinguish the class by separating the traing set from all the other
+try to distinguish novelty by separating the training set from all the other
 points in the input space. They try to achieve that by enclosing the training
-set by some structure (i.e. an hyper-sphere)~\cite{bennett2000support}.
+set by some structure (e.g.\ an hyper-sphere)~\cite{bennett2000support}.
 Those approaches have been made in line with Vapnik principle of not solving
@@
 They use the assumption that the class to be defined lies on a manifold embedded
-on the input space of higher dimensions. By using dimensional reduction
+on an input space of higher dimensions. By using dimensional reduction
 techniques, they try to defined that manifold and calculate how far a new sample
@@
 to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
-between error and rejection: the more the system tries to reduce rejection, the more
+between error and detection: the more the system tries to increase detection, the more
 prone to produce errors it becomes.
@@
 When dealing with a statistic point of view, noisy data or unstable features,
-a decision to classify a sample $x$ as novel reduces rejection rate by
-$P(novel|x)P(x)$ by introducing an associated error probability of $P(\overline{novel}|x)P(x)$.
+a decision to classify a sample $x$ as novel increases detection rate by
+$P(novel|x)P(x)$ and introduces an associated error rate of $P(\overline{novel}|x)P(x)$.
 Generalizing to a system that classifies a set of inputs $N$ as novel, the
@@
 As a given sample $x$ is included in $N$ both the true-positive and
-false-positive rates increase or stay equal\footnote{
-Note that both are described as summations of products of probabilities
-(non-negative functions).} describing the trade-of any detection system
-has to consider: by increasing it detection rate it also has to increase
-the error rate.
+false-positive rates never-decrease~\footnote{Note that both are described as
+summations of products of probabilities (non-negative functions).} describing an
+\emph{error and rejection tradeoff}~\cite{chow1970optimum} that any detection
+system has to consider.
 
@@
 rate possible. This is equivalent to the \emph{continuous knapsack problem}
-which allows a greedy solution by sorting the items by value per weight.
+which allows a greedy solution by sorting the items with a value per weight
+function.
 Therefore a novelty system before classifying a sample $a$ as novel should
@@
 $P(\overline{novel}|x)$ any constant factor can be dropped.
-Leaving a ratio, between a \emph{conditional} and
-\emph{unconditional probability}, suitable for implementing novelty detection
+Leaving a ratio between a \emph{conditional} and
+\emph{unconditional probability} suitable for implementing novelty detection
 via thresholding.
@@
 distribution, otherwise it will contain some bias and lead to an incorrect
-ordering. The bias, for example, can be introduced by an incorrect filtering or
-data balancing.
+ordering. The bias, for example, can be introduced by an incorrect filtering.
 
@@
 distribution must be chosen.
-Though, as discussed in \autoref{sec:unlabelled-data}, by using unlabelled data
-it becomes possible to obtain prior-information and achieve a better
-approximation.
+Though by using unlabelled data it becomes possible to obtain extra information
+and achieve a better approximation.
 
@@
 
+\clearpage
 \section{A Practical Example}
@@
 features $X$ that are directly sensed by the agent.
-All the sensed features $x$ are all independent given the room category.
+All the sensed features $x$ are independent given the room category.
 In whole there was 11 different room categories and 7 different feature types.
-Each feature can be sensed more than once (i.e. room shape is extracted from 2D
+Each feature can be sensed more than once (e.g.\ room shape is extracted from 2D
 laser scans in more than one position in the room), but all those sensed
@@
 The room categories were chosen to mimic as close as possible the real features
-and categories existing in reality (i.e. 1 person office, 2 person office,
+and categories existing in reality (i.e.\ 1 person office, 2 person office,
 hallway, robot lab, etc\dots). A table describing the used synthetic
@@
 
-The objective then was to draw a system that, although only trained with
+The objective then was to design a system that, although only trained with
 labelled data from 5 of the 11 room categories, was able to detect novel
@@
 
+% TODO: discuss this with Andrzej
 \begin{equation}
-\phi_X(r,x) = number of samples + C
+\phi_X(r,x) = \frac{\#(r,x)+C}{\sum_{i \in X}{\#(r,i)+C}}
 \end{equation}
 
-With those the probability of sensing a set of features $x$ on a room knowing
-room category is known to the agent can be modelled with a factor graph as
-illustrated on \autoref{fig:simple-cond-graph}.
+Where $\#(r,x)$ denotes the number of times a feature $x$ was sensed inside a
+room category $r$ and $C$ is a smoothing parameter that accounts for fixing
+the probabilities in case a given sample is never seen.
+With those the probability of sensing a set of features $x$ on a room category
+$r$ known for the agent can be modelled with a factor graph as illustrated on
+\autoref{fig:simple-cond-graph}.
 
@@
 \caption{\label{fig:simple-cond-graph}A factor graph modelling the conditional
-probability of sensing a set of features $x$ known that the room category $r$ is
+probability of sensing a set of features $x$ given that the room category $r$ is
 one of the known classes.}
@@
 \subsection{Unconditional Probability}
+\label{sec:sample-uncond-prob}
 With no knowledge on the unconditional probability the correct approach is to
@@
 In this practical example the access to unlabelled data is explored as access to
-it is very common on applications in robotics.
+it is common on applications in robotics.
 
@@
 the \emph{curse of dimensionality}: as the number of sensed features and feature
-types increases exponential ammounts of data are needed to describe it.
+types increases exponential amounts of data are needed to describe it.
 Having interest in avoiding the issues associated with an high-connected factor
@@
 
-Note than in the case there is only one sensed feature, the distribution
-probability modelled by an assumption of independent features correctly models
-unconditional probability. And it deviates from it as more features are sensed.
-Though the individual factors associated with each variable try to unwrapped any
-possible existent bias on each dimension and therefore it is expected to be a
+In the case there is only one sensed feature the distribution generated by an
+assumption of independent features correctly models the unconditional
+probability. And it deviates from it as more features are sensed.
+The individual factors associated with each variable can be seen as a scaling of
+each feature that tries to account for a possible existent bias on each
+of them. Therefore, although far from the desired factor, it is expected to be a
 better estimate than assuming a uniform distribution.
@@
 implementing a threshold when the samples are taken from dynamic
-distributions (e.g.: some samples where there is only access to room size versus
+distributions (e.g.\ some samples where there is only access to room size versus
 samples where there is a lot of information about the room properties).
@@
 able to model the existent dependence between the features. This becomes obvious
-as the number of features increases (e.g.: graph $I$ perfectly models $P(x)$ in the
+as the number of features increases (e.g.\ graph $I$ perfectly models $P(x)$ in the
 case where only 1 feature is sensed).

### 2011-06-07 thesis/novelty-intro.tex
Added related work with novelty detection. PLISS, bishop and boutell.

@@
 \subsection{Related Work}
-More in line with novelty detection in scene categorization and semantic
-structures, the work of \gls{PLISS}~\cite{ranganathan2010pliss}.\dots
+\cite{ranganathan2010pliss} presents a system able to perform place recognition
+and classification from visual clues. It is able to perform segmentation by
+exploiting time-coherency on video information. The approach is particularly
+interesting by its ability of keeping a fully probabilistic distribution of
+the place classification and segmentation. This is, the system never performs
+a deterministic decision that impacts any future , allowing it to adjust all
+the segmentations and place classification as more data becomes available.
+The system is also able to detect novel instances and methods how to adapt it
+to run on a constant amount of memory and computation are presented.
+
+\cite{boutell2006factor} presents a method to perform scene-classification
+from low-level region detectors using factor graphs. Those low-level detectors
+provide semantic (high-level) cues on the higher level scene-classification
+task. Due to the nature of the low-level detectors, they ought to be treated
+in a probabilistic way. For that their method uses factor graphs. The method
+also exploits spatial relation information between regions to improve the
+classification performance. A presented scheme using only pairwise relations
+between regions is show to have better performance than try to approach the
+correct scheme with a big single-factor. I.e.\ the pairwise relations can be
+approximated with the modest amounts of training data, where the single-factor
+demands an impractical amount of training data.
+No approach is made on it at novelty detection either of regions or scene
+categories.
+
+\cite{bishop1994novelty} works shows that assuming a uniform unconditional
+distribution novelty detection can be performed by triggering on the
+density probability of the training data. The system is employed to detect novel
+inputs for a neural-network.
 

### 2011-06-07 thesis/novelty.tex
Changed TikZ styles to draw factor graphs.
Several changes on novelty detection on the conceptual map. Still WIP.

@@
 
-Assume now that everything the conceptual layer knows holds true except the
-class of a single variable $a$, which the system possess no information on. 
-In that case, the correct approach is to replace the factors associated with
-$a$ with factors representing the prior-information the system has on
-the unconditional distribution of those.
-Let this new graph with unconditional factors on the variable
-$a$ be $U_a$. $P(x)$ is then approximated by $P_{U_a}(x)$.
-
-Note that in case no prior-information is available on the unconditional
-factors this means replacing them by a uniform factor\footnote{In case the
-unconditional factors are uniform, this is the same as omitting the factors
-from the graph structure.}. In case some information is known
-(i.e. unlabelled data) the correct approach is to incorporate it into the
-unconditional factors using the principle of maximum entropy as seen on
-\autoref{sec:unlabelled-data}.
-
 \begin{figure}[h]
 \centering
-\subfloat[Factor graph modelling the probability distribution in the case both
-          the variables types and graph structure hold true.]{
 \begin{tikzpicture}
-\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
-\&
-\&
-\node (room3) [latent] {$R_3$}; \& \\
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {$c$}; \& \\
     \& \& \& \\
     \&
-\node (room1) [latent] {$R_1$}; \& \&
-\node (room2) [latent] {$R_2$}; \\
+    \node (room1) [latent] {$a$}; \& \&
+    \node (room2) [latent] {$b$}; \\
     \& \& \& \\
@@
 
+  \begin{pgfonlayer}{background}
     \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
     \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
 
-\end{tikzpicture} %
-          }
-\qquad\qquad
-\subfloat[Factor graph modelling the probability distribution in the case
-          where only variable $a$ does not conforms to any class the system
-          has learned before.]{
+\caption{Factor graph modelling the probability distribution in the case both
+          the variables types and graph structure hold true.}
+\end{figure}
+
+Assume now that everything the conceptual layer knows holds true except the
+class of a single variable $a$, which the system has no information on. 
+In that case, the correct approach is to replace the factors associated with
+$a$ with a single factor connecting all the variables dependent on it as seen
+on Figure~\ref{fig:novel-a-correct}.
+
+Such an highly connected factor is hard to model. Though assuming the structure
+holds true, it is possible to approximate that highly connected factor with a
+variable $a^*$ with $K$ states.\footnote{Ideally those $K$ states would be all
+the classes of $a$.}
+Assuming no information is known on the unconditional distribution, the factors
+dependent on $a^*$ should be modelled with uniform factors and $K$ can be set to
+$1$.
+
+Let $U_a$ be this new graph based on $G$ with variable $a$ replaced by $a^*$.
+$P_{U_a}(x)$ models then the unconditional distribution of the feature set $x$
+without the assumption that $a$ is one of the known classes and so is used as
+an approximation for $P(x)$.
+
+\begin{figure}[h]
+\centering
+\subfloat[\label{fig:novel-a-correct}Graphical model able to correctly represent
+          the unconditional probability when variable $a$ can be novel.]{
 \begin{tikzpicture}
-\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {$c$}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [factor] {}; \& \&
+    \node (room2) [latent] {$b$}; \\
+    \& \& \& \\
     \&
     \&
-\node (room3) [latent] {$R_3$}; \& \\
+    \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+
+  \draw [-] (room1) -- (shape1l);
+  \draw [-] (room1) -- (appearance1l);
+  \draw [-] (room1) -- (object1l);
+  \draw [-] (room2) -- (prop2);
+
+  \draw [-] (room1) -- (room2);
+  \draw [-] (room1) -- (room3);
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+}
+\qquad
+\subfloat[By using independent features for modelling the unconditional
+          distribution an highly simplified graph is obtained.]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \&
+    \&
+    \node (room3) [latent] {$c$}; \& \\
     \& \& \& \\
     \&
-\node (room1) [latent] {$R^N_1$}; \& \&
-\node (room2) [latent] {$R_2$}; \\
+    \node (room1) [latent,dashed] {$a^*$}; \& \&
+    \node (room2) [latent] {$b$}; \\
     \& \& \& \\
@@
   };
-
-\draw [-] (room1) -- (shape1f) -- (shape1l);
-\draw [-] (room1) -- (appearance1f) -- (appearance1l);
-\draw [-] (room1) -- (object1f) -- (object1l);
+  \draw [-] (shape1f) -- (shape1l);
+  \draw [-] (appearance1f) -- (appearance1l);
+  \draw [-] (object1f) -- (object1l);
   \draw [-] (room2) -- (factor2) -- (prop2);
-
-\draw [-] (room1) -- (room2) node (r12f) [midway,factor] {};
-\draw [-] (room1) -- (room3) node (r13f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (room2) node (r12f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (room3) node (r13f) [midway,factor] {};
   \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+  \draw [-] (room2) -- (r12f);
+  \draw [-] (room3) -- (r13f);
   \node (captr12f) [above=0pt of r12f] {\footnotesize{$U$}};
@@
 
+  \draw [-,dashed] (room1) -- (shape1f);
+  \draw [-,dashed] (room1) -- (appearance1f);
+  \draw [-,dashed] (room1) -- (object1f);
+  \draw [-,dashed] (room1) -- (r12f);
+  \draw [-,dashed] (room1) -- (r13f);
+
+  \begin{pgfonlayer}{background}
     \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
     \plate{places2}{(factor2)(prop2)}{$\dots$}{};
-
-\end{tikzpicture} %
+  \end{pgfonlayer}
+\end{tikzpicture}
 }
-
-\caption{\label{fig:nd-single-variable}Graphs used to model novelty detection on
-         a class of a single variable $R_1$, assuming all the other variables
-         classes and graph structure are correctly modelled by the agent
-         knowledge.}
+\caption{Factor graph modelling the probability distribution in the case
+         where only variable $a$ is not necessarily a class the system has
+         learned before.}
 \end{figure}
 
-Having now defined $P_G(x)$ and $P_{U_a}$ as models models for
-$P(x|\overline{novel})$ and $P(x)$ (\autoref{fig:nd-single-variable}),
-$P_G(x)/P_{U_a}(x)$ is used to define an order-relation on the novelty of $a$.
+Having now defined $P_G$ and $P_{U_a}$ as models models for
+$P(x|\overline{novel})$ and $P(x)$, the ratio $P_G(x)/P_{U_a}(x)$ is used to
+define an order-relation for the novelty of $a$.
 Now all that is missing is a threshold $t_A$ to trigger on.
@@
 Also, since there is no reason to distinguish between different instances of
-a variable type, only one threshold by variable type needs to be trained.
+a variable type, only one threshold per variable type needs to be trained.
 
@@
 where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
-by a distribution where only $a$ is novel.
+by a distribution where only $a$ is not necessarily a class the system has
+learned before.
+
+\subsection{Novelty Detection as a MAP Operation}
+Having interest in integrating the presented method to detect novelty of a
+variable $a$ on a single graphical model some more steps can be taken.
+
+Let the known states of $a$ be $\{A_1,\dotsc, A_{n}\}$ and the unconditional variable
+$a^*$ be modelled with states $\{A^*_1,\dotsc, A^*_{k}\}$. I.e. if $k=1$ then the
+unconditional variable is being modelled with an assumption on features that do
+not dependent on the class.
+
+First, extend variable $a$ with new states $A_{n+k}$
+
+By incorporating novelty information back into the graphical model the presented
+method allows
 

### 2011-06-07 thesis/novelty-intro.tex
Related work on novelty detection.

@@
 therefore should proceed with caution.
-It has several applications such as fault detection, intrusion detection,
-detection of masses in mammograms, hand written digit recognition and many
-others.
-% TODO: list usage of novelty detection together with references to papers
-% with them.
+It has several applications such as fault detection~\cite{tarassenko1999novelty},
+intrusion detection~\cite{fan2001using},
+detection of masses in mammograms~\cite{tarassenko1995novelty}.
+Any normal classification application is also a good candidate for extending
+with novelty detection, as it allows to detect samples the system has not been
+trained on and therefore the classification is considered
+instable~\cite{devarakota2008reliability}.
+E.g.\ applied to digit-recognition~\cite{tax1998outlier}
+or to detection of novel inputs on a neural network~\cite{bishop1994novelty}.
 
@@
 \subsection{Related Work}
-\cite{ranganathan2010pliss} presents a system able to perform place recognition
-and classification from visual clues. It is able to perform segmentation by
-exploiting time-coherency on video information. The approach is particularly
-interesting by its ability of keeping a fully probabilistic distribution of
-the place classification and segmentation. This is, the system never performs
-a deterministic decision that impacts any future , allowing it to adjust all
-the segmentations and place classification as more data becomes available.
-The system is also able to detect novel instances and methods how to adapt it
-to run on a constant amount of memory and computation are presented.
+\cite{zhang2005probabilistic} present a probabilistic semantic approach to image
+annotation and retrieval with a multi-modal information. They introduce an
+hidden variable modelling a concept present on a scene that connects together
+information coming from both a bag of words and visual features extracted. By
+training that hidden-variable their system becomes aware of possible existent
+semantic categories and their relation between both textual and visual features.
+Although not able to perform novelty detection, their approach is interesting by
+the use of an hidden variable to probabilistic model the multi-modal sensors.
 
@@
 inputs for a neural-network.
+
+\cite{ranganathan2010pliss} presents a system able to perform place recognition
+and classification from visual clues. It is able to perform segmentation by
+exploiting time-coherency on video information. The approach is particularly
+interesting by its ability of keeping a fully probabilistic distribution of
+the place classification and segmentation. This is, the system never performs
+a deterministic decision that impacts any future result, allowing it to adjust
+the segmentation and place classification as more data becomes available.
+The system is also able to detect novel instances and methods how to adapt it
+to run on a constant amount of memory and computation are presented.
 

### 2011-06-07 thesis/novelty-intro.tex
added reference to novelty detection on documents.

@@
 intrusion detection~\cite{fan2001using},
-detection of masses in mammograms~\cite{tarassenko1995novelty}.
+detection of masses in mammograms~\cite{tarassenko1995novelty} or detection of
+novel and useful documents~\cite{zhang2002novelty}.
+
 Any normal classification application is also a good candidate for extending
@@
 
+% TODO: wuld be nice to have this one starting on a new page.. but still make
+% sure it looks good and not a forced break.
+% Reason to have this on a diferent page is to make it really clear novelty
+% content (and not only chit-chat) starts now.
+\clearpage
 \section{Novelty Detection as a Threshold}
@@
 
-\clearpage
 \section{A Practical Example}

### 2011-06-08 thesis/novelty.tex
Add how to use a MAP to perform novelty detection between 2 graphs.
This starts introducing the relation between graph-structure and novel probability.

@@
 conceptual map. As seen before on \autoref{chap:novelty-intro},
-novelty detection can be done by implementing a threshold on
-$P(x|\overline{novel})/P(x)$.
+assuming that $P(\overline{novel})$ is constant, novelty detection can be
+done by implementing a threshold on $P(x|\overline{novel})/P(x)$.
 In order to correctly measure novelty on a single variable both
@@
 $P_{U_a}(x)$ models then the unconditional distribution of the feature set $x$
-without the assumption that $a$ is one of the known classes and so is used as
-an approximation for $P(x)$.
+without the assumption that $a$ is one of the known classes and can be is used
+as an approximation for $P(x)$.
 
@@
 define an order-relation for the novelty of $a$.
-Now all that is missing is a threshold $t_A$ to trigger on.
+All that is missing is a threshold $t_A$ to trigger on.
 Such a threshold can be obtained by calculating which value of $t_A$
@@
 
-Concluding: novelty on a variable $a$ of type $A$ should be detected when:
+Concluding: assuming a constant $P(\overline{novel})$ for any specific set and
+structure of $x$, novelty on a variable $a$ of type $A$ should be detected when:
 
 \begin{equation}
+\label{eq:threshold-with-ta}
 \frac{P_G(x)}{P_{U_a}(x)} < t_A
@@
 where $a$ is known, and $P_{Ua}$ denotes the probability of $x$ being generated
-by a distribution where only $a$ is not necessarily a class the system has
+by a distribution where $a$ is not necessarily a class the system has
 learned before.
@@
 
-Let the known states of $a$ be $\{A_1,\dotsc, A_{n}\}$ and the unconditional variable
-$a^*$ be modelled with states $\{A^*_1,\dotsc, A^*_{k}\}$. I.e. if $k=1$ then the
-unconditional variable is being modelled with an assumption on features that do
-not dependent on the class.
+The method presented before obtains an unconditional model
+from the conditional model by replacing a variable $a \in \{A_1,\dotsc, A_n\}$
+and the associated factors $\phi_a$ with a variable
+$a^* \in \{A^*_1,\dots,A^*_k\}$ and related unconditional factors $\phi_{a^*}$
+and leaving all the remaining variables and factors equal.
 
-First, extend variable $a$ with new states $A_{n+k}$
+With this it becomes possible to merge both graphs into a single one $G$ by
+merging $a$ and $a^*$ into $b$ and factors $\phi_a$ and $\phi_{a^*}$ into
+$\phi_b$ as follows:
 
-By incorporating novelty information back into the graphical model the presented
-method allows
+\[
+b \in \{A_1,\dotsc,A_n,A^*_1,\dotsc,A^*_k\}
+\]
 
-\section{Copping with Multiple Novel Variables}
+\[
+\phi_b(b, \cdot) = \left\{
+  \begin{array}{l l}
+    \phi_a(b, \cdot), & \quad \text{if } b \in \{A_1,\dotsc,A_n\} \\
+    \phi_{a^*}(b, \cdot), & \quad \text{if } b \in \{A^*_1,\dotsc,A^*_k\}
+  \end{array} \right.
+\]
+
+Additionally it is possible to control variable $b$ states into being only one
+of the states of $a$ or $a^*$ by introducing a control variable $c \in {0, 1}$
+and factor $\phi_{c}$:
+\[
+\phi_{c}(b, c) = \left\{
+  \begin{array}{l l}
+    1, & \qquad \text{if } c = 0 \text{ and } b \in \{A_1,\dotsc,A_n\} \\
+    1, & \qquad \text{if } c = 1 \text{ and } b \in \{A^*_1,\dotsc,A^*_n\}\\
+    0, & \qquad \text{otherwise}
+  \end{array} \right.
+\]
+
+By conditioning the control variable $c$ it is possible to obtain a graph with
+potentials on any of the remaining variables equivalent to the potentials
+obtained with $G_A$ (i.e.\ $c=0$) or $G_{U_a}$ (i.e.\ $c=1$). By introducing
+an extra threshold factor $\phi_t(c)$ it also becomes possible to balance the
+potential of the distribution described by $B$ such that a \gls{MAP} operation
+on the $c$ variable decides on the novelty of $a$ with a desired threshold:
+
+\begin{eqnarray*}
+ P_B(c=0 | x)         &<& P_B(c=1 | x) \\
+\Phi_B(c=0 \land x)   &<& \Phi_B(x=1 \land x) \\
+\Phi_G(x) \phi_t(c=0) &<& \Phi_{U_a}(x) \phi_t(c=1) \\
+P_G(x) Z_G \phi_t(c=0) &<& P_{U_a}(x) Z_{U_a} \phi_t(c=1) \\
+\frac{P_G(x)}{P_{U_a}(x)} &<& \frac{\phi_t(c=1)}{\phi_t(c=0)}\frac{Z_{U_a}}{Z_G}
+\end{eqnarray*}
+
+Concluding, after merging graph $G$ and $U_a$ into graph $B$, the condition
+described in \autoref{eq:threshold-with-ta} can be test by performing a
+\gls{MAP} operation on variable $c$ of graph $B$ when the threshold factor
+$\phi_t$ is proportional to:
+\begin{equation}
+\phi_t(c) \propto \left\{
+  \begin{array}{l l}
+    1, & \qquad \text{if } c = 0 \\
+    t_A \frac{Z_G}{Z_{U_a}}, & \qquad \text{if } c = 1\\
+  \end{array} \right.
+\end{equation}
+
+This equation is not very practical due to the need of calculation of
+normalization factors $Z_G$ and $Z_{U_a}$. The presence of this factors
+not only turns computation more expensive, but also requires that a full model
+is known for all the factors disabling the ability to incorporate probabilistic
+sensing output from complex detectors such as \glspl{SVM}. Ultimately, it also
+removes capability to mix complex thresholds on several variables that would
+allow to do probabilistic inference on the novelty of multiple variables at once.
+
+\begin{figure}[h]
+\centering
+lalla
+\caption{\label{fig:merging-graphs}A condition based on the ratio between 2
+         factor graphs that share some structure can be performed by testing the
+         MAP probability of a variable on a merged graph.}
+\end{figure}
+
+Analysing the reason for the normalization factors playing a role on the
+threshold can be tracked down to the assumption that $P(novel)$ is
+considered constant over any set or structure of $x$.
+
+This requirement forces the threshold function to use the normalization factors
+to compensate for a bigger likelihood of certain structure being naturally more
+disposed to generate novel categories.
+
+This relation between graph structure and $P(novel)$ is discussed in more detail
+in the next section where also the novelty detection method is made to work
+without normalization factors, allowing probabilistic sensing, and extended to
+detect novelty on multiple variables.
+
+\section{Detecting and Copping with Novelty on Multiple Variable}
 On the previous section it was shown how to detect a novel class on a variable

### 2011-06-08 thesis/novelty.tex
Added coments on how does the sampling method to have a constant P(novel) looks like.

@@
 
+\subsection{Graph Structure Influences $P(novel)$}
 Analysing the reason for the normalization factors playing a role on the
@@
 considered constant over any set or structure of $x$.
-
 This requirement forces the threshold function to use the normalization factors
 to compensate for a bigger likelihood of certain structure being naturally more
-disposed to generate novel categories.
+disposed to generate novel categories (an example of such structure is given on
+\autoref{fig:structure-influences-novelty}).
 
-This relation between graph structure and $P(novel)$ is discussed in more detail
-in the next section where also the novelty detection method is made to work
-without normalization factors, allowing probabilistic sensing, and extended to
-detect novelty on multiple variables.
+A method to generate samples that match this \emph{criterium} of
+having a constant $P(novel)$ can be:
+
+\begin{algorithm}[Draw samples with equal $P(novel)$]
+\begin{enumerate}
+\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
+\item Decide novelty of $a$.
+\item Sample the remaining variables according to distribution $P_G(x|a)$.
+\end{enumerate}
+\end{algorithm}
+
+According to this sampling algorithm, given any structure the probability of
+variable $a$ being novel is constant. This is somehow unexpected as it means
+the sampling method is biased to adjust the real distribution $G$ such that
+exactly a constant percentage of drawn samples are novel.
+
+Alternative, the sampling method described below is considered more correct as
+it introduces no bias on the distribution $G$ describing the real distribution
+and classes of $a$:
+\begin{algorithm}[Draw samples with equal $P(novel)$]
+\begin{enumerate}
+\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
+\item Sample variables $x$ according to distribution $P_G(x)$.
+\end{enumerate}
+\end{algorithm}
+
+The next section shows how to generate a novelty detection method based on this
+unbiased sampling method where samples are drawn directly from an assumed real
+distribution aware of all the classes of $a$.
+
+\begin{figure}[h]
+\caption{\label{fig:structure-influences-novelty}Example showing how
+         structure can influence $P(novel)$}
+\end{figure}
 
 \section{Detecting and Copping with Novelty on Multiple Variable}
+
 On the previous section it was shown how to detect a novel class on a variable

### 2011-06-09 thesis/macros.tex
Added coments on how does the sampling method to have a constant P(novel) looks like.
Added more discussion on how P(novel) is dependent on structure.

@@
 \def\chapterautorefname{Chapter}
+
+\newtheorem{algorithm}{Algorithm}
 
@@
 factor/.style={rectangle, fill=red,minimum size=7pt, inner sep=0pt},
+yellow/.style={latent,minimum size=15pt,fill=yellow!75},
+blue/.style={latent,minimum size=15pt,fill=blue!75},
 -/.style={color=red, thick},

### 2011-06-09 thesis/novelty.tex
Added more discussion on how P(novel) is dependent on structure.

@@
 \begin{figure}[h]
-\caption{\label{fig:structure-influences-novelty}Example showing how
-         structure can influence $P(novel)$}
+
+\centering
+
+\subfloat[Training Data]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
+    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
+    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
+  };
+  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (f);
+\end{tikzpicture}
+\quad
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
+    \node (empty) [blue,fill=none,draw=none] {}; \& \\
+    \node (c) [blue] {}; \& \node (d) [yellow] {}; \\
+  };
+  \draw [-] (c) -- (a) -- (d) -- (e);
+\end{tikzpicture}
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
+    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
+    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
+  };
+  \draw [-] (a) -- (e) -- (f) -- (d) -- (c) -- (b) -- (a);
+\end{tikzpicture}
+\quad
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [blue] {}; \& \node (e) [yellow] {}; \\
+    \node (b) [yellow] {}; \& \node (f) [blue] {}; \\
+    \node (c) [blue] {}; \& \node (d) [blue] {}; \\
+  };
+  \draw [-] (c) -- (b) -- (a) -- (e) -- (f) -- (b);
+  \draw [-] (b) -- (d);
+\end{tikzpicture}
+\quad
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [yellow] {}; \& \node (e) [blue] {}; \\
+    \node (b) [blue] {}; \& \node (f) [yellow] {}; \\
+    \node (c) [yellow] {}; \& \node (d) [blue] {}; \\
+  };
+  \draw [-] (e) -- (a) -- (b) -- (c) -- (d);
+  \draw [-] (b) -- (f);
+\end{tikzpicture}
+
+}
+
+\subfloat[Sample A\label{fig:sample-a}]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [latent] {$a$}; \& \node (b) [latent] {}; \\
+    \node (c) [latent] {};    \& \node (d) [latent] {}; \\
+  };
+  \draw [-] (a) -- (b) -- (d) -- (c) -- (a);
+\end{tikzpicture}
+}
+\qquad
+\subfloat[Sample B\label{fig:sample-b}]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (a) [latent] {$a$}; \& \node (b) [latent] {}; \\
+    \node (c) [latent] {};    \& \\
+  };
+  \draw [-] (a) -- (b) -- (c) -- (a);
+\end{tikzpicture}
+}
+
+\caption{\label{fig:structure-influences-novelty}Given that the only known
+         classes are yellow and blue, and that two nodes of the same class
+         were never found connected together, it is reasonable to assume
+         that on sample B the probability of node $a$
+         be a novel class is greater than on sample A.}
 \end{figure}

### 2011-06-10 thesis/novelty.tex
Lots of content and changes on detecting novelty detection without assumption of P(novel) is constant on any structure.

@@
 
-\section{Detecting Novelty on a Single Variable}
+\section{Detecting Novelty with Ratio Between Two Graphs}
 This section focus on detecting a novel class on single variable of the
@@
   \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
-    \& \& \node (room3) [latent] {$c$}; \& \\
+    \& \& \node (room3) [latent] {}; \& \\
     \& \& \& \\
@@
     \node (room1) [latent] {$a$}; \& \&
-    \node (room2) [latent] {$b$}; \\
+    \node (room2) [latent] {}; \\
     \& \& \& \\
@@
     \node (room1) [factor] {}; \& \&
-    \node (room2) [latent] {$b$}; \\
+    \node (room2) [latent] {}; \\
     \& \& \& \\
@@
     \&
-    \node (room3) [latent] {$c$}; \& \\
+    \node (room3) [latent] {}; \& \\
     \& \& \& \\
@@
     \node (room1) [latent,dashed] {$a^*$}; \& \&
-    \node (room2) [latent] {$b$}; \\
+    \node (room2) [latent] {}; \\
     \& \& \& \\
@@
 \begin{equation}
+\label{eq:structure-ratio}
 \phi_t(c) \propto \left\{
@@
 \centering
-lalla
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \node (thresholdguide) [latent,draw=none] {}; \&
+    \&
+    \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \node (control) [latent,dashed] {$c$}; \&
+    \node (room1) [latent,dashed] {$b$}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \nofactor {shape1f}{}{$\phi_b$} {right=0pt}; \&
+    \nofactor {appearance1f}{} {$\phi_b$} {right=0pt}; \&
+    \nofactor {object1f}{} {$\phi_b$} {right=0pt}; \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+  \draw [-] (shape1f) -- (shape1l);
+  \draw [-] (appearance1f) -- (appearance1l);
+  \draw [-] (object1f) -- (object1l);
+  \draw [-] (room2) -- (factor2) -- (prop2);
+  \draw [-,draw=none] (room1) -- (room2) node (r12f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (room3) node (r13f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (control) node (controlf) [midway,factor] {};
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+  \draw [-] (room2) -- (r12f);
+  \draw [-] (room3) -- (r13f);
+  \node (captr12f) [above=0pt of r12f] {\footnotesize{$\phi_b$}};
+  \node (captr13f) [above=0pt of r13f] {\footnotesize{$\phi_b$}};
+
+  \draw [-,dashed] (room1) -- (shape1f);
+  \draw [-,dashed] (room1) -- (appearance1f);
+  \draw [-,dashed] (room1) -- (object1f);
+  \draw [-,dashed] (room1) -- (r12f);
+  \draw [-,dashed] (room1) -- (r13f);
+
+  \draw [-,draw=none] (room1) -- (control) node (controlf) [midway,factor] {};
+  \draw [-,draw=none] (control) -- (thresholdguide) node (thresholdf) [midway,factor] {};
+  \node (captcontrolf) [above=0pt of controlf] {\footnotesize{$\phi_c$}};
+  \node (captthresholdf) [above=0pt of thresholdf] {\footnotesize{$\phi_t$}};
+  \draw [-,dashed] (room1) -- (controlf) -- (control) -- (thresholdf);
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+
 \caption{\label{fig:merging-graphs}A condition based on the ratio between 2
@@
 
-\section{Detecting and Copping with Novelty on Multiple Variable}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Detecting Novel Classes on a Graph Variable}
+On the previous section it was shown how to detect novelty on a variable
+by assuming $P(novel)$ is constant across any structure. It was shown that
+considering that \emph{criterium} a need to compensate the likelihood of
+certain graph structures for generating novel classes arises and it requires
+calculating normalization factors between the structures.
+That \emph{criterium} is not only a nuisance but also represents
+a different problem. It can be seen as an attempt to recognize that a sample
+was drawn from a novel method instead of detecting whether a given variable on
+a graph is a class that the system is not aware.
+
+This section focus on the last one. It assumes there is a distribution modelled
+by a graph $G$ that is used to drawn samples. On that graph $G$ there is a
+variable $a \in \{A_1,\dotsc, A_K, A^*_{K+1},\dotsc, A^*_N\}$. The system is considered
+to known only the states $\{A_1,\dotsc, A_K\}$. The goal is then: given a sample subset
+$x$ of variables drawn from the graph, detect whether variable
+$a \in \{A_1,\dotsc, A_K\}$.
+
+That decision can be optimally modelled by the following condition:
+
+\begin{eqnarray*}
+P_G(a \in \{A_1,\dotsc, A_K\}|x) &<& P_G(a \in \{A^*_{K+1},\dotsc,A^*_N\}|x) \\
+P_G(a \in \{A_1,\dotsc, A_K\} \land x) &<& P_G(a \in \{A^*_{K+1},\dotsc,A^*_N\} \land x) \\
+2 P_G(a \in \{A_1,\dotsc, A_K\} \land x) &<& P_G(x)
+\end{eqnarray*}
+
+Similarly to the previous approach a graph $G'$ can be created based on $G$ such
+that conditioning on a variable $c$ allows to obtain either the conditional or
+unconditional probability of a given set of variables $x$.
+For that variable $a$ is extended into $a'$ by duplicating the known states $A_i$
+into $A^*_i$ and extending the factors associated with it.
+Additionally a control factor $\phi_c(c, a)$ is added to make $c$ enforce the
+desired constraints:
+
+\begin{eqnarray*}
+a' &\in& \{A_1, \dotsc, A_K,\quad A^*_1, \dots, A^*_K,\quad A^*_{K+1}, \dots, A^*_N\} \\
+c &\in& \{0, 1\}
+\end{eqnarray*}
+\begin{eqnarray*}
+\phi_c(c, a') &=& \left\{
+  \begin{array}{l l}
+    1, & \qquad \text{if } c = 0 \text{ and } a' \in \{A_1,\dotsc,A_K\} \\
+    1, & \qquad \text{if } c = 1 \text{ and } a' \in \{A^*_1,\dotsc,A^*_N\}\\
+    0, & \qquad \text{otherwise}
+  \end{array} \right. \\
+\end{eqnarray*}
+
+Note that with this transformation, the following potential equivalences hold:
+\begin{eqnarray*}
+\Phi_{G'}(x, c=0) &=& \Phi_{G}(x | a \in \{A_1, \dots, A_K\})\\
+\Phi_{G'}(x, c=1) &=& \Phi_{G}(x) 
+\end{eqnarray*}
+
+Although those potentials allow to determine novelty of $a$, they require
+knowledge of all the possible classes, which is not available.
+Nonetheless access to unlabelled data or other informations or assumptions on
+the unconditional distribution allow to model or approximate unconditional
+distributions of $a$. For that there is interest in trying to model the set
+of states $\{A^*_1, \dots, A^*_N\}$ with a set of states $\{A^{\dag}_1, \dotsc, A^{\dag}_M\}$.
+
+Moreover assuming that the set of states $A^\dag$ is able to perfectly
+characterize the distribution over any graph $G$,
+for a specific graph the potentials with states $A^\dag$ are a scale-factor
+away from the potentials with states $A$. And that scale-factor is dependent on
+the graph structure and factors of both $G$ and $G^\dag$:
+
+\begin{eqnarray}
+P_G(x) &=& P^\dag_G(x) \\
+\Phi_G(x) &=& \frac{Z_G}{Z^\dag_G} \Phi^\dag_G(x)
+\end{eqnarray}
+
+Note that this ratio, although seems similar to the one obtained in
+\autoref{eq:structure-ratio} it has a special property. Here both $G$ and
+$G^\dag$ model the same probability, so it is expected that both scale they
+normalization factor according to some relation on how a given factor differs
+from $G$ to $G^\dag$.\footnote{On \autoref{eq:structure-ratio} the ratio is
+between two different distributions: a conditional and an unconditional distribution.}
+
+If on any graph structure the marginal probabilities of all variables except $a$
+is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$)
+then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
+It can be proved that since the distribution over any graph structure has to be
+equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
+factor $\phi_i$:
+
+\begin{eqnarray*}
+P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
+\frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
+s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}}
+\end{eqnarray*}
+
+With this, the scale-factor that distinguish the potential on a graph
+$G$ from the potential on $G^\dag$ is given by:
+
+\begin{eqnarray}
+s &=& \frac{\Phi_G(x)}{\Phi^\dag_G(x)} \\
+  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
+           {\sum_{a \in A^\dag}\prod \phi^\dag(a^\dag, x_\phi)} \\
+  &=& \frac{\sum_{a \in A}\prod \phi(a, x_\phi)}
+           {\sum_{a \in A}\prod \phi(a, x_\phi) s_i} \\
+  &=& \frac{1}{\prod s_i}
+\end{eqnarray}
+
+By adding a condition variable $c \in \{0, 1\}$, a condition factorthat is added result can be converted into a graph $G$is particularly interesting as it allows to detect novelty of $a$ by
+having models to approach
+
+Similar to the procedure followed before, a decision variable $c \in \{0, 1\}$
+and a factor $\phi_c(c,a)$ will be added to $G$, creating graph $G'$.
+If $\phi_c$ is defined as:
+
+This is variable $c$ detects whether $a$ is a known or unknown class.
+Note also that by adding this condition variable and factor any marginal
+potential on $G$ is the same as the potential on $G'$. Since marginal potentials
+can be seen as sum of potentials on all variables it can be proved by showing
+the potential on any instance of all variables $X$ in $G$ is the same as the
+marginal potential of the same variables in $G'$:
+
+\begin{eqnarray*}
+\Phi_{G'}(x) &=& \sum_{c}\prod_{\phi \in G'}{\phi(x_\phi)} \\
+             &=& \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(0,a) 
+               + \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(1,a)\\
+             &=& \Phi_G(x)\phi_c(0,a) + \Phi_G(x)\phi_c(1,a)\\
+             &=& \Phi_G(x)\\
+\end{eqnarray*}
+
+An optimal decision on whether variable $a$ of $G$ is $\in \{A_1,\dotsc,A_K\}$
+can be performed by a map \gls{MAP} operation on it. And that decision can be
+done solely based on potentials in $G'$:
+
+\begin{eqnarray*}
+P_G(a \in \{A_1,\dotsc,A_K\}|x) &<& P_G(a \in \{A_{K+1},\dotsc,A_N\}|x) \\
+\Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c = 1 \land x) \\
+2 \Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c=0 \land x) + \phi_{G'}(c = 1 \land x)\\
+\end{eqnarray*}
+
+By extending $a$ with duplicated states $\{A_1, \dotsc, A_K\}$, extending
+the respective factors and adding a factor $\phi_f$ on the decision
+variable $c$, it is possible to create a graph $G^{''}$ where a MAP decision on
+$c$ will yield the same result. This graph $G^{''}$ is particullary 
+the respective factors and , it is possible
+to transform the previous condition on a map operation
+
+And therefore the same happens to marginal potentials as marginal potentials
+are the sum of potentials.
+
+to detect whether a given sample from the graph has variable
+$a$considered known to the system
+The system only knowns $k$ of those $N$ states.
+All the remaining ones are novel.
+
+Just like in previous sections it was assumed that: it is possible to estimate
+all the factors on the graph not dependent on $a$, using labelled data it is
+possible to estimate the factors dependent on $a$ given that
+$a \in \{A_1,\dotsc,A_K\}$. And using unlabelled data it is possible to estimate
+the factors on the graph $P(x)$ for $a \in \{A_1,\dotsc,A_N\}$.
+
+On this section it will be presented how to perform novelty detection of a
+given variable class assuming samples are drawn from a distribution described
+by the graph.
 

### 2011-06-10 thesis/novelty.tex
Finished explanation on how to detect novelty without assumption on P(novel) = constant.
No explanation on how to estimate the s_i parameters is given.

@@
 If on any graph structure the marginal probabilities of all variables except $a$
-is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$)
+is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$
 then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
@@
 \frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
-s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}}
+s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}} \quad \text{(constant)}
 \end{eqnarray*}
@@
 
-\begin{eqnarray}
+\begin{eqnarray*}
 s &=& \frac{\Phi_G(x)}{\Phi^\dag_G(x)} \\
@@
   &=& \frac{1}{\prod s_i}
+\end{eqnarray*}
+
+If factor $\phi_i$ appears $k_i$ times on a graph structure then:
+\begin{eqnarray}
+\Phi^\dag_G(x) \prod {s_i}^{k_i} = \Phi_G(x)
 \end{eqnarray}
 
-By adding a condition variable $c \in \{0, 1\}$, a condition factorthat is added result can be converted into a graph $G$is particularly interesting as it allows to detect novelty of $a$ by
-having models to approach
+Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x)/s_i$ are
+introduced then:
+\begin{eqnarray}
+\Phi^\ddag_G(x) = \Phi_G(x)
+\end{eqnarray}
 
-Similar to the procedure followed before, a decision variable $c \in \{0, 1\}$
-and a factor $\phi_c(c,a)$ will be added to $G$, creating graph $G'$.
-If $\phi_c$ is defined as:
-
-This is variable $c$ detects whether $a$ is a known or unknown class.
-Note also that by adding this condition variable and factor any marginal
-potential on $G$ is the same as the potential on $G'$. Since marginal potentials
-can be seen as sum of potentials on all variables it can be proved by showing
-the potential on any instance of all variables $X$ in $G$ is the same as the
-marginal potential of the same variables in $G'$:
+Having shown how to replace a variable $a$ for an equivalent $a^\dag$, and
+respectively change the factors dependent on $a$ from $\phi_i(a,\cdot)$ to
+$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials on the original
+graph are kept constant it becomes possible to go back to the graph $G'$ and
+replace the states $\{A^*_1,\dotsc,A^*_N\}$ for the equivalents $A^\ddag$
+and change the factors associated with $a'$ for:
 
 \begin{eqnarray*}
-\Phi_{G'}(x) &=& \sum_{c}\prod_{\phi \in G'}{\phi(x_\phi)} \\
-             &=& \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(0,a) 
-               + \left(\prod_{\phi \in G}{\phi(x_\phi)}\right)\phi_c(1,a)\\
-             &=& \Phi_G(x)\phi_c(0,a) + \Phi_G(x)\phi_c(1,a)\\
-             &=& \Phi_G(x)\\
+\phi_i(a', \cdot) &=& \left\{
+  \begin{array}{l l}
+    \phi_i(a',\cdot), & \qquad \text{if } a' \in \{A_1,\dotsc,A_K\} \\
+    \phi^\ddag_i(a', \cdot), & \qquad \text{if } a' \in A^\ddag
+  \end{array} \right. \\
 \end{eqnarray*}
 
-An optimal decision on whether variable $a$ of $G$ is $\in \{A_1,\dotsc,A_K\}$
-can be performed by a map \gls{MAP} operation on it. And that decision can be
-done solely based on potentials in $G'$:
+It is now possible to perform a decision on the novelty of $a$ with the condition:
+\begin{equation}
+2 \Phi_{G'}(x | c = 0) < \Phi_{G'}(x | c= 1)
+\end{equation}
 
-\begin{eqnarray*}
-P_G(a \in \{A_1,\dotsc,A_K\}|x) &<& P_G(a \in \{A_{K+1},\dotsc,A_N\}|x) \\
-\Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c = 1 \land x) \\
-2 \Phi_{G'}(c = 0 \land x) &<& \Phi_{G'}(c=0 \land x) + \phi_{G'}(c = 1 \land x)\\
-\end{eqnarray*}
+This condition is particularly useful since it does not depends on the normalization
+factors of $G'$. This allows to use \emph{intractable models} for some factors,
+e.g.\ using an \gls{SVM} for factors associated with sensed variables.
 
-By extending $a$ with duplicated states $\{A_1, \dotsc, A_K\}$, extending
-the respective factors and adding a factor $\phi_f$ on the decision
-variable $c$, it is possible to create a graph $G^{''}$ where a MAP decision on
-$c$ will yield the same result. This graph $G^{''}$ is particullary 
-the respective factors and , it is possible
-to transform the previous condition on a map operation
+\subsection{Estimating the Factor Scaling-factors $s_i$}
+On the previous section it was shown how to detect novelty on a variable $a$ by
+approximating the unconditional distribution factors with factors $\phi^\dag_i$
+that are a scalar $s_i$ away from the ones used on the real distribution.
 
-And therefore the same happens to marginal potentials as marginal potentials
-are the sum of potentials.
+This section address the problem of estimating those parameters.
 
-to detect whether a given sample from the graph has variable
-$a$considered known to the system
-The system only knowns $k$ of those $N$ states.
-All the remaining ones are novel.
-
-Just like in previous sections it was assumed that: it is possible to estimate
-all the factors on the graph not dependent on $a$, using labelled data it is
-possible to estimate the factors dependent on $a$ given that
-$a \in \{A_1,\dotsc,A_K\}$. And using unlabelled data it is possible to estimate
-the factors on the graph $P(x)$ for $a \in \{A_1,\dotsc,A_N\}$.
+\subsection{Old content}
 

### 2011-06-12 thesis/synthetic/Graph.cc
Started sample code for thesis.

@@ -1,0 +1,203 @@
+#include <boost/foreach.hpp>
+#include <cstdio>
+#include "Graph-internal.h"
+using namespace std;
+
+#define ROCS_DEBUG_STRINGIFY(x) #x
+#define ROCS_DEBUG_TO_STRING(x) ROCS_DEBUG_STRINGIFY(x)
+
+#define ROCS_DEBUG(...) { \
+  printf("[" __FILE__ ":" ROCS_DEBUG_TO_STRING(__LINE__) "] " __VA_ARGS__); \
+  printf("\n"); \
+}
+
+#define rocsDebug2(...) ROCS_DEBUG(__VA_ARGS__)
+#define rocsDebug1(...) ROCS_DEBUG(__VA_ARGS__)
+
+namespace {
+
+// Unfortunaly is_sorted is a SGI extension.
+// Therefore we need to define it.
+template<typename T>
+bool is_sorted(T begin, T end)
+{
+	if (begin == end)
+		return true;
+
+	T next = begin;
+	for (++next; next != end; begin = next, ++next)
+		if(*next < *begin)
+			return false;
+	return true;
+}
+
+// Builds a vector with the variable types of the given FactorData.
+// Returns false in case not all variables types are defined.
+bool GetVariableTypes(const FactorData &data,
+                      const map<string, VariableType> &known_types,
+                      vector<const VariableType*> *vartypes)
+{
+	typedef pair<string, string> PSS;
+	BOOST_FOREACH(const PSS &v, data.variables)
+	{
+		map<string, VariableType>::const_iterator it = known_types.find(v.second);
+		if (it == known_types.end())
+		{
+			rocsDebug2("Invalid FactorData: Unknown variable type '%s'",
+					v.second.c_str());
+			return false;
+		}
+		vartypes->push_back(&it->second);
+	}
+	return true;
+}
+}  // end namespace anonymous
+
+FactorData* GraphInformation::Add(const string &name, const FactorData &factor)
+{
+	rocsDebug2("Add factor type: '%s'", name.c_str());
+	if (factors.insert(make_pair(name, factor)).second == false)
+		rocsDebug1("Dropped factor: there is already a factor with that name '%s'",
+				name.c_str());
+    return &factors[name];
+}
+
+VariableType* GraphInformation::Add(const string &name, const VariableType &type)
+{
+	rocsDebug2("Add variable type: '%s'", name.c_str());
+	if (types.insert(make_pair(name, type)).second == false)
[…]

### 2011-06-12 thesis/synthetic/Query.cc

@@ -1,0 +1,244 @@
+// ==================================================================
+// ROCS - Toolkit for Robots Comprehending Space
+// Copyright (C) 2011  André Susano Pinto
+//
+// This file is part of ROCS.
+//
+// ROCS is free software: you can redistribute it and/or modify it
+// under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation, either version 3
+// of the License, or (at your option) any later version.
+//
+// ROCS is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with ROCS. If not, see <http://www.gnu.org/licenses/>.
+// ==================================================================
+#include <cstdio>
+
+#define ROCS_DEBUG_STRINGIFY(x) #x
+#define ROCS_DEBUG_TO_STRING(x) ROCS_DEBUG_STRINGIFY(x)
+
+#define ROCS_DEBUG(...) { \
+  printf("[" __FILE__ ":" ROCS_DEBUG_TO_STRING(__LINE__) "] " __VA_ARGS__); \
+  printf("\n"); \
+}
+
+#define rocsDebug3(...) ROCS_DEBUG(__VA_ARGS__)
+#define rocsDebug2(...) ROCS_DEBUG(__VA_ARGS__)
+#define rocsDebug1(...) ROCS_DEBUG(__VA_ARGS__)
+
+
+#include "Query.h"
+
+#define DAI_WITH_JTREE
+#define DAI_WITH_BP
+#include <dai/alldai.h>
+#include <dai/factorgraph.h>
+#include <dai/gibbs.h>
+#include <dai/properties.h>
+#include <map>
+#include <string>
+
+#include "Graph-internal.h"
+
+using namespace std;
+
+// This handles all internals of Query class.
+// It should totally hide its dependency on libDAI.
+class QueryInternal {
+  public:
+	QueryInternal()
+	{
+		size_t maxiter = 10000;
+		dai::Real  tol = 1e-9;
+		size_t verbose = 1;
+
+		daiOptions_.set("maxiter", maxiter);  // Maximum number of iterations
+		daiOptions_.set("tol", tol);          // Tolerance for convergence
+		daiOptions_.set("verbose", verbose);
+	}
+
[…]

### 2011-06-12 thesis/synthetic/Query.h

@@ -1,0 +1,70 @@
+// ==================================================================
+// ROCS - Toolkit for Robots Comprehending Space
+// Copyright (C) 2011  André Susano Pinto
+//
+// This file is part of ROCS.
+//
+// ROCS is free software: you can redistribute it and/or modify it
+// under the terms of the GNU Lesser General Public License as
+// published by the Free Software Foundation, either version 3
+// of the License, or (at your option) any later version.
+//
+// ROCS is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU Lesser General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with ROCS. If not, see <http://www.gnu.org/licenses/>.
+// ==================================================================
+
+/*!
+ * \author André Susano Pinto
+ * \file Query.h
+ */
+
+#ifndef _ROCS_CONCEPT_QUERY_H_
+#define _ROCS_CONCEPT_QUERY_H_
+
+#include <boost/scoped_ptr.hpp>
+#include <vector>
+#include <string>
+
+class Graph;
+class Variable;
+class QueryInternal;
+
+class Query {
+ public:
+	// This class does not takes ownership over Graph.
+	// Graph and all associated variables have to outlive this instance.
+	explicit Query(const Graph *graph);
+	~Query();
+
+	// Calculates probability for a given set of variables.
+	//
+	// Output is the probability for all combinations over the given variables.
+	// Output is lexicographic ordered by the value of its variables.
+	// This means:
+	//   output[0] = P( Var1 = v0, Var2 = v0, ... VarN = v0]
+	//   output[1] = P( Var1 = v0, var2 = v0, ... VarN = v1]
+	//   ...
+	//   output[last] = P( Var1 = vN, Var2 = vN, ... VarN = vN]
+	void Marginalize(const std::vector<const Variable*> &variables,
+	                 std::vector<double> *output);
+    
+    void Sample(const std::vector<const Variable*> &variables,
+                std::vector<std::string> *output);
+ protected:
+	const Graph *graph_;
+	boost::scoped_ptr<QueryInternal> internal_;
+
+	void BuildInternal();
+
+ private:
+	// DISALLOW COPY AND ASSIGN
+	Query(const Query&);
+	void operator=(const Query&);
+};
+
+#endif  // _ROCS_CONCEPT_QUERY_H_

### 2011-06-12 thesis/synthetic/main.cc

@@ -1,0 +1,187 @@
+/**
+ * Novelty Detection Scripts
+ *
+ * Copyright 2011: André Susano Pinto
+ *
+ * This program is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU General Public License
+ * as published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software Foundation,
+ * Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+ */
+#include "Graph.h"
+#include "Query.h"
+#include "util.h"
+
+#include <cassert>
+#include <set>
+#include <algorithm>
+using namespace std;
+
+#include <boost/foreach.hpp>
+#include <boost/scoped_ptr.hpp>
+using namespace boost;
+
+#include <dai/factorgraph.h>
+#include <dai/gibbs.h>
+#include <dai/properties.h>
+using namespace dai;
+
+GraphInformation gi;
+const VariableType* roomType;
+const VariableType* senseType;
+const FactorData* room_room;
+const FactorData* room_sense;
+
+const string ROOM_TYPE  = "ROOM";
+const string SENSE_TYPE = "SENSE";
+
+
+string room(int i)  { return "room" + to_string(i); }
+string sense(int i) { return "sense" + to_string(i); }
+
+// Initializes GI with a random real distribution.
+void gi_init_random_real(int roomCats, int senseCats) {
+  // Generate states for variables of type Room
+  {
+    VariableType roomCat;
+    for (int i = 0; i < roomCats; ++i)
+      roomCat.values.push_back(room(i));
+    sort(roomCat.values.begin(), roomCat.values.end());
+    roomType = gi.Add(ROOM_TYPE, roomCat);
+  }
+
+  // Generate states for sensed properties
+  {
+    VariableType senseCat;
+    for (int i = 0; i < senseCats; ++i)
+      senseCat.values.push_back(sense(i));
+    sort(senseCat.values.begin(), senseCat.values.end());
[…]

### 2011-06-12 thesis/synthetic/main.cc
Added dot output capabilities.

@@ -18,10 +18,12 @@
  * Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
  */
 #include "Graph.h"
+#include "Graph-internal.h"
 #include "Query.h"
 #include "util.h"
 
 #include <cassert>
+#include <cstdio>
 #include <set>
 #include <algorithm>
 using namespace std;
@@ -162,6 +164,37 @@ public:
     }
   }
 
+  void print(const vector<const Variable*> &obs_vars, const vector<string> &obs_vals) {
+     map<const Variable*, string> id;
+     BOOST_FOREACH(const Variable* var, vars)
+       id[var] = "n" + to_string(id.size());
+
+	 vector<const ::Factor*> factors;
+     ListFactors(&factors);
+     BOOST_FOREACH(const ::Factor* fac, factors)
+       BOOST_FOREACH(const Variable* var, fac->variables_)
+         if (id.count(var) == 0)
+           id[var] = "n" + to_string(id.size());
+
+     printf("graph sample {\n");
+     printf(" node [shape=circle];\n");
+     BOOST_FOREACH(const Variable* var, vars)
+       printf("  %s;\n", id[var].c_str());
+
+     for (size_t i = 0; i < obs_vars.size(); ++i)
+       printf("  %s[label=%s];\n", id[obs_vars[i]].c_str(), obs_vals[i].c_str());
+
+     int nf = 1;
+     printf(" node [shape=square];\n");
+     BOOST_FOREACH(const ::Factor* fac, factors) {
+       string fid = "f" + to_string(nf++);
+       printf("  %s;\n", fid.c_str());
+       BOOST_FOREACH(const Variable* var, fac->variables_)
+         printf("  %s -- %s;\n", fid.c_str(), id[var].c_str());
+     }
+     printf("}\n");
+  }
+
 };
 
 
@@ -172,15 +205,13 @@ int main() {
   assert(gi.CheckConsistency());
 
   GraphStructure s;
-  s.createRandom(4, 4, 1);
+  s.createRandom(3, 10, 1);
   MGraph g(s, roomType, senseType, room_room, room_sense);
 
   Query q(&g);
   vector<string> output;
   q.Sample(g.vars, &output);
-
-  BOOST_FOREACH(const string &s, output)
-    cout << s << endl;
+  g.print(g.vars, output);
 
   return 0;
 }

### 2011-06-12 thesis/novelty.tex
Some cleanup

@@
 \subsection{Old content}
-
-On this section it will be presented how to perform novelty detection of a
-given variable class assuming samples are drawn from a distribution described
-by the graph.
-
-On the previous section it was shown how to detect a novel class on a variable
-assuming that the rest of the graph variables and structure is correctly modeled
-by the agent knowledge.
-In this section methods are presented that not only allow to extended this to
-detection on multiple variables but also allow the system to incorporate
-novelty information back in the graphical model in order to continue its
-probabilistic and uncertain reasoning.
-
-When trying to extend the method to deal with multiple novel variables it may
-be tempting to follow a deterministic approach such as:
-when the agent determines a variable $a$ to have a novel class, it marks it as
-novel and proceeds from that moment by using a graphical model where the factors
-associated with the $a$ replaced by unconditional factors.
-But such an approach would not be able to optimally detect novel variables
-as later decisions are dependent on former ones.
-It would also not be able to adapt itself and reconsider previously decisions
-when more data becomes available for the agent.
-
-To handle those issues a fully probabilistic approach should be taken
-instead of a deterministic.
-That can be done by extending all the variables on the factor graph representing
-the semantic map with an additional state \emph{novel} and extending all the
-factors with information on the unconditional distribution.
-
-Such a factor graph would be able to handle novel variables and incorporate
-prior-information on the unconditional distributions since it is able to
-represent any state of the $X$ variables present on the graph where any
-subset of them is novel. But in order to add the ability to control the
-novelty thresholds a normalization factor $\phi_N$ needs to be introduced.
-
-\begin{figure}[h]
-\centering
-Yeah get a figure for here.
-\caption{\label{fig:multiple-big-threshold}Factor graph extended to be able to
-         represent novelty on any subset of variables. Note the added factor
-         $\phi_N$ that servers as a normalization factor such that probability
-         with any subset of $Y$ variables considered novel can be scaled with
-         any value such that any thresholding decision mechanism can be
-         implemented.}
-\end{figure}
-
-In order to give full control on the threshold on which any subset of
-variables is considered novel the factor $\phi_N$ needs to be dependent on all
-the $X$ variables and act as a weighting function that correctly scales the
-likelihood of a given instance of the $X$ variables.
-If the criterium for defining the threshold is the likelihood of all variables
-$X$ being non-novel has the max likelihood if and only if
-\autoref{eq:multiple-threshold} holds for all the non-empty subsets $Y$ of $X$.
-The weight for $\phi_N$ when the set $Y$ is classified as novel
-can be set for $t_Y$ and $\phi_N$ for the empty subset of novel variables is set
-to $1$.
-
-\begin{equation}
-\label{eq:multiple-threshold}
-\frac{P_G(x)}{P_{U_Y}(x)} > t_Y
-\end{equation}
-
-Still without further assumptions a threshold $t_Y$ would need to be defined for
-all the $2^N-1$ non-empty subsets.
-By taking a simple approach of considering the novelty on a variable independent
-from the novelty on any other variable $t_Y$ can be factorized:
-
-\begin{equation}
-t_Y = \prod_{i \in Y}{t_i}
-\end{equation}
-
-This factorization yields then the interesting property to allow to split the
-threshold factor $\phi_N$ into $N$ different $\phi_{N_i}$ associated with each
-variable of the graph. A graph able to handle novelty detection on several
-variables with the described factorization is illustrated on
-\autoref{fig:multi-factorized-threshold}.
-
-\begin{figure}[h]
-\centering
-Yeah get a figure for here.
-\caption{\label{fig:multi-factorized-threshold}Factor graph extended to be able
-         to represent novelty on any subset of variables. Note the added factor
-         $\phi_N$ has been split into a $\phi_{N_i}$ still it is able to perform
-         detection on multiple variables according to the factorization
-         described in the text.}
-\end{figure}
-
 To sum up, the method described above allows to integrate novelty detection
@@
 \section{Detecting New Structure}
+I do not belive there is any time left for this.

### 2011-06-12 thesis/semantic-mapping.tex
Content on the semantic mapping chapter.

@@
 
+% Reasoning on why to use semantic mapping
+% High-level processing
+Also due to uncertainity and complexity 
 % Advantages of Multi-modal approaches
@@
 
+It is therefore not only for efficiency reasons but also desirable to handle
+the information on an high-level probabilistic fashion.
+
 \section{Dora Architecture Overview}
+This thesis motivation lies on developing methods for detecting knowledge gaps
+on the semantic knowledge of a mobile agent. For that the semantic mapping system
+proposed by \cite{pronobis2011phd} was used as a base.
+
+Pronobis presents a system architecture working on indoor environments using
+non\hyp{}omnidirectional lasers and visual sensors, that by integrating cues such as geometry,
+object presence and appearances is able to perform inference across any semantic property,
+for example to handle place categorization.
+The introduced semantic representation has not only been tested on real scenarios
+and performance measured across several conditions~\cite{pronobis2007iros} but also shown
+to be tailored to effectively solve tasks arising on mobile robotics~\cite{hanheide2011ijcai}.
+Besides that the system architecture has been drawn with an high-focus on probabilistic
+and uncertain reasoning, human interaction and life-long learning capabilities.
+
+For that it was considered an excellent base defining not only the concepts and
+requirements of an high-level semantic representation but also the information flow
+between all the layers involved on the agent.
+The system is implemented on \Gls{Dora} and a brief overview of its architecture is
+given in the following sections.
+
 \subsection{System organization}
-\subsubsection{Sensory Layer}
-\subsubsection{Categorical Layer}
-\subsubsection{Place Layer}
-\subsubsection{Conceptual Layer}
+The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
+sub-systems, all of which actively use or maintain the spatial knowledge representation
+(see \autoref{fig:dora-architecture}):
 
-\section{Features}
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/dora-architecture.pdf}
+\caption{\label{fig:dora-architecture}Interaction of the sub-systems
+         in Dora with special focus on the conceptual layer.}
+\end{figure}
 
-\section{Conceptual Knowledge}
+\subsubsection*{Sensory Layer}
+The sensory layer is responsible for all the low-level sensors.
+It maintains an immediate and exact description of the environment by using
+data such as images from cameras, 2D\hyp{}laser scans and odometry from encoders in the
+movement motors. The information provided by it is used on the higher layers to extract
+tractable information.
+
+\subsubsection*{Place Layer}
+The place layer is responsible for discretizing the continuous low-level space into
+a finite number of places by spreading several virtual places markers around the whole
+environment and maintaining connectivity information between them.
+This virtual places also handle information associating the robot view point. Together
+with extracted properties they allow the system to perform a clever integration over the
+room properties, avoiding sampling bias, for example coming due to the robot spend too
+much time in a part of the room~\cite{pronobis2010ijrr}.
+
+\subsubsection*{Categorical Layer}
+The categorical layer is responsible for extracting high-level information from the
+sensors. It has classifiers trained to detect several properties from the visual and laser
+features captured on the sensory layer.
+It is responsible for detecting objects, doors, categorize the geometry and the appearance.
+The used methods and properties extracted by this layer are described in more detail in
+\autoref{sec:dora-features}.
+
+\subsubsection*{Conceptual Layer}
+The conceptual layer merges the sensed properties and places structure together
+with general knowledge under a unified probabilistic semantic representation: the
+\emph{conceptual map}.
+This conceptual map is the structure for handling the high-level semantic information
+and performing inferences. It is the layer where the novelty detection methods described
+on this thesis are meant to be integrated and is described in more detail in
+\autoref{sec:conceptual-map}.
+
+\subsection{Features}
+\label{sec:dora-features}
+
 \section{Conceptual Map}
-\label{sec:clues-from-low-level}
+\label{sec:conceptual-map}
+As \gls{Dora} moves through the environment its \emph{conceptual layer} builds a structural and
+probabilistic representation of the space instantiated as a \emph{graphical model}.
+It includes taxonomy of human\hyp{}compatible spatial concepts which are linked to the sensed 
+instances of these concepts drawn from lower layers. It is the conceptual layer which 
+contains the information that kitchens commonly contain cereal boxes and have certain 
+general appearance and allows the robot to infer that the cornflakes box in front of the 
+robot makes it more likely that the current room is a kitchen. The conceptual layer is 
+described in terms of a probabilistic ontology defining spatial concepts and linking 
+those concepts to instances of spatial entities (see the example of the ontology in
+\autoref{fig:dora-architecture}).
+
+Based on this design, a \emph{chain graph} model is proposed as a 
+representation for performing inferences on the knowledge represented in the conceptual 
+layer. Chain graphs are probabilistic graphical models that combine the properties of 
+both Bayesian Networks and Random Markov Fields. This results in an efficient approach to 
+probabilistic modeling and reasoning about conceptual knowledge.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.75\textwidth]{figures/chaingraph.pdf}
+\caption{\label{fig:chain-graph}Example chain-graph produced by the \emph{conceptual layer}.}
+\end{figure}
+
+%The conceptual layer structures the sensed environment together with the conceptual knowledge
+%in order to create a structured probabilistic representation of the world.
+An exemplary chain graph corresponding to the conceptual map ontology is presented
+in \autoref{fig:chain-graph}. 
+Each discrete place identified in the environment is represented by a set of random variables, 
+one for each class of relation linked to that place. These are each connected to a random variable
+over the categories of rooms, representing the ``is-a'' relation between rooms and their categories. 
+Moreover, the room category variables are connected by undirected links to one another according 
+to the topological map. The remaining variables represent: shape and appearance properties of space 
+as observed from each place, and the presence of objects. 
+These are connected to observations of features extracted directly from 
+the sensory input. Finally, the 
+distributions $p_{s}(\cdot|\cdot)$, $p_a(\cdot|\cdot)$, $p_{o_i}(\cdot|\cdot)$ 
+represent the common sense knowledge about shape, appearance, and object co-occurrence, respectively. 
+They allow for inference about other properties and room categories e.g. that the room is likely to be a kitchen,
+because you are likely to have observed cornflakes in it.
+
+The use of graphical models to describe distributions of variables has useful properties.
+First, they permit inference about uncertain conceptual knowledge. At the same time, they are 
+generative models and therefore allow to calculate the probability
+on any given subset of variables of the graph, allowing the system to work even when some
+information is missing.
+
+\subsection{Factor Graphs}
+Although the conceptual layer works with \emph{chain graphs}~\cite{lauritzen2002chain},
+those can be converted into \emph{factor graphs}~\cite{kschischang2001factor}.
+Factor graphs have been introduced in \autoref{sec:graphical-models} and are used
+throughout this thesis as they provide a easier manipulation due to factorization.
+
+Moreover, there exists efficient implementation of inference engines operating on factor
+graph representations~\cite{Mooij_libDAI_10}.
+Describing the distribution function in terms of graphs allows to use those engines to
+efficently calculate marginals on any given subset of variables by exploiting conditional
+independence between variables.
+
+\subsection{Uncertain Sensing and Implicit Factors}
+\label{sec:cues-from-low-level}
+In a realistic scenario classification tasks are by their nature uncertain.
+And although a trained classifier is able to decide which class is more likely, there is
+interest in handling uncertainty produced by those to turn decisions more robust.
+
+By modelling the classifiers as an implicit function $\phi_{classifier}(c, x)$ it becomes
+possible to model the uncertainty of those given the presence of the sensed low-level
+features $X$. The presence of those implicit models no longer allow to calculate the
+normalization factor over the whole graph as there is no explicit model for either the
+low\hyp{}level features $X$ (e.g.\ $X$ can be an image received from a visual sensor)
+or for $\phi_{classifier}$ (e.g.\ the classifier can be an \gls{SVM}).
+
+Nonetheless it is still possible to calculate any probability given that the features $X$
+are observed. For that instead of modelling the variables $X$ and explicitly describing
+the factor $\phi_{classifier}$ it is enough to replace it by an observed factor $\phi_{classifier}(c|x)$.
+
+\begin{figure}[h]
+\centering
+\subfloat[]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (graph) [] {\dots}; \& \& \\
+    \& \& \node (prop) [latent] {$c$}; \& \& \\
+    \& \& \node (fac) [factor] {}; \& \& \\
+    \& \& \& \& \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (prop) -- (graph);
+  \draw [-] (prop) -- (fac);
+  \draw [-] (fac) -- (x1);
+  \draw [-] (fac) -- (x2);
+  \draw [-] (fac) -- (x3);
+  \draw [-] (fac) -- (xi);
+  \draw [-] (fac) -- (xn);
+
+  \node (captfactor) [right=2pt of fac] {\footnotesize{$\phi_{classifier}(c, x)$}};
+\end{tikzpicture}
+}
+\qquad
+\subfloat[]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (graph) [] {\dots}; \& \& \\
+    \& \& \node (prop) [latent] {$c$}; \& \& \\
+    \& \& \node (fac) [factor] {}; \& \& \\
+    \& \& \node (hidden) [latent,draw=none] {}; \& \& \\
+  };
+  \draw [-] (prop) -- (graph);
+  \draw [-] (prop) -- (fac);
+  \node (captfactor) [right=2pt of fac] {\footnotesize{$\phi_{classifier}(c | x)$}};
+\end{tikzpicture}
+
+}
+
+\caption{\label{fig:implicit-factor}By modelling classifiers as implicit factors it is
+possible to perform inference on the graphical models handling the uncertainty measures
+produced by those classifiers.}
+\end{figure}

### 2011-06-12 thesis/background.tex
Semantic Mapping changes.

@@
 distance. Other variations exists as is the case of using the minimal distance
-to the average classification distance of each class~\citep{pronobis2007confidence}.
+to the average classification distance of each class~\citep{pronobis2007iros}.
 

### 2011-06-12 thesis/novelty.tex
Semantic Mapping changes.

@@
     1, & \qquad \text{if } c = 0 \text{ and } b \in \{A_1,\dotsc,A_n\} \\
-    1, & \qquad \text{if } c = 1 \text{ and } b \in \{A^*_1,\dotsc,A^*_n\}\\
+    1, & \qquad \text{if } c = 1 \text{ and } b \in \{A^*_1,\dotsc,A^*_k\}\\
     0, & \qquad \text{otherwise}
@@
 P_G(x) Z_G \phi_t(c=0) &<& P_{U_a}(x) Z_{U_a} \phi_t(c=1) \\
-\frac{P_G(x)}{P_{U_a}(x)} &<& \frac{\phi_t(c=1)}{\phi_t(c=0)}\frac{Z_{U_a}}{Z_G}
+\frac{P_G(x)}{P_{U_a}(x)} &<& \frac{Z_{U_a}}{Z_G}
 \end{eqnarray*}

### 2011-06-12 thesis/semantic-mapping.tex
Semantic Mapping changes.

@@
 
-% Reasoning on why to use semantic mapping
-% High-level processing
-Also due to uncertainity and complexity 
-% Advantages of Multi-modal approaches
-%
+There has been several efforts in the areas of artificial intelligence and robotics
+in creating robots that are able to interact with humans and their environments.
+One of the important aspects is to endow those robots with a deeper understanding
+of human environments, not just for the purpose of navigation and obstacle avoidance, 
+but also in terms of human semantics and functionality.
 
-It is therefore not only for efficiency reasons but also desirable to handle
-the information on an high-level probabilistic fashion.
+In that sense the need for semantic mapping arises, and it can be described as the
+relation between spatial areas and a set of existent concepts used by humans to
+describe them.
+Humans associate concepts to areas that relate to the functionality of those. That
+association is either useful by explaining where to find certain objects as well to
+understand the types of activities to perform on it.
+By being able to reason over those semantics an agent can be expected to better
+understand human environments and increase its efficiency when performing tasks on them.
 
-\section{Dora Architecture Overview}
+It is therefore clear that the need to handle information on an high-level semantic and
+probabilistic fashion arises not only for efficiency reasons but also from the desire
+of being able to communicate and reason on those aspects.
+In order to fully explore the semantic properties and their relations it is necessary to
+detect and classify semantic aspects as well to understand the relationships between them.
+That conceptual knowledge is important and any limitations on it will limit the capacities
+of an agent to reason.
+
 This thesis motivation lies on developing methods for detecting knowledge gaps
@@
 
+\section{Dora Architecture Overview}
 Pronobis presents a system architecture working on indoor environments using

### 2011-06-13 thesis/novelty.tex
Fixes here and there.

@@
 \begin{eqnarray*}
- P_B(c=0 | x)         &<& P_B(c=1 | x) \\
-\Phi_B(c=0 \land x)   &<& \Phi_B(x=1 \land x) \\
-\Phi_G(x) \phi_t(c=0) &<& \Phi_{U_a}(x) \phi_t(c=1) \\
-P_G(x) Z_G \phi_t(c=0) &<& P_{U_a}(x) Z_{U_a} \phi_t(c=1) \\
-\frac{P_G(x)}{P_{U_a}(x)} &<& \frac{Z_{U_a}}{Z_G}
+ \frac{P_G(x)}{P_{U_a}(x)}&<& t_A \\
+ \frac{\Phi_G(x)}{\Phi_{U_a}(x)}    &<& t_A \frac{Z_G}{Z_{U_a}} \\
+ \frac{\Phi_B(x, c=0)}{\Phi_{B}(x, c=1)}    &<& t_A \frac{Z_G}{Z_{U_a}} \\
 \end{eqnarray*}
@@
 is known for all the factors disabling the ability to incorporate probabilistic
-sensing output from complex detectors such as \glspl{SVM}. Ultimately, it also
-removes capability to mix complex thresholds on several variables that would
-allow to do probabilistic inference on the novelty of multiple variables at once.
+sensing output from complex detectors such as \glspl{SVM}.
 
@@
 \end{tikzpicture}
-\quad
+\qquad
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+\qquad
 \begin{tikzpicture}
@@
 \begin{eqnarray*}
-\Phi_{G'}(x, c=0) &=& \Phi_{G}(x | a \in \{A_1, \dots, A_K\})\\
+\Phi_{G'}(x, c=0) &=& \Phi_{G}(a \in \{A_1, \dots, A_K\}, x)\\
 \Phi_{G'}(x, c=1) &=& \Phi_{G}(x) 
@@
 knowledge of all the possible classes, which is not available.
-Nonetheless access to unlabelled data or other informations or assumptions on
-the unconditional distribution allow to model or approximate unconditional
-distributions of $a$. For that there is interest in trying to model the set
-of states $\{A^*_1, \dots, A^*_N\}$ with a set of states $\{A^{\dag}_1, \dotsc, A^{\dag}_M\}$.
+But by having access to unlabelled data, other informations or assumptions on
+the unconditional distribution it is possible to model the unconditional
+distributions of $x$. For that there is interest in trying to model the real set
+of states $\{A^*_1, \dots, A^*_N\}$ with a set of states $\{A^{\dag}_1, \dotsc, A^{\dag}_M\}$
+learned or assumed from information on the unconditional distribution.
 
-Moreover assuming that the set of states $A^\dag$ is able to perfectly
-characterize the distribution over any graph $G$,
-for a specific graph the potentials with states $A^\dag$ are a scale-factor
-away from the potentials with states $A$. And that scale-factor is dependent on
-the graph structure and factors of both $G$ and $G^\dag$:
+When changing a variable states from $A^*$ into $A^\dag$ it also becomes necessary to redefine
+the factors dependent on $a$ to accommodate the new states $A^\dag$.
+Moreover assuming that the change of variable states $A^\dag$ is able to perfectly
+characterize the distribution over any graph structure,
+for a specific graph $G$ the potentials $\Phi^\dag_G(x)$ with states $A^\dag$ are a
+scale\hyp{}factor away from the potentials $\Phi_G(x)$ with states $A$.
+And that scale-factor is dependent on the graph structure and factors of both $G$ and $G^\dag$:
 
@@
 \autoref{eq:structure-ratio} it has a special property. Here both $G$ and
-$G^\dag$ model the same probability, so it is expected that both scale they
+$G^\dag$ model the same distribution, so it is expected that both scale they
 normalization factor according to some relation on how a given factor differs
@@
 is the same either by using $A$ with $\phi$, or by using $A^\dag$ with $\phi^\dag$
-then $\sum_{A^\dag}\phi^\dag_i(a^\dag, \cdot) = s_i \sum_A \phi_i(a, \cdot)$.
-It can be proved that since the distribution over any graph structure has to be
+then it is possible to scale each individual factor $\phi_i^\dag$ with a constant $1/s_i$
+such that the potentials with $A$ will be equal to the potentials with $A^\dag$.
+
+Since the distribution over any graph structure has to be
 equal, then the same holds for a graph structure $P_{G_i}$ composed by a single
@@
 P^\dag_{G_i}(x) &=& P_{G_i}(x) \\
-\frac{\phi_i(x)}{Z^\dag_{G_i}}  &=& \frac{s_i \phi_i(x)}{Z_{G_i}} \\
+\frac{\sum_{a^\dag \in A^\dag}\phi^\dag_i(a^\dag, x)}{Z^\dag_{G_i}}  &=& \frac{\sum_{a \in A}\phi_i(x)}{Z_{G_i}} \\
 s_i &=& \frac{Z_{G_i}}{Z^\dag_{G_i}} \quad \text{(constant)}
@@
 \begin{eqnarray}
-\Phi^\dag_G(x) \prod {s_i}^{k_i} = \Phi_G(x)
+\frac{\Phi^\dag_G(x)}{\Phi_G(x)} =  \prod {s_i}^{k_i}
 \end{eqnarray}
 
-Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x)/s_i$ are
+Alternative, if factors $\phi^\ddag_i(a^\dag,x) = \phi^\dag_i(a^\dag,x) s_i$ are
 introduced then:
@@
 respectively change the factors dependent on $a$ from $\phi_i(a,\cdot)$ to
-$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials on the original
-graph are kept constant it becomes possible to go back to the graph $G'$ and
+$\phi^\ddag_i(a^\dag,\cdot)$ such that the potentials are kept equal equal
+it becomes possible to go back to the graph $G'$ and
 replace the states $\{A^*_1,\dotsc,A^*_N\}$ for the equivalents $A^\ddag$
@@
 \begin{equation}
-2 \Phi_{G'}(x | c = 0) < \Phi_{G'}(x | c= 1)
+2 \Phi_{G'}(x, c = 0) < \Phi_{G'}(x, c= 1)
 \end{equation}

### 2011-06-13 thesis/synthetic/Graph.cc
Recode and refactor

@@ -190,7 +190,6 @@ const Variable* Graph::CreateVariable(const VariableType *type)
 const Factor* Graph::CreateFactor(const FactorData *data,
                            const vector<const Variable*> &variables)
 {
-    rocsDebug2("Creating a factor of type %x", data);
 	Factor* factor = new Factor(data, variables);
 	factors_.push_back(factor);
 	return factor;

### 2011-06-13 thesis/synthetic/Query.cc

@@ -100,19 +100,19 @@ class QueryInternal {
 			if (factorGraph_.get() == NULL)
 				MakeFactorGraph();
 
-			rocsDebug3("Running Loopy Belief Propagation...");
+			//rocsDebug3("Running Loopy Belief Propagation...");
 			bp_.reset(new dai::BP(*factorGraph_,
 					daiOptions_("updates", string("SEQRND"))
 					           ("logdomain", false)));
 			bp_->init();
 			bp_->run();
-			rocsDebug3("Finished Loopy Belief Propagation...");
+			//rocsDebug3("Finished Loopy Belief Propagation...");
 		}
 	}
 
 	void MakeFactorGraph()
 	{
-		rocsDebug3("Creating DAI FactorGraph...");
+		//rocsDebug3("Creating DAI FactorGraph...");
 		factorGraph_.reset(new dai::FactorGraph(factors_));
 
 		// Invalidate BP
@@ -128,7 +128,7 @@ class QueryInternal {
 			dai::SmallSet<dai::Var>(dai_vars.begin(), dai_vars.end(),
 			                        dai_vars.size()));
 
-		rocsDebug3("Asking belief on set of %ld variables.", dai_varset.size());
+		//rocsDebug3("Asking belief on set of %ld variables.", dai_varset.size());
         dai::Factor marginal = bp_->belief(dai_varset);
 
 		// Convert output to original variable order.
@@ -153,12 +153,12 @@ class QueryInternal {
             gibbsProps.set("burnin", size_t(100));
             gibbsProps.set("verbose", size_t(0));
 	
-            rocsDebug3("Creating GibbsSampler...");
+            //rocsDebug3("Creating GibbsSampler...");
 			gibbs_.reset(new dai::Gibbs(*factorGraph_, gibbsProps));
             
 			gibbs_->init();
 			gibbs_->run();
-			rocsDebug3("Finished Loopy Belief Propagation...");
+			//rocsDebug3("Finished Loopy Belief Propagation...");
 		}
 	}
     void Sample(const vector<const Variable*> &vars,

### 2011-06-13 thesis/synthetic/main.cc

@@ -37,182 +37,11 @@ using namespace boost;
 #include <dai/properties.h>
 using namespace dai;
 
-GraphInformation gi;
-const VariableType* roomType;
-const VariableType* senseType;
-const FactorData* room_room;
-const FactorData* room_sense;
-
-const string ROOM_TYPE  = "ROOM";
-const string SENSE_TYPE = "SENSE";
-
-
-string room(int i)  { return "room" + to_string(i); }
-string sense(int i) { return "sense" + to_string(i); }
-
-// Initializes GI with a random real distribution.
-void gi_init_random_real(int roomCats, int senseCats) {
-  // Generate states for variables of type Room
-  {
-    VariableType roomCat;
-    for (int i = 0; i < roomCats; ++i)
-      roomCat.values.push_back(room(i));
-    sort(roomCat.values.begin(), roomCat.values.end());
-    roomType = gi.Add(ROOM_TYPE, roomCat);
-  }
-
-  // Generate states for sensed properties
-  {
-    VariableType senseCat;
-    for (int i = 0; i < senseCats; ++i)
-      senseCat.values.push_back(sense(i));
-    sort(senseCat.values.begin(), senseCat.values.end());
-    senseType = gi.Add(SENSE_TYPE, senseCat);
-  }
-
-  // Generate factors between Room Categories
-  {
-    // This factor is symmetric.
-    // This is the approach that is taken on the room connectivity
-    // otherwise a scheme to figure out on which direction to plug
-    // the factor would need to be devised.
-    FactorData rr;
-    rr.variables.push_back(make_pair(string("room1"), ROOM_TYPE));
-    rr.variables.push_back(make_pair(string("room2"), ROOM_TYPE));
-    vector<string> index(2);
-    for (int i = 0; i < roomCats; ++i)
-    for (int j = i; j < roomCats; ++j) {
-      double potential = random();
-      index[0] = room(i);
-      index[1] = room(j);
-      rr.potential[index] = potential;
-      swap(index[0], index[1]);
-      rr.potential[index] = potential;
-    }
-    room_room = gi.Add("room-room", rr);
-  }
-
-  // Generate factors for Room Category and Sensed Property
-  {
-    FactorData rs;
-    rs.variables.push_back(make_pair(string("room"), ROOM_TYPE));
-    rs.variables.push_back(make_pair(string("sense"), SENSE_TYPE));
[…]

### 2011-06-13 thesis/synthetic/single-factor.cc

@@ -1,0 +1,298 @@
+/**
+ * Novelty Detection Scripts
+ *
+ * Copyright 2011: André Susano Pinto
+ *
+ * This program is free software; you can redistribute it and/or
+ * modify it under the terms of the GNU General Public License
+ * as published by the Free Software Foundation; either version 2
+ * of the License, or (at your option) any later version.
+ *
+ * This program is distributed in the hope that it will be useful,
+ * but WITHOUT ANY WARRANTY; without even the implied warranty of
+ * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+ * GNU General Public License for more details.
+ *
+ * You should have received a copy of the GNU General Public License
+ * along with this program; if not, write to the Free Software Foundation,
+ * Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
+ */
+#include "Graph.h"
+#include "Graph-internal.h"
+#include "Query.h"
+#include "util.h"
+
+#include <cassert>
+#include <cstdio>
+#include <set>
+#include <iostream>
+#include <algorithm>
+using namespace std;
+
+#include <boost/foreach.hpp>
+#include <boost/scoped_ptr.hpp>
+using namespace boost;
+
+/**
+ * This experiment tries to approach a graph where variable
+ * a is only connected with a single type of factor \phi.
+ * 
+ */
+namespace {
+string room(int i) { return "r" + to_string(i); }
+string prop(int i) { return "p" + to_string(i); }
+
+vector<string> vector_of(const string &a, const string &b) {
+  vector<string> t(2);
+  t[0] = a, t[1] = b;
+  return t;
+}
+
+struct GraphStructure {
+  int rooms;
+  int senses;
+  set<pair<int,int> > edges;
+
+  void createRandom(int rooms, int senses, int extra_conn) {
+    this->rooms = rooms;
+    this->senses = senses;
+    // Creates a tree
+    for (int i = 1; i < rooms; ++i) {
+      int other = random() % i;
+      edges.insert(make_pair(other, i));
+    }
+    // Tries to add extra room connectivity
+    while (extra_conn--) {
+      int a = random() % rooms;
+      int b = random() % rooms;
+      if (a == b) continue;
+      if (a > b) swap(a, b);
+      edges.insert(make_pair(a, b));
+    }
[…]

### 2011-06-13 thesis/synthetic/single-factor.cc
Used gi.Add to create variables and then update.

@@ -141,6 +141,7 @@ public:
 
 void example_single_factor(int room_types, int property_types,
                            int n_known_rooms) {
+  int C = 1;
   GraphInformation gi;
   set<string> known_rooms;
 
@@ -148,35 +149,32 @@ void example_single_factor(int room_types, int property_types,
     known_rooms.insert(room(i));
 
   // Real room variable.
-  const VariableType *type_real_room = NULL;
+  VariableType *type_real_room = gi.Add("real_room", VariableType());
   {
-    VariableType type;
+    VariableType &type = *type_real_room;
     for (int i = 0; i < room_types; ++i)
       type.values.push_back(room(i));
     sort(type.values.begin(), type.values.end());
-    type_real_room = gi.Add("real_room", type);
   }
 
   // Real property variable.
-  const VariableType *type_real_prop = NULL;
+  VariableType *type_real_prop = gi.Add("real_prop", VariableType());
   {
-    VariableType type;
+    VariableType &type = *type_real_prop;
     for (int i = 0; i < room_types; ++i)
       type.values.push_back(prop(i));
     sort(type.values.begin(), type.values.end());
-    type_real_prop = gi.Add("real_prop", type);
   }
   
   // Real factor(room, property).
-  const FactorData *factor_rroom_rprop = NULL;
+  FactorData *factor_rroom_rprop = gi.Add("f(room, property)", FactorData());
   {
-    FactorData factor;
+    FactorData &factor = *factor_rroom_rprop;
     factor.variables.push_back(make_pair("room", "real_room"));
     factor.variables.push_back(make_pair("prop", "real_prop"));
     BOOST_FOREACH(const string &room, type_real_room->values)
     BOOST_FOREACH(const string &prop, type_real_prop->values)
       factor.potential[vector_of(room, prop)] = drand48();
-    factor_rroom_rprop = gi.Add("f(real_room, real_prop)", factor);
   }
 
   assert(type_real_room);
@@ -198,31 +196,18 @@ void example_single_factor(int room_types, int property_types,
     else conditional_samples[i].clear();
   }
 
-  // Generate unconditional samples.
-  vector< vector<string> > unconditional_samples(1000);
-  BOOST_FOREACH(vector<string> &sample, unconditional_samples) {
-    GraphStructure s;
-    int n_properties = 1 + (random() % 15);
-    s.createRandom(1, n_properties, 0);
[…]

### 2011-06-13 thesis/synthetic/main.cc
Splitted code in functions

@@ -37,8 +37,7 @@ using namespace boost;
 #include <dai/properties.h>
 using namespace dai;
 
-void example_single_factor(int room_types = 10, int property_types = 15,
-                           int known_rooms = 5);
+void example_single_factor();
 
 int main() {
   example_single_factor();

### 2011-06-13 thesis/synthetic/single-factor.cc

@@ -139,78 +139,85 @@ public:
 
 }
 
-void example_single_factor(int room_types, int property_types,
-                           int n_known_rooms) {
-  int C = 1;
-  GraphInformation gi;
-  set<string> known_rooms;
+int room_types = 10;
+int n_known_rooms = 5;
+int property_types = 8;
+int C = 1;
 
+set<string> known_rooms;
+
+
+GraphInformation gi;
+VariableType *type_real_room = NULL;
+VariableType *type_real_prop = NULL;
+FactorData *factor_rroom_rprop = NULL;
+
+VariableType *type_known_room = NULL;
+FactorData *factor_kroom_kprop = NULL;
+
+VariableType *type_any_room = NULL;
+FactorData *factor_aroom_kprop = NULL;
+
+void generate_real_distribution_vars_and_factors() {
   for (size_t i = 0; i < n_known_rooms; ++i)
     known_rooms.insert(room(i));
 
   // Real room variable.
-  VariableType *type_real_room = gi.Add("real_room", VariableType());
   {
-    VariableType &type = *type_real_room;
+    VariableType &type = *(type_real_room = gi.Add("real_room", VariableType()));
     for (int i = 0; i < room_types; ++i)
       type.values.push_back(room(i));
     sort(type.values.begin(), type.values.end());
   }
 
   // Real property variable.
-  VariableType *type_real_prop = gi.Add("real_prop", VariableType());
   {
-    VariableType &type = *type_real_prop;
+    VariableType &type = *(type_real_prop = gi.Add("real_prop", VariableType()));
     for (int i = 0; i < room_types; ++i)
       type.values.push_back(prop(i));
     sort(type.values.begin(), type.values.end());
   }
   
   // Real factor(room, property).
-  FactorData *factor_rroom_rprop = gi.Add("f(room, property)", FactorData());
   {
-    FactorData &factor = *factor_rroom_rprop;
+    FactorData &factor = *(factor_rroom_rprop = gi.Add("f(room, property)", FactorData()));
     factor.variables.push_back(make_pair("room", "real_room"));
     factor.variables.push_back(make_pair("prop", "real_prop"));
     BOOST_FOREACH(const string &room, type_real_room->values)
     BOOST_FOREACH(const string &prop, type_real_prop->values)
       factor.potential[vector_of(room, prop)] = drand48();
   }
+}
 
-  assert(type_real_room);
-  assert(type_real_prop);
-  assert(factor_rroom_rprop);
[…]

### 2011-06-13 thesis/synthetic/Query.cc
Single experiment coded finished.

@@ -172,6 +172,33 @@ class QueryInternal {
 	    BOOST_FOREACH(const Variable *v, vars)
           output->push_back(v->type_->values[state[ConvertToDai(v).label()]]);
     }
+    
+    double LogZ(const vector<const Variable*> &clamp_vars,
+                const vector<string> &clamp_state) {
+	    if (factorGraph_.get() == NULL)
+		    MakeFactorGraph();
+
+		vector<dai::Var> dai_vars = ConvertToDai(clamp_vars);
+		dai::VarSet dai_varset(
+			dai::SmallSet<dai::Var>(dai_vars.begin(), dai_vars.end(),
+			                        dai_vars.size()));
+
+
+		bp_.reset(new dai::BP(*factorGraph_,
+				daiOptions_("updates", string("SEQRND"))
+				           ("logdomain", false)));
+		bp_->init();
+        for (size_t i = 0; i < clamp_vars.size(); ++i) {
+            const VariableType *type = clamp_vars[i]->type_;
+            size_t index = lower_bound(type->values.begin(), type->values.end(),
+                                       clamp_state[i]) - type->values.begin();
+            assert(index < type->values.size());
+            assert(type->values[index] == clamp_state[i]);
+            bp_->clamp(dai_vars[i].label(), index);
+        }
+		bp_->run();
+        return bp_->logZ();
+    }
 
  protected:
 	dai::Var ConvertToDai(const Variable *var)
@@ -233,6 +260,14 @@ void Query::Marginalize(const vector<const Variable*> &variables,
 		BuildInternal();
 
 	return internal_->CalcMarginal(variables, output);
+}
+
+double Query::LogZ(const vector<const Variable*> &clamp_vars,
+                   const vector<string> &clamp_state)
+{
+    if (!internal_.get())
+        BuildInternal();
+    return internal_->LogZ(clamp_vars, clamp_state);
 }
 
 void Query::Sample(const vector<const Variable*> &variables,

### 2011-06-13 thesis/synthetic/Query.h

@@ -55,6 +55,9 @@ class Query {
     
     void Sample(const std::vector<const Variable*> &variables,
                 std::vector<std::string> *output);
+
+    double LogZ(const std::vector<const Variable*> &clamp_vars,
+                const std::vector<std::string> &clamp_state);
  protected:
 	const Graph *graph_;
 	boost::scoped_ptr<QueryInternal> internal_;

### 2011-06-13 thesis/synthetic/single-factor.cc

@@ -27,6 +27,7 @@
 #include <set>
 #include <iostream>
 #include <algorithm>
+#include <cmath>
 using namespace std;
 
 #include <boost/foreach.hpp>
@@ -139,9 +140,9 @@ public:
 
 }
 
-int room_types = 10;
-int n_known_rooms = 5;
-int property_types = 8;
+size_t room_types = 10;
+size_t n_known_rooms = 5;
+size_t property_types = 15;
 int C = 1;
 
 set<string> known_rooms;
@@ -246,6 +247,7 @@ void generate_unconditional_samples(vector<vector<string> > *output) {
 
 void learn_any_room_distribution(const vector<vector<string> > &samples) {
   // Create variable type for any room states.
+  type_any_room = gi.Add("any_room", VariableType());
   type_any_room->values.push_back("any");
 
   // Learn factor(room, property | room is any room) form unlabelled data.
@@ -259,20 +261,99 @@ void learn_any_room_distribution(const vector<vector<string> > &samples) {
 
     BOOST_FOREACH(const string &prop, type_real_prop->values)
       factor.potential[vector_of("any", prop)] += C;
+    
+    map<vector<string>,double>::iterator iter;
+    for (iter = factor.potential.begin(); iter != factor.potential.end(); ++iter) {
+      iter->second *= exp(-4.86222);
+      iter->second *= exp(0.399);
+    }
   }
+
 }
 
 void compare_real_normalization_factors() {
-  int n = 100;
-  while (n--) {
+  for (size_t i = 1; i < 100; ++i) {
     GraphStructure s;
-    int n_properties = 1 + (random() % 15);
+    int n_properties = i; //1 + (random() % 15);
     s.createRandom(1, n_properties, 0);
     
     MGraph real_g(s, type_real_room, type_real_prop, NULL, factor_rroom_rprop);
     MGraph any_g (s, type_any_room,  type_real_prop, NULL, factor_aroom_kprop);
-   
+
+    Query rq(&real_g);
+    Query aq(&any_g);
+
+    // Absolute difference:
+    vector<const Variable*> var;
+    vector<string> var_clamp;
+
+    double abs_rlogz = rq.LogZ(var, var_clamp);
+    double abs_alogz = aq.LogZ(var, var_clamp);
+    cout << n_properties << ": total_abs_diff = " << exp(abs_rlogz - abs_alogz) << endl;
[…]

### 2011-06-13 thesis/synthetic/plot.tex
Plotting code.

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{plotmarks}
\begin{document}

\section{Constant $P(novel)$}

\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    	\draw (0,0) -- coordinate (y axis mid) (0,30);
    	%ticks
    	\foreach \x in {0,...,20}
     		\draw (\x,1pt) -- (\x,-3pt)
			node[anchor=north] {\x};
    	\foreach \y in {0,5,...,30}
     		\draw (1pt,\y) -- (-3pt,\y) 
     			node[anchor=east] {\y}; 
	%labels      
	\node[below=0.8cm] at (x axis mid) {Graph Size};
	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};

	%plots
	\draw plot[mark=*, mark options={fill=white}] 
		file {result_fix_threshold_known.data};
	\draw plot[mark=triangle*, mark options={fill=black}] 
		file {result_fix_threshold_novel.data};

\end{tikzpicture}

\clearpage
\section{Dynamic $P(novel)$}

\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
 	%axis
	\draw (0,0) -- coordinate (x axis mid) (20,0);
    	\draw (0,0) -- coordinate (y axis mid) (0,30);
    	%ticks
    	\foreach \x in {0,...,20}
     		\draw (\x,1pt) -- (\x,-3pt)
			node[anchor=north] {\x};
    	\foreach \y in {0,5,...,30}
     		\draw (1pt,\y) -- (-3pt,\y) 
     			node[anchor=east] {\y}; 
	%labels      
	\node[below=0.8cm] at (x axis mid) {Graph Size};
	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};

	%plots
	\draw plot[mark=*, mark options={fill=white}] 
		file {result_dyn_threshold_known.data};
	\draw plot[mark=triangle*, mark options={fill=black}] 
		file {result_dyn_threshold_novel.data};

\end{tikzpicture}

\clearpage
\end{document}

### 2011-06-13 thesis/synthetic/single-factor.cc

@@ -26,6 +26,7 @@
 #include <cstdio>
 #include <set>
 #include <iostream>
+#include <fstream>
 #include <algorithm>
 #include <cmath>
 using namespace std;
@@ -291,28 +292,6 @@ void compare_real_normalization_factors() {
     double abs_alogz = aq.LogZ(var, var_clamp);
     cout << n_properties << ": total_abs_diff = " << exp(abs_rlogz - abs_alogz) << endl;
     cout << n_properties << ": total_rel_diff = " << (abs_rlogz - abs_alogz)/n_properties << endl;
-
-    // Sensed difference:
-    vector<double> log_diff;
-    for (int i = 0; i < 0; ++i) {
-      vector<const Variable*> r_prop(real_g.vars.begin()+1, real_g.vars.end());
-      vector<const Variable*> a_prop(any_g.vars.begin()+1,  any_g.vars.end());
-
-      vector<string> fsample;
-      rq.Sample(real_g.vars, &fsample);
-      vector<string> sample(fsample.begin()+1, fsample.end());
-
-      double rlogz = rq.LogZ(r_prop, sample);
-      double alogz = aq.LogZ(a_prop, sample);
-      log_diff.push_back(rlogz - alogz);
-      // cout << n_properties << ": sense: " << fsample[0] << endl;
-      // cout << n_properties << ": sense_abs_diff = " << (rlogz - alogz) << endl;
-      // cout << n_properties << ": sense_rel_diff = " << (rlogz - alogz)/n_properties << endl;
-    }
-    sort(log_diff.begin(), log_diff.end());
-    cout << n_properties << ":";
-    BOOST_FOREACH(const double a, log_diff) cout << " " << a;
-    cout << endl;
   }
 }
 
@@ -340,8 +319,8 @@ void compare_performance(const vector<vector<string> > &samples) {
     
     // log(2) + log(\phi_k(x)) < k_i log(s_i) + log(\phi_a(x))
     // log(s_i) > (log(2) + log(\phi_k(x)) - log(\phi_a(x))) / k_i
-    double dyn_threshold = (log(2) + klogZ - alogZ)/n_properties;
-    result_dyn.push_back(make_pair(make_pair(n_properties, dyn_threshold), known_rooms.count(sample[0]) == 1));
+    double dyn_threshold = (alogZ - klogZ);
[…]

### 2011-06-13 thesis/synthetic/Query.cc
More tires with plot... its now clear a static threshold is not correct.
Still theres something wrong with the other one.
More work needed to understand how to control the s_i and threshold parameters.

@@ -150,7 +150,7 @@ class QueryInternal {
 
             dai::PropertySet gibbsProps;
             gibbsProps.set("maxiter", size_t(100));   // number of Gibbs sampler iterations
-            gibbsProps.set("burnin", size_t(100));
+            gibbsProps.set("burnin", size_t(0));
             gibbsProps.set("verbose", size_t(0));
 	
             //rocsDebug3("Creating GibbsSampler...");

### 2011-06-13 thesis/synthetic/single-factor.cc

@@ -142,7 +142,7 @@ public:
 }
 
 size_t room_types = 10;
-size_t n_known_rooms = 5;
+size_t n_known_rooms = 3;
 size_t property_types = 15;
 int C = 1;
 
@@ -237,7 +237,7 @@ void learn_known_room_distribution(const vector<vector<string> > samples) {
 void generate_unconditional_samples(vector<vector<string> > *output) {
   BOOST_FOREACH(vector<string> &sample, *output) {
     GraphStructure s;
-    int n_properties = 1 + (random() % 15);
+    int n_properties = 1 + (random() % 17);
     s.createRandom(1, n_properties, 0);
     MGraph g(s, type_real_room, type_real_prop, NULL, factor_rroom_rprop);
    
@@ -265,8 +265,8 @@ void learn_any_room_distribution(const vector<vector<string> > &samples) {
     
     map<vector<string>,double>::iterator iter;
     for (iter = factor.potential.begin(); iter != factor.potential.end(); ++iter) {
-      iter->second *= exp(-4.86222);
-      iter->second *= exp(0.399);
+//      iter->second *= exp(-4.86222);
+//      iter->second *= exp(0.399);
     }
   }
 
@@ -295,6 +295,8 @@ void compare_real_normalization_factors() {
   }
 }
 
+double junk() { return (drand48()-0.5)/3; }
+
 void compare_performance(const vector<vector<string> > &samples) {
   vector<pair<pair<int,double>,bool> > result_fix;
   vector<pair<pair<int,double>,bool> > result_dyn;
@@ -319,7 +321,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     
     // log(2) + log(\phi_k(x)) < k_i log(s_i) + log(\phi_a(x))
     // log(s_i) > (log(2) + log(\phi_k(x)) - log(\phi_a(x))) / k_i
-    double dyn_threshold = (alogZ - klogZ);
+    double dyn_threshold = (klogZ - alogZ)/n_properties;
     result_dyn.push_back(make_pair(make_pair(n_properties, 10*exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
 
     // Assuming a constant probability of drawing a novel sample.
@@ -329,6 +331,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     double t_alogZ = aq.LogZ(no_prop, no_clamp);
     double fix_threshold = (klogZ - t_klogZ) - (alogZ - t_alogZ);
[…]

### 2011-06-14 thesis/synthetic/single-factor.cc
Some improves on drawing

@@ -234,10 +234,13 @@ void learn_known_room_distribution(const vector<vector<string> > samples) {
   }
 }
 
-void generate_unconditional_samples(vector<vector<string> > *output) {
+void generate_unconditional_samples(vector<vector<string> > *output, int size = -1) {
   BOOST_FOREACH(vector<string> &sample, *output) {
     GraphStructure s;
-    int n_properties = 1 + (random() % 17);
+    int n_properties = size;
+    if (n_properties == -1)
+      n_properties = 1 + (random() % 17);
+
     s.createRandom(1, n_properties, 0);
     MGraph g(s, type_real_room, type_real_prop, NULL, factor_rroom_rprop);
    
@@ -322,7 +325,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     // log(2) + log(\phi_k(x)) < k_i log(s_i) + log(\phi_a(x))
     // log(s_i) > (log(2) + log(\phi_k(x)) - log(\phi_a(x))) / k_i
     double dyn_threshold = (klogZ - alogZ)/n_properties;
-    result_dyn.push_back(make_pair(make_pair(n_properties, 10*exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
+    //result_dyn.push_back(make_pair(make_pair(n_properties, 10*exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
 
     // Assuming a constant probability of drawing a novel sample.
     vector<const Variable*> no_prop;
@@ -331,7 +334,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     double t_alogZ = aq.LogZ(no_prop, no_clamp);
     double fix_threshold = (klogZ - t_klogZ) - (alogZ - t_alogZ);
     result_fix.push_back(make_pair(make_pair(n_properties, 10*exp(fix_threshold)), known_rooms.count(sample[0]) == 1));
-    //result_dyn.push_back(make_pair(make_pair(n_properties, 10*exp(fix_threshold/n_properties)), known_rooms.count(sample[0]) == 1));
+    result_dyn.push_back(make_pair(make_pair(n_properties, exp(fix_threshold/n_properties)), known_rooms.count(sample[0]) == 1));
   }
   sort(result_dyn.begin(), result_dyn.end());
   sort(result_fix.begin(), result_fix.end());
@@ -341,9 +344,9 @@ void compare_performance(const vector<vector<string> > &samples) {
[…]

### 2011-06-14 thesis/synthetic/plot.tex
Some improves on drawing
Changed to pgfPlots package.

@@
 \usetikzlibrary{plotmarks}
+\usepackage{pgfplots}
+
 \begin{document}
@@
 \section{Constant $P(novel)$}
-
-\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
- 	%axis
-	\draw (0,0) -- coordinate (x axis mid) (20,0);
-    	\draw (0,0) -- coordinate (y axis mid) (0,30);
-    	%ticks
-    	\foreach \x in {0,...,20}
-     		\draw (\x,1pt) -- (\x,-3pt)
-			node[anchor=north] {\x};
-    	\foreach \y in {0,5,...,30}
-     		\draw (1pt,\y) -- (-3pt,\y) 
-     			node[anchor=east] {\y}; 
-	%labels      
-	\node[below=0.8cm] at (x axis mid) {Graph Size};
-	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};
-
-	%plots
-	\draw plot[mark=*, mark options={fill=white}] 
-		file {result_fix_threshold_known.data};
-	\draw plot[mark=triangle*, mark options={fill=black}] 
-		file {result_fix_threshold_novel.data};
-
+\begin{tikzpicture}[only marks,mark size=0.5pt]
+\begin{axis}[
+ xlabel={Threshold},
+ ylabel={Graph Size},
+ width=\textwidth,
+ height=1.5\textwidth]
+\addplot[mark=+,mark options={fill=white}] file {result_fix_threshold_known.data};
+\addplot[mark=-,mark options={fill=black}] file {result_fix_threshold_novel.data};
+\end{axis}
 \end{tikzpicture}
 
-\clearpage
 \section{Dynamic $P(novel)$}
-
-\begin{tikzpicture}[only marks,y=.1cm, x=.7cm]
- 	%axis
-	\draw (0,0) -- coordinate (x axis mid) (20,0);
-    	\draw (0,0) -- coordinate (y axis mid) (0,30);
-    	%ticks
-    	\foreach \x in {0,...,20}
-     		\draw (\x,1pt) -- (\x,-3pt)
-			node[anchor=north] {\x};
-    	\foreach \y in {0,5,...,30}
-     		\draw (1pt,\y) -- (-3pt,\y) 
-     			node[anchor=east] {\y}; 
-	%labels      
-	\node[below=0.8cm] at (x axis mid) {Graph Size};
-	\node[rotate=90, above=0.8cm] at (y axis mid) {Threshold function};
-
-	%plots
-	\draw plot[mark=*, mark options={fill=white}] 
-		file {result_dyn_threshold_known.data};
-	\draw plot[mark=triangle*, mark options={fill=black}] 
-		file {result_dyn_threshold_novel.data};
-
+\begin{tikzpicture}[only marks,mark size=0.5pt]
+\begin{axis}[
+ xlabel={Threshold},
+ ylabel={Graph Size},
+ width=\textwidth,
+ height=1.5\textwidth]
+\addplot[mark=+,mark options={fill=white}] file {result_dyn_threshold_known.data};
+\addplot[mark=-,mark options={fill=black}] file {result_dyn_threshold_novel.data};
+\end{axis}
 \end{tikzpicture}
 
-\clearpage
 \end{document}

### 2011-06-14 thesis/synthetic/single-factor.cc

@@ -298,7 +298,7 @@ void compare_real_normalization_factors() {
   }
 }
 
-double junk() { return (drand48()-0.5)/3; }
+double junk() { return 0.1 + drand48()/5; }
 
 void compare_performance(const vector<vector<string> > &samples) {
   vector<pair<pair<int,double>,bool> > result_fix;
@@ -325,7 +325,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     // log(2) + log(\phi_k(x)) < k_i log(s_i) + log(\phi_a(x))
     // log(s_i) > (log(2) + log(\phi_k(x)) - log(\phi_a(x))) / k_i
     double dyn_threshold = (klogZ - alogZ)/n_properties;
-    //result_dyn.push_back(make_pair(make_pair(n_properties, 10*exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
+    result_dyn.push_back(make_pair(make_pair(n_properties, exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
 
     // Assuming a constant probability of drawing a novel sample.
     vector<const Variable*> no_prop;
@@ -333,8 +333,8 @@ void compare_performance(const vector<vector<string> > &samples) {
     double t_klogZ = kq.LogZ(no_prop, no_clamp);
     double t_alogZ = aq.LogZ(no_prop, no_clamp);
     double fix_threshold = (klogZ - t_klogZ) - (alogZ - t_alogZ);
-    result_fix.push_back(make_pair(make_pair(n_properties, 10*exp(fix_threshold)), known_rooms.count(sample[0]) == 1));
-    result_dyn.push_back(make_pair(make_pair(n_properties, exp(fix_threshold/n_properties)), known_rooms.count(sample[0]) == 1));
+    result_fix.push_back(make_pair(make_pair(n_properties, exp(fix_threshold)), known_rooms.count(sample[0]) == 1));
+    //result_dyn.push_back(make_pair(make_pair(n_properties, fix_threshold/n_properties), known_rooms.count(sample[0]) == 1));
   }
   sort(result_dyn.begin(), result_dyn.end());
   sort(result_fix.begin(), result_fix.end());
@@ -344,9 +344,9 @@ void compare_performance(const vector<vector<string> > &samples) {
     ofstream os_novel("result_dyn_threshold_novel.data");
     for (size_t i = 0; i < result_dyn.size(); ++i)
       if (result_dyn[i].second)
[…]

### 2011-06-14 thesis/semantic-mapping.tex
Fixes here and there.
Initial fixes of semantic mapping chapter.
"Spatial Knowledge Representation" is now clear one concept.
And dora specific implementation another.

@@
 
-There has been several efforts in the areas of artificial intelligence and robotics
-in creating robots that are able to interact with humans and their environments.
-One of the important aspects is to endow those robots with a deeper understanding
-of human environments, not just for the purpose of navigation and obstacle avoidance, 
-but also in terms of human semantics and functionality.
-
-In that sense the need for semantic mapping arises, and it can be described as the
-relation between spatial areas and a set of existent concepts used by humans to
-describe them.
+% motivation
 Humans associate concepts to areas that relate to the functionality of those. That
 association is either useful by explaining where to find certain objects as well to
-understand the types of activities to perform on it.
-By being able to reason over those semantics an agent can be expected to better
-understand human environments and increase its efficiency when performing tasks on them.
+understand the types of activities to perform on it. In a
+
+Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
+knowledge about them it is expected to improve the agent 
+% semantic mapping definition
+In this context semantic mapping is seen as the process of building a representation of
+the environment which associates spatial entities with a set of defined spatial concepts (i.e.\ human concepts).
+
+% importance of conceptual knowledge
+% application of conceptual knowledge
+It is expected that by being able to reason over those semantics an agent can better
+understand human environments and increase its efficiency when performing human-tasks on them.
+By handling this conceptual knowledge the agent becomes aware of
+
+% challenges and need for representing uncertainty
+The goal of a spatial representation is to provide the agent with a view about the environment
+that allows it to reason beyond its sensory horizon.
+
+ Such a representation is by its nature
+imperfect, incomplete, inaccurate and invalid.
+
+world that allows the agent to perform reasoning over it and efficiently perform tasks.
+It should also give the agent ability to reason about parts 
+Such a representation cannot be perfect since the methods and information are uncertain
+
+% here a here a method is used that does semantic mappping and representes conceptual knowledge
+% on a probabilistic fashion.
 
@@
 on the semantic knowledge of a mobile agent. For that the semantic mapping system
-proposed by \cite{pronobis2011phd} was used as a base.
+implemented on \gls{Dora} proposed by \cite{pronobis2011phd} was used as a base.
 
-\section{Dora Architecture Overview}
-Pronobis presents a system architecture working on indoor environments using
-non\hyp{}omnidirectional lasers and visual sensors, that by integrating cues such as geometry,
-object presence and appearances is able to perform inference across any semantic property,
-for example to handle place categorization.
-The introduced semantic representation has not only been tested on real scenarios
-and performance measured across several conditions~\cite{pronobis2007iros} but also shown
+\section{Architecture Overview}
+Pronobis~\cite{pronobis2011phd} presents a system architecture working in indoor environments using
+non\hyp{}omnidirectional laser and visual sensors for semantic mapping. The system integrates cues such as geometry,
+object presence and appearances and is able to perform inference across any semantic properties,
+for example for the purpose of place categorization.
+The introduced system is built around a spatial knowledge representation~\cite{pronobis2010ias} and it has not only been tested on real-scenarios
+and performance tested across several conditions~\cite{pronobis2007iros} but also shown
 to be tailored to effectively solve tasks arising on mobile robotics~\cite{hanheide2011ijcai}.
-Besides that the system architecture has been drawn with an high-focus on probabilistic
+
+Besides that the spatial knowledge representation has been drawn with an high-focus on probabilistic
 and uncertain reasoning, human interaction and life-long learning capabilities.
-
 For that it was considered an excellent base defining not only the concepts and
@@
 between all the layers involved on the agent.
+
+Being able to perform probabilistic inference on all available information Such a representation is suitable to detect novel concepts.
+
 The system is implemented on \Gls{Dora} and a brief overview of its architecture is
@@
 
-\subsection{System organization}
-The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
-sub-systems, all of which actively use or maintain the spatial knowledge representation
-(see \autoref{fig:dora-architecture}):
+\section{Spatial Knowledge Representation}
+\cite{pronobis2010ias} proposes a spatial knowledge representation sub-divided into four layers.
+Each layer focus on different aspects of the world, abstractions levels of the spatial knowledge
+and different spatial scales. Each uses different spatial entities and relate to the agent
+goals in different ways:
+
+In the lowest abstraction level the sensory layer represents the most immediate and short term
+accurate representation of the world. Above, the place layer discretizes the continuous
+space into. The categorical layer focus on knowledge of low\hyp{}level, long\hyp{}term
+categorical models of the sensory information. And at the higher level the conceptual map
+associates concepts to areas such as rooms (see \autoref{fig:spatial-knowledge}).
 
@@
 \includegraphics[width=0.7\textwidth]{figures/dora-architecture.pdf}
-\caption{\label{fig:dora-architecture}Interaction of the sub-systems
-         in Dora with special focus on the conceptual layer.}
+\caption{\label{fig:spatial-knowledge}Layered structure of the spatial representation.
+         The conceptual layer illustrates part of the ontology representing both instance
+         and pre-defined world knowledge.}
 \end{figure}
@@
 \subsubsection*{Sensory Layer}
-The sensory layer is responsible for all the low-level sensors.
-It maintains an immediate and exact description of the environment by using
-data such as images from cameras, 2D\hyp{}laser scans and odometry from encoders in the
-movement motors. The information provided by it is used on the higher layers to extract
-tractable information.
+The sensory layer keeps a detailed representation of the immediate surrounds of the agent.
+That representation is based on directly sensed inputs as well on data-fusion for short
+time\hyp{}intervals.
+
+It is responsible for providing the agent with an exact position and to accurately position
+low\hyp{}features and landmarks. The information is stored with associated uncertainty it
+is highly susceptible to be replaced with newer versions. Also old or distant information
+is forgotten.
+By tracking the surrounds for short amount of times it provides an accurate representation
+that can be used to locate and guide the agent on low\hyp{}level movements.
 
 \subsubsection*{Place Layer}
-The place layer is responsible for discretizing the continuous low-level space into
-a finite number of places by spreading several virtual places markers around the whole
-environment and maintaining connectivity information between them.
-This virtual places also handle information associating the robot view point. Together
-with extracted properties they allow the system to perform a clever integration over the
-room properties, avoiding sampling bias, for example coming due to the robot spend too
-much time in a part of the room~\cite{pronobis2010ijrr}.
+The place layer is responsible for maintaining a bottom\hyp{}up discretization of the
+continuous space. On it the world is represented as a collection of discrete spatial
+entities, called places, together with connectivity information.
+
+Each place is defined with the features from the sensory layer and with spatial relations
+to other places. Connectivity here, stands for ability to travel directly between the
+places. Additionally places can be created for areas not yet explored allowing to introduce
+virtual place holders to places that would eventually be uncover if the agent moved there.
+
+The places are also considered stable on long\hyp{}term and can be used to help the agent
+performing localization and planning longer distance motion planning. This higher level
+representation although uncertain allows to model connectivity and relaxed spatial relations.
 
 \subsubsection*{Categorical Layer}
-The categorical layer is responsible for extracting high-level information from the
-sensors. It has classifiers trained to detect several properties from the visual and laser
-features captured on the sensory layer.
-It is responsible for detecting objects, doors, categorize the geometry and the appearance.
-The used methods and properties extracted by this layer are described in more detail in
-\autoref{sec:dora-features}.
+The categorical layer holds the generic long-term, low-level representations of categorical models
+of the agent's sensory information.
+The knowledge represented by this layer is not instance specific. It is the knowledge
+required to determine certain features indicate the presence of an entity of certain
+category. E.g.\ the knowledge required to detect objects, landmarks or describe room appearances.
+Other properties may as well be defined for example shape, color, edges and size properties
+are all defined in terms of low\hyp{}level features.
+
+This layer is responsible for representing knowledge on how to extract properties used
+by the conceptual layer.
+Although not mandatory, in many cases the categories represented by it will map to human concepts
+and not to the internal concepts of the agent.
+The type of properties handled by this layer may require incredible complex models and for
+that they might be trained on a supervised fashion.
 
 \subsubsection*{Conceptual Layer}
-The conceptual layer merges the sensed properties and places structure together
-with general knowledge under a unified probabilistic semantic representation: the
-\emph{conceptual map}.
-This conceptual map is the structure for handling the high-level semantic information
-and performing inferences. It is the layer where the novelty detection methods described
-on this thesis are meant to be integrated and is described in more detail in
-\autoref{sec:conceptual-map}.
+The conceptual layer provides an ontology that represents a taxonomy of the spatial concepts
+and their relations with the properties represented on the categorical layer.
+This associates semantic meaning to those properties that is useful for human-agent iteration
+and provides relations that can be verbalized and explained with the ontology.
+
+It provides also knowledge between the semantic concepts and their instances of those concepts.
+Including definitions of spatial concepts related to space segmentation as well to semantic
+categorization of those spatial entities.
+E.g.\ rooms are commonly split by doors, rooms exist in a floor, a building has floors, milk is
+likely to be found on kitchens, etc\dots
+
+By providing information for segmentation, semantic mapping to concepts and relations between those
+concepts, the knowledge represented on this layer allows the agent to explain and reason close to
+the human concepts. The knowledge represented here is considered to be valid through very long
+periods of time or even life\hyp{}long.
+
+\section{System Structure}
+The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
+sub-systems, all of which actively use or maintain the spatial knowledge representation
+(see \autoref{fig:dora-architecture}).
 
@@
 \label{sec:dora-features}
+The \emph{categorical layer} requires classifiers with the ability to categorize
 
@@
 \label{sec:conceptual-map}
-As \gls{Dora} moves through the environment its \emph{conceptual layer} builds a structural and
+As \gls{Dora} moves through the environment it builds a \emph{conceptual map}: a structural and
 probabilistic representation of the space instantiated as a \emph{graphical model}.
@@
 \includegraphics[width=0.75\textwidth]{figures/chaingraph.pdf}
-\caption{\label{fig:chain-graph}Example chain-graph produced by the \emph{conceptual layer}.}
+\caption{\label{fig:chain-graph}Example chain-graph instantiated from the \emph{conceptual layer}.}
 \end{figure}
@@
 Factor graphs have been introduced in \autoref{sec:graphical-models} and are used
-throughout this thesis as they provide a easier manipulation due to factorization.
+throughout this thesis as they provide an easier manipulation due to factorization.
 
-Moreover, there exists efficient implementation of inference engines operating on factor
+Moreover, there exist efficient implementations of inference engines operating on factor
 graph representations~\cite{Mooij_libDAI_10}.
-Describing the distribution function in terms of graphs allows to use those engines to
+Describing the distribution function in terms of graphs permits the use of those engines to
 efficently calculate marginals on any given subset of variables by exploiting conditional
@@
 \label{sec:cues-from-low-level}
-In a realistic scenario classification tasks are by their nature uncertain.
+In a realistic scenario, classification tasks are by their nature uncertain.
 And although a trained classifier is able to decide which class is more likely, there is
-interest in handling uncertainty produced by those to turn decisions more robust.
+interest in handling uncertainty produced by those to make decisions more robust.
 
 By modelling the classifiers as an implicit function $\phi_{classifier}(c, x)$ it becomes
-possible to model the uncertainty of those given the presence of the sensed low-level
-features $X$. The presence of those implicit models no longer allow to calculate the
+possible to model the uncertainty of the decision given the presence of the sensed low-level
+features $x$. The presence of those implicit models no longer allow calculation of the
 normalization factor over the whole graph as there is no explicit model for either the
@@
 are observed. For that instead of modelling the variables $X$ and explicitly describing
-the factor $\phi_{classifier}$ it is enough to replace it by an observed factor $\phi_{classifier}(c|x)$.
+the factor $\phi_{classifier}$ it is enough to replace it by an observed factor
+$\phi_{classifier}(c|x)$ as shown on \autoref{fig:implicit-factor}.
 

### 2011-06-14 thesis/semantic-mapping.tex
More changes on the semantic mapping chapter... only introduction section missing now.

@@
 \centering
-\includegraphics[width=0.7\textwidth]{figures/dora-architecture.pdf}
-\caption{\label{fig:spatial-knowledge}Layered structure of the spatial representation.
-         The conceptual layer illustrates part of the ontology representing both instance
-         and pre-defined world knowledge.}
+\includegraphics[width=0.7\textwidth]{figures/spatial-knowledge-representation.pdf}
+\caption{\label{fig:spatial-knowledge}Layered structure of the spatial representation.}
 \end{figure}
 
-\subsubsection*{Sensory Layer}
+\subsection{Sensory Layer}
 The sensory layer keeps a detailed representation of the immediate surrounds of the agent.
@@
 
-\subsubsection*{Place Layer}
+\subsection{Place Layer}
 The place layer is responsible for maintaining a bottom\hyp{}up discretization of the
@@
 
-\subsubsection*{Categorical Layer}
+\subsection{Categorical Layer}
 The categorical layer holds the generic long-term, low-level representations of categorical models
@@
 
-\subsubsection*{Conceptual Layer}
+\subsection{Conceptual Layer}
 The conceptual layer provides an ontology that represents a taxonomy of the spatial concepts
@@
 The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
-sub-systems, all of which actively use or maintain the spatial knowledge representation
-(see \autoref{fig:dora-architecture}).
+sub-systems~(see \autoref{fig:dora-architecture}),
+all of which actively use or maintain the \emph{spatial knowledge representation}.
+
+The layer structure of the spatial knowledge representation allow to implement
+data driven processes that control the flow of knowledge and information between the
+lower and higher\hyp{}levels layers. To reduce computation and make it feasible not
+all updates from lower\hyp{}levels are treated immediately. Instead the system waits
+until a substantial change has occurred.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.8\textwidth]{figures/dora-architecture.pdf}
+\caption{\label{fig:dora-architecture}Dora the Explorer as well as the elements and the data flow inside
+         the semantic mapping system.}
+\end{figure}
+
+Initially the low-level topology and metric mapping systems use a \gls{SLAM} algorithm~\cite{Folkesson07a}
+that is used as representation on the sensory layer. Those two systems are also responsible
+for creating the place map on the place layer, as that is highly related with the \gls{SLAM} goals.
+By using the landmarks (doors) and other available information the system also tries to
+group places and understand the concept of rooms.
+
+At the same time the spatial property categorization system is responsible for detecting
+instances of the categorical knowledge by running classifiers previously trained to detect
+a set of features listed on \autoref{sec:dora-features}. For that it uses the data available
+from a laser range finder and a camera and besides classifying the features it estimates also
+its confidence on the information.
+The extracted features together with the estimates is then integrated over time and space
+in order to increase its robustness.
+
+Finally the conceptual map system utilizes the acquired instance knowledge together with
+the high\hyp{}level conceptual knowledge to create a probabilistic graphical model that allows
+to run inferences.
+Since this conceptual mapping processing is responsible for creating the probabilistic
+graphical model it is described in detail in \autoref{sec:conceptual-map}.
 
@@
 \label{sec:dora-features}
-The \emph{categorical layer} requires classifiers with the ability to categorize
+The \emph{categorical layer} requires classifiers with the ability to categorize the
+low\hyp{}level features into categories.
+For that \gls{Dora} performs detection on the following categorical entities:
+\begin{description}
+\item[Doorway Detection] by using the laser range finder the system tries to detect doors.
+\item[Geometric Shape] by using features extracted from the laser data and pre-trained classifiers
+the system tries to classify rooms according to their shapes (e.g.\ rectangular, square).
+\item[Geometric Size] is classified once again recurring to laser data. In this case the system
+\item[Object Detection] is performed by running traditional object recognition classifiers.
+Those are previously trained on a supervised fashion to later be able to detect objects
+based on their visual features, e.g.\ by using \gls{SIFT} as described in \autoref{sec:sift}.
+tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
+\item[Visual Appearance] is extracted from the visual input. For that Dora uses a trained
+a set of classifiers on global visual features such as \gls{CRFH} to categorize the visual
+appearance of a room in categories such as corridor, office, library.
+\end{description}
 
@@
 those concepts to instances of spatial entities (see the example of the ontology in
-\autoref{fig:dora-architecture}).
+\autoref{fig:spatial-knowledge}).
 

### 2011-06-15 thesis/synthetic/plot.tex
Added plot of optimal decision.

@@
 
+\section{Optimal Decision}
+\begin{tikzpicture}[only marks,mark size=1.2pt]
+\begin{axis}[
+ xlabel={Threshold},
+ ylabel={Graph Size},
+ width=\textwidth,
+ height=1.5\textwidth]
+\addplot[mark=+,mark options={fill=white}] file {result_opt_threshold_known.data};
+\addplot[mark=-,mark options={fill=black}] file {result_opt_threshold_novel.data};
+\end{axis}
+\end{tikzpicture}
+
 \section{Constant $P(novel)$}
-\begin{tikzpicture}[only marks,mark size=0.5pt]
+\begin{tikzpicture}[only marks,mark size=1pt]
 \begin{axis}[
@@
 \section{Dynamic $P(novel)$}
-\begin{tikzpicture}[only marks,mark size=0.5pt]
+\begin{tikzpicture}[only marks,mark size=1pt]
 \begin{axis}[

### 2011-06-15 thesis/synthetic/single-factor.cc

@@ -141,9 +141,9 @@ public:
 
 }
 
-size_t room_types = 10;
-size_t n_known_rooms = 3;
-size_t property_types = 15;
+size_t room_types = 7;
+size_t n_known_rooms = 5;
+size_t property_types = 8;
 int C = 1;
 
 set<string> known_rooms;
@@ -175,7 +175,7 @@ void generate_real_distribution_vars_and_factors() {
   // Real property variable.
   {
     VariableType &type = *(type_real_prop = gi.Add("real_prop", VariableType()));
-    for (int i = 0; i < room_types; ++i)
+    for (int i = 0; i < property_types; ++i)
       type.values.push_back(prop(i));
     sort(type.values.begin(), type.values.end());
   }
@@ -187,7 +187,9 @@ void generate_real_distribution_vars_and_factors() {
     factor.variables.push_back(make_pair("prop", "real_prop"));
     BOOST_FOREACH(const string &room, type_real_room->values)
     BOOST_FOREACH(const string &prop, type_real_prop->values)
-      factor.potential[vector_of(room, prop)] = drand48();
+      factor.potential[vector_of(room, prop)] = 0.1;
+    BOOST_FOREACH(const string &room, type_real_room->values)
+      factor.potential[vector_of(room, type_real_prop->values[random()%property_types])] = 0.3 + drand48();
   }
 }
 
@@ -303,6 +305,7 @@ double junk() { return 0.1 + drand48()/5; }
 void compare_performance(const vector<vector<string> > &samples) {
   vector<pair<pair<int,double>,bool> > result_fix;
   vector<pair<pair<int,double>,bool> > result_dyn;
+  vector<pair<pair<int,double>,bool> > result_opt;
 
   BOOST_FOREACH(const vector<string> &sample, samples) {
     int n_properties = sample.size()-1;
@@ -310,13 +313,16 @@ void compare_performance(const vector<vector<string> > &samples) {
     GraphStructure s;
     s.createRandom(1, n_properties, 0);
     MGraph known_g(s, type_known_room, type_real_prop, NULL, factor_kroom_kprop);
-    MGraph any_g (s,  type_any_room,  type_real_prop, NULL, factor_aroom_kprop);
+    MGraph any_g  (s, type_any_room,   type_real_prop, NULL, factor_aroom_kprop);
+    MGraph real_g (s, type_real_room,  type_real_prop, NULL, factor_rroom_rprop);
 
     Query kq(&known_g);
[…]

### 2011-06-15 thesis/introduction.tex
Semantic Mapping chapter almost finished.

@@
 \chapter{Introduction}
+There has been several efforts in the areas of artificial intelligence and robotics
+in creating robots that are able to interact with humans and their environments.
+One of the important aspects is to endow those robots with a deeper understanding
+of human environments, not just for the purpose of navigation and obstacle avoidance, 
+but also in terms of human semantics and functionality.
+
+In that sense the need for semantic mapping arises, and it can be described as the
+relation between spatial areas and a set of existent concepts used by humans to
+describe them.
+
+An essential competence for cognitives robots is their ability to reason about space.
+Spatial knowledge plays an important role in navigation, reasoning, planning and episodic memory.
+
 % Introduce:

### 2011-06-15 thesis/semantic-mapping.tex
Semantic Mapping chapter almost finished.

@@
 association is either useful by explaining where to find certain objects as well to
-understand the types of activities to perform on it. In a
+understand the types of activities to perform on it. 
+Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
+knowledge about them it is expected to improve the agent ability to interact and communicate
+with humans and the ability perform human-like tasks.
 
-Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
-knowledge about them it is expected to improve the agent 
 % semantic mapping definition
 In this context semantic mapping is seen as the process of building a representation of
-the environment which associates spatial entities with a set of defined spatial concepts (i.e.\ human concepts).
+the environment which associates spatial entities with a set of defined spatial concepts (in this case the human concepts).
 
-% importance of conceptual knowledge
-% application of conceptual knowledge
-It is expected that by being able to reason over those semantics an agent can better
-understand human environments and increase its efficiency when performing human-tasks on them.
-By handling this conceptual knowledge the agent becomes aware of
+Missing:
+\begin{itemize}
+\item importance of conceptual knowledge
+% In order to build such a mapping of the environment, it is important to define the knowledge that allows the agent to reason
+% about.
 
-% challenges and need for representing uncertainty
-The goal of a spatial representation is to provide the agent with a view about the environment
-that allows it to reason beyond its sensory horizon.
+\item application of conceptual knowledge
+% It is expected that by being able to reason over those semantics an agent can better
+% understand human environments and increase its efficiency when performing human-tasks on them.
+% By handling this conceptual knowledge the agent becomes aware of
 
- Such a representation is by its nature
-imperfect, incomplete, inaccurate and invalid.
+\item challenges and need for representing uncertainty
+% The goal of a spatial representation is to provide the agent with a view about the environment
+% that allows it to reason beyond its sensory horizon.
 
-world that allows the agent to perform reasoning over it and efficiently perform tasks.
-It should also give the agent ability to reason about parts 
-Such a representation cannot be perfect since the methods and information are uncertain
+% Such a representation is by its nature imperfect, incomplete, inaccurate and invalid.
+
+% world that allows the agent to perform reasoning over it and efficiently perform tasks.
+% It should also give the agent ability to reason about parts 
+% Such a representation cannot be perfect since the methods and information are uncertain
+
+% It is therefore clear that the need to handle information on an high-level semantic and
+% probabilistic fashion arises not only for efficiency reasons but also from the desire
+% of being able to communicate and reason on those aspects.
+% In order to fully explore the semantic properties and their relations it is necessary to
+% detect and classify semantic aspects as well to understand the relationships between them.
+% That conceptual knowledge is important and any limitations on it will limit the capacities
+% of an agent to reason. 
+\end{itemize}
 
@@
 % on a probabilistic fashion.
-
-It is therefore clear that the need to handle information on an high-level semantic and
-probabilistic fashion arises not only for efficiency reasons but also from the desire
-of being able to communicate and reason on those aspects.
-In order to fully explore the semantic properties and their relations it is necessary to
-detect and classify semantic aspects as well to understand the relationships between them.
-That conceptual knowledge is important and any limitations on it will limit the capacities
-of an agent to reason.
-
-This thesis motivation lies on developing methods for detecting knowledge gaps
-on the semantic knowledge of a mobile agent. For that the semantic mapping system
-implemented on \gls{Dora} proposed by \cite{pronobis2011phd} was used as a base.
+This thesis motivation lies on developing methods for detecting knowledge gaps on the spatial
+semantic knowledge of a mobile agent. For that spatial knowledge needs to be defined, as well
+the methods used to perform the semantic mapping.
+In that context this chapter presents \gls{Dora}, a system that represents its spatial knowledge
+with the representation proposed by \cite{pronobis2010ias}, and performs semantic mapping using
+inference on a probabilistic graphical model.
 
@@
 
-Besides that the spatial knowledge representation has been drawn with an high-focus on probabilistic
+The \emph{spatial knowledge representation} has been drawn with an high-focus on probabilistic
 and uncertain reasoning, human interaction and life-long learning capabilities.
@@
 
-Being able to perform probabilistic inference on all available information Such a representation is suitable to detect novel concepts.
-
-The system is implemented on \Gls{Dora} and a brief overview of its architecture is
-given in the following sections.
+\gls{Dora}, the system, utilizes its spatial knowledge to perform semantic mapping of the environment
+using probabilist inferences on top of a graphical model.
+Due to its nature, it was considered suitable for implementing novelty detection of semantic concepts.
 

### 2011-06-15 thesis/background.tex
Changes on the background: classifiers and features.

@@
 
+One of the interesting facts about high-level problems is that they often require a
+broad overview of the whole system and several concepts. For example in the case of
+this thesis approach to novelty detection on semantic knowledge it becomes first
+important to understand how classification of already known categories is performed
+and how is knowledge of lower\hyp{}levels represented. I.e.\ understand how can
+lower\hyp{}level classifiers be represented, how to obtain those representations from
+training data, and how to use those them to infer new classifications from those models.
+It is also important to know how to model several of those classifications with
+probabilistic relations in a unified model and how to infer information on those.
+
+This chapter tries to introduce the reader to those aspects. The presented material
+is not complete, as such an attempt would be naive. In that sense this chapter limits
+to present the base ideas and main concepts behind and where appropriate it refers to
+articles that present the introduced techniques.
+Special attention is given to the \emph{probabilistic graphical models} as they are used
+as base tool for modelling the semantic map where novelty detection is then implemented.
+
 \section{Classification}
+Classification deals with the problem of identifying classes or groups that lie underneath
+the sensed data. It is expected that the presence of a underlying concept is involved on the
+generation of the data that is sensed, and the system tries to infer that underlying concept
+without directly observe it.
+
+Often the sensed data is just to large to be directly handled within a classifier. In those
+cases classifiers work on a small subset of features extracted from the data. Those features
+should avoid discard important information and if correctly employed should turn the classification
+easier.
+
 \subsection{Recognition and Categorization}
+Classification can target very different types of classes, as it is not clear on which classes
+the system is interested in distinguish. For example the system maybe interested in distinguish
+specific instances or interested in detecting a wide group of instances matching a
+common category. E.g. distinguish a specific room: room 304 in the 3rd floor, from a generic room
+category: a library.
+
+Based on the type of desired learning system, the available features and the required generalization
+specifications a wide range of machine-learning methods can be used. And no single method
+is expected to handle all the problems with optimal performance.
+The methods can go from statistical classification, neural networks, nearest neighbours,
+decision trees, support vector machines to others like graphical models, clustering,
+Gaussian mixture models, hidden Markov models, principal component analysis~\cite{bishop2006pattern}.
+
+This background chapter does not aims at describing all the classification methods, but at
+introducing the concept of classifier and to use them to extract information from low\hyp{}level
+features. In that sense only \glspl{SVM} will be generally described, for more information
+on others the reader is welcome to check the vast literature in pattern recognition and machine
+learning such as the reference cited above.
+
+\subsection{Supervised and Unsupervised Training}
+It is often impossible to manually specify the information need to correctly classify samples.
+There is also interest to make systems flexible and allow them to learn and adapt based solely
+on the available data.
+
+Methods to learn classifiers able to distinguish underlying classes on the data can be
+separated based on the model used to train it.
+
+\begin{description}
+\item[Supervised Methods] assume the existent of a supervisor that is able to give the
+ground\hyp{}truth. With it the algorithm tries to learn the best description that matches
+the given labelled data.
+\item[Unsupervised Methods] try to learn classes without any extra information. The methods
+have to figure out how many and which classes seem more reasonable to be modelled.
+\end{description}
+
+\subsection{Discriminative and Generative Models}
+\label{sec:discriminative-vs-generative}
+After constructing a classifier from the information available either in form of
+available knowledge on the task in hand and from acquired data samples the agent
+ends up with a model that represents the knowledge it believes to be suitable to
+solve the task.
+
+The produced model differs a lot based on the type of classifier but they can be
+distinguish in two different categories:
+\begin{description}
+\item[Discriminative models] are only able to classify samples in the known categories.
+\item[Generative models] model the full probability relation between sensed
+features and classes. With that it becomes possible to use the learned model to generate
+new samples.
+\end{description}
+
+\subsection{Multi-Class Classification}
+\label{sec:multiclass-classifiers}
+Most classifiers methods are designed as two-class classifier and do not
+natively support multi-class classification problems.
+The most common approach is to try to approach multi-class problems by
+by combining several two-class classifiers and perform a voting scheme or other
+integrating method based on the confidence of the individual two-class classifiers.
+For example:
+
+\begin{description}
+\item[One Against All] - in this method $c$ distinct classifiers are trained to
+distinguish any class of the remaining ones. The output of all those classifiers
+(distance to the separating hyperplane) is then used to categorize the output.
+The most common approach is to pick the class with the largest hyperplane
+distance. Other variations exists as is the case of using the minimal distance
+to the average classification distance of each class~\citep{pronobis2007iros}.
+
+\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to
+distinguish between each pair of classes. The final decision is based on the
+output of all those classifiers being common to use a majority vote strategy.
+\end{description}
+
 \subsection{Support Vector Machines}
@@
 \emph{maximal margin hyperplane}~(\autoref{fig:svm-sample}).
-
 Later they were extended to deal with noisy data by using \emph{soft margins}.
 And to handle non-linear spaces as seen on \autoref{sec:kernel-trick}.
-Although they only allow for two class-classification several methods have been
-proposed to build multi-class classifiers based on binary classifiers as seen on
-\autoref{sec:multiclass-classifiers}.
 
@@
 in fact a standard across machine learning techniques. Their efficiency, exact
-training results and generalization made their suitable for many tasks. Such as
+training results and generalization made them suitable for many tasks. Such as
 text categorization, digit-recognition, spam-classification.
 They have also been extensively used in visual place classification.
-
-\subsection{Multi-Class Classification}
-\label{sec:multiclass-classifiers}
-\Glspl{SVM} were designed as two-class classifier. And so they need to be
-adapted for using in multi-class problems. The most common usage is to design
-multi-class classifiers recurring at the usage of multiple two-class classifiers
-and methods for combining them. The most typical approaches for a multi-class
-classification on $c$ classes using \glspl{SVM} are:
-
-\begin{description}
-\item[One Against All] - in this method $c$ distinct classifiers are trained to
-distinguish any class of the remaining ones. The output of all those classifiers
-(distance to the separating hyperplane) is then used to categorize the output.
-The most common approach is to pick the class with the largest hyperplane
-distance. Other variations exists as is the case of using the minimal distance
-to the average classification distance of each class~\citep{pronobis2007iros}.
-
-\item[One Against One] - in this method $c*(c-1)/2$ classifiers are trained to
-distinguish between each pair of classes. The final decision is based on the
-output of all those classifiers being common to use a majority vote strategy.
-\end{description}
+% TODO: add citations.
 
@@
 \label{sec:kernel-trick}
-\gls{SVM} in their basic form are only able to handle linear spaces.
-Nonetheless the classes are most of the time not linearly separable in the input
+Several classifiers can be modelled by requiring only the concept of inner\hyp{}product
+between two samples. That product is often seen as a measure of distance or similarity
+between the samples on some space. By using kernels it becomes possible to define such
+space without explicit convert the samples to it. By abstracting the concept of distance
+on such a kernel function it also becomes possible to use those classification methods
+on strange data structures such as trees, strings or graphs.
+
+For example \glspl{SVM}, in their basic form, are only able to handle linear spaces.
+But the classes are most of the time not linearly separable in the input
 space. Although there might exists a transformation $\phi$ from the original
@@
 space: $K(x_i, x_j) = \phi(x_i)\cdot\phi(x_j)$.
-In that sense a Kernel function can be seen as a function that calculates some
-similarity measure between two inputs.
 
@@
 \item[Radial Basis Function] - $K(x, y) = e^{-\gamma\|x - y \|^2}$
-\item[Histogram intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where
+\item[Histogram Intersection] - $K(x, y) = e^{-\gamma \chi^2(x,y)}$, where
 $\chi^2(x,y) = \sum_{i=1}^{N}\frac{(x_i-y_i)}{x_i+y_i}$ introduced by
@@
 A feature is a piece of information which is expected to reveal information for
-solving a specific task. In that sense features are task-dependant and they will
+solving a specific task. Features are task-dependant and they will
 yield different performance based on the type of task they are applied to.
@@
 different impact in indoor scene recognition based on the type of scene
-matching. Namely was seen that some room-categories are more likely to be
+matching. Specifically it was seen that some room-categories are more likely to be
 recognized by the presence of some objects and others by it generic appearance.
@@
 In the context of robotics, sensors such as cameras, laser scans are used to
-sense the surrounding environment, and features can be extracted from all those.
-Nonetheless as described in \autoref{sec:visual-motivation} the visual sensor is
-incredibly rich and most of our features will come from it.
-
+sense the surrounding environment, and features can be extracted from those.
 Visual features can be seen as belonging to two categories:
 \emph{local features} and \emph{global features}.
-\emph{Local features} describe fine grain properties of a part of image.
-For example the existence of specific corner or an edge. \Gls{SIFT}
-(\autoref{sec:sift}) is an example of such feature and has been proven useful
-for matching points between images and subsequence extension to object detection.
-
-\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
-{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
-statistical analysis of features over all the image or by a structured
-distribution of textures findable in the image.
 
@@
 \label{sec:local-features}
+\emph{Local features} describe fine grain properties of a part of image.
+For example the existence of specific corner or an edge. \Gls{SIFT}
+(\autoref{sec:sift}) is an example of such feature and it been proven useful
+for matching points between images and subsequence extension to object detection.
 
-\subsubsection{SIFT}
+\subsubsection*{Interesting Points and SIFT}
 \label{sec:sift}
@@
 \begin{figure}[h]
-    \includegraphics[width=\textwidth]{figures/sift/sift.pdf}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/sift/sift.pdf}
     \caption{{SIFT} and other local features have been proven useful in object
@@
 \subsection{Global Features}
+\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
+{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
+statistical analysis of features over all the image or by a structured
+distribution of textures findable in the image.
 
-\subsubsection{Gist of a Scene}
+\subsubsection*{Gist of a Scene}
 \label{sec:gist}
@@
 
+As seen on \autoref{fig:gist} the gist is able to capture the dominant textural
+features of the overall image and their coarse spatial layout~\citep{murphy2006object}.
+With that, it is expected they serve the purpose of correctly describe the image
+textures without the need to directly stored the original image.
+
 \begin{figure}[h]
@@
 \includegraphics[width=0.60\textwidth]{figures/gist.jpg}
-\caption{An illustration of the gist of an image. Top row: original image I;
-         bottom row: noise image J for which gist(I) = gist(J). We see that the
-         gist captures the dominant textural features of the overall image, and
-         their coarse spatial layout~\citep{murphy2006object}.}
+\caption{\label{fig:gist}An illustration of the gist of an image. Top row: original image I;
+         bottom row: noise image J for which gist(I) = gist(J).}
 \end{figure}
 
-\subsubsection{{CRFH} - Composed Receptive Field Histograms}
+\subsubsection*{{CRFH} - Composed Receptive Field Histograms}
 \label{sec:crfh}\label{sec:global-features}
-Composed Receptive Field Histograms are a multidimensional statistical
+\gls{CRFH} are a multidimensional statistical
 representation of the occurrence of several image descriptors applied to an
@@
 how many pixels of the image have the cell response for the applied descriptors.
-Such high-dimensional histogram is expected to be able to better global
-information contained in the image by capturing several properties of the image
-as well the relations between them.
+Such high-dimensional histogram is expected to be able to describe global
+information contained in the image by capturing several properties that co-occur
+in a part of the image.
+By using some techniques~\cite{linde2004object} several operations on those
+high-dimensional histograms can be made computational efficient. This way not
+only this descriptor discards part of the local information present on the image
+but also allows to faster computations on similarity measures.
 
@@
 
-Several type of image descriptors can be applied.
-For example Gaussian derivatives (such as $L_x$, $L_y$, $L_{xx}$, $L_{yy}$) are
-partial derivatives of the image luminance after applying a Gaussian filter of a
-given scale ($\sigma$) on the image. Gradient magnitude descriptors
-(such as $|\nabla L|$) can also be used as a rotation invariant descriptor.
-
-One of the problems of high-dimensional histograms is the memory and
-computationally complexity needed to handle them.
-\cite{linde2004object} suggested using a sparse form to represent those by
-storing only the non-zero-cells in an array structure.
-
 Multidimensional histograms have proven to be useful in the context of object
@@
 
-\subsection{Laser Features}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\clearpage
 \section{Probability Theory}
@@
 distribution shall be picked, etc\dots
-
-% In case we expand in some finding structure in graphs maybe the following will
-% be useful
-\section{Label Propagation, Min-Cuts}

### 2011-06-16 thesis/synthetic/single-factor.cc
Made test data have equal number of known and novel samples.

@@ -139,6 +139,25 @@ public:
 
 };
 
+void FilterVariable(const set<string> &allowed, VariableType *type) {
+  type->values.clear();
+  type->values.reserve(allowed.size());
+  type->values.insert(type->values.end(), allowed.begin(), allowed.end());
+}
+
+
+void FilterFactor(int varid, const string &vartype, const VariableType &var, FactorData *data) {
+  (*data).variables[varid].second = vartype;
+
+  map<vector<string>, double> potential;
+  typedef pair<vector<string>,double> iter;
+  BOOST_FOREACH(const iter &it, data->potential)
+    if (find(var.values.begin(), var.values.end(), it.first[varid]) != var.values.end())
+      potential.insert(it);
+  potential.swap(data->potential);
+}
+
+
 }
 
 size_t room_types = 7;
@@ -147,31 +166,42 @@ size_t property_types = 8;
 int C = 1;
 
 set<string> known_rooms;
+set<string> novel_rooms;
 
 
 GraphInformation gi;
+// Real World
 VariableType *type_real_room = NULL;
 VariableType *type_real_prop = NULL;
 FactorData *factor_rroom_rprop = NULL;
 
+// Filtered Real Worlds
+VariableType *type_real_kroom = NULL;
+VariableType *type_real_nroom = NULL;
+FactorData *factor_rkroom_rprop = NULL;
+FactorData *factor_rnroom_rprop = NULL;
+
+// Conditional Learning
 VariableType *type_known_room = NULL;
 FactorData *factor_kroom_kprop = NULL;
 
+// Unconditional Learning
 VariableType *type_any_room = NULL;
 FactorData *factor_aroom_kprop = NULL;
 
 void generate_real_distribution_vars_and_factors() {
   for (size_t i = 0; i < n_known_rooms; ++i)
     known_rooms.insert(room(i));
+  for (size_t i = n_known_rooms; i < room_types; ++i)
+    novel_rooms.insert(room(i));
 
   // Real room variable.
   {
     VariableType &type = *(type_real_room = gi.Add("real_room", VariableType()));
-    for (int i = 0; i < room_types; ++i)
-      type.values.push_back(room(i));
+    type.values.insert(type.values.end(), known_rooms.begin(), known_rooms.end());
+    type.values.insert(type.values.end(), novel_rooms.begin(), novel_rooms.end());
     sort(type.values.begin(), type.values.end());
   }
-
   // Real property variable.
   {
     VariableType &type = *(type_real_prop = gi.Add("real_prop", VariableType()));
[…]

### 2011-06-16 thesis/background.tex
Changes on background chapter.
I am not exactly happy about it.
Ideally it should only present factor graphs in details, but the requirements
and available time made it look very generic and not really useful.
At least it should fulfill with some pages such that evaluators don't complain.

@@
 this thesis approach to novelty detection on semantic knowledge it becomes first
-important to understand how classification of already known categories is performed
-and how is knowledge of lower\hyp{}levels represented. I.e.\ understand how can
+important to understand how classification of already known categories is performed.
+I.e.\ understand how can
 lower\hyp{}level classifiers be represented, how to obtain those representations from
@@
 
-This chapter tries to introduce the reader to those aspects. The presented material
-is not complete, as such an attempt would be naive. In that sense this chapter limits
-to present the base ideas and main concepts behind and where appropriate it refers to
-articles that present the introduced techniques.
-Special attention is given to the \emph{probabilistic graphical models} as they are used
-as base tool for modelling the semantic map where novelty detection is then implemented.
+This chapter tries to briefly introduce the reader to those aspects. The presented material
+is not complete, it serves only to give directions and cues to the reader on how the several
+subproblems of visual place classification can be solved. 
+Where appropriate it refers to articles or textbooks that present the introduced techniques.
 
@@
 on others the reader is welcome to check the vast literature in pattern recognition and machine
-learning such as the reference cited above.
+learning such as the standard textbook cited above.
 
 \subsection{Supervised and Unsupervised Training}
-It is often impossible to manually specify the information need to correctly classify samples.
+It is often impossible to manually specify the information needed to correctly classify samples.
 There is also interest to make systems flexible and allow them to learn and adapt based solely
-on the available data.
-
-Methods to learn classifiers able to distinguish underlying classes on the data can be
-separated based on the model used to train it.
+on the available data. In that sense the option is to train classifiers from available data.
+The training methods are often separated as:
 
@@
 After constructing a classifier from the information available either in form of
-available knowledge on the task in hand and from acquired data samples the agent
+available knowledge on the task in hand and from acquired data samples, the agent
 ends up with a model that represents the knowledge it believes to be suitable to
@@
 The produced model differs a lot based on the type of classifier but they can be
-distinguish in two different categories:
+distinguished in two different categories:
 \begin{description}
@@
 \label{sec:local-features}
-\emph{Local features} describe fine grain properties of a part of image.
+Local features describe fine grain properties of a part of image.
 For example the existence of specific corner or an edge. \Gls{SIFT}
-(\autoref{sec:sift}) is an example of such feature and it been proven useful
+is an example of such feature and it been proven useful
 for matching points between images and subsequence extension to object detection.
@@
 \subsection{Global Features}
-\emph{Global features} such as \gls{CRFH} (\autoref{sec:crfh}) or
-{gist}~(\autoref{sec:gist}) try to describe the whole image. Either by
+Global features try to describe the whole image. E.g.\ either by
 statistical analysis of features over all the image or by a structured
@@
 \section{Probability Theory}
-\subsection{Bayes Theorem}
+An important aspect when making decisions with noisy or missing information is
+the uncertainty of the received information and of the final conclusion.
+Probability theory comes to help by providing a framework able to deal with those
+issues~\cite{bishop2006pattern}.
 
-\subsection{Categorical and Multinomial Distribution}
+The key concept on it is that of \emph{probability}, which can be seen as way of
+express the degree of belief that a certain event occurs.
+When combined with decision theory it allows to perform optimal decisions with the
+available information, even if it is ambiguous or incomplete.
+
+A probability $P(x)$ of a given event $x$ is a value between $0$ and $1$, where $0$ means
+the belief that $x$ will not occur and $1$ that it will certainly occur.
+In cases $x$ does not specifies all the variables states that define the sample space, $P(x)$
+is called a marginal probability.
+
+Often it is impossible to exactly describe the probability of a given event, in such cases
+the concept of likelihood comes to help.
+The likelihood of an event only has meaning when compared to one of another event.
+In such case the ratio between both likelihoods will denote on how more likely an event is
+over another.
+When dealing with discrete and countable sample spaces any likelihood function can be converted
+to a probability function by calculating the normalization factor such that the sum over all
+the sample space is $1$.
+
+\subsection{Conditional Probability}
+It is also possible to denote the known information when calculating probabilities. For that
+the notion of conditional probability is used $P(a|b)$. It denotes the probability of $a$
+knowing that $b$ has occur. Often conditional probability is also represented by: $P_b(a)$.
+In this thesis both notations will be used: the former for representing the information on
+sensed variables and the second for representing information or assumptions on other generic
+information such as graph structures.
+
+Marginal probabilities can be related with conditional probabilities as described in the equation:
+\begin{equation}
+P(a|b) P(b)  = P(a, b)
+\end{equation}
+
+\section{Principle of Maximum Entropy}
+\label{sec:max-entropy}
+The principle of maximum entropy~\cite{shore1980axiomatic} states that when given
+a set of distributions that are coherent with the acquired knowledge, the one which
+maximizes entropy should be picked.
+
+This can be used to select a distributions that most correctly describes the obtained
+knowledge. For example, if all that is known from a distribution is the mean and deviation
+then the correct approach is to model it with a normal distribution with the known parameters.
+In case of having no information available, the uniform distribution shall be assumed.
 
@@
 This type of graphs provide a generative model where the probability of any
-given scenario can be determined. As opposed to non-generative models where a
-given probability can only be calculated if certain data is given.
-This means when a graphical model is learned, it can be used to generate new
+given scenario can be determined.
+This means once a graphical model is learned, it can be used to generate new
 samples from the learned distribution.
@@
 
-Two main types of graphical models have been used: Bayesian Networks which model
+Two main types of graphical models are widely used: Bayesian Networks which model
 directed edges between variables and Markov Random Fields where variables are
@@
 determine most likely scenarios. In the case of directed edges a node blanket is
-also influenced by the direction of the edges and a more complex schemes need to
+also influenced by the direction of the edges and more complex schemes need to
 be used.
@@
 \begin{figure}[ht]
-    \begin{minipage}[b]{0.3\linewidth}
 \centering
-        \includegraphics[width=0.8\textwidth]{figures/graphical-models/BayesNet.pdf}
-        \caption{Bayes Networks}
-    \end{minipage}
-    \hspace{0.5cm}
-    \begin{minipage}[b]{0.3\linewidth}
-        \centering
-        \includegraphics[width=0.8\textwidth]{figures/graphical-models/MarkovRandomField.pdf}
-        \caption{Markov Random Field}
-    \end{minipage}
-    \hspace{0.5cm}
-    \begin{minipage}[b]{0.3\linewidth}
-        \centering
-        \includegraphics[width=\textwidth]{figures/graphical-models/FactorGraph.pdf}
-        \caption{Factor Graph}
-    \end{minipage}
+
+\subfloat[Bayes Networks] {
+  \includegraphics[width=0.2\textwidth]{figures/graphical-models/BayesNet.pdf}
+}
+\quad
+\subfloat[Markov Random Field] {
+  \includegraphics[width=0.2\textwidth]{figures/graphical-models/MarkovRandomField.pdf}
+}
+\quad
+\subfloat[Factor Graph] {
+  \includegraphics[width=0.3\textwidth]{figures/graphical-models/FactorGraph.pdf}
+}
 \end{figure}
@@
 \subsection{Factor Graphs}
-A \emph{factor graph} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
+A \emph{factor graph}~\cite{kschischang2001factor} is a bipartite graph connecting two sets of nodes $X_G$ and $F_G$
 representing random variables and factors.
@@
 
-More often we are only interested in a marginal of the probability distribution.
+\subsection{Inference}
+The goal by obtaining a graphical model between all the variables of a system is the ability
+to efficiently infer the most likely scenarios based on the available information.
+The sensed information can be used to clamp variables and the model used to calculate probabilities
+or marginal probabilities.
 
-\begin{equation}
-M_G(x) = \frac{1}{Z}\sum_{X_G \backslash x}{\prod_{\phi \in F_G}{\phi(x_{\phi})}}
-\end{equation}
-
-\subsection{Inference Engines}
-Exact Inference and Approximate Inference methods
-
-\section{Principle of Maximum Entropy}
-\label{sec:max-entropy}
-The principle of maximum entropy states that given a set of distributions that
-are coherent with the acquired knowledge, the one which maximizes entropy should
-be picked.
-
-In the case of non-available information this is the uniform distributions.
-In cases where we only know the mean and standard deviation a normal
-distribution shall be picked, etc\dots
+In that sense a basic operation on a graph is the \gls{MAP} operation, which allows to calculate
+which configuration on a variable or subset of variables maximizes a given function.

### 2011-06-17 thesis/novelty-intro.tex
Fixes here and there.
Novelty Detection by Thresholding - fixed acording to andrzej comments.

@@
 
-This chapter presents concepts that are used as base when adding novelty
-detection capabilities on the semantic representation in \autoref{chap:novelty}.
-It starts by providing the reader with a brief overview on novelty detection
-techniques and related works.
+This chapter presents the main work of this thesis: a method to augment the conceptual
+map with novelty detection capabilities.
 
-After it shows that an optimal novelty detection system can be implemented by
-thresholding on a order-relation defined over the inputs.
+The chapter starts by providing the reader with a brief overview on novelty detection
+techniques and related works. Then it shows that an optimal novelty detection system
+can be implemented by thresholding on a order-relation defined over the inputs.
 And that an equivalent order-relation can be imposed by a ratio between
@@
 \section{Novelty Detection Review and Related Work}
-Novelty detection, also called outlier or anomaly detection, is the
-classification problem related with identification of new or unknown data
+Novelty detection, also known as outlier or anomaly detection, is a
+classification problem related to identification of new or unknown data
 patterns that the system is not aware of~\cite{markou2003novelty}.
@@
 that is deployed to unknown or to uncontrolled environments, as it gives the
-system the ability to detect something is not conforming to it knowledge and
-therefore should proceed with caution.
+system the ability to detect that something is not conforming to its knowledge and
+therefore should be treated with caution.
 It has several applications such as fault detection~\cite{tarassenko1999novelty},
@@
 Any normal classification application is also a good candidate for extending
-with novelty detection, as it allows to detect samples the system has not been
-trained on and therefore the classification is considered
-instable~\cite{devarakota2008reliability}.
-E.g.\ applied to digit-recognition~\cite{tax1998outlier}
+with novelty detection, as the results for samples the system has not been
+trained on can be unreliable~\cite{devarakota2008reliability}.
+e.g.\ applied to digit-recognition~\cite{tax1998outlier}
 or to detection of novel inputs on a neural network~\cite{bishop1994novelty}.
 
-It is on the nature of unknown environments and in the infeasibility of
-training the system on all possible object classes that lies the complexity of
-novelty detection: it is only possible to obtain samples representing positives
+It is in the nature of unknown environments and in the infeasibility of
+the system to acquire examples of all possible classes where the complexity of
+novelty detection lies: it is only possible to obtain samples representing positives
 examples of known cases and the lack of negative examples renders normal
 classification methods unusable.
-
 In order to become practical in real world applications, novelty detection
@@
 
-% TODO: wuld be nice to have this one starting on a new page.. but still make
-% sure it looks good and not a forced break.
-% Reason to have this on a diferent page is to make it really clear novelty
-% content (and not only chit-chat) starts now.
-\clearpage
-\section{Novelty Detection as a Threshold}
-% Explain threshold approach to novelty detection
-Due to the desire of robustness, a novelty detection system is often given the option
-to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
-between error and detection: the more the system tries to increase detection, the more
-prone to produce errors it becomes.
+\section{Novelty Detection by Thresholding}
+The objective of a novelty detection system is to classify a given sample $x$ as
+either \emph{known}: $x$ being generated by a known class to the system, or \emph{novel}:
+$x$ generated by a class unknown for the system.
+Based on the ground\hyp{}truth of a sample four cases are possible:
 
-When dealing with a statistic point of view, noisy data or unstable features,
-a decision to classify a sample $x$ as novel increases detection rate by
-$P(novel|x)P(x)$ and introduces an associated error rate of $P(\overline{novel}|x)P(x)$.
-Generalizing to a system that classifies a set of inputs $N$ as novel, the
-true positive and false negative rates arise:
+\begin{description}
+\item[true positive]  - when a system correctly   flags a sample of an unknown class as \emph{novel}.
+\item[false positive] - when a system incorrectly flags a sample of a known class as \emph{novel}.
+\item[true negative]  - when a system correctly   flags a sample of a known class as \emph{known}.
+\item[false negative] - when a system incorrectly flags a sample of an unknown class as \emph{known}.
+\end{description}
 
-\begin{equation}
-\label{eq:true-positive}
-P(\textnormal{true positive}) = \sum_{x \in N}{P(novel|x)P(x)}
-\end{equation}
+Due to noisy data, unstable features and lack of information it is impossible to develop
+a method able to exactly guess the correct classification of a sample.
+By modelling the outcomes with probabilities it becomes possible to handle the uncertainty
+associated with each decision.
+The notation $P(novel|x)$ will be used to denote the probability that the sample $x$ is in fact a
+sample of an unknown class. $P(x)$ will denote the probability that sample $x$ is given to the
+system to be classified. $\overline{novel}$ is also defined in such a way that
+$P(\overline{novel}|x)$ measures the probability that a decision on classifying $x$ as novel is
+wrong.
 
-\begin{equation}
-\label{eq:false-positive}
-P(\textnormal{false positive}) = \sum_{x \in N}{P(\overline{novel}|x)P(x)}
-\end{equation}
+Additionally a decision on the novelty of a sample $x$ performed by a deterministic system
+will be fully determined by the sample itself.
+Which implies that any deterministic novelty detection system can be uniquely determined by the set $N$
+of samples that are accepted by the classifier as \emph{novel}.
+This way it is possible to define the probability of a true positive and false positive event for any
+deterministic novelty detector with the following equations:
 
-The performance of a novelty detection system can be measured on these two
-rates: the true-positive rate (\autoref{eq:true-positive}) describes the
-interest in detecting as much as possible the novel cases;
-and the false-positive rate (\autoref{eq:false-positive}) the interest in
-having the smallest error possible.
-As a given sample $x$ is included in $N$ both the true-positive and
-false-positive rates never-decrease~\footnote{Note that both are described as
-summations of products of probabilities (non-negative functions).} describing an
-\emph{error and rejection tradeoff}~\cite{chow1970optimum} that any detection
-system has to consider.
+\begin{eqnarray}
+P(\text{true positive})  &=& \sum_{x \in N}{P(novel|x)P(x)} \\
+P(\text{false positive}) &=& \sum_{x \in N}{P(\overline{novel}|x)P(x})
+\end{eqnarray}
 
-An optimal novelty detector can then be formulated as selecting a subset $N$
-that achieves a given true-positive rate with the minimal false-positive
-rate possible. This is equivalent to the \emph{continuous knapsack problem}
-which allows a greedy solution by sorting the items with a value per weight
-function.
-Therefore a novelty system before classifying a sample $a$ as novel should
-classify any sample $b$ satisfying \autoref{eq:knapsack} as that would achieve
-a lower false-positive rate given a fixed true-positive rate.
+Note that since probability functions are non-negative, it is impossible to decrease either the
+true positive or the false positive probability by using a set $N' \supset N$.
+This describes the base of the \emph{error and rejection tradeoff}~\cite{chow1970optimum}, which
+states that a system aiming at increasing the true-positive probability will eventually increase its
+false-positive error.
+The true positive probability can be described as the interest in detecting as much
+as possible the novel classes and the false positive probability the interest
+in not incurring in too many errors.
+By fixating one of those it is possible to define a novelty detection
+system that achieves the maximum or minimum of the other.
+
+This way an optimal detector can be formulated by achieving the maximum true-positive
+probability without its false-positive probability increase beyond a given limit.
+This is equivalent to a \emph{continuous knapsack problem} which allows a greedy
+solution by sorting the items with a value per weight function.
+In the case of detection system that can be defined as:
+
+\begin{eqnarray}
+value(x)  &=& P(\text{true positive}|x) \\
+weight(x) &=& P(\text{false positive}|x) \\
+cost(x)   &=& value(x)/cost(x) \\
+          &=& \frac{P(novel|x)P(x)}{P(\overline{novel}|x)P(x)}
+\end{eqnarray}
+
+Therefore a novelty system before classifying a sample $a$ as novel should (greedily)
+classify any sample $b$ with a small cost as that would achieve
+an higher true positive probability given a fixed false positive one.
 
@@
 \label{eq:knapsack}
-\frac{P(\overline{novel}|b)P(b)}{P(novel|b)P(b)} < \frac{P(\overline{novel}|a)P(a)}{P(novel|a)P(a)}
+\frac{P(novel|b)}{P(\overline{novel}|b)} < \frac{P(novel|a)}{P(\overline{novel}|a)}
 \end{equation}
 
-This relation between $a$ and $b$ can then be simplified into:
+This relation between $a$ and $b$ can further be simplified into:
 
@@
 interested in defining an order relation on all the possible inputs equivalent
-to the order defined by the error rate: $P(\overline{novel}|x)$.
+to the order defined by the function: $P(\overline{novel}|x)$.
 And any optimal detector can be described by the largest $P(\overline{novel}|x)$
@@
 \emph{unconditional probability} suitable for implementing novelty detection
-via thresholding.
+by thresholding.
 
@@
 The conditional probability $P(x|\overline{novel})$ describes the distribution
-of the non-novel samples on the original distribution. In case the labelled data
-available for the agent to learn a concept comes from the same distribution
-where the system will run the correct approach is to use it as prior-information
-for modelling the conditional probability.
+of the samples given that they are generated by a known class.
+In case the labelled data available for the agent to learn a concept comes from
+the same distribution where the system will run the correct approach is to use it
+as prior-information for modelling the conditional probability.
 
-Note that it is important for the labelled data to come from that same
-distribution, otherwise it will contain some bias and lead to an incorrect
-ordering. The bias, for example, can be introduced by an incorrect filtering.
+Note that it is important for the labelled data to be a filtered version
+of the underlying world distribution (with all classes) that does not contains
+any bias.
+Otherwise that bias will lead to incorrect modelling the conditional probability
+and a subsequent wrong ordering of the sample space.
 

### 2011-06-17 thesis/novelty-intro.tex
Changes on related work of novelty detection chapter

@@
 
-\subsection{Review}
+\subsection{Review of Novelty Detection Methods}
 Many approaches have been made based on statistical methods. Those often model
@@
 A common approach is to use density estimation and use the expected probability
-of a sample to trigger the sample as novel. Examples are usage of
-Gaussian Mixture Models and Parzen-window estimators. In order to be effective
-those use data as close as possible to the input features and use
-dimension-reduction techniques such as \gls{PCA} to make density
+of a sample to classify the sample as novel. Examples of such techniques are
+Gaussian Mixture Models and Parzen-window estimators. In order to be effective,
+those rely on data as close as possible to the input features and use
+dimensionality\hyp{}reduction techniques such as \gls{PCA} to make density
 estimation feasible. Note that as dimension increases an exponential number
@@
 \cite{bishop1994novelty} uses that approach by employing a Parzen-window to
-estimate the density of the training data on a given input feed into a
-neural-network. By triggering on the calculated density they detect samples
+estimate the density of the training data on a given input fed into a
+neural-network. Using the calculated density, they detect samples
 that differ from the training data and consider their neural-network output
@@
 
-A slightly different approach is done by one-class \gls{SVM} approaches that
+A slightly different approach is applied in case of by one-class \gls{SVM} approaches that
 try to distinguish novelty by separating the training set from all the other
@@
 They use the assumption that the class to be defined lies on a manifold embedded
-on an input space of higher dimensions. By using dimensional reduction
-techniques, they try to defined that manifold and calculate how far a new sample
-is from it.
-One of the most common methods used for that is
+into a sample space of higher dimensionality. By using dimensionality\hyp{}reduction
+techniques, a manifold is defined and the distance between the manifold and the new sample
+is calculated. One of the most common methods used for that purpose is
 \gls{K-PCA}~\cite{scholkopf1997kernel}, which uses the kernel-trick to extend
-\gls{PCA} and perform a nonlinear dimensionality reduction on the input.
-This technique has been successfully used in novelty detection by
-\cite{Hoffmann2007863}.
+\gls{PCA} and perform a nonlinear dimensionality reduction of the input.
+This technique has been successfully used for novelty detection in \cite{Hoffmann2007863}.
 
-\subsection{Related Work}
-\cite{zhang2005probabilistic} present a probabilistic semantic approach to image
-annotation and retrieval with a multi-modal information. They introduce an
-hidden variable modelling a concept present on a scene that connects together
-information coming from both a bag of words and visual features extracted. By
-training that hidden-variable their system becomes aware of possible existent
-semantic categories and their relation between both textual and visual features.
-Although not able to perform novelty detection, their approach is interesting by
-the use of an hidden variable to probabilistic model the multi-modal sensors.
-
-\cite{boutell2006factor} presents a method to perform scene-classification
-from low-level region detectors using factor graphs. Those low-level detectors
-provide semantic (high-level) cues on the higher level scene-classification
-task. Due to the nature of the low-level detectors, they ought to be treated
-in a probabilistic way. For that their method uses factor graphs. The method
-also exploits spatial relation information between regions to improve the
-classification performance. A presented scheme using only pairwise relations
-between regions is show to have better performance than try to approach the
-correct scheme with a big single-factor. I.e.\ the pairwise relations can be
-approximated with the modest amounts of training data, where the single-factor
-demands an impractical amount of training data.
-No approach is made on it at novelty detection either of regions or scene
-categories.
-
-\cite{bishop1994novelty} works shows that assuming a uniform unconditional
-distribution novelty detection can be performed by triggering on the
-density probability of the training data. The system is employed to detect novel
-inputs for a neural-network.
+\cite{japkowicz1995novelty} also follows a similar approach: \emph{Redundancy
+Compression and Non-Redundancy Differentiation} which is a process believed
+to happen in the hippocampus. They introduce an auto-encoder
+that learns to encode a sample on a considerable smaller description and later reconstructs
+the original sample from this smaller description. Their suggested system
+learns how to discard and compress redundant information while still be able to recover
+the original sample. By training it with samples from a given class, it
+is expected that it will badly reconstruct a sample from a different one.
 
@@
 to run on a constant amount of memory and computation are presented.
+
+\cite{boutell2006factor} presents a method to perform scene-classification
+from low-level region detectors using probabilistic graphical models.
+The information obtained by the region detectors is exploited together with the
+spatial relations between on the higher level scene-classification task.
+A presented scheme using only pairwise relations between regions is show to
+have better performance than the correct approach of modelling all regions connectivity
+with a large single-factor as the pairwise relations can be
+approximated with modest amounts of training data, where the single-factor
+demands an impractical amount of training data.
+No approach is made on it at novelty detection either of regions or scene
+categories.
 

### 2011-06-17 thesis/introduction.tex
Initial part of introduction chapter done.

--- 2011-06-01 thesis/introduction.tex
@@
 %  Requirements to detect novelty and handle it.
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us. Understand our concepts and
-be able to reason.
-
-They need to possess ability to adapt to situations as its infeasible to rely
-on extensive man-work to tag objects, map space and code all the aspects that
-make up our human-reality.
-
-% == Introduce the need of semantic representations ==
-%
-% There is a lot of low-level methods (or specific knowledge to robots)
-% But the creation of semantic representations allow to bridge that low-level
-% sensing with high-level concepts that facilitate several high-level Tasks.
 
@@
 %
+
+% Introduction
+There has been several efforts in the areas of artificial intelligence and mobile robotics
+in creating robots that are able to interact with humans and their environments.
+Those robots would be able to move into our houses and offices. They would be able to communicate
+with us and provide help solving our tasks.
+
+% Keep listing problems robots should be able to solve 
+\begin{itemize}
+\item Keep listing problems robots should be able to solve
+\item They would dynamically adapt to our environment and exhibit intelligence by understanding
+\item Pose problems robots cannot solve without understanding human environments
+\end{itemize}
+
+% Representation of environments are needed!
+In order to reason and understand about the surrounding environment a robot first has to
+obtain a manageable representation of it.
+Such a representation provides the robot with a understanding on how the several
+entities relate to each other, and, if persistent, allow the robot to reason beyond its sensory
+horizon.
+
+% Machine-representations are not useful to solve the required problem.
+One may argue that robots and computers can already represent and understand environments, but they
+do so by using representations that are not human friendly: e.g.\ a robot can uniquely identify the
+location of a certain object with coordinates, but those coordinates are not suitable for humans.
+A robot may be able to detect a tag placed on an object, but such an approach besides limited
+and hard-working does not scales: new objects are created everyday, the huge amount of already
+existent objects and concepts also forbids any tagging and concepts by their nature are dynamic.
+Summing up: methods are needed that allow robots to represent environments and allow human concepts
+to also be representable.
+
+% Semantic Mapping
+By gathering semantic knowledge a robot can learn how to map their internal representation to the
+given set of human concepts. Web and other common databases, that have been created by humans, are a
+great source of semantic information, and, in some sense have helped to solve this semantic mapping
+problem.
+Common approaches at semantic mapping are based on classification, this is: a robot has a set of
+defined categories and it picks the one that best represents the sensed data. This approach often
+gives satisfactory results, but only when applied on a controlled environment where the robot has
+knowledge of all the categories that exist.
+
+% Complete knowledge? Don't joke me.
+This assumption on knowledge of all the categories that an entity can be mapped to is too strong
+to hold on unknown environments. In specific, a mapping defined over the human semantics will always
+be incomplete due to infeasibility to map all the concepts or even due to its dynamic nature where
+concepts change not only on time but also between sociocultural aspects.
+
+% Novelty Detection
+In this sense a robot should be able to identify gaps on its knowledge by detecting novelty.
+Developing methods for this problem helps to provide robots with novelty signals that can be
+used to improve its reliability, decide when to initiate active-learning or even provide basic
+information for keeping knowledge continuously adapted in long-term or life-long scenarios.
+
+The rest of this chapter presents in which context is this thesis interested in detecting novelty.
+It presents then the specific problem it addresses and defines the goals to achieve.
+At last it presents the structure of the other chapters.
+
+\section{Context}
+% Spatial Semantics only!
+% Mobile Robotics
+% Room Categories
+
+Being able to deal with any 
+in this sense this thesis will focus only on the semantic spatial mapping problem.
+
+Humans organize spaces as areas, they characterize them by the type of
+
+% Yeah thats what you call dream :P
+In an extreme ideal case the agent should be able to go from zero-knowledge to
+understanding presence of certain objects (as generators of a set of sensed
+sensed properties grouped locally), understand areas within rooms (sink-area on
+on a kitchen), rooms as separated by doors, and understand environments such
+as office, home, warehouse, spaceship.
+
+% Semantic knowledge (spatial-knowledge)
 % It also holds an important role in long-term planning and stability as it
@@
 
-% Deal with uncertainty -> need for probabilistic models.
-% Deal with semantic    -> need of structured models such as graphical models.
+In that sense the need for semantic mapping arises, and it can be described as the
+relation between spatial areas and a set of existent concepts used by humans to
+describe them.
+
+, not just for the purpose of navigation and obstacle avoidance, 
+but also in terms of human semantics and functionality.
+
+An essential competence for cognitives robots is their ability to reason about space.
+Spatial knowledge plays an important role in navigation, reasoning, planning and episodic memory.
+% == Introduce the need of semantic representations ==
+%
+% There is a lot of low-level methods (or specific knowledge to robots)
+% But the creation of semantic representations allow to bridge that low-level
+% sensing with high-level concepts that facilitate several high-level Tasks.
+
+% Spatial knowledge representation
+% Semantic Mapping
+
+\section{Motivation}
+% Detect limitations and novel knowledge
+
+Deploy to new environments, long-term and active learning,
+understand of limitations of the existent knowledge
+reliability
 
@@
 
-\section{Problem and Goals}
+\section{Problem}
+% Deal with uncertainty -> need for probabilistic models.
+% Deal with semantic    -> need of structured models such as graphical models.
+
+They need to possess ability to adapt to situations as its infeasible to rely
+on extensive man-work to tag objects, map space and code all the aspects that
+make up our human-reality.
+
+Lack of methods that are able to detect gaps on their
+Study how structure changes novelty
+
+\section{Proposal}
+
 % Extend a probabilistic graphical modelling framework with capabilities to
@@
 
-% Yeah thats what you call dream :P
-In an extreme ideal case the agent should be able to go from zero-knowledge to
-understanding presence of certain objects (as generators of a set of sensed
-sensed properties grouped locally), understand areas within rooms (sink-area on
-on a kitchen), rooms as separated by doors, and understand environments such
-as office, home, warehouse, spaceship.
-
 For that the probabilistic semantic representation presented by
-\cite{andrzej2011phd} is used.
+\cite{pronobis2011phd} is used.
 
@@
 
-\autoref{chap:semantic-mapping} describes the system proposed by
-\cite{andrzej}. In special it introduces the \emph{conceptual map} of the system
-that is responsible for the creation of the \emph{probabilistic semantic
-representation} that this thesis aims at improving by developing methods with
-the capability to identify knowledge gaps.
+\autoref{chap:semantic-mapping} describes the semantic mapping process.
+It introduces a spatial knowledge representation and describes a system
+that uses it for performing semantic mapping using using a \emph{probabilistic graphical model}.
 
@@
 statistical view point and how to interpret it as a threshold function.
-A simple approach on how to perform novelty on very simple graphs
-is given together with some results on the impact of using semi-supervised
-novelty detection to improve the system performance.
-
-\autoref{chap:novelty} presents the developed techniques developed to identify
-novelty on the semantic representation.
+It presents then the main work of this thesis: detecting novel semantic
+categories using graphical models.
 

### 2011-06-17 thesis/introduction.tex
Context on introduction done.

@@
 \section{Context}
+% Spatial Representations
 % Spatial Semantics only!
-% Mobile Robotics
-% Room Categories
+% Room Categories actually.
 
-Being able to deal with any 
-in this sense this thesis will focus only on the semantic spatial mapping problem.
+%
+In mobile robotics the main need is on spatial information. Having a representation of all
+the spatial entities and relations between them allow the robot to navigate and interact
+with them. Maps are the default tool for representing those spaces, and computers excel at
+creating close to exact representations of spatial environments.
+But which concepts existent in the human spatial knowledge are needed in order to turn those
+representations a usable tool to represent human concepts?
 
-Humans organize spaces as areas, they characterize them by the type of
+Different spatial knowledge types can be identified depending on the scale, abstraction level,
+and required generalization. For example geometric type aspects which robots are already very
+good at representing, object knowledge which is incredibly basic for humans, delineation
+of spatial areas and relations between instances of all those knowledge types. e.g.\ a specific
+object (book) is on top of another object (table) that is located inside an area (room-library)
+that is inside a another area (university).
+As seen on this example humans do not limit themselves at representing entities
+in space and spatially relate them but they also attribute semantic value to them.
 
-% Yeah thats what you call dream :P
-In an extreme ideal case the agent should be able to go from zero-knowledge to
-understanding presence of certain objects (as generators of a set of sensed
-sensed properties grouped locally), understand areas within rooms (sink-area on
-on a kitchen), rooms as separated by doors, and understand environments such
-as office, home, warehouse, spaceship.
+It is this semantic value attributed by humans to entities that is interesting for endowing
+robots with knowledge on how to interpret and reason over human environments.
+By understanding which objects are expected to be found on a room, what are the properties that
+distinguish room categories, which room types are likely to be connected together a robot can
+aim at increase its effectiveness on indoor environments.
 
-% Semantic knowledge (spatial-knowledge)
-% It also holds an important role in long-term planning and stability as it
-% hides the space-time local details and allows to focus on high-level thinking.
-
-In that sense the need for semantic mapping arises, and it can be described as the
-relation between spatial areas and a set of existent concepts used by humans to
-describe them.
-
-, not just for the purpose of navigation and obstacle avoidance, 
-but also in terms of human semantics and functionality.
-
-An essential competence for cognitives robots is their ability to reason about space.
-Spatial knowledge plays an important role in navigation, reasoning, planning and episodic memory.
-% == Introduce the need of semantic representations ==
-%
-% There is a lot of low-level methods (or specific knowledge to robots)
-% But the creation of semantic representations allow to bridge that low-level
-% sensing with high-level concepts that facilitate several high-level Tasks.
-
-% Spatial knowledge representation
-% Semantic Mapping
+A basic concept on indoor environments is that of room. Rooms allow humans to segment areas
+in high\hyp{}level entities that can be categorized according to the properties and expected
+activities to be performed on them.
+e.g.\ kitchens are rooms where cornflakes are usually found, a library is identified by the presence
+of many books, a lecture hall is a place where lessons are given.
+Since room categories play an important role in understanding and reasoning about
+indoor human environments this thesis will focus on them.
 
@@
 Study how structure changes novelty
+
+Being able to represent all this type of knowledge is an huge task on itself and this thesis
+focus mainly on how to detect novelty on semantic values  
+Pronobis~\cite{pronobis2011phd} has tackled this problem by proposing
 

### 2011-06-17 thesis/synthetic/single-factor.cc
Plot fixes.

@@ -363,7 +363,7 @@ void compare_performance(const vector<vector<string> > &samples) {
     
     // log(2) + log(\phi_k(x)) < k_i log(s_i) + log(\phi_a(x))
     // log(s_i) > (log(2) + log(\phi_k(x)) - log(\phi_a(x))) / k_i
-    double dyn_threshold = (klogZ - alogZ)/n_properties;
+    double dyn_threshold = (-1+klogZ - alogZ)/n_properties;
     result_dyn.push_back(make_pair(make_pair(n_properties, exp(dyn_threshold)), known_rooms.count(sample[0]) == 1));
 
     // Assuming a constant probability of drawing a novel sample.
@@ -371,8 +371,8 @@ void compare_performance(const vector<vector<string> > &samples) {
     vector<string> no_clamp;
     double t_klogZ = kq.LogZ(no_prop, no_clamp);
     double t_alogZ = aq.LogZ(no_prop, no_clamp);
-    double fix_threshold = (klogZ - t_klogZ) - (alogZ - t_alogZ);
-    result_fix.push_back(make_pair(make_pair(n_properties, exp(fix_threshold)), known_rooms.count(sample[0]) == 1));
+    double fix_threshold = exp((klogZ - t_klogZ) - (alogZ - t_alogZ));
+    result_fix.push_back(make_pair(make_pair(n_properties, fix_threshold), known_rooms.count(sample[0]) == 1));
     //result_dyn.push_back(make_pair(make_pair(n_properties, fix_threshold/n_properties), known_rooms.count(sample[0]) == 1));
 
     // Real distribution
@@ -427,11 +427,11 @@ void compare_performance(const vector<vector<string> > &samples) {
 void example_single_factor() {
   generate_real_distribution_vars_and_factors();
 
-  vector<vector<string> > conditional_samples(100);
+  vector<vector<string> > conditional_samples(20*n_known_rooms);
   generate_conditional_ksamples(&conditional_samples);
   learn_known_room_distribution(conditional_samples);
  
-  vector<vector<string> > unconditional_samples(1000);
+  vector<vector<string> > unconditional_samples(20*room_types);
   generate_unconditional_samples(&unconditional_samples);
   learn_any_room_distribution(unconditional_samples);
 

### 2011-06-18 thesis/introduction.tex
Introduction chapter finished

@@
 % Detect limitations and novel knowledge
+It is highly desirable that a robot can be deployed to new and unknown environments. Aiming at that,
+a robot has to perform its own semantic mapping of the environment instead of relying on a provision
+of already labelled maps.
+But the semantic mapping ability of a robot is restricted by its spatial and semantic knowledge.
+This poses a problem as its unrealistic to believe that on unknown environments the robot knowledge
+is complete. So it becomes important to detect gaps on the robot knowledge, such that the robot can
+be aware of limitations of its models and devise techniques to not fail victim of its limitations.
 
-Deploy to new environments, long-term and active learning,
-understand of limitations of the existent knowledge
-reliability
+Additionally the detection of knowledge gaps, will serve as base for several other problems in
+artificial intelligence such as active-learning and maintenance of long and life-long knowledge.
 
-Example of what kind of an high-level thinking on space and its semantics can
-provide: Get me a beer/milk concept.
+\section{Proposal and Goals}
+This thesis aims at developing methods that can be used to detect gaps on the semantic knowledge
+of an agent. For that the spatial knowledge representation proposed in \cite{pronobis2010ias} will
+be used as base defining the agent knowledge.
+Given an instance world, that knowledge, is used to perform semantic mapping~\cite{pronobis2011semmap}
+on indoor environments using a probabilistic graphical model. That model is instantiated from the
+environment using the spatial knowledge and captures both semantic and structural information of
+space. It allows the agent to map all the modelled variables in the presence of uncertainty and 
+therefore this thesis will focus on study and developing methods that can detect novel semantic room
+categories from it.
 
-% == Reasoning about need for Novelty Detection ==
-%
-% Dynamic environment (inability to know the environment or even to be possible
-% to map all semantic concepts to interact with humans).
-%
-% Knowledge-Awareness (knowing what we know... And detecting novel cases)
-% Identify gaps in the semantic knowledge.
-% Automatic detection and learning of novel concepts.
-%
-% Increase robustness
-% Self-Extending
-%
-
-\section{Problem}
-% Deal with uncertainty -> need for probabilistic models.
-% Deal with semantic    -> need of structured models such as graphical models.
-
-They need to possess ability to adapt to situations as its infeasible to rely
-on extensive man-work to tag objects, map space and code all the aspects that
-make up our human-reality.
-
-Lack of methods that are able to detect gaps on their
-Study how structure changes novelty
-
-Being able to represent all this type of knowledge is an huge task on itself and this thesis
-focus mainly on how to detect novelty on semantic values  
-Pronobis~\cite{pronobis2011phd} has tackled this problem by proposing
-
-\section{Proposal}
-
-% Extend a probabilistic graphical modelling framework with capabilities to
-% detect novelty.
-% Both in terms of novel classes as of novel structure.
-%
-Given a probabilistic structure representing the sensed conceptual knowledge
-obtained by the agent develop methods able to detect novelty present
-either on new semantic concepts (new classes) or even on structure that was
-previously unknown.
-
-For that the probabilistic semantic representation presented by
-\cite{pronobis2011phd} is used.
-
+% Study how structure changes novelty.
 % If everything goes fine... 2 of the "Future Directions" proposed are 

### 2011-06-19 thesis/abstract.tex
Abstract in english added.

@@
 \chapter*{Abstract}
+For a long time humanity has fantasized that one day robots will walk among us.
+They will move and be able to interact with us, understand our concepts and
+be able to reason.
+An important step on that direction is endowing robots with conceptual knowledge
+of human semantics. However its unrealistic to believe on the possibility of extensively
+describing human concepts to robots on a pre-deploy phase, and so, robots must be able to adapt
+and learn when confronted with novel situations.
+Detection of novel situations, where the robot knowledge is not sufficient, plays an important
+role for the subsequent adapting and learning of new concepts and its this thesis main topic.
 
-This should probably be written as last thing.
+In the context of mobile robotics, spatial concepts and semantics play a main role to perform
+complex human-like tasks and human interactions, for handling that a robot builds a representation
+of space extended with semantic properties, process which is known as semantic mapping.
+More exactly the ability to relate areas with human semantics describing expected properties
+and actions that take place on it allow the robot to reason at a very high-level close to humans.
+For example in a room scale level, humans categorize the space in kitchens, bedrooms, corridors,
+office, theaters, etc\dots
+
+This thesis studies a proposed %~\cite{pronobis2011semmap}
+process that utilizes spatial and semantic knowledge for semantic mapping using a
+probabilistic graphical model,
+and develops methods to perform detection of room categories not previously
+known to the robot using that graphical model.
+At last it draws attention to the use of unlabelled data on the novelty detection process,
+which can be used to considerably increase performance.
 

### 2011-06-20 thesis/conclusion.tex
Splitted conclusion and started to organize content.

@@
-\chapter{Conclusions and Future Work}
-\section*{}
+\chapter{Conclusions and Future Work}\label{chap:conclusions}
 
-In this thesis
+% Summary of developed work
+This thesis studied the problem of detecting situations where the knowledge of the robot
+is not 
+Gaps those, that disable the robot to function 
+It did so, specializing 
+that allows novelty detection of semantic room
+categories using as base the semantic mapping process proposed by \cite{pronobis2011semmap}.
+
+\url{https://github.com/andresusanopinto/novelty-detection-thesis}
+
+On the rest of this chapter the main results and conclusions are presented.
+Additionally limitations of the presented method and directions for future work are given.
+
+\section{Results and Conclusions}
+
+\section{Limitations}
+
+\begin{itemize}
+\item Strong assumption on constant $P(novel)$
+\item Threshold should actually be dynamic with structure
+\item Unable to drop normalizations factors
+\item Handle multiple novel variables at the same time
+\end{itemize}
+
+\section{Future Work}
+
+\begin{itemize}
+\item Extend novelty to all categories on the graph (room, size, shape, etc..)
+\item Handle uncertain sensing
+\item Probabilistic Space Segmentation (learn how to structure the graph)
+\item Learn Concepts / hidden variable types - (areas, rooms, floors, environments, etc\dots)
+\item Explain why a certain sample is novel
+\item Generate most likely novel sample
+\end{itemize}

### 2011-06-20 thesis/abstract.tex
Fixed abstract (english reviewed by me).

--- 2011-05-27 thesis/abstract.tex
@@
 \chapter*{Abstract}
+For a long time humanity has fantasized that one day robots will walk among us.
+They will move and be able to interact with us, understand our concepts and
+be able to reason.
+An important step on that direction is endowing robots with knowledge of human concepts and
+semantics. However its unrealistic to believe on the possibility of extensively
+describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
+to adapt and learn when confronted with novel situations.
+Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
+role for the subsequent adapting and learning of new concepts and its this thesis main topic.
 
-This should probably be written as last thing.
+In the context of mobile robotics, spatial concepts and semantics play a main role to perform
+complex human-like tasks and human interactions. For handling those, a robot builds a representation
+of space extended with semantic properties, process which is known as semantic mapping.
+This representation identifies spatial entities and classifies them according to their meaning to
+humans allowing the robot to reason at a very high-level, closer to humans.
+For example, in a room scale level, humans categorize space in kitchens, bedrooms, corridors,
+office, theater, etc\dots, by relating room areas with those categories describing expected
+properties and actions that take place on them a robot can solve tasks like:
+``Where are the cornflakes?''
+
+This thesis studies a proposed %~\cite{pronobis2011semmap}
+process that utilizes spatial and semantic knowledge for semantic mapping using a
+probabilistic graphical model,
+and develops methods to perform detection of room categories not previously
+known to the robot using that graphical model.
+At last, it draws attention to the use of unlabelled data on the novelty detection process,
+which can be used to considerably increase performance.
 

### 2011-06-20 thesis/novelty-intro.tex
cleanup novelty intro.

@@
 \subsection{Review of Novelty Detection Methods}
-Many approaches have been made based on statistical methods. Those often model
-the training data with statistic distributions and then perform hypothesis tests.
-For example: testing whether two distributions are the same or using distance to
-mean value or quartiles.
-In some sense all they define a distance measure to normality and employ a
-threshold on that.
-
 A common approach is to use density estimation and use the expected probability

### 2011-06-20 thesis/conclusion.tex
Conclusion Chapter: very fast and highlevel writting.. needs long review.

--- 2011-05-24 thesis/conclusion.tex
@@
-\chapter{Conclusions and Future Work}
-\section*{}
+\chapter{Conclusions and Future Work}\label{chap:conclusions}
 
-In this thesis
+% Summary of developed work
+This thesis studied the problem of detecting novel situations where a robot lacks knowledge to
+correctly describe them.
+It did so on the area of semantic mapping on indoor spaces, by detecting that none of known room
+categories was able to correctly explain the sensed properties.
+
+For that it reviewed novelty detection and how an optimal detector can be implemented by
+thresholding. Showing after that, with the assumption of constant probability of seeing a
+novel case, an optimal ordering function for thresholding can be implemented based on
+the factor between a conditional and unconditional probability.
+
+It studied the semantic mapping process proposed by \cite{pronobis2011semmap} and presented a method
+to detect novel room categories based on probabilistic graphical models.
+Using a synthetic dataset, respecting the assumptions, it showed that such a method would be optimal
+if unconditional probability could be optimally approximated.
+Since in realistic conditions unconditional probability cannot be implemented due to the lack of
+knowing all the classes, some approaches were performed to approximate it with either a uniform
+assumption or by approximating with simplified models using unlabelled data.
+
+On the rest of this chapter the main results and conclusions are presented.
+Additionally limitations of the presented method and directions for future work are given.
+
+\section{Results and Conclusions}
+
+After studying the used semantic mapping process and studying novelty detection methods, this
+thesis proposed modelling a conditional and an unconditional probability distribution of sensed
+data using graphical models. Under the presented assumptions and assuming both can be approximated
+with perfect accuracy a ratio between those both probabilities would yield an optimal ordering
+function for implementing a novelty detector.
+
+This case, was showed to be optimal using a synthetic dataset that simulates a simplified
+environment of semantic categorization of room based on sense properties.
+Being the correct way to model the conditional probability with the probabilistic graphical model
+used by the semantic mapping process, this thesis moved forward with testing methods to approximate
+the unconditional probability. It tested the usage of a uniform distribution and explored the usage
+of unlabelled data to produce better models.
+
+Additionally it was studied how the performance of the presented methods change as more data
+is sensed from the environment. As expected all the methods increase their performance, but they
+move further from what would optimally be expected.
+Additionally the disadvantage between picking a uniform assumption and using an highly simplified
+model for the unlabelled data, starts reducing as more information is available.
+Nonetheless the model based on the usage of unlabelled data consistently lead to better detector
+performance, being a strong indicator to use unlabelled data whenever there is access to it.
+
+As a note, all the results presented on this thesis are directly reproducible from an
+online\footnote{\url{https://github.com/andresusanopinto/novelty-detection-thesis}}
+repository. That repository besides containing all the code and data for the results also
+includes tech-reports, presentations, articles and notes that have been produced during
+this thesis work. Additionally future research by the author will be correctly linked there,
+when appropriate.
+
+\section{Limitations}
+
+The presented methods use a strong assumption on a constant $P(novel)$, this is unrealistic and
+forbids one to exploit graph structural information that could show a given variable is more likely
+to be unknown to the system because the given graph structure is not easily explained by the
+current knowledge.
+An example of that can be seen as: if all the agent knows are two categories that are very often
+seen as a bi-colored graph, when presented with a non-bipartite graph, the agent should consider
+the likelihood of a new category greater than on bipartite graphs.
+
+Additionally, all the presented methods require calculation of $P(x|\overline{novel})$ and
+$P(x)$. This way a full model of the graphical model is needed and uncertain sensing as described in
+\autoref{sec:cues-from-low-level} cannot be easily incorporated.
+
+This two limitations are a single symptom of the decision on inverting the conditional probability
+$P(novel|x)$ to obtain a threshold that can be directly modelled by data.
+It is expected that by trying to model how the potentials in the factor graphs change between the
+conditional and unconditional graphical model, both limitations can be surpassed.
+
+\section{Future Work}
+
+Future work should try to study and or work around the limitations above presented. At the moment
+it is not clear on how the threshold changes when the assumption on constant $P(novel)$ cannot be
+made, and attempts to bypass the listed limitations will need to consider that by developing a
+method where a static threshold can have a more realistic and controlled behaviour through any graph
+structure.
+
+At a longer and larger scale the following paragraphs describe possible and interesting directions
+to exploit in the context of detecting knowledge gaps on artificial intelligence systems:
+
+% \subsection*{Extend Novelty to All Variable Types on the Graph}
+% Extend novelty to all categories on the graph (room, size, shape, etc..)
+% \subsection*{Handle Uncertain and Novel Sensing from Low-level Classifiers}
+% Handle uncertain sensing
+\subsubsection*{Generalized Framework}
+The presented method should be generalized by allowing novelty to be performed on any variable of
+graphical model. Additionally the novelty information could be incorporated back in the graph
+allowing the agent to probabilistically reason even when variables are considered unknown.
+That generalization should aim at being fully probabilistic such as the system presented by
+\cite{ranganathan2010pliss} and not deterministic by having to make decisions on which variables
+it considers novel.
+
+Additionally several methods exists that allow to produce novelty signals from the low-level
+classifiers. A generalized framework should try to fuse and handle all that information by
+incorporate it back in the graphical model in a similar fashion on how uncertain sensing is
+performed.
+
+% \subsection*{Explain Why a Certain Sample is Novel}
+% Explain why a certain sample is novel
+% \subsection*{Generation of most Likely Novel Samples}
+% Generate most likely novel sample
+\subsubsection*{Exploiting Generative Models}
+An interesting aspect that arises from the use of generative models is the ability to generate new
+samples according to it. It is expected that this can be exploited to achieve a better understanding
+of what the system is modelling and what understand common properties between several novel samples
+that allow to explain the novelty back to user. For example by generating the most
+likely novel samples, it maybe be possible to understand the limitations of its
+knowledge and use them for active learning.
+
+% \subsection*{Probabilistic Modelling of Space Segmentation}
+% \subsection*{Learn New Concepts}
+% Learn Concepts / hidden variable types - (areas, rooms, floors, environments, etc\dots)
+\subsubsection*{More than Novel Semantic Categories: Learning Graph Structures}
+Though this thesis only touched the problem of detecting novel semantic categories, other knowledge
+types should also be considered incomplete and are also candidates for novelty detection and
+modelling tools that allow to handle the associated uncertainty on them.
+More concretely, besides detecting novel semantic categories on given spatial concepts, there is
+also interest on detecting novel concepts: the presence of hidden variables of a new type, not known
+for the system: for example detecting room areas or types of environment.
+
+For that an initial step should be made using probabilistic graph structures to model the several
+possible space segmentations. Allowing to use a probabilistic approach instead of a deterministic
+approach where space is segmented based on the presence of doors or other landmarks.
+
+\subsubsection*{Beyond Detection of Knowledge Gaps}
+Having methods to detect gaps of knowledge is just one of the first step on creating long-term and
+life-long adaptable systems that are capable of learning through time. After detecting new
+situations, ways to learn and incorporating the knowledge back on the agent must be developed.

### 2011-06-20 thesis/abstract.tex
Small fixes

@@
 This representation identifies spatial entities and classifies them according to their meaning to
-humans allowing the robot to reason at a very high-level, closer to humans.
-For example, in a room scale level, humans categorize space in kitchens, bedrooms, corridors,
-office, theater, etc\dots, by relating room areas with those categories describing expected
-properties and actions that take place on them a robot can solve tasks like:
+humans allowing the robot to reason at a very high-level.
+For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
+office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
+properties and actions that take place on them, a robot can solve tasks like:
 ``Where are the cornflakes?''

### 2011-06-20 thesis/conclusion.tex
Small fixes

@@
 correctly describe them.
-It did so on the area of semantic mapping on indoor spaces, by detecting that none of known room
+It did so on the area of semantic mapping on indoor spaces, by detecting that none of the known room
 categories was able to correctly explain the sensed properties.
@@
 For that it reviewed novelty detection and how an optimal detector can be implemented by
-thresholding. Showing after that, with the assumption of constant probability of seeing a
-novel case, an optimal ordering function for thresholding can be implemented based on
+thresholding. Showing after, that with the assumption of constant probability of seeing a
+novel case an optimal ordering function for thresholding can be implemented based on
 the factor between a conditional and unconditional probability.
@@
 if unconditional probability could be optimally approximated.
-Since in realistic conditions unconditional probability cannot be implemented due to the lack of
+Since, in realistic conditions, unconditional probability cannot be implemented due to the lack of
 knowing all the classes, some approaches were performed to approximate it with either a uniform
@@
 
-After studying the used semantic mapping process and studying novelty detection methods, this
-thesis proposed modelling a conditional and an unconditional probability distribution of sensed
-data using graphical models. Under the presented assumptions and assuming both can be approximated
-with perfect accuracy a ratio between those both probabilities would yield an optimal ordering
-function for implementing a novelty detector.
+After studying the semantic mapping process and novelty detection methods, this
+thesis proposed modelling the conditional and unconditional probability distribution of sensed
+data using graphical models. Under the presented assumptions and assuming both can be exactly known
+a ratio between those probabilities would yield an optimal ordering function for implementing a
+novelty detector.
 
@@
 
-Additionally it was studied how the performance of the presented methods change as more data
+Additionally it has studied how the performance of the presented methods change as more data
 is sensed from the environment. As expected all the methods increase their performance, but they
-move further from what would optimally be expected.
+move further away from what could be obtained with exact information.
 Additionally the disadvantage between picking a uniform assumption and using an highly simplified

### 2011-06-21 thesis/novelty-intro.tex
Text explaining the method and reasoning behind

@@
 such that a threshold can be implemented.
+
+\section{Novelty Detection on the Conceptual Map}
+
+This section presents now how to use graphical models produced by the conceptual map and the
+novelty detection concepts introduced earlier on this chapter in order to detect room categories the
+system is not aware of.
+
+For that both models to approximate both the conditional and unconditional probability need to
+be defined and then the ratio between them used to define a function over which a threshold
+implements a novelty detection system.
+
+\subsection{Approximating the Conditional Probability}
+The actual semantic mapping of the environment happens using the \emph{conceptual map}
+introduced on \autoref{sec:conceptual-map}. During it the agent instantiates a
+\emph{chain graph} to model distribution of the sensed variables according to the known
+concepts and categories introduced as hidden variables on the graph. Using that graphical model
+the system is then able to propagate and find the most likely configuration of the represented
+variables, for instance the semantic category of a specific room is modelled as an hidden variable
+with states representing the semantic values, by calculating probabilities on that variable the
+system finds its belief on the room belonging to a specific category.
+
+Since the graphical model produced by the conceptual map tries to model the distribution of
+variables assuming the knowledge of the agent holds true, it represents the way to model
+$P(x|known)$, and allows to calculate the density probability that the given set of sensed
+features $x$ are sensed given that all the variables and graph structure, including the category of
+a specific room $a$ are correctly modelled by the agent knowledge.
+This way the distribution modelled by a factor graph equivalent to the chain-graph used by
+the conceptual map can be used as an approximation for the conditional probability of $x$.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [latent] {$a$}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \node (shape1f) [factor] {}; \&
+    \node (appearance1f) [factor] {}; \&
+    \node (object1f) [factor] {}; \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+
+  \draw [-] (room1) -- (shape1f) -- (shape1l);
+  \draw [-] (room1) -- (appearance1f) -- (appearance1l);
+  \draw [-] (room1) -- (object1f) -- (object1l);
+  \draw [-] (room2) -- (factor2) -- (prop2);
+
+  \draw [-] (room1) -- (room2) node (r12f) [midway,factor] {};
+  \draw [-] (room1) -- (room3) node (r13f) [midway,factor] {};
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+
+\caption{Factor graph modelling the conditional probability distribution, case where the
+         room category of room $a$ is known by the system.}
+\end{figure}
+
+\subsection{Approximating the Unconditional Probability}
+
+With no knowledge on the unconditional probability the correct approach is to
+assume a uniform distribution (\autoref{sec:max-entropy}).
+That is represented in factor graphs by a graph without any factors: \autoref{fig:uniform-graph}.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \node (x1) [obs] {$S_1$}; \&
+    \node (x2) [obs] {$A_1$}; \&
+    \node (x3) [obs] {$O_1$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$X_n$}; \\
+  };
+\end{tikzpicture}
+\caption{\label{fig:uniform-graph}A factor graph modelling a uniform
+         distribution over the sensed set of features $x$.}
+\end{figure}
+
+Thought, very often there is extra knowledge that can be obtained about the distribution
+of the variables that helps to model the unconditional distribution. With that, more accurate
+models can be produced.
+In this case, the system is interested in detecting only if a specific variable category is not
+known. With that in mind, it was assumed that an unknown category can only influence variables
+by using the same structure as the known variables. This is: it was assumed the structure of the
+graph was correct even when a room has an unknown category.
+
+\begin{figure}[h]
+\centering
+\subfloat[\label{fig:any-a}Since variable $a$ can be unknown, the agent does not has
+knowledge on how to model the dashed relations.]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \&
+    \&
+    \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [latent,dashed] {$a$}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \nofactor {shape1f}{}{} {right=0pt}; \&
+    \nofactor {appearance1f}{} {} {right=0pt}; \&
+    \nofactor {object1f}{} {} {right=0pt}; \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+  \draw [-] (shape1f) -- (shape1l);
+  \draw [-] (appearance1f) -- (appearance1l);
+  \draw [-] (object1f) -- (object1l);
+  \draw [-] (room2) -- (factor2) -- (prop2);
+  \draw [-,draw=none] (room1) -- (room2) node (r12f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (room3) node (r13f) [midway,factor] {};
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+  \draw [-] (room2) -- (r12f);
+  \draw [-] (room3) -- (r13f);
+
+  \draw [-,dashed] (room1) -- (shape1f);
+  \draw [-,dashed] (room1) -- (appearance1f);
+  \draw [-,dashed] (room1) -- (object1f);
+  \draw [-,dashed] (room1) -- (r12f);
+  \draw [-,dashed] (room1) -- (r13f);
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+}
+\qquad \qquad
+\subfloat[Without knowledge on all states of $a$, all variables dependent on $a$ get directly
+          dependent between each other, introducing a big single factor.]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [factor] {}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \&
+    \&
+    \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+
+  \draw [-] (room1) -- (shape1l);
+  \draw [-] (room1) -- (appearance1l);
+  \draw [-] (room1) -- (object1l);
+  \draw [-] (room2) -- (prop2);
+
+  \draw [-] (room1) -- (room2);
+  \draw [-] (room1) -- (room3);
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+}
+\caption{\label{fig:known-struct}Factor graph modelling distribution of sensed variables when variable $a$ can be unknown.}
+\end{figure}
+
+If the structure and all the other variables can be assumed to be correctly modelled by the
+knowledge of the agent, then all variables that were directly dependent on $a$ become directly
+dependent between them, introducing a single factor (\autoref{fig:known-struct}).
+By approximating this factor the agent can then obtain a model for the unconditional probability.
+
+In cases no other information is available the introduced factor can be considered uniform\footnote{Uniform factors do not influence the
+distribution represented by the factor graph and so can be removed from the graph representation}
+and dropped from the graph (see \autoref{fig:uniform-model}).
+Nonetheless it may be the case that due the cost of labelling data, the agent does not has knowledge
+on all the possible categories of $a$, but still has access to unlabelled data.
+In that case unlabelled samples may be gathered and explored to achieve a better approximation of
+the real distribution of variables.
+Though the factor that needs to be approximated requires handling of all the connected variables
+and enormous amounts of data may be needed to approximate it.
+
+A simple approach can be performed by assuming the variables connected to the factor to be independent
+but having a bias towards certain values. This approach is equivalent to connecting single factors
+on each of variables and approximating those factors using unlabelled data as seen on \autoref{fig:independent-model}.
+
+\begin{figure}[h]
+\centering
+\subfloat[\label{fig:uniform-model}Uniform Model]{
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [latent,draw=none] {}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \&
+    \&
+    \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+
+  \draw [-] (room2) -- (factor2) -- (prop2);
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+}
+\qquad
+\subfloat[\label{fig:independent-model}Independent Model] {
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=15pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room3) [latent] {}; \& \\
+    \& \& \& \\
+    \&
+    \node (room1) [latent,draw=none] {}; \& \&
+    \node (room2) [latent] {}; \\
+    \& \& \& \\
+    \node (shape1f) [factor] {}; \&
+    \node (appearance1f) [factor] {}; \&
+    \node (object1f) [factor] {}; \&
+    \node (factor2) [factor] {}; \\
+    \node (shape1l) [obs] {$S_p$}; \&
+    \node (appearance1l) [obs] {$A_p$}; \&
+    \node (object1l) [obs] {$O_p$}; \&
+    \node (prop2) [obs] {$X_p$}; \\
+  };
+
+  \draw [-] (shape1f) -- (shape1l);
+  \draw [-] (appearance1f) -- (appearance1l);
+  \draw [-] (object1f) -- (object1l);
+  \draw [-] (room2) -- (factor2) -- (prop2);
+
+  \draw [-,draw=none] (room1) -- (room2) node (r12f) [midway,factor] {};
+  \draw [-,draw=none] (room1) -- (room3) node (r13f) [midway,factor] {};
+  \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+  \draw [-] (room2) -- (r12f);
+  \draw [-] (room3) -- (r13f);
+
+  \begin{pgfonlayer}{background}
+    \plate{places1}{(shape1f)(shape1l)(object1l)}{$\forall p \in places(room1)$}{};
+    \plate{places2}{(factor2)(prop2)}{$\dots$}{};
+  \end{pgfonlayer}
+\end{tikzpicture}
+}
+
+\caption{Factor graph modelling the conditional probability distribution, case where the
+         room category of room $a$ is known by the system.}
+\end{figure}
+
+Other schemes may be tried to approximate the unconditional independent assuming that only
+a specific variable is not necessarily known. For example modelling unconditional probability
+with an hidden variable trained from the unlabelled data. Or by trying to split the single big
+factor on factorizations of other factors.
+Though the presented cases have interesting properties:
+
+\begin{description}
+\item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
+the graphical model and so special methods can be devised to calculate novelty on several variables
+by analysing the messages that are propagated in and out of each variable when performing inference.
+
+\item[independent model] - it can very easily be trained from unlabelled data. 
+\end{description}
+
+\subsection{Threshold}
+After defining a model for the conditional and unconditional distributions, a threshold can be applied
+over the ratio of those.
+
+For that
 

### 2011-06-22 thesis/novelty-intro.tex
Added comments on using a constant P(novel)

@@
 But in the case where the set of features $x$ is variable it cannot be
-discarded. There $P(x)$ also plays a role in levering all the conditional
+discarded. There $P(x)$ plays a role in levering all the conditional
 probabilities on different sets of variables into the same measure units
@@
 
+\subsection{Assumption on a constant $P(novel)$}
+The ratio between the conditional and unconditional probabilities of the sensed variables
+is only directly applicable to novelty detection under the assumption of a constant $P(novel)$.
+
+All the literature focus on novelty detection within a fixed scenario, where the set of variables
+used to model the distribution is fixed and defined at the moment the threshold is trained.
+Due to that, they can drop a constant $P(novel)$ and never need to directly or indirectly calculate
+it.
+To the best of the author knowledge there has not been studies on how to approximate it on dynamic
+sets and structures of sensed variables, and so a strong assumption on being constant through all
+the possible scenarios is used to allow dropping the factor.
+
+This assumption was considered questionable on whether it holds on a realistic scenario,
+and for that a sampling scheme able to comply with it is now presented as well an unbiased sampling
+scheme. A method to generate samples that match this \emph{criterium} of having a constant $P(novel)$ can be:
+
+\begin{algorithm}[Draw samples with equal $P(novel)$]
+\begin{enumerate}
+\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
+\item Decide novelty of $a$.
+\item Sample the remaining variables according to distribution $P_G(x|a)$.
+\end{enumerate}
+\end{algorithm}
+
+According to this sampling algorithm, given any structure the probability of
+variable $a$ being novel is constant. This is somehow unexpected as it means
+the sampling method is biased to adjust the real distribution $G$ such that
+exactly a constant percentage of drawn samples are novel.
+
+Alternative, the sampling method described below is considered more correct as
+it introduces no bias on the distribution $G$ describing the real distribution
+and classes of $a$:
+\begin{algorithm}[Draw samples with equal $P(novel)$]
+\begin{enumerate}
+\item Sample a graph structure $G$ (where variable $a$ can be any of the known or unknown classes)
+\item Sample variables $x$ according to distribution $P_G(x)$.
+\end{enumerate}
+\end{algorithm}
+
+Due to the highly bias needed to drawn a constant amount of novel sampes from any structure, the
+author defends this assumption to be very strong and points that future work should avoid.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Novelty Detection on the Conceptual Map}
@@
 concepts and categories introduced as hidden variables on the graph. Using that graphical model
-the system is then able to propagate and find the most likely configuration of the represented
-variables, for instance the semantic category of a specific room is modelled as an hidden variable
+the system is able to propagate and find the most likely configuration of the represented
+variables. For instance, the semantic category of a specific room is modelled as an hidden variable
 with states representing the semantic values, by calculating probabilities on that variable the
@@
 Since the graphical model produced by the conceptual map tries to model the distribution of
-variables assuming the knowledge of the agent holds true, it represents the way to model
-$P(x|known)$, and allows to calculate the density probability that the given set of sensed
-features $x$ are sensed given that all the variables and graph structure, including the category of
+variables assuming the knowledge of the agent holds true, it represents a method to model
+$P(x|known)$ since it allows to calculate the density probability that the set of features $x$ is
+sensed given that all the variables and graph structure, including the category of
 a specific room $a$ are correctly modelled by the agent knowledge.
@@
 With no knowledge on the unconditional probability the correct approach is to
-assume a uniform distribution (\autoref{sec:max-entropy}).
-That is represented in factor graphs by a graph without any factors: \autoref{fig:uniform-graph}.
+assume a uniform distribution (\autoref{sec:max-entropy}), which can be modelled by a
+factor graph without any factors: \autoref{fig:uniform-graph}.
 
@@
 In this case, the system is interested in detecting only if a specific variable category is not
-known. With that in mind, it was assumed that an unknown category can only influence variables
-by using the same structure as the known variables. This is: it was assumed the structure of the
-graph was correct even when a room has an unknown category.
+known. With that in mind, it was assumed that the graph structure is known and that an unknown
+category can only influence variables by using the same structure as the known variables.
+This way the structural information and all the other hidden variables the system is aware of
+can be used to more accurately approximate the unconditional distribution.
 
@@
 If the structure and all the other variables can be assumed to be correctly modelled by the
-knowledge of the agent, then all variables that were directly dependent on $a$ become directly
+knowledge of the agent, then all variables that are directly dependent on $a$ become directly
 dependent between them, introducing a single factor (\autoref{fig:known-struct}).
@@
     \&
-    \node (room1) [latent,draw=none] {}; \& \&
+    \node (room1) [factor] {}; \& \&
     \node (room2) [latent] {}; \\
@@
 
-  \draw [-] (room2) -- (factor2) -- (prop2);
+  \draw [-,dashed] (room1) -- (shape1l);
+  \draw [-,dashed] (room1) -- (appearance1l);
+  \draw [-,dashed] (room1) -- (object1l);
+  \draw [-] (room2) -- (prop2);
+
+  \draw [-,dashed] (room1) -- (room2);
+  \draw [-,dashed] (room1) -- (room3);
   \draw [-] (room2) -- (room3) node (r23f) [midway,factor] {};
+  \node (captroom1) [above=0pt of room1] {\footnotesize{U}};
 

### 2011-06-22 thesis/summary-en.tex
1st version of the summary of thesis in english.

%-----------------------------------------------
% Template para criação de resumos de projectos/dissertação
% jlopes AT fe.up.pt,   Fri Jul  3 11:08:59 2009
%-----------------------------------------------

\documentclass[9pt,a4paper]{extarticle}

%% English version: comment first, uncomment second
%\usepackage[portuguese]{babel}  % Portuguese
\usepackage[english]{babel}     % English
\usepackage{graphicx}           % images .png or .pdf w/ pdflatex OR .eps w/ latex
\usepackage{times}              % use Times type-1 fonts
\usepackage[utf8]{inputenc}     % 8 bits using UTF-8
\usepackage{url}                % URLs
\usepackage{multicol}           % twocolumn, etc
\usepackage{float}              % improve figures & tables floating
\usepackage[tableposition=top]{caption} % captions
%% English version: comment first (maybe)
%\usepackage{indentfirst}        % portuguese standard for paragraphs
%\usepackage{parskip}

%% page layout
\usepackage[a4paper,margin=30mm,noheadfoot]{geometry}

%% space between columns
\columnsep 12mm

%% headers & footers
\pagestyle{empty}

%% figure & table caption
\captionsetup{figurename=Fig.,tablename=Tab.,labelsep=endash,font=bf,skip=.5\baselineskip}

%% heading
\makeatletter
\renewcommand*{\@seccntformat}[1]{%
  \csname the#1\endcsname.\quad
}
\makeatother

%% avoid widows and orphans
\clubpenalty=300
\widowpenalty=300

\input{macros}

\begin{document}

\title{\vspace*{-8mm}\textbf{\textsc{Novelty Detection for Semantic Place Categorization}}}
\author{\emph{André Susano Pinto}$^\dag$\\[2mm]
        \small{Supervised by: Andrzej Pronobis$^\ddag$, Luis Paulo Reis$^\dag$}\\
        \small{$^\dag$Faculdade de Engenharia da Universidade do Porto, Portugal}\\
        \small{$^\ddag$Royal Institute of Technology (KTH), Stockholm, Sweden}}
%\url{andresusanopinto@gmail.com}\\
\date{}
\maketitle
%no page number 
\thispagestyle{empty}

\vspace*{-4mm}\noindent\rule{\textwidth}{0.4pt}\vspace*{4mm}

\begin{multicols}{2}

\section{Motivation}
For a long time humanity has fantasized that one day robots will walk among us.
They will move and be able to interact with us, understand our concepts and
be able to reason.
An important step on that direction is endowing robots with knowledge of human concepts and
semantics. However its unrealistic to believe on the possibility of extensively
describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
to adapt and learn when confronted with novel situations.
Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
role for the subsequent adapting and learning of new concepts and its this thesis main topic.

Additionally in the context of mobile robotics, spatial concepts and semantics play a main role to perform
complex human-like tasks and human interactions. For handling those, a robot builds a representation
of space extended with semantic properties, process which is known as semantic mapping.
This representation identifies spatial entities and classifies them according to their meaning to
humans allowing the robot to reason at a very high-level.
For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
properties and actions that take place on them, a robot can solve tasks like:
``Where are the cornflakes?''

\section{Problem and Goals}
This thesis focus then on the particular problem of identifying novel semantic categories for places
using
This thesis studies a proposed %~\cite{pronobis2011semmap}
process that utilizes spatial and semantic knowledge for semantic mapping using a
probabilistic graphical model,
and develops methods to perform detection of room categories not previously
known to the robot using that graphical model.

\section{Semantic Mapping}
Semantic mapping is the process by which an agent builds a representation of the environment and
enhances the identified entities with semantic values (i.e.\ human semantics).
In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} was used as based.

\subsection{Spatial Knowledge Representation}
In order to perform semantic mapping an agent needs to possess knowledge on how to segment, identify
entities existent in the space by analysing the sensed properties. For that Pronobis~\cite{pronobis2011semmap} presents a
layered knowledge structure:
the sensory layer maintains an accurate and exact description of environment in a machine-friendly representation,
the place layer employs knowledge on how to segment and identify areas such as rooms,
the categorical layer defines knowledge on properties and categories that allow to extract
higher-level information from the sensed data
and the conceptual layer maintains knowledge on entities and their relations between them as well
information on how to map those to human semantics.

\subsection{Conceptual Map}
The semantic mapping is then based on the instantiation of a conceptual map that uses all the
ontology defined on the conceptual layer in order to build a probabilistic graphical model that
relates and propagates probabilities across all the entities detect and properties detected by the
other layers.

By using probabilistic graphical models, the conceptual map becomes then a generative model that
allows several inferences on variable configurations to be performed and for that serves as a useful
tool for novelty detection of unknown categories of some of its variables.

\section{Novelty Detection}
Novelty detection, also known as outlier or anomaly detection, is a
classification problem related to identification of new or unknown data
patterns that the system is not aware of~\cite{markou2003novelty}.
The ability to identify novel cases is crucial in any autonomous system
that is deployed to unknown or to uncontrolled environments, as it gives the
system the ability to detect that something is not conforming to its knowledge and
therefore should be treated with caution.
% It has several applications such as fault detection~\cite{tarassenko1999novelty},
% intrusion detection~\cite{fan2001using},
% detection of masses in mammograms~\cite{tarassenko1995novelty} or detection of
% novel and useful documents~\cite{zhang2002novelty}.

\subsection{Novelty Detection as Threshold}
Due to the noisy, uncertain and insufficient nature of sensed data, novelty ought to be
treated in a probabilistic way in where each sample has a certain probability
of being generated by a class not known for the agent and a complementary probability $P(\overline{novel}|x)$
of being generated by a known class.

Additionally any detector can be uniquely described by the set $N$ of samples that it
classifies as novel. With that its possible to show that by including a new sample $a$ in $N$ an agent
can increase its detection rate at a cost of producing more mis-classifications.
This describes the base of the \emph{error and rejection tradeoff} described by \cite{chow1970optimum}.
Due to the interest of having the highest detection rate with a given minimum error rate, any optimal
detector can be described as a \emph{continuous knapsack problem} of the sample space.
The greedy ordering defined to solve it can be shown to be equivalent to:
\begin{eqnarray}
P(\overline{novel}|x) &=& \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)}
\end{eqnarray}

\subsection{Assumption on a constant $P(novel)$}
The ratio introduced on the previous section allows to implement an optimal detector.
Though the factors involved need to be defined according to available data.

$P(\overline{novel})$ was considered constant through all possible situations. Although this
sounds a weak assumption note that in the case hat uses the ratio

\subsection{Modeling Conditional Probability}
Since the graphical model produced by the conceptual map tries to model the distribution of
variables assuming the knowledge of the agent holds true, it represents a method to model
$P(x|\overline{novel})$ since it allows to calculate the density probability that the set of features $x$ is
sensed given that all the variables and graph structure, including the category of
a specific room $a$ are correctly modelled by the agent knowledge.
This way the distribution modelled by a factor graph equivalent to the chain-graph used by
the conceptual map can be used as an approximation for the conditional probability of $x$.

\subsection{Modeling Unconditional Probability}
With no knowledge on the real distribution of all samples, the correct approach is to model it
with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
category of a specific room is unknown and that the remaining knowledge holds trues.
In that case, the correct approach is to replace the variable $a$, that represents the room where to
detect novelty, with a single factor connecting all variables directly dependent on it.

Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
approximations have to be done. In this thesis work two approximations where tried: an uniform model
and an independent model as illustrated in \autoref{fig:models}. In the independent model, the
single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
approximation to the unconditional probability than an assumption on the uniform model.

\section{Results}
A synthetic dataset modelling the expected features when implementing on the real case and enforcing
a constant $P(novel)$ was used for testing the developed models.
The two suggested models where compared between them and to an optimal case across several situations
showing they behaviour as more information is sensed from the environment.

\begin{figure}[H]
\centering
\includegraphics[width=0.4\textwidth]{results/synthetic-all.pdf}
\caption{\label{fig:results}Comparison of the proposed models using samples from a synthetic dataset
         with variable number of sensed features.}
\end{figure}

\section{Conclusions}
This thesis studied the problem of novelty detection of semantic categories on the spatial knowledge
of mobile robots. To that purpose a semantic mapping process~\cite{pronobis2011semmap} was studied and
a novelty detection scheme based on the conceptual map used by it proposed.
Initial tests where performed on synthetic data showing an interest in using unlabelled data to
increase the detection performance.
Additionally all the assumptions and bases behind novelty detection were presented and several
directions on how to create a generalized framework for novelty detection were discussed.

\bibliographystyle{unsrt}
\bibliography{refs}

\end{multicols}

\end{document}

### 2011-06-22 thesis/summary-en.tex
Changes on the summary

@@
 \section{Problem and Goals}
-This thesis focus then on the particular problem of identifying novel semantic categories for places
-using
-This thesis studies a proposed %~\cite{pronobis2011semmap}
-process that utilizes spatial and semantic knowledge for semantic mapping using a
-probabilistic graphical model,
-and develops methods to perform detection of room categories not previously
-known to the robot using that graphical model.
+The semantic mapping process plays an important role for allowing agents to reason at an high-level
+and is totally defined by the knowledge the agent has on the spatial concepts and semantic categories.
+For that reason there is an high interest in creating methods that are able to alert the agent for
+gaps on its knowledge.
+And with that purpose, this thesis, studies novelty detection techniques and how to apply them to
+detect novel room categories during a semantic mapping process.
+To this end, the semantic mapping process proposed by \cite{pronobis2011phd} is studied and a method
+to detect novel categories using graphical models is suggested.
 
@@
 enhances the identified entities with semantic values (i.e.\ human semantics).
-In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} was used as based.
+In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} is used as base.
 
 \subsection{Spatial Knowledge Representation}
-In order to perform semantic mapping an agent needs to possess knowledge on how to segment, identify
-entities existent in the space by analysing the sensed properties. For that Pronobis~\cite{pronobis2011semmap} presents a
-layered knowledge structure:
-the sensory layer maintains an accurate and exact description of environment in a machine-friendly representation,
-the place layer employs knowledge on how to segment and identify areas such as rooms,
-the categorical layer defines knowledge on properties and categories that allow to extract
+In order to perform semantic mapping an agent needs to possess knowledge on how to segment and
+identify entities existent in space by analysing the sensed data. For that
+Pronobis~\cite{pronobis2011semmap} presents a layered knowledge structure:
+the \emph{sensory layer} maintains an accurate and exact description of environment in a machine-friendly representation,
+the \emph{place layer} employs knowledge on how to segment and identify areas such as rooms,
+the \emph{categorical layer} defines knowledge on properties and categories that allow to extract
 higher-level information from the sensed data
-and the conceptual layer maintains knowledge on entities and their relations between them as well
-information on how to map those to human semantics.
+and the \emph{conceptual layer} maintains knowledge on entities concepts, the relations between them
+and information on how to map them to human semantics.
 
@@
 ontology defined on the conceptual layer in order to build a probabilistic graphical model that
-relates and propagates probabilities across all the entities detect and properties detected by the
+relates and propagates probabilities across all the entities and properties detected by the
 other layers.
@@
 By using probabilistic graphical models, the conceptual map becomes then a generative model that
-allows several inferences on variable configurations to be performed and for that serves as a useful
-tool for novelty detection of unknown categories of some of its variables.
+allows to perform inference on variable configurations. For that it is used as a base tool
+for novelty detection of unknown categories on some of its variables.
 
@@
 \subsection{Assumption on a constant $P(novel)$}
-The ratio introduced on the previous section allows to implement an optimal detector.
-Though the factors involved need to be defined according to available data.
-
-$P(\overline{novel})$ was considered constant through all possible situations. Although this
-sounds a weak assumption note that in the case hat uses the ratio
+By considering $P(novel)$ to be constant a ratio between conditional and unconditional probabilities
+of the sensed variables is obtained.
+Although it sounds reasonable to assume the probability of specific room belong to a novel
+category is constant, this thesis points that this assumption might be very strong and not
+reasonable in a realistic scenario. The main point being that in this case the graph structure
+changes between samples and some might be more prone to produce novel samples, as opposed to normal
+classification scenarios where its clear that the category is indeed the only generator of features.
 
@@
 variables assuming the knowledge of the agent holds true, it represents a method to model
-$P(x|\overline{novel})$ since it allows to calculate the density probability that the set of features $x$ is
-sensed given that all the variables and graph structure, including the category of
-a specific room $a$ are correctly modelled by the agent knowledge.
-This way the distribution modelled by a factor graph equivalent to the chain-graph used by
-the conceptual map can be used as an approximation for the conditional probability of $x$.
+$P(x|\overline{novel})$.
+%since it allows to calculate the density probability that the set of features $x$ is
+%sensed given that all the variables and graph structure, including the category of
+%a specific room $a$ are correctly modelled by the agent knowledge.
+%This way the distribution modelled by a factor graph equivalent to the chain-graph used by
+%the conceptual map can be used as an approximation for the conditional probability of $x$.
 
 \subsection{Modeling Unconditional Probability}
-With no knowledge on the real distribution of all samples, the correct approach is to model it
+With no knowledge on the real distribution samples, the correct approach is to model it
 with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
 category of a specific room is unknown and that the remaining knowledge holds trues.
-In that case, the correct approach is to replace the variable $a$, that represents the room where to
-detect novelty, with a single factor connecting all variables directly dependent on it.
+In that case, the correct approach is to replace the variable that represents the room where to
+detect novelty with a single factor connecting all variables directly dependent on it.
 
 Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
-approximations have to be done. In this thesis work two approximations where tried: an uniform model
-and an independent model as illustrated in \autoref{fig:models}. In the independent model, the
+approximations have to be done. In this thesis two approximations were presented: an uniform model
+and an independent model. In the independent model, the
 single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
@@
 a constant $P(novel)$ was used for testing the developed models.
-The two suggested models where compared between them and to an optimal case across several situations
-showing they behaviour as more information is sensed from the environment.
+The two suggested models where compared between them and to an optimal case.
+The tests across several situations shows how they behaviour change as more information is sensed
+from the environment.
 

### 2011-06-23 thesis/summary-pt.tex
Added portuguese sumary version and scons rules.

--- 2011-06-22 thesis/summary-en.tex
@@
 %% English version: comment first, uncomment second
-%\usepackage[portuguese]{babel}  % Portuguese
-\usepackage[english]{babel}     % English
+\usepackage[portuguese]{babel}  % Portuguese
+%\usepackage[english]{babel}     % English
 \usepackage{graphicx}           % images .png or .pdf w/ pdflatex OR .eps w/ latex
@@
 %% English version: comment first (maybe)
-%\usepackage{indentfirst}        % portuguese standard for paragraphs
+\usepackage{indentfirst}        % portuguese standard for paragraphs
 %\usepackage{parskip}
@@
 \author{\emph{André Susano Pinto}$^\dag$\\[2mm]
-        \small{Supervised by: Andrzej Pronobis$^\ddag$, Luis Paulo Reis$^\dag$}\\
+        \small{Sob orientação de: Andrzej Pronobis$^\ddag$, Luis Paulo Reis$^\dag$}\\
         \small{$^\dag$Faculdade de Engenharia da Universidade do Porto, Portugal}\\
@@
 \section{Motivation}
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us, understand our concepts and
-be able to reason.
-An important step on that direction is endowing robots with knowledge of human concepts and
-semantics. However its unrealistic to believe on the possibility of extensively
-describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
-to adapt and learn when confronted with novel situations.
-Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
-role for the subsequent adapting and learning of new concepts and its this thesis main topic.
+Já à muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
+Acredita-se que eles serão capaz de ser mover e interagir connosco, compreender os nossos conceitos
+e serem capazes de pensar.
+Um passo importante nesta direcção é a criação de robôs com conhecimento de conceitos e significados
+próprios do mundo humano.
+No entanto é irrealista acreditar que será possivel descrever e representar em toda a sua extensão
+tais conceitos. Assim os agentes deverão ser capazes de aprender e adaptar o seu conhecimento quando
+confrontados com novas situações.
+A detecção destas novas situações onde o conhecimento de um agente não é suficiente têm um papel importante
+na adaptação e aprendizagem de novos conceitos e significados e é por isso o tópico principal desta tese.
 
-Additionally in the context of mobile robotics, spatial concepts and semantics play a main role to perform
-complex human-like tasks and human interactions. For handling those, a robot builds a representation
-of space extended with semantic properties, process which is known as semantic mapping.
-This representation identifies spatial entities and classifies them according to their meaning to
-humans allowing the robot to reason at a very high-level.
-For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
-office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
-properties and actions that take place on them, a robot can solve tasks like:
-``Where are the cornflakes?''
+Addicionalmente no contêxto de agentes moveis, para facilitar o deslocamento e planeamento, os
+agentes constroiem uma representação do ambiente na qual detectam conceitos e propriedades
+conceitos espaciais.
+Essa representação identifica conceitos e classifica-os de acordo com significados humanos,
+permitindo assim ao agente obter uma representação dos conceitos humanos e pensar e planear a um
+nível próximo dos mesmos.
+Ao processo de construção desta representação é attribuido o nome de \emph{semantic mapping}.
+Por exemplo, a uma escala de quartos, os humanos categorizam o espaço em: cozinhas, quartos, corredores,
+escritórios, sala de computadores, etc\dots
+Ao relacionar quartos com estas categorias que descrevem propriedades e acções percebidas por
+humanos, um agente consegue resolver tarefas como: ``Onde estão os ceriais?''
 
-\section{Problem and Goals}
-The semantic mapping process plays an important role for allowing agents to reason at an high-level
-and is totally defined by the knowledge the agent has on the spatial concepts and semantic categories.
-For that reason there is an high interest in creating methods that are able to alert the agent for
-gaps on its knowledge.
-And with that purpose, this thesis, studies novelty detection techniques and how to apply them to
-detect novel room categories during a semantic mapping process.
-To this end, the semantic mapping process proposed by \cite{pronobis2011phd} is studied and a method
-to detect novel categories using graphical models is suggested.
+\section{Problema e Objectivos}
+Como deliniado na motivação, o processo de \emph{semantic mapping} têm um papel importante na
+capacidade pensativa de um agente. Em simultâneo, esse processo está limitado pelo conhecimento
+do agente em conceitos e significados do mundo humano. É então importante estudar e analisar
+técnicas que sejam capazes de detectar falhas no conhecimento e permitir ao agente agir de maneira
+cuidadosa.
+É com esse objectivo que a tese aqui descrita se foca no estudo de deteção
+de situações novas num processo de \emph{semantic mapping} propôsto por \cite{pronobis2011phd}.
+Para esse fim a tese estuda como usar modelos graphicos para detectar novas categorias nas conhecimento
+de categorias do agente.
 
-\section{Semantic Mapping}
-Semantic mapping is the process by which an agent builds a representation of the environment and
-enhances the identified entities with semantic values (i.e.\ human semantics).
-In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} is used as base.
+\section{Mapeamento Semântico}
+Mapeamento semântico é o processo através do qual um agente constroi uma representação do ambiente
+extendida com conceitos e categories com significado especial (i.e.\ conceitos humanos).
+Neste tese o processo proposto por \cite{pronobis2011phd} foi usado como base.
 
-\subsection{Spatial Knowledge Representation}
-In order to perform semantic mapping an agent needs to possess knowledge on how to segment and
-identify entities existent in space by analysing the sensed data. For that
-Pronobis~\cite{pronobis2011semmap} presents a layered knowledge structure:
-the \emph{sensory layer} maintains an accurate and exact description of environment in a machine-friendly representation,
-the \emph{place layer} employs knowledge on how to segment and identify areas such as rooms,
-the \emph{categorical layer} defines knowledge on properties and categories that allow to extract
-higher-level information from the sensed data
-and the \emph{conceptual layer} maintains knowledge on entities concepts, the relations between them
-and information on how to map them to human semantics.
+\subsection{Conhecimento Espacial}
+O conhecimento espacial de agente define todo a informação necessária para o mapeamento semântico,
+e inclui informação em como segmentar e identificar entidades e propriedades através dos dados
+recolhidos.
+Para esse fim Pronobis~\cite{pronobis2011semmap} utiliza uma representação estruturada em camadas:
+a \emph{camada de sensores} é responsável por manter uma descrição exacta do ambiente numa representação
+de baixo nível;
+a \emph{camada de lugares} é responsável por segmentar e identificar areas e conceitos
+como quartos;
+a \emph{camada de categories} é responsável por definir conhecimento sobre propriadades e categories
+e permite ao agente obter informação de mais alto-nível;
+e a \emph{camada de conceitos} mantêm conhicimento sobre os diversoes conceitos, relações sobre os mesmos
+e informação de como os mapear para conceitos humanos.
 

### 2011-06-23 thesis/novelty-intro.tex
Notes on how to train a threshold.

@@
 \section{Novelty Detection by Thresholding}
+\label{sec:threshold}
 The objective of a novelty detection system is to classify a given sample $x$ as
@@
 \item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
-the graphical model and so special methods can be devised to calculate novelty on several variables
-by analysing the messages that are propagated in and out of each variable when performing inference.
+the graphical model.
 
-\item[independent model] - it can very easily be trained from unlabelled data. 
+\item[independent model] - due to the use of single connected factors, it can easily be trained from
+                           unlabelled data. This helps the system to account for an existent bias on
+                           each of the variables without allowing the system to overfit to the unlabelled
+                           data.
 \end{description}
@@
 over the ratio of those.
+Under the assumption of a constant $P(novel)$, defining a threshold over it leads to an optimal novelty
+detector, as showed on \autoref{sec:threshold}.
 
-For that
+If the training data does not contains any bias, by picking a threshold on the training data that achieves
+a certain true-positive and false-positive rate, it is expected that on the real distribution such a system
+will achieve the same performance.
+
+Note however that the performance of the system is dependent on how well the models approximate the conditional
+and unconditional distributions and so are dependent on how the graph structure is able to model the real distribution
+and how well the training data allows the factors to be approximated.
 

### 2011-06-23 thesis/summary-en.tex
Corrections in the summary-en

@@
 \section{Motivation}
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us, understand our concepts and
-be able to reason.
-An important step on that direction is endowing robots with knowledge of human concepts and
-semantics. However its unrealistic to believe on the possibility of extensively
-describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
-to adapt and learn when confronted with novel situations.
-Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
-role for the subsequent adapting and learning of new concepts and its this thesis main topic.
+For a long time humanity has dreamed that one day robots will be among us. They will explore our world
+and interact with us, understand our concepts and reason. An important step in that direction is endowing 
+robots with knowledge about human concepts and semantics. However, it is unrealistic to believe that the 
+human world can be fully modeled in the robot's brain at the design stage. Therefore, robots must be able 
+to adapt and learn when confronted with novel situations. Detection of novel situations, where the knowledge
+of the robot is not sufficient plays an important role in adaptation and learning of new concepts and 
+is the main topic of this thesis.
 
-Additionally in the context of mobile robotics, spatial concepts and semantics play a main role to perform
+In the context of mobile robotics, spatial concepts and semantics are crucial to enable the robot to perform
 complex human-like tasks and human interactions. For handling those, a robot builds a representation
@@
 This representation identifies spatial entities and classifies them according to their meaning to
-humans allowing the robot to reason at a very high-level.
-For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
-office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
-properties and actions that take place on them, a robot can solve tasks like:
-``Where are the cornflakes?''
+humans allowing the robot to reason at a high abstraction level. For example, humans categorize spaces as 
+kitchens, bedrooms, corridors, offices, theaters, computer labs, etc. based on spatial properties, 
+objects and actions that are characteristic of those spaces. Similarly, if a robot had a knowledge about
+those spatial concepts, it could answer such questions as: ``Where are cornflakes?''
 
 \section{Problem and Goals}
-The semantic mapping process plays an important role for allowing agents to reason at an high-level
-and is totally defined by the knowledge the agent has on the spatial concepts and semantic categories.
-For that reason there is an high interest in creating methods that are able to alert the agent for
-gaps on its knowledge.
-And with that purpose, this thesis, studies novelty detection techniques and how to apply them to
-detect novel room categories during a semantic mapping process.
-To this end, the semantic mapping process proposed by \cite{pronobis2011phd} is studied and a method
+The semantic mapping process plays an important role for allowing artificial mobile agents to reason at 
+a high level and is dependent on the knowledge that the agent has about the spatial concepts and semantic 
+categories. For that reason there is high interest in creating methods that are capable of detecting 
+gaps in the agent's knowledge. For that purpose, this thesis, studies novelty detection techniques and 
+how they can be applied to detection novel room categories during the semantic mapping process.
+To this end, the semantic mapping process proposed in \cite{pronobis2011phd} is studied and a method
 to detect novel categories using graphical models is suggested.
@@
 Semantic mapping is the process by which an agent builds a representation of the environment and
-enhances the identified entities with semantic values (i.e.\ human semantics).
-In this thesis the semantic mapping process proposed by \cite{pronobis2011phd} is used as base.
+augments the identified spatial entities with semantic values (i.e.\ human semantics).
+In this thesis the semantic mapping process proposed in \cite{pronobis2011phd} is used as base.
 
@@
 In order to perform semantic mapping an agent needs to possess knowledge on how to segment and
-identify entities existent in space by analysing the sensed data. For that
-Pronobis~\cite{pronobis2011semmap} presents a layered knowledge structure:
-the \emph{sensory layer} maintains an accurate and exact description of environment in a machine-friendly representation,
-the \emph{place layer} employs knowledge on how to segment and identify areas such as rooms,
-the \emph{categorical layer} defines knowledge on properties and categories that allow to extract
-higher-level information from the sensed data
-and the \emph{conceptual layer} maintains knowledge on entities concepts, the relations between them
-and information on how to map them to human semantics.
+identify entities existent in space by analysing the sensed data. For that, 
+Pronobis~\cite{pronobis2011semmap} presents a spatial knowledge representation consisting of four layers:
+the \emph{sensory layer} maintains an accurate and exact description of the environment in a machine-friendly 
+representation, the \emph{place layer} represents the world in terms of a set of discrete places and connectivity
+between them, the \emph{categorical layer} defines knowledge about properties and categories that allow to extract
+higher-level information from the sensed data, and the \emph{conceptual layer} maintains knowledge about 
+spatial entities and concepts as well as the relations between them.
 
 \subsection{Conceptual Map}
-The semantic mapping is then based on the instantiation of a conceptual map that uses all the
-ontology defined on the conceptual layer in order to build a probabilistic graphical model that
+The semantic mapping is based on the instantiation of a conceptual map that uses the
+ontology defined in the conceptual layer in order to build a probabilistic graphical model that
 relates and propagates probabilities across all the entities and properties detected by the
@@
 
-By using probabilistic graphical models, the conceptual map becomes then a generative model that
-allows to perform inference on variable configurations. For that it is used as a base tool
-for novelty detection of unknown categories on some of its variables.
+By using probabilistic graphical models, the conceptual map becomes a generative model that
+permits inference about different variables. This enables the use of the conceptual map 
+as a tool for detection of unknown categories.
 
@@
 The ability to identify novel cases is crucial in any autonomous system
-that is deployed to unknown or to uncontrolled environments, as it gives the
-system the ability to detect that something is not conforming to its knowledge and
+that is deployed in unknown or uncontrolled environments, as it permits 
+detection of cases that do not conform to the robot's knowledge and
 therefore should be treated with caution.
@@
 
-\subsection{Novelty Detection as Threshold}
-Due to the noisy, uncertain and insufficient nature of sensed data, novelty ought to be
-treated in a probabilistic way in where each sample has a certain probability
-of being generated by a class not known for the agent and a complementary probability $P(\overline{novel}|x)$
+\subsection{Novelty Detection as Thresholding}
+Due to the nature of the sensed data which are noisy and uncertain, novelty ought to be
+treated in a probabilistic way where each sample has certain probability
+of being generated by a class not known to the agent and a complementary probability $P(\overline{novel}|x)$
 of being generated by a known class.
 
-Additionally any detector can be uniquely described by the set $N$ of samples that it
-classifies as novel. With that its possible to show that by including a new sample $a$ in $N$ an agent
-can increase its detection rate at a cost of producing more mis-classifications.
-This describes the base of the \emph{error and rejection tradeoff} described by \cite{chow1970optimum}.
-Due to the interest of having the highest detection rate with a given minimum error rate, any optimal
+Additionally, any detector can be uniquely described by the set $N$ of samples that it
+classifies as novel. With that, it is possible to show that by including a new sample $a$ into 
+$N$, an agent can increase its detection rate at a cost of producing more misclassifications.
+This describes the base of the \emph{error and rejection tradeoff} described in \cite{chow1970optimum}.
+Due to the goal of having the highest detection rate with a given minimum error rate, any optimal
 detector can be described as a \emph{continuous knapsack problem} of the sample space.
-The greedy ordering defined to solve it can be shown to be equivalent to:
+The resulting greedy ordering can be shown to be equivalent to:
 \begin{eqnarray}
@@
 
-\subsection{Assumption on a constant $P(novel)$}
-By considering $P(novel)$ to be constant a ratio between conditional and unconditional probabilities
-of the sensed variables is obtained.
-Although it sounds reasonable to assume the probability of specific room belong to a novel
-category is constant, this thesis points that this assumption might be very strong and not
-reasonable in a realistic scenario. The main point being that in this case the graph structure
-changes between samples and some might be more prone to produce novel samples, as opposed to normal
-classification scenarios where its clear that the category is indeed the only generator of features.
+\subsection{Assumption about Constant $P(novel)$}
+By considering $P(novel)$ to be constant, a ratio between conditional and unconditional probabilities
+of the sensed variables is obtained. Although it sounds reasonable to assume that the probability of a specific 
+room belonging to a novel category is constant, this thesis points out that this assumption might be very strong 
+and not reasonable in realistic scenarios. The main problem is related to the fact that the graph structure
+changes as the robot gathers new samples and some structures might be more prone to produce novel samples. This
+is different from the normal classification scenarios where its clear that the category is indeed the only 
+generator of features.
 
@@
 Since the graphical model produced by the conceptual map tries to model the distribution of
-variables assuming the knowledge of the agent holds true, it represents a method to model
-$P(x|\overline{novel})$.
+variables assuming that the knowledge of the agent is representative of the world, it models $P(x|\overline{novel})$.
 %since it allows to calculate the density probability that the set of features $x$ is
@@
 \subsection{Modeling Unconditional Probability}
-With no knowledge on the real distribution samples, the correct approach is to model it
+With no knowledge about the real distribution of samples, the correct approach is to model it
 with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
-category of a specific room is unknown and that the remaining knowledge holds trues.
-In that case, the correct approach is to replace the variable that represents the room where to
-detect novelty with a single factor connecting all variables directly dependent on it.
+category of a specific room is unknown and that the remaining knowledge holds true.
+In that case, the correct approach is to replace the variable that represents the room where the 
+novelty is detected with a single factor connecting all variables directly dependent on it.
 
 Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
-approximations have to be done. In this thesis two approximations were presented: an uniform model
-and an independent model. In the independent model, the
-single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
-approximation to the unconditional probability than an assumption on the uniform model.
+approximations have to be done. In this thesis two approximations were presented: a uniform model
+and an independent model. In the independent model, the single\hyp{}connected factors are trained from 
+unlabelled data and its expected that they create a better approximation to the unconditional probability 
+than an assumption about the uniform model.
 
 \section{Results}
-A synthetic dataset modelling the expected features when implementing on the real case and enforcing
-a constant $P(novel)$ was used for testing the developed models.
-The two suggested models where compared between them and to an optimal case.
-The tests across several situations shows how they behaviour change as more information is sensed
-from the environment.
+A synthetic dataset modelling the features expected when deployed in a real scenario and enforcing
+a constant $P(novel)$ was used for testing the developed models. The two suggested models where compared 
+to each other and to an optimal case. The tests across several situations show how the behaviour changes 
+as more information is sensed from the environment.
 
@@
 \section{Conclusions}
-This thesis studied the problem of novelty detection of semantic categories on the spatial knowledge
-of mobile robots. To that purpose a semantic mapping process~\cite{pronobis2011semmap} was studied and
-a novelty detection scheme based on the conceptual map used by it proposed.
-Initial tests where performed on synthetic data showing an interest in using unlabelled data to
-increase the detection performance.
-Additionally all the assumptions and bases behind novelty detection were presented and several
-directions on how to create a generalized framework for novelty detection were discussed.
+This thesis studied the problem of detection of novel semantic categories of rooms in the context of
+semantic mapping with mobile robots. To that purpose, a semantic mapping process~\cite{pronobis2011semmap} 
+was studied and a novelty detection scheme based on a conceptual map was proposed.
+Initial tests where performed on synthetic data showing benefit of using unlabelled data to
+increase the detection performance. Additionally, all the assumptions behind novelty detection were 
+presented and several directions on how to create a generalized framework for novelty detection 
+were discussed.
 

### 2011-06-23 thesis/abstract.tex
Corrected the abstract

@@
 \chapter*{Abstract}
-For a long time humanity has fantasized that one day robots will walk among us.
-They will move and be able to interact with us, understand our concepts and
-be able to reason.
-An important step on that direction is endowing robots with knowledge of human concepts and
-semantics. However its unrealistic to believe on the possibility of extensively
-describing human concepts and semantics to robots on a pre-deploy phase, and so, robots must be able
-to adapt and learn when confronted with novel situations.
-Detection of novel situations, where the robot knowledge is not sufficient, plays then an important
-role for the subsequent adapting and learning of new concepts and its this thesis main topic.
 
-In the context of mobile robotics, spatial concepts and semantics play a main role to perform
+For a long time humanity has dreamed that one day robots will be among us. They will explore our world
+and interact with us, understand our concepts and reason. An important step in that direction is endowing 
+robots with knowledge about human concepts and semantics. However, it is unrealistic to believe that the 
+human world can be fully modeled in the robot's brain at the design stage. Therefore, robots must be able 
+to adapt and learn when confronted with novel situations. Detection of novel situations, where the knowledge
+of the robot is not sufficient plays an important role in adaptation and learning of new concepts and 
+is the main topic of this thesis.
+
+In the context of mobile robotics, spatial concepts and semantics are crucial to enable the robot to perform
 complex human-like tasks and human interactions. For handling those, a robot builds a representation
@@
 This representation identifies spatial entities and classifies them according to their meaning to
-humans allowing the robot to reason at a very high-level.
-For example, in a room scale, humans categorize space in kitchens, bedrooms, corridors,
-office, theater, computer lab, etc\dots, by relating room areas with those categories, that describing expected
-properties and actions that take place on them, a robot can solve tasks like:
-``Where are the cornflakes?''
+humans allowing the robot to reason at a high abstraction level. For example, humans categorize spaces as 
+kitchens, bedrooms, corridors, offices, theaters, computer labs, etc. based on spatial properties, 
+objects and actions that are characteristic of those spaces. Similarly, if a robot had a knowledge about
+those spatial concepts, it could answer such questions as: ``Where are cornflakes?''
 
-This thesis studies a proposed %~\cite{pronobis2011semmap}
-process that utilizes spatial and semantic knowledge for semantic mapping using a
-probabilistic graphical model,
-and develops methods to perform detection of room categories not previously
-known to the robot using that graphical model.
-At last, it draws attention to the use of unlabelled data on the novelty detection process,
-which can be used to considerably increase performance.
+This thesis studies a semantic mapping process that employs spatial semantic knowledge
+and represents it using a probabilistic graphical model. It develops methods 
+for performing detection of novel room categories which were not previously known to the robot applicable 
+to that graphical model. Finally, it draws attention to the usefullness of unlabelled data for the 
+novelty detection process in order to considerably increase the detection accuracy.
 

### 2011-06-23 thesis/results/ontology-explain.tex

\begin{tabular}{|c|ccccccc|cccc|cccc|cccc|cccc|cccc|cccc|ccc|ccc|}
\hline
 & \multicolumn{7}{|c|}{appearance property} & \multicolumn{4}{|c|}{object book} & \multicolumn{4}{|c|}{object cerealbox} & \multicolumn{4}{|c|}{object computer} & \multicolumn{4}{|c|}{object robot} & \multicolumn{4}{|c|}{object stapler} & \multicolumn{4}{|c|}{object toiletpaper} & \multicolumn{3}{|c|}{shape property} & \multicolumn{3}{|c|}{size property}\\ \hline
 & \begin{sideways}anteroom\end{sideways} & \begin{sideways}bathroom\end{sideways} & \begin{sideways}hallway\end{sideways} & \begin{sideways}kitchen\end{sideways} & \begin{sideways}lab\end{sideways} & \begin{sideways}meetingroom\end{sideways} & \begin{sideways}office\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}0\end{sideways} & \begin{sideways}1\end{sideways} & \begin{sideways}2\end{sideways} & \begin{sideways}3+\end{sideways} & \begin{sideways}elongated\end{sideways} & \begin{sideways}rectangular\end{sideways} & \begin{sideways}square\end{sideways} & \begin{sideways}large\end{sideways} & \begin{sideways}medium\end{sideways} & \begin{sideways}small\end{sideways}\\ \hline
\begin{sideways}anteroom\end{sideways} & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$\\ \hline
\begin{sideways}bathroom\end{sideways} & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $87.5\%$ & $11.7\%$ & $0.8\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $83.0\%$ & $15.5\%$ & $1.4\%$ & $0.1\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $32.2\%$ & $36.5\%$ & $20.7\%$ & $10.7\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$\\ \hline
\begin{sideways}computerlab\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $71.0\%$ & $24.3\%$ & $4.2\%$ & $0.5\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}conferencehall\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}doubleoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $12.1\%$ & $25.6\%$ & $27.0\%$ & $35.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}hallway\end{sideways} & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}kitchen\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $84.9\%$ & $13.9\%$ & $1.1\%$ & $0.1\%$ & $66.4\%$ & $27.2\%$ & $5.6\%$ & $0.8\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $78.1\%$ & $19.3\%$ & $2.4\%$ & $0.2\%$ & $61.8\%$ & $29.7\%$ & $7.2\%$ & $1.3\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}meetingroom\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}professorsoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $42.1\%$ & $36.4\%$ & $15.8\%$ & $5.7\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\begin{sideways}robotlab\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $2.0\%$ & $2.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $71.0\%$ & $24.3\%$ & $4.2\%$ & $0.5\%$ & $30.0\%$ & $36.1\%$ & $21.7\%$ & $12.1\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $30.0\%$ & $60.0\%$ & $60.0\%$ & $30.0\%$ & $10.0\%$\\ \hline
\begin{sideways}singleoffice\end{sideways} & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $2.0\%$ & $88.0\%$ & $57.2\%$ & $32.0\%$ & $8.9\%$ & $1.9\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $42.1\%$ & $36.4\%$ & $15.8\%$ & $5.7\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $19.6\%$ & $32.0\%$ & $26.0\%$ & $22.4\%$ & $90.0\%$ & $9.5\%$ & $0.5\%$ & $0.0\%$ & $10.0\%$ & $60.0\%$ & $30.0\%$ & $20.0\%$ & $60.0\%$ & $20.0\%$\\ \hline
\end{tabular}

### 2011-06-23 thesis/introduction.tex
Clean introduction of missing text
Corrected the intro.

@@
 in creating robots that are able to interact with humans and their environments.
-Those robots would be able to move into our houses and offices. They would be able to communicate
-with us and provide help solving our tasks.
+Those robots would be able to explore our houses and offices. They would be able to communicate
+with humans and perform typical human-like tasks.
 
 % Keep listing problems robots should be able to solve 
-\begin{itemize}
-\item Keep listing problems robots should be able to solve
-\item They would dynamically adapt to our environment and exhibit intelligence by understanding
-\item Pose problems robots cannot solve without understanding human environments
-\end{itemize}
+%\begin{itemize}
+%\item Keep listing problems robots should be able to solve
+%\item They would dynamically adapt to our environment and exhibit intelligence by understanding
+%\item Pose problems robots cannot solve without understanding human environments
+%\end{itemize}
 
 % Representation of environments are needed!
-In order to reason and understand about the surrounding environment a robot first has to
-obtain a manageable representation of it.
-Such a representation provides the robot with a understanding on how the several
-entities relate to each other, and, if persistent, allow the robot to reason beyond its sensory
-horizon.
+In order to reason about and perform actions in the surrounding environment, the robot first 
+has to obtain a manageable representation of it. Such a representation provides the robot with 
+understanding of how the spatial entities relate to each other and spatial concepts, and, 
+if persistent, allow the robot to reason beyond its sensory horizon.
 
@@
 do so by using representations that are not human friendly: e.g.\ a robot can uniquely identify the
-location of a certain object with coordinates, but those coordinates are not suitable for humans.
-A robot may be able to detect a tag placed on an object, but such an approach besides limited
-and hard-working does not scales: new objects are created everyday, the huge amount of already
-existent objects and concepts also forbids any tagging and concepts by their nature are dynamic.
-Summing up: methods are needed that allow robots to represent environments and allow human concepts
-to also be representable.
+location of a certain landmark using internal coordinates, but those coordinates are not understandable
+for humans. A robot may be able to detect a tag placed on an object, but such an approach does not scale
+outsize structured environments: new objects are created everyday, there is huge amount of already
+existing objects and concepts by their nature are dynamic. For this reason, methods need 
+to be developed that will allow robots to represent real-world environments and allow human concepts
+to be transferred into the robot representations.
 
 % Semantic Mapping
-By gathering semantic knowledge a robot can learn how to map their internal representation to the
-given set of human concepts. Web and other common databases, that have been created by humans, are a
-great source of semantic information, and, in some sense have helped to solve this semantic mapping
-problem.
-Common approaches at semantic mapping are based on classification, this is: a robot has a set of
-defined categories and it picks the one that best represents the sensed data. This approach often
-gives satisfactory results, but only when applied on a controlled environment where the robot has
-knowledge of all the categories that exist.
+By gathering semantic knowledge a robot can learn how to augment the internal representation with a
+set of human concepts. Web and other common databases, that have been created by humans, are a
+great source of semantic information, and can be a valuable source of knowledge for the semantic mapping
+problem. Common approaches to semantic mapping are based on classification, i.e., a robot has a set of
+defined categories and selects the one that best represents the sensed data. This approach often
+gives satisfactory results, but only when applied in a controlled environment where the robot has
+knowledge about all of the categories that exist.
 
 % Complete knowledge? Don't joke me.
-This assumption on knowledge of all the categories that an entity can be mapped to is too strong
-to hold on unknown environments. In specific, a mapping defined over the human semantics will always
-be incomplete due to infeasibility to map all the concepts or even due to its dynamic nature where
-concepts change not only on time but also between sociocultural aspects.
+This assumption about the complete knowledge of all the categories that an entity can be mapped to is 
+too strong to hold on real-world environments. More specifically, a mapping defined over the human semantics 
+will always be incomplete due to the infeasibility to map all the concepts or even due to its dynamic nature 
+where concepts change not only over time but also between sociocultural aspects.
 
 % Novelty Detection
-In this sense a robot should be able to identify gaps on its knowledge by detecting novelty.
+Consequently, a robot should be able to identify gaps in its own knowledge and detect novelty.
 Developing methods for this problem helps to provide robots with novelty signals that can be
-used to improve its reliability, decide when to initiate active-learning or even provide basic
+used to improve reliability, decide when to initiate active learning or even provide basic
 information for keeping knowledge continuously adapted in long-term or life-long scenarios.
 
-The rest of this chapter presents in which context is this thesis interested in detecting novelty.
-It presents then the specific problem it addresses and defines the goals to achieve.
-At last it presents the structure of the other chapters.
+The rest of this chapter presents the context in which this thesis addresses the problem of novelty detection.
+Then, it presents the specifics of the problem and defines the goals to achieve. Finally, it outlines the 
+structure of the remaining chapters.
 
@@
 %
-In mobile robotics the main need is on spatial information. Having a representation of all
-the spatial entities and relations between them allow the robot to navigate and interact
-with them. Maps are the default tool for representing those spaces, and computers excel at
-creating close to exact representations of spatial environments.
-But which concepts existent in the human spatial knowledge are needed in order to turn those
-representations a usable tool to represent human concepts?
+One of the most crucial problems in mobile robotics is that of representing space in which the robot operates. 
+Having a representation of all the spatial entities and relations between them allows the robot to navigate and 
+interact with the environment. Maps are the default tool for representing those spaces, and computers excel at
+creating close to exact representations of spatial environments. However, those representations usually fail to
+capture aspects important for human-like spatial understanding.
 
 Different spatial knowledge types can be identified depending on the scale, abstraction level,
-and required generalization. For example geometric type aspects which robots are already very
-good at representing, object knowledge which is incredibly basic for humans, delineation
-of spatial areas and relations between instances of all those knowledge types. e.g.\ a specific
+and required generalization. Examples include geometric aspects of the world, object knowledge 
+(which is fundamental for humans, but difficult to extract for robots), segmentation of
+spatial areas, and finally, relations between different spatial entities e.g. a specific
 object (book) is on top of another object (table) that is located inside an area (room-library)
-that is inside a another area (university).
-As seen on this example humans do not limit themselves at representing entities
-in space and spatially relate them but they also attribute semantic value to them.
+that is inside a another area (university). We see that humans do not limit themselves to representing 
+entities of space and relating them spatially, but they also attribute semantic concept to them.
 
-It is this semantic value attributed by humans to entities that is interesting for endowing
-robots with knowledge on how to interpret and reason over human environments.
-By understanding which objects are expected to be found on a room, what are the properties that
-distinguish room categories, which room types are likely to be connected together a robot can
-aim at increase its effectiveness on indoor environments.
+It is those semantic concepts attributed by humans to spatial entities that are interesting for 
+endowing robots with knowledge on how to interpret and reason about human environments.
+By understanding which objects are expected to be found in a room, what are the properties that
+distinguish room categories, or which room types are likely to be connected together, a robot can
+increase its communication abilities but also its performance on complex tasks in indoor environments.
 
-A basic concept on indoor environments is that of room. Rooms allow humans to segment areas
-in high\hyp{}level entities that can be categorized according to the properties and expected
-activities to be performed on them.
-e.g.\ kitchens are rooms where cornflakes are usually found, a library is identified by the presence
-of many books, a lecture hall is a place where lessons are given.
+A basic concept of indoor environments is that of a room. Rooms allow humans to segment areas
+into high\hyp{}level entities that can be categorized according to their properties and canonical
+activities performed there, e.g. kitchens are rooms where cornflakes are usually found, a library is 
+identified by the presence of many books, a lecture hall is a place where lessons are given.
 Since room categories play an important role in understanding and reasoning about
@@
 % Detect limitations and novel knowledge
-It is highly desirable that a robot can be deployed to new and unknown environments. Aiming at that,
-a robot has to perform its own semantic mapping of the environment instead of relying on a provision
-of already labelled maps.
-But the semantic mapping ability of a robot is restricted by its spatial and semantic knowledge.
-This poses a problem as its unrealistic to believe that on unknown environments the robot knowledge
-is complete. So it becomes important to detect gaps on the robot knowledge, such that the robot can
-be aware of limitations of its models and devise techniques to not fail victim of its limitations.
+It is highly desirable that a robot can be deployed in new and unknown environments. In order to achieve that,
+the robot has to perform its own semantic mapping of the environment instead of relying on available labelled maps.
+However, the semantic mapping ability of a robot is restricted by its spatial and semantic knowledge.
+This poses a problem as its unrealistic to assume that in an unknown environment, the robot's knowledge
+will be complete. Therefore, it becomes important to detect gaps in the robot's knowledge so that the robot can
+be aware of the limitations of its models and deal with the novel situations. Furthermore, the detection of 
+knowledge gaps can be extremely useful for several other problems in artificial intelligence such as 
+active learning and knowledge maintenance during life-long operation.
 
-Additionally the detection of knowledge gaps, will serve as base for several other problems in
-artificial intelligence such as active-learning and maintenance of long and life-long knowledge.
-
-\section{Proposal and Goals}
-This thesis aims at developing methods that can be used to detect gaps on the semantic knowledge
-of an agent. For that the spatial knowledge representation proposed in \cite{pronobis2010ias} will
-be used as base defining the agent knowledge.
-Given an instance world, that knowledge, is used to perform semantic mapping~\cite{pronobis2011semmap}
-on indoor environments using a probabilistic graphical model. That model is instantiated from the
-environment using the spatial knowledge and captures both semantic and structural information of
-space. It allows the agent to map all the modelled variables in the presence of uncertainty and 
-therefore this thesis will focus on study and developing methods that can detect novel semantic room
-categories from it.
+\section{Goals}
+This thesis aims at developing methods that can be used to detect gaps in spatial semantic knowledge
+of a robotic agent. To this end, the spatial knowledge representation proposed in \cite{pronobis2010ias} 
+is used as a base for defining the agent's spatial knowledge. Given a concrete environment, 
+a semantic mapping process is applied and a conceptual map representation of the environment it build according
+to~\cite{pronobis2011semmap}. The conceptual map is represented in terms of a probabilistic graphical model
+capturing both semantic and structural information of space. This thesis focuses on novelty detection methods 
+from this probabilistic graphical model representation. In particular, emphasis is placed on detection of 
+novel room categories given the observed spatial properties of the environment.
 
@@
 
-\autoref{chap:background} introduces background concepts for understanding the
-presented thesis. In particular \autoref{sec:graphical-models} introduces
-\emph{probabilistic graphical models} that lay the base modelling tool for
-the used structured semantic representation.
+\autoref{chap:background} introduces fundamental concepts important for understanding the problems 
+addressed in the thesis. In particular \autoref{sec:graphical-models} introduces \emph{probabilistic 
+graphical models} that lay the foundation for the conceptual map.
 
-\autoref{chap:semantic-mapping} describes the semantic mapping process.
-It introduces a spatial knowledge representation and describes a system
-that uses it for performing semantic mapping using using a \emph{probabilistic graphical model}.
+\autoref{chap:semantic-mapping} describes the semantic mapping process. It introduces the spatial 
+knowledge representation and describes a system that uses it for performing semantic mapping using 
+using the \emph{probabilistic graphical model}.
 
-\autoref{chap:novelty-intro} introduces novelty detection under a
-statistical view point and how to interpret it as a threshold function.
-It presents then the main work of this thesis: detecting novel semantic
-categories using graphical models.
+\autoref{chap:novelty-intro} introduces novelty detection from the statistical point of view and 
+shows how to interpret it as a thresholding function. Then, it the main contribution of this thesis:
+methods for detecting novel semantic categories using graphical models.
 
-\autoref{chap:conclusions} draws conclusions on the developed work and presents
-interesting directions for future works.
+\autoref{chap:conclusions} draws conclusions on the developed work and presents possible 
+directions for future works.

### 2011-06-23 thesis/summary-pt.tex
Added summary in PT.

@@
 
-\section{Motivation}
+\section{Motivação}
 Já à muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
-Acredita-se que eles serão capaz de ser mover e interagir connosco, compreender os nossos conceitos
+Acredita-se que eles serão capaz de se mover e interagir connosco, compreender os nossos conceitos
 e serem capazes de pensar.
@@
 próprios do mundo humano.
-No entanto é irrealista acreditar que será possivel descrever e representar em toda a sua extensão
+No entanto é irrealista acreditar que será possível descrever e representar em toda a sua extensão
 tais conceitos. Assim os agentes deverão ser capazes de aprender e adaptar o seu conhecimento quando
 confrontados com novas situações.
-A detecção destas novas situações onde o conhecimento de um agente não é suficiente têm um papel importante
+A detecção destas novas situações, onde o conhecimento de um agente não é suficiente, têm um papel importante
 na adaptação e aprendizagem de novos conceitos e significados e é por isso o tópico principal desta tese.
 
-Addicionalmente no contêxto de agentes moveis, para facilitar o deslocamento e planeamento, os
-agentes constroiem uma representação do ambiente na qual detectam conceitos e propriedades
-conceitos espaciais.
-Essa representação identifica conceitos e classifica-os de acordo com significados humanos,
-permitindo assim ao agente obter uma representação dos conceitos humanos e pensar e planear a um
-nível próximo dos mesmos.
-Ao processo de construção desta representação é attribuido o nome de \emph{semantic mapping}.
+No contexto de agentes moveis, para facilitar o deslocamento e planeamento, os
+agentes constroem uma representação do ambiente. Quando essa representação é estendida com
+conceitos e categorias com significado para humanos, as tarefas de alto nível ficam mais fáceis.
 Por exemplo, a uma escala de quartos, os humanos categorizam o espaço em: cozinhas, quartos, corredores,
-escritórios, sala de computadores, etc\dots
+escritórios, etc\dots
 Ao relacionar quartos com estas categorias que descrevem propriedades e acções percebidas por
-humanos, um agente consegue resolver tarefas como: ``Onde estão os ceriais?''
+humanos, um agente consegue resolver tarefas como: ``Onde estão os cereais?''
 
 \section{Problema e Objectivos}
-Como deliniado na motivação, o processo de \emph{semantic mapping} têm um papel importante na
-capacidade pensativa de um agente. Em simultâneo, esse processo está limitado pelo conhecimento
-do agente em conceitos e significados do mundo humano. É então importante estudar e analisar
-técnicas que sejam capazes de detectar falhas no conhecimento e permitir ao agente agir de maneira
+O processo de mapeamento semântico têm um papel importante na capacidade pensativa de um agente.
+Em simultâneo, esse processo está limitado pelo conhecimento em conceitos e significados do mundo humano.
+É então importante estudar e analisar técnicas que sejam capazes de detectar falhas no conhecimento e permitir ao agente agir de maneira
 cuidadosa.
-É com esse objectivo que a tese aqui descrita se foca no estudo de deteção
-de situações novas num processo de \emph{semantic mapping} propôsto por \cite{pronobis2011phd}.
-Para esse fim a tese estuda como usar modelos graphicos para detectar novas categorias nas conhecimento
-de categorias do agente.
+É com esse objectivo que a tese aqui descrita se foca no estudo de técnicas de detecção de situações novas
+e de como as aplicar num processo de mapeamento semântico com vista a detectar novas categorias de lugares
+em ambientes interiores. Para esse fim, a tese estuda o processo de mapeamento semântico proposto por \cite{pronobis2011phd},
+e procura solucionar a detecção de novas categorias usando modelos gráficos.
 
 \section{Mapeamento Semântico}
-Mapeamento semântico é o processo através do qual um agente constroi uma representação do ambiente
-extendida com conceitos e categories com significado especial (i.e.\ conceitos humanos).
+Mapeamento semântico é o processo através do qual um agente constrói uma representação do ambiente
+estendida com conceitos e categorias com significado especial (i.e.\ conceitos humanos).
 Neste tese o processo proposto por \cite{pronobis2011phd} foi usado como base.
@@
 \subsection{Conhecimento Espacial}
-O conhecimento espacial de agente define todo a informação necessária para o mapeamento semântico,
+O conhecimento espacial de agente define a informação necessária para o mapeamento semântico,
 e inclui informação em como segmentar e identificar entidades e propriedades através dos dados
 recolhidos.
-Para esse fim Pronobis~\cite{pronobis2011semmap} utiliza uma representação estruturada em camadas:
-a \emph{camada de sensores} é responsável por manter uma descrição exacta do ambiente numa representação
+Para esse fim \cite{pronobis2011semmap} utiliza uma representação estruturada em camadas:
+a \emph{camada de sensores} mantém uma descrição exacta do ambiente numa representação
 de baixo nível;
-a \emph{camada de lugares} é responsável por segmentar e identificar areas e conceitos
-como quartos;
-a \emph{camada de categories} é responsável por definir conhecimento sobre propriadades e categories
-e permite ao agente obter informação de mais alto-nível;
-e a \emph{camada de conceitos} mantêm conhicimento sobre os diversoes conceitos, relações sobre os mesmos
+a \emph{camada de lugares} segmenta e identifica áreas e conceitos como quartos;
+a \emph{camada de categorias} é responsável por definir conhecimento sobre propriedades e categorias
+e permite ao agente obter informação de mais alto nível;
+e a \emph{camada de conceitos} mantêm conhecimento sobre os diversos conceitos, relações entre os mesmos
 e informação de como os mapear para conceitos humanos.
 
-\subsection{Conceptual Map}
-The semantic mapping is then based on the instantiation of a conceptual map that uses all the
-ontology defined on the conceptual layer in order to build a probabilistic graphical model that
-relates and propagates probabilities across all the entities and properties detected by the
-other layers.
-
-By using probabilistic graphical models, the conceptual map becomes then a generative model that
-allows to perform inference on variable configurations. For that it is used as a base tool
-for novelty detection of unknown categories on some of its variables.
+\subsection{Mapa Conceptual}
+Durante o processo de mapeamento semântico o agente utiliza a ontologia da camada conceptual para produzir
+um mapa conceptual. Este é produzido usando modelos gráficos probabilísticos que relacionam
+os conceitos detectados e propagam probabilidades entre os mesmos de acordo com o conhecimento do agente.
+A detecção de uma nova categoria no conhecimento do agente traduz-se na detecção de um novo estado nas
+variáveis de tal grafo e portanto os mapas conceptuais são usados como ferramenta para detecção de novas
+categorias.
 
 \section{Novelty Detection}
-Novelty detection, also known as outlier or anomaly detection, is a
-classification problem related to identification of new or unknown data
-patterns that the system is not aware of~\cite{markou2003novelty}.
-The ability to identify novel cases is crucial in any autonomous system
-that is deployed to unknown or to uncontrolled environments, as it gives the
-system the ability to detect that something is not conforming to its knowledge and
-therefore should be treated with caution.
-% It has several applications such as fault detection~\cite{tarassenko1999novelty},
-% intrusion detection~\cite{fan2001using},
-% detection of masses in mammograms~\cite{tarassenko1995novelty} or detection of
-% novel and useful documents~\cite{zhang2002novelty}.
+\emph{Novelty detection}, também conhecido por \emph{detecção de anomalias} ou \emph{detecção de outliers},
+é o problema de detecção que os dados foram produzidos por uma nova classe,
+sobre a qual o agente não têm conhecimento~\cite{markou2003novelty}.
+A habilidade de detectar estes casos é de fulcral importância em sistemas autónomos que operam em ambientes
+desconhecidos. Pois permite detectar que algo está fora do normal e assim actuar com cuidado.
 
-\subsection{Novelty Detection as Threshold}
-Due to the noisy, uncertain and insufficient nature of sensed data, novelty ought to be
-treated in a probabilistic way in where each sample has a certain probability
-of being generated by a class not known for the agent and a complementary probability $P(\overline{novel}|x)$
-of being generated by a known class.
+\subsection{\emph{Novelty Detection} através de \emph{Thresholds}}
+Devido a ruídos e incerteza existentes nos dados, a detecção de casos novos é tratada com
+probabilidades. Assim para cada amostra de dados $x$ existe uma certa probabilidade de a mesma
+ser gerada por uma classe desconhecida e uma probabilidade complementar $P(\overline{novel}|x)$
+de ser por uma classe conhecida.
 
-Additionally any detector can be uniquely described by the set $N$ of samples that it
-classifies as novel. With that its possible to show that by including a new sample $a$ in $N$ an agent
-can increase its detection rate at a cost of producing more mis-classifications.
-This describes the base of the \emph{error and rejection tradeoff} described by \cite{chow1970optimum}.
-Due to the interest of having the highest detection rate with a given minimum error rate, any optimal
-detector can be described as a \emph{continuous knapsack problem} of the sample space.
-The greedy ordering defined to solve it can be shown to be equivalent to:
+Além disso, qualquer detector por ser descrito unicamente pelo conjunto $N$ de amostras que classifica
+como novas. Com isso é possível demonstrar que ao expandir o conjunto $N$ com um novo elemento $a$,
+o agente aumenta a taxa de detecção a um custo de produzir mais classificações erradas.
+Isto descreve o principio de troca entre erro e rejeição descrito por \cite{chow1970optimum}.
+Um detector óptimo, é assim equivalente ao problema da mochila com quantidades fraccionarias, pois têm
+como objectivo obter a maior taxa de detecção com uma taxa de erro máximas. Assim qualquer detector
+óptimo pode ser descrito através de um limite imposto sobre a relação gulosa dado por:
 \begin{eqnarray}
@@
 
-\subsection{Assumption on a constant $P(novel)$}
-By considering $P(novel)$ to be constant a ratio between conditional and unconditional probabilities
-of the sensed variables is obtained.
-Although it sounds reasonable to assume the probability of specific room belong to a novel
-category is constant, this thesis points that this assumption might be very strong and not
-reasonable in a realistic scenario. The main point being that in this case the graph structure
-changes between samples and some might be more prone to produce novel samples, as opposed to normal
-classification scenarios where its clear that the category is indeed the only generator of features.
+\subsection{Critério: $P(novel)$ como Constante}
+Ao considerar $P(novel)$ constante, uma relação equivalente pode ser obtida usando somente a fracção entre $P(x|\overline{novel})$ e $P(x)$.
+Apesar de parecer inofensivo, o caso apresentado neste tese é diferente dos casos normais em que todos as amostras
+são geradas através de uma categoria. Aqui toda estrutura e relação entre as variáveis toma um papel na geração
+de amostras e portanto não é claro como tal critério se comporta na realidade.
 
-\subsection{Modeling Conditional Probability}
-Since the graphical model produced by the conceptual map tries to model the distribution of
-variables assuming the knowledge of the agent holds true, it represents a method to model
-$P(x|\overline{novel})$.
-%since it allows to calculate the density probability that the set of features $x$ is
-%sensed given that all the variables and graph structure, including the category of
-%a specific room $a$ are correctly modelled by the agent knowledge.
-%This way the distribution modelled by a factor graph equivalent to the chain-graph used by
-%the conceptual map can be used as an approximation for the conditional probability of $x$.
+\subsection{Modelar $P(x|\overline{novel})$}
+O modelo gráfico produzido durante o mapeamento semântico, modela a distribuição de todas as variáveis
+assumindo que o conhecimento do agente é representativo da realidade, como tal pode ser usado
+para aproximar $P(x|\overline{novel})$.
 
-\subsection{Modeling Unconditional Probability}
-With no knowledge on the real distribution samples, the correct approach is to model it
-with a uniform distribution. Nonetheless in the presented case, it was assumed that only the room
-category of a specific room is unknown and that the remaining knowledge holds trues.
-In that case, the correct approach is to replace the variable that represents the room where to
-detect novelty with a single factor connecting all variables directly dependent on it.
+\subsection{Modelar $P(x)$}
+Sem conhecimento sobre a distribuição real é correcto modela-la usando uma distribuição
+uniforme. No entanto, o agente considera que todo o seu conhecimento é representativo
+da realidade à excepção da categoria de uma das suas variáveis. Nesse caso a aproximação correcta é substituir
+os factores dependentes dessa variável por um único factor.
+Devido à dimensão do mesmo, é inviável aproximá-lo sem realizar simplificações.
+A tese apresenta assim duas aproximações: um modelo uniforme e um modelo independente. No modelo independente
+esse factor é aproximado com vários factores dependentes somente numa variável. Esses podem facilmente ser
+aproximados e é esperado que obtenham melhores resultados ao compensar por uma tendência do ambiente em gerar
+certas \emph{features}.
 
-Due to its high-dimensionality, approximating that factor with unlabelled data is infeasible and
-approximations have to be done. In this thesis two approximations were presented: an uniform model
-and an independent model. In the independent model, the
-single\hyp{}connected factors are trained from unlabelled data and its expected that they create a better
-approximation to the unconditional probability than an assumption on the uniform model.
-
-\section{Results}
-A synthetic dataset modelling the expected features when implementing on the real case and enforcing
-a constant $P(novel)$ was used for testing the developed models.
-The two suggested models where compared between them and to an optimal case.
-The tests across several situations shows how they behaviour change as more information is sensed
-from the environment.
+\section{Resultados}
+Para testes um conjunto de dados sintético foi preparado. Este aproxima as propriedades
+esperadas num cenário realístico como por exemplo um número variável de \emph{features}.
+Os modelos sugeridos foram comparados entre eles e entre um detector óptimo, e os testes realizados
+mostram o comportamento dos modelos em diversas situações, por exemplo conforme mais informação
+é recolhida pelo agente.
 
@@
 \includegraphics[width=0.4\textwidth]{results/synthetic-all.pdf}
-\caption{\label{fig:results}Comparison of the proposed models using samples from a synthetic dataset
-         with variable number of sensed features.}
+\caption{\label{fig:results}Comparação dos modelos propostos usando dados sintéticos com número variável de \emph{features}.}
 \end{figure}
 
-\section{Conclusions}
-This thesis studied the problem of novelty detection of semantic categories on the spatial knowledge
-of mobile robots. To that purpose a semantic mapping process~\cite{pronobis2011semmap} was studied and
-a novelty detection scheme based on the conceptual map used by it proposed.
-Initial tests where performed on synthetic data showing an interest in using unlabelled data to
-increase the detection performance.
-Additionally all the assumptions and bases behind novelty detection were presented and several
-directions on how to create a generalized framework for novelty detection were discussed.
+\section{Conclusão}
+Esta tese estudou o problema da detecção de novas categorias de espaços interiores no contexto de agentes moveis.
+Para tal o processo de mapeamento semântico~\cite{pronobis2011semmap} foi estudado e um mecanismo de detecção,
+baseado no mapa conceptual usado pelo mesmo, foi proposto.
+Foram realizados testes iniciais com dados sintéticos, mostrando o benefício de usar dados não supervisionados.
+Adicionalmente toda a teoria e assunções por trás do método apresentado foram documentadas, incluindo
+discussão de diversas direcções futuras em como criar uma \emph{framework} generalizada para detecção de casos
+novos no conhecimento semântico e espacial de agentes.
 

### 2011-06-23 thesis/conclusion.tex
Conclusions and rel. work corrected.

--- 2011-06-20 thesis/conclusion.tex
@@
 % Summary of developed work
-This thesis studied the problem of detecting situations where the knowledge of the robot
-is not 
-Gaps those, that disable the robot to function 
-It did so, specializing 
-that allows novelty detection of semantic room
-categories using as base the semantic mapping process proposed by \cite{pronobis2011semmap}.
+This thesis studied the problem of detecting novel situations in case of which the robot
+lacks sufficient knowledge to explain the sensory information. It did so in the area of 
+semantic mapping on indoor spaces, and more precisely, for the problem of detecting the situtation when
+none of the known room category models were able to correctly explain the sensed spatial properties.
 
-\url{https://github.com/andresusanopinto/novelty-detection-thesis}
+It reviewed novelty detection and how an optimal detector can be implemented by
+thresholding. It also showed that with the assumption of constant probability of seeing a
+novel case, an optimal ordering function for thresholding can be implemented based on
+the factor between conditional and unconditional probability.
 
-On the rest of this chapter the main results and conclusions are presented.
-Additionally limitations of the presented method and directions for future work are given.
+It studied the semantic mapping process proposed in \cite{pronobis2011semmap} and presented 
+a method to detect novel room categories based on probabilistic graphical models.
+Using a synthetic dataset, respecting the assumptions, it showed that such a method would be 
+optimal if unconditional probability could be optimally approximated. Since, in realistic conditions, 
+unconditional probability cannot be obtained due to the lack of knowledge about all possible classes, various
+techniques were used to approximate it with either a uniform assumption or simplified models built from 
+unlabelled data.
+
+The rest of this chapter presents the main results and conclusions. Additionally limitations of the 
+presented methods and directions for future work are discussed.
 
@@
 
+After studying the semantic mapping process and novelty detection methods, this
+thesis proposed modelling the conditional and unconditional probability distribution of sensed
+data using graphical models. Under the presented assumptions, and assuming both distributions can be 
+exactly known, a ratio between those probabilities would yield an optimal ordering function for implementing a
+novelty detector. This case was showed to be optimal using a synthetic dataset that simulates a simplified
+environment for semantic categorization of rooms based on sensed properties.
+As the next step, this thesis tested methods to approximate the unconditional probability. 
+It tested the usage of a uniform distribution and explored the usage of unlabelled data to 
+produce better models.
+
+Additionally it studied how the performance of the presented methods changed as more data
+was sensed from the environment. As expected, all the methods increase their performance, but they
+move further away from what could be obtained with exact information.
+Additionally, the disadvantage of the uniform assumption compared to using a highly simplified
+model for the unlabelled data starts reducing as more information is available.
+Nonetheless the model based on the usage of unlabelled data consistently lead to better detector
+performance, which is a strong indication that unlabelled data should be used whenever it is available.
+
+As a note, all the results presented in this thesis are directly reproducible from an
+online\footnote{\url{https://github.com/andresusanopinto/novelty-detection-thesis}}
+repository. That repository besides containing all the code and data for the results also
+includes tech-reports, presentations, articles and notes that have been produced during
+this thesis work. Additionally, future research by the author will be correctly linked there,
+when appropriate.
+
 \section{Limitations}
 
-\begin{itemize}
-\item Strong assumption on constant $P(novel)$
-\item Threshold should actually be dynamic with structure
-\item Unable to drop normalizations factors
-\item Handle multiple novel variables at the same time
-\end{itemize}
+The presented methods use a strong assumption on a constant $P(novel)$, this is unrealistic and
+forbids one to exploit graph structural information that could show a given variable is more likely
+to be unknown to the system because the given graph structure is not easily explained by the
+current knowledge. As as example of that, consider the following: if all the agent knows are two 
+categories that are very often seen as a bi-colored graph, when presented with a non-bipartite graph, 
+the agent should consider the likelihood of a new category greater than on bipartite graphs.
+
+Additionally, all the presented methods require calculation of $P(x|\overline{novel})$ and
+$P(x)$. This way a full graphical model is needed and uncertain sensing as described in
+\autoref{sec:cues-from-low-level} cannot be easily incorporated.
+
+These two limitations are a single symptom of the decision on inverting the conditional probability
+$P(novel|x)$ to obtain a threshold that can be directly modelled by the data.
+It is expected that by trying to model how the potentials in the factor graphs change between the
+conditional and unconditional graphical model, both limitations can be surpassed.
 
@@
 
-\begin{itemize}
-\item Extend novelty to all categories on the graph (room, size, shape, etc..)
-\item Handle uncertain sensing
-\item Probabilistic Space Segmentation (learn how to structure the graph)
-\item Learn Concepts / hidden variable types - (areas, rooms, floors, environments, etc\dots)
-\item Explain why a certain sample is novel
-\item Generate most likely novel sample
-\end{itemize}
+Future work will try to study and work around the limitations presented above. At the moment
+it is not clear how the threshold changes when the assumption about constant $P(novel)$ cannot be
+made. Any attempts to bypass the listed limitations will need to consider that problem by 
+developing a method where a threshold can have a more realistic and controlled behaviour 
+for any graph structure.
+
+The following paragraphs describe possible and interesting directions to explore in the context 
+of detecting knowledge gaps in artificial intelligent systems:
+
+% \subsection*{Extend Novelty to All Variable Types on the Graph}
+% Extend novelty to all categories on the graph (room, size, shape, etc..)
+% \subsection*{Handle Uncertain and Novel Sensing from Low-level Classifiers}
+% Handle uncertain sensing
+\subsubsection*{Generalized the Framework}
+The presented method should be generalized by allowing novelty detection to be performed on any 
+variable of the graphical model. Additionally, the novelty information could be incorporated back 
+into the graph allowing the agent to probabilistically reason even when variables are considered 
+unknown. That generalization should aim at being fully probabilistic such as the system presented by
+\cite{ranganathan2010pliss} and not deterministic by having to make decisions on which variables
+it considers novel.
+
+Additionally, several methods exists that allow to produce novelty signals from the low-level
+classifiers. A generalized framework should try to fuse and handle all that information by
+incorporating it into the graphical model in a similar fashion as how uncertain sensing outcomes are incorporated.
+
+% \subsection*{Explain Why a Certain Sample is Novel}
+% Explain why a certain sample is novel
+% \subsection*{Generation of most Likely Novel Samples}
+% Generate most likely novel sample
+\subsubsection*{Exploiting Generative Models}
+An interesting aspect that arises from the use of generative models is the ability to generate new
+samples from the models. It is expected that this can be exploited to achieve better understanding
+of what the system is modelling and what explain the character of the novelty back to the user.
+Moreover, this feature can be used to refine and actively update the models by confronting the captured 
+knowledge with the human understanding of the modeled concepts.
+
+% \subsection*{Probabilistic Modelling of Space Segmentation}
+% \subsection*{Learn New Concepts}
+% Learn Concepts / hidden variable types - (areas, rooms, floors, environments, etc\dots)
+\subsubsection*{Beyond Novel Semantic Categories: Learning Graph Structures}
+Although this thesis only touched the problem of detecting novel semantic categories, other knowledge
+types should also be considered incomplete and are also candidates for novelty detection.
+More concretely, besides detecting novel semantic categories for given spatial concepts, there is
+also interest in detecting novel concepts: the presence of hidden variables of a new type, not known
+to the system: for example detecting rooms or types of environment.
+
+An initial step towards that goal could be probabilistic learning of graph structures to model various
+possible space segmentations. This would permit the use of a probabilistic approach instead of the currently used
+deterministic approach where space is segmented based on the presence of doors or other landmarks.
+
+\subsubsection*{Beyond Detection of Knowledge Gaps}
+Having methods to detect gaps of knowledge is just one of the first steps towards creating long-term and
+life-long adaptable systems that are capable of learning over time. After detecting new
+situations, ways to learn and incorporate the knowledge into the models must be developed.
+
+g

### 2011-06-23 thesis/abstract.tex
Portuguese abstract finished.

@@
 for performing detection of novel room categories which were not previously known to the robot applicable 
-to that graphical model. Finally, it draws attention to the usefullness of unlabelled data for the 
+to that graphical model. Finally, it draws attention to the usefulness of unlabelled data for the 
 novelty detection process in order to considerably increase the detection accuracy.
@@
 % Portuguese Version of the above.
+
 \chapter*{Resumo}
-Assume this text to be $T_{E \rightarrow P}(abstract)$.
-\footnote{What I mean is that this will be a translation of the English text
-present on abstract. Also don't expect it to be filled in before the final
-version of this document!}
+\selectlanguage{portuguese}
+
+Já à muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
+Acredita-se que eles serão capaz de se mover e interagir connosco, compreender os nossos conceitos
+e serem capazes de pensar.
+Um passo importante nesta direcção é a criação de robôs com conhecimento de conceitos e significados
+próprios do mundo humano.
+No entanto é irrealista acreditar que será possível descrever e representar em toda a sua extensão
+tais conceitos. Assim os agentes deverão ser capazes de aprender e adaptar o seu conhecimento quando
+confrontados com novas situações.
+A detecção destas novas situações, onde o conhecimento de um agente não é suficiente, têm um papel importante
+na adaptação e aprendizagem de novos conceitos e significados e é por isso o tópico principal desta tese.
+
+No contexto de agentes moveis, para facilitar o deslocamento e planeamento, os
+agentes constroem uma representação do ambiente. Quando essa representação é estendida com
+conceitos e categorias com significado para humanos, as tarefas de alto nível ficam mais fáceis.
+Por exemplo, a uma escala de quartos, os humanos categorizam o espaço em: cozinhas, quartos, corredores,
+escritórios, etc\dots
+Ao relacionar quartos com estas categorias que descrevem propriedades e acções percebidas por
+humanos, um agente consegue resolver tarefas como: ``Onde estão os cereais?''
+
+Esta tese estuda um processo, que usa conhecimento de conceitos espaciais e categorias com significado
+para humanos, para realizar mapeamento semântico do ambiente através de modelos gráficos probabilísticos.
+Apresenta depois técnicas para realizar detecção de novas categorias. Chama também a atenção para a
+utilização de dados não supervisionados para efectivamente melhorar a exactidão da detecção.
+
+\selectlanguage{english}

### 2011-06-23 thesis/novelty-intro.tex
Part 1 of the corrections on chapter 4

@@
 techniques and related works. Then it shows that an optimal novelty detection system
-can be implemented by thresholding on a order-relation defined over the inputs.
-And that an equivalent order-relation can be imposed by a ratio between
+can be implemented by thresholding an order relation defined over the inputs.
+Moreover, an equivalent order relation can be imposed by a ratio between
 conditional and unconditional probabilities.
-Some discussion on how to interpret the meaning of those factors is also
-presented.
 
-At last a practical example using semantic data and probabilistic graphical
+Finally,  a practical example using semantic data and probabilistic graphical
 models is presented: it shows how to use the introduced ratio to obtain a
@@
 
-\section{Novelty Detection Review and Related Work}
+\section{Novelty Detection - Review and Related Work}
 Novelty detection, also known as outlier or anomaly detection, is a
@@
 Any normal classification application is also a good candidate for extending
-with novelty detection, as the results for samples the system has not been
-trained on can be unreliable~\cite{devarakota2008reliability}.
-e.g.\ applied to digit-recognition~\cite{tax1998outlier}
+with novelty detection, as the results for samples on which the system has not been
+trained on can be unreliable~\cite{devarakota2008reliability}
+e.g. when applied to digit-recognition~\cite{tax1998outlier}
 or to detection of novel inputs on a neural network~\cite{bishop1994novelty}.
@@
 the system to acquire examples of all possible classes where the complexity of
-novelty detection lies: it is only possible to obtain samples representing positives
+novelty detection lies: it is only possible to obtain samples representing positive
 examples of known cases and the lack of negative examples renders normal
@@
 be resistant to noisy features,
-ability to scale in feature dimension,
-deal with multiple classes and performing detection efficiently as many
+scale in feature dimension,
+deal with multiple classes and perform detection efficiently as many
 autonomous systems require real-time or close to real-time performance.
@@
 try to distinguish novelty by separating the training set from all the other
-points in the input space. They try to achieve that by enclosing the training
-set by some structure (e.g.\ an hyper-sphere)~\cite{bennett2000support}.
-Those approaches have been made in line with Vapnik principle of not solving
-something hard. As although having access to a perfect probability distribution
+points in the input space. This is achieved by enclosing the training
+set with some structure (e.g.\ an hyper-sphere)~\cite{bennett2000support}.
+Those approaches have been made in line with Vapnik's principle of never to solve a 
+problem which is more general than the one we actually need to solve~\cite{scholkopf2000support}.
+Although having access to a perfect probability distribution
 of the input would solve the problem, creating such a function is harder than
-simply creating a boundary between known data and novel data
-\cite{scholkopf2000support}.
+simply creating a boundary between known data and novel data.
 
@@
 exploiting time-coherency on video information. The approach is particularly
-interesting by its ability of keeping a fully probabilistic distribution of
-the place classification and segmentation. This is, the system never performs
+interesting by its ability of keep a fully probabilistic distribution of
+the place classification and segmentation. In other words, the system never performs
 a deterministic decision that impacts any future result, allowing it to adjust
@@
 The information obtained by the region detectors is exploited together with the
-spatial relations between on the higher level scene-classification task.
-A presented scheme using only pairwise relations between regions is show to
-have better performance than the correct approach of modelling all regions connectivity
-with a large single-factor as the pairwise relations can be
-approximated with modest amounts of training data, where the single-factor
+spatial relations for the higher level scene-classification task.
+A scheme using only pairwise relations between regions is shown to
+have better performance than the correct approach of modelling connectivity of all regions 
+with a large single factor as the pairwise relations can be
+approximated with modest amounts of training data, where the single factor
 demands an impractical amount of training data.
-No approach is made on it at novelty detection either of regions or scene
+No approach is made for novelty detection either of the regions or the scene
 categories.
@@
 The objective of a novelty detection system is to classify a given sample $x$ as
-either \emph{known}: $x$ being generated by a known class to the system, or \emph{novel}:
-$x$ generated by a class unknown for the system.
-Based on the ground\hyp{}truth of a sample four cases are possible:
+either \emph{known}: $x$ being generated by a class known to the system, or \emph{novel}:
+$x$ generated by a class unknown to the system. Based on the ground\hyp{}truth of a sample four cases are possible:
 
@@
 
-Due to noisy data, unstable features and lack of information it is impossible to develop
-a method able to exactly guess the correct classification of a sample.
-By modelling the outcomes with probabilities it becomes possible to handle the uncertainty
+Due to noisy data, unstable features and lack of information, it is impossible to develop
+a method able to always exactly guess the correct classification of a sample.
+By modelling the outcomes with probabilities, it becomes possible to handle the uncertainty
 associated with each decision.
@@
 system to be classified. $\overline{novel}$ is also defined in such a way that
-$P(\overline{novel}|x)$ measures the probability that a decision on classifying $x$ as novel is
-wrong.
+$P(\overline{novel}|x)$ measures the probability that the sample $x$ is generated by a known class.
 
@@
 false-positive error.
-The true positive probability can be described as the interest in detecting as much
-as possible the novel classes and the false positive probability the interest
-in not incurring in too many errors.
-By fixating one of those it is possible to define a novelty detection
-system that achieves the maximum or minimum of the other.
+The true positive probability is related to the goal of detecting as many novel samples as possible.
+At the same time, the false positive probability is related to the goal of not making too many errors.
+By fixing one of those, it is possible to define a novelty detection system that achieves the maximum or minimum of the other.
 
 This way an optimal detector can be formulated by achieving the maximum true-positive
-probability without its false-positive probability increase beyond a given limit.
+probability without its false-positive probability increasing beyond a given limit.
 This is equivalent to a \emph{continuous knapsack problem} which allows a greedy
-solution by sorting the items with a value per weight function.
-In the case of detection system that can be defined as:
+solution by sorting the items with a value per weight function. In the case of 
+detection system that can be defined as:
 
@@
 weight(x) &=& P(\text{false positive}|x) \\
-cost(x)   &=& value(x)/cost(x) \\
+cost(x)   &=& value(x)/weight(x) \\
           &=& \frac{P(novel|x)P(x)}{P(\overline{novel}|x)P(x)}
@@
 
-Therefore a novelty system before classifying a sample $a$ as novel should (greedily)
-classify any sample $b$ with a small cost as that would achieve
-an higher true positive probability given a fixed false positive one.
+Therefore a novelty detection system before classifying a sample $a$ as novel should (greedily)
+classify any sample $b$ with a smaller cost as that would achieve a higher true positive probability 
+given a fixed false positive one.
 
@@
 And any optimal detector can be described by the largest $P(\overline{novel}|x)$
-accepted by it. Which is seen as threshold.
+accepted by it. This can be seen as thresholding.
 
@@
 
-On the previous section it was shown that an optimal novelty detector can be
-implemented with a threshold on top of the order-relation defined by
+In the previous section, it was shown that an optimal novelty detector can be
+implemented by placing a threshold on the order relation defined by
 $P(\overline{novel}|x)$ over $x$. Performing some manipulations with
-Bayes theorem and assuming a constant $P(\overline{novel})$ a more usable
+Bayes theorem and assuming a constant $P(\overline{novel})$, a more usable
 form can be attained:
@@
 
-Since there is only interest in maintaining the same order-relation as
-$P(\overline{novel}|x)$ any constant factor can be dropped.
+Since the interest is only in maintaining the order relation defined by
+$P(\overline{novel}|x)$, any constant factor can be dropped.
 Leaving a ratio between a \emph{conditional} and
@@
 of the samples given that they are generated by a known class.
-In case the labelled data available for the agent to learn a concept comes from
-the same distribution where the system will run the correct approach is to use it
-as prior-information for modelling the conditional probability.
+In case when the labelled data used to learn a concept come from
+the same distribution from the test samples come, the correct approach is to use it
+as prior information for modelling the conditional probability.
 
 Note that it is important for the labelled data to be a filtered version
-of the underlying world distribution (with all classes) that does not contains
-any bias.
-Otherwise that bias will lead to incorrect modelling the conditional probability
-and a subsequent wrong ordering of the sample space.
+of the underlying world distribution (one modeling all the concepts in the world) 
+that does not contain any bias. Otherwise that bias will lead to incorrect modelling of 
+the conditional probability and in consequence wrong ordering of the sample space.
 
 \subsection{Unconditional Probability}
-The unconditional probability $P(x)$ plays an important role on obtaining a
+The unconditional probability $P(x)$ plays an important role in obtaining a
 correct order relation for performing novelty detection.
-It serves as a normalizing component that allows the system to figure out
-whether a given sample conditional probability arises from it belonging to
+It serves as a normalizing component that allows the system to decide
+whether conditional probability of a given sample arises from it belonging to
 the known concept or from the likelihood of being sampled.
@@
 the principle of maximum entropy (\autoref{sec:max-entropy}) a uniform
-distribution must be chosen.
-Though by using unlabelled data it becomes possible to obtain extra information
-and achieve a better approximation.
+distribution must be chosen. However, by using unlabelled data, it becomes possible 
+to obtain extra information and achieve a better approximation.
 
 %\subsection{Novelty Detection on a Variable Set of Features}
-Note also that often novelty detection is applied on a fixed set of features
+Note also that often novelty detection is applied to a fixed set of features
 together with an assumption of a uniform unconditional probability.
-On those cases $P(x)$ becomes a constant and therefore a novelty threshold
-can be directly applied on $P(x|\overline{novel})$ as is the case in \cite{bishop1994novelty}.
-But in the case where the set of features $x$ is variable it cannot be
-discarded. There $P(x)$ plays a role in levering all the conditional
-probabilities on different sets of variables into the same measure units
-such that a threshold can be implemented.
+In those cases, $P(x)$ becomes a constant and therefore a novelty threshold
+can be directly applied to $P(x|\overline{novel})$ as is the approach presented in \cite{bishop1994novelty}.
+However, in case where the set of features $x$ is variable, it cannot be
+discarded. There, $P(x)$ plays a role in levering all the conditional
+probabilities on different sets of variables into the same units so that a threshold can be obtained.
 
@@
 
+\begin{sidewaystable}[h]
+\begin{center}
+\scalebox{0.40}{\input{results/ontology-explain}}
+\end{center}
+\caption{\label{extra:synthetic-distribution}Distribution used on the synthetic experiment. Each column cell shows $P(feature|class)$}
+\end{sidewaystable}
+
 \subsection{Conditional Probability}

### 2011-06-24 thesis/acknows.tex
Added Acknowledgements chapter.

@@
 \chapter*{Acknowledgements}
+This thesis would have not been possible without the help of some other persons.
+I would like to thank to Andrzej Pronobis, for introducing me to the problem and
+directing it towards my interests in computer science, for the meetings, insights
+and endless reviews of my work. Also to Carl Henrik Ek, for enthusiastically taking part
+on the meetings discussing graphical models and related work.
 
-Aliquam id dui. Nulla facilisi. Nullam ligula nunc, viverra a, iaculis
-at, faucibus quis, sapien. Cum sociis natoque penatibus et magnis dis
-parturient montes, nascetur ridiculus mus. Curabitur magna ligula,
-ornare luctus, aliquam non, aliquet at, tortor. Donec iaculis nulla
-sed eros. Sed felis. Nam lobortis libero. Pellentesque
-odio. Suspendisse potenti. Morbi imperdiet rhoncus magna. Morbi
-vestibulum interdum turpis. Pellentesque varius. Morbi nulla urna,
-euismod in, molestie ac, placerat in, orci.
+The Computer Vision and Active Perception Lab at Royal Institute of Technology
+in Stockholm were also a key piece for providing such a staff and facilities that
+surrounded me during one year of thesis and exchange studies, and whose
+courses replenish my interests in machine learning.
+Also Faculdade de Engenharia of University of Porto who made it possible
+for me to study and perform my thesis abroad.
+Thanks to Luis Paulo Reis, for accept supervising my thesis and pushing me to submit a paper.
 
-Ut convallis. Suspendisse luctus pharetra sem. Sed sit amet mi in diam
-luctus suscipit. Nulla facilisi. Integer commodo, turpis et semper
-auctor, nisl ligula vestibulum erat, sed tempor lacus nibh at
-turpis. Quisque vestibulum pulvinar justo. Class aptent taciti
-sociosqu ad litora torquent per conubia nostra, per inceptos
-himenaeos. Nam sed tellus vel tortor hendrerit pulvinar. Phasellus
-eleifend, augue at mattis tincidunt, lorem lorem sodales arcu, id
-volutpat risus est id neque. Phasellus egestas ante. Nam porttitor
-justo sit amet urna. Suspendisse ligula nunc, mollis ac, elementum
-non, venenatis ut, mauris. Mauris augue risus, tempus scelerisque,
-rutrum quis, hendrerit at, nunc. Nulla posuere porta orci. Nulla dui.
+Virgile Högman, for besides handling me as a work colleague and chatter-box, becoming a friend.
+And Erik Ass for all the interesting coding and math discussions on totally random problems.
+I can't also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
+that kept chatting with me through the nights during my thesis work.
 
-Fusce gravida placerat sem. Aenean ipsum diam, pharetra vitae, ornare
-et, semper sit amet, nibh. Nam id tellus. Etiam ultrices. Praesent
-gravida. Aliquam nec sapien. Morbi sagittis vulputate dolor. Donec
-sapien lorem, laoreet egestas, pellentesque euismod, porta at,
-sapien. Integer vitae lacus id dui convallis blandit. Mauris non
-sem. Integer in velit eget lorem scelerisque vehicula. Etiam tincidunt
-turpis ac nunc. Pellentesque a justo. Mauris faucibus quam id
-eros. Cras pharetra. Fusce rutrum vulputate lorem. Cras pretium magna
-in nisl. Integer ornare dui non pede.
+Thanks to Mariatorget people who kept asking how my thesis was going and when it was going to be finished.
+And for all the parties, beer relaxing time and nights out! Thanks to Elina Säfsten, for being a cool neighbour
+with whom I had the pleasure to spent quite some hours talking and partying with her and her Swedish friends.
+Thanks to Manuel Gattermayr and Tommaso Facchini for being awesome neighbours as well.
+
+I am also grateful to Bernhard Schwaighofer, Manuel and Tommaso and others for an amazing
+road trip, night fires, sunsets and sunrises on Sweden. Special thanks to Bernie for during the road trip
+promising to wake me up before my sleeping bag starts to burn and melt with my skin and eventually turning me water proof.
+
+And thanks to all the party people, not forgetting Andrzej, who showed me how to do a PhD party in case I ever decide
+to seek one.
+
+Last but not the least, thanks to my family and friends back in Porto and other parts of the world,
+who kept talking and cheered me up.
+
+% PS: This list does not contains all the persons the author of this thesis would like to thank, and the author should not be held liable for it.
 
 \vspace{10mm}
-\flushleft{The Name of the Author}
+\flushleft{
+"I am thankful to all the awesome people who were part of this Stockholm chapter of my life"
+--
+André Susano Pinto}

### 2011-06-24 thesis/acknows.tex
Semantic mapping missing points and other small fixes.

@@
 And Erik Ass for all the interesting coding and math discussions on totally random problems.
-I can't also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
+I cannot also avoid to thank my neighbours back in Nockeby, in special to Diogo Gonçalves and Elise Löbker
 that kept chatting with me through the nights during my thesis work.
@@
 Thanks to Mariatorget people who kept asking how my thesis was going and when it was going to be finished.
-And for all the parties, beer relaxing time and nights out! Thanks to Elina Säfsten, for being a cool neighbour
-with whom I had the pleasure to spent quite some hours talking and partying with her and her Swedish friends.
+And for all the parties, drinking, relaxing and nights out! Thanks to Elina Säfsten, for being a cool neighbour
+with whom I had the pleasure to spent quite some time talking and partying with her and her Swedish friends.
 Thanks to Manuel Gattermayr and Tommaso Facchini for being awesome neighbours as well.
 
-I am also grateful to Bernhard Schwaighofer, Manuel and Tommaso and others for an amazing
+I am also grateful to Bernhard Schwaighofer, Manuel, Tommaso and others for an amazing
 road trip, night fires, sunsets and sunrises on Sweden. Special thanks to Bernie for during the road trip
@@
 --
-André Susano Pinto}
+\emph{André Susano Pinto}}

### 2011-06-24 thesis/semantic-mapping.tex
moved line in place. >.<
Semantic mapping missing points and other small fixes.

@@
 association is either useful by explaining where to find certain objects as well to
-understand the types of activities to perform on it. 
+understand the types of activities to perform on certain areas. 
 Those human-concepts are meant to be meaningful for humans and by incorporating conceptual
 knowledge about them it is expected to improve the agent ability to interact and communicate
-with humans and the ability perform human-like tasks.
+with humans and to perform human-like tasks.
 
@@
 In this context semantic mapping is seen as the process of building a representation of
-the environment which associates spatial entities with a set of defined spatial concepts (in this case the human concepts).
+the environment which associates spatial entities with a set of defined spatial concepts (in this case human concepts).
 
-Missing:
-\begin{itemize}
-\item importance of conceptual knowledge
-% In order to build such a mapping of the environment, it is important to define the knowledge that allows the agent to reason
-% about.
+% Importance of spatial and conceptual knowledge
+In order to build such a semantic map, it is important to define the spatial and conceptual knowledge
+that allows an agent to map between its machine-friendly representation of the world and the high-level semantics
+characteristic of the human world.
+That knowledge needs to define concepts and how to segment and identify them from the low-level representation,
+it needs also to define how the concepts relate to each other and attribute meanings to them.
+In the case of indoor environments this can be seen for example as: rooms are a unit of division of area and are
+often separated by doors, each room can also be characterized by the objects inside, size and shape, and they
+can be categorized in meaningful concepts such as: kitchen, library, corridor, etc\dots
 
-\item application of conceptual knowledge
-% It is expected that by being able to reason over those semantics an agent can better
-% understand human environments and increase its efficiency when performing human-tasks on them.
-% By handling this conceptual knowledge the agent becomes aware of
+% Why to treat it in a probabilistic way
+The low-level representation and the knowledge of the agent, will always be to a certain a degree imperfect,
+incomplete, inaccurate and invalid. Due to that both the concepts and their relations
+cannot be quantified in a rule base system and ought to be treated in a probabilistic way in order to allow
+the agent to still reason with some accuracy about the world.
 
-\item challenges and need for representing uncertainty
-% The goal of a spatial representation is to provide the agent with a view about the environment
-% that allows it to reason beyond its sensory horizon.
-
-% Such a representation is by its nature imperfect, incomplete, inaccurate and invalid.
-
-% world that allows the agent to perform reasoning over it and efficiently perform tasks.
-% It should also give the agent ability to reason about parts 
-% Such a representation cannot be perfect since the methods and information are uncertain
-
-% It is therefore clear that the need to handle information on an high-level semantic and
-% probabilistic fashion arises not only for efficiency reasons but also from the desire
-% of being able to communicate and reason on those aspects.
-% In order to fully explore the semantic properties and their relations it is necessary to
-% detect and classify semantic aspects as well to understand the relationships between them.
-% That conceptual knowledge is important and any limitations on it will limit the capacities
-% of an agent to reason. 
-\end{itemize}
-
-% here a here a method is used that does semantic mappping and representes conceptual knowledge
+% here a here a method is used that does semantic mapping and represents conceptual knowledge
 % on a probabilistic fashion.
@@
 \item[Geometric Size] is classified once again recurring to laser data. In this case the system
+tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
 \item[Object Detection] is performed by running traditional object recognition classifiers.
@@
 based on their visual features, e.g.\ by using \gls{SIFT} as described in \autoref{sec:sift}.
-tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
 \item[Visual Appearance] is extracted from the visual input. For that Dora uses a trained

### 2011-06-25 thesis/conclusion.tex
Thanks Diogo for spotting that.

@@
 lacks sufficient knowledge to explain the sensory information. It did so in the area of 
-semantic mapping on indoor spaces, and more precisely, for the problem of detecting the situtation when
+semantic mapping on indoor spaces, and more precisely, for the problem of detecting the situation when
 none of the known room category models were able to correctly explain the sensed spatial properties.
@@
 situations, ways to learn and incorporate the knowledge into the models must be developed.
-
-g

### 2011-06-26 thesis/novelty-intro.tex
Some changes in the novelty section.

@@
 
-\subsection{Assumption on a constant $P(novel)$}
+\subsection{Assumption about Constant $P(novel)$}
 The ratio between the conditional and unconditional probabilities of the sensed variables
@@
 
-All the literature focus on novelty detection within a fixed scenario, where the set of variables
+All the literature focuses on novelty detection within a fixed scenario, where the set of variables
 used to model the distribution is fixed and defined at the moment the threshold is trained.
-Due to that, they can drop a constant $P(novel)$ and never need to directly or indirectly calculate
-it.
-To the best of the author knowledge there has not been studies on how to approximate it on dynamic
-sets and structures of sensed variables, and so a strong assumption on being constant through all
-the possible scenarios is used to allow dropping the factor.
+Due to that, it is possible to drop the constant $P(novel)$ and there is no need to calculate
+it directly or indirectly.
+To the best of the knowledge of the author there has been no studies on how to approximate it on dynamic
+sets and structures of sensed variables and a strong assumption has always been used about the factor being 
+constant through all possible scenarios.
 
-This assumption was considered questionable on whether it holds on a realistic scenario,
+It is questionable whether this assumption holds in realistic scenarios.
+\textbf{[Read the section from here on.]}
 and for that a sampling scheme able to comply with it is now presented as well an unbiased sampling

### 2011-06-26 thesis/novelty-intro.tex
Changed section on constant P(novel)

@@
 It is questionable whether this assumption holds in realistic scenarios.
-\textbf{[Read the section from here on.]}
-and for that a sampling scheme able to comply with it is now presented as well an unbiased sampling
-scheme. A method to generate samples that match this \emph{criterium} of having a constant $P(novel)$ can be:
+Note that in order to match this \emph{criterium} of having a constant $P(novel)$ samples needs to be
+drawn with a method like:
 
@@
 
-According to this sampling algorithm, given any structure the probability of
-variable $a$ being novel is constant. This is somehow unexpected as it means
-the sampling method is biased to adjust the real distribution $G$ such that
-exactly a constant percentage of drawn samples are novel.
-
-Alternative, the sampling method described below is considered more correct as
-it introduces no bias on the distribution $G$ describing the real distribution
-and classes of $a$:
+Alternative, it is expected that in a realistic scenario, samples are drawn with the following
+unbiased method:
 \begin{algorithm}[Draw samples with equal $P(novel)$]
@@
 
-Due to the highly bias needed to drawn a constant amount of novel sampes from any structure, the
-author defends this assumption to be very strong and points that future work should avoid.
+If in reality samples are drawn according to this unbiased method, assuming a constant $P(novel)$
+will negatively impact the exactness of the detector. For that reason the author defends this assumption
+to be very strong and points that future work should try to develop methods to include structure information
+on the probabilistic model to allow the system to deal with certain structures being more prone to produce
+novel samples.
 

### 2011-06-26 thesis/novelty-intro.tex
Next corrections.

@@
 
-It is questionable whether this assumption holds in realistic scenarios.
-Note that in order to match this \emph{criterium} of having a constant $P(novel)$ samples needs to be
-drawn with a method like:
+It is questionable whether this assumption holds in realistic scenarios. Note that in order to match the 
+\emph{criterium} of having a constant $P(novel)$, samples needs to be drawn with a method like:
 
@@
 
-Alternative, it is expected that in a realistic scenario, samples are drawn with the following
+Alternatively, it is expected that in a realistic scenario, samples are drawn with the following
 unbiased method:
@@
 If in reality samples are drawn according to this unbiased method, assuming a constant $P(novel)$
-will negatively impact the exactness of the detector. For that reason the author defends this assumption
-to be very strong and points that future work should try to develop methods to include structure information
-on the probabilistic model to allow the system to deal with certain structures being more prone to produce
+will negatively impact the exactness of the detector. For that reason, the author believes this assumption
+is very strong and points out that future work should try to develop methods to include structure information
+about the probabilistic model to allow the system to deal with certain structures being more prone to produce
 novel samples.
@@
 This section presents now how to use graphical models produced by the conceptual map and the
-novelty detection concepts introduced earlier on this chapter in order to detect room categories the
+novelty detection concepts introduced earlier in this chapter in order to detect room categories the
 system is not aware of.
 
-For that both models to approximate both the conditional and unconditional probability need to
-be defined and then the ratio between them used to define a function over which a threshold
-implements a novelty detection system.
+For that purpose, two models approximating both the conditional and unconditional probability need to
+be defined. Then, the ratio between the resulting probabilities is used to define a function over which 
+a threshold is specified.
 
 \subsection{Approximating the Conditional Probability}
-The actual semantic mapping of the environment happens using the \emph{conceptual map}
-introduced on \autoref{sec:conceptual-map}. During it the agent instantiates a
-\emph{chain graph} to model distribution of the sensed variables according to the known
-concepts and categories introduced as hidden variables on the graph. Using that graphical model
+The semantic mapping algorithm uses the \emph{conceptual map}
+introduced in \autoref{sec:conceptual-map} to represent the environment on the high level of abstraction. 
+During the semantic mapping process, the agent instantiates a
+\emph{chain graph} to model the distribution of the sensed variables according to the known
+concepts and categories introduced as hidden variables in the graph. Using that graphical model,
 the system is able to propagate and find the most likely configuration of the represented
-variables. For instance, the semantic category of a specific room is modelled as an hidden variable
-with states representing the semantic values, by calculating probabilities on that variable the
-system finds its belief on the room belonging to a specific category.
+variables. For instance, the semantic category of a specific room is modelled as a hidden variable
+with states representing the semantic values. By calculating probabilities of that variable, the
+system obtains the belief about the room belonging to a specific category.
 
-Since the graphical model produced by the conceptual map tries to model the distribution of
-variables assuming the knowledge of the agent holds true, it represents a method to model
-$P(x|known)$ since it allows to calculate the density probability that the set of features $x$ is
+Since the graphical model produced by the conceptual map models the distribution of
+the variables assuming the knowledge of the agent holds true, it corresponds to the model of
+$P(x|\overline{novel})$. It allows to calculate the probability density that the set of features $x$ is
 sensed given that all the variables and graph structure, including the category of
-a specific room $a$ are correctly modelled by the agent knowledge.
-This way the distribution modelled by a factor graph equivalent to the chain-graph used by
+a specific room $a$ are correctly modelled by the knowledge of the agent.
+This way, the distribution modelled by a factor graph equivalent to the chain graph used by
 the conceptual map can be used as an approximation for the conditional probability of $x$.
@@
 
-With no knowledge on the unconditional probability the correct approach is to
+With no knowledge about the unconditional probability, the correct approach is to
 assume a uniform distribution (\autoref{sec:max-entropy}), which can be modelled by a
@@
 
-Thought, very often there is extra knowledge that can be obtained about the distribution
-of the variables that helps to model the unconditional distribution. With that, more accurate
+However, very often additional knowledge can be obtained
+that helps to model the unconditional distribution. Given that knowledge, more accurate
 models can be produced. 
-In this case, the system is interested in detecting only if a specific variable category is not
+In this case, the system only aims at detecting if a specific variable category is not
 known. With that in mind, it was assumed that the graph structure is known and that an unknown
-category can only influence variables by using the same structure as the known variables.
-This way the structural information and all the other hidden variables the system is aware of
+category can only influence variables using the same structure as the known variables.
+Therefore, the structural information and all the other hidden variables the system is aware of
 can be used to more accurately approximate the unconditional distribution.
@@
 \centering
-\subfloat[\label{fig:any-a}Since variable $a$ can be unknown, the agent does not has
-knowledge on how to model the dashed relations.]{
+\subfloat[\label{fig:any-a}Since variable $a$ can be unknown, the agent has no
+knowledge about how to model the dashed relations.]{
 \begin{tikzpicture}
@@
 \qquad \qquad
-\subfloat[Without knowledge on all states of $a$, all variables dependent on $a$ get directly
-          dependent between each other, introducing a big single factor.]{
+\subfloat[Without knowledge about all the states of $a$, all variables dependent on $a$ are directly
+          dependent on each other, introducing a big single factor.]{
 \begin{tikzpicture}
@@
 
-If the structure and all the other variables can be assumed to be correctly modelled by the
-knowledge of the agent, then all variables that are directly dependent on $a$ become directly
-dependent between them, introducing a single factor (\autoref{fig:known-struct}).
-By approximating this factor the agent can then obtain a model for the unconditional probability.
+If the knowledge of the agent is only lacking information about all the states
+of hidden variable $a$, it can still be used to approximate the sample distribution 
+by avoiding the need to directly represent $a$. For that, all the variables that were 
+directly dependent on $a$ become directly dependent between each other introducing a 
+single big factor connecting all of them (\autoref{fig:known-struct}).
+By approximating this factor, the agent can then obtain a model for the unconditional probability.
 
-In cases no other information is available the introduced factor can be considered uniform\footnote{Uniform factors do not influence the
-distribution represented by the factor graph and so can be removed from the graph representation}
-and dropped from the graph (see \autoref{fig:uniform-model}).
-Nonetheless it may be the case that due the cost of labelling data, the agent does not has knowledge
-on all the possible categories of $a$, but still has access to unlabelled data.
-In that case unlabelled samples may be gathered and explored to achieve a better approximation of
-the real distribution of variables.
-Though the factor that needs to be approximated requires handling of all the connected variables
-and enormous amounts of data may be needed to approximate it.
+In cases when no other information is available, the introduced factor can
+be considered uniform\footnote{Uniform factors do not influence the
+distribution represented by the factor graph and so can be removed
+from the graph representation} (see \autoref{fig:uniform-model}).
+Nonetheless, it may be the case that due to the cost of labelling data,
+the agent does not has knowledge on all the possible categories of $a$,
+but still has access to unlabelled data.
+In that case unlabelled samples may be gathered and exploited to
+achieve a better approximation of the real distribution of variables.
+However, the factor that needs to be approximated requires handling of
+all the connected variables and enormous amounts of data may be
+needed to approximate it.
 
-A simple approach can be performed by assuming the variables connected to the factor to be independent
-but having a bias towards certain values. This approach is equivalent to connecting single factors
-on each of variables and approximating those factors using unlabelled data as seen on \autoref{fig:independent-model}.
+A simple approach can be used by assuming the variables connected
+to the factor are independent, but still have a bias towards certain values. 
+This approach is equivalent to connecting single factors
+to each of variables and approximating those factors using unlabelled
+data as seen in \autoref{fig:independent-model}.
 
@@
 
-Other schemes may be tried to approximate the unconditional independent assuming that only
+Other schemes may be used to approximate the unconditional independent assuming that only
 a specific variable is not necessarily known. For example modelling unconditional probability
@@
 After defining a model for the conditional and unconditional distributions, a threshold can be applied
-over the ratio of those.
-Under the assumption of a constant $P(novel)$, defining a threshold over it leads to an optimal novelty
-detector, as showed on \autoref{sec:threshold}.
+to the ratio of those. Under the assumption of a constant $P(novel)$, defining a threshold over the ratio 
+leads to an optimal novelty detector, as showed on \autoref{sec:threshold}.
 
-If the training data does not contains any bias, by picking a threshold on the training data that achieves
-a certain true-positive and false-positive rate, it is expected that on the real distribution such a system
-will achieve the same performance.
-
-Note however that the performance of the system is dependent on how well the models approximate the conditional
-and unconditional distributions and so are dependent on how the graph structure is able to model the real distribution
-and how well the training data allows the factors to be approximated.
+If the training data do not contain any bias, it is expected that by picking a threshold on the training 
+data that achieves certain true-positive and false-positive rate, a system will achieve the same performance 
+on the real distribution. However, the performance of the system is dependent on how well the models 
+approximate the conditional and unconditional distributions and so are dependent on how the graph structure 
+is able to model the real distribution and how well the training data allows the factors to be approximated.
 
@@
 \label{sec:unlabelled-data}
-In order to sum up the presented concepts on novelty detection with a threshold
+In order to summarize the presented concepts related to novelty detection with a threshold
 function and exemplify how to use graphical models in the context of
-multi-modality room classification a synthetic dataset was generated.
+multi-modality room classification, a synthetic dataset was generated.
 The dataset was kept simple by only modelling directly sensed features from a
-room, skipping any structural knowledge or sensing model such as room
-connectivity and extra hidden variables.
+room, skipping any structural knowledge (room connectivity) and extra hidden variables.
 
-In this dataset a room $r$ is seen as an hidden-variable generator of a set of
+In this dataset, a room $r$ is seen as a hidden-variable generator of a set of
 features $X$ that are directly sensed by the agent.
 All the sensed features $x$ are independent given the room category.
-In whole there was 11 different room categories and 7 different feature types.
-Each feature can be sensed more than once (e.g.\ room shape is extracted from 2D
+In total, there were 11 different room categories and 7 different feature types.
+Each feature can be sensed more than once (e.g. room shape is extracted from 2D
 laser scans in more than one position in the room), but all those sensed
@@
 The room categories were chosen to mimic as close as possible the real features
-and categories existing in reality (i.e.\ 1 person office, 2 person office,
-hallway, robot lab, etc\dots). A table describing the used synthetic
+and categories (i.e.\ 1 person office, 2 persons office,
+hallway, robot lab, etc.). A table describing synthetic
 distribution is given in \autoref{extra:synthetic-distribution}.
 
-The objective then was to design a system that, although only trained with
+The objective was to design a system that, although only trained with
 labelled data from 5 of the 11 room categories, was able to detect novel
-room categories.
-For that 100 labelled samples for the 5 known categories were drawn and 1000 
-unlabelled samples were drawn from all the room categories for learning the
-unconditional probability distribution and measure effect of using unlabelled
-data.
+room categories. To this end, 100 labelled samples for the 5 known categories 
+were drawn and 1000 unlabelled samples were drawn from all the room categories for 
+learning the unconditional probability distribution and measure effect of using 
+unlabelled data.
 
@@
 \end{center}
-\caption{\label{extra:synthetic-distribution}Distribution used on the synthetic experiment. Each column cell shows $P(feature|class)$}
+\caption{\label{extra:synthetic-distribution}Distribution used for the synthetic data experiment. Each column cell shows $P(feature|class)$}
 \end{sidewaystable}
@@
 Using the labelled samples, 7 factors $\phi_X(r,x)$ were created, one for each
-feature type, to represent the potential of sensing features $x$ of type $X$ on
-a room of category $r$.
+feature type, to represent the potential of sensing features $x$ of type $X$ for
+the room of category $r$.
 
@@
 the probabilities in case a given sample is never seen.
-With those the probability of sensing a set of features $x$ on a room category
-$r$ known for the agent can be modelled with a factor graph as illustrated on
+With those, the probability of sensing a set of features $x$ for a room category
+$r$ known to the agent can be modelled with a factor graph illustrated in
 \autoref{fig:simple-cond-graph}.
@@
 \label{sec:sample-uncond-prob}
-With no knowledge on the unconditional probability the correct approach is to
-assume a uniform distribution.
-That is represented in factor graphs by a graph without any factors.
+With no knowledge about the unconditional probability, the correct approach is to
+assume a uniform distribution. That is represented as a factor graphs without any factors.
 
@@
 Very often there is extra knowledge that can be obtained about the distribution
-of the variable that helps to model the unconditional distribution.
-In this practical example the access to unlabelled data is explored as access to
-it is common on applications in robotics.
+of the variables that helps to model the unconditional distribution.
+In this practical example, the access to unlabelled data is exploited. The availability of
+unlabeled data is common in practical robotic applications.
 
-Note that the sensed variables are dependent between each other when the room
-category $r$ is not known. With that the correct approach would be to train a
-factor that is able to correlate all the sensed variables as seen on
+Note that the sensed variables are dependent on each other when the room
+category $r$ is not known. The correct approach would be to train a
+factor that is able to correlate all the sensed variables as seen in
 \autoref{fig:simple-all-dependent}.
@@
 \caption{\label{fig:simple-all-dependent}A general factor graph able
-         to model any unconditional distribution on the sensed variables
+         to model any unconditional distribution of the sensed variables
          requires a factor connecting all of them.}
@@
 the \emph{curse of dimensionality}: as the number of sensed features and feature
-types increases exponential amounts of data are needed to describe it.
-Having interest in avoiding the issues associated with an high-connected factor
-an assumption on independent features can be done and then the unconditional
-probability can be modelled with a fully disconnected variables, but with
-factors that account for existent bias on each single feature (as visible on
-graph in \autoref{fig:simple-independent-graph}).
+types increase, exponential amounts of data is needed to describe it.
+In order to avoid those the issues, an assumption about independent features can be made
+and then the unconditional probability can be modelled with fully disconnected variables, 
+but with factors that account for existent bias on each single feature (as shown in
+the graph in \autoref{fig:simple-independent-graph}).
 
-In the case there is only one sensed feature the distribution generated by an
+In case there is only one sensed feature, the distribution generated by the
 assumption of independent features correctly models the unconditional
-probability. And it deviates from it as more features are sensed.
+probability and it deviates as more features are sensed.
 The individual factors associated with each variable can be seen as a scaling of
-each feature that tries to account for a possible existent bias on each
-of them. Therefore, although far from the desired factor, it is expected to be a
-better estimate than assuming a uniform distribution.
+each feature which tries to account for a possible existent bias for each
+of them. Therefore, it is expected to be a better estimate than assuming a uniform distribution.
 
@@
 \item[exact] 
-- since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
+- since the distribution is synthetic, there is access to $P(x)$ and $P(x|concept)$
 and the ordering function $P(x|\overline{novel})/P(x)$ could be created
-to test how far the other presented thresholds are from optimal.
+to test how far the other presented thresholds are from the optimum.
 
 \item[uniform]
-- by assuming a uniform unconditional distribution the ordering function is
+- by assuming a uniform unconditional distribution, the ordering function is
   given by $P_G(x)/P_U(x)$.
@@
 \item[independent]
-- using the unlabelled data the unconditional distribution can be approximated.
+- using the unlabelled data, the unconditional distribution can be approximated.
   In this case, it was approximated with an independent distribution of the
@@
 % Show that the thresholds are suitable functions for implementing a static threshold.
-As first tryout, the performance of the novelty threshold selection was plotted for a set
-of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
-Those samples contain different number of sensed properties for each room, mimicking
+First, the performance of the novelty threshold selection was plotted for a set
+of 1000 samples taken from the whole distribution (\autoref{fig:synthetic-roc}).
+Those samples contain different numbers of sensed properties for each room, mimicking
 the dynamic properties expected to see when implemented on a robot.

### 2011-06-26 thesis/novelty-intro.tex
Final fixes.

@@
 
-Other schemes may be used to approximate the unconditional independent assuming that only
-a specific variable is not necessarily known. For example modelling unconditional probability
-with an hidden variable trained from the unlabelled data. Or by trying to split the single big
-factor on factorizations of other factors.
-Though the presented cases have interesting properties:
+Other approaches may be used to approximate the single factor described above
+and allow modeling the distribution when there is no knowledge about all the states
+of a specific variable. For example: by training a hidden variable in an
+unsupervised fashion or by using other factorization schemes for the factor.
+The presented cases have interesting properties:
 
 \begin{description}
-\item[uniform model] - by assuming the factor to be uniform, it can be considered as not existent on
-the graphical model.
+\item[uniform model] - by assuming the factor to be uniform, it can be
+considered as not existent in the graphical model.
 
-\item[independent model] - due to the use of single connected factors, it can easily be trained from
-                           unlabelled data. This helps the system to account for an existent bias on
-                           each of the variables without allowing the system to overfit to the unlabelled
-                           data.
+\item[independent model] - due to the use of single connected factors,
+it can easily be trained from unlabelled data. This helps the system to
+account for existing bias for each of the variables without the
+system overfitting to the unlabelled data.
 \end{description}
@@
 
-\caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
-         under samples with variable size of sensed properties.}
+\caption{\label{fig:synthetic-roc}ROC curve comparing the novelty detection performance
+         for samples with variable size of sensed properties.}
 \end{figure}
 
-The convex shape for the optimal threshold shows that the ratio between conditional
-and unconditional probability is indeed an optimal detector and is suitable for
-implementing a threshold when the samples are taken from dynamic
+The convex shape of the optimal threshold shows that the ratio between conditional
+and unconditional probability is indeed an optimal detector \textbf{[Why does it show it's optimal?? Maybe too strong word?]}
+and is suitable for implementing a threshold when the samples are taken from dynamic
 distributions (e.g.\ some samples where there is only access to room size versus
@@
 That is probably explained by the semantic properties being highly
-biased towards some values. And shows that bias plays an important step
+biased towards some values. This shows that bias plays an important role
 in detecting whether a given sensed value is a valuable cue about the room category.
@@
 % Measure performance of the thresholds as more information becomes available.
-\subsection{Performance Changes With Amount of Available Information}
+\subsection{Influence of the Amount of Available Information}
 In order to measure the performance impact as more semantic information becomes
-available ROC curves were plotted for samples grouped by the number of sensed
+available, ROC curves were plotted for samples grouped by the number of sensed
 semantic features.
@@
 \caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
-         presented novelty detection method on graphs generated for different amount of
+         presented novelty detection method for graphs generated for different amount of
          sensed features.}
@@
 
-It is possible to see that as the system gains more semantic information it
-becomes easier to detect novelty. The input space size increases and allows the
-several existing classes to become more easily distinguished.
+It is possible to see that as the system gains more semantic information, it
+becomes easier to detect novelty. The size of the input space increases and allows the
+existing classes to become more easily distinguished.
 
@@
 
-The uniform threshold shows a poor performance specially on small size samples
-where its performs almost no better than random.
-It performance increases as the size of sensed features increases but nonetheless
-its very small when compared to how optimal a threshold could be.
+The uniform threshold shows poor performance especially for samples with small amount of features
+where it performs almost no better than random.
+The performance increases as the size of sensed features increases but nonetheless
+is very small when compared to the optimal threshold.

### 2011-06-26 thesis/novelty-intro.tex
Address final comments by andrzej

@@
 of 1000 samples taken from the whole distribution (\autoref{fig:synthetic-roc}).
-Those samples contain different numbers of sensed properties for each room, mimicking
+The samples where uniformly generated by graphs with 5, 10, 15, 20, 35, 50 features.
+Additionally the feature types were also uniformly sampling, for that it is possible
+that in certain samples some feature types were sensed more than once and other were not
+sensed at all.
+This was chosen to mimic
 the dynamic properties expected to see when implemented on a robot.
@@
 \caption{\label{fig:synthetic-roc}ROC curve comparing the novelty detection performance
-         for samples with variable size of sensed properties.}
+         for samples with variable number of sensed properties.}
 \end{figure}
@@
 The convex shape of the optimal threshold shows that the ratio between conditional
-and unconditional probability is indeed an optimal detector \textbf{[Why does it show it's optimal?? Maybe too strong word?]}
-and is suitable for implementing a threshold when the samples are taken from dynamic
-distributions (e.g.\ some samples where there is only access to room size versus
+and unconditional probability is indeed a suitable detector for implementing a threshold when
+the samples are taken from dynamic distributions when $P(novel)$ is constant
+(e.g.\ some samples where there is only access to room size versus
 samples where there is a lot of information about the room properties).

### 2011-06-26 thesis/novelty-intro.tex
small typo

@@
 The samples where uniformly generated by graphs with 5, 10, 15, 20, 35, 50 features.
-Additionally the feature types were also uniformly sampling, for that it is possible
+Additionally the feature types were also uniformly sampled, for that it is possible
 that in certain samples some feature types were sensed more than once and other were not

### 2011-07-01 semantic-nd/article.tex
Review all till befoore the results.

@@
 supervised and unsupervised ways and by using thresholds on those distributions
-based solely on the conditional probability as seen on Bishop~\cite{bishop1994novelty}
+based solely on the conditional probability, as seen on Bishop~\cite{bishop1994novelty},
 a novelty system can be trivially implemented.
@@
 
-Based on this design, a \emph{chain graph} model is proposed as a 
+Based on this design, a \emph{chain graph}~\cite{lauritzen2002chain} model is proposed as a 
 representation for performing inferences on the knowledge represented in the conceptual 
@@
 The use of graphical models to describe distributions of variables has useful properties.
-First, the permit inference about uncertain conceptual knowledge. At the same time, they are 
+First, they permit inference about uncertain conceptual knowledge. At the same time, they are 
 generative models and therefore allow to calculate the probability
@@
 Describing the distribution function in terms of graphs allows to use those engines to
-efficently calculate marginals on any given subset of variables by exploiting conditional
+efficiently calculate marginals on any given subset of variables by exploiting conditional
 independence between variables.
@@
 % Explain threshold approach to novelty detection
-Due to the desire of robustness, a novelty detection system is often given the option
-to classify non-novel samples as novel. This is implemented as a threshold that describes a trade
-between error and rejection: the more the system tries to reduce rejection, the more
-prone to produce errors it becomes.
+Due to the nature of the sensed data which are noisy and uncertain, novelty ought to be
+treated in a probabilistic way where each sample has certain probability
+of being generated by a class not known to the agent and a complementary probability $P(\overline{novel}|x)$
+of being generated by a known class.
+The true positive and false positive rate of a novelty detection system which classifies
+the set $N$ of samples as novel is given by:
 
-When dealing with a statistic point of view, noisy data or unstable features, a decision to
-classify a sample as novel has an associated error probability of $P(\overline{novel}|x)$.
-And the error rate of a system that classifies the set of inputs $N$ as novel is given by
-the equation:
+\begin{eqnarray}
+P(\textnormal{true positive})  &=& \sum_{x \in N}{P(novel|x)P(x)} \\
+P(\textnormal{false positive}) &=& \sum_{x \in N}{P(\overline{novel}|x)P(x})
+\end{eqnarray}
+
+It leads that by extending the set $N$ with new samples, the true positive and
+false positive rate can never be decreased, leading to the problem where in order
+to increase detection a system needs to increase its error. This describes the
+base of the \emph{error and rejection tradeoff}~\cite{chow1970optimum}, which
+states that a system aiming at increasing the true-positive probability will eventually increase its
+false-positive error.
+
+This way an optimal detector can be formulated by achieving the maximum true-positive
+probability without its false-positive probability increasing beyond a given limit.
+This is equivalent to a \emph{continuous knapsack problem} which allows a greedy
+solution by sorting the items with a value per weight function. In the case of 
+detection system that can be defined as:
+
+\begin{eqnarray}
+value(x)  &=& P(novel|x) \\
+weight(x) &=& P(\overline{novel}|x) \\
+cost(x)   &=& value(x)/weight(x) \\
+          &=& \frac{P(novel|x)P(x)}{P(\overline{novel}|x)P(x)}
+\end{eqnarray}
+
+Therefore a novelty detection system before classifying a sample $a$ as novel should (greedily)
+classify any sample $b$ with a smaller cost as that would achieve a higher true positive probability 
+given a fixed false positive one.
 
 \begin{equation}
-P(e) = \frac{\sum_{n \in N}{P(\overline{novel}|n)P(n)}}{\sum_{n \in N}{P(n)}}
+\label{eq:knapsack}
+\frac{P(novel|b)}{P(\overline{novel}|b)} < \frac{P(novel|a)}{P(\overline{novel}|a)}
 \end{equation}
 
-Therefore any novelty system that decides to acknowledge a sample
-$a$ as novel should also detect novelty on any sample $b$ satisfying
-$P(\overline{novel}|b) < P(\overline{novel}|a)$.
-As that would always reduce the error probability.
-This way a novelty detection system is interested in defining an
-order relation on all the possible inputs equivalent to the order defined by
-the error rate: $P(\overline{novel}|x)$.
+This relation between $a$ and $b$ can further be simplified into:
+
+\begin{equation}
+P(\overline{novel}|b) < P(\overline{novel}|a)
+\end{equation}
+
+Based on this, it can be said that an optimal novelty detection system is
+interested in defining an order relation on all the possible inputs equivalent
+to the order defined by the function: $P(\overline{novel}|x)$.
+And any optimal detector can be described by the largest $P(\overline{novel}|x)$
+accepted by it. This can be seen as thresholding.
 
@@
 
+Note, however that in the case of dynamic graph structures, an assumption on a constant
+$P(novel)$ is quite strong. And although on this first approach it is assumed to be
+constant, the authors acknowledge that structure plays an important role and should
+be used as prior-information when calculating $P(novel)$.
+
 % Explain approach of using conditional probability
@@
 \label{sec:conditional-prob}
-
 The conditional probability models the distribution of variables given that the sample
@@
 \centering
-\includegraphics[width=0.70\textwidth]{figures/conditional-prob-graph.pdf}
-\caption{\label{fig:conditional-prob-graph}A sample of a factor graph modelling a distribution
-         of a set $x$ of sensed variables.}
+\includegraphics[width=0.65\textwidth]{figures/conditional-prob-graph.pdf}
+\caption{\label{fig:conditional-prob-graph}Illustration of a factor graph modelling
+         a distribution of a set $x$ of sensed variables.}
 \end{figure}
 
-\autoref{fig:conditional-prob-graph} presents a possible graph $G$ built from the conceptual
+\autoref{fig:conditional-prob-graph} illustrates a graph $G$ built from the conceptual
 layer to represent the conditional probability on the sensed variables $x$.
@@
 common knowledge to build potentials describing how likely it is that a specific set of
-values for a set of variables is more likely to occur over than any other set of values.
+values for a set of variables is likely to occur.
 For instance: it is very likely to find a cereal box in a kitchen; and it is unlikely to find
 bathroom connected to another bathroom.
-
 We propose using such a graph $G$ built by the conceptual layer for modelling $P_G(x)$
@@
 
-\begin{figure}
+\begin{figure}[h]
 \centering
-\includegraphics[width=0.70\textwidth]{figures/uniform-prob-graph.pdf}
+\includegraphics[width=0.65\textwidth]{figures/uniform-prob-graph.pdf}
 \caption{\label{fig:uniform-prob-graph}Without any existing factors, this graph $U$ represents a
@@
 For example the conditional probability would yield very small values on large sets of variables
-$x$ than in small sets due to the spreading over the dimensions of the input space.
+$x$ than in small sets due to the spreading over the dimensions of the sample space.
 As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
@@
 \subsection{Semi Supervised: Using Unlabelled Data}
-Nonetheless, its often the case that there is access to large amounts of unlabelled data.
-Under that it becomes possible to obtain a better approach to the unconditional probability
-distribution than the uniform one.
+Nonetheless, its often the case that there is access to extra data that allows to
+obtain a better approximation to the unconditional probability than the uniform one.
+In specific, all the knowledge of the agent can be considered to hold true apart from
+complete knowledge on the categories of a room.
+In that case a single big factor can be used to model all the variables directly
+dependent on the room which the category cannot be assumed to be known as illustrated in
+\autoref{fig:unconditional-prob-graph}.
 
-For practical reasons it was assumed that the unconditional distribution
-could be modelled with all the variables independent from each other.
-Which translates as $P(x)=\prod{P_{x_i}(x_i)}$.
-From the unlabelled data, only the probability of each feature $x_i$ needs to be estimated.
-\autoref{fig:semisupervised-threshold} illustrates a graph $I$ to model independent variables.
+\begin{figure}
+\centering
+\includegraphics[width=0.65\textwidth]{figures/unconditional-prob-graph.pdf}
+\caption{\label{fig:unconditional-prob-graph}Without being able to model variable $R_1$
+all the variables directly dependent on it become dependent between each other introducing
+a single big factor.}
+\end{figure}
+
+For practical reasons it is impossible to train such a factor, and simplifications need
+to be performed. And here, it was assumed that it can be approximated by factorizing
+it in several single factors such that all variables become independent.
+Additionally those single factors can easily be trained by using unlabelled data,
+obtaining this way a graphical model $I$,
+as illustrated in \autoref{fig:single-unconditional-prob-graph},
+to be used as approximation for the unconditional probability.
 
@@
 \centering
-\includegraphics[width=0.70\textwidth]{figures/independent-prob-graph.pdf}
-\caption{\label{fig:semisupervised-threshold}Using unlabelled data, its possible to estimate
-         the unconditional probability for each variable and this way try to compensate
-         for a possible existent bias on some of the variables.}
+\includegraphics[width=0.65\textwidth]{figures/single-unconditional-prob-graph.pdf}
+\caption{\label{fig:single-unconditional-prob-graph}By factorizing the single factor
+introduced by room 1 not being necessarily known, several single factors are obtained
+that can be trained from unlabelled data.}
 \end{figure}
@@
 in account.
+In specific the point on this initial work was to test the correctness of the
+presented threshold function and approximation capabilities by using an uniform or
+independent model.
+
 The synthetic distribution assumes that an independent and variable size set of features

### 2011-07-01 semantic-nd/article.tex
Reviewed text about experiments.

@@
 In order to verify the performance of the proposed threshold functions a synthetic dataset
-was generated. To keep it simple only information regarding direct features of
-a room were modelled and no structured knowledge such as room connectivity was taken
+was generated. As the point on this initial work was only to test the correctness of
+the presented threshold function and approximation capabilities by using a uniform or
+an independent model, only information regarding direct features of a room were
+modelled and no structured knowledge such as room connectivity was taken
 in account.
-In specific the point on this initial work was to test the correctness of the
-presented threshold function and approximation capabilities by using an uniform or
-independent model.
 
-The synthetic distribution assumes that an independent and variable size set of features
+The synthetic distribution assumes that an independent and variable number of features
 $x$ is generated by a given room category.
-In whole there was 11 different room categories and 7 different measured feature
-types. Each feature type can be present more than once in (e.g.: room shape is
-extracted from 2D laser scans in more than one position in the room).
+In whole there was 11 different room categories and 9 different measured feature
+types. The number of sensed features is dynamic and mimic the type of information
+extracted when running on the robot. Due to that it is possible that on a given sample
+a certain feature type can be present more than once or not be present at all
+(e.g.: room shape is extracted from 2D laser scans in more than one position in the room,
+information about detected objects is only present when the robot previously tried to
+detect objects on a given room).
 
-The room categories were chosen to mimic as close as possible the real features and
-categories existing in reality. And they model different room categories with
-different levels of detail. For example: 1 person office, 2 person office, hallway,
-robot lab.
+The sensed properties and room categories were chosen to mimic as close as possible
+the reality and they are based on a previously build ontology from web data.
+There is in total 11 different room categories ranging from: corridor, hallway,
+1 person office, 2 person office, bathroom, conference hall, etc\dots, and there is
+9 different extracted features: room size, room shape, room appearance and 6 different objects
+(e.g. book, cereal box, computer).
 
 From the distribution, 100 labelled samples for 5 of the 11 room categories were
-drawn to represent the known concepts and 1000 unlabelled samples were drawn from
+drawn to represent the known categories and 1000 unlabelled samples were drawn from
 all the room categories for learning the unconditional probability distribution.
@@
 and $P_G(x)/P_I(x)$ assuming an independent unconditional distribution.
-Since the distribution is synthetic there is access to $P(x)$ and $P(x|concept)$
+Since the distribution is synthetic there is access to $P(x)$ and $P(x|\overline{novel})$
 and a perfect threshold function could also be created to test how far the
@@
 % Show that the thresholds are suitable functions for implementing a static threshold.
-As first tryout, the performance of the novelty threshold selection was plotted for a set
-of 1000 samples taken out from the whole distribution (\autoref{fig:synthetic-roc}).
-Those samples contain different number of sensed properties for each room, mimicking
+First, the performance of the novelty threshold selection was plotted for a set
+of 1000 samples taken from the whole distribution (\autoref{fig:synthetic-roc}).
+The samples where uniformly generated by graphs with 5, 10, 15, 20, 35, 50 features.
+Additionally the feature types were also uniformly sampled, for that it is possible
+that in certain samples some feature types were sensed more than once and other were not
+sensed at all.
+This was chosen to mimic
 the dynamic properties expected to see when implemented on a robot.
@@
 
-The convex shape for the optimal threshold shows that the ratio between conditional
-and unconditional probability is indeed an optimal detector and is suitable for
-implementing a threshold when the samples are taken from dynamic
-distributions (e.g.: some samples where there is only access to room size versus
+The convex shape of the optimal threshold shows that the ratio between conditional
+and unconditional probability is indeed a suitable detector for implementing a threshold when
+the samples are taken from dynamic distributions when $P(novel)$ is constant
+(e.g.\ some samples where there is only access to room size versus
 samples where there is a lot of information about the room properties).
@@
 In order to measure the performance impact as more semantic information becomes
-available ROC curves were plotted for samples grouped by the number of sensed
+available, ROC curves were plotted for samples grouped by the number of sensed
 semantic features.
@@
 
-We can see that as the system gains more semantic information it becomes easier
-to detect novelty. The input space size increases and allows the several existing
-classes to become more easily distinguished.
+It is possible to see that as the system gains more semantic information, it
+becomes easier to detect novelty. The size of the input space increases and allows the
+existing classes to become more easily distinguished.
 
@@
 able to model the existent dependence between the features. This becomes obvious
-as the number of features increases (e.g.: graph $I$ perfectly models $P(x)$ in the
+as the number of features increases (e.g.\ graph $I$ perfectly models $P(x)$ in the
 case where only 1 feature is sensed).
 
-The uniform threshold shows a poor performance specially on small size samples
-where its performs almost no better than random.
-It performance increases as the size of sensed features increases but nonetheless
-its very small when compared to how optimal a threshold could be.
+The uniform threshold shows poor performance especially for samples with small amount of features
+where it performs almost no better than random.
+The performance increases as the size of sensed features increases but nonetheless
+is very small when compared to the optimal threshold.
 

### 2011-07-01 semantic-nd/article.tex
Small fixes on conclusion and future work.

@@
 unconditional probability and when perfect information exists it performs an optimal
-novelty detection threshold.
+novelty detection threshold under the assumption that $P(novel)$ is constant across
+all the graph structures.
 
 It was also shown that a correct estimation of unconditional probability plays an
-important role specially on small input spaces. Moreover, semi-supervised techniques
-implemented with the access to unlabelled data can be used to significantly improve
+important role specially on small input spaces. Moreover semi-supervised techniques,
+implemented with the access to unlabelled data, can be used to significantly improve
 novelty detection performance.

### 2011-07-01 semantic-nd/article.tex
Added authors and institutes
First part of corrections of secrion 1-3.

@@
 % the running heads and the author index.
-\author{Removed for Blind Review}
-%\author{André Susano Pinto, Andrzej Pronobis, Luis Paulo Reis}
+\author{André Susano Pinto\inst{1}\thanks{Corresponding author: \email{andresusanopinto@gmail.com}} \and Andrzej Pronobis\inst{2} \and Luis Paulo Reis\inst{1}}
 %
@@
 % unless you accept that it will be published
-\institute{Removed for Blind Review}
-%\institute{Faculdade de Engenharia da Universidade do Porto, Portugal\\
-%\url{andresusanopinto@gmail.com}\\
-%\url{http://url}}
+\institute{Faculdade de Engenharia da Universidade do Porto, Portugal,
+           \and Royal Institute of Technology, Stockholm, Sweden}
 
@@
 room categorization i.e. detecting whether the area in the environment identified as a
-separate room can be assigned one of the semantic labels that the robot knows (e.g. as \emph{a kitchen} 
-or \emph{an office}) or belongs to a yet unknown semantic category.
+separate room can be assigned to one of the semantic labels that the robot knows (e.g. \emph{a kitchen} 
+or \emph{an office}) or belongs to an unknown semantic category.
 
-The novelty detection algorithm is implemented within a cognitive robot Dora the Explorer,
-which already possesses a developed architecture based on probabilistic graphical models 
-oriented towards dealing with uncertain semantic information and reasoning about it~\cite{ijcai}.
-One of the major problems in using the probabilistic models of the environments for novelty
+The novelty detection algorithm is implemented on a cognitive robot Dora the Explorer,
+which already uses a developed architecture based on probabilistic graphical models 
+oriented towards dealing with and reasoning about uncertain semantic information~\cite{ijcai}.
+One of the major problems with using the probabilistic models of the environments for novelty
 detection is selecting the optimal threshold above which the test sample is considered
@@
 grows as the robot explores the environment. Methods are required to find
-optimal parameters given the current structure of the model. To this end, this work 
+optimal parameters given the current structure of the model which constantly changes. To this end, this work 
 studies methods for novelty threshold selection using probabilistic graphical models.
 
-The rest of this paper is structured as follows. First, we briefly review the related work on
+The rest of this paper is structured as follows. First, we briefly review the related work related to
 room categorization and novelty detection (Section~\ref{sec:related}). Then, we give an outline of the 
-architecture of Dora and the structure of the conceptual map representing the semantic information
+architecture of the Dora system and the structure of the conceptual map representing the semantic information
 (Section~\ref{sec:overview}). Next, we discuss the methods for novelty detection (Section~\ref{sec:novelty-detection}) 
@@
 The problem of room categorization based on visual information was first addressed
-in the computer vision community. In this case, the research focused mainly on the
+in the computer vision community. The research focused mainly on the
 problem of classifying single images captured in indoor or outdoor environments
@@
 \includegraphics[width=0.9\textwidth]{figures/processes.pdf}
-\caption{\label{fig:conceptual-layer}Interaction of the sub-systems
-         in Dora with special focus on the conceptual layer.}
+\caption{\label{fig:conceptual-layer}Interaction between the sub-systems
+         of Dora with special focus on the conceptual layer.}
 \end{figure}
@@
  \item[Doorway detection] is used to segment the continuous space into rooms and map connectivity between them.
- \item[Room size and shape] are classified by using 2D laser scans data from laser range finder mounted on the robot and are used as properties of a given room.
+ \item[Room size and shape] are obtained by classifying 2D laser scans from laser range finder mounted on the robot and are used as properties of a room.
 The system utilizes pre-trained set of classifiers to extract rooms sizes (either large, medium or small) and shapes (rectangular or elongated).
- \item[Object detection] is performed in the images acquired by the robot through its camera.
+ \item[Object detection] is performed in images acquired by the robot through its camera.
 The system keeps track of the number of objects of each type in each room.
@@
 As Dora moves through the environment its \emph{conceptual layer} builds a structural and
-probabilistic representation of the space instantiated as a \emph{graphical model}.
+probabilistic representation of space instantiated as a \emph{graphical model}.
 It includes taxonomy of human-compatible spatial concepts which are linked to the sensed 
@@
 \includegraphics[width=0.80\textwidth]{figures/chaingraph.pdf}
-\caption{\label{fig:chain-graph}Example chain-graph produced by the \emph{conceptual layer}.}
+\caption{\label{fig:chain-graph}Example of a chain-graph produced by the \emph{conceptual layer}.}
 \end{figure}
@@
 easier manipulation due to factorization.
-Moreover, there exists efficient implementation of inference engines operating on factor graph
+Moreover, there exist efficient implementations of inference engines operating on factor graph
 representations~\cite{Mooij_libDAI_10}.
 Describing the distribution function in terms of graphs allows to use those engines to
-efficiently calculate marginals on any given subset of variables by exploiting conditional
+efficiently calculate marginals over any given subset of variables by exploiting conditional
 independence between variables.
@@
 representing random variables and factors.
-Each factor is described by function $\phi$ dependent only on the variables $x_\phi$
+Each factor is described by a function $\phi$ dependent only on the variables $x_\phi$
 to which the factor is connected.

### 2011-07-01 semantic-nd/article.tex
Next part until 4.1 (excluding).

@@
 % Introduce novelty detection
-Novelty detection deals with detecting that a data sample belongs to a class
-originated by a distribution other than the ones the system knows about~\cite{markou2003novelty}.
+The aim of novelty detection is to identify data samples originating from a distribution different than one of those
+the system knows about~\cite{markou2003novelty}.
 It is harder than classification as only positive samples of the class are available
-rendering normal classification methods unusable.
-
-Adding capabilities to an agent to detect novel samples allows to increase its
-reliability. Novelty signal can then be interpreted by the system to proceed
-with more caution as its knowledge does not correctly describe reality.
+rendering normal classification methods unusable. Adding novelty detection capabilities allows to increase 
+reliability of the system. Novelty signal can be used to inform the system that it should proceed
+with caution as its knowledge does not correctly describe the environment.
 
@@
 Due to the nature of the sensed data which are noisy and uncertain, novelty ought to be
-treated in a probabilistic way where each sample has certain probability
+treated in a probabilistic fashion. In such case, each sample is associated with certain probability
 of being generated by a class not known to the agent and a complementary probability $P(\overline{novel}|x)$
@@
 
-It leads that by extending the set $N$ with new samples, the true positive and
+It follows that by extending the set $N$ with new samples, the true positive and
 false positive rate can never be decreased, leading to the problem where in order
-to increase detection a system needs to increase its error. This describes the
+to increase detection, the system needs to increase its error. This describes the
 base of the \emph{error and rejection tradeoff}~\cite{chow1970optimum}, which
-states that a system aiming at increasing the true-positive probability will eventually increase its
-false-positive error.
+states that a system aiming at increasing the true-positive probability will 
+eventually increase its false-positive error.
 
@@
 probability without its false-positive probability increasing beyond a given limit.
-This is equivalent to a \emph{continuous knapsack problem} which allows a greedy
+This is equivalent to the \emph{continuous knapsack problem} which allows for a greedy
 solution by sorting the items with a value per weight function. In the case of 
-detection system that can be defined as:
+a detection system, this can be defined as:
 
@@
 
-Therefore a novelty detection system before classifying a sample $a$ as novel should (greedily)
+Therefore, a novelty detection system before classifying a sample $a$ as novel, should (greedily)
 classify any sample $b$ with a smaller cost as that would achieve a higher true positive probability 
@@
 
-Based on this, it can be said that an optimal novelty detection system is
-interested in defining an order relation on all the possible inputs equivalent
+Based on this, it can be said that an optimal novelty detection system is aims at 
+defining an order relation on all the possible inputs equivalent
 to the order defined by the function: $P(\overline{novel}|x)$.
-And any optimal detector can be described by the largest $P(\overline{novel}|x)$
-accepted by it. This can be seen as thresholding.
+Then, the detector can be described by the largest $P(\overline{novel}|x)$ accepted which is the principle of thresholding.
 
-Using Bayes rule and assuming a constant $P(\overline{novel})$
-a ratio between a conditional and unconditional probabilities of the input $x$ is obtained.
+Using Bayes rule and assuming a constant $P(\overline{novel})$,
+a ratio between conditional and unconditional probabilities of the input $x$ is obtained.
 Such a ratio is a suitable function for implementing a novelty detector system with optimal
@@
 Note, however that in the case of dynamic graph structures, an assumption on a constant
-$P(novel)$ is quite strong. And although on this first approach it is assumed to be
+$P(novel)$ is quite strong. Although in our first approach, it is assumed to be
 constant, the authors acknowledge that structure plays an important role and should
-be used as prior-information when calculating $P(novel)$.
+be used as prior information when calculating $P(novel)$.
 

### 2011-07-01 semantic-nd/article.tex
English fixes, many left, but I got no time or patience.

@@
 % the running heads and the author index.
-\author{André Susano Pinto\inst{1}\thanks{Corresponding author: \email{andresusanopinto@gmail.com}} \and Andrzej Pronobis\inst{2} \and Luis Paulo Reis\inst{1}}
+\author{André Susano Pinto\inst{1}\thanks{Email: \email{andresusanopinto@gmail.com}} \and Andrzej Pronobis\inst{2} \and Luis Paulo Reis\inst{1}}
 %
@@
 grows as the robot explores the environment. Methods are required to find
-optimal parameters given the current structure of the model which constantly changes. To this end, this work 
+the right threshold given the current structure of the model which constantly changes.
+To this end, this work 
 studies methods for novelty threshold selection using probabilistic graphical models.
@@
 
-Based on this, it can be said that an optimal novelty detection system is aims at 
+Based on this, it can be said that a novelty detection system aims at 
 defining an order relation on all the possible inputs equivalent
@@
 \label{sec:conditional-prob}
-The conditional probability models the distribution of variables given that the sample
-is not novel for the agent.
-This is: it should model the distribution of the sensed features considering that the
-knowledge the agent possesses holds true. Under that its natural to model it with a graphical
-model that combines the information acquired with the relations that the describe the
-relations learned by the agent between the variables.
+The conditional probability models the distribution of variables given that the agent
+knowledge holds true and can be used to describe the generating classes of the sensed
+sample.
+Under that its natural to model it with a graphical model that combines the learned
+variables and categories as well with the relations between them.
 
@@
 common knowledge to build potentials describing how likely it is that a specific set of
-values for a set of variables is likely to occur.
+values of certain variable types is likely to occur.
 For instance: it is very likely to find a cereal box in a kitchen; and it is unlikely to find
@@
 discarding the factor.
-Nonetheless, here the distributions are dynamically changing as the system learns
+Nonetheless, here the distributions are dynamically growing as the system learns
 more on the environment.
@@
 $x$ than in small sets due to the spreading over the dimensions of the sample space.
-As so a novelty measure is seen as a ratio on how much introducing the known concepts helps to
+A novelty measure is seen as a ratio on how much introducing the known concepts helps to
 understand the observed result.
@@
 In that case a single big factor can be used to model all the variables directly
-dependent on the room which the category cannot be assumed to be known as illustrated in
+dependent on the possibly novel room as illustrated in
 \autoref{fig:unconditional-prob-graph}.
@@
 
-For practical reasons it is impossible to train such a factor, and simplifications need
-to be performed. And here, it was assumed that it can be approximated by factorizing
+For practical reasons, it is impossible to train such a factor, and simplifications need
+to be performed. Here, it was assumed that it can be approximated by factorizing
 it in several single factors such that all variables become independent.
-Additionally those single factors can easily be trained by using unlabelled data,
-obtaining this way a graphical model $I$,
-as illustrated in \autoref{fig:single-unconditional-prob-graph},
+Additionally those single factors can easily be trained by using unlabelled data.
+Obtaining this way a graphical model $I$,
+illustrated in \autoref{fig:single-unconditional-prob-graph},
 to be used as approximation for the unconditional probability.
@@
 
-In order to verify the performance of the proposed threshold functions a synthetic dataset
+In order to verify the performance of the proposed threshold functions, a synthetic dataset
 was generated. As the point on this initial work was only to test the correctness of
@@
 (e.g.: room shape is extracted from 2D laser scans in more than one position in the room,
-information about detected objects is only present when the robot previously tried to
+and information about detected objects is only present when the robot previously tried to
 detect objects on a given room).
@@
 The sensed properties and room categories were chosen to mimic as close as possible
-the reality and they are based on a previously build ontology from web data.
+the reality and they are based on a previously built ontology from web data.
 There is in total 11 different room categories ranging from: corridor, hallway,
@@
 unconditional probability and when perfect information exists it performs an optimal
-novelty detection threshold under the assumption that $P(novel)$ is constant across
+novelty detection under the assumption that $P(novel)$ is constant across
 all the graph structures.

### 2011-07-04 thesis/abstract.tex
Some language fixes by LPR and Michael.

@@
 and represents it using a probabilistic graphical model. It develops methods 
-for performing detection of novel room categories which were not previously known to the robot applicable 
-to that graphical model. Finally, it draws attention to the usefulness of unlabelled data for the 
+for performing detection of novel room categories which were not previously known to the robot,
+using for that graphical models. Finally, it draws attention to the usefulness of unlabelled data for the 
 novelty detection process in order to considerably increase the detection accuracy.
@@
 
-Já à muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
+Já há muitos anos que a humanidade sonha que um dia robôs caminharam lado a lado com os humanos.
 Acredita-se que eles serão capaz de se mover e interagir connosco, compreender os nossos conceitos

### 2011-07-04 thesis/background.tex
Some language fixes by LPR and Michael.

@@
 But the classes are most of the time not linearly separable in the input
-space. Although there might exists a transformation $\phi$ from the original
+space. Although there might exist a transformation $\phi$ from the original
 space into a space $H$ where the input becomes linearly separable.
@@
 
-The most used method was presented by \cite{lowe1999object}. And its based on
+The most used method was presented by \cite{lowe1999object} and its based on
 building a feature vector for each image. Each of those features is based on
-\emph{interesting points} detected by detecting maxima and minima of a
+\emph{interesting points} detected by the presence of maxima and minima of
 difference of Gaussian functions applied in a scale-space.
@@
 \label{sec:gist}
-\cite{oliva2006building} argue that fast scene recognition does not need to be
+Oliva and Torralba~\cite{oliva2006building} argue that fast scene recognition does not need to be
 built on top of object recognition but can be analyzed by scene-centered
@@
 by a product of all the factors. In order to represent the probability,
-a normalization factor needs to be introduced, resulting in the following equation:
+a normalization factor $Z$ needs to be introduced, resulting in the following equation:
 

### 2011-07-04 thesis/semantic-mapping.tex
Some language fixes by LPR and Michael.

@@
 % motivation
-Humans associate concepts to areas that relate to the functionality of those. That
+Humans associate concepts to areas that relate to their functionality. That
 association is either useful by explaining where to find certain objects as well to
@@
 characteristic of the human world.
-That knowledge needs to define concepts and how to segment and identify them from the low-level representation,
-it needs also to define how the concepts relate to each other and attribute meanings to them.
+That knowledge needs to define and describe how to segment and identify concepts from the low-level representation.
+It also needs to define how the concepts relate to each other and attribute meanings to them.
 In the case of indoor environments this can be seen for example as: rooms are a unit of division of area and are
@@
 This thesis motivation lies on developing methods for detecting knowledge gaps on the spatial
-semantic knowledge of a mobile agent. For that spatial knowledge needs to be defined, as well
+semantic knowledge of a mobile agent. For that, spatial knowledge needs to be defined, as well
 the methods used to perform the semantic mapping.
@@
 \label{sec:cues-from-low-level}
-In a realistic scenario, classification tasks are by their nature uncertain.
+In a realistic scenario, classification tasks are, by their nature, uncertain.
 And although a trained classifier is able to decide which class is more likely, there is

### 2011-07-04 semantic-nd/article.tex
Disabled colorlinks so it looks nice when printing.
Fixed affliations on article.

@@
 }      
+\hypersetup{colorlinks=false}
 
@@
 % the running heads and the author index.
-\author{André Susano Pinto\inst{1}\thanks{Email: \email{andresusanopinto@gmail.com}} \and Andrzej Pronobis\inst{2} \and Luis Paulo Reis\inst{1}}
+\author{André Susano Pinto\inst{1,2} \and Andrzej Pronobis\inst{2} \and Luis Paulo Reis\inst{1,3}}
 %
@@
 % unless you accept that it will be published
-\institute{Faculdade de Engenharia da Universidade do Porto, Portugal,
-           \and Royal Institute of Technology, Stockholm, Sweden}
+\institute{Dep.\ Informatics Engineering, Faculty of Engineering of the University of Porto
+           \and Centre for Autonomous Systems, The Royal Institute of Technology (KTH)\\
+           SE100-44 Stockholm, Sweden
+           \and LIACC - Artificial Intelligence And Computer Science Lab.\ - University of Porto \\
+           Rua Dr. Roberto Frias, s/n 4200-465 Porto, Portugal}
 

### 2011-07-04 thesis/background.tex
Added summary on end of background and semantic mapping chapters.

@@
 which configuration on a variable or subset of variables maximizes a given function.
+
+\section{Summary}
+This chapter has introduced concepts used as base during the process of semantic mapping.
+For that, it introduced classification tasks in a generic way, it explains the
+difference between supervised and unsupervised approaches, how to obtain multi-class
+classifiers on top of two class classifications, and how to perform those operations
+in implicit spaces by using a kernel function.
+It also presented both local and global visual features that are often used in
+computer vision for either detecting objects or classifying appearances.
+Some probability theory bases were presented together with graphical models that
+allow to perform inference on several variables by creating a model of how variables
+directly influence each others.
+
+All this concepts were presented and papers and standard text books referenced, 
+with the purpose to make the reader familiar with the tools that are used
+to extract the higher level features that the semantic mapping process and how does
+the process infers semantic labels on the detected areas.

### 2011-07-04 thesis/semantic-mapping.tex
Added summary on end of background and semantic mapping chapters.

@@
 \end{figure}
+
+\section{Summary}
+This chapter has introduced the semantic mapping process that is used as base platform
+where to perform detection of novel semantic place categories.
+It presented the spatial knowledge representation used by it, and shows how the knowledge
+is used during the mapping process to identify entities and classify places according
+to a set of human defined semantics.

### 2011-07-08 thesis/synthetic/Graph-internal.h
Cleaning thesis chapter on a novel detection method based on potentials.
Due to lack of time this was not included on the thesis.
Plan is to write it decently for a nice article, explaining how to
perform novelty detection directly on P(novel|x) and workaround the
constant P(novel) assumption by using potentials of factor graphs.

@@ -1,23 +1,0 @@
-#ifndef _GRAPH_INTERNAL_H_
-#define _GRAPH_INTERNAL_H_
-
-#include <vector>
-#include "Graph.h"
-
-class Variable {
- public:
-	explicit Variable(const VariableType *type):type_(type) {}
-
-	const VariableType *type_;
-};
-
-class Factor {
- public:
-	Factor(const FactorData *data, const std::vector<const Variable*> &variables)
-			:data_(data), variables_(variables) {}
-
-	const FactorData *data_;
-	std::vector<const Variable*> variables_;
-};
-
-#endif // _GRAPH_INTERNAL_H

### 2011-07-08 thesis/synthetic/Graph.cc

@@ -1,202 +1,0 @@
-#include <boost/foreach.hpp>
-#include <cstdio>
-#include "Graph-internal.h"
-using namespace std;
-
-#define ROCS_DEBUG_STRINGIFY(x) #x
-#define ROCS_DEBUG_TO_STRING(x) ROCS_DEBUG_STRINGIFY(x)
-
-#define ROCS_DEBUG(...) { \
-  printf("[" __FILE__ ":" ROCS_DEBUG_TO_STRING(__LINE__) "] " __VA_ARGS__); \
-  printf("\n"); \
-}
-
-#define rocsDebug2(...) ROCS_DEBUG(__VA_ARGS__)
-#define rocsDebug1(...) ROCS_DEBUG(__VA_ARGS__)
-
-namespace {
-
-// Unfortunaly is_sorted is a SGI extension.
-// Therefore we need to define it.
-template<typename T>
-bool is_sorted(T begin, T end)
-{
-	if (begin == end)
-		return true;
-
-	T next = begin;
-	for (++next; next != end; begin = next, ++next)
-		if(*next < *begin)
-			return false;
-	return true;
-}
-
-// Builds a vector with the variable types of the given FactorData.
-// Returns false in case not all variables types are defined.
-bool GetVariableTypes(const FactorData &data,
-                      const map<string, VariableType> &known_types,
-                      vector<const VariableType*> *vartypes)
-{
-	typedef pair<string, string> PSS;
-	BOOST_FOREACH(const PSS &v, data.variables)
-	{
-		map<string, VariableType>::const_iterator it = known_types.find(v.second);
-		if (it == known_types.end())
-		{
-			rocsDebug2("Invalid FactorData: Unknown variable type '%s'",
-					v.second.c_str());
-			return false;
-		}
-		vartypes->push_back(&it->second);
-	}
-	return true;
-}
-}  // end namespace anonymous
-
-FactorData* GraphInformation::Add(const string &name, const FactorData &factor)
-{
-	rocsDebug2("Add factor type: '%s'", name.c_str());
-	if (factors.insert(make_pair(name, factor)).second == false)
-		rocsDebug1("Dropped factor: there is already a factor with that name '%s'",
-				name.c_str());
-    return &factors[name];
-}
-
-VariableType* GraphInformation::Add(const string &name, const VariableType &type)
-{
-	rocsDebug2("Add variable type: '%s'", name.c_str());
-	if (types.insert(make_pair(name, type)).second == false)
[…]

### 2011-07-08 thesis/synthetic/Graph.h

@@ -1,77 +1,0 @@
-#ifndef _GRAPH_H_
-#define _GRAPH_H_
-
-#include <algorithm>
-#include <boost/ptr_container/ptr_vector.hpp>
-#include <map>
-#include <string>
-#include <vector>
-
-struct VariableType {
-	// This vector must be sorted. This allows to use it in a simple
-	// way to calculate variable indexs.
-	std::vector<std::string> values;
-};
-
-struct FactorData {
-	std::vector<std::pair<std::string, std::string> > variables;
-	std::map<std::vector<std::string>, double> potential;
-};
-
-class GraphInformation {
- public:
-	std::map<std::string, VariableType> types;
-	std::map<std::string, FactorData>   factors;
-
-	VariableType* Add(const std::string &name, const VariableType &type);
-	FactorData*   Add(const std::string &name, const FactorData &factor);
-
-	// Returns true if the defined types and factors are consistent.
-	// Debug information is produced in the negative case.
-	bool CheckConsistency() const;
-
-	bool CheckVarTypeConsistency(const std::string &name,
-	                             const VariableType &type) const;
-
-	bool CheckFactorConsistency(const std::string &name,
-	                            const FactorData &factor) const;
-};
-
-
-class Variable;  // This class definition is never exposed externaly.
-class Factor;    // This class definition is never exposed externaly.
-class Query;
-
-class Graph {
- public:
-	Graph();
-	virtual ~Graph();
-
-	// Creates a new variable of the given type.
-	// Returns a reference to the newly created variable.
-	// VariableType must exist for the lifetime of this Graph.
-	// Graph maintains ownership on the returned variable.
-	virtual const Variable* CreateVariable(const VariableType* type);
-
-	// Creates a factor between the given variables.
-	// FactorData must exist for the lifetime of this Graph.
-	// Graph maintains ownership on the returned factor.
-	virtual const Factor* CreateFactor(const FactorData *factorData,
-	                           const std::vector<const Variable*> &vars);
-
-    // Appends all factors of this graph to the given output vector.
-	virtual void ListFactors(std::vector<const Factor*> *output) const;
-
- protected:
-	boost::ptr_vector<Variable> variables_;
-	boost::ptr_vector<Factor> factors_;
-
- private:
-	// DISALLOW COPY AND ASSIGN
-	Graph(const Graph &);
-	void operator = (const Graph &);
-};
-
-#endif  // _GRAPH_H_
-
-

### 2011-07-08 thesis/synthetic/Query.cc

@@ -1,279 +1,0 @@
-// ==================================================================
-// ROCS - Toolkit for Robots Comprehending Space
-// Copyright (C) 2011  André Susano Pinto
-//
-// This file is part of ROCS.
-//
-// ROCS is free software: you can redistribute it and/or modify it
-// under the terms of the GNU Lesser General Public License as
-// published by the Free Software Foundation, either version 3
-// of the License, or (at your option) any later version.
-//
-// ROCS is distributed in the hope that it will be useful,
-// but WITHOUT ANY WARRANTY; without even the implied warranty of
-// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-// GNU Lesser General Public License for more details.
-//
-// You should have received a copy of the GNU General Public License
-// along with ROCS. If not, see <http://www.gnu.org/licenses/>.
-// ==================================================================
-#include <cstdio>
-
-#define ROCS_DEBUG_STRINGIFY(x) #x
-#define ROCS_DEBUG_TO_STRING(x) ROCS_DEBUG_STRINGIFY(x)
-
-#define ROCS_DEBUG(...) { \
-  printf("[" __FILE__ ":" ROCS_DEBUG_TO_STRING(__LINE__) "] " __VA_ARGS__); \
-  printf("\n"); \
-}
-
-#define rocsDebug3(...) ROCS_DEBUG(__VA_ARGS__)
-#define rocsDebug2(...) ROCS_DEBUG(__VA_ARGS__)
-#define rocsDebug1(...) ROCS_DEBUG(__VA_ARGS__)
-
-
-#include "Query.h"
-
-#define DAI_WITH_JTREE
-#define DAI_WITH_BP
-#include <dai/alldai.h>
-#include <dai/factorgraph.h>
-#include <dai/gibbs.h>
-#include <dai/properties.h>
-#include <map>
-#include <string>
-
-#include "Graph-internal.h"
-
-using namespace std;
-
-// This handles all internals of Query class.
-// It should totally hide its dependency on libDAI.
-class QueryInternal {
-  public:
-	QueryInternal()
-	{
-		size_t maxiter = 10000;
-		dai::Real  tol = 1e-9;
-		size_t verbose = 1;
-
-		daiOptions_.set("maxiter", maxiter);  // Maximum number of iterations
-		daiOptions_.set("tol", tol);          // Tolerance for convergence
-		daiOptions_.set("verbose", verbose);
-	}
-
[…]

### 2011-07-08 thesis/synthetic/Query.h

@@ -1,73 +1,0 @@
-// ==================================================================
-// ROCS - Toolkit for Robots Comprehending Space
-// Copyright (C) 2011  André Susano Pinto
-//
-// This file is part of ROCS.
-//
-// ROCS is free software: you can redistribute it and/or modify it
-// under the terms of the GNU Lesser General Public License as
-// published by the Free Software Foundation, either version 3
-// of the License, or (at your option) any later version.
-//
-// ROCS is distributed in the hope that it will be useful,
-// but WITHOUT ANY WARRANTY; without even the implied warranty of
-// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-// GNU Lesser General Public License for more details.
-//
-// You should have received a copy of the GNU General Public License
-// along with ROCS. If not, see <http://www.gnu.org/licenses/>.
-// ==================================================================
-
-/*!
- * \author André Susano Pinto
- * \file Query.h
- */
-
-#ifndef _ROCS_CONCEPT_QUERY_H_
-#define _ROCS_CONCEPT_QUERY_H_
-
-#include <boost/scoped_ptr.hpp>
-#include <vector>
-#include <string>
-
-class Graph;
-class Variable;
-class QueryInternal;
-
-class Query {
- public:
-	// This class does not takes ownership over Graph.
-	// Graph and all associated variables have to outlive this instance.
-	explicit Query(const Graph *graph);
-	~Query();
-
-	// Calculates probability for a given set of variables.
-	//
-	// Output is the probability for all combinations over the given variables.
-	// Output is lexicographic ordered by the value of its variables.
-	// This means:
-	//   output[0] = P( Var1 = v0, Var2 = v0, ... VarN = v0]
-	//   output[1] = P( Var1 = v0, var2 = v0, ... VarN = v1]
-	//   ...
-	//   output[last] = P( Var1 = vN, Var2 = vN, ... VarN = vN]
-	void Marginalize(const std::vector<const Variable*> &variables,
-	                 std::vector<double> *output);
-    
-    void Sample(const std::vector<const Variable*> &variables,
-                std::vector<std::string> *output);
-
-    double LogZ(const std::vector<const Variable*> &clamp_vars,
-                const std::vector<std::string> &clamp_state);
- protected:
-	const Graph *graph_;
-	boost::scoped_ptr<QueryInternal> internal_;
-
[…]

### 2011-07-08 thesis/synthetic/main.cc

@@ -1,46 +1,0 @@
-/**
- * Novelty Detection Scripts
- *
- * Copyright 2011: André Susano Pinto
- *
- * This program is free software; you can redistribute it and/or
- * modify it under the terms of the GNU General Public License
- * as published by the Free Software Foundation; either version 2
- * of the License, or (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, write to the Free Software Foundation,
- * Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
- */
-#include "Graph.h"
-#include "Graph-internal.h"
-#include "Query.h"
-#include "util.h"
-
-#include <cassert>
-#include <cstdio>
-#include <set>
-#include <algorithm>
-using namespace std;
-
-#include <boost/foreach.hpp>
-#include <boost/scoped_ptr.hpp>
-using namespace boost;
-
-#include <dai/factorgraph.h>
-#include <dai/gibbs.h>
-#include <dai/properties.h>
-using namespace dai;
-
-void example_single_factor();
-
-int main() {
-  example_single_factor();
-  return 0;
-}
-

### 2011-07-08 thesis/synthetic/single-factor.cc

@@ -1,460 +1,0 @@
-/**
- * Novelty Detection Scripts
- *
- * Copyright 2011: André Susano Pinto
- *
- * This program is free software; you can redistribute it and/or
- * modify it under the terms of the GNU General Public License
- * as published by the Free Software Foundation; either version 2
- * of the License, or (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, write to the Free Software Foundation,
- * Inc., 51 Franklin Street, Fifth Floor, Boston, MA 02110-1301, USA.
- */
-#include "Graph.h"
-#include "Graph-internal.h"
-#include "Query.h"
-#include "util.h"
-
-#include <cassert>
-#include <cstdio>
-#include <set>
-#include <iostream>
-#include <fstream>
-#include <algorithm>
-#include <cmath>
-using namespace std;
-
-#include <boost/foreach.hpp>
-#include <boost/scoped_ptr.hpp>
-using namespace boost;
-
-/**
- * This experiment tries to approach a graph where variable
- * a is only connected with a single type of factor \phi.
- * 
- */
-namespace {
-string room(int i) { return "r" + to_string(i); }
-string prop(int i) { return "p" + to_string(i); }
-
-vector<string> vector_of(const string &a, const string &b) {
-  vector<string> t(2);
-  t[0] = a, t[1] = b;
-  return t;
-}
-
-struct GraphStructure {
-  int rooms;
-  int senses;
-  set<pair<int,int> > edges;
-
-  void createRandom(int rooms, int senses, int extra_conn) {
-    this->rooms = rooms;
-    this->senses = senses;
-    // Creates a tree
-    for (int i = 1; i < rooms; ++i) {
-      int other = random() % i;
-      edges.insert(make_pair(other, i));
-    }
-    // Tries to add extra room connectivity
-    while (extra_conn--) {
-      int a = random() % rooms;
-      int b = random() % rooms;
-      if (a == b) continue;
-      if (a > b) swap(a, b);
[…]

### 2011-07-08 thesis/synthetic/util.h

@@ -1,10 +1,0 @@
-#include <string>
-#include <sstream>
-
-template<class T>
-std::string to_string(const T a) {
-  std::ostringstream os;
-  os << a;
-  return os.str();
-}
-

### 2011-07-10 thesis/presentation.tex
Presentation layout and scons rules added.

\documentclass[compress]{beamer}
\usepackage[utf8x]{inputenc}
\usepackage{default}
\usepackage{graphics}
\useinnertheme{rounded}
\usecolortheme{whale}
\useoutertheme[subsection=false]{miniframes}

\title{Novelty Detection for Semantic Place Categorization}
\author{André Susano Pinto\inst{1,2}}
\date{18 July 2011}

\institute {
 Supervised by: Andrzej Pronobis\inst{1}, Luis Paulo Reis\inst{2}
 \and
 \inst{1}The Royal Institute of Technology (KTH), Sweden \\
 \inst{2}Faculdade de Engenharia da Universidade do Porto
}

\begin{document}

\begin{frame}
 \titlepage
\end{frame}

\begin{frame}{Outline}
 \tableofcontents
\end{frame}

\section{Motivation}
\begin{frame}
\cite{pronobis2011phd}
\end{frame}

\section{Semantic Mapping}
\subsection{Spatial Knowledge}
\subsection{Conceptual Map}

\section{Novelty Detection}
\subsection{Novelty Detection as Thresholding}
\subsection{Conditional and Unconditional Ratio}

\section{Conclusion}
\subsection{Limitations}
\subsection{Future Work}

\appendix

\begin{frame}[allowframebreaks]{Bibliography}
\bibliographystyle{alpha}
\bibliography{refs}
\end{frame}

\begin{frame}
 \titlepage
\end{frame}

\end{document}

### 2011-07-13 pdis/presentation2/cold-database-sample.tex
Fixed scons rule for PDIS/presentation2

@@
 \begin{center}
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR}
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_PA}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_LO}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_2PO}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_CR.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_TR.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_PA.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_LO.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_LAB.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_2PO.jpg}
 
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_RL}
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_PA}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_CR}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_RL.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_PA.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_1PO.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_TL.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_CR.jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_PA.jpg}
 
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_CR(1)}
-\includegraphics[width=0.15\textwidth]{cold/Saarbruecken_TR(1)}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_1PO(1)}
-\includegraphics[width=0.15\textwidth]{cold/Freiburg_TL(1)}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_LAB(1)}
-\includegraphics[width=0.15\textwidth]{cold/Ljubljana_PA(1)}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_CR(1).jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Saarbruecken_TR(1).jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_1PO(1).jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Freiburg_TL(1).jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_LAB(1).jpg}
+\includegraphics[width=0.15\textwidth]{figures/cold/Ljubljana_PA(1).jpg}
 \end{center}

### 2011-07-13 pdis/presentation2/presentation.tex
Fixed scons rule for PDIS/presentation2

@@
 \useoutertheme[subsection=false]{miniframes}
-\graphicspath{{figures/}}
 
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{place_classification}
+\includegraphics[width=0.75\textwidth]{figures/place_classification.png}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=0.75\textwidth]{novelty_detection}
+\includegraphics[width=0.75\textwidth]{figures/novelty_detection.png}
 \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=.9\textwidth]{classification.pdf}
+        \includegraphics[width=.9\textwidth]{figures/classification.pdf}
     \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=.75\textwidth]{Svm_max_sep_hyperplane_with_margin}
+        \includegraphics[width=.75\textwidth]{figures/Svm_max_sep_hyperplane_with_margin.png}
     \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=\textwidth]{sift/sift}
+        \includegraphics[width=\textwidth]{figures/sift/sift.pdf}
         \caption{SIFT and other local features have been proven useful in object detection.}
@@
     \begin{figure}
-    \includegraphics[width=\textwidth]{crfh_model}
+    \includegraphics[width=\textwidth]{figures/crfh_model.jpg}
     \caption{Composed Receptive Field Histograms can be used to extract appearance.}
@@
     \begin{figure}
-    \includegraphics[width=0.60\textwidth]{laser_and_depth_input}
+    \includegraphics[width=0.60\textwidth]{figures/laser_and_depth_input.pdf}
     \caption{Lasers can be used to extract useful features.}
@@
     \begin{figure}
-        \includegraphics[width=0.75\textwidth]{novelty.pdf}
+        \includegraphics[width=0.75\textwidth]{figures/novelty.pdf}
     \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=0.9\textwidth]{ringlinesquare_kpca}
+        \includegraphics[width=0.9\textwidth]{figures/ringlinesquare_kpca.jpg}
     \end{figure}
@@
         \begin{figure}
-            \includegraphics[width=0.5\textwidth]{graphical-models/BayesNet.pdf}
+            \includegraphics[width=0.5\textwidth]{figures/graphical-models/BayesNet.pdf}
             \caption{Bayes Networks}
@@
         \begin{figure}
-            \includegraphics[width=0.5\textwidth]{graphical-models/MarkovRandomField}
+            \includegraphics[width=0.5\textwidth]{figures/graphical-models/MarkovRandomField}
             \caption{Markov Random Field}
@@
         \begin{figure}
-            \includegraphics[width=0.60\textwidth]{graphical-models/FactorGraph.pdf}
+            \includegraphics[width=0.60\textwidth]{figures/graphical-models/FactorGraph.pdf}
             \caption{Factor Graph}
@@
     \begin{figure}
-        \includegraphics[width=0.65\textwidth]{system-representation/layers.pdf}
+        \includegraphics[width=0.65\textwidth]{figures/system-representation/layers.pdf}
     \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=\textwidth]{system-representation/graph.pdf}
+        \includegraphics[width=\textwidth]{figures/system-representation/graph.pdf}
     \end{figure}
@@
     \begin{figure}
-        \includegraphics[width=\textwidth]{system-representation/novelty-detection.pdf}
+        \includegraphics[width=\textwidth]{figures/system-representation/novelty-detection.pdf}
     \end{figure}
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{gantt.pdf}
+\includegraphics[width=\textwidth]{figures/gantt.pdf}
 \end{figure}

### 2011-07-14 thesis/introduction.tex
little typo

@@
 for humans. A robot may be able to detect a tag placed on an object, but such an approach does not scale
-outsize structured environments: new objects are created everyday, there is huge amount of already
+outside structured environments: new objects are created everyday, there is huge amount of already
 existing objects and concepts by their nature are dynamic. For this reason, methods need 

### 2011-07-14 thesis/presentation.tex
Added draft of thesis presentation.
Most images come from wikipedia.

@@
 \documentclass[compress]{beamer}
-\usepackage[utf8x]{inputenc}
 \usepackage{default}
 \usepackage{graphics}
+\usepackage{ulem}
+\usepackage{multicol}
+
 \useinnertheme{rounded}
@@
 \useoutertheme[subsection=false]{miniframes}
+
+%%%%%%%% begin tikz %%%%%%
+\usepackage{tikz}
+\usetikzlibrary{shapes.callouts,decorations.pathmorphing}
+\usetikzlibrary{shapes,decorations,shadows}
+\usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{decorations.shapes}
+\usetikzlibrary{fadings}
+\usetikzlibrary{patterns}
+\usetikzlibrary{calc}
+\usetikzlibrary{decorations.text}
+\usetikzlibrary{decorations.footprints}
+\usetikzlibrary{decorations.fractals}
+\usetikzlibrary{shapes.gates.logic.IEC}
+\usetikzlibrary{shapes.gates.logic.US}
+\usetikzlibrary{fit,chains}
+\usetikzlibrary{positioning}
+\usepgflibrary{shapes}
+\usetikzlibrary{scopes}
+\usetikzlibrary{arrows}
+\usetikzlibrary{backgrounds}
+
+\pgfdeclarelayer{background}
+\pgfdeclarelayer{foreground}
+\pgfsetlayers{background,main,foreground}
+
+\tikzset{latent/.style={circle,fill=white,draw=red,thick,inner sep=1pt, 
+minimum size=20pt, font=\fontsize{10}{10}\selectfont},
+obs/.style={latent,fill=gray!25},
+const/.style={rectangle, inner sep=0pt},
+factor/.style={rectangle, fill=red,minimum size=7pt, inner sep=0pt},
+yellow/.style={latent,minimum size=15pt,fill=yellow!75},
+blue/.style={latent,minimum size=15pt,fill=blue!75},
+-/.style={color=red, thick},
+>={triangle 45}}
+
+% shapename, fitlist, caption, pos
+\newcommand{\plate}[4]{
+\node (invis#1) [draw, transparent, inner sep=1pt,rectangle,fit=#2] {};
+\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=0pt,yshift=-9pt] {\raisebox{0pt}[0pt]{\footnotesize{#3}}};
+\node (#1) [draw=black!50,thick,inner sep=3pt,rectangle,rounded corners,fit=(invis#1) (capt#1),#4] {};
+}
+
+\newcommand{\shiftedplate}[5]{
+\node (invis#1) [draw, transparent, inner sep=0 pt,rectangle,fit=#2] {};
+\node (capt#1) [#5, xshift=2pt] {\footnotesize{#3}};
+\node (#1) [draw,inner sep=2pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+}
+%shapename, pos, caption, in, out, captpos
+\newcommand{\factor}[6]{
+\node (#1) [factor] at #2 {};
+\node (capt#1) [#6 of #1]{\footnotesize{#3}};
+\draw [-] (#4) -- (#1) ;
+\draw [->,thick] (#1) -- (#5);
+}
+
+% name, --, caption, pos
+\newcommand{\nofactor}[4]{
+\node (#1) [factor, #2]  {};
+\node (capt#1) [#4 of #1]{\footnotesize{#3}};
+}
 
@@
 \begin{document}
-
 \begin{frame}
@@
 \begin{frame}{Outline}
+  \begin{multicols}{2}
     \tableofcontents
+  \end{multicols}
 \end{frame}
 
-\section{Motivation}
-\begin{frame}
-\cite{pronobis2011phd}
+\section{Introduction}
+\subsection{Motivation}
+\begin{frame}{Motivation}
+
+\begin{columns}[t]
+  \column{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/extra/PR2.jpg}
+
+  \centering
+  PR2
+
+  \column{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/extra/servingbot.jpg}
+
+  \centering
+  Serving Robot
+
+  \column{0.3\textwidth}
+  \includegraphics[width=\textwidth]{figures/extra/wakamaru.jpg}
+
+  \centering
+  Wakamaru
+\end{columns}
+\vfill
+
+\begin{itemize}
+  \item Robots are useful: can make our lives easier
+  \item Robots are entering our Houses and Offices
+  \item They need to understand our world and communicate with us.
+\end{itemize}
+\note{
+1 - Humans dream on creating robots. 
+Both by the challenge on creating intelligent systems on itself but also to make our life easier.
+% I should have an example list on how they could be useful
+
+2 - It is becoming a reality and robots will enter our houses and offices.
+The question is no longer if we will have a robot at home/office, but how many.
+
+3 - They will need to understand and communicate with us and our human worlds.
+The need of semantic information!
+}
 \end{frame}
 
+\subsection{Problem}
+\begin{frame}{Problem}
+
+  Enhance world with semantic information is \alert{not a scalable solution}.
+
+  \begin{block}{}
+  \begin{columns}
+  \column{0.3\textwidth}
+    \begin{figure}
+      \includegraphics[width=\textwidth]{figures/extra/RFID.pdf} \\
+      RFID for object tagging
+    \end{figure}
+  \column{0.3\textwidth}
+    \begin{figure}
+      \includegraphics[width=\textwidth]{figures/extra/floorplan.jpg} \\
+      Maps extended with semantic information
+    \end{figure}
+  \end{columns}
+  \end{block}
+
+\note{
+1 - A solution could pass by tagging our world with extra information, such as maps, rfids, but such a solution
+is not practical and scalable.
+}
+\end{frame}
+
+\begin{frame}{Problem}
+\begin{block}{Machine Learning}
+Give us methods to solve this problem in reverse by providing robots with knowledge
+on how to detect and classify concepts and categories from the underlying sensed world.
+\end{block}
+
+\begin{block}{Unreliable on Dynamic and Unknown Environments}
+It is unrealistic to believe on the possibility to fully describe the required knowledge
+for the agent in all situations.
+\end{block}
+
+\note{
+Machine Learning tries to solve the reverse semantic mapping problem.
+
+It endows robots with knowledge that allows them to map what they sense to human concepts
+and categories.
+
+Leading to the need of methods that are able to inspect and smartly identifying information directly
+from low-level sensors.
+
+For that several machine learning techniques have been developed, but all of those require and represent
+knowledge that is given a-priori.
+}
+\end{frame}
+
+\begin{frame}{Problem}
+
+\begin{block}{Detection of Knowledge-Gaps}
+Detection of situations where the agent knowledge does not suffices.
+\end{block}
+
+\begin{block}{Focus on Detection of Novel Semantic Categories of Places}
+
+\end{block}
+\note{
+- It is unrealistic to believe on the possibility to describe all the required knowledge
+
+- When dealing with novel and unknown scenarios an agent has to be able to detect something
+is not conforming to what it has learn.
+
+In order to be reliable agents need to detect situations where their knowledge
+And agents be reliable even when situations
+- Scalability
+- Dynamic Aspects
+- Long Term
+- Life Long
+
+It is unrealistic to believe on the possibility to describe knowledge that holds to all the possible
+scenarios, and when dealing with long-term and life-long agents
+
+- Variability, dynamic aspects of environment, 
+- Should be able to deal with novel cases
+- 1st step is to detect them
+}
+\end{frame}
+
+\subsection{Goals}
+\begin{frame}{Goals}
+  \begin{itemize}
+    \item Study a semantic mapping process in the context of mobile robots \cite{pronobis2011phd}.
+    \item Propose a method based on the studied system to detect novel semantic categories of places.
+  \end{itemize}
+
+\note{
+This thesis goes on the direction of detecting knowledge gaps on mobile robots.
+
+It studies a process of semantic mapping introduced by Pronobis
+and studies how to detect those situations where the existent knowledge does not describes
+the reality.
+But being such a large problem, this thesis focus then specifically on detecting novel room categories.
+}
+\end{frame}
+
+\subsection{Outline}
+\begin{frame}{Outline}
+  \begin{multicols}{2}
+    \tableofcontents
+  \end{multicols}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Semantic Mapping}
+
+\subsection{Semantic Mapping}
+\begin{frame}{Mapping - Machine Friendly}
+\begin{columns}[c]
+\column{0.5\textwidth}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{figures/machinefriendly.jpg}
+  \end{figure}
+
+\column{0.5\textwidth}
+  \only<1>{
+  \begin{itemize}
+    \item World Representation
+    \item Bypass Sensory Horizon
+    \item Planning Tool
+  \end{itemize}
+  }
+
+  \only<2>{
+  \begin{figure}
+  \begin{tikzpicture}
+    % \draw [help lines] grid(3,2);
+
+    \node (human) at (0,0) {\reflectbox{\includegraphics[width=0.4\textwidth]{figures/extra/bart.pdf}}};
+    \node (robot) at (3,0) {\reflectbox{\includegraphics[width=0.2\textwidth]{figures/robot.png}}};
+
+    \node[ellipse callout, draw, callout absolute pointer={(human.70)}]
+          (humanvoice) at (1,3) {Where is Wally?};
+
+    \node[ellipse callout, draw, callout absolute pointer={(robot.north)},
+          text width=0.2\textwidth]
+          (robotvoice) at (3,2) {X=32, Y=17};
+
+  \end{tikzpicture}
+  \end{figure}
+  }
+\end{columns}
+\end{frame}
+
+\begin{frame}{Semantic Mapping}
+\begin{columns}[c]
+
+\column{0.5\textwidth}
+  \begin{figure}
+    \includegraphics[width=\textwidth]{figures/semanticfriendly.jpg}
+  \end{figure}
+
+\column{0.5\textwidth}
+  \only<1>{
+  \begin{figure}
+  \begin{tikzpicture}
+    \node (human) at (0,0) {\reflectbox{\includegraphics[width=0.4\textwidth]{figures/extra/bart.pdf}}};
+    \node (robot) at (3,0) {\reflectbox{\includegraphics[width=0.2\textwidth]{figures/robot.png}}};
+
+    \node[ellipse callout, draw, callout absolute pointer={(human.70)}]
+          (humanvoice) at (1,3) {Where is Wally?};
+
+    \node[ellipse callout, draw, callout absolute pointer={(robot.north)},
+          text width=0.2\textwidth]
+          (robotvoice) at (3,2) {In the kitchen};
+
+  \end{tikzpicture}
+  \end{figure}
+  }
+
+  \only<2>{
+  \begin{itemize}
+    \item High Level Reasoning
+    \item Easier Planning \cite{hanheide2011ijcai}
+    \item Closer to Human Semantic
+  \end{itemize}
+  }
+
+\end{columns}
+\end{frame}
+
 \subsection{Spatial Knowledge}
+\begin{frame}{Spatial Knowledge}
+  \begin{columns}[c]
+  \column{0.5\textwidth}
+    \begin{figure}
+    \includegraphics[width=0.8\textwidth]{figures/spatial-knowledge-representation.pdf}
+    \end{figure}
+
+  \column{0.5\textwidth}
+    \begin{block}{Layered Spatial Knowledge Representation~\cite{pronobis2010ias}}
+    \begin{itemize}
+      \item Sensory Layer
+      \item Place Layer
+      \item Categorical Layer
+      \item Conceptual Layer
+    \end{itemize}
+    \end{block}
+  \end{columns}
+\end{frame}
+
+\subsection{Process}
+\begin{frame}{A Semantic Mapping Process}
+  \begin{figure}
+    \includegraphics[width=0.8\textwidth]{figures/dora-architecture.pdf}
+  \end{figure}
+  \cite{pronobis2011semmap} presents a process that uses the spatial knowledge
+  to perform probabilistic and multi-modal semantic mapping of the environment
+  using graphical models.
+\end{frame}
+
 \subsection{Conceptual Map}
+\begin{frame}{Conceptual Map}
+  \begin{columns}[c]
+  \column{0.5\textwidth}
+    \begin{figure}
+      \includegraphics[width=\textwidth]{figures/chaingraph.pdf}
+    \end{figure}
+  \column{0.5\textwidth}
+    \begin{itemize}
+      \item Uncertain Ontology
+      \item Probabilistic Inference
+      \item Use of Graphical Models
+    \end{itemize}
+  \end{columns}
 
+  \note{
+    The semantic mapping process proposed by Pronobis uses the previously presented spatial knowledge
+    to produce a conceptual map that relates all the identified concepts and relates them ...
+
+    This tool allows to perform semantic mapping of the environment, but is not able to handle gaps
+    on the agent knowledge.
+  }
+\end{frame}
+
+\begin{frame}{Probabilistic Graphical Models}
+Provides us methods to relate and combine the several information,
+Can be approximated, handle uncertainties, etc\dots
+
+\begin{block}{Several Methods}
+Bayes Networks, Random Markov Fields, Chain Graphs, Factor Graphs
+\end{block}
+
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Novelty Detection}
+
+\subsection{Fast Review}
+\begin{frame}{Novelty Detection - Fast Review}
+  \begin{block}{Definition}
+    Detection of novel patterns or signals presents on the data that the system was not trained
+    to handle -- \cite{markou2003novelty}
+  \end{block}
+  \begin{block}{Single Class Classification}
+    Lack of negative samples makes them suitable for cases where it is not possible to
+    provide negative samples: anomaly detection, intrusion detection, etc\dots
+  \end{block}
+  \begin{block}{Several Methods}
+    K-PCA, One class SVM, Nearest Neighbours, etc\dots
+  \end{block}
+\end{frame}
+
 \subsection{Novelty Detection as Thresholding}
+\begin{frame}{Novelty Detection via Thresholding}
+
+\begin{block}{Threshold on Probability Density}
+For a fixed distribution and without any prior knowledge on the distribution,
+novelty can be detected by thresholding on the density of
+the training data as shown on \cite{bishop1994novelty}.
+\end{block}
+
+\begin{block}{Dynamic Sample-Space}
+As more information becomes available, the .
+
+The question is on how to define a threshold under this circumstances. 
+\end{block}
+\end{frame}
+
 \subsection{Conditional and Unconditional Ratio}
+\begin{frame}{Ordering the Sample Space}
+
+\begin{block}{Back to the Roots}
+An optimal detector can be implemented by defining a threshold over a
+relation order defined by $P(\overline{novel}|x)$.
+\end{block}
+\vfill
+
+\begin{block}{Conditional and Unconditional Probability Ratio}
+\[P(\overline{novel}|x) = \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)}\]
+\end{block}
+\note{
+Going back to a definition of a novelty detector, it can be shown that any optimal
+detector can be seen as a solving a continuous knapsack problem, and that a threshold
+can be set over the order relation defined by the probability of a sample being novel.
+}
+\end{frame}
+
+\subsection{Models}
+\begin{frame}{Conditional Probability}
+\begin{columns}[c]
+  \column{0.5\textwidth}
+    \input{conditional_model.tikz}
+  \column{0.5\textwidth}
+    \begin{block}{$P(x|\overline{novel})$}
+      The graphical model used by the conceptual map represents the distribution of the
+      variables given that the agent knowledge holds true.
+
+%      The conditional probability is then the probability of the sensed features being
+%      generated by a distribution modelled according to the agent knowledge.
+    \end{block}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Unconditional Probability Model}
+\only<1>{
+\begin{columns}[c]
+  \column{0.5\textwidth}
+  \input{uniform-distribution.tikz}
+  \column{0.5\textwidth}
+  \input{bigfactor-distribution.tikz}
+\end{columns}
+}
+\only<2>{
+\begin{columns}[c]
+ \column{0.5\textwidth}
+  \input{uniformfac-distribution.tikz}
+
+  \centering
+  Uniform Model
+ \column{0.5\textwidth}
+  \input{independent-distribution.tikz}
+
+  \centering
+  Independent Model
+\end{columns}
+}
+\end{frame}
+
+\subsection{Results}
+\begin{frame}{Results}
+  \begin{figure}
+    \includegraphics[width=0.7\textwidth]{results/synthetic-all.pdf}
+  \end{figure}
+\end{frame}
 
 \section{Conclusion}
+\begin{frame}{Conclusion}
+\end{frame}
+
 \subsection{Limitations}
+\begin{frame}{Limitations}
+\begin{itemize}
+\item Incorporate structural information (non const P(novel))
+\item Incorporate uncertain sensing
+\end{itemize}
+\end{frame}
+
 \subsection{Future Work}
+\begin{frame}{Future Work}
+\begin{itemize}
+\item Generalize the Framework
+\item Exploiting Generative Models
+\item Learning Graph Structures
+\item After Detection of Knowledge Gaps
+\end{itemize}
+\end{frame}
+
+\subsection{Extra Information}
+\begin{frame}{Extra}
+\begin{block}{Reproducible Research}
+All produced results reproducible from online repository:
+\url{https://github.com/andresusanopinto/novelty-detection-thesis}
+\end{block}
+
+\begin{block}{Article on EPIA 2011}
+Part of this thesis work will be presented in EPIA 2011 (October)
+\emph{Novelty Detection Using Graphical Models for
+Semantic Room Classification}
+\end{block}
+\end{frame}
 
 \appendix
-
 \begin{frame}[allowframebreaks]{Bibliography}

### 2011-07-14 thesis/presentation.tex
more updates

@@
 \documentclass[compress]{beamer}
+\usepackage[utf8x]{inputenc}
 \usepackage{default}
@@
 \useoutertheme[subsection=false]{miniframes}
+\setbeamertemplate{footline}[frame number]
+\beamertemplatenavigationsymbolsempty
 
@@
 
-\begin{block}{Focus on Detection of Novel Semantic Categories of Places}
-
+\begin{block}{Focus on Semantic Categories of Places}
+Humans attribute meaning to the areas that describe the expected properties,
+objects and actions of them. E.g.\ cups are found in kitchens. 
 \end{block}
@@
 
-\subsection{Semantic Mapping}
+\subsection{Mapping}
 \begin{frame}{Mapping - Machine Friendly}
@@
 
+\subsection{Semantic Mapping}
 \begin{frame}{Semantic Mapping}
@@
 
-\subsection{Novelty Detection as Thresholding}
+\subsection{Thresholding}
 \begin{frame}{Novelty Detection via Thresholding}
@@
 \[P(\overline{novel}|x) = \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)}\]
+\uncover<2>{\alert{Simplified with an assumption on constant $P(novel)$}}
 \end{block}
@@
 \subsection{Results}
+\begin{frame}{Synthetic Dataset}
+
+\end{frame}
+
 \begin{frame}{Results}
@@
 \section{Conclusion}
-\begin{frame}{Conclusion}
+
+\subsection{Summary}
+\begin{frame}{Summary}
+\begin{itemize}
+\item Studies a semantic mapping process
+\item Studies novelty detection
+\item Proposes a method to perform detection of novel categories
+      of a variable type of a graphical model when the graph structure
+      is dynamic.
+\end{itemize}
 \end{frame}
@@
 \begin{frame}{Limitations}
-\begin{itemize}
-\item Incorporate structural information (non const P(novel))
-\item Incorporate uncertain sensing
-\end{itemize}
+\begin{block}{Assumption on constant $P(novel)$}
+Graph structure plays a role on the likelihood a sample with a novel
+category is drawn. Assuming it to be constant is a very strong assumption
+that is not expected to hold in all scenarios.
+\end{block}
+%\begin{block}{Uncertain Sensing}
+%The presented methods require calculation of normalized probabilities turning it impossible
+%to handle implicit factors.
+%\end{block}
 \end{frame}
@@
 \begin{frame}{Future Work}
-\begin{itemize}
-\item Generalize the Framework
-\item Exploiting Generative Models
-\item Learning Graph Structures
-\item After Detection of Knowledge Gaps
-\end{itemize}
+  \begin{block}{Generalize the Framework}
+  Handle novelty on any and more than one variable of the graph and fusing
+  all novelty information in a single method in a probabilistic fashion (e.g.\ MAP operation).
+  \end{block}
+  \begin{block}{Exploit Generative Models}
+  Develop methods to be able to explain what the system considers to be known or novel.
+  Such that the reason causing novelty can be interpreted.
+  \end{block}
+\end{frame}
+\begin{frame}{Future Work}
+  \begin{block}{Learning Graph Structures}
+  Graph structure should be considered probabilistic. Methods to handle and probabilistic
+  represent knowledge on how the agent detect structures must be develop.
+  \end{block}
+  \begin{block}{After Detection of Knowledge Gaps}
+  What can be done with the detection of novel situations? Can the system be adapted
+  to learn and group novel situations? How to perform self-extension?
+  \end{block}
 \end{frame}
@@
 \begin{block}{Reproducible Research}
-All produced results reproducible from online repository:
+All reports and results reproducible from online repository:
 \url{https://github.com/andresusanopinto/novelty-detection-thesis}
@@
 \begin{block}{Article on EPIA 2011}
-Part of this thesis work will be presented in EPIA 2011 (October)
-\emph{Novelty Detection Using Graphical Models for
-Semantic Room Classification}
+\textit{Novelty Detection Using Graphical Models for Semantic Room Classification}
 \end{block}

### 2011-07-14 thesis/presentation.tex
More changes in thesis presentation

@@
 \begin{frame}{Mapping - Machine Friendly}
-\begin{columns}[c]
+\begin{columns}[tc]
 \column{0.5\textwidth}
   \begin{figure}
-    \includegraphics[width=\textwidth]{figures/machinefriendly.jpg}
+    \only<1>{\includegraphics[width=\textwidth]{figures/extra/pointcloud.jpg}}
+    \only<2>{\includegraphics[width=\textwidth]{figures/machinefriendly.pdf}}
   \end{figure}
@@
   \only<2>{
-  \begin{figure}
   \begin{tikzpicture}
@@
   \end{tikzpicture}
-  \end{figure}
   }
@@
 \begin{frame}{Semantic Mapping}
-\begin{columns}[c]
+\begin{columns}[tt]
 
@@
   \begin{figure}
-    \includegraphics[width=\textwidth]{figures/semanticfriendly.jpg}
+    \includegraphics[width=\textwidth]{figures/semanticfriendly.pdf}
   \end{figure}
@@
           text width=0.2\textwidth]
-          (robotvoice) at (3,2) {In the kitchen};
+          (robotvoice) at (3,2) {In the office};
 
@@
 
-\begin{block}{Dynamic Sample-Space}
-As more information becomes available, the .
-
-The question is on how to define a threshold under this circumstances. 
+\begin{block}{Dynamic Sample Space}
+As more information becomes available the probability mass spreads around and
+the threshold for novelty detection needs to be adjusted.
+\alert<2->{How to do it?}
 \end{block}
@@
 \begin{frame}{Ordering the Sample Space}
-
 \begin{block}{Back to the Roots}
@@
 \only<1>{
-\begin{columns}[c]
+\begin{columns}[t]
   \column{0.5\textwidth}
   \input{uniform-distribution.tikz}
+
+  \centering
+  No prior knowledge on $P(x)$. \\
+  Uniform distribution~\cite{shore1980axiomatic}
+
 \column{0.5\textwidth}
   \input{bigfactor-distribution.tikz}
+
+  \centering
+  Assume structure and knowledge on other variables still hold true on $P(x)$
 \end{columns}
@@
 
+\begin{columns}[c]
+\column{0.5\textwidth}
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room) [latent] {$r$}; \& \& \\
+    \& \& \& \& \\
+    \node (f1) [factor] {}; \&
+    \node (f2) [factor] {}; \&
+    \node (f3) [factor] {}; \&
+    \node (fi) [] {\dots}; \&
+    \node (fn) [factor] {}; \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (room) -- (f1) -- (x1);
+  \draw [-] (room) -- (f2) -- (x2);
+  \draw [-] (room) -- (f3) -- (x3);
+  \draw [-] (room) -- (fi);
+  \draw [-] (room) -- (fn) -- (xn);
+
+  \node (captf1) [right=2pt of f1] {\footnotesize{$\phi_{X_1}$}};
+  \node (captf2) [right=2pt of f2] {\footnotesize{$\phi_{X_2}$}};
+  \node (captf3) [right=2pt of f3] {\footnotesize{$\phi_{X_3}$}};
+  \node (captfn) [right=2pt of fn] {\footnotesize{$\phi_{X_n}$}};
+\end{tikzpicture}
+\column{0.5\textwidth}
+
+\begin{block}{Room class generator of features}
+\begin{itemize}
+  \item 11 - Room Categories
+  \item 2 - Room Shapes (laser)
+  \item 3 - Room Size (laser)
+  \item 10 - Room Appearance (CRFH)
+  \item 7 - Object Detectors (SIFT)
+\end{itemize}
+\end{block}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Synthetic Dataset - Conditional Model}
+\begin{block}{Conditional Model}
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \node (room) [latent] {$r$}; \& \& \\
+    \& \& \& \& \\
+    \node (f1) [factor] {}; \&
+    \node (f2) [factor] {}; \&
+    \node (f3) [factor] {}; \&
+    \node (fi) [] {\dots}; \&
+    \node (fn) [factor] {}; \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (room) -- (f1) -- (x1);
+  \draw [-] (room) -- (f2) -- (x2);
+  \draw [-] (room) -- (f3) -- (x3);
+  \draw [-] (room) -- (fi);
+  \draw [-] (room) -- (fn) -- (xn);
+
+  \node (captf1) [right=2pt of f1] {\footnotesize{$\phi_{X_1}$}};
+  \node (captf2) [right=2pt of f2] {\footnotesize{$\phi_{X_2}$}};
+  \node (captf3) [right=2pt of f3] {\footnotesize{$\phi_{X_3}$}};
+  \node (captfn) [right=2pt of fn] {\footnotesize{$\phi_{X_n}$}};
+\end{tikzpicture}
+
+Trained from data sampled from 5 of the 11 room categories.
+\end{block}
+\end{frame}
+
+\begin{frame}{Synthetic Dataset - Unconditional Models}
+
+\begin{block}{Uniform Model}
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+\end{tikzpicture}
+\end{block}
+
+\begin{block}{Independent Model}
+\centering
+\begin{tikzpicture}
+  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
+    \& \& \& \& \\
+    \node (f1) [factor] {}; \&
+    \node (f2) [factor] {}; \&
+    \node (f3) [factor] {}; \&
+    \node (fi) [] {\dots}; \&
+    \node (fn) [factor] {}; \\
+    \node (x1) [obs] {$x_1$}; \&
+    \node (x2) [obs] {$x_2$}; \&
+    \node (x3) [obs] {$x_3$}; \&
+    \node (xi) [] {\dots}; \&
+    \node (xn) [obs] {$x_n$}; \\
+  };
+  \draw [-] (f1) -- (x1);
+  \draw [-] (f2) -- (x2);
+  \draw [-] (f3) -- (x3);
+  \draw [-] (fi);
+  \draw [-] (fn) -- (xn);
+\end{tikzpicture}
+
+Factors were trained using samples drawn from the synthetic distribution where the label
+for the generating class was hidden.
+\end{block}
 \end{frame}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
 \item Proposes a method to perform detection of novel categories
-      of a variable type of a graphical model when the graph structure
-      is dynamic.
+      of a variable when the graph structure is dynamic.
 \end{itemize}

### 2011-07-14 thesis/presentation.tex
More updates to presentation.
Probably not final state yet, but I am running out of time.

@@
   \column{0.5\textwidth}
-    \begin{itemize}
-      \item Uncertain Ontology
-      \item Probabilistic Inference
-      \item Use of Graphical Models
-    \end{itemize}
+    \begin{block}{Conceptual Map}
+      The ontology is used to identify instances of the concepts and categories
+      And perform probabilistic semantic mapping using graphical models.
+    \end{block}
   \end{columns}
@@
 \begin{frame}{Probabilistic Graphical Models}
-Provides us methods to relate and combine the several information,
-Can be approximated, handle uncertainties, etc\dots
+Provides us methods to relate and combine the several variables in a probabilistic way.
+Allowing to deal with uncertain information and infer variable states.
+
+\vfill
 
@@
 
+\begin{frame}{Results - Variable Graph Size}
+
+\begin{columns}[c]
+  \column{0.5\textwidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{results/synthetic-3features.pdf} \\ 3 sensed features
+  \column{0.5\textwidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{results/synthetic-5features.pdf} \\ 5 sensed features
+\end{columns}
+\begin{columns}[c]
+  \column{0.5\textwidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{results/synthetic-10features.pdf} \\ 10 sensed features
+  \column{0.5\textwidth}
+  \centering
+  \includegraphics[width=0.8\textwidth]{results/synthetic-50features.pdf} \\ 50 sensed features
+\end{columns}
+\end{frame}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
       of a variable when the graph structure is dynamic.
+\item Shows access to unlabelled data can greatly increase performance.
 \end{itemize}

### 2011-07-15 thesis/presentation.tex

@@
 \note{
-1 - A solution could pass by tagging our world with extra information, such as maps, rfids, but such a solution
+1 - A solution could pass by tagging our world with extra information, such as maps, RFIDs, but such a solution
 is not practical and scalable.
@@
 \begin{block}{Machine Learning}
-Give us methods to solve this problem in reverse by providing robots with knowledge
+Gives us methods to solve this problem in reverse by providing robots with knowledge
 on how to detect and classify concepts and categories from the underlying sensed world.
@@
       The ontology is used to identify instances of the concepts and categories
-      And perform probabilistic semantic mapping using graphical models.
+      and perform probabilistic semantic mapping using graphical models.
     \end{block}

### 2011-10-07 semantic-nd/presentation.tex
Working on presentation for EPIA-2011.

--- 2011-07-14 thesis/presentation.tex
@@
 
-\title{Novelty Detection for Semantic Place Categorization}
-\author{André Susano Pinto\inst{1,2}}
+\title{Novelty Detection Using Graphical Models for
+       Semantic Room Classification}
+\author{André Susano Pinto\inst{1,2} \and Andrzej Pronobis\inst{1} \and Luis Paulo Reis\inst{2}}
 \date{18 July 2011}
@@
 \institute {
- Supervised by: Andrzej Pronobis\inst{1}, Luis Paulo Reis\inst{2}
- \and
  \inst{1}The Royal Institute of Technology (KTH), Sweden \\
@@
 
+\section{Summary}
+\begin{frame}{Summary - Plan}
+    \begin{block}{Semantic mapping of indoor enviroments}
+        \begin{itemize}
+            \item What is it? What for?
+            \item How to, using graphical models
+            \item Limited on the agent knowledge of variable categories
+        \end{itemize}
+    \end{block}
+    \begin{block}{Novelty Detection}
+        \begin{itemize}
+            \item Trigger on $P(x|known)$
+            \item Dynamic graphs
+            \item $P(x|known)/P(x)$
+            \item Approximating $P(x)$
+        \end{itemize}
+    \end{block}
+\end{frame}
+
+\begin{frame}{Goals}
+  Brief introduction
+  \begin{itemize}
+    \item Study a semantic mapping process in the context of mobile robots.
+    \item Propose a method based on the studied system to detect novel semantic categories of places.
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{Outline}
@@
 
-\section{Introduction}
-\subsection{Motivation}
-\begin{frame}{Motivation}
-
-\begin{columns}[t]
-  \column{0.3\textwidth}
-  \includegraphics[width=\textwidth]{figures/extra/PR2.jpg}
-
-  \centering
-  PR2
-
-  \column{0.3\textwidth}
-  \includegraphics[width=\textwidth]{figures/extra/servingbot.jpg}
-
-  \centering
-  Serving Robot
-
-  \column{0.3\textwidth}
-  \includegraphics[width=\textwidth]{figures/extra/wakamaru.jpg}
-
-  \centering
-  Wakamaru
-\end{columns}
-\vfill
-
-\begin{itemize}
-  \item Robots are useful: can make our lives easier
-  \item Robots are entering our Houses and Offices
-  \item They need to understand our world and communicate with us.
-\end{itemize}
-\note{
-1 - Humans dream on creating robots. 
-Both by the challenge on creating intelligent systems on itself but also to make our life easier.
-% I should have an example list on how they could be useful
-
-2 - It is becoming a reality and robots will enter our houses and offices.
-The question is no longer if we will have a robot at home/office, but how many.
-
-3 - They will need to understand and communicate with us and our human worlds.
-The need of semantic information!
-}
-\end{frame}
-
-\subsection{Problem}
-\begin{frame}{Problem}
-
-  Enhance world with semantic information is \alert{not a scalable solution}.
-
-  \begin{block}{}
-  \begin{columns}
-  \column{0.3\textwidth}
-    \begin{figure}
-      \includegraphics[width=\textwidth]{figures/extra/RFID.pdf} \\
-      RFID for object tagging
-    \end{figure}
-  \column{0.3\textwidth}
-    \begin{figure}
-      \includegraphics[width=\textwidth]{figures/extra/floorplan.jpg} \\
-      Maps extended with semantic information
-    \end{figure}
-  \end{columns}
-  \end{block}
-
-\note{
-1 - A solution could pass by tagging our world with extra information, such as maps, rfids, but such a solution
-is not practical and scalable.
-}
-\end{frame}
-
-\begin{frame}{Problem}
-\begin{block}{Machine Learning}
-Give us methods to solve this problem in reverse by providing robots with knowledge
-on how to detect and classify concepts and categories from the underlying sensed world.
-\end{block}
-
-\begin{block}{Unreliable on Dynamic and Unknown Environments}
-It is unrealistic to believe on the possibility to fully describe the required knowledge
-for the agent in all situations.
-\end{block}
-
-\note{
-Machine Learning tries to solve the reverse semantic mapping problem.
-
-It endows robots with knowledge that allows them to map what they sense to human concepts
-and categories.
-
-Leading to the need of methods that are able to inspect and smartly identifying information directly
-from low-level sensors.
-
-For that several machine learning techniques have been developed, but all of those require and represent
-knowledge that is given a-priori.
-}
-\end{frame}
-
-\begin{frame}{Problem}
-
-\begin{block}{Detection of Knowledge-Gaps}
-Detection of situations where the agent knowledge does not suffices.
-\end{block}
-
-\begin{block}{Focus on Semantic Categories of Places}
-Humans attribute meaning to the areas that describe the expected properties,
-objects and actions of them. E.g.\ cups are found in kitchens. 
-\end{block}
-\note{
-- It is unrealistic to believe on the possibility to describe all the required knowledge
-
-- When dealing with novel and unknown scenarios an agent has to be able to detect something
-is not conforming to what it has learn.
-
-In order to be reliable agents need to detect situations where their knowledge
-And agents be reliable even when situations
-- Scalability
-- Dynamic Aspects
-- Long Term
-- Life Long
-
-It is unrealistic to believe on the possibility to describe knowledge that holds to all the possible
-scenarios, and when dealing with long-term and life-long agents
-
-- Variability, dynamic aspects of environment, 
-- Should be able to deal with novel cases
-- 1st step is to detect them
-}
-\end{frame}
-
-\subsection{Goals}
-\begin{frame}{Goals}
-  \begin{itemize}
-    \item Study a semantic mapping process in the context of mobile robots \cite{pronobis2011phd}.
-    \item Propose a method based on the studied system to detect novel semantic categories of places.
-  \end{itemize}
-
-\note{
-This thesis goes on the direction of detecting knowledge gaps on mobile robots.
-
-It studies a process of semantic mapping introduced by Pronobis
-and studies how to detect those situations where the existent knowledge does not describes
-the reality.
-But being such a large problem, this thesis focus then specifically on detecting novel room categories.
-}
-\end{frame}
-
-\subsection{Outline}
-\begin{frame}{Outline}
-  \begin{multicols}{2}
-    \tableofcontents
-  \end{multicols}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Semantic Mapping}
-
-\subsection{Mapping}
-\begin{frame}{Mapping - Machine Friendly}
-\begin{columns}[tc]
-\column{0.5\textwidth}
-  \begin{figure}
-    \only<1>{\includegraphics[width=\textwidth]{figures/extra/pointcloud.jpg}}
-    \only<2>{\includegraphics[width=\textwidth]{figures/machinefriendly.pdf}}
-  \end{figure}
-
-\column{0.5\textwidth}
-  \only<1>{
-  \begin{itemize}
-    \item World Representation
-    \item Bypass Sensory Horizon
-    \item Planning Tool
-  \end{itemize}
-  }
-
-  \only<2>{
-  \begin{tikzpicture}
-    % \draw [help lines] grid(3,2);
-
-    \node (human) at (0,0) {\reflectbox{\includegraphics[width=0.4\textwidth]{figures/extra/bart.pdf}}};
-    \node (robot) at (3,0) {\reflectbox{\includegraphics[width=0.2\textwidth]{figures/robot.png}}};
-
-    \node[ellipse callout, draw, callout absolute pointer={(human.70)}]
-          (humanvoice) at (1,3) {Where is Wally?};
-
-    \node[ellipse callout, draw, callout absolute pointer={(robot.north)},
-          text width=0.2\textwidth]
-          (robotvoice) at (3,2) {X=32, Y=17};
-
-  \end{tikzpicture}
-  }
-\end{columns}
-\end{frame}
-
-\subsection{Semantic Mapping}
+\subsection{Introduction}
 \begin{frame}{Semantic Mapping}
     \begin{columns}[tt]
-
         \column{0.5\textwidth}
@@
 \only<1>{
+        \begin{block}{Use of Human Semantics}
+            Expanding space representation with semantics. 
+        \end{block}
+        \begin{block}{Focus on Room Categories}
+            Semantic categories help describing expected properties and functionality.
+        \end{block}
+}
+\only<2>{
         \begin{figure}
@@
 }
-
-  \only<2>{
-  \begin{itemize}
-    \item High Level Reasoning
-    \item Easier Planning \cite{hanheide2011ijcai}
-    \item Closer to Human Semantic
-  \end{itemize}
-  }
-
     \end{columns}
@@
 
-\subsection{Spatial Knowledge}
-\begin{frame}{Spatial Knowledge}
-  \begin{columns}[c]
-  \column{0.5\textwidth}
-    \begin{figure}
-    \includegraphics[width=0.8\textwidth]{figures/spatial-knowledge-representation.pdf}
-    \end{figure}
-
-  \column{0.5\textwidth}
-    \begin{block}{Layered Spatial Knowledge Representation~\cite{pronobis2010ias}}
-    \begin{itemize}
-      \item Sensory Layer
-      \item Place Layer
-      \item Categorical Layer
-      \item Conceptual Layer
-    \end{itemize}
+\subsection{Dora Architecture Overview}
+\begin{frame}{Dora Architecture Overview}
+    \begin{block}{Multi-modal approach}
+        Combining semantic data extracted from diferent sources is expected to yield better performance.
     \end{block}
-  \end{columns}
+    \begin{block}{Probabilistic approach}
+        TODO: Need to defend use of probabilistic approach and graphical models.
+    \end{block}
 \end{frame}
 
-\subsection{Process}
-\begin{frame}{A Semantic Mapping Process}
+\begin{frame}{Dora Architecture Overview}
   \begin{figure}
-    \includegraphics[width=0.8\textwidth]{figures/dora-architecture.pdf}
+    \includegraphics[width=0.8\textwidth]{figures/processes.pdf}
   \end{figure}
-  \cite{pronobis2011semmap} presents a process that uses the spatial knowledge
-  to perform probabilistic and multi-modal semantic mapping of the environment
-  using graphical models.
 \end{frame}
 
-\subsection{Conceptual Map}
-\begin{frame}{Conceptual Map}
+\subsection{Spatial Knowledge and Conceptual Map}
+\begin{frame}{Spatial Knowledge and Conceptual Map}
   \begin{columns}[c]
@@
   \column{0.5\textwidth}
-    \begin{itemize}
-      \item Uncertain Ontology
-      \item Probabilistic Inference
-      \item Use of Graphical Models
-    \end{itemize}
-  \end{columns}
-
-  \note{
-    The semantic mapping process proposed by Pronobis uses the previously presented spatial knowledge
-    to produce a conceptual map that relates all the identified concepts and relates them ...
-
-    This tool allows to perform semantic mapping of the environment, but is not able to handle gaps
-    on the agent knowledge.
-  }
-\end{frame}
-
-\begin{frame}{Probabilistic Graphical Models}
-Provides us methods to relate and combine the several information,
-Can be approximated, handle uncertainties, etc\dots
-
-\begin{block}{Several Methods}
-Bayes Networks, Random Markov Fields, Chain Graphs, Factor Graphs
-\end{block}
-
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Novelty Detection}
-
-\subsection{Fast Review}
-\begin{frame}{Novelty Detection - Fast Review}
-  \begin{block}{Definition}
-    Detection of novel patterns or signals presents on the data that the system was not trained
-    to handle -- \cite{markou2003novelty}
-  \end{block}
-  \begin{block}{Single Class Classification}
-    Lack of negative samples makes them suitable for cases where it is not possible to
-    provide negative samples: anomaly detection, intrusion detection, etc\dots
-  \end{block}
-  \begin{block}{Several Methods}
-    K-PCA, One class SVM, Nearest Neighbours, etc\dots
-  \end{block}
-\end{frame}
-
-\subsection{Thresholding}
-\begin{frame}{Novelty Detection via Thresholding}
-
-\begin{block}{Threshold on Probability Density}
-For a fixed distribution and without any prior knowledge on the distribution,
-novelty can be detected by thresholding on the density of
-the training data as shown on \cite{bishop1994novelty}.
-\end{block}
-
-\begin{block}{Dynamic Sample Space}
-As more information becomes available the probability mass spreads around and
-the threshold for novelty detection needs to be adjusted.
-\alert<2->{How to do it?}
-\end{block}
-\end{frame}
-
-\subsection{Conditional and Unconditional Ratio}
-\begin{frame}{Ordering the Sample Space}
-\begin{block}{Back to the Roots}
-An optimal detector can be implemented by defining a threshold over a
-relation order defined by $P(\overline{novel}|x)$.
-\end{block}
-\vfill
-
-\begin{block}{Conditional and Unconditional Probability Ratio}
-\[P(\overline{novel}|x) = \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)}\]
-\uncover<2>{\alert{Simplified with an assumption on constant $P(novel)$}}
-\end{block}
-\note{
-Going back to a definition of a novelty detector, it can be shown that any optimal
-detector can be seen as a solving a continuous knapsack problem, and that a threshold
-can be set over the order relation defined by the probability of a sample being novel.
-}
-\end{frame}
-
-\subsection{Models}
-\begin{frame}{Conditional Probability}
-\begin{columns}[c]
-  \column{0.5\textwidth}
-    \input{conditional_model.tikz}
-  \column{0.5\textwidth}
-    \begin{block}{$P(x|\overline{novel})$}
-      The graphical model used by the conceptual map represents the distribution of the
-      variables given that the agent knowledge holds true.
-
-%      The conditional probability is then the probability of the sensed features being
-%      generated by a distribution modelled according to the agent knowledge.
+    \begin{block}{Conceptual Map}
+      The ontology is used to identify instances of the concepts and categories
+      and perform probabilistic semantic mapping using graphical models.
     \end{block}
@@
 
-\begin{frame}{Unconditional Probability Model}
-\only<1>{
-\begin{columns}[t]
+\subsection{Graphical Models}
+\begin{frame}{Graphical models}
+    Multi-modal probabilistic semantic mapping using graphical models.
+    Uncertain semantic information
+    Conceptual map
+    Probabilistic inference on variables
+
+            $\text{known} = \{\text{corridor}, \text{office}, \cdots\}$
+            $R_i \in \text{known}$
+
+    TODO: picture explaining that maybe roomA is not known and showing X (observed features) and introduced novel and not novel, and showing graph structure.
+
+\end{frame}
+
+\section{Novelty Detection}
+\subsection{Introduction}
+\begin{frame}{Novelty Detection}
+    \begin{columns}[c]
         \column{0.5\textwidth}
-  \input{uniform-distribution.tikz}
-
         \centering
-  No prior knowledge on $P(x)$. \\
-  Uniform distribution~\cite{shore1980axiomatic}
+        \includegraphics[width=\textwidth]{figures/conditional-prob-graph.pdf}
 
         \column{0.5\textwidth}
-  \input{bigfactor-distribution.tikz}
-
-  \centering
-  Assume structure and knowledge on other variables still hold true on $P(x)$
-\end{columns}
+        \only<1>{
+        \begin{itemize}
+            \item Deploy on new and unknown environments
+            \item Active-learning and self-extension
+            \item Detect Knowledge-Gaps
+        \end{itemize}
         }
         \only<2>{
-\begin{columns}[c]
- \column{0.5\textwidth}
-  \input{uniformfac-distribution.tikz}
+        \begin{block}
+            {Novel room category?}
+            \begin{itemize}
+                \item Detect \alert{$R_1\notin \text{known}$}
+            \end{itemize}
+        \end{block}
+        }
+    \end{columns}
+\end{frame}
 
-  \centering
-  Uniform Model
- \column{0.5\textwidth}
-  \input{independent-distribution.tikz}
+\subsection{Novelty Detection via Thresholding}
+\begin{frame}
+    $\frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)} \propto P(x|\overline{novel})$
+    \begin{block}{$P(novel)$ constant}
+        Warning: assuming any graph structure has the same probability of generating
+        novel classes.
+    \end{block}
+    \begin{block}
+        Assumption on uniform $P(x)$.
+    \end{block}
+\end{frame}
 
-  \centering
-  Independent Model
-\end{columns}
-}
+\subsection{Dynamic distributions}
+\begin{frame}
+    The set of observed features and graph structure is always changing.
+    Threshold on $P(x|known)$ \cite{bishop1994novelty}
+    Dynamic graphs
+\end{frame}
+
+\subsection{Semi-supervised}
+\begin{frame}
+
 \end{frame}
@@
 \subsection{Results}
-\begin{frame}{Synthetic Dataset}
-
-\begin{columns}[c]
-\column{0.5\textwidth}
-\begin{tikzpicture}
-  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
-    \& \& \node (room) [latent] {$r$}; \& \& \\
-    \& \& \& \& \\
-    \node (f1) [factor] {}; \&
-    \node (f2) [factor] {}; \&
-    \node (f3) [factor] {}; \&
-    \node (fi) [] {\dots}; \&
-    \node (fn) [factor] {}; \\
-    \node (x1) [obs] {$x_1$}; \&
-    \node (x2) [obs] {$x_2$}; \&
-    \node (x3) [obs] {$x_3$}; \&
-    \node (xi) [] {\dots}; \&
-    \node (xn) [obs] {$x_n$}; \\
-  };
-  \draw [-] (room) -- (f1) -- (x1);
-  \draw [-] (room) -- (f2) -- (x2);
-  \draw [-] (room) -- (f3) -- (x3);
-  \draw [-] (room) -- (fi);
-  \draw [-] (room) -- (fn) -- (xn);
-
-  \node (captf1) [right=2pt of f1] {\footnotesize{$\phi_{X_1}$}};
-  \node (captf2) [right=2pt of f2] {\footnotesize{$\phi_{X_2}$}};
-  \node (captf3) [right=2pt of f3] {\footnotesize{$\phi_{X_3}$}};
-  \node (captfn) [right=2pt of fn] {\footnotesize{$\phi_{X_n}$}};
-\end{tikzpicture}
-\column{0.5\textwidth}
-
-\begin{block}{Room class generator of features}
-\begin{itemize}
-  \item 11 - Room Categories
-  \item 2 - Room Shapes (laser)
-  \item 3 - Room Size (laser)
-  \item 10 - Room Appearance (CRFH)
-  \item 7 - Object Detectors (SIFT)
-\end{itemize}
-\end{block}
-\end{columns}
+\begin{frame}
+    Synthetic dataset where 1 room category is responsible for generating a variable set of observed features.
 \end{frame}
 
-\begin{frame}{Synthetic Dataset - Conditional Model}
-\begin{block}{Conditional Model}
-\centering
-\begin{tikzpicture}
-  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
-    \& \& \node (room) [latent] {$r$}; \& \& \\
-    \& \& \& \& \\
-    \node (f1) [factor] {}; \&
-    \node (f2) [factor] {}; \&
-    \node (f3) [factor] {}; \&
-    \node (fi) [] {\dots}; \&
-    \node (fn) [factor] {}; \\
-    \node (x1) [obs] {$x_1$}; \&
-    \node (x2) [obs] {$x_2$}; \&
-    \node (x3) [obs] {$x_3$}; \&
-    \node (xi) [] {\dots}; \&
-    \node (xn) [obs] {$x_n$}; \\
-  };
-  \draw [-] (room) -- (f1) -- (x1);
-  \draw [-] (room) -- (f2) -- (x2);
-  \draw [-] (room) -- (f3) -- (x3);
-  \draw [-] (room) -- (fi);
-  \draw [-] (room) -- (fn) -- (xn);
-
-  \node (captf1) [right=2pt of f1] {\footnotesize{$\phi_{X_1}$}};
-  \node (captf2) [right=2pt of f2] {\footnotesize{$\phi_{X_2}$}};
-  \node (captf3) [right=2pt of f3] {\footnotesize{$\phi_{X_3}$}};
-  \node (captfn) [right=2pt of fn] {\footnotesize{$\phi_{X_n}$}};
-\end{tikzpicture}
-
-Trained from data sampled from 5 of the 11 room categories.
+\section{Conclusion}
+\subsection{Conclusion}
+\begin{frame}
+    \begin{block}
+        {Semantic Mapping Process}
+        A multi-modal semantic mapping process based on \emph{probabilistic graphical models}.
     \end{block}
-\end{frame}
-
-\begin{frame}{Synthetic Dataset - Unconditional Models}
-
-\begin{block}{Uniform Model}
-\centering
-\begin{tikzpicture}
-  \node [matrix,matrix anchor=mid, column sep=10pt, row sep=10pt,ampersand replacement=\&] {
-    \node (x1) [obs] {$x_1$}; \&
-    \node (x2) [obs] {$x_2$}; \&
-    \node (x3) [obs] {$x_3$}; \&
-    \node (xi) [] {\dots}; \&
-    \node (xn) [obs] {$x_n$}; \\
-  };
-\end{tikzpicture}
+    \begin{block}
+        {Novelty detection via thresholding}
+        Presented a method to detect novel classes on a variable of \emph{probabilistic graphical models}
+        dinamicaly instatiated from sensed semantic data.
     \end{block}
-
-\begin{block}{Independent Model}
-\centering
-\begin{tikzpicture}
-  \node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt,ampersand replacement=\&] {
-    \& \& \& \& \\
-    \node (f1) [factor] {}; \&
-    \node (f2) [factor] {}; \&
-    \node (f3) [factor] {}; \&
-    \node (fi) [] {\dots}; \&
-    \node (fn) [factor] {}; \\
-    \node (x1) [obs] {$x_1$}; \&
-    \node (x2) [obs] {$x_2$}; \&
-    \node (x3) [obs] {$x_3$}; \&
-    \node (xi) [] {\dots}; \&
-    \node (xn) [obs] {$x_n$}; \\
-  };
-  \draw [-] (f1) -- (x1);
-  \draw [-] (f2) -- (x2);
-  \draw [-] (f3) -- (x3);
-  \draw [-] (fi);
-  \draw [-] (fn) -- (xn);
-\end{tikzpicture}
-
-Factors were trained using samples drawn from the synthetic distribution where the label
-for the generating class was hidden.
-\end{block}
-\end{frame}
-
-\begin{frame}{Results}
-  \begin{figure}
-    \includegraphics[width=0.7\textwidth]{results/synthetic-all.pdf}
-  \end{figure}
-\end{frame}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-
-\subsection{Summary}
-\begin{frame}{Summary}
-\begin{itemize}
-\item Studies a semantic mapping process
-\item Studies novelty detection
-\item Proposes a method to perform detection of novel categories
-      of a variable when the graph structure is dynamic.
-\end{itemize}
-\end{frame}
-
-\subsection{Limitations}
-\begin{frame}{Limitations}
-\begin{block}{Assumption on constant $P(novel)$}
-Graph structure plays a role on the likelihood a sample with a novel
-category is drawn. Assuming it to be constant is a very strong assumption
-that is not expected to hold in all scenarios.
-\end{block}
-%\begin{block}{Uncertain Sensing}
-%The presented methods require calculation of normalized probabilities turning it impossible
-%to handle implicit factors.
-%\end{block}
 \end{frame}
@@
 \begin{frame}{Future Work}
-  \begin{block}{Generalize the Framework}
-  Handle novelty on any and more than one variable of the graph and fusing
-  all novelty information in a single method in a probabilistic fashion (e.g.\ MAP operation).
-  \end{block}
-  \begin{block}{Exploit Generative Models}
-  Develop methods to be able to explain what the system considers to be known or novel.
-  Such that the reason causing novelty can be interpreted.
-  \end{block}
-\end{frame}
-\begin{frame}{Future Work}
-  \begin{block}{Learning Graph Structures}
-  Graph structure should be considered probabilistic. Methods to handle and probabilistic
-  represent knowledge on how the agent detect structures must be develop.
-  \end{block}
-  \begin{block}{After Detection of Knowledge Gaps}
-  What can be done with the detection of novel situations? Can the system be adapted
-  to learn and group novel situations? How to perform self-extension?
-  \end{block}
-\end{frame}
-
-\subsection{Extra Information}
-\begin{frame}{Extra}
-\begin{block}{Reproducible Research}
-All reports and results reproducible from online repository:
-\url{https://github.com/andresusanopinto/novelty-detection-thesis}
-\end{block}
-
-\begin{block}{Article on EPIA 2011}
-\textit{Novelty Detection Using Graphical Models for Semantic Room Classification}
+    \begin{block}
+        {Incorporate novelty signals}
     \end{block}
@@
 \appendix
-\begin{frame}[allowframebreaks]{Bibliography}
-\bibliographystyle{alpha}
-\bibliography{refs}
-\end{frame}
-
+% \begin{frame}[allowframebreaks]{Bibliography}
+% \bibliographystyle{alpha}
+% \bibliography{refs}
+% \end{frame}
+% 
 \begin{frame}

### 2011-10-10 semantic-nd/presentation.tex
Presentation for EPIA-2011 (finished maybe).

@@
 \usepackage{multicol}
+\usepackage{subfig}
 
@@
 \section{Summary}
-\begin{frame}{Summary - Plan}
-    \begin{block}{Semantic mapping of indoor enviroments}
-        \begin{itemize}
-            \item What is it? What for?
-            \item How to, using graphical models
-            \item Limited on the agent knowledge of variable categories
-        \end{itemize}
-    \end{block}
-    \begin{block}{Novelty Detection}
-        \begin{itemize}
-            \item Trigger on $P(x|known)$
-            \item Dynamic graphs
-            \item $P(x|known)/P(x)$
-            \item Approximating $P(x)$
-        \end{itemize}
-    \end{block}
-\end{frame}
+%\begin{frame}{Summary - Plan}
+%    \begin{block}{Semantic mapping of indoor enviroments}
+%        \begin{itemize}
+%            \item What is it? What for?
+%            \item How to, using graphical models
+%            \item Limited on the agent knowledge of variable categories
+%        \end{itemize}
+%    \end{block}
+%    \begin{block}{Novelty Detection}
+%        \begin{itemize}
+%            \item Trigger on $P(x|known)$
+%            \item Dynamic graphs
+%            \item $P(x|known)/P(x)$
+%            \item Approximating $P(x)$
+%        \end{itemize}
+%    \end{block}
+%\end{frame}
 
-\begin{frame}{Goals}
+\subsection{Context}
+\begin{frame}{Context}
   Brief introduction
@@
 
+\subsection{Outline}
 \begin{frame}{Outline}
@@
     \begin{block}{Probabilistic approach}
-        TODO: Need to defend use of probabilistic approach and graphical models.
+        Uses probabilistic approach in classification in order to handle uncertainty.
     \end{block}
@@
 \begin{frame}{Spatial Knowledge and Conceptual Map}
-  \begin{columns}[c]
-  \column{0.5\textwidth}
+    Using the spatial knowledge the agent instatiates a conceptual map, here represented as a chain-graph:
     \begin{figure}
-      \includegraphics[width=\textwidth]{figures/chaingraph.pdf}
+      \includegraphics[width=0.5\textwidth]{figures/chaingraph.pdf}
     \end{figure}
-  \column{0.5\textwidth}
-    \begin{block}{Conceptual Map}
-      The ontology is used to identify instances of the concepts and categories
-      and perform probabilistic semantic mapping using graphical models.
-    \end{block}
-  \end{columns}
-\end{frame}
-
-\subsection{Graphical Models}
-\begin{frame}{Graphical models}
-    Multi-modal probabilistic semantic mapping using graphical models.
-    Uncertain semantic information
-    Conceptual map
-    Probabilistic inference on variables
-
-            $\text{known} = \{\text{corridor}, \text{office}, \cdots\}$
-            $R_i \in \text{known}$
-
-    TODO: picture explaining that maybe roomA is not known and showing X (observed features) and introduced novel and not novel, and showing graph structure.
-
 \end{frame}
@@
 \subsection{Novelty Detection via Thresholding}
-\begin{frame}
-    $\frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)} \propto P(x|\overline{novel})$
-    \begin{block}{$P(novel)$ constant}
-        Warning: assuming any graph structure has the same probability of generating
-        novel classes.
+\begin{frame}{Thresholding}
+    \begin{block}
+        {Optimal detector}
+        \begin{itemize}
+            \item Threshold on $P(R_i \in \text{known}|x) = P(\overline{novel}|x)$
+            \item Optimizes true-positives given a fixed false-positive ratio.
+        \end{itemize}
     \end{block}
     \begin{block}
-        Assumption on uniform $P(x)$.
+        {Usually aproximated by $P(x|\overline{novel})$}
+        \begin{itemize}
+            \item Bayes rule
+            \item Assumption on constant $P(\overline{novel})$
+            \item Assumption on constant $P(x)$
+        \end{itemize}
+
+        \begin{equation*}
+            \frac{P(x|\overline{novel})P(\overline{novel})}{P(x)} \propto \frac{P(x|\overline{novel})}{P(x)} \propto P(x|\overline{novel})
+        \end{equation*}
     \end{block}
@@
 
-\subsection{Dynamic distributions}
-\begin{frame}
-    The set of observed features and graph structure is always changing.
-    Threshold on $P(x|known)$ \cite{bishop1994novelty}
-    Dynamic graphs
+\begin{frame}{Dynamic distributions}
+    \begin{columns}[c]
+        \column{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{figures/conditional-prob-graph.pdf}
+
+        \column{0.6\textwidth}
+        \begin{itemize}
+            \item Distribution changes with graph structure and sensed features.
+            \item A threshold trained on $P(x|\overline{novel})$ is only valid for a specific
+                graph.
+        \end{itemize}
+    \end{columns}
+    \begin{center}
+        \uncover<2>{\alert{$P(x)$ is not constant - needs to be approximated.}}
+    \end{center}
 \end{frame}
 
-\subsection{Semi-supervised}
+\subsection{Unconditional Probability}
+\begin{frame}{Unconditional Probability}
+    Assume $P(x)$ to be uniform:
+    \vfill
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.50\textwidth]{figures/uniform-prob-graph.pdf}
+    \end{figure}
+    \vfill
+    Expected to approximate poorly as it discards knowledge on the graph structure and other variables.
+\end{frame}
+
+\begin{frame}{Unconditional Probability}
+    \only<1>{
+    Unability to model $R_1$ leads to all factors dependent on it become between themselves:
+    \vfill
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.50\textwidth]{figures/unconditional-prob-graph.pdf}
+    \end{figure}
+    }
+    \only<2>{
+    Approach the unknown factor with independent single factors that can easily be trained from unlabelled data:
+    \vfill
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.50\textwidth]{figures/single-unconditional-prob-graph.pdf}
+    \end{figure}
+    }
+\end{frame}
+
+\subsection{Synthetic Distribution}
 \begin{frame}
-
+    {Results}
+    \begin{itemize}
+        \item Synthetic distribution.
+        \item 11 diferent room categories and 9 types of room properties.
+        \item Database of properties build from web data.
+        \item A variable number of independent features is generated from a given room category.
+        \item Only 5 room categories are given on labelled data.
+    \end{itemize}
+    \only<1>{
+        \begin{figure}[h]
+            \centering
+            \includegraphics[width=0.40\textwidth]{figures/simple-cond-graph.pdf}
+        \end{figure}
+    }
+    \only<2>{
+        \begin{figure}[h]
+            \centering
+            \subfloat[Uniform approach]{\includegraphics[width=0.30\textwidth]{figures/simple-uniform-graph.pdf}}
+            \qquad \qquad
+            \subfloat[Independent approach]{\includegraphics[width=0.30\textwidth]{figures/simple-independent-graph.pdf}}
+        \end{figure}
+    }
 \end{frame}
@@
 \subsection{Results}
-\begin{frame}
-    Synthetic dataset where 1 room category is responsible for generating a variable set of observed features.
+\begin{frame}{Results}
+    \begin{figure}[h]
+    \centering
+    \includegraphics[width=0.60\textwidth]{results/synthetic-all.pdf}
+
+    \caption{\label{fig:synthetic-roc}ROC curve comparing novelty detection performance
+             under samples with variable size of sensed properties.}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Results}
+    \begin{figure}[h]
+        \centering
+
+        \subfloat[3 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-3features.pdf}}
+        \qquad
+        \subfloat[5 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-5features.pdf}}
+
+        \subfloat[10 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-10features.pdf}}
+        \qquad
+        \subfloat[50 sensed features]{\includegraphics[width=0.35\textwidth]{results/synthetic-50features.pdf}}
+
+%        \caption{\label{fig:synthetic-roc-breakdown}ROC curves plotted showing performance of the
+%        presented novelty detection method for graphs generated for different amount of
+%        sensed features.}
+    \end{figure}
 \end{frame}
@@
 \section{Conclusion}
-\subsection{Conclusion}
-\begin{frame}
-    \begin{block}
-        {Semantic Mapping Process}
-        A multi-modal semantic mapping process based on \emph{probabilistic graphical models}.
+\subsection{Summary}
+\begin{frame}{Summary}
+    \begin{itemize}
+        \item Studies a semantic mapping process
+        \item Studies novelty detection
+        \item Proposes a method to perform detection of novel categories
+              of a variable when the graph structure is dynamic.
+        \item Shows access to unlabelled data can greatly increase performance.
+    \end{itemize}
+\end{frame}
+
+\subsection{Limitations}
+\begin{frame}{Limitations and Future Work}
+    \begin{block}{Assumption on constant $P(novel)$}
+        Graph structure plays a role on the likelihood a sample with a novel
+        category is drawn. Assuming it to be constant is a very strong assumption
+        that is not expected to hold in all scenarios.
     \end{block}
     \begin{block}
-        {Novelty detection via thresholding}
-        Presented a method to detect novel classes on a variable of \emph{probabilistic graphical models}
-        dinamicaly instatiated from sensed semantic data.
+        {Generalize the framework}
+        Handle probabilistic classification from low-level classifiers and incorporate novelty signals from them.
     \end{block}
@@
 
-\subsection{Future Work}
-\begin{frame}{Future Work}
-    \begin{block}
-        {Incorporate novelty signals}
+\subsection{Extra Information}
+\begin{frame}{Extra}
+    \begin{block}{Reproducible Research}
+        All reports and results reproducible from online repository:
+        \url{https://github.com/andresusanopinto/novelty-detection-thesis}
     \end{block}
